### 2023-06-21 TKT_RNAseq/R Code/EdgeR_Analysis.Rmd
TKT edgeR analysis update

---
title: "TKT RNAseq Experiment: Differential Expression Analysis"
author: "John Santiago"
date: "2022-06-20"
output: 
  bookdown::html_document2:
    number_sections: false
---

```{r setup, include=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## [Return to Home Page](https://johncsantiago.github.io/TKT_RNAseq_preliminary.html)

<br /><br />
\newpage

## <span style="text-decoration:underline">Table of Contents</span>

### All samples processed together
- #### Figure \@ref(fig:fig1): MDS plot, all samples

- #### Figure \@ref(fig:fig2): Barplot of Total DEGs

### Testing potential outlier samples

- #### Figure \@ref(fig:fig3): PCA of GR.F CxDf DEGs

- #### Figure \@ref(fig:fig4): PCA of GR.F CxOE DEGs

- #### Figure \@ref(fig:fig5): PCA of GR.FC x WT.FC DEGs

- #### Figure \@ref(fig:fig6): PCA of WT.F CxDf DEGs

- #### Figure \@ref(fig:fig7): PCA of WT.F CxOE DEGs

- #### Figure \@ref(fig:fig8): PCA of TktDfGR.F x TktDfWT.F DEGs

### Male and female samples processed together with TEs and outlier libraries removed
- #### Figure \@ref(fig:fig9): MDS plot, all samples

- #### Figure \@ref(fig:fig10): Barplot of Total DEGs

### Male and Female samples processed separately with TEs and outlier libraries removed
- #### Figure \@ref(fig:fig11): MDS plot of female samples only

- #### Figure \@ref(fig:fig12): MDS plot of male samples only

- #### Figure \@ref(fig:fig13): Barplot of Total DEGs

- #### Figure \@ref(fig:fig14): Barplot of Total DEGs for each comparison under the different trimming conditions

<br /><br /><br />
\newpage

### Conclusion: GR.F3, WT.F3 and TktDfWT.F3 are outlier libraries that should be removed before further analysis but there is no need to process male and female samples separately

<br /><br /><br /><br /><br /><br />
\newpage

```{r echo=F, include=FALSE}

library(edgeR)
library(plotly)
library(heatmaply)

##Clean Trait Data and calculate CPMs
countdata=read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT.counttable.csv",row.names=1)
groups=read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/All.TKT.metadata.csv",row.names=1)

convert=read.csv("https://raw.githubusercontent.com/DavidRandLab/Santiago-et-al-2021-BMC-Genomics/main/Data%20Files/FBgnConversionTable.csv",row.names=1)

##normalize data
countdata=countdata[, row.names(groups)]
x <- countdata
gr <- factor(groups$Group)
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
cpmdata=cpm(z)

design<-model.matrix(~0 + gr)
colnames(design) = unique(gr)
z = estimateDisp(z, design, robust = T)

fit <- glmQLFit(z, design)
plotQLDisp(fit)

##Will calculate fold-change with the first over the second

##GR.F TKT control vs TKT altered
compare = makeContrasts(TktDfGR.F-GR.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.F.CxDf=DEout
sGR.F.CxDf=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.F-GR.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.F.CxOE=DEout
sGR.F.CxOE=DEout[DEout$FDR<.05,]

##WT.F TKT control vs TKT altered
compare = makeContrasts(TktDfWT.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.F.CxDf=DEout
sWT.F.CxDf=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEWT.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.F.CxOE=DEout
sWT.F.CxOE=DEout[DEout$FDR<.05,]

##WT.F vs GR.F
compare = makeContrasts(GR.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.F=DEout
sGRxWT.F=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.F-TktDfWT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.FDf=DEout
sGRxWT.FDf=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.F-TktOEWT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.FOE=DEout
sGRxWT.FOE=DEout[DEout$FDR<.05,]

##GR.M TKT control vs TKT altered
compare = makeContrasts(TktDfGR.M-GR.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.M.CxDf=DEout
sGR.M.CxDf=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.M-GR.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.M.CxOE=DEout
sGR.M.CxOE=DEout[DEout$FDR<.05,]

##WT.M TKT control vs TKT altered
compare = makeContrasts(TktDfWT.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.M.CxDf=DEout
sWT.M.CxDf=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEWT.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.M.CxOE=DEout
sWT.M.CxOE=DEout[DEout$FDR<.05,]

##WT.M vs GR.M
compare = makeContrasts(GR.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.M=DEout
sGRxWT.M=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.M-TktDfWT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.MDf=DEout
sGRxWT.MDf=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.M-TktOEWT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.MOE=DEout
sGRxWT.MOE=DEout[DEout$FDR<.05,]

TKT.EdgeR = data.frame(GRF.CxDf = GR.F.CxDf[row.names(cpmdata),'FDR'],
                       GRF.CxOE = GR.F.CxOE[row.names(cpmdata),'FDR'],

                       WTF.CxDf = WT.F.CxDf[row.names(cpmdata),'FDR'],
                       WTF.CxOE = WT.F.CxOE[row.names(cpmdata),'FDR'],

                       GRxWT.FC = GRxWT.F[row.names(cpmdata),'FDR'],
                       GRxWT.FDf = GRxWT.FDf[row.names(cpmdata),'FDR'],
                       GRxWT.FOE = GRxWT.FOE[row.names(cpmdata),'FDR'],

                       GRM.CxDf = GR.M.CxDf[row.names(cpmdata),'FDR'],
                       GRM.CxOE = GR.M.CxOE[row.names(cpmdata),'FDR'],

                       WTM.CxDf = WT.M.CxDf[row.names(cpmdata),'FDR'],
                       WTM.CxOE = WT.M.CxOE[row.names(cpmdata),'FDR'],

                       GRxWT.MC = GRxWT.M[row.names(cpmdata),'FDR'],
                       GRxWT.MDf = GRxWT.MDf[row.names(cpmdata),'FDR'],
                       GRxWT.MOE = GRxWT.MOE[row.names(cpmdata),'FDR'])

row.names(TKT.EdgeR) = row.names(cpmdata)
##write.csv(TKT.EdgeR, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT.EdgeR.Table.csv")

Female = c(nrow(sGR.F.CxDf), nrow(sGR.F.CxOE),
           nrow(sWT.F.CxDf), nrow(sWT.F.CxOE),
           nrow(sGRxWT.F), nrow(sGRxWT.FDf), nrow(sGRxWT.FOE))

Male = c(nrow(sGR.M.CxDf), nrow(sGR.M.CxOE),
         nrow(sWT.M.CxDf), nrow(sWT.M.CxOE),
         nrow(sGRxWT.M), nrow(sGRxWT.MDf), nrow(sGRxWT.MOE))

MDSdata = plotMDS(z, top = 13582, col=as.numeric(factor(gr)))

temp=convert[,"Symbol"]
temp2=setdiff(row.names(countdata),convert[,2])
names(temp)=convert[,2]
names(temp2)=temp2
convert=c(temp,temp2)

Comparisons = list(row.names(sGR.F.CxDf), row.names(sGR.F.CxOE),
                row.names(sWT.F.CxDf), row.names(sWT.F.CxOE),
                row.names(sGRxWT.F), row.names(sGRxWT.FDf), row.names(sGRxWT.FOE),
                row.names(sGR.M.CxDf), row.names(sGR.M.CxOE),
                row.names(sWT.M.CxDf), row.names(sWT.M.CxOE),
                row.names(sGRxWT.M), row.names(sGRxWT.MDf), row.names(sGRxWT.MOE))

GroupsToDEGs = matrix("", nrow = 28, ncol = 3)

GroupsToDEGs[,1] = c("GR.F", "TktDfGR.F",
                     "GR.F", "TktOEGR.F",
                     "WT.F", "TktDfWT.F",
                     "WT.F", "TktOEWT.F",
                     "GR.F", "WT.F",
                     "TktDfGR.F", "TktDfWT.F",
                     "TktOEGR.F", "TktOEWT.F",
                     "GR.M", "TktDfGR.M",
                     "GR.M", "TktOEGR.M",
                     "WT.M", "TktDfWT.M",
                     "WT.M", "TktOEWT.M",
                     "GR.M", "WT.M",
                     "TktDfGR.M", "TktDfWT.M",
                     "TktOEGR.M", "TktOEWT.M")

GroupsToDEGs[,2] = c("TktDfGR.F", "GR.F",
                     "TktOEGR.F", "GR.F", 
                     "TktDfWT.F", "WT.F", 
                     "TktOEWT.F", "WT.F", 
                     "WT.F", "GR.F", 
                     "TktDfWT.F", "TktDfGR.F", 
                     "TktOEWT.F", "TktOEGR.F", 
                     "TktDfGR.M", "GR.M", 
                     "TktOEGR.M", "GR.M", 
                     "TktDfWT.M", "WT.M", 
                     "TktOEWT.M", "WT.M", 
                     "WT.M", "GR.M", 
                     "TktDfWT.M", "TktDfGR.M", 
                     "TktOEWT.M", "TktOEGR.M")

GroupsToDEGs[,3] = c(1,1,2,2,3,3,4,4,5,5,6,6,7,7,
                     8,8,9,9,10,10,11,11,12,12,13,13,14,14)

```

## <span style="text-decoration:underline">Do the samples group by condition?</span> 
```{r fig1, echo = F, warning=F, fig.height=10, fig.width=10, fig.cap= 'MDS plot using all samples. Of these, samples GR.F3, WT.F3 and TktDfWT.F3 stand out as potential outliers.'}

MDSplot = data.frame(x = MDSdata$x,
                     y = MDSdata$y,
                     sample = row.names(groups))
MDSplot = cbind(MDSplot, groups[MDSplot$sample,])
MDSplot$GenoSex = paste0(MDSplot$Genotype,MDSplot$Sex)
MDSplot = MDSplot[order(MDSplot$TKT),]
MDSplot = MDSplot[order(MDSplot$Sex),]
MDSplot = MDSplot[order(MDSplot$GenoSex),]
MDSplot$Group = factor(MDSplot$Group, levels = unique(MDSplot$Group))
##MDSplot$GenoSex = factor(MDSplot$GenoSex, levels = unique(MDSplot$GenoSex))

  fig  = plot_ly(data = MDSplot, 
                 x = ~x, 
                 y = ~y,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Group,
                 ##legendrank = ~ as.numeric(Group),
                 symbols = rep(c('circle', 'triangle-down', 'triangle-up'),4),
                 colors = c("red","firebrick3","deeppink",
                            "lightskyblue","royalblue","cyan",
                            "gold", "darkgoldenrod1", "yellow",
                            "limegreen", "darkgreen", "chartreuse"),
                 marker = list(size = 25,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", MDSplot$Sample)) %>%
    layout(xaxis = list(title = paste0("Leading logFC dim 1 ", "(", round((MDSdata$var.explained[1]*100),1), "%)")),
           yaxis = list(title = paste0("Leading logFC dim 2 ", "(", round((MDSdata$var.explained[2]*100),1), "%)")))

fig

```

<br /><br /><br /><br /><br /><br />
\newpage

```{r fig2, echo = F, fig.cap='Barplot of Total DEGs observed with edgeR when using all samples.'}

Female = c(nrow(sGR.F.CxDf), nrow(sGR.F.CxOE),
           nrow(sWT.F.CxDf), nrow(sWT.F.CxOE),
           nrow(sGRxWT.F), nrow(sGRxWT.FDf), nrow(sGRxWT.FOE))

Male = c(nrow(sGR.M.CxDf), nrow(sGR.M.CxOE),
         nrow(sWT.M.CxDf), nrow(sWT.M.CxOE),
         nrow(sGRxWT.M), nrow(sGRxWT.MDf), nrow(sGRxWT.MOE))

Comparison = c("GR.C x GR.Df", "GR.C x GR.OE",
               "WT.C x WT.Df", "WT.C x WT.OE",
               "GR.C x WT.C", "GR.Df x WT.Df", "GR.OE x WT.OE")

data = data.frame(Female, Male, Comparison)

data$Comparison = factor(data$Comparison, levels = data$Comparison)

fig <- plot_ly(data, 
               x = ~Comparison, y = ~Female, 
               type = 'bar',
               name = "Female",
               marker = list(color =c("firebrick"),
                             line = list(color = "black", width = 1.5)))

fig <- fig %>% add_trace(y = ~Male, name = 'Males',
                         marker = list(color = c("royalblue"),
                                       line = list(color = "black", width = 1.5)))

fig <- fig %>% layout(yaxis = list(title ="Total DEGs"),
                      xaxis = list(title = NA),
                      title = "Total DEGs")

fig

```

<br /><br /><br /><br /><br /><br />
\newpage

```{r, echo =F, include=F}
##Supervised PCAs

##Samples to plot
##genotype = "GR"                       ##options = "GR", "WT", or "all"
##sex      = "F"                         ##options = "M", "F", or "all"
##tkt      = c("Control", "DF")                    ##options = "Control", "DF", "OE", or "all"

                                       ##must choose genotype of first and second condition
##DEGs.genotypes = c("GR", "GR")         ##length = 2; options = "GR", "WT", or "all"

                                       ##a single sex, won't compare between sexs
##DEGs.sex       = "F"                   ##length = 1; options = "M", "F"

                                       ##choose tkt expression of first and second condition
##DEGs.tkt       = c("Control", "DF")    ##length = 1; "Control", "DF", "OE"

                                  ##must choose genotype of first and second condition
genotypes = c("GR", "GR")         ##length = 2; options = "GR", "WT", or "all"

                                  ##a single sex, won't compare between sexs
sex       = "F"                   ##length = 1; options = "M", "F"

                                  ##choose tkt expression of first and second condition
tkt       = c("Control", "DF")    ##length = 1; "Control", "DF", "OE"

##SupervisedPCA = function(genotype, sex, tkt, DEGs.genotypes, DEGs.sex, DEGs.tkt){
SupervisedPCA = function(genotypes, sex, tkt){
Group1=unique(groups[(groups$Genotype == genotypes[1] & ##DEGs.genotypes[1] &
                      groups$Sex == sex & ##DEGs.sex &
                      groups$TKT == tkt[1]), ##DEGs.tkt[1]), 
                      "Group"])

Group2=unique(groups[(groups$Genotype == genotypes[2] & ##DEGs.genotypes[2] &
                      groups$Sex == sex & ##DEGs.sex &
                      groups$TKT == tkt[2]), ##DEGs.tkt[2]), 
                      "Group"])

DEGs = GroupsToDEGs[(GroupsToDEGs[,1] == Group1 &
                     GroupsToDEGs[,2] == Group2),
                    3]

DEGs = Comparisons[[as.numeric(DEGs)]]

    use.groups =  groups[groups$Group == Group1 |
                         groups$Group == Group2,]

    ##if(genotype == "GR" | genotype == "WT"){
      ##use.groups = use.groups[use.groups$Genotype == genotype,]
    ##}

    ##if(sex == "M" | sex == "F"){
      ##use.groups = use.groups[use.groups$Sex == sex,]
    ##}

    ##if(length(tkt) == 1 & length(intersect(tkt, "all")) == 0){
      ##use.groups = use.groups[use.groups$TKT == tkt,]
    ##}

    ##if(length(tkt) == 2){
      ##use.groups = use.groups[use.groups$TKT == tkt[1]|
                                ##use.groups$TKT == tkt[2],]
    ##}

    use.data   = cpmdata[DEGs,row.names(use.groups)]
    pca <- prcomp(t(use.data), scale.=TRUE) 
    gr <- as.data.frame(use.groups)
    PC1=scale(pca$x[,1])
    PC2=scale(pca$x[,2])
    eigs <- pca$sdev^2
    ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(Sample=row.names(gr),
                        PC1=scale(pca$x[,1]),
                        PC2=scale(pca$x[,2]),
                        sdev=pca$sdev,
                        Genotype=gr$Genotype, 
                        Sex=gr$Sex, 
                        TKT=gr$TKT,
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group=gr$Group)

  fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~TKT,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 colors = c("firebrick","gold","deepskyblue"),
                 marker = list(size = 25,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Sample)) %>%
    layout(xaxis = list(title = paste0("PC1", "(", ve[1], "%)")),
           yaxis = list(title = paste0("PC2", "(", ve[2], "%)")))
  return(fig)
  }

```

## <span style="text-decoration:underline">Is the expression profile for GR.F3 an outlier in genes found DE GR.FC and GR.FDf?</span> 
```{r fig3, echo = F, warning=F, fig.height=8, fig.width=8, echo = F, fig.cap='PCA of GR.F CxDf DEGs. Testing if GR.F3 is an outlier library.'}
##Supervised PCAs
##Samples to plot
                                  ##must choose genotype of first and second condition
genotypes = c("GR", "GR")         ##length = 2; options = "GR", "WT", or "all"

                                  ##a single sex, won't compare between sexs
sex       = "F"                   ##length = 1; options = "M", "F"

                                  ##choose tkt expression of first and second condition
tkt       = c("Control", "DF")    ##length = 1; "Control", "DF", "OE"

SupervisedPCA(genotypes, sex, tkt)

```

<br /><br /><br /><br /><br /><br />
\newpage

## <span style="text-decoration:underline">Is the expression profile for GR.F3 an outlier in genes found DE GR.FC and GR.FOE?</span> 
```{r fig4, echo = F, warning=F, fig.height=8, fig.width=8, fig.cap='PCA of GR.F CxOE DEGs. Testing if GR.F3 is an outlier library.'}
##Supervised PCAs
##Samples to plot
                                  ##must choose genotype of first and second condition
genotypes = c("GR", "GR")         ##length = 2; options = "GR", "WT", or "all"

                                  ##a single sex, won't compare between sexs
sex       = "F"                   ##length = 1; options = "M", "F"

                                  ##choose tkt expression of first and second condition
tkt       = c("Control", "OE")    ##length = 1; "Control", "DF", "OE"

SupervisedPCA(genotypes, sex, tkt)

```

<br /><br /><br /><br /><br /><br />
\newpage

## <span style="text-decoration:underline">Are the expression profiles for GR.F3 and WT.F3 outliers in genes found DE GR.FC and WT.FC?</span> 
```{r fig5, echo = F, warning=F, fig.height=8, fig.width=8, fig.cap='PCA of GR.FC x WT.FC DEGs. Testing if GR.F3 and WT.F3 are outlier libraries.'}
##Supervised PCAs
##Samples to plot
                                  ##must choose genotype of first and second condition
genotypes = c("GR", "WT")         ##length = 2; options = "GR", "WT", or "all"

                                  ##a single sex, won't compare between sexs
sex       = "F"                   ##length = 1; options = "M", "F"

                                  ##choose tkt expression of first and second condition
tkt       = c("Control", "Control")    ##length = 1; "Control", "DF", "OE"

SupervisedPCA(genotypes, sex, tkt)

```

<br /><br /><br /><br /><br /><br />
\newpage

## <span style="text-decoration:underline">Are the expression profiles for WT.F3 and TktDfWT.F3 outliers in genes found WT.FC and TktDfWT.F?</span> 
```{r fig6, echo = F, warning=F, fig.height=8, fig.width=8, fig.cap='PCA of WT.F CxDf DEGs. Testing if WT.F3 and TktDfWT.F3 are outlier libraries.'}
##Supervised PCAs
##Samples to plot
                                  ##must choose genotype of first and second condition
genotypes = c("WT", "WT")         ##length = 2; options = "GR", "WT", or "all"

                                  ##a single sex, won't compare between sexs
sex       = "F"                   ##length = 1; options = "M", "F"

                                  ##choose tkt expression of first and second condition
tkt       = c("Control", "DF")    ##length = 1; "Control", "DF", "OE"

SupervisedPCA(genotypes, sex, tkt)

```

<br /><br /><br /><br /><br /><br />
\newpage

## <span style="text-decoration:underline">Is the expression profile for WT.F3 an outlier in genes found DE WT.FC and WT.FOE?</span> 
```{r fig7, echo = F, warning=F, fig.height=8, fig.width=8, fig.cap='PCA of WT.F CxOE DEGs. Testing if WT.F3 is an outlier library.'}
##Supervised PCAs
##Samples to plot
                                  ##must choose genotype of first and second condition
genotypes = c("WT", "WT")         ##length = 2; options = "GR", "WT", or "all"

                                  ##a single sex, won't compare between sexs
sex       = "F"                   ##length = 1; options = "M", "F"

                                  ##choose tkt expression of first and second condition
tkt       = c("Control", "OE")    ##length = 1; "Control", "DF", "OE"

SupervisedPCA(genotypes, sex, tkt)

```

<br /><br /><br /><br /><br /><br />
\newpage

## <span style="text-decoration:underline">Is the expression profile for TktDfWT.F3 an outlier in genes found DE TktDfGR.F and TktDfWT.F?</span> 
```{r fig8, echo = F, warning=F, fig.height=8, fig.width=8, fig.cap='PCA of TktDfGR.F x TktDfWT.F DEGs. Testing if TktDfWT.F3 is an outlier library.'}
##Supervised PCAs
##Samples to plot
                                  ##must choose genotype of first and second condition
genotypes = c("GR", "WT")         ##length = 2; options = "GR", "WT", or "all"

                                  ##a single sex, won't compare between sexs
sex       = "F"                   ##length = 1; options = "M", "F"

                                  ##choose tkt expression of first and second condition
tkt       = c("DF", "DF")    ##length = 1; "Control", "DF", "OE"

SupervisedPCA(genotypes, sex, tkt)

```

<br /><br /><br /><br /><br /><br />
\newpage

```{r, echo = F, include = F}
###EdgeR with TEs and outlier libraries removed

groups2 = groups[setdiff(row.names(groups), c("GR.F3", "TktDfWT.F3")),]##"WT.F3",

countdata2 = countdata[, row.names(groups2)]

countdata2 = countdata2[-(grep("ti",row.names(countdata2))),]
countdata2 = countdata2[-(grep("trans",row.names(countdata2))),]

x <- countdata2
gr <- factor(groups2$Group)
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
cpmdata2=cpm(z)

design<-model.matrix(~0 + gr)
colnames(design) = unique(gr)
z = estimateDisp(z, design, robust = T)

fit <- glmQLFit(z, design)
plotQLDisp(fit)

##Will calculate fold-change with the first over the second

##GR.F TKT control vs TKT altered
compare = makeContrasts(TktDfGR.F-GR.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.F.CxDf2=DEout
sGR.F.CxDf2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.F-GR.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.F.CxOE2=DEout
sGR.F.CxOE2=DEout[DEout$FDR<.05,]

##WT.F TKT control vs TKT altered
compare = makeContrasts(TktDfWT.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.F.CxDf2=DEout
sWT.F.CxDf2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEWT.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.F.CxOE2=DEout
sWT.F.CxOE2=DEout[DEout$FDR<.05,]

##WT.F vs GR.F
compare = makeContrasts(GR.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.F2=DEout
sGRxWT.F2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.F-TktDfWT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.FDf2=DEout
sGRxWT.FDf2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.F-TktOEWT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.FOE2=DEout
sGRxWT.FOE2=DEout[DEout$FDR<.05,]

##GR.M TKT control vs TKT altered
compare = makeContrasts(TktDfGR.M-GR.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.M.CxDf2=DEout
sGR.M.CxDf2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.M-GR.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.M.CxOE2=DEout
sGR.M.CxOE2=DEout[DEout$FDR<.05,]

##WT.M TKT control vs TKT altered
compare = makeContrasts(TktDfWT.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.M.CxDf2=DEout
sWT.M.CxDf2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEWT.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.M.CxOE2=DEout
sWT.M.CxOE2=DEout[DEout$FDR<.05,]

##WT.M vs GR.M
compare = makeContrasts(GR.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.M2=DEout
sGRxWT.M2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.M-TktDfWT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.MDf2=DEout
sGRxWT.MDf2=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.M-TktOEWT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.MOE2=DEout
sGRxWT.MOE2=DEout[DEout$FDR<.05,]

MDSdata2 = plotMDS(z, top = 13086, col=as.numeric(factor(gr)))

Comparisons2 = list(row.names(sGR.F.CxDf2), row.names(sGR.F.CxOE2),
                row.names(sWT.F.CxDf2), row.names(sWT.F.CxOE2),
                row.names(sGRxWT.F2), row.names(sGRxWT.FDf2), row.names(sGRxWT.FOE2),
                row.names(sGR.M.CxDf2), row.names(sGR.M.CxOE2),
                row.names(sWT.M.CxDf2), row.names(sWT.M.CxOE2),
                row.names(sGRxWT.M2), row.names(sGRxWT.MDf2), row.names(sGRxWT.MOE2))

```

## <span style="text-decoration:underline">Does processing the samples after removing the 3 outlier libraries and TEs improve sample normalization?</span> 
```{r fig9, echo = F, warning=F, fig.height=10, fig.width=10, fig.cap= 'MDS plot using all samples, with the TEs and outlier libraries removed.'}

MDSplot = data.frame(x = MDSdata2$x,
                     y = MDSdata2$y,
                     sample = row.names(groups2))
MDSplot = cbind(MDSplot, groups[MDSplot$sample,])
MDSplot$GenoSex = paste0(MDSplot$Genotype,MDSplot$Sex)
MDSplot = MDSplot[order(MDSplot$TKT),]
MDSplot = MDSplot[order(MDSplot$Sex),]
MDSplot = MDSplot[order(MDSplot$GenoSex),]
MDSplot$Group = factor(MDSplot$Group, levels = unique(MDSplot$Group))
##MDSplot$GenoSex = factor(MDSplot$GenoSex, levels = unique(MDSplot$GenoSex))

  fig  = plot_ly(data = MDSplot, 
                 x = ~x, 
                 y = ~y,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Group,
                 ##legendrank = ~ as.numeric(Group),
                 symbols = rep(c('circle', 'triangle-down', 'triangle-up'),4),
                 colors = c("red","firebrick3","deeppink",
                            "lightskyblue","royalblue","cyan",
                            "gold", "darkgoldenrod1", "yellow",
                            "limegreen", "darkgreen", "chartreuse"),
                 marker = list(size = 25,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", MDSplot$Sample)) %>%
    layout(xaxis = list(title = paste0("Leading logFC dim 1 ", "(", round((MDSdata2$var.explained[1]*100),1), "%)")),
           yaxis = list(title = paste0("Leading logFC dim 2 ", "(", round((MDSdata2$var.explained[2]*100),1), "%)")))

fig

```

<br /><br /><br /><br /><br /><br />
\newpage

```{r fig10, echo = F, fig.cap='Barplot of Total DEGs found when using all samples with the TEs and outlier libraries removed.'}

Female = c(nrow(sGR.F.CxDf2), nrow(sGR.F.CxOE2),
           nrow(sWT.F.CxDf2), nrow(sWT.F.CxOE2),
           nrow(sGRxWT.F2), nrow(sGRxWT.FDf2), nrow(sGRxWT.FOE2))

Male = c(nrow(sGR.M.CxDf2), nrow(sGR.M.CxOE2),
         nrow(sWT.M.CxDf2), nrow(sWT.M.CxOE2),
         nrow(sGRxWT.M2), nrow(sGRxWT.MDf2), nrow(sGRxWT.MOE2))

Comparison = c("GR.C x GR.Df", "GR.C x GR.OE",
               "WT.C x WT.Df", "WT.C x WT.OE",
               "GR.C x WT.C", "GR.Df x WT.Df", "GR.OE x WT.OE")

data = data.frame(Female, Male, Comparison)

data$Comparison = factor(data$Comparison, levels = data$Comparison)

fig <- plot_ly(data, 
               x = ~Comparison, y = ~Female, 
               type = 'bar',
               name = "Female",
               marker = list(color =c("firebrick"),
                             line = list(color = "black", width = 1.5)))

fig <- fig %>% add_trace(y = ~Male, name = 'Males',
                         marker = list(color = c("royalblue"),
                                       line = list(color = "black", width = 1.5)))

fig <- fig %>% layout(yaxis = list(title ="Total DEGs"),
                      xaxis = list(title = NA),
                      title = "Total DEGs")

fig

```

<br /><br /><br /><br /><br /><br />
\newpage

```{r, include = F}
###EdgeR with TEs and outlier libraries removed, Females only

groups3 = groups[setdiff(row.names(groups), c("GR.F3", "WT.F3","TktDfWT.F3")),]##

groups3 = groups3[groups3$Sex == "F",]

countdata3 = countdata[, row.names(groups3)]

countdata3 = countdata3[-(grep("ti",row.names(countdata3))),]
countdata3 = countdata3[-(grep("trans",row.names(countdata3))),]

x <- countdata3
gr <- factor(groups3$Group)
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
cpmdata3=cpm(z)

design<-model.matrix(~0 + gr)
colnames(design) = unique(gr)
z = estimateDisp(z, design, robust = T)

fit <- glmQLFit(z, design)
plotQLDisp(fit)

##Will calculate fold-change with the first over the second

##GR.F TKT control vs TKT altered
compare = makeContrasts(TktDfGR.F-GR.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.F.CxDf3=DEout
sGR.F.CxDf3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.F-GR.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.F.CxOE3=DEout
sGR.F.CxOE3=DEout[DEout$FDR<.05,]

##WT.F TKT control vs TKT altered
compare = makeContrasts(TktDfWT.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.F.CxDf3=DEout
sWT.F.CxDf3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEWT.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.F.CxOE3=DEout
sWT.F.CxOE3=DEout[DEout$FDR<.05,]

##WT.F vs GR.F
compare = makeContrasts(GR.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.F3=DEout
sGRxWT.F3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.F-TktDfWT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.FDf3=DEout
sGRxWT.FDf3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.F-TktOEWT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.FOE3=DEout
sGRxWT.FOE3=DEout[DEout$FDR<.05,]

MDSdata3 = plotMDS(z, top = 10968, col=as.numeric(factor(gr)))

```

## <span style="text-decoration:underline">Does sex specifically processing the samples after removing the 3 outlier libraries and TEs improve sample normalization?</span> 
```{r fig11, echo = F, warning=F, fig.height=10, fig.width=10, fig.cap= 'MDS plot when the female samples are processed alone. The TEs and outlier libraries are removed.'}

MDSplot = data.frame(x = MDSdata3$x,
                     y = MDSdata3$y,
                     sample = row.names(groups3))
MDSplot = cbind(MDSplot, groups[MDSplot$sample,])
MDSplot$GenoSex = paste0(MDSplot$Genotype,MDSplot$Sex)
MDSplot = MDSplot[order(MDSplot$TKT),]
MDSplot = MDSplot[order(MDSplot$Sex),]
MDSplot = MDSplot[order(MDSplot$GenoSex),]
MDSplot$Group = factor(MDSplot$Group, levels = unique(MDSplot$Group))
##MDSplot$GenoSex = factor(MDSplot$GenoSex, levels = unique(MDSplot$GenoSex))

  fig  = plot_ly(data = MDSplot, 
                 x = ~x, 
                 y = ~y,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Group,
                 ##legendrank = ~ as.numeric(Group),
                 symbols = rep(c('circle', 'triangle-down', 'triangle-up'),4),
                 colors = c("red","firebrick3","deeppink",
                            "gold", "darkgoldenrod1", "yellow"),
                 marker = list(size = 25,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", MDSplot$Sample)) %>%
    layout(xaxis = list(title = paste0("Leading logFC dim 1 ", "(", round((MDSdata3$var.explained[1]*100),1), "%)")),
           yaxis = list(title = paste0("Leading logFC dim 2 ", "(", round((MDSdata3$var.explained[2]*100),1), "%)")))

fig

```

<br /><br /><br /><br /><br /><br />
\newpage

```{r, include = F}
###EdgeR with TEs and outlier libraries removed, Males only

groups4 = groups[groups$Sex == "M",]

countdata4 = countdata[, row.names(groups4)]

countdata4 = countdata4[-(grep("ti",row.names(countdata4))),]
countdata4 = countdata4[-(grep("trans",row.names(countdata4))),]

x <- countdata4
gr <- factor(groups4$Group)
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
cpmdata4=cpm(z)

design<-model.matrix(~0 + gr)
colnames(design) = unique(gr)
z = estimateDisp(z, design, robust = T)

fit <- glmQLFit(z, design)
plotQLDisp(fit)

##Will calculate fold-change with the first over the second

##GR.M TKT control vs TKT altered
compare = makeContrasts(TktDfGR.M-GR.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.M.CxDf3=DEout
sGR.M.CxDf3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.M-GR.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.M.CxOE3=DEout
sGR.M.CxOE3=DEout[DEout$FDR<.05,]

##WT.F TKT control vs TKT altered
compare = makeContrasts(TktDfWT.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.M.CxDf3=DEout
sWT.M.CxDf3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEWT.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.M.CxOE3=DEout
sWT.M.CxOE3=DEout[DEout$FDR<.05,]

##WT.F vs GR.F
compare = makeContrasts(GR.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.M3=DEout
sGRxWT.M3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.M-TktDfWT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.MDf3=DEout
sGRxWT.MDf3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.M-TktOEWT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.MOE3=DEout
sGRxWT.MOE3=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRMDfxWTM=DEout
sGRMDfxWTM=DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.M-TktDfWT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRMDfxWTMDf=DEout
sGRMDfxWTMDf=DEout[DEout$FDR<.05,]

MDSdata4 = plotMDS(z, top = 12764, col=as.numeric(factor(gr)))

all.cpmgenes = unique(c(row.names(cpmdata3),row.names(cpmdata4)))

all.cpmdata = matrix(NA, ncol = 33, nrow = length(all.cpmgenes))
colnames(all.cpmdata) = c(colnames(cpmdata3),colnames(cpmdata4))
row.names(all.cpmdata) = all.cpmgenes

all.cpmdata[row.names(cpmdata3),colnames(cpmdata3)] = cpmdata3
all.cpmdata[row.names(cpmdata4),colnames(cpmdata4)] = cpmdata4

all.groups = groups[colnames(all.cpmdata),]

all.TKT.EdgeR = all.cpmdata[, 1:14]
colnames(all.TKT.EdgeR) = c('GRF.CxDf', 'GRF.CxOE', 
                            'WTF.CxDf', 'WTF.CxOE', 
                            'GRxWT.FC', 'GRxWT.FDf', 'GRxWT.FOE',
                            'GRM.CxDf', 'GRM.CxOE', 
                            'WTM.CxDf', 'WTM.CxOE', 
                            'GRxWT.MC', 'GRxWT.MDf', 'GRxWT.MOE')

row.names(all.TKT.EdgeR) = row.names(all.cpmdata)

all.TKT.EdgeR[,1] = GR.F.CxDf3[row.names(all.cpmdata),'FDR']
all.TKT.EdgeR[,2] = GR.F.CxOE3[row.names(all.cpmdata),'FDR']

all.TKT.EdgeR[,3] = WTF.CxDf = WT.F.CxDf3[row.names(all.cpmdata),'FDR']
all.TKT.EdgeR[,4] = WTF.CxOE = WT.F.CxOE3[row.names(all.cpmdata),'FDR']

all.TKT.EdgeR[,5] = GRxWT.FC = GRxWT.F3[row.names(all.cpmdata),'FDR']
all.TKT.EdgeR[,6] = GRxWT.FDf = GRxWT.FDf3[row.names(all.cpmdata),'FDR']
all.TKT.EdgeR[,7] = GRxWT.FOE = GRxWT.FOE3[row.names(all.cpmdata),'FDR']

all.TKT.EdgeR[,8] = GRM.CxDf = GR.M.CxDf3[row.names(all.cpmdata),'FDR']
all.TKT.EdgeR[,9] = GRM.CxOE = GR.M.CxOE3[row.names(all.cpmdata),'FDR']

all.TKT.EdgeR[,10] = WTM.CxDf = WT.M.CxDf3[row.names(all.cpmdata),'FDR']
all.TKT.EdgeR[,11] = WTM.CxOE = WT.M.CxOE3[row.names(all.cpmdata),'FDR']

all.TKT.EdgeR[,12] = GRxWT.MC = GRxWT.M3[row.names(all.cpmdata),'FDR']
all.TKT.EdgeR[,13] = GRxWT.MDf = GRxWT.MDf3[row.names(all.cpmdata),'FDR']
all.TKT.EdgeR[,14] = GRxWT.MOE = GRxWT.MOE3[row.names(all.cpmdata),'FDR']

##write.csv(all.cpmdata, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT_cpmdata.csv")
##write.csv(all.groups, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT.metadata.csv")
##write.csv(all.TKT.EdgeR, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT.EdgeR.Table.csv")

```

```{r fig12, echo = F, warning=F, fig.height=10, fig.width=10, fig.cap= 'MDS plot when the male samples are process alone only. The TEs are removed.'}

MDSplot = data.frame(x = MDSdata4$x,
                     y = MDSdata4$y,
                     sample = row.names(groups4))
MDSplot = cbind(MDSplot, groups[MDSplot$sample,])
MDSplot$GenoSex = paste0(MDSplot$Genotype,MDSplot$Sex)
MDSplot = MDSplot[order(MDSplot$TKT),]
MDSplot = MDSplot[order(MDSplot$Sex),]
MDSplot = MDSplot[order(MDSplot$GenoSex),]
MDSplot$Group = factor(MDSplot$Group, levels = unique(MDSplot$Group))
##MDSplot$GenoSex = factor(MDSplot$GenoSex, levels = unique(MDSplot$GenoSex))

  fig  = plot_ly(data = MDSplot, 
                 x = ~x, 
                 y = ~y,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Group,
                 ##legendrank = ~ as.numeric(Group),
                 symbols = rep(c('circle', 'triangle-down', 'triangle-up'),4),
                 colors = c("lightskyblue","royalblue","cyan",
                            "limegreen", "darkgreen", "chartreuse"),
                 marker = list(size = 25,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", MDSplot$Sample)) %>%
    layout(xaxis = list(title = paste0("Leading logFC dim 1 ", "(", round((MDSdata4$var.explained[1]*100),1), "%)")),
           yaxis = list(title = paste0("Leading logFC dim 2 ", "(", round((MDSdata4$var.explained[2]*100),1), "%)")))

fig

```

<br /><br /><br /><br /><br /><br />
\newpage

```{r fig13, echo = F, fig.cap='Barplot of Total DEGs when male and female samples are processed separately. The TEs and outlier libraries were removed.'}

Female = c(nrow(sGR.F.CxDf3), nrow(sGR.F.CxOE3),
           nrow(sWT.F.CxDf3), nrow(sWT.F.CxOE3),
           nrow(sGRxWT.F3), nrow(sGRxWT.FDf3), nrow(sGRxWT.FOE3))

Male = c(nrow(sGR.M.CxDf3), nrow(sGR.M.CxOE3),
         nrow(sWT.M.CxDf3), nrow(sWT.M.CxOE3),
         nrow(sGRxWT.M3), nrow(sGRxWT.MDf3), nrow(sGRxWT.MOE3))

Comparison = c("GR.C x GR.Df", "GR.C x GR.OE",
               "WT.C x WT.Df", "WT.C x WT.OE",
               "GR.C x WT.C", "GR.Df x WT.Df", "GR.OE x WT.OE")

data = data.frame(Female, Male, Comparison)

data$Comparison = factor(data$Comparison, levels = data$Comparison)

fig <- plot_ly(data, 
               x = ~Comparison, y = ~Female, 
               type = 'bar',
               name = "Female",
               marker = list(color =c("firebrick"),
                             line = list(color = "black", width = 1.5)))

fig <- fig %>% add_trace(y = ~Male, name = 'Males',
                         marker = list(color = c("royalblue"),
                                       line = list(color = "black", width = 1.5)))

fig <- fig %>% layout(yaxis = list(title ="Total DEGs"),
                      xaxis = list(title = NA),
                      title = "Total DEGs")

fig

```

<br /><br /><br /><br /><br /><br />
\newpage

## <span style="text-decoration:underline">How do the different processing methods alter the number of total DEGs?</span> 
```{r fig14, echo = F, fig.cap='Barplot of Total DEGs for each comparison under the different trimming conditions. Different trimming conditions were: (1) all samples processed together with no TEs or outlier libraries removed; (2) trimmed data with TEs and outlier libraries removed, male and female samples processed together; (3) trimmed data with TEs and outlier libraries removed, male and female samples processed separately.'}

Female.trim = c(nrow(sGR.F.CxDf2), nrow(sGR.F.CxOE2),
           nrow(sWT.F.CxDf2), nrow(sWT.F.CxOE2),
           nrow(sGRxWT.F2), nrow(sGRxWT.FDf2), nrow(sGRxWT.FOE2))

Male.trim = c(nrow(sGR.M.CxDf2), nrow(sGR.M.CxOE2),
         nrow(sWT.M.CxDf2), nrow(sWT.M.CxOE2),
         nrow(sGRxWT.M2), nrow(sGRxWT.MDf2), nrow(sGRxWT.MOE2))

Female.Only.trim = c(nrow(sGR.F.CxDf3), nrow(sGR.F.CxOE3),
           nrow(sWT.F.CxDf3), nrow(sWT.F.CxOE3),
           nrow(sGRxWT.F3), nrow(sGRxWT.FDf3), nrow(sGRxWT.FOE3))

Male.Only.trim = c(nrow(sGR.M.CxDf3), nrow(sGR.M.CxOE3),
         nrow(sWT.M.CxDf3), nrow(sWT.M.CxOE3),
         nrow(sGRxWT.M3), nrow(sGRxWT.MDf3), nrow(sGRxWT.MOE3))

Female = c(nrow(sGR.F.CxDf), nrow(sGR.F.CxOE),
           nrow(sWT.F.CxDf), nrow(sWT.F.CxOE),
           nrow(sGRxWT.F), nrow(sGRxWT.FDf), nrow(sGRxWT.FOE))

Male = c(nrow(sGR.M.CxDf), nrow(sGR.M.CxOE),
         nrow(sWT.M.CxDf), nrow(sWT.M.CxOE),
         nrow(sGRxWT.M), nrow(sGRxWT.MDf), nrow(sGRxWT.MOE))

Comparison = c("GR.C x GR.Df", "GR.C x GR.OE",
               "WT.C x WT.Df", "WT.C x WT.OE",
               "GR.C x WT.C", "GR.Df x WT.Df", "GR.OE x WT.OE")

data = data.frame(Female, Female.trim, Male, Male.trim, Comparison)

data$Comparison = factor(data$Comparison, levels = data$Comparison)

fig <- plot_ly(data, 
               x = ~Comparison, y = ~Female, 
               type = 'bar',
               name = "Female",
               marker = list(color =c("firebrick"),
                             line = list(color = "black", width = 1.5)))

fig <- fig %>% add_trace(y = ~Female.trim, name = 'Trim Female',
                         marker = list(color = c("red"),
                                       line = list(color = "black", width = 1.5)))

fig <- fig %>% add_trace(y = ~Female.Only.trim, name = 'Trim Only Female',
                         marker = list(color = c("pink"),
                                       line = list(color = "black", width = 1.5)))

fig <- fig %>% add_trace(y = ~Male, name = 'Males',
                         marker = list(color = c("blue"),
                                       line = list(color = "black", width = 1.5)))

fig <- fig %>% add_trace(y = ~Male.trim, name = 'Trim Males',
                         marker = list(color = c("cornflowerblue"),
                                       line = list(color = "black", width = 1.5)))

fig <- fig %>% add_trace(y = ~Male.Only.trim, name = 'Trim Only Males',
                         marker = list(color = c("skyblue"),
                                       line = list(color = "black", width = 1.5)))

fig <- fig %>% layout(yaxis = list(title ="Total DEGs"),
                      xaxis = list(title = NA),
                      title = "Total DEGs")

fig

```

```{r echo = F, include = F, fig.height=10, fig.cap='Heatmaps of Total DEGs in GR.FC x GR.FDf.'}

Samples = row.names(groups2)[groups2$Sex =='M'
                            & groups2$Genotype == 'GR']

hmdata = cpmdata2[Comparisons2[[8]],Samples]

hmdata=hmdata[2500:3000,]

hmdata=hmdata[intersect(row.names(hmdata),row.names(cpmdata)),]

hmdata=cbind(hmdata, cpmdata[row.names(hmdata),colnames(hmdata)])

colnames(hmdata)[10:18] = paste0("Old.",colnames(hmdata)[10:18])

row.names(hmdata) = convert[row.names(hmdata)]

cpm.text = round(hmdata,1)
i=1
while(i<=ncol(cpm.text)){
  cpm.text[,i] = paste0("CPM: ", cpm.text[,i])
  i = i+1
}

  heatmaply(hmdata,
            trace="none",
            col=RdYlBu(100)[100:1],
            scale="row",
            dendrogram = "both",
            Rowv=T,
            Colv=T,
            cexRow = .75, 
            na.color="grey",
            cexCol = .75,
            labRow = as.character(row.names(hmdata)),
            key = T,
            column_text_angle = 90,
            ##col_side_colors=groups[Samples,1:3],
            ##col_side_palette=Spectral,
            main= "GR.FC x GR.FDf",
            custom_hovertext = cpm.text)

```