### 2023-07-07 GeneralDataFiles/GenesInGOTerms.Rmd

---
title: "Lean Top PC1 Function Figures"
author: "John Santiago"
date: "11/28/2020"
output: html_document
---

```{r setup, include=FALSE}
library(rgl)
knitr::opts_chunk$set(echo = FALSE, fig.align="center")
knitr::knit_hooks$set(webgl = hook_webgl)
```

```{r include=F}
##Load Libraries

library(biomaRt)

library(org.Dm.eg.db)

library(GO.db)

```

```{r}
##FB2Ensembl
x1=as.list(org.Dm.egENSEMBL2EG)
##Ensembl2Symbol
x2=as.list(org.Dm.egSYMBOL)
##Symbol2Ensembl
x3=as.list(org.Dm.egSYMBOL2EG)
##Ensembl2FB
x4=as.list(org.Dm.egENSEMBL)
##ensembl to FBgn
x5=as.list(org.Dm.egFLYBASE)

##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/TKT_RNAseq/CountTables/"
cpmdata = read.csv(paste0(git.dir,"TKT_cpmdata.csv"),row.names = 1)
groups  = read.csv(paste0(git.dir,"TKT.metadata.csv"),row.names = 1)
TKT.EdgeR = read.csv(paste0(git.dir,"TKT.EdgeR.Table.csv"),row.names = 1)
convert=read.csv("https://raw.githubusercontent.com/DavidRandLab/Santiago-et-al-2021-BMC-Genomics/main/Data%20Files/FBgnConversionTable.csv",row.names=1)

FBgn2ensembl = row.names(cpmdata)
names(FBgn2ensembl) = row.names(cpmdata)
tempEG = unlist(x1[lengths(x1)==1][row.names(cpmdata)])
FBgn2ensembl[names(tempEG)]=tempEG

ensembl2symbol = row.names(cpmdata)
names(ensembl2symbol) = FBgn2ensembl
tempSymbol = unlist(x2[FBgn2ensembl])
ensembl2symbol[names(tempSymbol)]=tempSymbol
symbol2ensembl = names(ensembl2symbol)
names(symbol2ensembl) = ensembl2symbol

extras = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/UnknownIDGenes.csv")
row.names(extras)=extras$Symbol
extras$ensembl = extras$FBgn
tempExtras = unlist(x3[extras$Symbol])
extras[names(tempExtras), "ensembl"] = tempExtras
row.names(extras) = extras$FBgn

ensembl = FBgn2ensembl
ensembl[row.names(extras)] = extras$ensembl

symbol = ensembl2symbol
symbol[row.names(extras)] = extras$Symbol

KEY = data.frame(FBgn = row.names(cpmdata),
                 Symbol = symbol,
                 ensembl = ensembl)
row.names(KEY) = KEY$FBgn

##write.csv(KEY,"/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneIDKey.csv")

ccpar=as.list(GOCCPARENTS)
ccchild=as.list(GOCCCHILDREN)

bppar=as.list(GOBPPARENTS)
bpchild=as.list(GOBPCHILDREN)

mfpar=as.list(GOMFPARENTS)
mfchild=as.list(GOMFCHILDREN)

term=as.list(GOTERM)

genesingo=as.list(org.Dm.egGO2ALLEGS)

GOi = names(genesingo)[1]
genesingoi = intersect(row.names(KEY),unlist(x4[genesingo[[1]]]))

GO2FBgn = data.frame(FB.ID = genesingoi,
                     Symbol = KEY[genesingoi,"Symbol"],
                     GO.ID = GOi,
                     GO.Term = term[[names(genesingo)[1]]]@'Term',
                     Ontology = term[[names(genesingo)[1]]]@'Ontology',
                     GO.Definition = term[[names(genesingo)[1]]]@'Definition')

i=2
while(i<=length(genesingo)){
  GOi = names(genesingo)[i]
  genesingoi = intersect(row.names(KEY),unlist(x4[genesingo[[i]]]))

  if(length(genesingoi) > 0){
    tempGO2FBgn = data.frame(FB.ID = genesingoi,
                             Symbol = KEY[genesingoi,"Symbol"],
                             GO.ID = GOi,
                             GO.Term = term[[names(genesingo)[i]]]@'Term',
                             Ontology = term[[names(genesingo)[i]]]@'Ontology',
                             GO.Definition = term[[names(genesingo)[i]]]@'Definition')
    GO2FBgn = rbind(GO2FBgn, tempGO2FBgn)
  }
  i=i+1
}

##write.csv(GO2FBgn, "/Users/johncsantiago/Documents/GitHub/WhartonLab/GenesInGOTerms.csv")

Key = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)

kegg=keggLink("pathway","dme")
names(kegg)=gsub("dme:Dmel_","",names(kegg))
i=1
temp = as.list(kegg[(unique(names(kegg)))])
while(i<=length(temp)){
  temp[[i]] = as.character(kegg[grep(names(temp)[i], names(kegg))])
  i = i+1
}
kegg = temp
x=as.list(org.Dm.egFLYBASECG2EG)
xx=as.list(org.Dm.egSYMBOL)
names(kegg)=as.vector(unlist(xx[unlist(x[names(kegg)])]))

##saveRDS(kegg, "/Users/johncsantiago/Documents/GitHub/WhartonLab/Deficiency Screen/DataFiles/kegg.symbol2path.RData")

```