### 2023-07-10 DeficiencyScreen/KEGG_Analysis.Rmd
update

---
title: "Defficiency Screen Functional Analysis"
author: "John Santiago"
date: "11/28/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, include=F}
##Load Libraries
library(plotly)
library(pathview)
```

```{r echo = F, include =F}
##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"
GeneIDKey = read.csv(paste0(git.dir,"GeneralDataFiles/GeneIDKey.csv"),row.names = 1)
Df.data = read.csv(paste0(git.dir,"GeneralDataFiles/DeficiencyModifierGenes.csv"),row.names = 1)
Df.data$CG = GeneIDKey[row.names(Df.data),"CG"]
Df.data$Repress = 0
Df.data$Repress[Df.data$Moderate == 1|Df.data$Strong == 1] = 1
kegg = readRDS(gzcon(url(paste0(git.dir,"DeficiencyScreen/DataFiles/kegg.symbol2path.RData"))))
kegg.names = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/KEGG.names.csv"), row.names = 1)

Moderate.KEGG = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/ModerateKEGGAnalysis.csv"), row.names = 1)
Strong.KEGG = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/StrongKEGGAnalysis.csv"), row.names = 1)
Repress.KEGG = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/RepressorKEGGAnalysis.csv"), row.names = 1)
Lethal.KEGG = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/LethalKEGGAnalysis.csv"), row.names = 1)
Enhancer.KEGG = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/EnhancerKEGGAnalysis.csv"), row.names = 1)

```

```{r, include = F, echo = F}

if(intersect(list.files(), "KEGG_Image_Files") != "KEGG_Image_Files"){
  dir.create(paste0(dirname(rstudioapi::getSourceEditorContext()$path), "/KEGG_Image_Files/"))
}

Home = dirname(rstudioapi::getSourceEditorContext()$path)
OutputDir = paste0(Home, "/KEGG_Image_Files/")

ShowKEGG = function(PathwayID){

download.kegg(pathway.id= PathwayID, species = "dme", kegg.dir = OutputDir)
xml.file=paste0(OutputDir, "dme",PathwayID,".xml")
node.data=node.info(xml.file)

all.path = names(kegg[grep(PathwayID,kegg)])

Moderate = Df.data[Df.data$Moderate == 1, "Symbol"]
Moderate = intersect(Moderate, all.path)
Moderate = paste("Dmel_", GeneIDKey[GeneIDKey$Symbol %in% Moderate, "CG"],sep="")

Strong = Df.data[Df.data$Strong == 1, "Symbol"]
Strong = intersect(Strong, all.path)
Strong = paste("Dmel_", GeneIDKey[GeneIDKey$Symbol %in% Strong, "CG"],sep="")

Enhancer = Df.data[Df.data$Enhancer == 1, "Symbol"]
Enhancer = intersect(Enhancer, all.path)
Enhancer = paste("Dmel_", GeneIDKey[GeneIDKey$Symbol %in% Enhancer, "CG"],sep="")

Lethal = Df.data[Df.data$Lethal == 1, "Symbol"]
Lethal = intersect(Lethal, all.path)
Lethal = paste("Dmel_", GeneIDKey[GeneIDKey$Symbol %in% Strong, "CG"],sep="")

plot.data.gene=node.map(mol.data=c(Moderate, Strong, Enhancer, Lethal), node.data, node.types="gene")
cols.ts.gene = node.color(plot.data.gene, limit=2, bins=2)
cols.ts.gene[1:length(cols.ts.gene)] = 0

pdg.moderate=node.map(mol.data=Moderate, node.data, node.types="gene")
temp = node.color(pdg.moderate, limit=2, bins=2)
ctg.moderate = cols.ts.gene
ctg.moderate[temp!="#FFFFFF"]= 1 ##"gold2"

pdg.strong=node.map(mol.data=Strong, node.data, node.types="gene")
temp = node.color(pdg.strong, limit=2, bins=2)
ctg.strong = cols.ts.gene
ctg.strong[temp!="#FFFFFF"]= 1 ##"brown2"

pdg.enhancer=node.map(mol.data=Enhancer, node.data, node.types="gene")
temp = node.color(pdg.enhancer, limit=2, bins=2)
ctg.enhancer = cols.ts.gene
ctg.enhancer[temp!="#FFFFFF"]= 1 ##"deepskyblue"

pdg.lethal=node.map(mol.data=Lethal, node.data, node.types="gene")
temp = node.color(pdg.lethal, limit=2, bins=2)
ctg.lethal = cols.ts.gene
ctg.lethal[temp!="#FFFFFF"]= 1 ##"navyblue"

color.combos = data.frame(Moderate = as.numeric(ctg.moderate),
                          Strong = as.numeric(ctg.strong),
                          Enhancer = as.numeric(ctg.enhancer),
                          Lethal = as.numeric(ctg.lethal))
row.names(color.combos)= names(cols.ts.gene)

cols.ts.gene[cols.ts.gene=="0"]="darkseagreen1"

cols.ts.gene[row.names(color.combos)[color.combos$Moderate == 1]] = "gold2"
cols.ts.gene[row.names(color.combos)[color.combos$Strong   == 1]] = "brown1"
cols.ts.gene[row.names(color.combos)[color.combos$Enhancer   == 1]] = "lightsteelblue"
cols.ts.gene[row.names(color.combos)[color.combos$Lethal   == 1]] = "cornflowerblue"
cols.ts.gene[row.names(color.combos)[(color.combos$Moderate   == 1 | color.combos$Strong   == 1) & (color.combos$Enhancer   == 1 | color.combos$Lethal   == 1) ]] = "grey60"

setwd(OutputDir)
pv.pars= keggview.native(plot.data.gene=plot.data.gene,
                         cols.ts.gene=cols.ts.gene, 
                         node.data=node.data,
                         pathway.name=paste0("dme", PathwayID),
                         same.layer=T, plot.col.key=F, out.suffix = kegg.names[paste0("dme", PathwayID),1],
                         kegg.dir = OutputDir)
setwd(Home)
}

```

```{r, echo = F, include = F}

KEGG20 = function(data){

  data = data[1:20, c("category", "Name","adjp", "numDEInCat", "numInCat")]

data = data[20:1,]
data$Score = -log10(data$adjp)
data$term = factor(data$Name, levels = c(unique(data$Name)))

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

fig <- plot_ly(data, 
               x = ~numDEInCat, y = ~term, 
               type = 'bar',
               name = "Name",
               marker = list(colorscale = list(c(0,.5,1), c("brown","gold", "steelblue")),
    reversescale =T,
                             colorbar = list(title = "-log10(FDR)",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                             color = ~Score,
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$Name,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = "Top 20 Most Enriched KEGG Terms",
           margin = m)

return(fig)

}

```

```{r fig.width=10}

##Options: Moderate.KEGG, Strong.KEGG, Repress.KEGG, Lethal.KEGG, Enhancer.KEGG
data = Repress.KEGG

KEGG20(data)

```

```{r echo = F, include = F}

PathwayID = "04141"

ShowKEGG(PathwayID)

```

```{r echo = F}

knitr::include_graphics(paste0(OutputDir, "dme",PathwayID, ".", kegg.names[paste0("dme", PathwayID),1], ".png"))

```

```{r include = F, echo = F}

GenesInKEGG = Df.data[Df.data$Symbol %in% names(kegg[grep(PathwayID, kegg)]),]
write.csv(GenesInKEGG, paste0(OutputDir,"GenesInKEGG_dme", PathwayID,".csv"))

```