### 2023-07-10 DeficiencyScreen/GO_Analysis.Rmd
update

---
title: "Defficiency Screen Functional Analysis"
author: "John Santiago"
date: "11/28/2020"
output: html_document
---

```{r setup, include=FALSE}
library(rgl)
knitr::opts_chunk$set(echo = FALSE, fig.align="center")
knitr::knit_hooks$set(webgl = hook_webgl)
```

```{r echo = F, include=F}
##Load Libraries

##library(biomaRt)
##library(org.Dm.eg.db)
##library(goseq)
##library(GO.db)
##library(visNetwork)
library(rrvgo)
library(plotly)
##library(KEGGREST)
##library(pathview)
##library(RColorBrewer)
```

```{r echo = F, include =F}
##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"
GeneIDKey = read.csv(paste0(git.dir,"GeneralDataFiles/GeneIDKey.csv"),row.names = 1)
Df.data = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/DeficiencyModifierGenes.csv"),row.names = 1)
##kegg = readRDS(gzcon(url(paste0(git.dir,"DeficiencyScreen/DataFiles/kegg.symbol2path.RData"))))
##kegg.names = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/KEGG.names.csv"), row.names = 1)

Moderate.GO = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/ModerateGOAnalysis.csv"), row.names = 1)
Strong.GO = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/StrongGOAnalysis.csv"), row.names = 1)
Repress.GO = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/RepressorGOAnalysis.csv"), row.names = 1)
Lethal.GO = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/LethalGOAnalysis.csv"), row.names = 1)
Enhancer.GO = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/EnhancerGOAnalysis.csv"), row.names = 1)

```

```{r include=F, echo=F}

##sigs = Df.data[Df.data$Strong == 1 | Df.data$Moderate == 1 , "Symbol"]
##genes = setNames(rep(0, nrow(Df.data)), Df.data$Symbol)
##genes[sigs] = 1
##pwf=nullp(genes,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol")
##GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
##row.names(GO.wall) = GO.wall$category
##Repress.GO = GO.wall
##write.csv(Repress.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/Deficiency Screen/DataFiles/RepressorGOAnalysis.csv")

##KEGG term enrichment is KEGG
##KEGG=goseq(pwf,gene2cat=kegg)
##KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
##row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
##KEGG$Name=kegg.names[row.names(KEGG),1]
##Repress.KEGG = KEGG
##write.csv(Repress.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/Deficiency Screen/DataFiles/RepressorKEGGAnalysis.csv")

##sigs = Df.data[Df.data$Moderate == 1 , "Symbol"]
##genes = setNames(rep(0, nrow(Df.data)), Df.data$Symbol)
##genes[sigs] = 1
##pwf=nullp(genes,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol")
##GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
##row.names(GO.wall) = GO.wall$category
##Moderate.GO = GO.wall
##write.csv(Moderate.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/Deficiency Screen/DataFiles/ModerateGOAnalysis.csv")

##KEGG term enrichment is KEGG
##KEGG=goseq(pwf,gene2cat=kegg)
##KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
##row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
##KEGG$Name=kegg.names[row.names(KEGG),1]
##Moderate.KEGG = KEGG
##write.csv(Moderate.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/Deficiency Screen/DataFiles/ModerateKEGGAnalysis.csv")

##sigs = Df.data[Df.data$Strong == 1 , "Symbol"]
##genes = setNames(rep(0, nrow(Df.data)), Df.data$Symbol)
##genes[sigs] = 1
##pwf=nullp(genes,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol")
##GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
##row.names(GO.wall) = GO.wall$category
##Strong.GO = GO.wall
##write.csv(Strong.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/Deficiency Screen/DataFiles/StrongGOAnalysis.csv")

##KEGG term enrichment is KEGG
##KEGG=goseq(pwf,gene2cat=kegg)
##KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
##row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
##KEGG$Name=kegg.names[row.names(KEGG),1]
##Strong.KEGG = KEGG
##write.csv(Strong.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/Deficiency Screen/DataFiles/StrongKEGGAnalysis.csv")

##sigs = Df.data[Df.data$Lethal == 1, "Symbol"]
##genes = setNames(rep(0, nrow(Df.data)), Df.data$Symbol)
##genes[sigs] = 1
##pwf=nullp(genes,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol")
##GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
##row.names(GO.wall) = GO.wall$category
##Lethal.GO = GO.wall
##write.csv(Lethal.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/Deficiency Screen/DataFiles/LethalGOAnalysis.csv")

##KEGG term enrichment is KEGG
##KEGG=goseq(pwf,gene2cat=kegg)
##KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
##row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
##KEGG$Name=kegg.names[row.names(KEGG),1]
##Lethal.KEGG = KEGG
##write.csv(Lethal.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/Deficiency Screen/DataFiles/LethalKEGGAnalysis.csv")

##sigs = Df.data[Df.data$Enhancer == 1, "Symbol"]
##genes = setNames(rep(0, nrow(Df.data)), Df.data$Symbol)
##genes[sigs] = 1
##pwf=nullp(genes,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol")
##GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
##row.names(GO.wall) = GO.wall$category
##Enhance.GO = GO.wall
##write.csv(Enhance.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/Deficiency Screen/DataFiles/EnhancerGOAnalysis.csv")

##KEGG term enrichment is KEGG
##KEGG=goseq(pwf,gene2cat=kegg)
##KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
##row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
##EGG$Name=kegg.names[row.names(KEGG),1]
##Enhance.KEGG = KEGG
##write.csv(Enhance.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/Deficiency Screen/DataFiles/EnhancerKEGGAnalysis.csv")

```

```{r echo = F, include =F}

GO20 = function(GO.data, GO.ontology){
GO.data = GO.data[GO.data$ontology == GO.ontology,
                  c("category", "term","adjp", "numDEInCat", "numInCat")]

data = GO.data[20:1,c("term","adjp","numDEInCat", "numInCat", "category")]
data$Score = -log10(data$adjp)
data$term = factor(data$term, levels = c(unique(data$term)))

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

fig <- plot_ly(data, 
               x = ~numDEInCat, y = ~term, 
               type = 'bar',
               name = "term",
               marker = list(colorscale = list(c(0,.5,1), c("brown","gold", "steelblue")),
    reversescale =T,
                             colorbar = list(title = "-log10(FDR)",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                             color = ~Score,
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$term,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = "Top 20 Most Enriched GO Terms",
           margin = m)

  return(fig)
}

```

```{r echo = F, include =F}

GO.data = Repress.GO
GO.ontology = "BP"

GO.fig = GO20(GO.data, GO.ontology)

```

```{r, fig.width= 10}

GO.fig

```

```{r include=F, echo=F}

go_analysis = GO.data$category[GO.data$adjp <= .05]
scores = setNames(-log10(GO.data$adjp), go_analysis)

simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

reducedTerms <- reduceSimMatrix(simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r fig.height=7, fig.width=10}

heatmapPlot(simMatrix,
            reducedTerms,
            annotateParent=TRUE,
            annotationLabel="parentTerm",
            fontsize=6,
            main = "Repressor Gene Enriched Biological Process GO terms")

```

```{r, fig.width=10}
 use.data   = simMatrix
    pca <- prcomp(t(use.data), scale.=TRUE) 
    pc1=scale(pca$x[,1])
    pc2=scale(pca$x[,2])
    eigs <- pca$sdev^2
    ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(PC1=pc1,
                        PC2=pc2,
                        sdev=pca$sdev,
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group = reducedTerms[row.names(use.data), "parentTerm"],
                        Score = reducedTerms[row.names(use.data), "score"],
                        Term = reducedTerms[row.names(use.data), "term"],
                        TotalDE = GO.data[row.names(use.data), "numDEInCat"],
                        Total = GO.data[row.names(use.data), "numInCat"],
                        SigVal = signif(GO.data[row.names(use.data), "adjp"],3))

    pca.data$Group = factor(pca.data$Group, levels = unique(pca.data$Group))
    pca.data$Score = 20*(1+pca.data$Score/mean(pca.data$Score))

  fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 ##symbol = ~Group,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c("firebrick","gold","deepskyblue"),
                 marker = list(size = pca.data$Score,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Term, 
                                   "\nTotal DE: ", pca.data$TotalDE, 
                                   "\nTotal in Cat.: ", pca.data$Total, 
                                   "\nFDR: ", pca.data$SigVal)) %>%
    layout(xaxis = list(title = paste0("PC1", "(", ve[1], "%)")),
           yaxis = list(title = paste0("PC2", "(", ve[2], "%)")),
           title = "Enriched GO term similarity")

  fig

```

```{r fig.align='left'}

treemapPlot(reducedTerms)

```