### 2023-07-17 TKT_RNAseq/R Code/GO_Analysis.Rmd

---
title: "Defficiency Screen Functional Analysis"
author: "John Santiago"
date: "11/28/2020"
output: html_document
---

```{r setup, include=FALSE}
library(rgl)
knitr::opts_chunk$set(echo = FALSE, fig.align="center")
knitr::knit_hooks$set(webgl = hook_webgl)
```

```{r echo = F, include=F}
##Load Libraries

library(biomaRt)
library(org.Dm.eg.db)
library(goseq)
library(GO.db)
library(visNetwork)
library(rrvgo)
library(plotly)
library(KEGGREST)
library(pathview)
##library(RColorBrewer)
```

```{r echo = F, include =F}
##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"
cpmdata = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
groups  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
TKT.EdgeR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.Table.csv"),row.names = 1)
GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)
kegg = readRDS(gzcon(url(paste0(git.dir,"GeneralDataFiles/kegg.symbol2path.RData"))))
kegg.names = read.csv(paste0(git.dir, "GeneralDataFiles/KEGG.names.csv"), row.names = 1)

##keggID=keggConv("dme","ncbi-geneid")

datadir = "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/"

GRF.CxDf.GO  = read.csv(paste0(datadir, "GRF.CxDf.GOAnalysis.csv"), row.names =1)
GRF.CxOE.GO  = read.csv(paste0(datadir, "GRF.CxOE.GOAnalysis.csv"), row.names =1)
WTF.CxDf.GO  = read.csv(paste0(datadir, "WTF.CxDf.GOAnalysis.csv"), row.names =1)
WTF.CxOE.GO  = read.csv(paste0(datadir, "WTF.CxOE.GOAnalysis.csv"), row.names =1)

GRM.CxDf.GO  = read.csv(paste0(datadir, "GRM.CxDf.GOAnalysis.csv"), row.names =1)
GRM.CxOE.GO  = read.csv(paste0(datadir, "GRM.CxOE.GOAnalysis.csv"), row.names =1)
WTM.CxDf.GO  = read.csv(paste0(datadir, "WTM.CxDf.GOAnalysis.csv"), row.names =1)
WTM.CxOE.GO  = read.csv(paste0(datadir, "WTM.CxOE.GOAnalysis.csv"), row.names =1)

GR.FxM.GO    = read.csv(paste0(datadir, "GR.FxM.GOAnalysis.csv"), row.names =1)
WT.FxM.GO    = read.csv(paste0(datadir, "WT.FxM.GOAnalysis.csv"), row.names =1)

GRxWT.FC.GO  = read.csv(paste0(datadir, "GRxWT.FC.GOAnalysis.csv"), row.names =1)
GRxWT.FDf.GO = read.csv(paste0(datadir, "GRxWT.FDf.GOAnalysis.csv"), row.names =1)
GRxWT.FOE.GO = read.csv(paste0(datadir, "GRxWT.FOE.GOAnalysis.csv"), row.names =1)

GRxWT.MC.GO  = read.csv(paste0(datadir, "GRxWT.MC.GOAnalysis.csv"), row.names =1)
GRxWT.MDf.GO = read.csv(paste0(datadir, "GRxWT.MDf.GOAnalysis.csv"), row.names =1)
GRxWT.MOE.GO = read.csv(paste0(datadir, "GRxWT.MOE.GOAnalysis.csv"), row.names =1)

SigDEGs = TKT.EdgeR
SigDEGs[,] = 0

SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$GRF.CxDf <= .05) & (apply(cpmdata[,groups$Group == "GR.F" | groups$Group == "TktDfGR.F"],1,min) > 5)], "GRF.CxDf"] = 1

SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$GRF.CxOE <= .05) & (apply(cpmdata[,groups$Group == "GR.F" | groups$Group == "TktOEGR.F"],1,min) > 5)], "GRF.CxOE"] = 1

SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$WTF.CxDf <= .05) & (apply(cpmdata[,groups$Group == "WT.F" | groups$Group == "TktDfWT.F"],1,min) > 5)], "WTF.CxDf"] = 1

SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$WTF.CxOE <= .05) & (apply(cpmdata[,groups$Group == "WT.F" | groups$Group == "TktOEWT.F"],1,min) > 5)], "WTF.CxOE"] = 1

SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$GRxWT.FC <= .05) & (apply(cpmdata[,groups$Group == "GR.F" | groups$Group == "WT.F"],1,min) > 5)], "GRxWT.FC"] = 1

SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$GRxWT.FDf <= .05) & (apply(cpmdata[,groups$Group == "TktDfGR.F" | groups$Group == "TktDfWT.F"],1,min) > 5)], "GRxWT.FDf"] = 1

SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$GRxWT.FOE <= .05) & (apply(cpmdata[,groups$Group == "TktOEGR.F" | groups$Group == "TktOEWT.F"],1,min) > 5)], "WTF.CxOE"] =1

SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$GRM.CxDf <= .05) & (apply(cpmdata[,groups$Group == "GR.M" | groups$Group == "TktDfGR.M"],1,min) > 5)], "GRM.CxDf"] = 1

SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$GRM.CxOE <= .05) & (apply(cpmdata[,groups$Group == "GR.M" | groups$Group == "TktOEGR.M"],1,min) > 5)], "GRM.CxOE"] = 1

SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$WTM.CxDf <= .05) & (apply(cpmdata[,groups$Group == "WT.M" | groups$Group == "TktDfWT.M"],1,min) > 5)], "WTM.CxDf"] = 1

SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$WTM.CxOE <= .05) & (apply(cpmdata[,groups$Group == "WT.M" | groups$Group == "TktOEWT.M"],1,min) > 5)], "WTM.CxOE"] = 1

SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$GRxWT.MC <= .05) & (apply(cpmdata[,groups$Group == "GR.M" | groups$Group == "WT.F"],1,min) > 5)], "GRxWT.MC"] = 1

SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$GRxWT.MDf <= .05) & (apply(cpmdata[,groups$Group == "TktDfGR.M" | groups$Group == "TktDfWT.M"],1,min) > 5)], "GRxWT.MDf"] = 1

SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$GRxWT.MOE <= .05) & (apply(cpmdata[,groups$Group == "TktOEGR.M" | groups$Group == "TktOEWT.M"],1,min) > 5)], "GRxWT.MOE"] = 1
SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$GR.FxM  <= .05)], "GR.FxM"] = 1
SigDEGs[row.names(TKT.EdgeR)[(TKT.EdgeR$WT.FxM  <= .05)], "WT.FxM"] = 1
```

##GOterm and KEGG enrichment analysis for all EdgeR comparisons
##```{r include=F, echo=F}

##############
## GRF.CxDf ##
##############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$GRF.CxDf <= .05) & (apply(cpmdata[,groups$Group == "GR.F" | groups$Group == "TktDfGR.F"],1,min) > 5)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
GRF.CxDf.GO = GO.wall
write.csv(GRF.CxDf.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRF.CxDf.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
GRF.CxDf.KEGG = KEGG
write.csv(GRF.CxDf.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRF.CxDf.KEGGAnalysis.csv")

##############
## GRF.CxOE ##
##############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$GRF.CxOE <= .05) & (apply(cpmdata[,groups$Group == "GR.F" | groups$Group == "TktOEGR.F"],1,min) > 5)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
GRF.CxOE.GO = GO.wall
write.csv(GRF.CxOE.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRF.CxOE.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
GRF.CxOE.KEGG = KEGG
write.csv(GRF.CxOE.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRF.CxOE.KEGGAnalysis.csv")

##############
## WTF.CxDf ##
##############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$WTF.CxDf <= .05) & (apply(cpmdata[,groups$Group == "WT.F" | groups$Group == "TktDfWT.F"],1,min) > 5)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
WTF.CxDf.GO = GO.wall
write.csv(WTF.CxDf.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/WTF.CxDf.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
WTF.CxDf.KEGG = KEGG
write.csv(WTF.CxDf.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/WTF.CxDf.KEGGAnalysis.csv")

##############
## WTF.CxOE ##
##############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$WTF.CxOE <= .05) & (apply(cpmdata[,groups$Group == "WT.F" | groups$Group == "TktOEWT.F"],1,min) > 5)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
WTF.CxOE.GO = GO.wall
write.csv(WTF.CxOE.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/WTF.CxOE.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
WTF.CxOE.KEGG = KEGG
write.csv(WTF.CxOE.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/WTF.CxOE.KEGGAnalysis.csv")

##############
## GRM.CxDf ##
##############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$GRM.CxDf <= .05) & (apply(cpmdata[,groups$Group == "GR.M" | groups$Group == "TktDfGR.M"],1,min) > 5)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
GRM.CxDf.GO = GO.wall
write.csv(GRM.CxDf.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRM.CxDf.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
GRM.CxDf.KEGG = KEGG
write.csv(GRM.CxDf.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRM.CxDf.KEGGAnalysis.csv")

##############
## GRM.CxOE ##
##############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$GRM.CxOE <= .05) & (apply(cpmdata[,groups$Group == "GR.M" | groups$Group == "TktOEGR.M"],1,min) > 5)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
GRM.CxOE.GO = GO.wall
write.csv(GRM.CxOE.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRM.CxOE.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
GRM.CxOE.KEGG = KEGG
write.csv(GRM.CxOE.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRM.CxOE.KEGGAnalysis.csv")

##############
## WTM.CxDf ##
##############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$WTM.CxDf <= .05) & (apply(cpmdata[,groups$Group == "WT.M" | groups$Group == "TktDfWT.M"],1,min) > 5)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
WTM.CxDf.GO = GO.wall
write.csv(WTM.CxDf.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/WTM.CxDf.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
WTM.CxDf.KEGG = KEGG
write.csv(WTM.CxDf.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/WTM.CxDf.KEGGAnalysis.csv")

##############
## WTM.CxOE ##
##############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$WTM.CxOE <= .05) & (apply(cpmdata[,groups$Group == "WT.M" | groups$Group == "TktOEWT.M"],1,min) > 5)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
WTM.CxOE.GO = GO.wall
write.csv(WTM.CxOE.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/WTM.CxOE.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
WTM.CxOE.KEGG = KEGG
write.csv(WTM.CxOE.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/WTM.CxOE.KEGGAnalysis.csv")

##############
## GRxWT.FC ##
##############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$GRxWT.FC <= .05) & (apply(cpmdata[,groups$Group == "GR.F" | groups$Group == "WT.F"],1,min) > 5)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
GRxWT.FC.GO = GO.wall
write.csv(GRxWT.FC.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRxWT.FC.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
GRxWT.FC.KEGG = KEGG
write.csv(GRxWT.FC.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRxWT.FC.KEGGAnalysis.csv")

###############
## GRxWT.FDf ##
###############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$GRxWT.FDf <= .05) & (apply(cpmdata[,groups$Group == "TktDfGR.F" | groups$Group == "TktDfWT.F"],1,min) > 5)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
GRxWT.FDf.GO = GO.wall
write.csv(GRxWT.FDf.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRxWT.FDf.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
GRxWT.FDf.KEGG = KEGG
write.csv(GRxWT.FDf.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRxWT.FDf.KEGGAnalysis.csv")

###############
## GRxWT.FOE ##
###############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$GRxWT.FOE <= .05) & (apply(cpmdata[,groups$Group == "TktOEGR.F" | groups$Group == "TktOEWT.F"],1,min) > 5)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
GRxWT.FOE.GO = GO.wall
write.csv(GRxWT.FOE.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRxWT.FOE.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
GRxWT.FOE.KEGG = KEGG
write.csv(GRxWT.FOE.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRxWT.FOE.KEGGAnalysis.csv")

##############
## GRxWT.MC ##
##############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$GRxWT.MC <= .05) & (apply(cpmdata[,groups$Group == "GR.M" | groups$Group == "WT.M"],1,min) > 5)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
GRxWT.MC.GO = GO.wall
write.csv(GRxWT.MC.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRxWT.MC.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
GRxWT.MC.KEGG = KEGG
write.csv(GRxWT.MC.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRxWT.MC.KEGGAnalysis.csv")

###############
## GRxWT.MDf ##
###############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$GRxWT.MDf  <= .05) & (apply(cpmdata[,groups$Group == "TktDfGR.M" | groups$Group == "TktDfWT.M"],1,min) > 5)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
GRxWT.MDf.GO = GO.wall
write.csv(GRxWT.MDf.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRxWT.MDf.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
GRxWT.MDf.KEGG = KEGG
write.csv(GRxWT.MDf.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRxWT.MDf.KEGGAnalysis.csv")

###############
## GRxWT.MOE ##
###############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$GRxWT.MOE  <= .05) & (apply(cpmdata[,groups$Group == "TktOEGR.M" | groups$Group == "TktOEWT.M"],1,min) > 5)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
GRxWT.MOE.GO = GO.wall
write.csv(GRxWT.MOE.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRxWT.MOE.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
GRxWT.MOE.KEGG = KEGG
write.csv(GRxWT.MOE.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GRxWT.MOE.KEGGAnalysis.csv")

############
## GR.FxM ##
############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$GR.FxM  <= .05)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
GR.FxM.GO = GO.wall
write.csv(GR.FxM.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GR.FxM.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
GR.FxM.KEGG = KEGG
write.csv(GR.FxM.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/GR.FxM.KEGGAnalysis.csv")

############
## WT.FxM ##
############
sigs = GeneIDKey[row.names(TKT.EdgeR)[(TKT.EdgeR$WT.FxM  <= .05)], "Symbol"]
genes = setNames(rep(0, nrow(cpmdata)), GeneIDKey[row.names(cpmdata), "Symbol"])
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
WT.FxM.GO = GO.wall
write.csv(WT.FxM.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/WT.FxM.GOAnalysis.csv")

KEGG=goseq(pwf,gene2cat=kegg)
KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
KEGG$Name=kegg.names[row.names(KEGG),1]
WT.FxM.KEGG = KEGG
write.csv(WT.FxM.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/WT.FxM.KEGGAnalysis.csv")

##```

##GO20 Function
```{r echo = F, include =F}

GO20 = function(GO.data, GO.ontology){
GO.data = GO.data[GO.data$ontology == GO.ontology,
                  c("category", "term","adjp", "numDEInCat", "numInCat")]

data = GO.data[20:1,c("term","adjp","numDEInCat", "numInCat", "category")]
data$Score = -log10(data$adjp)
data$term = factor(data$term, levels = c(unique(data$term)))

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

fig <- plot_ly(data, 
               x = ~numDEInCat, y = ~term, 
               type = 'bar',
               name = "term",
               marker = list(colorscale = list(c(0,.5,1), c("brown","gold", "steelblue")),
    reversescale =T,
                             colorbar = list(title = "-log10(FDR)",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                             color = ~Score,
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$term,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = "Top 20 Most Enriched GO Terms",
           margin = m)

  return(fig)
}

```

##KEGG20 Function
```{r, echo = F, include = F}

KEGG20 = function(data){

  data = data[1:20, c("category", "Name","adjp", "numDEInCat", "numInCat")]

data = data[20:1,]
data$Score = -log10(data$adjp)
data$term = factor(data$Name, levels = c(unique(data$Name)))

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

fig <- plot_ly(data, 
               x = ~numDEInCat, y = ~term, 
               type = 'bar',
               name = "Name",
               marker = list(colorscale = list(c(0,.5,1), c("brown","gold", "steelblue")),
    reversescale =T,
                             colorbar = list(title = "-log10(FDR)",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                             color = ~Score,
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$Name,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = "Top 20 Most Enriched KEGG Terms",
           margin = m)

return(fig)

}

```

##Pathview function
```{r, include = F, echo = F}

if(length(intersect(list.files(), "KEGG_Image_Files") != "KEGG_Image_Files") == 0){
  dir.create(paste0(dirname(rstudioapi::getSourceEditorContext()$path), "/KEGG_Image_Files/"))
}

Home = dirname(rstudioapi::getSourceEditorContext()$path)
OutputDir = paste0(Home, "/KEGG_Image_Files/")

ShowKEGG = function(PathwayID, DEG1, DEG2){

  if(length(intersect(list.files(OutputDir), paste0("dme",PathwayID,".xml")))==0){
    download.kegg(pathway.id= PathwayID, species = "dme", kegg.dir = OutputDir)
  }  
  xml.file=paste0(OutputDir, "dme",PathwayID,".xml")
  node.data=node.info(xml.file)

  all.path = names(kegg[grep(PathwayID,kegg)])
  all.path = intersect(GeneIDKey[row.names(TKT.EdgeR),"Symbol"],all.path)

  DEG1 = GeneIDKey[row.names(SigDEGs)[SigDEGs[,DEG1] == 1], "Symbol"]
  DEG1 = intersect(DEG1, all.path)
  DEG1 = paste("Dmel_", GeneIDKey[GeneIDKey$Symbol %in% DEG1, "CG"],sep="")

  if(length(intersect(DEG2, colnames(SigDEGs))) ==1 ){
  DEG2 = GeneIDKey[row.names(SigDEGs)[SigDEGs[,DEG2] == 1], "Symbol"]
  DEG2 = intersect(DEG2, all.path)
  DEG2 = paste("Dmel_", GeneIDKey[GeneIDKey$Symbol %in% DEG2, "CG"],sep="")    
  }

plot.data.gene=node.map(mol.data=c(DEG1, DEG2), node.data, node.types="gene")
cols.ts.gene = node.color(plot.data.gene, limit=2, bins=2)
cols.ts.gene[1:length(cols.ts.gene)] = 0

pdg.DEG1 = node.map(mol.data = DEG1, node.data, node.types="gene")
temp = node.color(pdg.DEG1, limit=2, bins=2)
ctg.DEG1 = cols.ts.gene
ctg.DEG1[temp!="#FFFFFF"]= 1

pdg.DEG2 = node.map(mol.data = DEG2, node.data, node.types="gene")
temp = node.color(pdg.DEG2, limit=2, bins=2)
ctg.DEG2 = cols.ts.gene
ctg.DEG2[temp!="#FFFFFF"]= 1

color.combos = data.frame(DEG1 = as.numeric(ctg.DEG1),
                          DEG2 = as.numeric(ctg.DEG2))
row.names(color.combos)= names(cols.ts.gene)

cols.ts.gene[cols.ts.gene=="0"]="gainsboro"

cols.ts.gene[row.names(color.combos)[color.combos$DEG1 == 1]] = "cornflowerblue"
cols.ts.gene[row.names(color.combos)[color.combos$DEG2   == 1]] = "firebrick"
cols.ts.gene[row.names(color.combos)[color.combos$DEG2   == 1 & color.combos$DEG1 == 1]] = "gold"

setwd(OutputDir)
pv.pars= keggview.native(plot.data.gene=plot.data.gene,
                         cols.ts.gene=cols.ts.gene, 
                         node.data=node.data,
                         pathway.name=paste0("dme", PathwayID),
                         same.layer=T, plot.col.key=F, out.suffix = kegg.names[paste0("dme", PathwayID),1],
                         kegg.dir = OutputDir)
setwd(Home)
}

```

```{r echo = F, include =F}

GO.data = GRF.CxDf.GO
GO.ontology = "BP"
BP.GO.fig = GO20(GO.data, GO.ontology)

GO.ontology = "CC"
CC.GO.fig = GO20(GO.data, GO.ontology)

GO.ontology = "MF"
MF.GO.fig = GO20(GO.data, GO.ontology)

KEGG.data = GRxWT.FDf.KEGG
KEGG.fig = KEGG20(KEGG.data)

```

```{r, fig.width= 10}

BP.GO.fig

```

```{r, fig.width= 10}

CC.GO.fig

```

```{r, fig.width= 10}

MF.GO.fig

```

```{r fig.width=10}

KEGG.fig

```

```{r echo = F, include = F}
PathwayID = "04310"
PathwayID = "00480" ##Glutathione Metabolism
PathwayID = "00230" ##Purine metabolism
DEG1 = "GRF.CxDf"
DEG2 = "GRF.CxOE"
ShowKEGG(PathwayID, DEG1, DEG2)

```

```{r}
knitr::include_graphics(paste0(OutputDir,"/dme",PathwayID, ".", kegg.names[paste0("dme", PathwayID),1], ".png"))
```

```{r  echo = F}

geneset = GeneIDKey[row.names(TKT.EdgeR)[TKT.EdgeR$GRF.CxDf <= .05],]

GenesInKEGG = geneset[geneset$Symbol %in% names(kegg[grep(PathwayID, kegg)]), ]

hmdata = cpmdata[row.names(GenesInKEGG), groups$Group %in% c("GR.F",  "TktDfGR.F")]##, "TktDfGR.F","TktDfWT.F")]

heatmaply(hmdata,
            trace="none",
            col=RdYlBu(100)[100:1],
            scale="row",
            dendrogram = "both",
            show_dendrogram = c(F,T),
            Rowv=T,
            Colv=T,
            na.color="grey",
            labRow = GeneIDKey[row.names(hmdata), "Symbol"],
            cexCol = 1,
            key = T,
            column_text_angle = 90)

```

```{r echo =F, include =F}

GO.data = GRxWT.FC.GO
GO.ontology = "BP"

go_analysis = GO.data$category[GO.data$adjp <= .05 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$adjp), go_analysis)

simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

reducedTerms <- reduceSimMatrix(simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r fig.height=7, fig.width=10}

heatmapPlot(simMatrix,
            reducedTerms,
            annotateParent=TRUE,
            annotationLabel="parentTerm",
            fontsize=6,
            main = "Repressor Gene Enriched Biological Process GO terms")

```

```{r, fig.align='left', include=F}

scatterPlot(simMatrix, reducedTerms)

```

```{r, fig.width=10}
 use.data   = simMatrix
    pca <- prcomp(t(use.data), scale.=TRUE) 
    pc1=scale(pca$x[,1])
    pc2=scale(pca$x[,2])
    eigs <- pca$sdev^2
    ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(PC1=pc1,
                        PC2=pc2,
                        sdev=pca$sdev,
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group = reducedTerms[row.names(use.data), "parentTerm"],
                        Score = reducedTerms[row.names(use.data), "score"],
                        Term = reducedTerms[row.names(use.data), "term"],
                        TotalDE = GO.data[row.names(use.data), "numDEInCat"],
                        Total = GO.data[row.names(use.data), "numInCat"],
                        SigVal = signif(GO.data[row.names(use.data), "adjp"],3))

    pca.data$Group = factor(pca.data$Group, levels = unique(pca.data$Group))
    pca.data$Score = 20*(1+pca.data$Score/mean(pca.data$Score))

  fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 ##symbol = ~Group,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c("firebrick","gold","deepskyblue"),
                 marker = list(size = pca.data$Score,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Term, 
                                   "\nTotal DE: ", pca.data$TotalDE, 
                                   "\nTotal in Cat.: ", pca.data$Total, 
                                   "\nFDR: ", pca.data$SigVal)) %>%
    layout(xaxis = list(title = paste0("PC1", "(", ve[1], "%)")),
           yaxis = list(title = paste0("PC2", "(", ve[2], "%)")),
           title = "Enriched GO term similarity")

  fig

```

```{r fig.align='left'}

treemapPlot(reducedTerms)

```