### 2023-07-26 Metabolomics/MetaboanalystRCode.Rmd
update

---
title: "Metaboanalyst"
author: "John Santiago"
date: "2023-07-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, include = F}

library("MetaboAnalystR")

mSet<-InitDataObjects("pktable", "stat", FALSE)

mSet = Read.TextData(mSet, "/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/MetabolomicsForMetaboanalyst.txt", "colu", "disc")

mSet = SanityCheckData(mSet)

mSet<-ReplaceMin(mSet)
mSet<-SanityCheckData(mSet)
mSet<-FilterVariable(mSet, "none", -1, "F", 25, F)
mSet<-PreparePrenormData(mSet)

mSet<-Normalization(mSet, "CompNorm", "NULL", "NULL", "TIC", ratio=FALSE, ratioNum=20)
##mSet<-PlotNormSummary(mSet, "norm_0_", "png", 72, width=NA)
##mSet<-PlotSampleNormSummary(mSet, "snorm_0_", "png", 72, width=NA)

```

```{r}

mSet<-GetGroupNames(mSet, "")
feature.nm.vec <- c("")
smpl.nm.vec <- c("")
grp.nm.vec <- c("C","E")
mSet<-UpdateData(mSet)
mSet<-PreparePrenormData(mSet)
mSet<-Normalization(mSet, "CompNorm", "NULL", "NULL", "TIC", ratio=FALSE, ratioNum=20)
##mSet<-PlotNormSummary(mSet, "norm_1_", "png", 72, width=NA)
##mSet<-PlotSampleNormSummary(mSet, "snorm_1_", "png", 72, width=NA)
mSet<-FC.Anal(mSet, 2.0, 0, FALSE)
##mSet<-PlotFC(mSet, "fc_0_", "png", 72, width=NA)

mSet$analSet$fc
```