### 2023-08-07 Metabolomics/MetaboanalystRCode.Rmd

---
title: "Metaboanalyst"
author: "John Santiago"
date: "2023-07-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, include = F}

library("MetaboAnalystR")

rm(mSet)

mSet<-InitDataObjects("pktable", "stat", FALSE)

mSet = Read.TextData(mSet, "/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/MetabolomicsForMetaboanalyst.txt", "colu", "disc")

mSet = SanityCheckData(mSet)

mSet<-ReplaceMin(mSet)
mSet<-SanityCheckData(mSet)
mSet<-FilterVariable(mSet, "none", -1, "F", 25, F)
mSet<-PreparePrenormData(mSet)

##Normalization just takes the value, divides by TIC and multiply by 1000
mSet<-Normalization(mSet, "CompNorm", "NULL", "NULL", "TIC", ratio=FALSE, ratioNum=20)
##mSet<-PlotNormSummary(mSet, "norm_0_", "png", 72, width=NA)
##mSet<-PlotSampleNormSummary(mSet, "snorm_0_", "png", 72, width=NA)

```

```{r}

mSet<-GetGroupNames(mSet, "")
feature.nm.vec <- c("")
smpl.nm.vec <- c("")
grp.nm.vec <- c("C","E")
mSet<-UpdateData(mSet)
mSet<-PreparePrenormData(mSet)
mSet<-Normalization(mSet, "CompNorm", "NULL", "NULL", "TIC", ratio=FALSE, ratioNum=20)
##mSet<-PlotNormSummary(mSet, "norm_1_", "png", 72, width=NA)
##mSet<-PlotSampleNormSummary(mSet, "snorm_1_", "png", 72, width=NA)

##FC.Anal takes the mean for the group and divides by the mean from the other group
mSet<-FC.Anal(mSet, 2.0, 0, FALSE)
##mSet<-PlotFC(mSet, "fc_0_", "png", 72, width=NA)

mSet<-Ttests.Anal(mSet, F, .05, F, T, "fdr", F)

##can use t.test(x= cond1 ,y = cond2, paired = F, var.equal = T) to get the t and p values
tempt = function(data){
  p = t.test(data[1:4], data[5:8], paired = F, var.equal = T)
  return(p[[3]])
}
temp = apply(tempdata, 2, tempt)
tempfdr = p.adjust(temp, "BH", length(temp))

```

```{r}

rawdata = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/MetabolomicsForMetaboanalyst.txt", sep = "\t", row.names = 1)

KEGG.Key = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/RawMetabolomicData.csv", row.names = 1)
KEGG.Key = setNames(KEGG.Key[-c(1:2), 2], row.names(KEGG.Key)[-c(1:2)])

genotypes = setNames(c("WT", "GR", "TktOEWT", "TktOEGR", "TktDfWT", "TktDfGR"), c("C", "E", "A", "F", "B", "D"))

metadata = data.frame(Sample = colnames(rawdata), 
                      Group = as.character(rawdata[1,]), 
                      TIC = as.numeric(rawdata[2,]), 
                      Genotypes = genotypes[as.character(rawdata[1,])])

norm.func = function(data){
  nd = (as.numeric(data)/metadata$TIC)*1000
  return(nd)
}

norm.data = t(apply(rawdata[3:nrow(rawdata),], 1, norm.func))
colnames(norm.data) = colnames(rawdata)

group.mean=function(data){
  gm = mean(na.omit(data))
}

mean.data = matrix(0, nrow=nrow(norm.data), ncol = length(genotypes))
row.names(mean.data) = row.names(norm.data)
colnames(mean.data) = unique(metadata$Group)

i=1
while(i<=ncol(mean.data)){
  mean.data[,i] = apply(norm.data[,metadata[metadata$Group == colnames(mean.data)[i], "Sample"]], 1, group.mean)
  i = i+1
}

compare.conditions = function(condition1, condition2){
  columns1 = metadata[metadata$Group == condition1, "Sample"]
  columns2 = metadata[metadata$Group == condition2, "Sample"]
  p=setNames(rep(NA, nrow(norm.data)), row.names(norm.data))
  i=1
  while(i<=length(p)){
    if(length(na.omit(norm.data[i,columns1]))>2 & 
       length(na.omit(norm.data[i,columns2]))>2){
      temp = t.test(na.omit(norm.data[i,columns1]),
                    na.omit(norm.data[i,columns2]), paired = F, var.equal = T)
      p[i] = temp[[3]]
      }
    i=i+1
    }

  fdr = p.adjust(p, "BH", length(p))

  fc = mean.data[,condition1]/mean.data[,condition2]

  comparison.table = data.frame(p = p, 
                                FDR = fdr, 
                                FC = fc, 
                                KEGG = KEGG.Key[row.names(mean.data)])

  comparison.table = comparison.table[order(comparison.table$FDR),]
}

GRxWT.C = compare.conditions("E", "C")
GRxWT.Df = compare.conditions("D", "B")
GRxWT.OE = compare.conditions("F", "A")

KEGGID = na.omit(GRxWT.C[GRxWT.C$FDR<=.1, "KEGG"])
KEGGID.C = KEGGID[-grep(" / ", KEGGID)]

KEGGID = na.omit(GRxWT.Df[GRxWT.Df$FDR<=.1, "KEGG"])
KEGGID.Df = KEGGID[-grep(" / ", KEGGID)]

KEGGID = na.omit(GRxWT.OE[GRxWT.OE$FDR<=.1, "KEGG"])
KEGGID.OE = KEGGID[-grep(" / ", KEGGID)]

```

```{r echo = F}
KEGGID = KEGGID.Df

rm(mSet)
mSet<-InitDataObjects("conc", "pathora", FALSE)
cmpd.vec<-KEGGID
mSet<-Setup.MapData(mSet, cmpd.vec)
mSet<-CrossReferencing(mSet, "kegg")
mSet<-CreateMappingResultTable(mSet)
mSet<-SetKEGG.PathLib(mSet, "dme", "current")
mSet<-SetMetabolomeFilter(mSet, F)
mSet<-CalculateOraScore(mSet, "rbc", "hyperg")

##pathway plot data
ppd = mSet$api$ora.results
colnames(ppd)[4:5] = c("p","neg.log.p")

ppd = as.data.frame(ppd)
ppd$Pathway = row.names(ppd)    
ppd$Pathway = factor(ppd$Pathway, levels= unique(ppd$Pathway))
ppd$Size = 5*log(100*ppd$Impact +1)+10

col.gradient=colorRampPalette(c("white", "yellow","red", "red3"), space = "Lab", bias = 3)

  fig  = plot_ly(data = ppd, 
                 x = ~Impact, 
                 y = ~neg.log.p,
                 name = ~Pathway,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = F,
                 color = ~neg.log.p,
                 colors = col.gradient(nrow(ppd)),
                 marker = list(size = as.numeric(ppd$Size), opacity = 0.75, line = list(color = "black", width = 1)),
                 hoverinfo = "text",
                 hovertext = paste("Pathway:", ppd$Pathway, 
                                   "\nHits: ", ppd$Hits, 
                                   "\nTotal in Cat.: ", ppd$Total, 
                                   "\nFDR: ", ppd$FDR)) %>%
    hide_colorbar()

fig
```

```{r}

```