### 2023-08-07 TKT_RNAseq/R Code/G85RvsWT_Analysis.Rmd

---
title: "G85R vs WT (M + F)"
author: "John Santiago"
date: "2023-07-11"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, include=F}
##Load Libraries

library(biomaRt)
library(org.Dm.eg.db)
library(goseq)
library(GO.db)
library(visNetwork)
library(rrvgo)
library(plotly)
library(KEGGREST)
library(pathview)
library(VennDiagram)
library(heatmaply)
```

```{r, echo = F, include = F}

##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"
cpmdata = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
groups  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
TKT.EdgeR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)
TKT.EdgeR.FC = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FCTable.csv"),row.names = 1)

GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)
kegg = readRDS(gzcon(url(paste0(git.dir,"GeneralDataFiles/kegg.symbol2path.RData"))))
kegg.names = read.csv(paste0(git.dir, "GeneralDataFiles/KEGG.names.csv"), row.names = 1)

genesingo=as.list(org.Dm.egGO2ALLEGS)

##mean.cpm; Mean cpm for each gene in each condition (groups$Group). Used for filtering
mean.cpm = matrix(0, ncol = length(unique(groups$Group)), nrow = nrow(cpmdata))
colnames(mean.cpm) = unique(groups$Group)
row.names(mean.cpm) = row.names(cpmdata)
i=1
while(i<=ncol(mean.cpm)){
  mean.cpm[,i]= apply(cpmdata[,row.names(groups[groups$Group == colnames(mean.cpm)[i],])], 1, mean)
  i=i+1
}

i=1
for(i in 1:ncol(TKT.EdgeR)){
  TKT.EdgeR[TKT.EdgeR[,i] == 0,i] = 2
  i=i+1
}

```

##GO and KEGG Enrichment function

```{r echo = F, include = F}

Enrichment = function(sigs, bg){
  genes = setNames(rep(0, length(bg)), bg)
  genes[sigs] = 1
  pwf=nullp(genes,"dm3","geneSymbol")
  GO.wall=goseq(pwf,"dm3","geneSymbol")
  GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
  row.names(GO.wall) = GO.wall$category
  KEGG=goseq(pwf,gene2cat=kegg)
  KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
  row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
  KEGG$Name=kegg.names[row.names(KEGG),1]

  enrichment.data = list(GO.wall, KEGG)
  names(enrichment.data) = c("GO", "KEGG")
  return(enrichment.data)
}

```

##GO.Sim.PCA function
```{r echo = F, include = F}
GO.Sim.PCA = function(use.data){
    pca <- prcomp(t(use.data), scale.=TRUE) 
    pc1=scale(pca$x[,1])
    pc2=scale(pca$x[,2])
    eigs <- pca$sdev^2
    ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(PC1=pc1,
                        PC2=pc2,
                        sdev=pca$sdev,
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group = reducedTerms[row.names(use.data), "parentTerm"],
                        Score = reducedTerms[row.names(use.data), "score"],
                        Term = reducedTerms[row.names(use.data), "term"],
                        TotalDE = GO.data[row.names(use.data), "numDEInCat"],
                        Total = GO.data[row.names(use.data), "numInCat"],
                        SigVal = signif(GO.data[row.names(use.data), "adjp"],3))

    pca.data$Group = factor(pca.data$Group, levels = unique(pca.data$Group))
    pca.data$Score = 20*(1+pca.data$Score/mean(pca.data$Score))

  fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 ##symbol = ~Group,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c("firebrick","gold","deepskyblue"),
                 marker = list(size = pca.data$Score,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Parent Group: ", pca.data$Group,
                                   "\nSample: ", pca.data$Term, 
                                   "\nTotal DE: ", pca.data$TotalDE, 
                                   "\nTotal in Cat.: ", pca.data$Total, 
                                   "\nFDR: ", pca.data$SigVal)) %>%
    layout(xaxis = list(title = paste0("PC1", "(", ve[1], "%)")),
           yaxis = list(title = paste0("PC2", "(", ve[2], "%)")),
           title = "Enriched GO term similarity")

  return(fig)
}

```

##GO20 Function
```{r echo = F, include =F}

GO20 = function(GO.data, GO.ontology){
GO.data = GO.data[GO.data$ontology == GO.ontology,
                  c("category", "term","adjp", "numDEInCat", "numInCat")]

data = GO.data[20:1,c("term","adjp","numDEInCat", "numInCat", "category")]
data$Score = -log10(data$adjp)
data$term = factor(data$term, levels = c(unique(data$term)))

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

fig <- plot_ly(data, 
               x = ~numDEInCat, y = ~term, 
               type = 'bar',
               name = "term",
               marker = list(colorscale = list(c(0,.5,1), c("brown","gold", "steelblue")),
    reversescale =T,
                             colorbar = list(title = "-log10(FDR)",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                             color = ~Score,
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$term,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = "Top 20 Most Enriched GO Terms",
           margin = m)

  return(fig)
}

```

##KEGG20 Function
```{r, echo = F, include = F}

KEGG20 = function(data){

  data = data[1:20, c("category", "Name","adjp", "numDEInCat", "numInCat")]

data = data[20:1,]
data$Score = -log10(data$adjp)
data$term = factor(data$Name, levels = c(unique(data$Name)))

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

fig <- plot_ly(data, 
               x = ~numDEInCat, y = ~term, 
               type = 'bar',
               name = "Name",
               marker = list(colorscale = list(c(0,.5,1), c("brown","gold", "steelblue")),
    reversescale =T,
                             colorbar = list(title = "-log10(FDR)",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                             color = ~Score,
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$Name,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = "Top 20 Most Enriched KEGG Terms",
           margin = m)

return(fig)

}

```

```{r echo = F, include = F}

##GR vs WT Females

sigs1 = row.names(TKT.EdgeR)[TKT.EdgeR[,"GRxWT.FC"] <= .05]
bg1 = row.names(TKT.EdgeR)[TKT.EdgeR[, "GRxWT.FC"]<= 1]

sigs2 = row.names(TKT.EdgeR)[TKT.EdgeR[,"GRxWT.MC"] <= .05]
bg2 = row.names(TKT.EdgeR)[TKT.EdgeR[, "GRxWT.MC"]<= 1]

sigs = intersect(sigs1, sigs2)
bg = unique(c(bg1, bg2))

sigsKEY = GeneIDKey[sigs,]
sigs = sigsKEY$Symbol
bgKEY = GeneIDKey[bg,]
bg = bgKEY$Symbol

sig.cats = Enrichment(sigs, bg)

```

```{r echo =F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "BP"
go_analysis = GO.data$category[GO.data$adjp <= .05 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$adjp), go_analysis)

simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

reducedTerms <- reduceSimMatrix(simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r echo = F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "BP"
BP.GO.fig = GO20(GO.data, GO.ontology)

GO.ontology = "CC"
CC.GO.fig = GO20(GO.data, GO.ontology)

GO.ontology = "MF"
MF.GO.fig = GO20(GO.data, GO.ontology)

KEGG.fig = KEGG20(sig.cats[["KEGG"]])

```

```{r, fig.width= 10}

BP.GO.fig

```

```{r, fig.width= 10}

CC.GO.fig

```

```{r, fig.width= 10}

MF.GO.fig

```

```{r, fig.width= 10}

KEGG.fig

```

```{r fig.height=7, fig.width=10}

heatmapPlot(simMatrix,
            reducedTerms,
            annotateParent=TRUE,
            annotationLabel="parentTerm",
            fontsize=6,
            main = "Repressor Gene Enriched Biological Process GO terms")

```

```{r warning=F, fig.width=10}

use.data   = simMatrix
GO.Sim.PCA(use.data)

```

```{r fig.align='left'}

treemapPlot(reducedTerms)

```

```{r include = F, echo = F}

GOID = "GO:0005634"
GenesInGO = genesingo[GOID]

GenesInGO = sigsKEY[sigsKEY$ensembl %in% GenesInGO[[1]], "FBgn"]

```

```{r include = F, echo = F}

ParentID = "GO:0044271"
GOIDs = row.names(reducedTerms)[reducedTerms$parent == ParentID]
GOID = GOIDs[1]
GenesInParent= unique(genesingo[[GOID]])
i=2
while(i<=length(GOIDs)){
  GOID = GOIDs[i]
  GenesInParent= unique(GenesInParent, genesingo[[GOID]])
  i=i+1
}

GenesInGO = sigsKEY[sigsKEY$ensembl %in% GenesInParent, "FBgn"]

```

```{r echo =F, warnings = F}

comparison = "GRxWT.FC"
comparison2 = "GRxWT.MC"

hmdata = cpmdata[GenesInGO, groups$Group == "GR.F" | groups$Group == "WT.F"]
hmdata = cpmdata[sigsKEY$FBgn, groups$Group == "GR.F" | groups$Group == "WT.F"]

hover.text = round(hmdata, digits = 1)

FDR = hmdata
FDR[,1:ncol(FDR)] = TKT.EdgeR[row.names(hmdata), comparison]

FBgn = hmdata
FBgn[,1:ncol(FBgn)] = row.names(FBgn)

i=1
while(i<=ncol(hover.text)){
  hover.text[,i] = paste("CPM = ", hover.text[,i],
                         "\nFDR = ", FDR[,i],
                         "\nFBgn = ", FBgn[,i])
  i=i+1
}

hmdata2 = cpmdata[GenesInGO, groups$Group == "GR.M" | groups$Group == "WT.M"]
hmdata2 = cpmdata[sigsKEY$FBgn, groups$Group == "GR.M" | groups$Group == "WT.M"]

hover.text2 = round(hmdata2, digits = 1)

FDR2 = hmdata2
FDR2[,1:ncol(FDR2)] = TKT.EdgeR[row.names(hmdata2), comparison2]

FBgn2 = hmdata2
FBgn2[,1:ncol(FBgn2)] = row.names(FBgn2)

i=1
while(i<=ncol(hover.text2)){
  hover.text2[,i] = paste("CPM = ", hover.text2[,i],
                         "\nFDR = ", FDR2[,i],
                         "\nFBgn = ", FBgn2[,i])
  i=i+1
}

hmdata = cbind(hmdata, hmdata2)
hover.text = cbind(hover.text, hover.text2)

heatmaply(hmdata,
            trace="none",
            col=RdYlBu(100)[100:1],
            scale="row",
            dendrogram = "both",
            show_dendrogram = c(F,T),
            Rowv=T,
            Colv=T,
            ##cexRow = 1/(as.numeric(input$tophits)/25),
            na.color="grey",
            labRow = GeneIDKey[row.names(hmdata), "Symbol"],
            ##cexCol = col.text,
            cexCol = 1,
            key = T,
            ##margins = c(50,350,NA,0),
            ##row_dend_left = T,
            column_text_angle = 90,

            ##col_side_colors=color.groups,
            ##col_side_palette=Spectral,
            ##main= "",
            custom_hovertext = hover.text)

```

##############
## GRxWT Df ##
##############

```{r echo = F, include = F}

comparison1 = "GRxWT.FDf"
comparison2 = "GRxWT.MDf"

sigs1 = row.names(TKT.EdgeR)[TKT.EdgeR[,comparison1] <= .05]

sigs1 = sigs1[apply(cpmdata[sigs1,groups$Group == "TktDfGR.F" | groups$Group == "TktDfWT.F"], 1, min) > 5]

sigs2 = row.names(TKT.EdgeR)[TKT.EdgeR[,comparison2] <= .05]

sigs2 = sigs2[apply(cpmdata[sigs2,groups$Group == "TktDfGR.M" | groups$Group == "TktDfGR.M"], 1, min) > 5]

sigs= intersect(sigs1, sigs2)

sigsKEY = GeneIDKey[sigs,]

sigs = sigsKEY$Symbol

sig.cats = Enrichment(sigs)

```

```{r echo = F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "BP"
BP.GO.fig = GO20(GO.data, GO.ontology)

GO.ontology = "CC"
CC.GO.fig = GO20(GO.data, GO.ontology)

GO.ontology = "MF"
MF.GO.fig = GO20(GO.data, GO.ontology)

KEGG.fig = KEGG20(sig.cats[["KEGG"]])

```

```{r, fig.width= 10}

BP.GO.fig

```

```{r, fig.width= 10}

CC.GO.fig

```

```{r, fig.width= 10}

MF.GO.fig

```

```{r, fig.width= 10}

KEGG.fig

```

```{r echo =F, warnings = F}

comparison = "GRxWT.FDf"
comparison2 = "GRxWT.MDf"

##hmdata = cpmdata[GenesInGO, groups$Group == "GR.F" | groups$Group == "WT.F"]
hmdata = cpmdata[sigsKEY$FBgn, groups$Group == "GR.F" | groups$Group == "WT.F"]

hover.text = round(hmdata, digits = 1)

FDR = hmdata
FDR[,1:ncol(FDR)] = TKT.EdgeR[row.names(hmdata), comparison]

FBgn = hmdata
FBgn[,1:ncol(FBgn)] = row.names(FBgn)

i=1
while(i<=ncol(hover.text)){
  hover.text[,i] = paste("CPM = ", hover.text[,i],
                         "\nFDR = ", FDR[,i],
                         "\nFBgn = ", FBgn[,i])
  i=i+1
}

##hmdata2 = cpmdata[GenesInGO, groups$Group == "GR.M" | groups$Group == "WT.M"]
hmdata2 = cpmdata[sigsKEY$FBgn, groups$Group == "GR.M" | groups$Group == "WT.M"]

hover.text2 = round(hmdata2, digits = 1)

FDR2 = hmdata2
FDR2[,1:ncol(FDR2)] = TKT.EdgeR[row.names(hmdata2), comparison2]

FBgn2 = hmdata2
FBgn2[,1:ncol(FBgn2)] = row.names(FBgn2)

i=1
while(i<=ncol(hover.text2)){
  hover.text2[,i] = paste("CPM = ", hover.text2[,i],
                         "\nFDR = ", FDR2[,i],
                         "\nFBgn = ", FBgn2[,i])
  i=i+1
}

hmdata = cbind(hmdata, hmdata2)
hover.text = cbind(hover.text, hover.text2)

heatmaply(hmdata,
            trace="none",
            col=RdYlBu(100)[100:1],
            scale="row",
            dendrogram = "both",
            show_dendrogram = c(F,T),
            Rowv=T,
            Colv=T,
            ##cexRow = 1/(as.numeric(input$tophits)/25),
            na.color="grey",
            labRow = GeneIDKey[row.names(hmdata), "Symbol"],
            ##cexCol = col.text,
            cexCol = 1,
            key = T,
            ##margins = c(50,350,NA,0),
            ##row_dend_left = T,
            column_text_angle = 90,

            ##col_side_colors=color.groups,
            ##col_side_palette=Spectral,
            ##main= "",
            custom_hovertext = hover.text)

```

##############
## GRxWT OE ##
##############

```{r echo = F, include = F}

comparison1 = "GRxWT.FOE"
comparison2 = "GRxWT.MOE"

sigs1 = row.names(TKT.EdgeR)[TKT.EdgeR[,comparison1] <= .05]

sigs1 = sigs1[apply(cpmdata[sigs1,groups$Group == "TktOEGR.F" | groups$Group == "TktOEWT.F"], 1, min) > 5]

sigs2 = row.names(TKT.EdgeR)[TKT.EdgeR[,comparison2] <= .05]

sigs2 = sigs2[apply(cpmdata[sigs2,groups$Group == "TktOEGR.M" | groups$Group == "TktOEGR.M"], 1, min) > 5]

sigs= intersect(sigs1, sigs2)

sigsKEY = GeneIDKey[sigs,]

sigs = sigsKEY$Symbol

sig.cats = Enrichment(sigs)

```

```{r echo =F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "BP"
go_analysis = GO.data$category[GO.data$adjp <= .05 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$adjp), go_analysis)

simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

reducedTerms <- reduceSimMatrix(simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r echo = F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "BP"
BP.GO.fig = GO20(GO.data, GO.ontology)

GO.ontology = "CC"
CC.GO.fig = GO20(GO.data, GO.ontology)

GO.ontology = "MF"
MF.GO.fig = GO20(GO.data, GO.ontology)

KEGG.fig = KEGG20(sig.cats[["KEGG"]])

```

```{r, fig.width= 10}

BP.GO.fig

```

```{r, fig.width= 10}

CC.GO.fig

```

```{r, fig.width= 10}

MF.GO.fig

```

```{r, fig.width= 10}

KEGG.fig

```

```{r fig.height=7, fig.width=10}

heatmapPlot(simMatrix,
            reducedTerms,
            annotateParent=TRUE,
            annotationLabel="parentTerm",
            fontsize=6,
            main = "Repressor Gene Enriched Biological Process GO terms")

```

```{r warning=F, fig.width=10}

use.data   = simMatrix
GO.Sim.PCA(use.data)

```

```{r fig.align='left'}

treemapPlot(reducedTerms)

```

```{r include = F, echo = F}

GOID = "GO:0003723"
GenesInGO = genesingo[GOID]

GenesInGO = sigsKEY[sigsKEY$ensembl %in% GenesInGO[[1]], "FBgn"]

```

```{r include = F, echo = F}
ParentID = "GO:0034249"
GOIDs = row.names(reducedTerms)[reducedTerms$parent == ParentID]
GOID = GOIDs[1]
GenesInParent= unique(genesingo[[GOID]])
i=2
while(i<=length(GOIDs)){
  GOID = GOIDs[i]
  GenesInParent= unique(GenesInParent, genesingo[[GOID]])
  i=i+1
}

GenesInGO = sigsKEY[sigsKEY$ensembl %in% GenesInParent, "FBgn"]

```

```{r echo =F, warnings = F}

comparison = "GRxWT.FOE"
comparison2 = "GRxWT.MOE"

hmdata = cpmdata[GenesInGO, groups$Group == "TktOEGR.F" | groups$Group == "TktOEWT.F"]
hmdata = cpmdata[sigsKEY$FBgn, groups$Group == "TktOEGR.F" | groups$Group == "TktOEWT.F"]

hover.text = round(hmdata, digits = 1)

FDR = hmdata
FDR[,1:ncol(FDR)] = TKT.EdgeR[row.names(hmdata), comparison]

FBgn = hmdata
FBgn[,1:ncol(FBgn)] = row.names(FBgn)

i=1
while(i<=ncol(hover.text)){
  hover.text[,i] = paste("CPM = ", hover.text[,i],
                         "\nFDR = ", FDR[,i],
                         "\nFBgn = ", FBgn[,i])
  i=i+1
}

hmdata2 = cpmdata[GenesInGO, groups$Group == "TktOEGR.M" | groups$Group == "TktOEWT.M"]
hmdata2 = cpmdata[sigsKEY$FBgn, groups$Group == "TktOEGR.M" | groups$Group == "TktOEWT.M"]

hover.text2 = round(hmdata2, digits = 1)

FDR2 = hmdata2
FDR2[,1:ncol(FDR2)] = TKT.EdgeR[row.names(hmdata2), comparison2]

FBgn2 = hmdata2
FBgn2[,1:ncol(FBgn2)] = row.names(FBgn2)

i=1
while(i<=ncol(hover.text2)){
  hover.text2[,i] = paste("CPM = ", hover.text2[,i],
                         "\nFDR = ", FDR2[,i],
                         "\nFBgn = ", FBgn2[,i])
  i=i+1
}

hmdata = cbind(hmdata, hmdata2)
hover.text = cbind(hover.text, hover.text2)

heatmaply(hmdata,
            trace="none",
            col=RdYlBu(100)[100:1],
            scale="row",
            dendrogram = "both",
            show_dendrogram = c(F,T),
            Rowv=T,
            Colv=T,
            ##cexRow = 1/(as.numeric(input$tophits)/25),
            na.color="grey",
            labRow = GeneIDKey[row.names(hmdata), "Symbol"],
            ##cexCol = col.text,
            cexCol = 1,
            key = T,
            ##margins = c(50,350,NA,0),
            ##row_dend_left = T,
            column_text_angle = 90,

            ##col_side_colors=color.groups,
            ##col_side_palette=Spectral,
            ##main= "",
            custom_hovertext = hover.text)

```

#########################
## Comparing the three ##
#########################

```{r include = F, echo = F}

comparison1 = "GRxWT.FC"
comparison2 = "GRxWT.MC"

sigs1 = row.names(TKT.EdgeR)[TKT.EdgeR[,comparison1] <= .05]

sigs1 = sigs1[apply(cpmdata[sigs1,groups$Group == "GR.F" | groups$Group == "WT.F"], 1, min) > 5]

sigs2 = row.names(TKT.EdgeR)[TKT.EdgeR[,comparison2] <= .05]

sigs2 = sigs2[apply(cpmdata[sigs2,groups$Group == "GR.M" | groups$Group == "GR.M"], 1, min) > 5]

sigs= intersect(sigs1, sigs2)

sigsKEY = GeneIDKey[sigs,]

C.sigs = sigsKEY$Symbol

comparison1 = "GRxWT.FDf"
comparison2 = "GRxWT.MDf"

sigs1 = row.names(TKT.EdgeR)[TKT.EdgeR[,comparison1] <= .05]

sigs1 = sigs1[apply(cpmdata[sigs1,groups$Group == "TktDfGR.F" | groups$Group == "TktDfWT.F"], 1, min) > 5]

sigs2 = row.names(TKT.EdgeR)[TKT.EdgeR[,comparison2] <= .05]

sigs2 = sigs2[apply(cpmdata[sigs2,groups$Group == "TktDfGR.M" | groups$Group == "TktDfGR.M"], 1, min) > 5]

sigs= intersect(sigs1, sigs2)

sigsKEY = GeneIDKey[sigs,]

Df.sigs = sigsKEY$Symbol

comparison1 = "GRxWT.FOE"
comparison2 = "GRxWT.MOE"

sigs1 = row.names(TKT.EdgeR)[TKT.EdgeR[,comparison1] <= .05]

sigs1 = sigs1[apply(cpmdata[sigs1,groups$Group == "TktOEGR.F" | groups$Group == "TktOEWT.F"], 1, min) > 5]

sigs2 = row.names(TKT.EdgeR)[TKT.EdgeR[,comparison2] <= .05]

sigs2 = sigs2[apply(cpmdata[sigs2,groups$Group == "TktOEGR.M" | groups$Group == "TktOEGR.M"], 1, min) > 5]

sigs= intersect(sigs1, sigs2)

sigsKEY = GeneIDKey[sigs,]

OE.sigs = sigsKEY$Symbol

##Sample titles as strings. Only fill in up to your number of selected categories
set1="Control"
set2="Defficient"
set3="Over Expression"
##set4="GSC2.Stem"

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1 = C.sigs
s2 = Df.sigs
s3 = OE.sigs
##s4=row.names(temp)[temp[,4]==1]

```

```{r}

data = data.frame(Condition = c("Control", "Defficient", "Over Expression"), DEGs = c(length(C.sigs), length(Df.sigs), length(OE.sigs)))

fig <- plot_ly(data, 
               x = ~Condition, y = ~DEGs, 
               type = 'bar',
               name = ~Condition,
               color = ~Condition,
               colors = c("coral3","gold2", "steelblue3"),
               marker = list(line = list(color = "black", width = 1.5)))

fig

```

```{r}

##Triple Venn
grid.newpage()
draw.triple.venn(area1=length(s1),area2=length(s2),area3=length(s3),n12=length(intersect(s1,s2)),n23 = length(intersect(s2,s3)),n13 = length(intersect(s1,s3)),n123 = length(intersect(intersect(s1,s2),s3)),category=c(set1,set2,set3),fill = c("brown1","gold2", "steelblue2"),cex=1.5,cat.cex = 1)

```