### 2024-01-17 Anna qPCR data/20231220_Anna_GRWT_IDGF6.PEPCK.RP49.Rmd

---
title: "Anna IDGF6 and PEPCK Primer qPCR Plate"
author: "John Santiago"
date: "2023-12-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, include = F}
library(plotly)
IDGF6.data = read.csv("/Users/johncsantiago/Documents/20231220_Anna_GRWT_IDGF6.csv")
IDGF6.data$Genotype = substr(IDGF6.data$Sample,1,2)
IDGF6.data$Concentration = substr(IDGF6.data$Sample,5,nchar(IDGF6.data$Sample))
IDGF6.data$Rep = substr(IDGF6.data$Sample,3,3)
IDGF6.data$Primer = 'IDGF6'

IDGF6.data = IDGF6.data[IDGF6.data$Well != "A7", c('Well', 'Sample', 'Cq', 'Genotype', 'Primer', 'Rep', 'Concentration')]

RP49.PEPCK.data = read.csv("/Users/johncsantiago/Documents/20231220_Anna_GRWT_RP49.PEPCK.csv")
RP49.PEPCK.data$Genotype = substr(RP49.PEPCK.data$Sample,1,2)
RP49.PEPCK.data$Concentration = substr(RP49.PEPCK.data$Sample,6,(nchar(RP49.PEPCK.data$Sample)-3))
RP49.PEPCK.data$Rep = substr(RP49.PEPCK.data$Sample,3,3)
RP49.PEPCK.data$Primer = RP49.PEPCK.data$Target

RP49.PEPCK.data = RP49.PEPCK.data[RP49.PEPCK.data$Well != "A6" &
                                    RP49.PEPCK.data$Well != "A7" &
                                    RP49.PEPCK.data$Well != "G3" &
                                    RP49.PEPCK.data$Well != "G4" &
                                    RP49.PEPCK.data$Well != "G5" &
                                    RP49.PEPCK.data$Well != "G6" &
                                    RP49.PEPCK.data$Well != "C9", 
                                  c('Well', 'Sample', 'Cq', 'Genotype', 'Primer', 'Rep', 'Concentration')]

RP49.PEPCK.data$Cq = as.numeric(RP49.PEPCK.data$Cq)

LDH.PDH.PCB.data = read.csv("/Users/johncsantiago/Documents/20231221_Anna_GRWT_PCB.PDH.LDH.csv")

LDH.PDH.PCB.data[grep('idgf', LDH.PDH.PCB.data$Sample), 'Sample'] = paste0('KD', substr(LDH.PDH.PCB.data[grep('idgf', LDH.PDH.PCB.data$Sample), 'Sample'] , 12, 18))

LDH.PDH.PCB.data[grep('GR13', LDH.PDH.PCB.data$Sample), 'Sample'] = paste0('GR3', substr(LDH.PDH.PCB.data[grep('GR13', LDH.PDH.PCB.data$Sample), 'Sample'],5, 10))

LDH.PDH.PCB.data$Genotype = substr(LDH.PDH.PCB.data$Sample,1,2)
LDH.PDH.PCB.data$Concentration = substr(LDH.PDH.PCB.data$Sample,6,(nchar(LDH.PDH.PCB.data$Sample)-3))
LDH.PDH.PCB.data$Rep = substr(LDH.PDH.PCB.data$Sample,3,3)
LDH.PDH.PCB.data$Primer = LDH.PDH.PCB.data$Target

LDH.PDH.PCB.data = LDH.PDH.PCB.data[LDH.PDH.PCB.data$Well != "D6" &
                                    LDH.PDH.PCB.data$Well != "D1" &
                                    LDH.PDH.PCB.data$Well != "D10" &
                                    LDH.PDH.PCB.data$Well != "D12" &
                                    LDH.PDH.PCB.data$Well != "E4", 
                                  c('Well', 'Sample', 'Cq', 'Genotype', 'Primer', 'Rep', 'Concentration')]

LDH.PDH.PCB.data$Cq = as.numeric(LDH.PDH.PCB.data$Cq)

```

```{r echo = F}

plotdata = IDGF6.data
plotdata = plotdata[order(paste0(plotdata$Genotype, plotdata$Rep)),]

plotdata$x = factor(plotdata$Sample, 
                    levels = c(unique(plotdata$Sample)))

plotdata$Cq = as.numeric(plotdata$Cq)

boxplot(plotdata$Cq~plotdata$x,
     ylab = "Quantification Cycle (Cq)",
     xlab = "",
     main = "IDGF6",
     las = 2)

points(x = plotdata$x,
       y = plotdata$Cq,
       cex = 1.5,
       pch = 21,
       bg = "steelblue",
       col = "black")

```

```{r echo = F}

mean.cq = c(1:18)
names(mean.cq) = unique(IDGF6.data$Sample)
i = 1
while(i <= length(mean.cq)){
  mean.cq[i] = mean(IDGF6.data[IDGF6.data$Sample == names(mean.cq)[i], 'Cq'])
  i = i+1
}

idgf.mean.cq = mean.cq[-grep(' 2.5', names(mean.cq))]
names(idgf.mean.cq) = substr(names(idgf.mean.cq), 1,3)

dCq = c(1:9)
names(dCq) = unique(paste0(IDGF6.data$Genotype, IDGF6.data$Rep))
i = 1
while(i <= length(dCq)){
  temp = mean.cq[substr(names(mean.cq),1,3) == names(dCq)[i]]
  dCq[i] = temp[2] - temp[1]
  i=i+1
}

FC = 2^dCq

barplot(names = names(FC),
     height = FC,
     ylab = "Fold Change",
     xlab = "",
     main = "IDGF6 FC between 2.5ng and 5ng cDNA",
     las = 2,
     col = "steelblue",
     ylim = c(0,2))

```

```{r echo = F}

plotdata = RP49.PEPCK.data[RP49.PEPCK.data$Primer == 'RP49',]
plotdata = plotdata[order(paste0(plotdata$Genotype, plotdata$Rep)),]

plotdata$x = factor(plotdata$Sample, 
                    levels = c(unique(plotdata$Sample)))

plotdata$Cq = as.numeric(plotdata$Cq)

boxplot(plotdata$Cq~plotdata$x,
     ylab = "Quantification Cycle (Cq)",
     xlab = "",
     main = "RP49",
     las = 2)

points(x = plotdata$x,
       y = plotdata$Cq,
       cex = 1.5,
       pch = 21,
       bg = "firebrick",
       col = "black")

```

```{r echo = F}

plotdata = RP49.PEPCK.data[RP49.PEPCK.data$Primer == 'Pepck-1',]
plotdata = plotdata[order(paste0(plotdata$Genotype, plotdata$Rep)),]

plotdata$x = factor(plotdata$Sample, 
                    levels = c(unique(plotdata$Sample)))

plotdata$Cq = as.numeric(plotdata$Cq)

boxplot(plotdata$Cq~plotdata$x,
     ylab = "Quantification Cycle (Cq)",
     xlab = "",
     main = "PEPCK",
     las = 2)

points(x = plotdata$x,
       y = plotdata$Cq,
       cex = 1.5,
       pch = 21,
       bg = "darkgreen",
       col = "black")

```

```{r echo = F}

plotdata = LDH.PDH.PCB.data[LDH.PDH.PCB.data$Primer == 'PCB',]
plotdata = plotdata[order(paste0(plotdata$Genotype, plotdata$Rep)),]

plotdata$x = factor(plotdata$Sample, 
                    levels = c(unique(plotdata$Sample)))

plotdata$Cq = as.numeric(plotdata$Cq)

boxplot(plotdata$Cq~plotdata$x,
     ylab = "Quantification Cycle (Cq)",
     xlab = "",
     main = "PCB",
     las = 2)

points(x = plotdata$x,
       y = plotdata$Cq,
       cex = 1.5,
       pch = 21,
       bg = "gold",
       col = "black")

```

```{r echo = F}

plotdata = LDH.PDH.PCB.data[LDH.PDH.PCB.data$Primer == 'PDH',]
plotdata = plotdata[order(paste0(plotdata$Genotype, plotdata$Rep)),]

plotdata$x = factor(plotdata$Sample, 
                    levels = c(unique(plotdata$Sample)))

plotdata$Cq = as.numeric(plotdata$Cq)

boxplot(plotdata$Cq~plotdata$x,
     ylab = "Quantification Cycle (Cq)",
     xlab = "",
     main = "PDH",
     las = 2)

points(x = plotdata$x,
       y = plotdata$Cq,
       cex = 1.5,
       pch = 21,
       bg = "orchid4",
       col = "black")

```

```{r echo = F}

plotdata = LDH.PDH.PCB.data[LDH.PDH.PCB.data$Primer == 'LDH',]
plotdata = plotdata[order(paste0(plotdata$Genotype, plotdata$Rep)),]

plotdata$x = factor(plotdata$Sample, 
                    levels = c(unique(plotdata$Sample)))

plotdata$Cq = as.numeric(plotdata$Cq)

boxplot(plotdata$Cq~plotdata$x,
     ylab = "Quantification Cycle (Cq)",
     xlab = "",
     main = "LDH",
     las = 2)

points(x = plotdata$x,
       y = plotdata$Cq,
       cex = 1.5,
       pch = 21,
       bg = "darkorange",
       col = "black")

```

```{r echo = F}

rp49.data = RP49.PEPCK.data[RP49.PEPCK.data$Primer == 'RP49',]

mean.cq = c(1:9)
names(mean.cq) = unique(rp49.data$Sample)
i = 1
while(i <= length(mean.cq)){
  mean.cq[i] = mean(rp49.data[rp49.data$Sample == names(mean.cq)[i], 'Cq'])
  i = i+1
}

rp49.mean.cq = mean.cq
names(rp49.mean.cq) = substr(names(rp49.mean.cq),1,3)

pepck.data = RP49.PEPCK.data[RP49.PEPCK.data$Primer == 'Pepck-1',]

mean.cq = c(1:10)
names(mean.cq) = unique(pepck.data$Sample)
i = 1
while(i <= length(mean.cq)){
  mean.cq[i] = mean(pepck.data[pepck.data$Sample == names(mean.cq)[i], 'Cq'])
  i = i+1
}

pepck.mean.cq = mean.cq
names(pepck.mean.cq) = substr(names(pepck.mean.cq),1,3)
pepck.FC = 2^(pepck.mean.cq[10] - pepck.mean.cq[9])
pepck.mean.cq = pepck.mean.cq[1:8]
pepck.mean.cq = pepck.mean.cq[names(rp49.mean.cq)]
names(pepck.mean.cq) = names(rp49.mean.cq)

ldh.data = LDH.PDH.PCB.data[LDH.PDH.PCB.data$Primer == 'LDH',]

mean.cq = c(1:10)
names(mean.cq) = unique(ldh.data$Sample)
i = 1
while(i <= length(mean.cq)){
  mean.cq[i] = mean(ldh.data[ldh.data$Sample == names(mean.cq)[i], 'Cq'])
  i = i+1
}

ldh.mean.cq = mean.cq
ldh.FC = 2^(ldh.mean.cq['Wt1 (2.5ng)'] - ldh.mean.cq['WT1 (5ng)'])
ldh.mean.cq = ldh.mean.cq[names(ldh.mean.cq) != 'Wt1 (2.5ng)']
names(ldh.mean.cq) = substr(names(ldh.mean.cq),1,3)
ldh.mean.cq = ldh.mean.cq[names(rp49.mean.cq)]
names(ldh.mean.cq) = names(rp49.mean.cq)

pdh.data = LDH.PDH.PCB.data[LDH.PDH.PCB.data$Primer == 'PDH',]

mean.cq = c(1:10)
names(mean.cq) = unique(pdh.data$Sample)
i = 1
while(i <= length(mean.cq)){
  mean.cq[i] = mean(pdh.data[pdh.data$Sample == names(mean.cq)[i], 'Cq'])
  i = i+1
}

pdh.mean.cq = mean.cq
pdh.FC = 2^(pdh.mean.cq['Wt1 (2.5ng)'] - pdh.mean.cq['WT1 (5ng)'])
pdh.mean.cq = pdh.mean.cq[names(pdh.mean.cq) != 'Wt1 (2.5ng)']
names(pdh.mean.cq) = substr(names(pdh.mean.cq),1,3)
pdh.mean.cq = pdh.mean.cq[names(rp49.mean.cq)]
names(pdh.mean.cq) = names(rp49.mean.cq)

pcb.data = LDH.PDH.PCB.data[LDH.PDH.PCB.data$Primer == 'PCB',]

mean.cq = c(1:10)
names(mean.cq) = unique(pcb.data$Sample)
i = 1
while(i <= length(mean.cq)){
  mean.cq[i] = mean(pcb.data[pcb.data$Sample == names(mean.cq)[i], 'Cq'])
  i = i+1
}

pcb.mean.cq = mean.cq
pcb.FC = 2^(pcb.mean.cq['Wt1 (2.5ng)'] - pcb.mean.cq['WT1 (5ng)'])
pcb.mean.cq = pcb.mean.cq[names(pcb.mean.cq) != 'Wt1 (2.5ng)']
names(pcb.mean.cq) = substr(names(pcb.mean.cq),1,3)
pcb.mean.cq = pcb.mean.cq[names(rp49.mean.cq)]
names(pcb.mean.cq) = names(rp49.mean.cq)

```

```{r echo = F}

barplotdata = c(pepck.FC, pcb.FC, pdh.FC, ldh.FC)
names(barplotdata) = c ('PEPCK', 'PCB', 'PDH', 'LDH')

barplot(height = barplotdata,
     ylab = "Fold Change",
     xlab = "",
     main = "Pepck-1 FC between 2.5ng and 10ng cDNA",
     las = 1,
     col = c("darkgreen", 'gold', 'orchid4', 'darkorange'),
     ylim = c(0,4.1))
abline(h = 4, lty = 2, lwd = 2)
abline(h = 2, lty = 2, lwd = 2)

```

```{r echo = F}

plotdata = rbind(IDGF6.data, RP49.PEPCK.data, LDH.PDH.PCB.data)
plotdata = plotdata[plotdata$Genotype == "GR" & plotdata$Concentration == 5,]

plotdata$x = factor(paste0(plotdata$Primer, '-', plotdata$Rep), 
                    levels = c(unique(paste0(plotdata$Primer, '-', plotdata$Rep))[order(unique(paste0(plotdata$Primer, '-', plotdata$Rep)))]))

plotdata$Cq = as.numeric(plotdata$Cq)

boxplot(plotdata$Cq~plotdata$x,
     ylab = "Quantification Cycle (Cq)",
     xlab = "",
     main = "G85R",
     las = 2)

points(x = plotdata$x,
       y = plotdata$Cq,
       cex = 1.5,
       pch = 21,
       bg = c(rep('steelblue',8), 
              rep('firebrick',7),
              rep("darkgreen",9),
              rep("gold", 9),
              rep('orchid4', 9),
              rep('darkorange', 9)),
       col = "black")

```

```{r echo = F}

plotdata = rbind(IDGF6.data, RP49.PEPCK.data, LDH.PDH.PCB.data)
plotdata = plotdata[plotdata$Genotype == "WT" & plotdata$Concentration == 5,]

plotdata$x = factor(paste0(plotdata$Primer, '-', plotdata$Rep), 
                    levels = c(unique(paste0(plotdata$Primer, '-', plotdata$Rep))[order(unique(paste0(plotdata$Primer, '-', plotdata$Rep)))]))

plotdata$Cq = as.numeric(plotdata$Cq)

boxplot(plotdata$Cq~plotdata$x,
     ylab = "Quantification Cycle (Cq)",
     xlab = "",
     main = "WT",
     las = 2)

points(x = plotdata$x,
       y = plotdata$Cq,
       cex = 1.5,
       pch = 21,
       bg = c(rep('steelblue',9), 
              rep('firebrick',9),
              rep("darkgreen",6),
              rep("gold", 7),
              rep('orchid4', 9),
              rep('darkorange', 9)),
       col = "black")
```

```{r echo = F}

plotdata = rbind(IDGF6.data, RP49.PEPCK.data, LDH.PDH.PCB.data)
plotdata = plotdata[plotdata$Genotype == "KD" & plotdata$Concentration == 5,]

plotdata$x = factor(paste0(plotdata$Primer, '-', plotdata$Rep), 
                    levels = c(unique(paste0(plotdata$Primer, '-', plotdata$Rep))[order(unique(paste0(plotdata$Primer, '-', plotdata$Rep)))]))

plotdata$Cq = as.numeric(plotdata$Cq)

boxplot(plotdata$Cq~plotdata$x,
     ylab = "Quantification Cycle (Cq)",
     xlab = "",
     main = "KD",
     las = 2)

points(x = plotdata$x,
       y = plotdata$Cq,
       cex = 1.5,
       pch = 21,
       bg = c(rep('steelblue',9), 
              rep('firebrick',8),
              rep("darkgreen",9),
              rep("gold", 9),
              rep('orchid4', 9),
              rep('darkorange', 9)),
       col = "black")
```

```{r echo = F}

dCt = data.frame(Genotype = substr(names(idgf.mean.cq),1,2),
                 IDGF6 = idgf.mean.cq,
                 RP49 = rp49.mean.cq,
                 PEPCK = pepck.mean.cq,
                 PDH = pdh.mean.cq,
                 PCB = pcb.mean.cq,
                 LDH = ldh.mean.cq)
dCt$Genotype = factor(dCt$Genotype, levels = c('WT', 'GR', 'KD'))

dCt$IDGF6.dCt = dCt$IDGF6 - dCt$RP49
dCt$IDGF6.FC = 2^-dCt$IDGF6.dCt

dCt$PEPCK.dCt = dCt$PEPCK - dCt$RP49
dCt$PEPCK.FC = 2^-dCt$PEPCK.dCt

dCt$PDH.dCt = dCt$PDH - dCt$RP49
dCt$PDH.FC = 2^-dCt$PDH.dCt

dCt$PCB.dCt = dCt$PCB - dCt$RP49
dCt$PCB.FC = 2^-dCt$PCB.dCt

dCt$LDH.dCt = dCt$LDH - dCt$RP49
dCt$LDH.FC = 2^-dCt$LDH.dCt

GRvsWT = t.test(dCt$IDGF6.dCt[dCt$Genotype == "WT"], dCt$IDGF6.dCt[dCt$Genotype == "GR"])[[3]]
KDvsWT = t.test(dCt$IDGF6.dCt[dCt$Genotype == "WT"], dCt$IDGF6.dCt[dCt$Genotype == "KD"])[[3]]

boxplot(IDGF6.dCt~Genotype, data = dCt, ylim = c(2,7))

points(x = dCt$Genotype,
       y = dCt$IDGF6.dCt,
       cex = 1.5,
       pch = 21,
       bg = c(rep('steelblue',3), 
              rep('firebrick',3),
              rep("darkgreen",3)),
       col = "black")

text(x=c(2,3),y=c(6.75,6.75), labels = c(paste0('pval = ', signif(GRvsWT,3)), 
                                   paste0('pval = ', signif(KDvsWT,3))))

```

```{r echo = F}

GRvsWT = t.test(dCt$PEPCK.dCt[dCt$Genotype == "WT"], dCt$PEPCK.dCt[dCt$Genotype == "GR"])[[3]]
KDvsWT = t.test(dCt$PEPCK.dCt[dCt$Genotype == "WT"], dCt$PEPCK.dCt[dCt$Genotype == "KD"])[[3]]

boxplot(PEPCK.dCt~Genotype, data = dCt, ylim = c(4, 7))

points(x = dCt$Genotype,
       y = dCt$PEPCK.dCt,
       cex = 1.5,
       pch = 21,
       bg = c(rep('steelblue',3), 
              rep('firebrick',3),
              rep("darkgreen",3)),
       col = "black")

text(x=c(2,3),y=c(6.75,6.75), labels = c(paste0('pval = ', signif(GRvsWT,3)), 
                                   paste0('pval = ', signif(KDvsWT,3))))

```

```{r echo = F}

GRvsWT = t.test(dCt$PDH.dCt[dCt$Genotype == "WT"], dCt$PDH.dCt[dCt$Genotype == "GR"])[[3]]
KDvsWT = t.test(dCt$PDH.dCt[dCt$Genotype == "WT"], dCt$PDH.dCt[dCt$Genotype == "KD"])[[3]]

KDvsGR = t.test(dCt$PDH.dCt[dCt$Genotype == "GR"], dCt$PDH.dCt[dCt$Genotype == "KD"])[[3]]

boxplot(PDH.dCt~Genotype, data = dCt)

points(x = dCt$Genotype,
       y = dCt$PDH.dCt,
       cex = 1.5,
       pch = 21,
       bg = c(rep('steelblue',3), 
              rep('firebrick',3),
              rep("darkgreen",3)),
       col = "black")

text(x=c(2,3,2.5),y=c(11,11, 14), labels = c(paste0('pval = ', signif(GRvsWT,3)), 
                                   paste0('pval = ', signif(KDvsWT,3)),
                                   paste0('KD vs GR\npval = ', signif(KDvsGR,3))))

```

```{r echo = F}

GRvsWT = t.test(dCt$PCB.dCt[dCt$Genotype == "WT"], dCt$PCB.dCt[dCt$Genotype == "GR"])[[3]]
KDvsWT = t.test(dCt$PCB.dCt[dCt$Genotype == "WT"], dCt$PCB.dCt[dCt$Genotype == "KD"])[[3]]

KDvsGR = t.test(dCt$PCB.dCt[dCt$Genotype == "GR"], dCt$PCB.dCt[dCt$Genotype == "KD"])[[3]]

boxplot(PCB.dCt~Genotype, data = dCt)

points(x = dCt$Genotype,
       y = dCt$PCB.dCt,
       cex = 1.5,
       pch = 21,
       bg = c(rep('steelblue',3), 
              rep('firebrick',3),
              rep("darkgreen",3)),
       col = "black")

text(x=c(2,3, 2.5),y=c(9, 9, 14), labels = c(paste0('pval = ', signif(GRvsWT,3)), 
                                   paste0('pval = ', signif(KDvsWT,3)),
                                   paste0('KD vs GR\npval = ', signif(KDvsGR,3))))

```

```{r echo = F}

GRvsWT = t.test(dCt$LDH.dCt[dCt$Genotype == "WT"], dCt$LDH.dCt[dCt$Genotype == "GR"])[[3]]
KDvsWT = t.test(dCt$LDH.dCt[dCt$Genotype == "WT"], dCt$LDH.dCt[dCt$Genotype == "KD"])[[3]]

KDvsGR = t.test(dCt$LDH.dCt[dCt$Genotype == "GR"], dCt$LDH.dCt[dCt$Genotype == "KD"])[[3]]

boxplot(LDH.dCt~Genotype, data = dCt)

points(x = dCt$Genotype,
       y = dCt$LDH.dCt,
       cex = 1.5,
       pch = 21,
       bg = c(rep('steelblue',3), 
              rep('firebrick',3),
              rep("darkgreen",3)),
       col = "black")

text(x=c(2,3, 2.5),y=c(8, 8, 11), labels = c(paste0('pval = ', signif(GRvsWT,3)), 
                                   paste0('pval = ', signif(KDvsWT,3)),
                                   paste0('KD vs GR\npval = ', signif(KDvsGR,3))))

```

```{r echo = F}

KDvsGR = t.test(dCt$PDH.dCt[dCt$Genotype == "GR"], dCt$PDH.dCt[dCt$Genotype == "KD"])[[3]]

boxplotdata = dCt[row.names(dCt) != 'WT1' & row.names(dCt) != 'WT2',]

boxplot(PDH.dCt~Genotype, data = boxplotdata, ylim = c(7.5,11))

points(x = dCt$Genotype,
       y = dCt$PDH.dCt,
       cex = 1.5,
       pch = 21,
       bg = c(rep('steelblue',3), 
              rep('firebrick',3),
              rep("darkgreen",3)),
       col = "black")
text(x=c(2.5),y=c(10.5), labels = c(paste0('KD vs GR\npval = ', signif(KDvsGR,3))))

```

```{r echo = F}

KDvsGR = t.test(dCt$PCB.dCt[dCt$Genotype == "GR"], dCt$PCB.dCt[dCt$Genotype == "KD"])[[3]]

GRvsWT = t.test(dCt$PCB.dCt[dCt$Genotype == "GR"], dCt$PCB.dCt[dCt$Genotype == "WT"])[[3]]

boxplotdata = dCt[row.names(dCt) != 'WT1' & row.names(dCt) != 'WT2',]

boxplot(PCB.dCt~Genotype, data = boxplotdata, ylim = c(6,9))

points(x = dCt$Genotype,
       y = dCt$PCB.dCt,
       cex = 1.5,
       pch = 21,
       bg = c(rep('steelblue',3), 
              rep('firebrick',3),
              rep("darkgreen",3)),
       col = "black")
text(x=c(2.5),y=c(8.5), labels = c(paste0('KD vs GR\npval = ', signif(KDvsGR,3))))

```

```{r echo = F}

KDvsGR = t.test(dCt$LDH.dCt[dCt$Genotype == "GR"], dCt$LDH.dCt[dCt$Genotype == "KD"])[[3]]

boxplotdata = dCt[row.names(dCt) != 'WT1' & row.names(dCt) != 'WT2',]

boxplot(LDH.dCt~Genotype, data = boxplotdata, ylim = c(4.5,8))

points(x = dCt$Genotype,
       y = dCt$LDH.dCt,
       cex = 1.5,
       pch = 21,
       bg = c(rep('steelblue',3), 
              rep('firebrick',3),
              rep("darkgreen",3)),
       col = "black")
text(x=c(2.5),y=c(7.5), labels = c(paste0('KD vs GR\npval = ', signif(KDvsGR,3))))

```

```{r echo = F}

mean.dCt = data.frame(Genotype = c('WT','GR', 'KD'),
                     IDGF6.dCT = c(mean(dCt$IDGF6.dCt[dCt$Genotype == 'WT']),
                                   mean(dCt$IDGF6.dCt[dCt$Genotype == 'GR']),
                                   mean(dCt$IDGF6.dCt[dCt$Genotype == 'KD'])),
                     PEPCK.dCT = c(mean(na.omit(dCt$PEPCK.dCt[dCt$Genotype == 'WT'])),
                                   mean(dCt$PEPCK.dCt[dCt$Genotype == 'GR']),
                                   mean(dCt$PEPCK.dCt[dCt$Genotype == 'KD'])),
                     PDH.dCT = c(dCt['WT3', 'PDH.dCt'],
                                   mean(dCt$PDH.dCt[dCt$Genotype == 'GR']),
                                   mean(dCt$PDH.dCt[dCt$Genotype == 'KD'])),
                     PCB.dCT = c(dCt['WT3', 'PCB.dCt'],
                                   mean(dCt$PCB.dCt[dCt$Genotype == 'GR']),
                                   mean(dCt$PCB.dCt[dCt$Genotype == 'KD'])),
                     LDH.dCT = c(dCt['WT3', 'LDH.dCt'],
                                   mean(dCt$LDH.dCt[dCt$Genotype == 'GR']),
                                   mean(na.omit(dCt$LDH.dCt[dCt$Genotype == 'KD']))))

IDGF6.ddCT = c((mean.dCt$IDGF6.dCT[2] - mean.dCt$IDGF6.dCT[1]),
               (mean.dCt$IDGF6.dCT[3] - mean.dCt$IDGF6.dCT[1]))
names(IDGF6.ddCT) = c('GR/WT', 'KD/WT')

PEPCK.ddCT = c((mean.dCt$PEPCK.dCT[2] - mean.dCt$PEPCK.dCT[1]),
               (mean.dCt$PEPCK.dCT[3] - mean.dCt$PEPCK.dCT[1]))
names(PEPCK.ddCT) = c('GR/WT', 'KD/WT')

PDH.ddCT = c((mean.dCt$PDH.dCT[2] - mean.dCt$PDH.dCT[1]),
               (mean.dCt$PDH.dCT[3] - mean.dCt$PDH.dCT[1]))
names(PDH.ddCT) = c('GR/WT', 'KD/WT')

PCB.ddCT = c((mean.dCt$PCB.dCT[2] - mean.dCt$PCB.dCT[1]),
               (mean.dCt$PCB.dCT[3] - mean.dCt$PCB.dCT[1]))
names(PCB.ddCT) = c('GR/WT', 'KD/WT')

LDH.ddCT = c((mean.dCt$LDH.dCT[2] - mean.dCt$LDH.dCT[1]),
               (mean.dCt$LDH.dCT[3] - mean.dCt$LDH.dCT[1]))
names(PDH.ddCT) = c('GR/WT', 'KD/WT')

barplot(height = c(2^-IDGF6.ddCT, NA, 
                   2^-PEPCK.ddCT, NA, 
                   2^-PDH.ddCT, NA, 
                   2^-PCB.ddCT, NA, 
                   2^-LDH.ddCT),
        names.arg = c('GR IDGF6', 'KD IDGF6', NA, 
                      'GR PEPCK', 'KD PEPCK', NA,
                      'GR PDH', 'KD PDH', NA,
                      'GR PCB', 'KD PCB', NA,
                      'GR LDH', 'KD LDH'),
        ylab = 'Fold Change (relative to WT)',
        ylim = c(0,2.5),
        las = 2,
        cex.names = .9)
abline(h = 1, lty = 2)

```