### 2024-01-17 GeneExpressionBoxPlots.Rmd

---
title: "Box Plots Using All Data"
author: "John Santiago"
date: "2023-11-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = F, echo = F}

library(plotly)
library(edgeR)

##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"
cpmdata = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
groups  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
TKT.EdgeR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)
TKT.EdgeR.FC = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FCTable.csv"),row.names = 1)

groups = groups[colnames(cpmdata),]

gbbOE = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbOE_CountTable.csv", row.names = 1)
gbbOE.groups = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbOE_Metadata.csv", row.names = 1)
gbbOE = gbbOE[,row.names(gbbOE.groups)]
gbbOE.groups$group = paste0(gbbOE.groups$genotype, "_",gbbOE.groups$stage)

##normalize data
countdata=gbbOE
x <- countdata
gr <- factor(gbbOE.groups$group, levels=c(unique(gbbOE.groups$group)))
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
gbbOE.cpmdata=cpm(z)

gbbOEmean = gbbOE.cpmdata[,1:6]
colnames(gbbOEmean) = unique(gr)
i=1
while(i<=6){
  temp= gbbOE.cpmdata[,row.names(gbbOE.groups[gbbOE.groups$group == colnames(gbbOEmean)[i],])]
  gbbOEmean[,i] = apply(temp, 1, mean)
  i=i+1
}

design<-model.matrix(~0 + gr)
colnames(design) = unique(gr)
z = estimateDisp(z, design, robust = T)

fit <- glmQLFit(z, design)
plotQLDisp(fit)

compare = makeContrasts(G85R_LL3-LoxP_LL3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.LL3=DEout
sGRxWT.LL3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_ML3-LoxP_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.ML3=DEout
sGRxWT.ML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_LL3-G85R_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.LL3xML3=DEout
sGR.LL3xML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(LoxP_LL3-LoxP_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.LL3xML3=DEout
sWT.LL3xML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_gbb_LL3-G85R_gbb_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
G85R_gbb.LL3xML3=DEout
sG85R_gbb.LL3xML3=DEout[DEout$FDR<.05,]

gbbKO = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbKO_CountTable.csv", row.names = 1)
gbbKO.groups = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbKO_Metadata.csv", row.names = 1)

##normalize data
countdata=gbbKO
x <- countdata
gr <- factor(gbbKO.groups$Genotype, levels=c(unique(gbbKO.groups$Genotype)))
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
gbbKO.cpmdata=cpm(z)

GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)

```

```{r}

##relish FBgn0014018
##attacin FBgn0041579
##cecropin FBgn0000276
##diptericin FBgn0004240
##drosocin FBgn0010388
##defensin FBgn0010385
##drosomycin FBgn0283461
##metchnikowin FBgn0014865
##sod1 FBgn0003462
##sod2 FBgn0010213
##catalase FBgn0000261
##gstd1 FBgn0001149

##cgs = c('CG10592','CG10827','CG1462','CG16771','CG1809','CG3264','CG3290','CG3292','CG5150','CG5361','CG5656','CG8105','CG8147')

FBgn = 'FBgn0004362'
name = ''

gene = data.frame(cpm = as.numeric(cpmdata[FBgn,]), condition = groups$Group, group = paste0(groups$Genotype, groups$Sex))

gene$group = factor(gene$group, levels = c(unique(gene$group)))
gene$condition = factor(gene$condition, levels = c('GR.F', "TktDfGR.F", "TktOEGR.F",
                                                 'WT.F', "TktDfWT.F", "TktOEWT.F",
                                                 'GR.M', "TktDfGR.M", "TktOEGR.M",
                                                 'WT.M', "TktDfWT.M", "TktOEWT.M"))

boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"gold","darkgreen","dodgerblue")[as.numeric(factor(gene$group))])

```

```{r}

gene = data.frame(cpm = as.numeric(gbbOE.cpmdata[FBgn,]), condition = gbbOE.groups$group, group = gbbOE.groups$genotype)

gene$group = factor(gene$group, levels = c(unique(gene$group)))
gene$condition = factor(gene$condition, levels = c('LoxP_ML3', "LoxP_LL3",
                                                 'G85R_ML3', "G85R_LL3",
                                                 'G85R_gbb_ML3', "G85R_gbb_LL3"))

boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"dodgerblue","gold")[as.numeric(factor(gene$group))])

```

```{r}

## WT = W1118
## KO = GBB1/GBB2 (both null alleles)
## Aaron and Catherine legend for RNAseq
## Whole larvae

gene = data.frame(cpm = as.numeric(gbbKO.cpmdata[FBgn,]), condition = gbbKO.groups$Genotype)

gene$condition = factor(gene$condition, levels = c('WT', "KO"))

boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"dodgerblue","gold")[as.numeric(factor(gene$condition))])

```