### 2024-01-17 TKT_RNAseq/R Code/GlycolysisNetwork.Rmd

---
title: "Glycolysis Network"
author: "John Santiago"
date: "2023-12-06"
output: 
  bookdown::pdf_document2:
    number_sections: false
---

```{r setup, include=FALSE}
##knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(dev='svg', fig.width=9, fig.height=12, out.width = "9in", out.height = "12in")
```

- #### Figure \@ref(fig:fig1): GR.FC vs WT.FC

- #### Figure \@ref(fig:fig2): GR.MC vs WT.MC

- #### Figure \@ref(fig:fig3): GR.FDf vs WT.FDf

- #### Figure \@ref(fig:fig4): GR.MDf vs WT.MDf

- #### Figure \@ref(fig:fig5): GR.FOE vs WT.FOE

- #### Figure \@ref(fig:fig6): GR.MOE vs WT.MOE

- #### Figure \@ref(fig:fig7): GRF: C vs Df

- #### Figure \@ref(fig:fig8): GRM: C vs Df

- #### Figure \@ref(fig:fig9): WTF: C vs Df

- #### Figure \@ref(fig:fig10): WTM: C vs Df

- #### Figure \@ref(fig:fig11): GRF: C vs OE

- #### Figure \@ref(fig:fig12): GRM: C vs OE

- #### Figure \@ref(fig:fig13): WTF: C vs OE

- #### Figure \@ref(fig:fig14): WTM: C vs OE

<br><br><br>

```{r echo = F, include = F}
library(plotly)
library(org.Dm.eg.db)
library(ggpubr)
library(performance)
library(pathview)
library(visNetwork)
library(org.Dm.eg.db)

##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"
cpmdata = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
groups  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
TKT.EdgeR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)
TKT.EdgeR.FC = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FCTable.csv"),row.names = 1)

GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)
kegg = readRDS(gzcon(url(paste0(git.dir,"GeneralDataFiles/kegg.symbol2path.RData"))))
kegg.names = read.csv(paste0(git.dir, "GeneralDataFiles/KEGG.names.csv"), row.names = 1)

genesingo=as.list(org.Dm.egGO2ALLEGS)

##mean.cpm; Mean cpm for each gene in each condition (groups$Group). Used for filtering
mean.cpm = matrix(0, ncol = length(unique(groups$Group)), nrow = nrow(cpmdata))
colnames(mean.cpm) = unique(groups$Group)
row.names(mean.cpm) = row.names(cpmdata)
i=1
while(i<=ncol(mean.cpm)){
  mean.cpm[,i]= apply(cpmdata[,row.names(groups[groups$Group == colnames(mean.cpm)[i],])], 1, mean)
  i=i+1
}

i=1
for(i in 1:ncol(TKT.EdgeR)){
  TKT.EdgeR[TKT.EdgeR[,i] == 0,i] = 2
  i=i+1
}

rawdata = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/MetabolomicsForMetaboanalyst.txt", sep = "\t", row.names = 1)

KEGG.Key = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/RawMetabolomicData.csv", row.names = 1)
KEGG.Key = setNames(KEGG.Key[-c(1:2), 2], row.names(KEGG.Key)[-c(1:2)])

genotypes = setNames(c("WT", "GR", "TktOEWT", "TktOEGR", "TktDfWT", "TktDfGR"), c("C", "E", "A", "F", "B", "D"))

metadata = data.frame(Sample = colnames(rawdata), 
                      Group = as.character(rawdata[1,]), 
                      TIC = as.numeric(rawdata[2,]), 
                      Genotypes = genotypes[as.character(rawdata[1,])])

row.names(metadata) = metadata$Sample

norm.func = function(data){
  nd = (as.numeric(data)/metadata$TIC)*1000
  return(nd)
}

norm.data = t(apply(rawdata[3:nrow(rawdata),], 1, norm.func))
colnames(norm.data) = colnames(rawdata)

group.mean=function(data){
  gm = mean(na.omit(data))
}

mean.data = matrix(0, nrow=nrow(norm.data), ncol = length(genotypes))
row.names(mean.data) = row.names(norm.data)
colnames(mean.data) = unique(metadata$Group)

i=1
while(i<=ncol(mean.data)){
  mean.data[,i] = apply(norm.data[,metadata[metadata$Group == colnames(mean.data)[i], "Sample"]], 1, group.mean)
  i = i+1
}

mean.cpm = as.data.frame(mean.cpm)

compare.conditions = function(condition1, condition2){
  columns1 = metadata[metadata$Group == condition1, "Sample"]
  columns2 = metadata[metadata$Group == condition2, "Sample"]
  p=setNames(rep(NA, nrow(norm.data)), row.names(norm.data))
  i=1
  while(i<=length(p)){
    if(length(na.omit(norm.data[i,columns1]))>2 & 
       length(na.omit(norm.data[i,columns2]))>2){
      temp = t.test(na.omit(norm.data[i,columns1]),
                    na.omit(norm.data[i,columns2]), paired = F, var.equal = T)
      p[i] = temp[[3]]
      }
    i=i+1
    }

  fdr = p.adjust(p, "BH", length(p))

  fc = mean.data[,condition1]/mean.data[,condition2]

  comparison.table = data.frame(p = p, 
                                FDR = fdr, 
                                FC = fc, 
                                KEGG = KEGG.Key[row.names(mean.data)])

  comparison.table = comparison.table[order(comparison.table$FDR),]
}

GRxWT.C = compare.conditions("E", "C")
GRxWT.Df = compare.conditions("D", "B")
GRxWT.OE = compare.conditions("F", "A")

GR.CxDf = compare.conditions("D", "E")
WT.CxDf = compare.conditions("B", "C")

GR.CxOE = compare.conditions("F", "E")
WT.CxOE = compare.conditions("A", "C")

##Comparison cpm means
ccm = data.frame(GRF.CxDf = apply(mean.cpm[,c("TktDfGR.F","GR.F")], MARGIN = 1, mean), 
                 GRF.CxOE = apply(mean.cpm[,c("TktOEGR.F","GR.F")], MARGIN = 1, mean),
                 WTF.CxDf = apply(mean.cpm[,c("TktDfWT.F","WT.F")], MARGIN = 1, mean),
                 WTF.CxOE = apply(mean.cpm[,c("TktOEWT.F","WT.F")], MARGIN = 1, mean),
                 GRxWT.FC = apply(mean.cpm[,c("GR.F","WT.F")], MARGIN = 1, mean), 
                 GRxWT.FDf = apply(mean.cpm[,c("TktDfGR.F","TktDfWT.F")], MARGIN = 1, mean),
                 GRxWT.FOE = apply(mean.cpm[,c("TktOEGR.F","TktOEWT.F")], MARGIN = 1, mean), 
                 GRM.CxDf = apply(mean.cpm[,c("TktDfGR.M","GR.M")], MARGIN = 1, mean), 
                 GRM.CxOE = apply(mean.cpm[,c("TktOEGR.M","GR.M")], MARGIN = 1, mean),
                 WTM.CxDf = apply(mean.cpm[,c("TktDfWT.M","WT.M")], MARGIN = 1, mean),
                 WTM.CxOE = apply(mean.cpm[,c("TktOEWT.M","WT.M")], MARGIN = 1, mean),
                 GRxWT.MC = apply(mean.cpm[,c("GR.M","WT.M")], MARGIN = 1, mean), 
                 GRxWT.MDf = apply(mean.cpm[,c("TktDfGR.M","TktDfWT.M")], MARGIN = 1, mean),
                 GRxWT.MOE = apply(mean.cpm[,c("TktOEGR.M","TktOEWT.M")], MARGIN = 1, mean),
                 GR.FxM = apply(mean.cpm[,c("GR.F","GR.M")], MARGIN = 1, mean),
                 WT.FxM = apply(mean.cpm[,c("WT.F","WT.M")], MARGIN = 1, mean),
                 row.names = row.names(mean.cpm))

comparison.means = list(GRF.CxDf = c("TktDfGR.F","GR.F"), 
                 GRF.CxOE = c("TktOEGR.F","GR.F"),
                 WTF.CxDf = c("TktDfWT.F","WT.F"),
                 WTF.CxOE = c("TktOEWT.F","WT.F"),
                 GRxWT.FC = c("GR.F","WT.F"), 
                 GRxWT.FDf = c("TktDfGR.F","TktDfWT.F"),
                 GRxWT.FOE = c("TktOEGR.F","TktOEWT.F"), 
                 GRM.CxDf = c("TktDfGR.M","GR.M"), 
                 GRM.CxOE = c("TktOEGR.M","GR.M"),
                 WTM.CxDf = c("TktDfWT.M","WT.M"),
                 WTM.CxOE = c("TktOEWT.M","WT.M"),
                 GRxWT.MC = c("GR.M","WT.M"), 
                 GRxWT.MDf = c("TktDfGR.M","TktDfWT.M"),
                 GRxWT.MOE = c("TktOEGR.M","TktOEWT.M"),
                 GR.FxM = c("GR.F","GR.M"),
                 WT.FxM = c("WT.F","WT.M"))

```

```{r include = F}

##genes in the kegg pathway and their enzyme IDs
glycolysis.genes = read.csv('/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GlycolysisGenesKEGGinfo.csv')
##Edge data for genes connecting metabolites
glycolysis.edges = read.csv('/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GlycolysisEnzymeEdgesKEGGinfo.csv')
##Metabolites in the KEGG pathway with their ID
glycolysis.metabs = read.csv('/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GlycolysisMetaboliteKEGGinfo.csv')

node.data = node.info('/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/R Code/dme00010.xml')

glycolysis.network = function(comparison.T, comparison.M){

##extracting position data for nodes
node.info = data.frame(ID = unlist(node.data$kegg.names, use.names = F),
                       number = rep(names(node.data$kegg.names),
                                    lengths(node.data$kegg.names)),
                       x = 0,
                       y = 0,
                       color = 0,
                       names = "")
node.info$x = node.data$x[node.info$number]
node.info$y = node.data$y[node.info$number]

##make new columns for transcriptomic data
glycolysis.genes$FBgn = ""
glycolysis.genes$Symbol = ""
glycolysis.genes$FC = 0
glycolysis.genes$pval = 1

##Isolate gene ID info for goi
temp = GeneIDKey[GeneIDKey$CG %in% glycolysis.genes$CG,]
row.names(temp) = temp$CG

##fill in new columns with transcriptomic data
glycolysis.genes$FBgn = temp[glycolysis.genes$CG,'FBgn']
glycolysis.genes$Symbol = temp[glycolysis.genes$CG,'Symbol']
glycolysis.genes$FC = TKT.EdgeR.FC[glycolysis.genes$FBgn, comparison.T]
glycolysis.genes$FC = TKT.EdgeR[glycolysis.genes$FBgn, comparison.T]

##make new columns for metabolomic data
row.names(glycolysis.metabs) = glycolysis.metabs$KEGG.ID
glycolysis.metabs$metabolomicsID = ""
glycolysis.metabs$FC = 0
glycolysis.metabs$pval = 1

##retrieve name used in metabolomic data for metabolite IDs in path
##including IDs associated with multiple metabolomic names
i = 1
while(i<=nrow(glycolysis.metabs)){
  if(length(grep(glycolysis.metabs$KEGG.ID[i], KEGG.Key)) > 0){
    glycolysis.metabs$metabolomicsID[i]= names(KEGG.Key[grep(glycolysis.metabs$KEGG.ID[i], KEGG.Key)])
  }
  i = i +1
}

##retrieve metabolite fold change and pvalue data
glycolysis.metabs$FC = log2(comparison.M[glycolysis.metabs$metabolomicsID, "FC"])
glycolysis.metabs$pval = -log10(comparison.M[glycolysis.metabs$metabolomicsID, "p"])

##using the positions for nodes (metabolites) only
metab.info = node.info[node.info$ID %in% glycolysis.metabs$KEGG.ID,]

metab.info$FC = glycolysis.metabs[metab.info$ID, 'FC']
metab.info$pval = glycolysis.metabs[metab.info$ID, 'pval']
metab.info$names = glycolysis.metabs[metab.info$ID, 'Metabolite']
metab.info$color = glycolysis.metabs[metab.info$ID, 'FC']
metab.info$size = glycolysis.metabs[metab.info$ID, 'pval']

color.pallette = colorRampPalette(c('royalblue', 'grey', "firebrick"))(401)

metab.info$color[metab.info$color > 1] = 1
metab.info$color[metab.info$color < -1] = -1
metab.info$color = (metab.info$color * 200) +201
metab.info$color = color.pallette[metab.info$color]
metab.info$color[is.na(metab.info$color)] = 'dimgrey'

metab.info$size = 10 + 10*abs(metab.info$size)
metab.info$size[is.na(metab.info$size)] = 10

metab.info$title = paste0(glycolysis.metabs[metab.info$ID, "metabolomicsID"],
                          "<br>log2(FC): ", signif(metab.info$FC,3),
                          "<br>pval: ", signif(10^-metab.info$pval,3))

nodes = data.frame(id = paste0(metab.info$names),
                   color = metab.info$color,
                   size = metab.info$size,
                   x = 2*metab.info$x,
                   y = 2*metab.info$y,
                   physics = F,
                   group = "",
                   title = metab.info$title)

glycolysis.edges$FBgn = ''
temp.glycolysis.path = glycolysis.edges

i = 1
while(i<=nrow(glycolysis.edges)){
  enzymeID = paste0(glycolysis.edges$enzyme[i], "$")
  if(nrow(glycolysis.genes[grep(enzymeID, glycolysis.genes$Enzyme),])>0){
    temp = glycolysis.genes[grep(enzymeID, glycolysis.genes$Enzyme), "CG"]
    temp = row.names(GeneIDKey[GeneIDKey$CG %in% temp,])
    temp2 = glycolysis.edges[rep(i, length(temp)),]
    temp2$FBgn = temp
  }else{
    temp2 = glycolysis.edges[i,]
  }
  temp.glycolysis.path = rbind(temp.glycolysis.path,temp2)

  i = i +1
}

glycolysis.edges = temp.glycolysis.path[(nrow(glycolysis.edges)+1):nrow(temp.glycolysis.path),]

glycolysis.edges$size = log10(ccm[glycolysis.edges[,'FBgn'], comparison.T])
glycolysis.edges$size[glycolysis.edges$size <1] = 1
glycolysis.edges$size[is.na(glycolysis.edges$size)] = 1
glycolysis.edges$size = 2*glycolysis.edges$size

glycolysis.edges$color = TKT.EdgeR.FC[glycolysis.edges[,'FBgn'], comparison.T]
## max color intensity is at FC of 2
glycolysis.edges$color[glycolysis.edges$color>2] = 2
glycolysis.edges$color[glycolysis.edges$color<(-2)] = -2
glycolysis.edges$color = 201+(glycolysis.edges$color *100)
glycolysis.edges$color = color.pallette[glycolysis.edges$color]
glycolysis.edges$color[is.na(glycolysis.edges$color)] = 'dimgrey'

glycolysis.edges$title = paste0(GeneIDKey[glycolysis.edges$FBgn, "Symbol"],
                          "<br>log2(FC): ", signif(TKT.EdgeR.FC[glycolysis.edges[,'FBgn'], comparison.T],3),
                          "<br>FDR: ", signif(TKT.EdgeR[glycolysis.edges[,'FBgn'], comparison.T],3),
                          "<br>", comparison.means[[comparison.T]][1], " cpm: ", signif(mean.cpm[glycolysis.edges[,'FBgn'],comparison.means[[comparison.T]][1]],3),
                          "<br>", comparison.means[[comparison.T]][2], " cpm: ", signif(mean.cpm[glycolysis.edges[,'FBgn'],comparison.means[[comparison.T]][2]],3))

edges = data.frame(from = glycolysis.edges$from.name,
                   to = glycolysis.edges$to.name,
                   arrows = glycolysis.edges$arrow,
                   color = glycolysis.edges$color,
                   width = glycolysis.edges$size,
                   title = glycolysis.edges$title,
                   physics = T)

nodes = data.frame(id = paste0(nodes$id),
                   color = nodes$color,
                   size = nodes$size,
                   x = nodes$x,
                   y = nodes$y,
                   physics = F,
                   group = "",
                   title = nodes$title)

OutsidePath = data.frame(id = c('Pentose Phosphate Pathway',
                               'Propanoate Metabolism',
                               'Starch and Sucrose Metabolism',
                               'Citrate Cycle'),
                        color = "black",
                        size = 15,
                        x= c(1350,1254,966,175),
                        y=c(1112,1822,110,1700),
                        physics = F,
                        group = 'Outside Path',
                        title = "")

nodes = rbind(nodes, OutsidePath)

visNetwork(nodes, edges, main = paste0(comparison.means[[comparison.T]][1], " vs ", comparison.means[[comparison.T]][2]))%>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE)%>%
  visEdges(arrows = edges$arrows)%>%
  visGroups(groupname = "Outside Path", shape = 'box', physics=F, size = 25, font=list(color="white"))

}

```

```{r echo = F, fig.height=12, fig.width=9, fig.align='left', fig1, fig.cap = 'G85R relative to WT. Transcriptomic data is from females and metabolomic data is from both sexes. Grey is no data; Red is increase; Blue is decreae; Node size is relative to pval; Node color is relative to fold change (range: -1 to 1); Edge width is relative to FDR (max -log10(4)); edge color is relative to fold change (range: -2 to 2)'}

glycolysis.network(comparison.T = 'GRxWT.FC', comparison.M = GRxWT.C)

```

<br><br><br>

```{r, echo =F, fig2, fig.height=12, fig.width=9, fig.align='left', fig.cap = 'G85R relative to WT. Transcriptomic data is from males and metabolomic data is from both sexes.Grey is no data; Red is increase; Blue is decreae; Node size is relative to pval; Node color is relative to fold change (range: -1 to 1); Edge width is relative to FDR (max -log10(4)); edge color is relative to fold change (range: -2 to 2)'}

glycolysis.network(comparison.T = 'GRxWT.MC', comparison.M = GR.CxDf)

```

<br><br><br>

```{r, echo =F, fig3, fig.height=12, fig.width=9, fig.align='left', fig.cap = 'G85R TKT-Df relative to WT TKT-Df. Transcriptomic data is from females and metabolomic data is from both sexes. Grey is no data; Red is increase; Blue is decreae; Node size is relative to pval; Node color is relative to fold change (range: -1 to 1); Edge width is relative to FDR (max -log10(4)); edge color is relative to fold change (range: -2 to 2)'}

glycolysis.network(comparison.T = 'GRxWT.FDf', comparison.M = GRxWT.Df)

```

<br><br><br>

```{r, echo =F, fig4, fig.height=12, fig.width=9, fig.align='left', fig.cap = 'G85R TKT-Df relative to WT TKT-Df. Transcriptomic data is from males and metabolomic data is from both sexes. Grey is no data; Red is increase; Blue is decreae; Node size is relative to pval; Node color is relative to fold change (range: -1 to 1); Edge width is relative to FDR (max -log10(4)); edge color is relative to fold change (range: -2 to 2)'}

glycolysis.network(comparison.T = 'GRxWT.MDf', comparison.M = GRxWT.Df)

```

<br><br><br>

```{r, echo =F, fig5, fig.height=12, fig.width=9, fig.align='left', fig.cap = 'G85R TKT-OE relative to WT TKT-OE. Transcriptomic data is from females and metabolomic data is from both sexes. Grey is no data; Red is increase; Blue is decreae; Node size is relative to pval; Node color is relative to fold change (range: -1 to 1); Edge width is relative to FDR (max -log10(4)); edge color is relative to fold change (range: -2 to 2)'}

glycolysis.network(comparison.T = 'GRxWT.FOE', comparison.M = GRxWT.OE)

```

<br><br><br>

```{r, echo =F, fig6, fig.height=12, fig.width=9, fig.align='left', fig.cap = 'G85R TKT-OE relative to WT TKT-OE. Transcriptomic data is from males and metabolomic data is from both sexes. Grey is no data; Red is increase; Blue is decreae; Node size is relative to pval; Node color is relative to fold change (range: -1 to 1); Edge width is relative to FDR (max -log10(4)); edge color is relative to fold change (range: -2 to 2)'}

glycolysis.network(comparison.T = 'GRxWT.MOE', comparison.M = GRxWT.OE)

```

<br><br><br>

```{r, echo =F, fig7, fig.height=12, fig.width=9, fig.align='left', fig.cap = 'G85R control relative to G85R TKT-Df. Transcriptomic data is from females and metabolomic data is from both sexes. Grey is no data; Red is increase; Blue is decreae; Node size is relative to pval; Node color is relative to fold change (range: -1 to 1); Edge width is relative to FDR (max -log10(4)); edge color is relative to fold change (range: -2 to 2)'}

glycolysis.network(comparison.T = 'GRF.CxDf', comparison.M = GR.CxDf)

```

<br><br><br>

```{r, echo =F, fig8, fig.height=12, fig.width=9, fig.align='left', fig.cap = 'G85R control relative to G85R TKT-Df. Transcriptomic data is from males and metabolomic data is from both sexes. Grey is no data; Red is increase; Blue is decreae; Node size is relative to pval; Node color is relative to fold change (range: -1 to 1); Edge width is relative to FDR (max -log10(4)); edge color is relative to fold change (range: -2 to 2)'}

glycolysis.network(comparison.T = 'GRM.CxDf', comparison.M = GR.CxDf)

```

<br><br><br>

```{r, echo =F, fig9, fig.height=12, fig.width=9, fig.align='left', fig.cap = 'WT control relative to WT TKT-Df. Transcriptomic data is from females and metabolomic data is from both sexes. Grey is no data; Red is increase; Blue is decreae; Node size is relative to pval; Node color is relative to fold change (range: -1 to 1); Edge width is relative to FDR (max -log10(4)); edge color is relative to fold change (range: -2 to 2)'}

glycolysis.network(comparison.T = 'WTF.CxDf', comparison.M = WT.CxDf)

```

<br><br><br>

```{r, echo =F, fig10, fig.height=12, fig.width=9, fig.align='left', fig.cap = 'WT control relative to WT TKT-Df. Transcriptomic data is from males and metabolomic data is from both sexes. Grey is no data; Red is increase; Blue is decreae; Node size is relative to pval; Node color is relative to fold change (range: -1 to 1); Edge width is relative to FDR (max -log10(4)); edge color is relative to fold change (range: -2 to 2)'}

glycolysis.network(comparison.T = 'WTM.CxDf', comparison.M = WT.CxDf)

```

<br><br><br>

```{r, echo =F, fig11, fig.height=12, fig.width=9, fig.align='left', fig.cap = 'G85R control relative to G85R TKT-OE. Transcriptomic data is from females and metabolomic data is from both sexes. Grey is no data; Red is increase; Blue is decreae; Node size is relative to pval; Node color is relative to fold change (range: -1 to 1); Edge width is relative to FDR (max -log10(4)); edge color is relative to fold change (range: -2 to 2)'}

glycolysis.network(comparison.T = 'GRF.CxOE', comparison.M = GR.CxOE)

```

<br><br><br>

```{r, echo =F, fig12, fig.height=12, fig.width=9, fig.align='left', fig.cap = 'G85R control relative to G85R TKT-OE. Transcriptomic data is from males and metabolomic data is from both sexes. Grey is no data; Red is increase; Blue is decreae; Node size is relative to pval; Node color is relative to fold change (range: -1 to 1); Edge width is relative to FDR (max -log10(4)); edge color is relative to fold change (range: -2 to 2)'}

glycolysis.network(comparison.T = 'GRM.CxOE', comparison.M = GR.CxOE)

```

<br><br><br>

```{r, echo =F, fig13, fig.height=12, fig.width=9, fig.align='left', fig.cap = 'WT control relative to WT TKT-OE. Transcriptomic data is from females and metabolomic data is from both sexes. Grey is no data; Red is increase; Blue is decreae; Node size is relative to pval; Node color is relative to fold change (range: -1 to 1); Edge width is relative to FDR (max -log10(4)); edge color is relative to fold change (range: -2 to 2)'}

glycolysis.network(comparison.T = 'WTF.CxOE', comparison.M = WT.CxOE)

```

<br><br><br>

```{r, echo =F, fig14, fig.height=12, fig.width=9, fig.align='left', fig.cap = 'WT control relative to WT TKT-OE. Transcriptomic data is from males and metabolomic data is from both sexes. Grey is no data; Red is increase; Blue is decreae; Node size is relative to pval; Node color is relative to fold change (range: -1 to 1); Edge width is relative to FDR (max -log10(4)); edge color is relative to fold change (range: -2 to 2)'}

glycolysis.network(comparison.T = 'WTM.CxOE', comparison.M = WT.CxOE)

```

<br><br><br>