### 2024-01-17 TKT_RNAseq/R Code/PurineMetabolismNetwork.Rmd

---
title: "TKT_MetabPaper_Analyses"
author: "John Santiago"
date: "2023-09-29"
output: 
  bookdown::html_document2:
    number_sections: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

- #### Figure \@ref(fig:fig1): GR.FC vs WT.FC

- #### Figure \@ref(fig:fig2): GR.MC vs WT.MC

- #### Figure \@ref(fig:fig3): GR.FDf vs WT.FDf

- #### Figure \@ref(fig:fig4): GR.MDf vs WT.MDf

- #### Figure \@ref(fig:fig5): GR.FOE vs WT.FOE

- #### Figure \@ref(fig:fig6): GR.MOE vs WT.MOE

- #### Figure \@ref(fig:fig7): GRF: C vs Df

- #### Figure \@ref(fig:fig8): GRM: C vs Df

- #### Figure \@ref(fig:fig9): WTF: C vs Df

- #### Figure \@ref(fig:fig10): WTM: C vs Df

- #### Figure \@ref(fig:fig11): GRF: C vs OE

- #### Figure \@ref(fig:fig12): GRM: C vs OE

- #### Figure \@ref(fig:fig13): WTF: C vs OE

- #### Figure \@ref(fig:fig14): WTM: C vs OE

```{r echo = F, include = F}
library(plotly)
library(org.Dm.eg.db)
library(ggpubr)
library(performance)

##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"
cpmdata = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
groups  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
TKT.EdgeR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)
TKT.EdgeR.FC = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FCTable.csv"),row.names = 1)

GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)
kegg = readRDS(gzcon(url(paste0(git.dir,"GeneralDataFiles/kegg.symbol2path.RData"))))
kegg.names = read.csv(paste0(git.dir, "GeneralDataFiles/KEGG.names.csv"), row.names = 1)

Synaptic = read.csv("/Users/johncsantiago/Documents/SynapticGenesandTFs.csv")

genesingo=as.list(org.Dm.egGO2ALLEGS)

##mean.cpm; Mean cpm for each gene in each condition (groups$Group). Used for filtering
mean.cpm = matrix(0, ncol = length(unique(groups$Group)), nrow = nrow(cpmdata))
colnames(mean.cpm) = unique(groups$Group)
row.names(mean.cpm) = row.names(cpmdata)
i=1
while(i<=ncol(mean.cpm)){
  mean.cpm[,i]= apply(cpmdata[,row.names(groups[groups$Group == colnames(mean.cpm)[i],])], 1, mean)
  i=i+1
}

i=1
for(i in 1:ncol(TKT.EdgeR)){
  TKT.EdgeR[TKT.EdgeR[,i] == 0,i] = 2
  i=i+1
}

rawdata = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/MetabolomicsForMetaboanalyst.txt", sep = "\t", row.names = 1)

KEGG.Key = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/RawMetabolomicData.csv", row.names = 1)
KEGG.Key = setNames(KEGG.Key[-c(1:2), 2], row.names(KEGG.Key)[-c(1:2)])

genotypes = setNames(c("WT", "GR", "TktOEWT", "TktOEGR", "TktDfWT", "TktDfGR"), c("C", "E", "A", "F", "B", "D"))

metadata = data.frame(Sample = colnames(rawdata), 
                      Group = as.character(rawdata[1,]), 
                      TIC = as.numeric(rawdata[2,]), 
                      Genotypes = genotypes[as.character(rawdata[1,])])

row.names(metadata) = metadata$Sample

norm.func = function(data){
  nd = (as.numeric(data)/metadata$TIC)*1000
  return(nd)
}

norm.data = t(apply(rawdata[3:nrow(rawdata),], 1, norm.func))
colnames(norm.data) = colnames(rawdata)

group.mean=function(data){
  gm = mean(na.omit(data))
}

mean.data = matrix(0, nrow=nrow(norm.data), ncol = length(genotypes))
row.names(mean.data) = row.names(norm.data)
colnames(mean.data) = unique(metadata$Group)

i=1
while(i<=ncol(mean.data)){
  mean.data[,i] = apply(norm.data[,metadata[metadata$Group == colnames(mean.data)[i], "Sample"]], 1, group.mean)
  i = i+1
}

mean.cpm = as.data.frame(mean.cpm)

compare.conditions = function(condition1, condition2){
  columns1 = metadata[metadata$Group == condition1, "Sample"]
  columns2 = metadata[metadata$Group == condition2, "Sample"]
  p=setNames(rep(NA, nrow(norm.data)), row.names(norm.data))
  i=1
  while(i<=length(p)){
    if(length(na.omit(norm.data[i,columns1]))>2 & 
       length(na.omit(norm.data[i,columns2]))>2){
      temp = t.test(na.omit(norm.data[i,columns1]),
                    na.omit(norm.data[i,columns2]), paired = F, var.equal = T)
      p[i] = temp[[3]]
      }
    i=i+1
    }

  fdr = p.adjust(p, "BH", length(p))

  fc = mean.data[,condition1]/mean.data[,condition2]

  comparison.table = data.frame(p = p, 
                                FDR = fdr, 
                                FC = fc, 
                                KEGG = KEGG.Key[row.names(mean.data)])

  comparison.table = comparison.table[order(comparison.table$FDR),]
}

GRxWT.C = compare.conditions("E", "C")
GRxWT.Df = compare.conditions("D", "B")
GRxWT.OE = compare.conditions("F", "A")

GR.CxDf = compare.conditions("D", "E")
WT.CxDf = compare.conditions("B", "C")

GR.CxOE = compare.conditions("F", "E")
WT.CxOE = compare.conditions("A", "C")

```

```{r echo = F, include = F}

tempF = apply(mean.cpm[,c("GR.F", "WT.F")], MARGIN = 1, mean)
tempM = apply(mean.cpm[,c("GR.M", "WT.M")], MARGIN = 1, mean)
tempF = names(tempF)[tempF>10]
tempM = names(tempM)[tempM>10]
cutoff.genes = unique(c(tempF, tempM))

subset.genes = row.names(TKT.EdgeR)[TKT.EdgeR$GRxWT.FC < .05 | TKT.EdgeR$GRxWT.MC < .05]
subset.genes = setdiff(subset.genes, row.names(TKT.EdgeR)[TKT.EdgeR$GRxWT.FC < .05 & TKT.EdgeR$GRxWT.MC < .05])
subset.genes = intersect(subset.genes, cutoff.genes)

FCx = TKT.EdgeR.FC[subset.genes, "GRxWT.FC"]
FCy = TKT.EdgeR.FC[subset.genes, "GRxWT.MC"]

set.size = F

volcano.data = data.frame(Symbol = GeneIDKey[subset.genes, "Symbol"], 
                          FCx = FCx,
                          FCy = FCy,
                          Color = 'grey',
                          size = 2.5,
                          FDR.F = TKT.EdgeR[subset.genes, "GRxWT.FC"],
                          FDR.M = TKT.EdgeR[subset.genes, "GRxWT.MC"],
                          GR.FC.cpm = mean.cpm[subset.genes, "GR.F"],
                          WT.FC.cpm = mean.cpm[subset.genes, "WT.F"],
                          GR.MC.cpm = mean.cpm[subset.genes, "GR.M"],
                          WT.MC.cpm = mean.cpm[subset.genes, "WT.M"],
                          tempcolor = FCx - FCy)

row.names(volcano.data) = subset.genes

volcano.data$Symbol[is.na(volcano.data$Symbol)] = row.names(volcano.data)[is.na(volcano.data$Symbol)]

main.title = "Female and Male G85R/WT FC, Sig. DEG in Males and Females G85R vs WT"

maxx= max(c(abs(min(FCx)),abs(max(FCx))))
maxy= max(c(abs(min(FCy)),abs(max(FCy))))

fig = plot_ly(data = volcano.data,
              x = ~FCx,
              y = ~FCy,
              type = 'scatter',
              mode = 'markers',
              marker = list(colorscale = list(c(0,.7,1), c("firebrick","ivory", "royalblue")),
                            reversescale =T,
                             colorbar = list(title = "delta FC",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                            color = ~tempcolor,
                            line = list(color = "black", width = 1.5),
                            size = 5*volcano.data$size),
              hoverinfo = "text",
              hovertext = paste("Gene:", volcano.data$Symbol,
                      "\nFemale FDR: ", round(volcano.data$FDR.F,4),
                      "\nMale FDR: ", round(volcano.data$FDR.M,4),
                      "\nFemale mean cpm (G85R/WT): ", round(volcano.data$GR.FC.cpm, 1),
                      "/", round(volcano.data$WT.FC.cpm, 1),
                      "\nMale mean cpm(G85R/WT): ", round(volcano.data$GR.MC.cpm, 1),
                      "/", round(volcano.data$WT.MC.cpm, 1)))
fig = fig %>% layout(xaxis = list(title = "Female log2(FC)"),
                     yaxis = list(title = "Male log2(FC)"),
                     title = main.title,
                     shapes = list(
                       type = "line",
                       line = list(color = "black", dash = 'dash', width = 2),
                       x0 = floor(-maxx),
                       x1 = ceiling(maxx),
                       y0 = floor(-maxy),
                       y1 = ceiling(maxy)))

fig

```

```{r include = F}

library(visNetwork)
library(org.Dm.eg.db)
keggpath = as.list(org.Dm.egPATH2EG)
nodeID = as.list(org.Dm.egENZYME)
nodeID['40528'] = '3.5.4.3'
nodeID['41318'] = '2.7.4.10'
nodeID['42273'] = '2.7.1.145'
nodeID['33924'] = '3.1.3.89'
nodeID[['31309']] = c("3.1.4.17", '3.1.4.53')
nodeID['37741'] = '3.1.4.53'

purine.path = data.frame(enzyme = c('1.17.1.4', '1.17.3.2'),
                         from = c('xanthine', 'xanthine'),
                         to = c(rep('uric acid',2)))

xanthine = data.frame(enzyme = c('1.17.1.4', '1.17.3.2', '2.4.2.1', '3.5.4.3'),
                      from = c('hypoxanthine', 'hypoxanthine', 'xanthosine', 'guanine'),
                      to = c(rep('xanthine',4)))
purine.path = rbind(purine.path, xanthine)

hypoxanthine = data.frame(enzyme = c('2.4.2.1', '2.4.2.1'),
                          from = c('deoxyinosine', 'inosine'),
                          to = c(rep('hypoxanthine', 2)))
purine.path = rbind(purine.path, hypoxanthine)

xanthosine = data.frame(enzyme = c('3.1.3.5'),
                        from = c('XMP'),
                        to = 'xanthosine')
purine.path = rbind(purine.path, xanthosine)

XMP = data.frame(enzyme = c('1.1.1.205'),
                 from = c('IMP'),
                 to = 'XMP')
purine.path = rbind(purine.path, XMP)

IMP = data.frame(enzyme = c('3.6.1.6', '3.5.4.6'),
                 from = c('IDP', 'AMP'),
                 to = c(rep('IMP', 2)))
purine.path = rbind(purine.path, IMP)

adenylosuccinate = data.frame(enzyme = c('6.3.4.4'),
                              from = c('IMP'),
                              to = 'adenylosuccinate')
purine.path = rbind(purine.path, adenylosuccinate)

inosine = data.frame(enzyme = c('2.4.2.1', '3.5.4.4', '3.1.3.5'),
                     from = c('hypoxanthine', 'adenosine', 'IMP'),
                     to = c(rep('inosine', 3)))
purine.path = rbind(purine.path, inosine)

deoxyinosine = data.frame(enzyme = c('2.4.2.1', '3.5.4.4'),
                          from = c('hypoxanthine', 'deoxyadenosine'),
                          to = c(rep('deoxyinosine', 2)))
purine.path = rbind(purine.path, deoxyinosine)

adenosine = data.frame(enzyme = c('2.4.2.1', '3.1.3.5', '2.7.1.20'),
                       from = c('adenine', 'AMP', 'AMP'),
                       to = c(rep('adenosine', 3)))
purine.path = rbind(purine.path, adenosine)

AMP = data.frame(enzyme = c('4.3.2.2', '2.7.1.20', '2.4.2.7', '2.7.4.3', '2.7.4.10', '3.1.4.17', '3.1.4.53'),
                 from = c('adenylosuccinate', 'adenosine', 'adenine', 'ADP', 'ADP', 'cAMP', 'cAMP'),
                 to = c(rep('AMP', 7)))
purine.path = rbind(purine.path, AMP)

adenine = data.frame(enzyme = c('2.4.2.1', '2.4.2.7', '2.4.2.1'),
                     from = c('adenosine', 'AMP', 'deoxyadenosine'),
                     to = c(rep('adenine', 3)))
purine.path = rbind(purine.path, adenine)

ATP = data.frame(enzyme = c('2.7.4.6'),
                 from = c('ADP'),
                 to = c(rep('ATP', 1)))
purine.path = rbind(purine.path, ATP)

dATP = data.frame(enzyme = c('2.7.4.6'),
                  from = c('dADP'),
                  to = c(rep('dATP', 1)))
purine.path = rbind(purine.path, dATP)

ADP = data.frame(enzyme = c('2.7.4.3', '2.7.4.10', '2.7.4.6', '3.6.1.15'),
                 from = c('AMP', 'AMP', 'ATP', 'ATP'),
                 to = c(rep('ADP', 4)))
purine.path = rbind(purine.path, ADP)

dADP = data.frame(enzyme = c('2.7.4.3', '1.17.4.1', '2.7.4.6'),
                  from = c('dAMP', 'ADP', 'dATP'),
                  to = c(rep('dADP', 3)))
purine.path = rbind(purine.path, dADP)

dAMP = data.frame(enzyme = c('2.7.1.145', '2.7.4.3'),
                  from = c('deoxyadenosine', 'dADP'),
                  to = c(rep('dAMP', 2)))
purine.path = rbind(purine.path, dAMP)

deoxyadenosine = data.frame(enzyme = c('2.4.2.1', '3.1.3.5', '3.1.3.89'),
                            from = c('adenine', 'dAMP', 'dAMP'),
                            to = c(rep('deoxyadenosine', 3)))
purine.path = rbind(purine.path, deoxyadenosine)

dGTP = data.frame(enzyme = c('2.7.4.6'),
                  from = c('dGDP'),
                  to = c('dGTP'))
purine.path = rbind(purine.path, dGTP)

guanine = data.frame(enzyme = c('2.4.2.1', '2.4.2.1'),
                     from = c('guanosine', 'deoxyguanosine'),
                     to = c(rep('guanine', 2)))
purine.path = rbind(purine.path, guanine)

dGDP = data.frame(enzyme = c('2.7.4.6', '1.17.4.1', '2.7.4.8'),
                  from = c('dGTP', 'GDP', 'dGMP'),
                  to = c(rep('dGDP', 3)))
purine.path = rbind(purine.path, dGDP)

guanosine = data.frame(enzyme = c('2.4.2.1', '3.1.3.5'),
                       from = c('guanine', 'GMP'),
                       to = c(rep('guanosine', 2)))
purine.path = rbind(purine.path, guanosine)

GTP = data.frame(enzyme = c('2.7.4.6'),
                 from = c('GDP'),
                 to = c(rep('GTP', 1)))
purine.path = rbind(purine.path, GTP)

GDP = data.frame(enzyme = c('2.7.4.6', '2.7.4.8'),
                 from = c('GTP', 'GMP'),
                 to = c(rep('GDP', 2)))
purine.path = rbind(purine.path, GDP)

GMP = data.frame(enzyme = c('6.3.5.2','3.6.1.6', '2.7.4.8'),
                 from = c('XMP', 'GDP', 'GDP'),
                 to = c(rep('GMP', 3)))
purine.path = rbind(purine.path, GMP)

dGMP = data.frame(enzyme = c('2.7.4.8', '2.7.1.145'),
                  from = c('dGDP', 'deoxyguanosine'),
                  to = c(rep('dGMP', 2)))
purine.path = rbind(purine.path, dGMP)

cAMP = data.frame(enzyme = '4.6.1.1',
                  from = 'ATP',
                  to = 'cAMP')
purine.path = rbind(purine.path, cAMP)

FGAM = data.frame(c(),
               c())

purine.path$arrows = 'to'
temp = purine.path
temp[,5] = paste0(purine.path[,1], purine.path[,2], purine.path[,3])
temp[,6] = paste0(purine.path[,1], purine.path[,3], purine.path[,2])
temp$keep = TRUE

i=1
while(i<=nrow(temp)){
  if(length(intersect(temp[i,6], temp[temp$keep==TRUE,5]))==1){
        temp$keep[i] = FALSE
        temp[temp[,5] == temp[i,6], 'arrows'] = 'to;from'
  }
  i=i+1
}

purine.path = temp[temp$keep, 1:4]
purine.path$FBgn = ''

temp.purine.path = purine.path

i=1
while(i<=nrow(purine.path)){  
  enzymeID = purine.path$enzyme[i]
  temp = names(nodeID[grep(enzymeID, nodeID)])
  temp = setNames(unlist(nodeID[temp], use.names = F), rep(names(nodeID[temp]), lengths(nodeID[temp])))
  temp = names(temp[temp == enzymeID])
  temp = row.names(GeneIDKey[GeneIDKey$ensembl %in% temp, ])
  temp2 = purine.path[rep(i, length(temp)),]
  temp2$FBgn = temp
  temp.purine.path = rbind(temp.purine.path,temp2)
  i = i +1
}

purine.path = temp.purine.path[(nrow(purine.path)+1):nrow(temp.purine.path),]
purine.path = purine.path[!duplicated(purine.path[,2:5]),]

TerminalPoints = data.frame(enzyme = "Excretion",
                            from = 'uric acid',
                            to = 'Excretion')

DNA.A = data.frame(enzyme = c('DNA.A'),
                 from = c('dATP'),
                 to = c('DNA.A'))
TerminalPoints = rbind(TerminalPoints, DNA.A)

DNA.G = data.frame(enzyme = c('DNA.G'),
                 from = c('dGTP'),
                 to = c('DNA.G'))
TerminalPoints = rbind(TerminalPoints, DNA.G)

RNA.A = data.frame(enzyme = c('RNA.A'),
                 from = c('ATP'),
                 to = c('RNA.A'))
TerminalPoints = rbind(TerminalPoints, RNA.A)

RNA.G = data.frame(enzyme = c('RNA.G'),
                 from = c('GTP'),
                 to = c('RNA.G'))
TerminalPoints = rbind(TerminalPoints, RNA.G)

Riboflavin.Metabolism = data.frame(enzyme = c('Riboflavin.Metabolism'),
                 from = c('GTP'),
                 to = c('Riboflavin.Metabolism'))
TerminalPoints = rbind(TerminalPoints, Riboflavin.Metabolism)

Folate.Biosynthesis = data.frame(enzyme = c('Folate.Biosynthesis'),
                 from = c('GTP'),
                 to = c('Folate.Biosynthesis'))
TerminalPoints = rbind(TerminalPoints, Folate.Biosynthesis)

PPP = data.frame(enzyme = c('Pentose Phosphate'),
                 from = c('Pentose Phosphate'),
                 to = c('IMP'))
TerminalPoints = rbind(TerminalPoints, PPP)

TerminalPoints = data.frame(enzyme = TerminalPoints$enzyme,
                            from = TerminalPoints$from,
                            to = TerminalPoints$to,
                            arrows = 'to',
                            FBgn = NA,
                            color = 'black',
                            width = 1,
                            symbol = NA,
                            fdr = NA,
                            GR = NA,
                            WT = NA,
                            title = '')

```

```{r include = F}

network.fun = function(comparison.T, comparison.M, cond1, cond2){
temp = colorRampPalette(c('royalblue',"dodgerblue",'red', "firebrick"))(401)

visedges = purine.path
visedges$color = round(TKT.EdgeR.FC[purine.path$FBgn, comparison.T],2)
##visedges$width = .5 + abs(visedges$color)
visedges$color[visedges$color < (-2)] = -2
visedges$color[visedges$color > (2)] = 2
visedges$color = (visedges$color * 100) +201
visedges$color = temp[visedges$color]
visedges$symbol = GeneIDKey[visedges$FBgn, 'Symbol']
visedges$fdr = signif(TKT.EdgeR[visedges$FBgn, comparison.T], 2)
visedges$GR = round(mean.cpm[visedges$FBgn, cond1],1)
visedges$WT = round(mean.cpm[visedges$FBgn, cond2],1)
visedges$width = .01 + (1*abs(log10(visedges$WT)))

visedges$title = paste0(visedges$symbol,
                        "<br>FDR: ", visedges$fdr, 
                        "<br>",cond1," CPM: ", visedges$GR, 
                        "<br>",cond2," CPM: ", visedges$WT)
visedges = na.omit(visedges)
visedges = rbind(visedges, TerminalPoints)

node.names = unique(c(visedges$from, visedges$to))
node.data = data.frame(id = node.names,
                       color = 'dimgrey',
                       size = 10,
                       FC = NA,
                       pval = NA,
                       row.names = node.names)

row.names(comparison.M)[row.names(comparison.M) == "cAMP(2',3') / cAMP(3',5')"] = 'cAMP'

node.colors = na.omit(comparison.M[intersect(row.names(comparison.M), node.names),])
node.pval = setNames(signif((node.colors$p),2), row.names(node.colors))
node.data[names(node.pval), 'pval'] = node.pval
node.colors = setNames(round(log2(node.colors$FC),2), row.names(node.colors))
node.data[names(node.colors), "FC"] = node.colors
node.data[names(node.colors),'size'] = 
  node.data[names(node.colors),'size'] + 7.5*abs(node.colors)
node.colors[node.colors < (-2)] = -2
node.colors[node.colors>2] = 2

node.colors = (node.colors*100)+201

node.colors = setNames(temp[node.colors], names(node.colors)) 
node.data[names(node.colors), 'color'] = node.colors

node.data$title = paste0(node.data$id,
                         "<br>FC: ", node.data$FC,
                         "<br>pval: ", node.data$pval)

node.data[unique(TerminalPoints$enzyme), 'color'] = 'black'

node.data$group = 'metabolite'
node.data[node.data$color == 'black', 'group'] = "Outside Path"
node.data[node.data$color == 'dimgrey', 'group'] = "no data"

node.data$x = 1
node.data$y = 1
node.data["Excretion", 'x'] = -200
node.data["Excretion", 'y'] = 500

node.data["Pentose Phosphate" , 'x'] = 50
node.data["Pentose Phosphate" , 'y'] = -400

node.data["RNA.A", 'x'] = 850
node.data["RNA.A", 'y'] = -250

node.data["RNA.G", 'x'] = -1000
node.data["RNA.G", 'y'] = -250

node.data["Folate.Biosynthesis", 'x'] = -1000
node.data["Folate.Biosynthesis", 'y'] = -350

node.data["Riboflavin.Metabolism", 'x'] = -1000
node.data["Riboflavin.Metabolism", 'y'] = -150

node.data["DNA.A", 'x'] = 850
node.data["DNA.A", 'y'] = 250

node.data["DNA.G", 'x'] = -1000
node.data["DNA.G", 'y'] = 250

# passing custom nodes and/or edges
lnodes <- data.frame(label = c("Increased", "Decreased", "No Data"), 
shape = 'dot', color = c("firebrick", "royalblue",'dimgrey'),
 title = "")

nodes = node.data
visedges = visedges[,c('from', 'to', 'arrows', 'color', 'width', 'title')]

visNetwork(nodes,visedges,
                  main = paste0(cond1, " vs ", cond2))%>%
  visNodes(physics = T)%>%
  visLayout(hierarchical=F, improvedLayout=T)%>%
  ##visPhysics(solver="forceAtlas2Based", forceAtlas2Based = list(gravitationalConstant=-50),enabled=T, minVelocity = .5)%>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE)%>%
  visEdges(arrows = visedges$arrows)%>%
  visGroups(groupname = "Outside Path", shape = 'box', physics=F, size = 25, font=list(color="white"))%>%
    visLegend(addNodes = lnodes, useGroups = FALSE, width = .115)

}

```

```{r echo = F}

##tempF = apply(mean.cpm[,c("GR.F", "WT.F")], MARGIN = 1, mean)
##tempM = apply(mean.cpm[,c("GR.M", "WT.M")], MARGIN = 1, mean)
##tempF = names(tempF)[tempF>10]
##tempM = names(tempM)[tempM>10]
##cutoff.genes = unique(c(tempF, tempM))

##subset.genes = row.names(TKT.EdgeR)[TKT.EdgeR$GRxWT.FC < .05 | TKT.EdgeR$GRxWT.MC < .05]
##subset.genes = setdiff(subset.genes, row.names(TKT.EdgeR)[TKT.EdgeR$GRxWT.FC < .05 & TKT.EdgeR$GRxWT.MC < .05])
##subset.genes = intersect(subset.genes, cutoff.genes)

subset.genes = intersect(row.names(TKT.EdgeR.FC), unique(purine.path$FBgn))

FCx = TKT.EdgeR.FC[subset.genes, "GRxWT.FC"]
FCy = TKT.EdgeR.FC[subset.genes, "GRxWT.MC"]

set.size = F

volcano.data = data.frame(Symbol = GeneIDKey[subset.genes, "Symbol"], 
                          FCx = FCx,
                          FCy = FCy,
                          Color = 'grey',
                          size = 2.5,
                          FDR.F = TKT.EdgeR[subset.genes, "GRxWT.FC"],
                          FDR.M = TKT.EdgeR[subset.genes, "GRxWT.MC"],
                          GR.FC.cpm = mean.cpm[subset.genes, "GR.F"],
                          WT.FC.cpm = mean.cpm[subset.genes, "WT.F"],
                          GR.MC.cpm = mean.cpm[subset.genes, "GR.M"],
                          WT.MC.cpm = mean.cpm[subset.genes, "WT.M"],
                          tempcolor = FCx - FCy)

row.names(volcano.data) = subset.genes

volcano.data$Symbol[is.na(volcano.data$Symbol)] = row.names(volcano.data)[is.na(volcano.data$Symbol)]

main.title = "Female and Male G85R/WT FC, Sig. DEG in Males and Females G85R vs WT"

maxx= max(c(abs(min(FCx)),abs(max(FCx))))
maxy= max(c(abs(min(FCy)),abs(max(FCy))))

fig = plot_ly(data = volcano.data,
              x = ~FCx,
              y = ~FCy,
              type = 'scatter',
              mode = 'markers',
              marker = list(colorscale = list(c(0,.7,1), c("firebrick","ivory", "royalblue")),
                            reversescale =T,
                             colorbar = list(title = "delta FC",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                            color = ~tempcolor,
                            line = list(color = "black", width = 1.5),
                            size = 5*volcano.data$size),
              hoverinfo = "text",
              hovertext = paste("Gene:", volcano.data$Symbol,
                      "\nFemale FDR: ", round(volcano.data$FDR.F,4),
                      "\nMale FDR: ", round(volcano.data$FDR.M,4),
                      "\nFemale mean cpm (G85R/WT): ", round(volcano.data$GR.FC.cpm, 1),
                      "/", round(volcano.data$WT.FC.cpm, 1),
                      "\nMale mean cpm(G85R/WT): ", round(volcano.data$GR.MC.cpm, 1),
                      "/", round(volcano.data$WT.MC.cpm, 1)))
fig = fig %>% layout(xaxis = list(title = "Female log2(FC)"),
                     yaxis = list(title = "Male log2(FC)"),
                     title = main.title,
                     shapes = list(
                       type = "line",
                       line = list(color = "black", dash = 'dash', width = 2),
                       x0 = floor(-maxx),
                       x1 = ceiling(maxx),
                       y0 = floor(-maxy),
                       y1 = ceiling(maxy)))

fig

```

```{r echo = F, fig.height=9, fig.width=12, fig.align='left', fig1, fig.cap = 'G85R relative to WT. Transcriptomic data is from females and metabolomic data is from both sexes. Node and arrow size and color '}

comparison.T = "GRxWT.FC"
cond1 = 'GR.F'
cond2 = "WT.F"
comparison.M = GRxWT.C

network.fun(comparison.T, comparison.M, cond1, cond2)

```

<br><br><br>

```{r echo = F, fig.height=9, fig.width=12, fig.align='left', fig2, fig.cap = 'G85R relative to WT. Transcriptomic data is from males and metabolomic data is from both sexes. Node and arrow size and color '}

comparison.T = "GRxWT.MC"
cond1 = 'GR.M'
cond2 = "WT.M"
comparison.M = GRxWT.C

network.fun(comparison.T, comparison.M, cond1, cond2)

```

<br><br><br>

```{r echo = F, fig.height=9, fig.width=12, fig.align='left', fig3, fig.cap = 'G85R TKT-Df relative to WT TKT-Df. Transcriptomic data is from females and metabolomic data is from both sexes. Node and arrow size and color '}

comparison.T = "GRxWT.FDf"
cond1 = "TktDfGR.F"
cond2 = "TktDfWT.F"
comparison.M = GRxWT.Df

network.fun(comparison.T, comparison.M, cond1, cond2)

```

<br><br><br>

```{r echo = F, fig.height=9, fig.width=12, fig.align='left', fig4, fig.cap = 'G85R TKT-Df relative to WT TKT-Df. Transcriptomic data is from males and metabolomic data is from both sexes. Node and arrow size and color '}

comparison.T = "GRxWT.MDf"
cond1 = "TktDfGR.M"
cond2 = "TktDfWT.M"
comparison.M = GRxWT.Df

network.fun(comparison.T, comparison.M, cond1, cond2)

```

<br><br><br>

```{r echo = F, fig.height=9, fig.width=12, fig.align='left', fig5, fig.cap = 'G85R TKT-OE relative to WT TKT-OE. Transcriptomic data is from females and metabolomic data is from both sexes. Node and arrow size and color '}

comparison.T = "GRxWT.FOE"
cond1 = "TktOEGR.F"
cond2 = "TktOEWT.F"
comparison.M = GRxWT.OE

network.fun(comparison.T, comparison.M, cond1, cond2)

```

<br><br><br>

```{r echo = F, fig.height=9, fig.width=12, fig.align='left', fig6, fig.cap = 'G85R TKT-Df relative to WT TKT-Df. Transcriptomic data is from males and metabolomic data is from both sexes. Node and arrow size and color '}

comparison.T = "GRxWT.MOE"
cond1 = "TktOEGR.M"
cond2 = "TktOEWT.M"
comparison.M = GRxWT.OE

network.fun(comparison.T, comparison.M, cond1, cond2)

```
<br><br><br>

```{r echo = F, fig.height=9, fig.width=12, fig.align='left', fig7, fig.cap = 'G85R control relative to G85R TKT-Df. Transcriptomic data is from females and metabolomic data is from both sexes. Node and arrow size and color '}

comparison.T = "GRF.CxDf"
cond2 = 'GR.F'
cond1 = "TktDfGR.F"
comparison.M = GR.CxDf

network.fun(comparison.T, comparison.M, cond1, cond2)

```

<br><br><br>

```{r echo = F, fig.height=9, fig.width=12, fig.align='left', fig8, fig.cap = 'G85R control relative to G85R TKT-Df. Transcriptomic data is from males and metabolomic data is from both sexes. Node and arrow size and color '}

comparison.T = "GRM.CxDf"
cond2 = 'GR.M'
cond1 = "TktDfGR.M"
comparison.M = GR.CxDf

network.fun(comparison.T, comparison.M, cond1, cond2)

```

<br><br><br>

```{r echo = F, fig.height=9, fig.width=12, fig.align='left', fig9, fig.cap = 'WT control relative to WT TKT-Df. Transcriptomic data is from females and metabolomic data is from both sexes. Node and arrow size and color '}

comparison.T = "WTF.CxDf"
cond2 = 'WT.F'
cond1 = "TktDfWT.F"
comparison.M = WT.CxDf

network.fun(comparison.T, comparison.M, cond1, cond2)

```

<br><br><br>

```{r echo = F, fig.height=9, fig.width=12, fig.align='left', fig10, fig.cap = 'WT control relative to WT TKT-Df. Transcriptomic data is from males and metabolomic data is from both sexes. Node and arrow size and color '}

comparison.T = "WTM.CxDf"
cond2 = 'WT.M'
cond1 = "TktDfWT.M"
comparison.M = WT.CxDf

network.fun(comparison.T, comparison.M, cond1, cond2)

```

<br><br><br>

```{r echo = F, fig.height=9, fig.width=12, fig.align='left', fig11, fig.cap = 'G85R control relative to G85R TKT-OE. Transcriptomic data is from females and metabolomic data is from both sexes. Node and arrow size and color '}

comparison.T = "GRF.CxOE"
cond2 = 'GR.F'
cond1 = "TktOEGR.F"
comparison.M = GR.CxOE

network.fun(comparison.T, comparison.M, cond1, cond2)

```

<br><br><br>

```{r echo = F, fig.height=9, fig.width=12, fig.align='left', fig12, fig.cap = 'G85R control relative to G85R TKT-OE. Transcriptomic data is from males and metabolomic data is from both sexes. Node and arrow size and color '}

comparison.T = "GRM.CxOE"
cond2 = 'GR.M'
cond1 = "TktOEGR.M"
comparison.M = GR.CxOE

network.fun(comparison.T, comparison.M, cond1, cond2)

```

<br><br><br>

```{r echo = F, fig.height=9, fig.width=12, fig.align='left', fig13, fig.cap = 'WT control relative to WT TKT-OE. Transcriptomic data is from females and metabolomic data is from both sexes. Node and arrow size and color '}

comparison.T = "WTF.CxOE"
cond2 = 'WT.F'
cond1 = "TktOEWT.F"
comparison.M = WT.CxOE

network.fun(comparison.T, comparison.M, cond1, cond2)

```

<br><br><br>

```{r echo = F, fig.height=9, fig.width=12, fig.align='left', fig14, fig.cap = 'WT control relative to WT TKT-Df. Transcriptomic data is from males and metabolomic data is from both sexes. Node and arrow size and color '}

comparison.T = "WTM.CxOE"
cond2 = 'WT.M'
cond1 = "TktOEWT.M"
comparison.M = WT.CxOE

network.fun(comparison.T, comparison.M, cond1, cond2)

```