### 2024-01-17 mtDNA-nucDNA analysis/20231214_John_WT.GR_RpL32.mtCoI.Rmd

---
title: "John CoI and RpL32 in G85R and WT Analysis"
author: "John Santiago"
date: "2023-12-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, include = F}
library(plotly)
CoI.data = read.csv("/Users/johncsantiago/Documents/20231214_John_WT.GR_RpL32.mtCoI.csv")
CoI.data$Genotype = substr(CoI.data$Sample,1,2)
CoI.data$Sex = substr(CoI.data$Sample,3,3)
CoI.data$Rep = substr(CoI.data$Sample,4,nchar(CoI.data$Sample))
CoI.data$Primer = CoI.data$Target

CoI.data = CoI.data[CoI.data$Well != 'F1',]

```

```{r echo = F}

plotdata = CoI.data[CoI.data$Target == "RpL32" & nchar(CoI.data$Rep) == 1,]

plotdata$x = factor(paste0(plotdata$Genotype, plotdata$Sex, plotdata$Rep), 
                    levels = c(unique(paste0(plotdata$Genotype, plotdata$Sex, plotdata$Rep))))

plotdata$Cq = as.numeric(plotdata$Cq)

boxplot(plotdata$Cq~plotdata$x,
     ylab = "Quantification Cycle (Cq)",
     xlab = "Genotype",
     main = "RpL32",
     las = 2)

points(x = plotdata$x,
       y = plotdata$Cq,
       cex = 1.5,
       pch = 21,
       bg = "firebrick",
       col = "black")

```

```{r echo = F}

plotdata = CoI.data[CoI.data$Target == "CoI" & nchar(CoI.data$Rep) == 1,]

plotdata$x = factor(paste0(plotdata$Genotype, plotdata$Sex, plotdata$Rep), 
                    levels = c(unique(paste0(plotdata$Genotype, plotdata$Sex, plotdata$Rep))))

plotdata$Cq = as.numeric(plotdata$Cq)

boxplot(plotdata$Cq~plotdata$x,
     ylab = "Quantification Cycle (Cq)",
     xlab = "Genotype",
     main = "CoI",
     las = 2)

points(x = plotdata$x,
       y = plotdata$Cq,
       cex = 1.5,
       pch = 21,
       bg = "steelblue",
       col = "black")

```

```{r echo = F}

plotdata = CoI.data[CoI.data$Genotype == "WT" & 
                      CoI.data$Sex == "M" &
                      CoI.data$Rep != '2' &
                      CoI.data$Rep != '3' &
                      CoI.data$Primer == 'RpL32',]
plotdata$x = factor(plotdata$Sample, 
                    levels = c(unique(plotdata$Sample, plotdata$Primer)))
plotdata$Cq = as.numeric(plotdata$Cq)

boxplot(plotdata$Cq~plotdata$x,
     ylab = "Quantification Cycle (Cq)",
     xlab = "",
     main = "RpL32 Efficiency",
     las = 2,
     cex = .1)

points(x = plotdata$x,
       y = plotdata$Cq,
       cex = 1.5,
       pch = 21,
       bg = "firebrick",
       col = "black")

```

```{r echo = F}

plotdata = CoI.data[CoI.data$Genotype == "WT" & 
                      CoI.data$Sex == "M" &
                      CoI.data$Rep != '2' &
                      CoI.data$Rep != '3' &
                      CoI.data$Primer == 'RpL32',]
plotdata$concentration = plotdata$Rep
plotdata$concentration[plotdata$Rep == 1] = 5
plotdata$concentration[plotdata$Rep == '1_2.5'] = 2.5
plotdata$concentration[plotdata$Rep == '1_10'] = 10

plotdata$x = log2(as.numeric(plotdata$concentration))
plotdata$Cq = as.numeric(plotdata$Cq)

plot(y = plotdata$Cq,
     x = plotdata$x, 
     type = "p",  
     cex = 1, 
     pch = 21, 
     bg = "firebrick",
     col = "black",
     ylab = "Quantification Cycle (Cq)",
     xlab = "log2(cDNA Conc. (ng/uL)")

linearmodel=lm(as.numeric(plotdata$Cq)~plotdata$x)
abline(a=linearmodel[[1]][[1]],b=linearmodel[[1]][2],lty=2,col="firebrick",lwd=2)
text(x = 1.75, y = 20,paste0("RpL32 slope = ", signif(linearmodel[[1]][2],3)),cex=.8)
RpL32.slope = linearmodel[[1]][2]

RpL32.E = (-1+2^(-1/RpL32.slope))*100
text(x = 1.75, y = 19.85,paste0("RpL32 Efficiency = ", signif(RpL32.E,3)),cex=.8)

```

```{r echo = F}

plotdata = CoI.data[CoI.data$Genotype == "WT" & 
                      CoI.data$Sex == "M" &
                      CoI.data$Rep != '2' &
                      CoI.data$Rep != '3' &
                      CoI.data$Primer == 'CoI',]
plotdata$x = factor(plotdata$Sample, 
                    levels = c(unique(plotdata$Sample, plotdata$Primer)))
plotdata$Cq = as.numeric(plotdata$Cq)

boxplot(plotdata$Cq~plotdata$x,
     ylab = "Quantification Cycle (Cq)",
     xlab = "",
     main = "CoI Efficiency",
     las = 2,
     cex = .1)

points(x = plotdata$x,
       y = plotdata$Cq,
       cex = 1.5,
       pch = 21,
       bg = "steelblue",
       col = "black")

```

```{r echo = F}

plotdata = CoI.data[CoI.data$Genotype == "WT" & 
                      CoI.data$Sex == "M" &
                      CoI.data$Rep != '2' &
                      CoI.data$Rep != '3' &
                      CoI.data$Primer == 'CoI',]
plotdata$concentration = plotdata$Rep
plotdata$concentration[plotdata$Rep == 1] = 5
plotdata$concentration[plotdata$Rep == '1_2.5'] = 2.5
plotdata$concentration[plotdata$Rep == '1_10'] = 10

plotdata$x = log2(as.numeric(plotdata$concentration))
plotdata$Cq = as.numeric(plotdata$Cq)

plot(y = plotdata$Cq,
     x = plotdata$x, 
     type = "p",  
     cex = 1, 
     pch = 21, 
     bg = "steelblue",
     col = "black",
     ylab = "Quantification Cycle (Cq)",
     xlab = "log2(cDNA Conc. (ng/uL)")

linearmodel=lm(as.numeric(plotdata$Cq)~plotdata$x)
abline(a=linearmodel[[1]][[1]],b=linearmodel[[1]][2],lty=2,col="steelblue",lwd=2)
text(x = 1.75, y = 13.65,paste0("CoI slope = ", signif(linearmodel[[1]][2],3)),cex=.8)
CoI.slope = linearmodel[[1]][2]

CoI.E = (-1+2^(-1/CoI.slope))*100
text(x = 1.75, y = 13.5,paste0("CoI Efficiency = ", signif(CoI.E,3)),cex=.8)

```

```{r echo = F, include = F}

CoI.only = CoI.data[CoI.data$Primer == "CoI" & nchar(CoI.data$Rep) == 1,
                    c("Sample", "Cq", "Primer")]
RpL32.only = CoI.data[CoI.data$Primer == "RpL32" & nchar(CoI.data$Rep) == 1,
                      c("Sample", "Cq", "Primer")]

plotdata = rbind(CoI.only, RpL32.only)

plotdata$Cq = as.numeric(plotdata$Cq)
plotdata$Primer = as.factor(plotdata$Primer)

plot.colors = c("steelblue", "firebrick")

fig =  plot_ly(data = plotdata,
          x = ~Sample,
          y = ~Cq,
          name = ~Primer,
          type = 'scatter',
          mode = 'markers',
          color = ~Primer,
          colors = plot.colors,
          marker = list(size = 10,
                        line = list(color = 'black', width = .5)))

fig

```

```{r echo = F}

plotdata = CoI.data[nchar(CoI.data$Rep) == 1,
                    c("Sample", "Cq", "Primer")]
plotdata$SP = paste0(plotdata$Sample, plotdata$Primer)
plotdata$Cq = as.numeric(plotdata$Cq)
mean.Ct = data.frame(Sample = unique(plotdata$Sample),
                     Genotype = factor(substr(unique(plotdata$Sample),1,3), 
                                       levels = c('WTF', 'WTM', 'GRF', 'GRM')),
                     Sex = substr(unique(plotdata$Sample), 3,3),
                     CoI = 0,
                     RpL32 = 0)
mean.Ct = mean.Ct[order(mean.Ct$Genotype),]

i=1
while(i <= nrow(mean.Ct)){
  mean.Ct$CoI[i] = mean(plotdata[plotdata$SP == paste0(mean.Ct$Sample[i], 'CoI'),'Cq'])
  mean.Ct$RpL32[i] = mean(plotdata[plotdata$SP == paste0(mean.Ct$Sample[i], 'RpL32'),'Cq'])
  i = i + 1
}

mean.Ct$dCt = mean.Ct$CoI - mean.Ct$RpL32
mean.Ct$FC = 2^-mean.Ct$dCt

boxplot(mean.Ct$dCt~mean.Ct$Genotype,
     ylab = "dCT",
     xlab = "Genotype",
     main = "dCT")

points(x = mean.Ct$Genotype,
       y = mean.Ct$dCt,
       cex = 1.5,
       pch = 21,
       bg = c(rep("firebrick", 6), rep('steelblue', 6)),
       col = "black")

```

```{r echo = F}

boxplot(mean.Ct$FC~mean.Ct$Genotype,
     ylab = "FC",
     xlab = "Genotype",
     main = "FC")

points(x = mean.Ct$Genotype,
       y = mean.Ct$FC,
       cex = 1.5,
       pch = 21,
       bg = c(rep("firebrick", 6), rep('steelblue', 6)),
       col = "black")

```
\newpage

```{r}

dCT.aov = aov(mean.Ct$dCt~mean.Ct$Genotype)

summary(dCT.aov)
```

```{r}

TukeyHSD(dCT.aov)

```

```{r echo = F}

mean.dCt = data.frame(Genotype = unique(mean.Ct$Genotype), 
                      mean.dCt = 0)
mean.dCt$mean.dCt[1] = mean(mean.Ct[mean.Ct$Genotype == mean.dCt$Genotype[1], 'dCt'])
mean.dCt$mean.dCt[2] = mean(mean.Ct[mean.Ct$Genotype == mean.dCt$Genotype[2], 'dCt'])
mean.dCt$mean.dCt[3] = mean(mean.Ct[mean.Ct$Genotype == mean.dCt$Genotype[3], 'dCt'])
mean.dCt$mean.dCt[4] = mean(mean.Ct[mean.Ct$Genotype == mean.dCt$Genotype[4], 'dCt'])
row.names(mean.dCt) = mean.dCt$Genotype

Male_fc = 2^-(mean.dCt['GRM',2] - mean.dCt['WTM',2])
Female_fc = 2^-(mean.dCt['GRF',2] - mean.dCt['WTF',2])

print(paste0("Relative mtDNA levels in Male G85R/WT = ", signif(Male_fc, 3)))

print(paste0("Relative mtDNA levels in Female G85R/WT = ", signif(Female_fc, 3)))

```