### 2024-03-22 A4V BMP Signaling Genes Analysis.Rmd
overdue update

---
title: "BMP Signaling Genes"
author: "John Santiago"
date: "2024-02-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, message = F}

library(heatmaply)
library(dendextend)

counts = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.countdata.csv", row.names = 1)

cpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.cpmdata.csv", row.names = 1)

meancpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.meancpmdata.csv", row.names = 1)

FDR = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FDRdata.csv", row.names = 1)

FC = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FCdata.csv", row.names = 1)

meta = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.metadata.csv", row.names = 1)

git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"
GR.cpm = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
GR.meta  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
GR.FDR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)

bmp = read.csv('/Users/johncsantiago/Documents/GitHub/WhartonLab/BMP Signaling Genes.csv', row.names = 1, header = F)

bmp$fb = row.names(bmp)

bmpcpm = cpm[intersect(row.names(bmp), row.names(cpm)),]

GR.bmp = GR.cpm[intersect(row.names(GR.cpm), row.names(bmp)),]

GR.mean = matrix(rep(0, 12*nrow(GR.bmp)), ncol = 12)
colnames(GR.mean) = unique(GR.meta$Group)
row.names(GR.mean) = row.names(GR.bmp)

i = 1
while(i <= ncol(GR.mean)){
  GR.mean[,i] = apply(GR.bmp[,row.names(GR.meta)[GR.meta$Group == colnames(GR.mean)[i]]], 1, mean)
  i = i +1
}

GR.mean.meta = GR.meta[GR.meta$Replicate == 1, c(1:3,5)]
row.names(GR.mean.meta) = GR.mean.meta$Group

```

```{r echo = F}

hmgroups = meta[meta$Section=='Thorax' & meta$Sex == "F",]
hmgroups$Age[nchar(hmgroups$Age) ==1] = paste0("0",hmgroups$Age[nchar(hmgroups$Age) ==1])
hmgroups$ID = paste0(substr(hmgroups$Genotype,1,1), hmgroups$Age, hmgroups$Sex, "T_rep", hmgroups$Replicate)
hmgroups=hmgroups[order(hmgroups$ID),]
hmdata = bmpcpm[, row.names(hmgroups)]
colnames(hmdata) = hmgroups$ID
Fhmdata=t(scale(t(hmdata)))
Fsidecols=data.frame("Tissue" = "Thorax",
                     "Sex" = hmgroups$Sex,
                     "Genotype" = hmgroups$Genotype,
                     "Age" = hmgroups$Age)
row.names(Fsidecols)=colnames(Fhmdata)

hmgroups = meta[meta$Section=='Thorax' & meta$Sex == "M",]
hmgroups$Age[nchar(hmgroups$Age) ==1] = paste0("0",hmgroups$Age[nchar(hmgroups$Age) ==1])
hmgroups$ID = paste0(substr(hmgroups$Genotype,1,1), hmgroups$Age, hmgroups$Sex, "T_rep", hmgroups$Replicate)
hmgroups=hmgroups[order(hmgroups$ID),]
hmdata = bmpcpm[, row.names(hmgroups)]
colnames(hmdata) = hmgroups$ID
Mhmdata=t(scale(t(hmdata)))
Msidecols=data.frame("Tissue" = "Thorax",
                     "Sex" = hmgroups$Sex,
                     "Genotype" = hmgroups$Genotype,
                     "Age" = hmgroups$Age)
row.names(Msidecols)=colnames(Mhmdata)

Thmdata = cbind(Fhmdata, Mhmdata)
row.names(Thmdata) = bmp[row.names(Thmdata),1]
Tsidecols = rbind(Fsidecols, Msidecols)

hmgroups = meta[meta$Section=='Head' & meta$Sex == "F",]
hmgroups$Age[nchar(hmgroups$Age) ==1] = paste0("0",hmgroups$Age[nchar(hmgroups$Age) ==1])
hmgroups$ID = paste0(substr(hmgroups$Genotype,1,1), hmgroups$Age, hmgroups$Sex, "H_rep", hmgroups$Replicate)
hmgroups=hmgroups[order(hmgroups$ID),]
hmdata = bmpcpm[, row.names(hmgroups)]
colnames(hmdata) = hmgroups$ID
Fhmdata=t(scale(t(hmdata)))
Fsidecols=data.frame("Tissue" = "Head",
                     "Sex" = hmgroups$Sex,
                     "Genotype" = hmgroups$Genotype,
                     "Age" = hmgroups$Age)
row.names(Fsidecols)=colnames(Fhmdata)

hmgroups = meta[meta$Section=='Head' & meta$Sex == "M",]
hmgroups$Age[nchar(hmgroups$Age) ==1] = paste0("0",hmgroups$Age[nchar(hmgroups$Age) ==1])
hmgroups$ID = paste0(substr(hmgroups$Genotype,1,1), hmgroups$Age, hmgroups$Sex, "H_rep", hmgroups$Replicate)
hmgroups=hmgroups[order(hmgroups$ID),]
hmdata = bmpcpm[, row.names(hmgroups)]
colnames(hmdata) = hmgroups$ID
Mhmdata=t(scale(t(hmdata)))
Msidecols=data.frame("Tissue" = "Head",
                     "Sex" = hmgroups$Sex,
                     "Genotype" = hmgroups$Genotype,
                     "Age" = hmgroups$Age)
row.names(Msidecols)=colnames(Mhmdata)

Hhmdata = cbind(Fhmdata, Mhmdata)
row.names(Hhmdata) = bmp[row.names(Hhmdata),1]
Hsidecols = rbind(Fsidecols, Msidecols)

hmgroups = meta[meta$Section=='Abdomen' & meta$Sex == "F",]
hmgroups$Age[nchar(hmgroups$Age) ==1] = paste0("0",hmgroups$Age[nchar(hmgroups$Age) ==1])
hmgroups$ID = paste0(substr(hmgroups$Genotype,1,1), hmgroups$Age, hmgroups$Sex, "A_rep", hmgroups$Replicate)
hmgroups=hmgroups[order(hmgroups$ID),]
hmdata = bmpcpm[, row.names(hmgroups)]
colnames(hmdata) = hmgroups$ID
Fhmdata=t(scale(t(hmdata)))
Fsidecols=data.frame("Tissue" = "Abdomen",
                     "Sex" = hmgroups$Sex,
                     "Genotype" = hmgroups$Genotype,
                     "Age" = hmgroups$Age)
row.names(Fsidecols)=colnames(Fhmdata)

hmgroups = meta[meta$Section=='Abdomen' & meta$Sex == "M",]
hmgroups$Age[nchar(hmgroups$Age) ==1] = paste0("0",hmgroups$Age[nchar(hmgroups$Age) ==1])
hmgroups$ID = paste0(substr(hmgroups$Genotype,1,1), hmgroups$Age, hmgroups$Sex, "A_rep", hmgroups$Replicate)
hmgroups=hmgroups[order(hmgroups$ID),]
hmdata = bmpcpm[, row.names(hmgroups)]
colnames(hmdata) = hmgroups$ID
Mhmdata=t(scale(t(hmdata)))
Msidecols=data.frame("Tissue" = "Abdomen",
                     "Sex" = hmgroups$Sex,
                     "Genotype" = hmgroups$Genotype,
                     "Age" = hmgroups$Age)
row.names(Msidecols)=colnames(Mhmdata)

Ahmdata = cbind(Fhmdata, Mhmdata)
row.names(Ahmdata) = bmp[row.names(Ahmdata),1]
Asidecols = rbind(Fsidecols, Msidecols)

```

```{r echo = F, include = F}

hmdata = cbind(cbind(Hhmdata, Thmdata), Ahmdata)
sidecols = rbind(rbind(Hsidecols, Tsidecols), Asidecols)

heatmaply(hmdata,
          trace="none",
          col=RdYlBu(100)[100:1],
          scale="none",
          dendrogram = "row",
          Rowv=T,
          Colv=F,
          cexRow = .75,
          na.value="black",
          labRow = row.names(hmdata),
          col_side_colors = sidecols)

```

```{r echo = F}

mean.meta = meta[meta$Replicate == '1', c(1:4,6)]
row.names(mean.meta) = mean.meta$Group
bmpmean = meancpm[intersect(row.names(meancpm),row.names(bmp)),]

hmgroups = mean.meta[mean.meta$Section=='Thorax' & mean.meta$Sex == "F",]
hmgroups$Age[nchar(hmgroups$Age) ==1] = paste0("0",hmgroups$Age[nchar(hmgroups$Age) ==1])
hmgroups$ID = paste0(substr(hmgroups$Genotype,1,1), hmgroups$Age, hmgroups$Sex, substr(hmgroups$Section,1,1))
hmgroups=hmgroups[order(hmgroups$ID),]
hmdata = bmpmean[, row.names(hmgroups)]
colnames(hmdata) = hmgroups$ID

TFhmdata=t(scale(t(hmdata)))
TFsidecols=data.frame("Tissue" = "Thorax",
                     "Sex" = hmgroups$Sex,
                     "Genotype" = hmgroups$Genotype,
                     "Age" = hmgroups$Age)
row.names(TFsidecols)=colnames(TFhmdata)

hmgroups = mean.meta[mean.meta$Section=='Thorax' & mean.meta$Sex == "M",]
hmgroups$Age[nchar(hmgroups$Age) ==1] = paste0("0",hmgroups$Age[nchar(hmgroups$Age) ==1])
hmgroups$ID = paste0(substr(hmgroups$Genotype,1,1), hmgroups$Age, hmgroups$Sex, substr(hmgroups$Section,1,1))
hmgroups=hmgroups[order(hmgroups$ID),]
hmdata = bmpmean[, row.names(hmgroups)]
colnames(hmdata) = hmgroups$ID

TMhmdata=t(scale(t(hmdata)))
TMsidecols=data.frame("Tissue" = "Thorax",
                     "Sex" = hmgroups$Sex,
                     "Genotype" = hmgroups$Genotype,
                     "Age" = hmgroups$Age)
row.names(TMsidecols)=colnames(TMhmdata)

Thmdata = cbind(TFhmdata, TMhmdata)
row.names(Thmdata) = bmp[row.names(Thmdata),1]
Tsidecols = rbind(TFsidecols, TMsidecols)

hmgroups = mean.meta[mean.meta$Section=='Head' & mean.meta$Sex == "F",]
hmgroups$Age[nchar(hmgroups$Age) ==1] = paste0("0",hmgroups$Age[nchar(hmgroups$Age) ==1])
hmgroups$ID = paste0(substr(hmgroups$Genotype,1,1), hmgroups$Age, hmgroups$Sex, substr(hmgroups$Section,1,1))
hmgroups=hmgroups[order(hmgroups$ID),]
hmdata = bmpmean[, row.names(hmgroups)]
colnames(hmdata) = hmgroups$ID

HFhmdata=t(scale(t(hmdata)))
HFsidecols=data.frame("Tissue" = "Thorax",
                     "Sex" = hmgroups$Sex,
                     "Genotype" = hmgroups$Genotype,
                     "Age" = hmgroups$Age)
row.names(HFsidecols)=colnames(HFhmdata)

hmgroups = mean.meta[mean.meta$Section=='Head' & mean.meta$Sex == "M",]
hmgroups$Age[nchar(hmgroups$Age) ==1] = paste0("0",hmgroups$Age[nchar(hmgroups$Age) ==1])
hmgroups$ID = paste0(substr(hmgroups$Genotype,1,1), hmgroups$Age, hmgroups$Sex, substr(hmgroups$Section,1,1))
hmgroups=hmgroups[order(hmgroups$ID),]
hmdata = bmpmean[, row.names(hmgroups)]
colnames(hmdata) = hmgroups$ID

HMhmdata=t(scale(t(hmdata)))
HMsidecols=data.frame("Tissue" = "Thorax",
                     "Sex" = hmgroups$Sex,
                     "Genotype" = hmgroups$Genotype,
                     "Age" = hmgroups$Age)
row.names(HMsidecols)=colnames(HMhmdata)

Hhmdata = cbind(HFhmdata, HMhmdata)
row.names(Hhmdata) = bmp[row.names(Hhmdata),1]
Hsidecols = rbind(HFsidecols, HMsidecols)

hmgroups = mean.meta[mean.meta$Section=='Abdomen' & mean.meta$Sex == "F",]
hmgroups$Age[nchar(hmgroups$Age) ==1] = paste0("0",hmgroups$Age[nchar(hmgroups$Age) ==1])
hmgroups$ID = paste0(substr(hmgroups$Genotype,1,1), hmgroups$Age, hmgroups$Sex, substr(hmgroups$Section,1,1))
hmgroups=hmgroups[order(hmgroups$ID),]
hmdata = bmpmean[, row.names(hmgroups)]
colnames(hmdata) = hmgroups$ID

AFhmdata=t(scale(t(hmdata)))
AFsidecols=data.frame("Tissue" = "Thorax",
                     "Sex" = hmgroups$Sex,
                     "Genotype" = hmgroups$Genotype,
                     "Age" = hmgroups$Age)
row.names(AFsidecols)=colnames(AFhmdata)

hmgroups = mean.meta[mean.meta$Section=='Abdomen' & mean.meta$Sex == "M",]
hmgroups$Age[nchar(hmgroups$Age) ==1] = paste0("0",hmgroups$Age[nchar(hmgroups$Age) ==1])
hmgroups$ID = paste0(substr(hmgroups$Genotype,1,1), hmgroups$Age, hmgroups$Sex, substr(hmgroups$Section,1,1))
hmgroups=hmgroups[order(hmgroups$ID),]
hmdata = bmpmean[, row.names(hmgroups)]
colnames(hmdata) = hmgroups$ID

AMhmdata=t(scale(t(hmdata)))
AMsidecols=data.frame("Tissue" = "Thorax",
                     "Sex" = hmgroups$Sex,
                     "Genotype" = hmgroups$Genotype,
                     "Age" = hmgroups$Age)
row.names(AMsidecols)=colnames(AMhmdata)

Ahmdata = cbind(AFhmdata, AMhmdata)
row.names(Ahmdata) = bmp[row.names(Ahmdata),1]
Asidecols = rbind(AFsidecols, AMsidecols)

```

```{r echo = F, include = F}

hmdata = TMhmdata
row.names(hmdata) = bmp[row.names(hmdata),1]
sidecols = TMsidecols

heatmaply(hmdata,
          trace="none",
          col=RdYlBu(100)[100:1],
          scale="none",
          dendrogram = "row",
          Rowv=T,
          Colv=F,
          cexRow = .75,
          na.value="black",
          labRow = row.names(hmdata),
          col_side_colors = sidecols)

```

```{r echo = F}

hmcluster = function(clusternum, allscaled, Title){

dnd=cut(as.dendrogram(hc),h=cutheight)
cluster=row.names(scaled)[unlist(dnd$lower[[clusternum]])]

row.names(allscaled) = bmp[row.names(allscaled), 1]
df = allscaled[cluster[length(cluster):1],] %>%
  reshape2::melt() %>%
  dplyr::rename(row=Var1,col=Var2)

meandf = bmpmean
row.names(meandf) = bmp[row.names(meandf),1]
meandf = meandf[cluster[length(cluster):1],]
temp = hm.meta[hm.meta$Replicate == 1,]
row.names(temp) = temp$Group
colnames(meandf) = temp[colnames(meandf), 'ID']
meandf =  as.matrix(meandf[, colnames(allscaled)])
meandf = meandf %>%
  reshape2::melt() %>%
  dplyr::rename(row=Var1,col=Var2)

rep1df = bmpcpm
row.names(rep1df) = bmp[row.names(rep1df),1]
rep1df = rep1df[cluster[length(cluster):1], row.names(hm.meta)[hm.meta$Replicate == 1]]
colnames(rep1df) = hm.meta[colnames(rep1df), 'ID']
rep1df =  as.matrix(rep1df[, colnames(allscaled)])
rep1df = rep1df %>%
  reshape2::melt() %>%
  dplyr::rename(row=Var1,col=Var2)

rep2df = bmpcpm
row.names(rep2df) = bmp[row.names(rep2df),1]
rep2df = rep2df[cluster[length(cluster):1], row.names(hm.meta)[hm.meta$Replicate == 2]]
colnames(rep2df) = hm.meta[colnames(rep2df), 'ID']
rep2df =  as.matrix(rep2df[, colnames(allscaled)])
rep2df = rep2df %>%
  reshape2::melt() %>%
  dplyr::rename(row=Var1,col=Var2)

rep3df = bmpcpm
row.names(rep3df) = bmp[row.names(rep3df),1]
rep3df = rep3df[cluster[length(cluster):1], row.names(hm.meta)[hm.meta$Replicate == 3]]
colnames(rep3df) = hm.meta[colnames(rep3df), 'ID']
rep3df =  as.matrix(rep3df[, colnames(allscaled)])
rep3df = rep3df %>%
  reshape2::melt() %>%
  dplyr::rename(row=Var1,col=Var2)

FBgn = bmp
row.names(FBgn) = FBgn[,1]

df$mean = meandf$value
df$rep1 = rep1df$value
df$rep2 = rep2df$value
df$rep3 = rep3df$value
df$fbgn = FBgn[as.character(df$row), 'fb']

fig = plot_ly(x = df$col, 
              y = df$row,
              z = df$value,
              type = "heatmap",
              colors = RdYlBu(100)[100:1],
              hoverinfo = "text",
              hovertext = paste("Sample:", df$col,
                                "\nGene Symbol: ", df$row,
                                "\nGene FBgn: ", df$fbgn,
                                "\nMean CPM: ", signif(df$mean, 3),
                                "\nRep1 CPM: ", signif(df$rep1, 3),
                                "\nRep2 CPM: ", signif(df$rep2, 3),
                                "\nRep3 CPM: ", signif(df$rep3, 3)))

fig = fig %>% layout(title = paste(Title, " cluster ",clusternum, " genes"),
                     xaxis = list(title = ""),
                     yaxis = list(title = ""),
                     shapes = list(
                       list(
                         type = "rect",
                         fillcolor = 'deeppink',
                         line = list(color = "deeppink", dash = 'none', width = 1),
                         opacity = 1,
                         x0 = -.5,
                         x1 = 2.5,
                         y0 = length(cluster)-.5,
                         y1 = length(cluster)+1,
                         textposition = "inside",
                         textinfo = "label"),
                       list(
                         type = "rect",
                         fillcolor = 'khaki',
                         line = list(color = "khaki", dash = 'none', width = 1),
                         opacity = 1,
                         x0 = 2.5,
                         x1 = 5.5,
                         y0 = length(cluster)-.5,
                         y1 = length(cluster)+1),
                       list(
                         type = "rect",
                         fillcolor = 'deepskyblue',
                         line = list(color = "deepskyblue", dash = 'none', width = 1),
                         opacity = 1,
                         x0 = 5.5,
                         x1 = 7.5,
                         y0 = length(cluster)-.5,
                         y1 = length(cluster)+1),
                       list(
                         type = "rect",
                         fillcolor = 'slategray',
                         line = list(color = "slategray", dash = 'none', width = 1),
                         opacity = 1,
                         x0 = 7.5,
                         x1 = 9.5,
                         y0 = length(cluster)-.5,
                         y1 = length(cluster)+1),
                       list(
                         type = "rect",
                         line = list(color = "black", dash = 'none', width = 6),
                         opacity = 1,
                         x0 = -.5,
                         x1 = 9.5,
                         y0 = -.5,
                         y1 = length(cluster)+1),
                       list(
                         type = 'line',
                         line = list(color = "black", dash = 'none', width = 6),
                         x0 = -.5,
                         x1 = 9.5,
                         y0 = length(cluster)-.5,
                         y1 = length(cluster)-.5),
                       list(
                         type = 'line',
                         line = list(color = "black", dash = 'none', width = 6),
                         x0 = 2.5,
                         x1 = 2.5,
                         y0 = -.5,
                         y1 = length(cluster)+1),
                       list(
                         type = 'line',
                         line = list(color = "black", dash = 'none', width = 6),
                         x0 = 7.5,
                         x1 = 7.5,
                         y0 = -.5,
                         y1 = length(cluster)+1),
                       list(
                         type = 'line',
                         line = list(color = "black", dash = 'none', width = 6),
                         x0 = 5.5,
                         x1 = 5.5,
                         y0 = -.5,
                         y1 = length(cluster)+1)))

fig = fig %>% add_annotations(x = c(1, 4, 6.5, 8.5), 
                              y = length(cluster)+.25, 
                              text = c("A4V Females", 
                                       "Silent Females", 
                                       "A4V Males", 
                                       "Silent Males"), 
                              font = list(color = '#000000', size = 14), 
                              showarrow = F)

return(fig)

}

```

```{r echo = F, fig.height=10}

scaled = HFhmdata
allscaled = cbind(HFhmdata, HMhmdata)
Title = "Female Head"
branches=3

hm.meta = meta
hm.meta$Age[nchar(hm.meta$Age) ==1] = paste0("0",hm.meta$Age[nchar(hm.meta$Age) ==1])
hm.meta$ID = paste0(substr(hm.meta$Genotype,1,1), hm.meta$Age, hm.meta$Sex, substr(hm.meta$Section,1,1))

row.names(scaled) = bmp[row.names(scaled),1]

allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')

colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = Title)

```

```{r echo = F}

hmcluster(clusternum = 1, allscaled, Title)

```

```{r echo = F}

hmcluster(clusternum = 2, allscaled, Title)

```

```{r echo = F}

hmcluster(clusternum = 3, allscaled, Title)

```

```{r echo = F, fig.height=10}

scaled = HMhmdata
Title = "Male Head"
branches=3

row.names(scaled) = bmp[row.names(scaled),1]

colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = Title)

```

```{r echo = F}

hmcluster(clusternum = 1, allscaled, Title)

```

```{r echo = F}

hmcluster(clusternum = 2, allscaled, Title)

```

```{r echo = F}

hmcluster(clusternum = 3, allscaled, Title)

```

```{r echo = F, fig.height=10}

scaled = TFhmdata
allscaled = cbind(TFhmdata, TMhmdata)
Title = "Female Thorax"
branches=3

row.names(scaled) = bmp[row.names(scaled),1]

colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = Title)

```

```{r echo = F}

hmcluster(clusternum = 1, allscaled, Title)

```

```{r echo = F}

hmcluster(clusternum = 2, allscaled, Title)

```

```{r echo = F}

hmcluster(clusternum = 3, allscaled, Title)

```

```{r echo = F, fig.height=10}

scaled = TMhmdata
Title = "Male Thorax"
branches=3

row.names(scaled) = bmp[row.names(scaled),1]

colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = Title)

```

```{r echo = F}

hmcluster(clusternum = 1, allscaled, Title)

```

```{r echo = F}

hmcluster(clusternum = 2, allscaled, Title)

```

```{r echo = F}

hmcluster(clusternum = 3, allscaled, Title)

```

```{r echo = F, fig.height=10}

scaled = AFhmdata
allscaled = cbind(AFhmdata, AMhmdata)
Title = "Female Abdomen"
branches=3

row.names(scaled) = bmp[row.names(scaled),1]

colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = Title)

```

```{r echo = F}

hmcluster(clusternum = 1, allscaled, Title)

```

```{r echo = F}

hmcluster(clusternum = 2, allscaled, Title)

```

```{r echo = F}

hmcluster(clusternum = 3, allscaled, Title)

```

```{r echo = F, fig.height=10}

scaled = AMhmdata
Title = "Male Abdomen"
branches=3

row.names(scaled) = bmp[row.names(scaled),1]

colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = Title)

```

```{r echo = F}

hmcluster(clusternum = 1, allscaled, Title)

```

```{r echo = F}

hmcluster(clusternum = 2, allscaled, Title)

```

```{r echo = F}

hmcluster(clusternum = 3, allscaled, Title)

```

```{r echo = F, fig.height=10}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('GR.M', 'TktDfGR.M', 'TktOEGR.M',
                      'WT.M', 'TktDfWT.M', 'TktOEWT.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('GR.F', 'TktDfGR.F', 'TktOEGR.F',
                      'WT.F', 'TktDfWT.F', 'TktOEWT.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

allscaled = cbind(Fhmdata,Mhmdata)

Fsidecols=data.frame("Sex" = hmgroups$Sex,
                     "Genotype" = hmgroups$Genotype)
row.names(Fsidecols)=colnames(Fhmdata)

scaled = Fhmdata
Title = "Female Larvae"
branches=3

row.names(scaled) = bmp[row.names(scaled),1]

colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = Title)

```

```{r echo = F}

GR.hmcluster = function(clusternum, allscaled, Title){
dnd=cut(as.dendrogram(hc),h=cutheight)
cluster=row.names(scaled)[unlist(dnd$lower[[clusternum]])]

row.names(allscaled) = bmp[row.names(allscaled), 1]
df = allscaled[cluster[length(cluster):1],] %>%
  reshape2::melt() %>%
  dplyr::rename(row=Var1,col=Var2)

meandf = GR.mean
row.names(meandf) = bmp[row.names(meandf),1]
meandf = meandf[cluster[length(cluster):1],]
meandf =  as.matrix(meandf[, colnames(allscaled)])
meandf = meandf %>%
  reshape2::melt() %>%
  dplyr::rename(row=Var1,col=Var2)

FBgn = bmp
row.names(FBgn) = FBgn[,1]

df$mean = meandf$value
df$fbgn = FBgn[as.character(df$row), 'fb']

fig = plot_ly(x = df$col, 
              y = df$row,
              z = df$value,
              type = "heatmap",
              colors = RdYlBu(100)[100:1],
              hoverinfo = "text",
              hovertext = paste("Sample:", df$col,
                                "\nGene Symbol: ", df$row,
                                "\nGene FBgn: ", df$fbgn,
                                "\nMean CPM: ", signif(df$mean, 3)))

fig = fig %>% layout(title = paste(Title, " cluster ",clusternum, " genes"),
                     xaxis = list(title = ""),
                     yaxis = list(title = ""),
                     shapes = list(
                       list(
                         type = "rect",
                         fillcolor = 'deeppink',
                         line = list(color = "deeppink", dash = 'none', width = 1),
                         opacity = 1,
                         x0 = -.5,
                         x1 = 2.5,
                         y0 = length(cluster)-.5,
                         y1 = length(cluster)+1),
                       list(
                         type = "rect",
                         fillcolor = 'khaki',
                         line = list(color = "khaki", dash = 'none', width = 1),
                         opacity = 1,
                         x0 = 2.5,
                         x1 = 5.5,
                         y0 = length(cluster)-.5,
                         y1 = length(cluster)+1),
                       list(
                         type = "rect",
                         fillcolor = 'deepskyblue',
                         line = list(color = "deepskyblue", dash = 'none', width = 1),
                         opacity = 1,
                         x0 = 5.5,
                         x1 = 8.5,
                         y0 = length(cluster)-.5,
                         y1 = length(cluster)+1),
                       list(
                         type = "rect",
                         fillcolor = 'slategray',
                         line = list(color = "slategray", dash = 'none', width = 1),
                         opacity = 1,
                         x0 = 8.5,
                         x1 = 11.5,
                         y0 = length(cluster)-.5,
                         y1 = length(cluster)+1),
                       list(
                         type = "rect",
                         line = list(color = "black", dash = 'none', width = 6),
                         opacity = 1,
                         x0 = -.5,
                         x1 = 11.5,
                         y0 = -.5,
                         y1 = length(cluster)+1),
                       list(
                         type = 'line',
                         line = list(color = "black", dash = 'none', width = 6),
                         x0 = -.5,
                         x1 = 11.5,
                         y0 = length(cluster)-.5,
                         y1 = length(cluster)-.5),
                       list(
                         type = 'line',
                         line = list(color = "black", dash = 'none', width = 6),
                         x0 = 2.5,
                         x1 = 2.5,
                         y0 = -.5,
                         y1 = length(cluster)+1),
                       list(
                         type = 'line',
                         line = list(color = "black", dash = 'none', width = 6),
                         x0 = 8.5,
                         x1 = 8.5,
                         y0 = -.5,
                         y1 = length(cluster)+1),
                       list(
                         type = 'line',
                         line = list(color = "black", dash = 'none', width = 6),
                         x0 = 5.5,
                         x1 = 5.5,
                         y0 = -.5,
                         y1 = length(cluster)+1)))

fig = fig %>% add_annotations(x = c(1, 4, 7, 10), 
                              y = length(cluster)+.25, 
                              text = c("G85R Females", 
                                       "Silent Females", 
                                       "G85R Males", 
                                       "Silent Males"), 
                              font = list(color = '#000000', size = 14), 
                              showarrow = F)

return(fig)

}

```

```{r echo = F}

GR.hmcluster(clusternum = 1, allscaled, Title)

```

```{r echo = F}

GR.hmcluster(clusternum = 2, allscaled, Title)

```

```{r echo = F}

GR.hmcluster(clusternum = 3, allscaled, Title)

```

```{r echo = F, fig.height=10}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('GR.F', 'TktDfGR.F', 'TktOEGR.F',
                      'WT.F', 'TktDfWT.F', 'TktOEWT.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('GR.M', 'TktDfGR.M', 'TktOEGR.M',
                      'WT.M', 'TktDfWT.M', 'TktOEWT.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

allscaled = cbind(Fhmdata,Mhmdata)

Msidecols=data.frame("Sex" = hmgroups$Sex,
                     "Genotype" = hmgroups$Genotype)
row.names(Msidecols)=colnames(Mhmdata)

scaled = Mhmdata
Title = "Male Larvae"
branches=3

row.names(scaled) = bmp[row.names(scaled),1]

colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = Title)

```

```{r echo = F}

GR.hmcluster(clusternum = 1, allscaled, Title)

```

```{r echo = F}

GR.hmcluster(clusternum = 2, allscaled, Title)

```

```{r echo = F}

GR.hmcluster(clusternum = 3, allscaled, Title)

```

##Heatmap of Fold Changes
```{r echo = F, fig.height=10}

hmdata = FC[intersect(row.names(FC), row.names(bmp)), c("AvS3FH", "AvS9FH", "AvS40FH",
                                                        "AvS3MH", "AvS9MH",
                                                        "AvS3FT", "AvS9FT", "AvS40FT",
                                                        "AvS3MT", "AvS9MT",
                                                        "AvS3FA", "AvS9FA", "AvS40FA",
                                                        "AvS3MA", "AvS9MA")]

hmFDR = FDR[row.names(hmdata), colnames(hmdata)]

colnames(hmdata) = c('A3FH', 'A9FH', 'A40FH',
                     'A3MH', 'A9MH',
                     'A3FT', 'A9FT', 'A40FT',
                     'A3MT', 'A9MT',
                     'A3FA', 'A9FA', 'A40FA',
                     'A3MA', 'A9MA')
scaled=t(scale(t(hmdata)))

Title = "Fold Change Data"
branches=6

row.names(scaled) = bmp[row.names(scaled),1]

colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

scaled[hmFDR>.05] = NA

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = Title)

```