### 2024-03-22 A4V_BodySections_RNAseq/R code/A4V_DEG_Analyses.Rmd

---
title: "A4V PCA"
author: "John Santiago"
date: "2024-01-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=F, include=F}

library(plotly)
library(edgeR)
library(goseq)
library(org.Dm.eg.db)
library(rrvgo)
library(VennDiagram)

```

```{r}

cpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.cpmdata.csv", row.names = 1)

meancpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.meancpmdata.csv", row.names = 1)

FDR = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FDRdata.csv", row.names = 1)

FC = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FCdata.csv", row.names = 1)

meta = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.metadata.csv", row.names = 1)

GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)

```

##GO and KEGG Enrichment function
```{r echo = F, include = F}

Enrichment = function(sigs, bg){
  genes = setNames(rep(0, length(bg)), bg)
  genes[sigs] = 1
  pwf=nullp(genes,"dm3","geneSymbol")
  GO.wall=goseq(pwf,"dm3","geneSymbol")
  GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
  row.names(GO.wall) = GO.wall$category
  KEGG=goseq(pwf,gene2cat=kegg)
  KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
  row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
  KEGG$Name=kegg.names[row.names(KEGG),1]

  enrichment.data = list(GO.wall, KEGG)
  names(enrichment.data) = c("GO", "KEGG")
  return(enrichment.data)
}

```

##GO.Sim.PCA function
```{r echo = F, include = F}
GO.Sim.PCA = function(use.data){
    pca <- prcomp(t(use.data), scale.=TRUE) 
    pc1=scale(pca$x[,1])
    pc2=scale(pca$x[,2])
    eigs <- pca$sdev^2
    ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(PC1=pc1,
                        PC2=pc2,
                        sdev=pca$sdev,
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group = reducedTerms[row.names(use.data), "parentTerm"],
                        Score = reducedTerms[row.names(use.data), "score"],
                        Term = reducedTerms[row.names(use.data), "term"],
                        TotalDE = GO.data[row.names(use.data), "numDEInCat"],
                        Total = GO.data[row.names(use.data), "numInCat"],
                        SigVal = signif(GO.data[row.names(use.data), "adjp"],3))

    pca.data$Group = factor(pca.data$Group, levels = unique(pca.data$Group))
    pca.data$Score = 20*(1+pca.data$Score/mean(pca.data$Score))

  fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 ##symbol = ~Group,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c("firebrick","gold","deepskyblue"),
                 marker = list(size = pca.data$Score,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Parent Group: ", pca.data$Group,
                                   "\nSample: ", pca.data$Term, 
                                   "\nTotal DE: ", pca.data$TotalDE, 
                                   "\nTotal in Cat.: ", pca.data$Total, 
                                   "\nFDR: ", pca.data$SigVal)) %>%
    layout(xaxis = list(title = paste0("PC1", "(", ve[1], "%)")),
           yaxis = list(title = paste0("PC2", "(", ve[2], "%)")),
           title = "Enriched GO term similarity")

  return(fig)
}

```

##GO20 Function
```{r echo = F, include =F}

GO20 = function(GO.data, GO.ontology){
GO.data = GO.data[GO.data$ontology == GO.ontology,
                  c("category", "term","adjp", "numDEInCat", "numInCat")]

data = GO.data[20:1,c("term","adjp","numDEInCat", "numInCat", "category")]
data$Score = -log10(data$adjp)
data$term = factor(data$term, levels = c(unique(data$term)))

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

fig <- plot_ly(data, 
               x = ~numDEInCat, y = ~term, 
               type = 'bar',
               name = "term",
               marker = list(colorscale = list(c(0,.5,1), c("brown","gold", "steelblue")),
    reversescale =T,
                             colorbar = list(title = "-log10(FDR)",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                             color = ~Score,
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$term,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = "Top 20 Most Enriched GO Terms",
           margin = m)

  return(fig)
}

```

##KEGG20 Function
```{r, echo = F, include = F}

KEGG20 = function(data){

  data = data[1:20, c("category", "Name","adjp", "numDEInCat", "numInCat")]

data$term = data$Name
if(length(unique(data$term)) < length(data$term)){
  data$term[duplicated(data$term)] = paste(data$term[duplicated(data$term)], "2")
}  
data = data[20:1,]
data$Score = -log10(data$adjp)
data$term = factor(data$term, levels = c(unique(data$term)))

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

fig <- plot_ly(data, 
               x = ~numDEInCat, y = ~term, 
               type = 'bar',
               name = "Name",
               marker = list(colorscale = list(c(0,.5,1), c("brown","gold", "steelblue")),
    reversescale =T,
                             colorbar = list(title = "-log10(FDR)",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                             color = ~Score,
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$Name,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = "Top 20 Most Enriched KEGG Terms",
           margin = m)

return(fig)

}

```

```{r echo = F}

sigsA3FvMT = row.names(FDR)[FDR[,'A3FvMT']<.05]
sigsS3FvMT = row.names(FDR)[FDR[,'S3FvMT']<.05]
sigsA9FvMT = row.names(FDR)[FDR[,'A9FvMT']<.05]
sigsS9FvMT = row.names(FDR)[FDR[,'S9FvMT']<.05]

GR.cpm = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
GR.meta  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
GR.FDR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)

sigsGR.FxM = row.names(GR.FDR)[GR.FDR$GR.FxM <.05]
sigsWT.FxM = row.names(GR.FDR)[GR.FDR$WT.FxM <.05]

A3 = setdiff(sigsA3FvMT, sigsS3FvMT)
A9 = setdiff(sigsA9FvMT, sigsS9FvMT)
GR = setdiff(sigsGR.FxM, sigsWT.FxM)

##Sample titles as strings. Only fill in up to your number of selected categories
set1="A4V 3-day"
set2="A4V 9-day"
set3="G85R Larvae"
set4=""

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1=A3
s2=A9
s3=GR
s4=""

T3 = A3
T9 = A9

```

```{r}

##Triple Venn
grid.newpage()
draw.triple.venn(area1=length(s1),area2=length(s2),area3=length(s3),n12=length(intersect(s1,s2)),n23 = length(intersect(s2,s3)),n13 = length(intersect(s1,s3)),n123 = length(intersect(intersect(s1,s2),s3)),category=c(set1,set2,set3),fill = c("green", "blue","red"),cex=1.5,cat.cex = 1)

```

```{r echo = F}

sigsA3FvMH = row.names(FDR)[FDR[,'A3FvMH']<.05]
sigsS3FvMH = row.names(FDR)[FDR[,'S3FvMH']<.05]
sigsA9FvMH = row.names(FDR)[FDR[,'A9FvMH']<.05]
sigsS9FvMH = row.names(FDR)[FDR[,'S9FvMH']<.05]

GR.cpm = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
GR.meta  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
GR.FDR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)

sigsGR.FxM = row.names(GR.FDR)[GR.FDR$GR.FxM <.05]
sigsWT.FxM = row.names(GR.FDR)[GR.FDR$WT.FxM <.05]

A3 = setdiff(sigsA3FvMH, sigsS3FvMH)
A9 = setdiff(sigsA9FvMH, sigsS9FvMH)
GR = setdiff(sigsGR.FxM, sigsWT.FxM)

##Sample titles as strings. Only fill in up to your number of selected categories
set1="A4V 3-day"
set2="A4V 9-day"
set3="G85R Larvae"
set4=""

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1=A3
s2=A9
s3=GR
s4=""

H3=A3
H9=A9

```

```{r}

##Triple Venn
grid.newpage()
draw.triple.venn(area1=length(s1),area2=length(s2),area3=length(s3),n12=length(intersect(s1,s2)),n23 = length(intersect(s2,s3)),n13 = length(intersect(s1,s3)),n123 = length(intersect(intersect(s1,s2),s3)),category=c(set1,set2,set3),fill = c("green", "blue","red"),cex=1.5,cat.cex = 1)

```

```{r echo = F}

sigsA3FvMA = row.names(FDR)[FDR[,'A3FvMA']<.05]
sigsS3FvMA = row.names(FDR)[FDR[,'S3FvMA']<.05]
sigsA9FvMA = row.names(FDR)[FDR[,'A9FvMA']<.05]
sigsS9FvMA = row.names(FDR)[FDR[,'S9FvMA']<.05]

GR.cpm = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
GR.meta  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
GR.FDR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)

sigsGR.FxM = row.names(GR.FDR)[GR.FDR$GR.FxM <.05]
sigsWT.FxM = row.names(GR.FDR)[GR.FDR$WT.FxM <.05]

A3 = setdiff(sigsA3FvMA, sigsS3FvMA)
A9 = setdiff(sigsA9FvMA, sigsS9FvMA)
GR = setdiff(sigsGR.FxM, sigsWT.FxM)

##Sample titles as strings. Only fill in up to your number of selected categories
set1="A4V 3-day"
set2="A4V 9-day"
set3="G85R Larvae"
set4=""

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1=A3
s2=A9
s3=GR
s4=""

A3=A3
A9=A9

```

```{r}

##Triple Venn
grid.newpage()
draw.triple.venn(area1=length(s1),area2=length(s2),area3=length(s3),n12=length(intersect(s1,s2)),n23 = length(intersect(s2,s3)),n13 = length(intersect(s1,s3)),n123 = length(intersect(intersect(s1,s2),s3)),category=c(set1,set2,set3),fill = c("green", "blue","red"),cex=1.5,cat.cex = 1)

```

```{r echo = F, include = F}

sigs = setdiff(setdiff(T9,T3),GR)
bg = row.names(FDR)

sigsKEY = GeneIDKey[sigs,]
sigs = na.omit(sigsKEY$Symbol)
bgKEY = GeneIDKey[bg,]
bg = na.omit(bgKEY$Symbol)

sig.cats = Enrichment(sigs, bg)

```

```{r echo =F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "BP"
go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .005 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

reducedTerms <- reduceSimMatrix(simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r echo = F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "BP"
BP.GO.fig = GO20(GO.data, GO.ontology)

GO.ontology = "CC"
CC.GO.fig = GO20(GO.data, GO.ontology)

GO.ontology = "MF"
MF.GO.fig = GO20(GO.data, GO.ontology)

KEGG.fig = KEGG20(sig.cats[["KEGG"]])

```

```{r, fig.width= 10}

BP.GO.fig

```

```{r, fig.width= 10}

CC.GO.fig

```

```{r, fig.width= 10}

MF.GO.fig

```

```{r, fig.width= 10}

KEGG.fig

```

```{r fig.height=7, fig.width=10}

heatmapPlot(simMatrix,
            reducedTerms,
            annotateParent=TRUE,
            annotationLabel="parentTerm",
            fontsize=6,
            main = "Repressor Gene Enriched Biological Process GO terms")

```

```{r warning=F, fig.width=10}

use.data   = simMatrix
GO.Sim.PCA(use.data)

```

```{r fig.align='left'}

treemapPlot(reducedTerms)

```

```{r echo=F, include=F}

pca <- prcomp(t(A4V.cpm), scale.=TRUE) 
gr <- colnames(A4V.cpm)
PC1=scale(pca$x[,1])
PC2=scale(pca$x[,2])
eigs <- pca$sdev^2
ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(Sample=gr,
                        ##PC1=scale(pca$x[,1]),
                        ##PC2=scale(pca$x[,2]),
                        PC1=scale(pca$x[,1]),
                        PC2=scale(pca$x[,2]),
                        sdev=pca$sdev,
                        Genotype=substr(gr, 1, 1), 
                        Sex=substr(gr, nchar(gr)-2, nchar(gr)-2), 
                        Age=substr(gr, 2, nchar(gr)-3),
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group= paste0(substr(gr, 1, 1),
                                      substr(gr, nchar(gr)-2, nchar(gr)-1)),
                        pca.color = substr(gr, nchar(gr)-1, nchar(gr)-1),
                        pca.shape = substr(gr, nchar(gr)-2, nchar(gr)-2))

fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~pca.color,
                 symbol = ~pca.shape,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c(input$color1, input$color2, input$color3)[c(1:length(levels(pca.color)))],
                 marker = list(size = 15,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Sample)) %>%
    layout(xaxis = list(title = paste0("PC 1", "(", pca.data$Percent1[1], "%)")),
           yaxis = list(title = paste0("PC 2", "(", pca.data$Percent2[2], "%)")))
```

```{r}

  fig

```

```{r echo=F, include=F}

pca.data = cpmdata[,substr(colnames(cpmdata), nchar(colnames(cpmdata))-1, nchar(colnames(cpmdata))-1) == "H"]

remove = apply(pca.data, MARGIN = 1, sum) != 0

pca.data = pca.data[remove,]

pca <- prcomp(t(pca.data), scale.=TRUE) 
gr <- colnames(pca.data)
PC1=scale(pca$x[,1])
PC2=scale(pca$x[,2])
eigs <- pca$sdev^2
ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(Sample=gr,
                        ##PC1=scale(pca$x[,1]),
                        ##PC2=scale(pca$x[,2]),
                        PC1=scale(pca$x[,1]),
                        PC2=scale(pca$x[,2]),
                        sdev=pca$sdev,
                        Genotype=substr(gr, 1, 1), 
                        Sex=substr(gr, nchar(gr)-2, nchar(gr)-2), 
                        Age=substr(gr, 2, nchar(gr)-3),
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group= paste0(substr(gr, 1, 1),
                                      substr(gr, nchar(gr)-2, nchar(gr)-1)))

fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Genotype,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c(input$color1, input$color2, input$color3)[c(1:length(levels(pca.color)))],
                 marker = list(size = 15,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Sample)) %>%
    layout(xaxis = list(title = paste0("PC 1", "(", pca.data$Percent1[1], "%)")),
           yaxis = list(title = paste0("PC 2", "(", pca.data$Percent2[2], "%)")))
```

```{r}

  fig

```

```{r echo=F, include=F}

pca.data = cpmdata[,substr(colnames(cpmdata), nchar(colnames(cpmdata))-1, nchar(colnames(cpmdata))-1) == "T"]

remove = apply(pca.data, MARGIN = 1, sum) != 0

pca.data = pca.data[remove,]

pca <- prcomp(t(pca.data), scale.=TRUE) 
gr <- colnames(pca.data)
PC1=scale(pca$x[,1])
PC2=scale(pca$x[,2])
eigs <- pca$sdev^2
ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(Sample=gr,
                        ##PC1=scale(pca$x[,1]),
                        ##PC2=scale(pca$x[,2]),
                        PC1=scale(pca$x[,1]),
                        PC2=scale(pca$x[,2]),
                        sdev=pca$sdev,
                        Genotype=substr(gr, 1, 1), 
                        Sex=substr(gr, nchar(gr)-2, nchar(gr)-2), 
                        Age=substr(gr, 2, nchar(gr)-3),
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group= paste0(substr(gr, 1, 1),
                                      substr(gr, nchar(gr)-2, nchar(gr)-1)))

fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Genotype,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c(input$color1, input$color2, input$color3)[c(1:length(levels(pca.color)))],
                 marker = list(size = 15,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Sample)) %>%
    layout(xaxis = list(title = paste0("PC 1", "(", pca.data$Percent1[1], "%)")),
           yaxis = list(title = paste0("PC 2", "(", pca.data$Percent2[2], "%)")))
```

```{r}

  fig

```

```{r echo=F, include=F}

pca.data = A4V.cpm[,substr(colnames(A4V.cpm), nchar(colnames(A4V.cpm))-1, nchar(colnames(A4V.cpm))-1) == "A"]

remove = apply(pca.data, MARGIN = 1, sum) != 0

pca.data = pca.data[remove,]

pca <- prcomp(t(pca.data), scale.=TRUE) 
gr <- colnames(pca.data)
PC1=scale(pca$x[,1])
PC2=scale(pca$x[,2])
eigs <- pca$sdev^2
ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(Sample=gr,
                        ##PC1=scale(pca$x[,1]),
                        ##PC2=scale(pca$x[,2]),
                        PC1=scale(pca$x[,1]),
                        PC2=scale(pca$x[,2]),
                        sdev=pca$sdev,
                        Genotype=substr(gr, 1, 1), 
                        Sex=substr(gr, nchar(gr)-2, nchar(gr)-2), 
                        Age=substr(gr, 2, nchar(gr)-3),
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group= paste0(substr(gr, 1, 1),
                                      substr(gr, nchar(gr)-2, nchar(gr)-1)))

fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Genotype,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c(input$color1, input$color2, input$color3)[c(1:length(levels(pca.color)))],
                 marker = list(size = 15,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Sample)) %>%
    layout(xaxis = list(title = paste0("PC 1", "(", pca.data$Percent1[1], "%)")),
           yaxis = list(title = paste0("PC 2", "(", pca.data$Percent2[2], "%)")))
```

```{r}

  fig

```