### 2024-03-22 A4V_BodySections_RNAseq/R code/A4V_PCA.Rmd

---
title: "A4V PCA"
author: "John Santiago"
date: "2024-01-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=F, include=F}

library(plotly)
library(edgeR)
library(goseq)
library(org.Dm.eg.db)
library(rrvgo)
library(VennDiagram)

```

```{r include =F}
sample.names = list.files("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/Count_Tables/")

meta = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.metadata.csv", row.names = 1)

temp = read.csv(paste0("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/Count_Tables/", sample.names[1]), sep = '\t', row.names = 1)

countdata = data.frame(temp = temp)
colnames(countdata)[1] = substr(sample.names[1], 1, nchar(sample.names[1])-15)

i=2
while(i <= length(sample.names)){
  temp = read.csv(paste0("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/Count_Tables/", sample.names[i]), sep = '\t', row.names = 1)
  countdata = cbind(countdata, temp[row.names(countdata),1])
  colnames(countdata)[i] = substr(sample.names[i], 1, nchar(sample.names[i])-15)
  i=i+1
}

countdata = countdata[-(grep("ti",row.names(countdata))),]
countdata = countdata[-(grep("trans",row.names(countdata))),]

##normalize data
x <- countdata
gr <- factor(substr(colnames(countdata), 1, nchar(colnames(countdata))-1))
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- normLibSizes(y)
##normalized cpm
cpmdata=cpm(z)

##updated edger analysis pipeline
design<-model.matrix(~0 + gr)
colnames(design) = unique(gr)
z = estimateDisp(z, design)
fit <- glmQLFit(z, design)

contr = makeContrasts(A3FH - S3FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS3FH = DEdata
sigAvS3FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3MH - S3MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS3MH = DEdata
sigAvS3MH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FH - S9FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS9FH = DEdata
sigAvS9FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9MH - S9MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS9MH = DEdata
sigAvS9MH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A40FH - S40FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS40FH = DEdata
sigAvS40FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FH - S9FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v9FH = DEdata
sigS3v9FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FH - A9FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v9FH = DEdata
sigA3v9FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3MH - S9MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v9MH = DEdata
sigS3v9MH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3MH - A9MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v9MH = DEdata
sigA3v9MH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S9FH - S40FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S9v40FH = DEdata
sigS9v40FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FH - A40FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A9v40FH = DEdata
sigA9v40FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FH - S40FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v40FH = DEdata
sigS3v40FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FH - A40FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v40FH = DEdata
sigA3v40FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FT - S3FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS3FT = DEdata
sigAvS3FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3MT - S3MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS3MT = DEdata
sigAvS3MT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FT - S9FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS9FT = DEdata
sigAvS9FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9MT - S9MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS9MT = DEdata
sigAvS9MT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A40FT - S40FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS40FT = DEdata
sigAvS40FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FT - S9FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v9FT = DEdata
sigS3v9FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FT - A9FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v9FT = DEdata
sigA3v9FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3MT - S9MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v9MT = DEdata
sigS3v9MT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3MT - A9MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v9MT = DEdata
sigA3v9MT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S9FT - S40FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S9v40FT = DEdata
sigS9v40FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FT - A40FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A9v40FT = DEdata
sigA9v40FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FT - S40FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v40FT = DEdata
sigS3v40FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FT - A40FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v40FT = DEdata
sigA3v40FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FA - S3FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS3FA = DEdata
sigAvS3FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3MA - S3MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS3MA = DEdata
sigAvS3MA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FA - S9FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS9FA = DEdata
sigAvS9FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9MA - S9MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS9MA = DEdata
sigAvS9MA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A40FA - S40FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS40FA = DEdata
sigAvS40FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FA - S9FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v9FA = DEdata
sigS3v9FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FA - A9FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v9FA = DEdata
sigA3v9FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3MA - S9MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v9MA = DEdata
sigS3v9MA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3MA - A9MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v9MA = DEdata
sigA3v9MA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S9FA - S40FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S9v40FA = DEdata
sigS9v40FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FA - A40FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A9v40FA = DEdata
sigA9v40FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FA - S40FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v40FA = DEdata
sigS3v40FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FA - A40FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v40FA = DEdata
sigA3v40FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FH - A3MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3FvMH = DEdata
sigA3FvMH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FH - S3MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3FvMH = DEdata
sigS3FvMH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FH - A9MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A9FvMH = DEdata
sigA9FvMH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S9FH - S9MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S9FvMH = DEdata
sigS9FvMH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FT - A3MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3FvMT = DEdata
sigA3FvMT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FT - S3MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3FvMT = DEdata
sigS3FvMT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FT - A9MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A9FvMT = DEdata
sigA9FvMT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S9FT - S9MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S9FvMT = DEdata
sigS9FvMT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FA - A3MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3FvMA = DEdata
sigA3FvMA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FA - S3MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3FvMA = DEdata
sigS3FvMA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FA - A9MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A9FvMA = DEdata
sigA9FvMA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S9FA - S9MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S9FvMA = DEdata
sigS9FvMA = DEdata[DEdata$FDR < 0.05, ]

order = row.names(cpmdata)
A4V.FDR = data.frame(AvS3FH = AvS3FH[order, "FDR"],
                     AvS3MH = AvS3MH[order, "FDR"],
                     AvS9FH = AvS9FH[order, "FDR"],
                     AvS9MH = AvS9MH[order, "FDR"],
                     AvS40FH = AvS40FH[order, "FDR"],
                     A3v9FH = A3v9FH[order, "FDR"],
                     S3v9FH = S3v9FH[order, "FDR"],
                     A3v9MH = A3v9MH[order, "FDR"],
                     S3v9MH = S3v9MH[order, "FDR"],
                     A9v40FH = A9v40FH[order, "FDR"],
                     S9v40FH = S9v40FH[order, "FDR"],
                     A3v40FH = A3v40FH[order, "FDR"],
                     S3v40FH = S3v40FH[order, "FDR"],

                     AvS3FT = AvS3FT[order, "FDR"],
                     AvS3MT = AvS3MT[order, "FDR"],
                     AvS9FT = AvS9FT[order, "FDR"],
                     AvS9MT = AvS9MT[order, "FDR"],
                     AvS40FT = AvS40FT[order, "FDR"],
                     A3v9FT = A3v9FT[order, "FDR"],
                     S3v9FT = S3v9FT[order, "FDR"],
                     A3v9MT = A3v9MT[order, "FDR"],
                     S3v9MT = S3v9MT[order, "FDR"],
                     A9v40FT = A9v40FT[order, "FDR"],
                     S9v40FT = S9v40FT[order, "FDR"],
                     A3v40FT = A3v40FT[order, "FDR"],
                     S3v40FT = S3v40FT[order, "FDR"],

                     AvS3FA = AvS3FA[order, "FDR"],
                     AvS3MA = AvS3MA[order, "FDR"],
                     AvS9FA = AvS9FA[order, "FDR"],
                     AvS9MA = AvS9MA[order, "FDR"],
                     AvS40FA = AvS40FA[order, "FDR"],
                     A3v9FA = A3v9FA[order, "FDR"],
                     S3v9FA = S3v9FA[order, "FDR"],
                     A3v9MA = A3v9MA[order, "FDR"],
                     S3v9MA = S3v9MA[order, "FDR"],
                     A9v40FA = A9v40FA[order, "FDR"],
                     S9v40FA = S9v40FA[order, "FDR"],
                     A3v40FA = A3v40FA[order, "FDR"],
                     S3v40FA = S3v40FA[order, "FDR"],

                     A3FvMH = A3FvMH[order, "FDR"],
                     S3FvMH = S3FvMH[order, "FDR"],
                     A9FvMH = A9FvMH[order, "FDR"],
                     S9FvMH = S9FvMH[order, "FDR"],
                     A3FvMT = A3FvMT[order, "FDR"],
                     S3FvMT = S3FvMT[order, "FDR"],
                     A9FvMT = A9FvMT[order, "FDR"],
                     S9FvMT = S9FvMT[order, "FDR"],
                     A3FvMA = A3FvMA[order, "FDR"],
                     S3FvMA = S3FvMA[order, "FDR"],
                     A9FvMA = A9FvMA[order, "FDR"],
                     S9FvMA = S9FvMA[order, "FDR"])
row.names(A4V.FDR) = order

A4V.FC = data.frame(AvS3FH = AvS3FH[order, "logFC"],
                     AvS3MH = AvS3MH[order, "logFC"],
                     AvS9FH = AvS9FH[order, "logFC"],
                     AvS9MH = AvS9MH[order, "logFC"],
                     AvS40FH = AvS40FH[order, "logFC"],
                     A3v9FH = A3v9FH[order, "logFC"],
                     S3v9FH = S3v9FH[order, "logFC"],
                     A3v9MH = A3v9MH[order, "logFC"],
                     S3v9MH = S3v9MH[order, "logFC"],
                     A9v40FH = A9v40FH[order, "logFC"],
                     S9v40FH = S9v40FH[order, "logFC"],
                     A3v40FH = A3v40FH[order, "logFC"],
                     S3v40FH = S3v40FH[order, "logFC"],

                     AvS3FT = AvS3FT[order, "logFC"],
                     AvS3MT = AvS3MT[order, "logFC"],
                     AvS9FT = AvS9FT[order, "logFC"],
                     AvS9MT = AvS9MT[order, "logFC"],
                     AvS40FT = AvS40FT[order, "logFC"],
                     A3v9FT = A3v9FT[order, "logFC"],
                     S3v9FT = S3v9FT[order, "logFC"],
                     A3v9MT = A3v9MT[order, "logFC"],
                     S3v9MT = S3v9MT[order, "logFC"],
                     A9v40FT = A9v40FT[order, "logFC"],
                     S9v40FT = S9v40FT[order, "logFC"],
                     A3v40FT = A3v40FT[order, "logFC"],
                     S3v40FT = S3v40FT[order, "logFC"],

                     AvS3FA = AvS3FA[order, "logFC"],
                     AvS3MA = AvS3MA[order, "logFC"],
                     AvS9FA = AvS9FA[order, "logFC"],
                     AvS9MA = AvS9MA[order, "logFC"],
                     AvS40FA = AvS40FA[order, "logFC"],
                     A3v9FA = A3v9FA[order, "logFC"],
                     S3v9FA = S3v9FA[order, "logFC"],
                     A3v9MA = A3v9MA[order, "logFC"],
                     S3v9MA = S3v9MA[order, "logFC"],
                     A9v40FA = A9v40FA[order, "logFC"],
                     S9v40FA = S9v40FA[order, "logFC"],
                     A3v40FA = A3v40FA[order, "logFC"],
                     S3v40FA = S3v40FA[order, "logFC"],

                     A3FvMH = A3FvMH[order, "logFC"],
                     S3FvMH = S3FvMH[order, "logFC"],
                     A9FvMH = A9FvMH[order, "logFC"],
                     S9FvMH = S9FvMH[order, "logFC"],
                     A3FvMT = A3FvMT[order, "logFC"],
                     S3FvMT = S3FvMT[order, "logFC"],
                     A9FvMT = A9FvMT[order, "logFC"],
                     S9FvMT = S9FvMT[order, "logFC"],
                     A3FvMA = A3FvMA[order, "logFC"],
                     S3FvMA = S3FvMA[order, "logFC"],
                     A9FvMA = A9FvMA[order, "logFC"],
                     S9FvMA = S9FvMA[order, "logFC"])
row.names(A4V.FC) = order

cpmdata = cpmdata[grep("FBgn",row.names(cpmdata)),]
A4V.FDR = A4V.FDR[row.names(cpmdata),]

A4V.meancpm = cpmdata[,1:30]
colnames(A4V.meancpm) = unique(meta$Group)
i=1
while(i<=nrow(A4V.meancpm)){
  j=1
  while(j<=ncol(A4V.meancpm)){
    A4V.meancpm[i,j]=mean(cpmdata[i,row.names(meta)[meta$Group == colnames(A4V.meancpm)[j]]])
    j=j+1
  }
  i=i+1
}

##write.csv(cpmdata, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.cpmdata.csv")
##write.csv(A4V.FDR, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FDRdata.csv")
##write.csv(A4V.meancpm, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.meancpmdata.csv")
##write.csv(A4V.FC, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FCdata.csv")

```

```{r}

cpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.cpmdata.csv", row.names = 1)

meancpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.meancpmdata.csv", row.names = 1)

FDR = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FDRdata.csv", row.names = 1)

FC = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FCdata.csv", row.names = 1)

meta = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.metadata.csv", row.names = 1)

GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)

```

```{r echo = F}

## arc1 FBgn = 'FBgn0033926'
##spn FBgn = 'FBgn0010905'
FBgn = 'FBgn0035348'
name = ''

par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
gene = data.frame(cpm = as.numeric(cpm[FBgn,]), 
                  condition = substr(row.names(meta),1, nchar(row.names(meta))-1), 
                  group = paste(meta$Genotype, meta$Sex))

gene$condition = factor(gene$condition, levels = c("A3FH", "A9FH", "A40FH",
                                                   "S3FH", "S9FH", "S40FH",
                                                   "A3FT", "A9FT", "A40FT",
                                                   "S3FT", "S9FT", "S40FT",
                                                   "A3FA", "A9FA", "A40FA",
                                                   "S3FA", "S9FA", "S40FA",
                                                   "A3MH", "A9MH",
                                                   "S3MH", "S9MH",
                                                   "A3MT", "A9MT",
                                                   "S3MT", "S9MT",
                                                   "A3MA", "A9MA",
                                                   "S3MA", "S9MA"))
gene$group = factor(gene$group, levels = c(unique(gene$group)))

boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"gold","darkgreen","dodgerblue")[as.numeric(factor(gene$group))])
legend('topright',
       inset=c(-0.175,0),
       legend = c('A4V F', 'Silent F', 'A4V M', 'Silent F'), 
       fill = c('firebrick',"darkgreen","gold","dodgerblue"),
       cex = .65,
       bty = 'n',
       pt.cex = .5)

```

##GO and KEGG Enrichment function
```{r echo = F, include = F}

Enrichment = function(sigs, bg){
  genes = setNames(rep(0, length(bg)), bg)
  genes[sigs] = 1
  pwf=nullp(genes,"dm3","geneSymbol")
  GO.wall=goseq(pwf,"dm3","geneSymbol")
  GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
  row.names(GO.wall) = GO.wall$category
  KEGG=goseq(pwf,gene2cat=kegg)
  KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
  row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
  KEGG$Name=kegg.names[row.names(KEGG),1]

  enrichment.data = list(GO.wall, KEGG)
  names(enrichment.data) = c("GO", "KEGG")
  return(enrichment.data)
}

```

##GO.Sim.PCA function
```{r echo = F, include = F}
GO.Sim.PCA = function(use.data){
    pca <- prcomp(t(use.data), scale.=TRUE) 
    pc1=scale(pca$x[,1])
    pc2=scale(pca$x[,2])
    eigs <- pca$sdev^2
    ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(PC1=pc1,
                        PC2=pc2,
                        sdev=pca$sdev,
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group = reducedTerms[row.names(use.data), "parentTerm"],
                        Score = reducedTerms[row.names(use.data), "score"],
                        Term = reducedTerms[row.names(use.data), "term"],
                        TotalDE = GO.data[row.names(use.data), "numDEInCat"],
                        Total = GO.data[row.names(use.data), "numInCat"],
                        SigVal = signif(GO.data[row.names(use.data), "adjp"],3))

    pca.data$Group = factor(pca.data$Group, levels = unique(pca.data$Group))
    pca.data$Score = 20*(1+pca.data$Score/mean(pca.data$Score))

  fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 ##symbol = ~Group,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c("firebrick","gold","deepskyblue"),
                 marker = list(size = pca.data$Score,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Parent Group: ", pca.data$Group,
                                   "\nSample: ", pca.data$Term, 
                                   "\nTotal DE: ", pca.data$TotalDE, 
                                   "\nTotal in Cat.: ", pca.data$Total, 
                                   "\nFDR: ", pca.data$SigVal)) %>%
    layout(xaxis = list(title = paste0("PC1", "(", ve[1], "%)")),
           yaxis = list(title = paste0("PC2", "(", ve[2], "%)")),
           title = "Enriched GO term similarity")

  return(fig)
}

```

##GO20 Function
```{r echo = F, include =F}

GO20 = function(GO.data, GO.ontology){
GO.data = GO.data[GO.data$ontology == GO.ontology,
                  c("category", "term","adjp", "numDEInCat", "numInCat")]

data = GO.data[20:1,c("term","adjp","numDEInCat", "numInCat", "category")]
data$Score = -log10(data$adjp)
data$term = factor(data$term, levels = c(unique(data$term)))

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

fig <- plot_ly(data, 
               x = ~numDEInCat, y = ~term, 
               type = 'bar',
               name = "term",
               marker = list(colorscale = list(c(0,.5,1), c("brown","gold", "steelblue")),
    reversescale =T,
                             colorbar = list(title = "-log10(FDR)",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                             color = ~Score,
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$term,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = "Top 20 Most Enriched GO Terms",
           margin = m)

  return(fig)
}

```

##KEGG20 Function
```{r, echo = F, include = F}

KEGG20 = function(data){

  data = data[1:20, c("category", "Name","adjp", "numDEInCat", "numInCat")]

data$term = data$Name
if(length(unique(data$term)) < length(data$term)){
  data$term[duplicated(data$term)] = paste(data$term[duplicated(data$term)], "2")
}  
data = data[20:1,]
data$Score = -log10(data$adjp)
data$term = factor(data$term, levels = c(unique(data$term)))

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

fig <- plot_ly(data, 
               x = ~numDEInCat, y = ~term, 
               type = 'bar',
               name = "Name",
               marker = list(colorscale = list(c(0,.5,1), c("brown","gold", "steelblue")),
    reversescale =T,
                             colorbar = list(title = "-log10(FDR)",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                             color = ~Score,
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$Name,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = "Top 20 Most Enriched KEGG Terms",
           margin = m)

return(fig)

}

```

```{r echo = F, include = F}

i=1

while(i<=ncol(FDR)){
compare = colnames(FDR)[i]

sigs = row.names(FDR)[FDR[,i]<.05]
bg = row.names(FDR)

sigsKEY = GeneIDKey[sigs,]
sigs = na.omit(sigsKEY$Symbol)
bgKEY = GeneIDKey[bg,]
bg = na.omit(bgKEY$Symbol)

sig.cats = Enrichment(sigs, bg)

##write.csv(sig.cats[[1]], paste0("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/GOterm Enrichment/", colnames(FDR)[i],"_GOTermAnalysis.csv"))

##write.csv(sig.cats[[2]], paste0("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/KEGG Enrichment/", colnames(FDR)[i],"KEGGAnalysis.csv"))

i = i +1

}

```

```{r}

sigsA3FvMT = row.names(FDR)[FDR[,'A3FvMT']<.05]
sigsS3FvMT = row.names(FDR)[FDR[,'S3FvMT']<.05]
sigsA9FvMT = row.names(FDR)[FDR[,'A9FvMT']<.05]
sigsS9FvMT = row.names(FDR)[FDR[,'S9FvMT']<.05]

GR.cpm = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
GR.meta  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
GR.FDR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)

sigsGR.FxM = row.names(GR.FDR)[GR.FDR$GR.FxM <.05]
sigsWT.FxM = row.names(GR.FDR)[GR.FDR$WT.FxM <.05]

A3 = setdiff(sigsA3FvMT, sigsS3FvMT)
A9 = setdiff(sigsA9FvMT, sigsS9FvMT)
GR = setdiff(sigsGR.FxM, sigsWT.FxM)

##Sample titles as strings. Only fill in up to your number of selected categories
set1="A4V 3-day"
set2="A4V 9-day"
set3="G85R Larvae"
set4=""

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1=A3
s2=A9
s3=GR
s4=""

T3 = A3
T9 = A9

```

```{r echo=F, include=F}

FC.subset = FC[, c("AvS3FH", "AvS9FH", "AvS40FH",
                   "AvS3MH", "AvS9MH",
                   "AvS3FT", "AvS9FT", "AvS40FT",
                   "AvS3MT", "AvS9MT",
                   "AvS3FA", "AvS9FA", "AvS40FA",
                   "AvS3MA", "AvS9MA")]
gr <- c("03FH", "09FH", "40FH",
        "03MH", "09MH",
        "03FT", "09FT", "40FT",
        "03MT", "09MT",
        "03FA", "09FA", "40FA",
        "03MA", "09MA")
colnames(FC.subset) = gr

pca <- prcomp(t(FC.subset), scale.=TRUE)

PC1=scale(pca$x[,1])
PC2=scale(pca$x[,2])
eigs <- pca$sdev^2
ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(Sample=gr,
                        ##PC1=scale(pca$x[,1]),
                        ##PC2=scale(pca$x[,2]),
                        PC1=scale(pca$x[,1]),
                        PC2=scale(pca$x[,2]),
                        sdev=pca$sdev,
                        Tissue=substr(gr, 4, 4), 
                        Sex=substr(gr, 3, 3), 
                        Age=substr(gr, 1, 2),
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group= paste0(substr(gr, 3, 3),
                                      substr(gr, 4, 4)),
                        pca.color = substr(gr, 3, 3),
                        pca.shape = substr(gr, 4, 4))

fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~pca.color,
                 symbol = ~pca.shape,
                 marker = list(size = 15,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Sample)) %>%
    layout(xaxis = list(title = paste0("PC 1", "(", pca.data$Percent1[1], "%)")),
           yaxis = list(title = paste0("PC 2", "(", pca.data$Percent2[2], "%)")))
```

```{r}

  fig

```

```{r echo=F, include=F}

pca.data = cpmdata[,substr(colnames(cpmdata), nchar(colnames(cpmdata))-1, nchar(colnames(cpmdata))-1) == "H"]

remove = apply(pca.data, MARGIN = 1, sum) != 0

pca.data = pca.data[remove,]

pca <- prcomp(t(pca.data), scale.=TRUE) 
gr <- colnames(pca.data)
PC1=scale(pca$x[,1])
PC2=scale(pca$x[,2])
eigs <- pca$sdev^2
ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(Sample=gr,
                        ##PC1=scale(pca$x[,1]),
                        ##PC2=scale(pca$x[,2]),
                        PC1=scale(pca$x[,1]),
                        PC2=scale(pca$x[,2]),
                        sdev=pca$sdev,
                        Genotype=substr(gr, 1, 1), 
                        Sex=substr(gr, nchar(gr)-2, nchar(gr)-2), 
                        Age=substr(gr, 2, nchar(gr)-3),
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group= paste0(substr(gr, 1, 1),
                                      substr(gr, nchar(gr)-2, nchar(gr)-1)))

fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Genotype,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c(input$color1, input$color2, input$color3)[c(1:length(levels(pca.color)))],
                 marker = list(size = 15,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Sample)) %>%
    layout(xaxis = list(title = paste0("PC 1", "(", pca.data$Percent1[1], "%)")),
           yaxis = list(title = paste0("PC 2", "(", pca.data$Percent2[2], "%)")))
```

```{r}

  fig

```

```{r echo=F, include=F}

pca.data = cpmdata[,substr(colnames(cpmdata), nchar(colnames(cpmdata))-1, nchar(colnames(cpmdata))-1) == "T"]

remove = apply(pca.data, MARGIN = 1, sum) != 0

pca.data = pca.data[remove,]

pca <- prcomp(t(pca.data), scale.=TRUE) 
gr <- colnames(pca.data)
PC1=scale(pca$x[,1])
PC2=scale(pca$x[,2])
eigs <- pca$sdev^2
ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(Sample=gr,
                        ##PC1=scale(pca$x[,1]),
                        ##PC2=scale(pca$x[,2]),
                        PC1=scale(pca$x[,1]),
                        PC2=scale(pca$x[,2]),
                        sdev=pca$sdev,
                        Genotype=substr(gr, 1, 1), 
                        Sex=substr(gr, nchar(gr)-2, nchar(gr)-2), 
                        Age=substr(gr, 2, nchar(gr)-3),
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group= paste0(substr(gr, 1, 1),
                                      substr(gr, nchar(gr)-2, nchar(gr)-1)))

fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Genotype,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c(input$color1, input$color2, input$color3)[c(1:length(levels(pca.color)))],
                 marker = list(size = 15,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Sample)) %>%
    layout(xaxis = list(title = paste0("PC 1", "(", pca.data$Percent1[1], "%)")),
           yaxis = list(title = paste0("PC 2", "(", pca.data$Percent2[2], "%)")))
```

```{r}

  fig

```

```{r echo=F, include=F}

pca.data = A4V.cpm[,substr(colnames(A4V.cpm), nchar(colnames(A4V.cpm))-1, nchar(colnames(A4V.cpm))-1) == "A"]

remove = apply(pca.data, MARGIN = 1, sum) != 0

pca.data = pca.data[remove,]

pca <- prcomp(t(pca.data), scale.=TRUE) 
gr <- colnames(pca.data)
PC1=scale(pca$x[,1])
PC2=scale(pca$x[,2])
eigs <- pca$sdev^2
ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(Sample=gr,
                        ##PC1=scale(pca$x[,1]),
                        ##PC2=scale(pca$x[,2]),
                        PC1=scale(pca$x[,1]),
                        PC2=scale(pca$x[,2]),
                        sdev=pca$sdev,
                        Genotype=substr(gr, 1, 1), 
                        Sex=substr(gr, nchar(gr)-2, nchar(gr)-2), 
                        Age=substr(gr, 2, nchar(gr)-3),
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group= paste0(substr(gr, 1, 1),
                                      substr(gr, nchar(gr)-2, nchar(gr)-1)))

fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Genotype,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c(input$color1, input$color2, input$color3)[c(1:length(levels(pca.color)))],
                 marker = list(size = 15,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Sample)) %>%
    layout(xaxis = list(title = paste0("PC 1", "(", pca.data$Percent1[1], "%)")),
           yaxis = list(title = paste0("PC 2", "(", pca.data$Percent2[2], "%)")))
```

```{r}

  fig

```