### 2024-03-22 A4V_BodySections_RNAseq/R code/A4V_Tables.Rmd

---
title: "A4V PCA"
author: "John Santiago"
date: "2024-01-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=F, include=F}

library(edgeR)
library(goseq)
library(org.Dm.eg.db)

```

```{r include =F}
sample.names = list.files("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/Count_Tables/")

meta = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.metadata.csv", row.names = 1)

temp = read.csv(paste0("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/Count_Tables/", sample.names[1]), sep = '\t', row.names = 1)

countdata = data.frame(temp = temp)
colnames(countdata)[1] = substr(sample.names[1], 1, nchar(sample.names[1])-15)

i=2
while(i <= length(sample.names)){
  temp = read.csv(paste0("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/Count_Tables/", sample.names[i]), sep = '\t', row.names = 1)
  countdata = cbind(countdata, temp[row.names(countdata),1])
  colnames(countdata)[i] = substr(sample.names[i], 1, nchar(sample.names[i])-15)
  i=i+1
}

countdata = countdata[-(grep("ti",row.names(countdata))),]
countdata = countdata[-(grep("trans",row.names(countdata))),]

countdata=countdata[substr(row.names(countdata),1,2)=="FB",]

##normalize data
x <- countdata
gr <- factor(substr(colnames(countdata), 1, nchar(colnames(countdata))-1))
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- normLibSizes(y)
##normalized cpm
cpmdata=cpm(z)

##updated edger analysis pipeline
design<-model.matrix(~0 + gr)
colnames(design) = unique(gr)
z = estimateDisp(z, design)
fit <- glmQLFit(z, design)

contr = makeContrasts(A3FH - S3FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS3FH = DEdata
sigAvS3FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3MH - S3MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS3MH = DEdata
sigAvS3MH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FH - S9FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS9FH = DEdata
sigAvS9FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9MH - S9MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS9MH = DEdata
sigAvS9MH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A40FH - S40FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS40FH = DEdata
sigAvS40FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FH - S9FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v9FH = DEdata
sigS3v9FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FH - A9FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v9FH = DEdata
sigA3v9FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3MH - S9MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v9MH = DEdata
sigS3v9MH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3MH - A9MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v9MH = DEdata
sigA3v9MH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S9FH - S40FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S9v40FH = DEdata
sigS9v40FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FH - A40FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A9v40FH = DEdata
sigA9v40FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FH - S40FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v40FH = DEdata
sigS3v40FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FH - A40FH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v40FH = DEdata
sigA3v40FH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FT - S3FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS3FT = DEdata
sigAvS3FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3MT - S3MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS3MT = DEdata
sigAvS3MT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FT - S9FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS9FT = DEdata
sigAvS9FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9MT - S9MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS9MT = DEdata
sigAvS9MT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A40FT - S40FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS40FT = DEdata
sigAvS40FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FT - S9FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v9FT = DEdata
sigS3v9FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FT - A9FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v9FT = DEdata
sigA3v9FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3MT - S9MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v9MT = DEdata
sigS3v9MT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3MT - A9MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v9MT = DEdata
sigA3v9MT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S9FT - S40FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S9v40FT = DEdata
sigS9v40FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FT - A40FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A9v40FT = DEdata
sigA9v40FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FT - S40FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v40FT = DEdata
sigS3v40FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FT - A40FT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v40FT = DEdata
sigA3v40FT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FA - S3FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS3FA = DEdata
sigAvS3FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3MA - S3MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS3MA = DEdata
sigAvS3MA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FA - S9FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS9FA = DEdata
sigAvS9FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9MA - S9MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS9MA = DEdata
sigAvS9MA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A40FA - S40FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
AvS40FA = DEdata
sigAvS40FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FA - S9FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v9FA = DEdata
sigS3v9FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FA - A9FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v9FA = DEdata
sigA3v9FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3MA - S9MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v9MA = DEdata
sigS3v9MA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3MA - A9MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v9MA = DEdata
sigA3v9MA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S9FA - S40FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S9v40FA = DEdata
sigS9v40FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FA - A40FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A9v40FA = DEdata
sigA9v40FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FA - S40FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3v40FA = DEdata
sigS3v40FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FA - A40FA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3v40FA = DEdata
sigA3v40FA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FH - A3MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3FvMH = DEdata
sigA3FvMH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FH - S3MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3FvMH = DEdata
sigS3FvMH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FH - A9MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A9FvMH = DEdata
sigA9FvMH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S9FH - S9MH, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S9FvMH = DEdata
sigS9FvMH = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FT - A3MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3FvMT = DEdata
sigA3FvMT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FT - S3MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3FvMT = DEdata
sigS3FvMT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FT - A9MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A9FvMT = DEdata
sigA9FvMT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S9FT - S9MT, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S9FvMT = DEdata
sigS9FvMT = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A3FA - A3MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A3FvMA = DEdata
sigA3FvMA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S3FA - S3MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S3FvMA = DEdata
sigS3FvMA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(A9FA - A9MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
A9FvMA = DEdata
sigA9FvMA = DEdata[DEdata$FDR < 0.05, ]

contr = makeContrasts(S9FA - S9MA, levels=design)
qlf = glmQLFTest(fit, contrast=contr)
DEdata = topTags(qlf,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEdata = DEdata$table
S9FvMA = DEdata
sigS9FvMA = DEdata[DEdata$FDR < 0.05, ]

order = row.names(cpmdata)
A4V.FDR = data.frame(AvS3FH = AvS3FH[order, "FDR"],
                     AvS3MH = AvS3MH[order, "FDR"],
                     AvS9FH = AvS9FH[order, "FDR"],
                     AvS9MH = AvS9MH[order, "FDR"],
                     AvS40FH = AvS40FH[order, "FDR"],
                     A3v9FH = A3v9FH[order, "FDR"],
                     S3v9FH = S3v9FH[order, "FDR"],
                     A3v9MH = A3v9MH[order, "FDR"],
                     S3v9MH = S3v9MH[order, "FDR"],
                     A9v40FH = A9v40FH[order, "FDR"],
                     S9v40FH = S9v40FH[order, "FDR"],
                     A3v40FH = A3v40FH[order, "FDR"],
                     S3v40FH = S3v40FH[order, "FDR"],

                     AvS3FT = AvS3FT[order, "FDR"],
                     AvS3MT = AvS3MT[order, "FDR"],
                     AvS9FT = AvS9FT[order, "FDR"],
                     AvS9MT = AvS9MT[order, "FDR"],
                     AvS40FT = AvS40FT[order, "FDR"],
                     A3v9FT = A3v9FT[order, "FDR"],
                     S3v9FT = S3v9FT[order, "FDR"],
                     A3v9MT = A3v9MT[order, "FDR"],
                     S3v9MT = S3v9MT[order, "FDR"],
                     A9v40FT = A9v40FT[order, "FDR"],
                     S9v40FT = S9v40FT[order, "FDR"],
                     A3v40FT = A3v40FT[order, "FDR"],
                     S3v40FT = S3v40FT[order, "FDR"],

                     AvS3FA = AvS3FA[order, "FDR"],
                     AvS3MA = AvS3MA[order, "FDR"],
                     AvS9FA = AvS9FA[order, "FDR"],
                     AvS9MA = AvS9MA[order, "FDR"],
                     AvS40FA = AvS40FA[order, "FDR"],
                     A3v9FA = A3v9FA[order, "FDR"],
                     S3v9FA = S3v9FA[order, "FDR"],
                     A3v9MA = A3v9MA[order, "FDR"],
                     S3v9MA = S3v9MA[order, "FDR"],
                     A9v40FA = A9v40FA[order, "FDR"],
                     S9v40FA = S9v40FA[order, "FDR"],
                     A3v40FA = A3v40FA[order, "FDR"],
                     S3v40FA = S3v40FA[order, "FDR"],

                     A3FvMH = A3FvMH[order, "FDR"],
                     S3FvMH = S3FvMH[order, "FDR"],
                     A9FvMH = A9FvMH[order, "FDR"],
                     S9FvMH = S9FvMH[order, "FDR"],
                     A3FvMT = A3FvMT[order, "FDR"],
                     S3FvMT = S3FvMT[order, "FDR"],
                     A9FvMT = A9FvMT[order, "FDR"],
                     S9FvMT = S9FvMT[order, "FDR"],
                     A3FvMA = A3FvMA[order, "FDR"],
                     S3FvMA = S3FvMA[order, "FDR"],
                     A9FvMA = A9FvMA[order, "FDR"],
                     S9FvMA = S9FvMA[order, "FDR"])
row.names(A4V.FDR) = order

A4V.FC = data.frame(AvS3FH = AvS3FH[order, "logFC"],
                     AvS3MH = AvS3MH[order, "logFC"],
                     AvS9FH = AvS9FH[order, "logFC"],
                     AvS9MH = AvS9MH[order, "logFC"],
                     AvS40FH = AvS40FH[order, "logFC"],
                     A3v9FH = A3v9FH[order, "logFC"],
                     S3v9FH = S3v9FH[order, "logFC"],
                     A3v9MH = A3v9MH[order, "logFC"],
                     S3v9MH = S3v9MH[order, "logFC"],
                     A9v40FH = A9v40FH[order, "logFC"],
                     S9v40FH = S9v40FH[order, "logFC"],
                     A3v40FH = A3v40FH[order, "logFC"],
                     S3v40FH = S3v40FH[order, "logFC"],

                     AvS3FT = AvS3FT[order, "logFC"],
                     AvS3MT = AvS3MT[order, "logFC"],
                     AvS9FT = AvS9FT[order, "logFC"],
                     AvS9MT = AvS9MT[order, "logFC"],
                     AvS40FT = AvS40FT[order, "logFC"],
                     A3v9FT = A3v9FT[order, "logFC"],
                     S3v9FT = S3v9FT[order, "logFC"],
                     A3v9MT = A3v9MT[order, "logFC"],
                     S3v9MT = S3v9MT[order, "logFC"],
                     A9v40FT = A9v40FT[order, "logFC"],
                     S9v40FT = S9v40FT[order, "logFC"],
                     A3v40FT = A3v40FT[order, "logFC"],
                     S3v40FT = S3v40FT[order, "logFC"],

                     AvS3FA = AvS3FA[order, "logFC"],
                     AvS3MA = AvS3MA[order, "logFC"],
                     AvS9FA = AvS9FA[order, "logFC"],
                     AvS9MA = AvS9MA[order, "logFC"],
                     AvS40FA = AvS40FA[order, "logFC"],
                     A3v9FA = A3v9FA[order, "logFC"],
                     S3v9FA = S3v9FA[order, "logFC"],
                     A3v9MA = A3v9MA[order, "logFC"],
                     S3v9MA = S3v9MA[order, "logFC"],
                     A9v40FA = A9v40FA[order, "logFC"],
                     S9v40FA = S9v40FA[order, "logFC"],
                     A3v40FA = A3v40FA[order, "logFC"],
                     S3v40FA = S3v40FA[order, "logFC"],

                     A3FvMH = A3FvMH[order, "logFC"],
                     S3FvMH = S3FvMH[order, "logFC"],
                     A9FvMH = A9FvMH[order, "logFC"],
                     S9FvMH = S9FvMH[order, "logFC"],
                     A3FvMT = A3FvMT[order, "logFC"],
                     S3FvMT = S3FvMT[order, "logFC"],
                     A9FvMT = A9FvMT[order, "logFC"],
                     S9FvMT = S9FvMT[order, "logFC"],
                     A3FvMA = A3FvMA[order, "logFC"],
                     S3FvMA = S3FvMA[order, "logFC"],
                     A9FvMA = A9FvMA[order, "logFC"],
                     S9FvMA = S9FvMA[order, "logFC"])
row.names(A4V.FC) = order

cpmdata = cpmdata[grep("FBgn",row.names(cpmdata)),]
A4V.FDR = A4V.FDR[row.names(cpmdata),]

A4V.meancpm = cpmdata[,1:30]
colnames(A4V.meancpm) = unique(meta$Group)
i=1
while(i<=nrow(A4V.meancpm)){
  j=1
  while(j<=ncol(A4V.meancpm)){
    A4V.meancpm[i,j]=mean(cpmdata[i,row.names(meta)[meta$Group == colnames(A4V.meancpm)[j]]])
    j=j+1
  }
  i=i+1
}

##write.csv(cpmdata, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.cpmdata.csv")
##write.csv(countdata, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.countdata.csv")
##write.csv(A4V.FDR, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FDRdata.csv")
##write.csv(A4V.meancpm, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.meancpmdata.csv")
##write.csv(A4V.FC, "/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FCdata.csv")

```

##GO and KEGG Enrichment function
```{r echo = F, include = F}

Enrichment = function(sigs, bg){
  genes = setNames(rep(0, length(bg)), bg)
  genes[sigs] = 1
  pwf=nullp(genes,"dm3","geneSymbol")
  GO.wall=goseq(pwf,"dm3","geneSymbol")
  GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
  row.names(GO.wall) = GO.wall$category
  KEGG=goseq(pwf,gene2cat=kegg)
  KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
  row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
  KEGG$Name=kegg.names[row.names(KEGG),1]

  enrichment.data = list(GO.wall, KEGG)
  names(enrichment.data) = c("GO", "KEGG")
  return(enrichment.data)
}

```

```{r echo = F, include = F}

i=1

while(i<=ncol(FDR)){
compare = colnames(FDR)[i]

sigs = row.names(FDR)[FDR[,i]<.05]
bg = row.names(FDR)

sigsKEY = GeneIDKey[sigs,]
sigs = na.omit(sigsKEY$Symbol)
bgKEY = GeneIDKey[bg,]
bg = na.omit(bgKEY$Symbol)

sig.cats = Enrichment(sigs, bg)

##write.csv(sig.cats[[1]], paste0("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/GOterm Enrichment/", colnames(FDR)[i],"_GOTermAnalysis.csv"))

##write.csv(sig.cats[[2]], paste0("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/KEGG Enrichment/", colnames(FDR)[i],"KEGGAnalysis.csv"))

i = i +1

}

```