### 2024-03-22 Cami_qpcr 2_28_24 - Sheet1.Rmd

---
title: "Cami qPCR 20240228"
author: "John Santiago"
date: "2024-02-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}

cami = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Cami_qpcr 2_28_24 - Sheet1.csv", header = T)

```

```{r}

plotdata1 = cami[cami$Genotype=='OK371>G4C2' & cami$Primer == 'vap33',]
plotdata2 = cami[cami$Genotype=='OK371>G4C2' & cami$Primer == 'Rp49',]
plotdata3 = cami[cami$Genotype=='OK371>G4C2' & cami$Primer == 'xbp1',] 
plotdata4 = cami[cami$Genotype=='OK371>G4C2' & cami$Primer == 'atf4',] 
plotdata5 = cami[cami$Genotype=='OK371>G4C2' & cami$Primer == 'atf6',]

plot(x = log2(plotdata1$Concentration), 
     y = plotdata1$Cq.Mean, 
     type = "p",  
     cex = 1, 
     pch = 21, 
     bg = "firebrick",
     col = "black",
     ylab = "Quantification Cycle (Cq)",
     xlab = "log2(cDNA Conc. (ng/uL)",
     ylim = c(16,28))

points(x = log2(plotdata2$Concentration),
       y = plotdata2$Cq.Mean,
       cex = 1,
       pch = 21,
       bg = "steelblue",
       col = "black")

points(x = log2(plotdata3$Concentration),
       y = plotdata3$Cq.Mean,
       cex = 1,
       pch = 21,
       bg = "gold",
       col = "black")

points(x = log2(plotdata4$Concentration),
       y = plotdata4$Cq.Mean,
       cex = 1,
       pch = 21,
       bg = "darkgreen",
       col = "black")

points(x = log2(plotdata5$Concentration),
       y = plotdata5$Cq.Mean,
       cex = 1,
       pch = 21,
       bg = "grey",
       col = "black")

linearmodel=lm(as.numeric(plotdata1$Cq.Mean)~log2(plotdata1$Concentration))
abline(a=linearmodel[[1]][[1]],b=linearmodel[[1]][2],lty=2,col="firebrick",lwd=2)
text(x = 2, y = 21,paste0("vap33 slope = ", signif(linearmodel[[1]][2],3)),cex=.8)

linearmodel=lm(as.numeric(plotdata2$Cq.Mean)~log2(plotdata2$Concentration))
abline(a=linearmodel[[1]][[1]],b=linearmodel[[1]][2],lty=2,col="steelblue",lwd=2)
text(x = 2, y = 20.5,paste0("Rp49 slope = ", signif(linearmodel[[1]][2],3)),cex=.8)

linearmodel=lm(as.numeric(plotdata3$Cq.Mean)~log2(plotdata3$Concentration))
abline(a=linearmodel[[1]][[1]],b=linearmodel[[1]][2],lty=2,col="gold",lwd=2)
text(x = 2, y = 20,paste0("xbp1 slope = ", signif(linearmodel[[1]][2],3)),cex=.8)

linearmodel=lm(as.numeric(plotdata4$Cq.Mean)~log2(plotdata4$Concentration))
abline(a=linearmodel[[1]][[1]],b=linearmodel[[1]][2],lty=2,col="darkgreen",lwd=2)
text(x = 2, y = 19.5,paste0("atf4 slope = ", signif(linearmodel[[1]][2],3)),cex=.8)

linearmodel=lm(as.numeric(plotdata5$Cq.Mean)~log2(plotdata5$Concentration))
abline(a=linearmodel[[1]][[1]],b=linearmodel[[1]][2],lty=2,col="grey",lwd=2)
text(x = 2, y = 19,paste0("atf6 slope = ", signif(linearmodel[[1]][2],3)),cex=.8)

legend("topright", legend = c("vap33", "Rp49", 'xbp1', 'atf4', 'atf5'), pch = 21, col = "black", pt.bg = c("firebrick", "steelblue", 'gold', 'darkgreen', 'grey'), bty = "n")

```

```{r}

cami2 = cami
cami2$Concentration[cami2$Concentration == 5] = 0
cami2$Concentration[cami2$Concentration == 2.5] = 5
cami2$Concentration[cami2$Concentration == 0] = 2.5

plotdata1 = cami2[cami2$Genotype=='OK371>G4C2' & cami2$Primer == 'vap33',]
plotdata2 = cami2[cami2$Genotype=='OK371>G4C2' & cami2$Primer == 'Rp49',]
plotdata3 = cami2[cami2$Genotype=='OK371>G4C2' & cami2$Primer == 'xbp1',] 
plotdata4 = cami2[cami2$Genotype=='OK371>G4C2' & cami2$Primer == 'atf4',] 
plotdata5 = cami2[cami2$Genotype=='OK371>G4C2' & cami2$Primer == 'atf6',]

plot(x = log2(plotdata1$Concentration), 
     y = plotdata1$Cq.Mean, 
     type = "p",  
     cex = 1, 
     pch = 21, 
     bg = "firebrick",
     col = "black",
     ylab = "Quantification Cycle (Cq)",
     xlab = "log2(cDNA Conc. (ng/uL)",
     ylim = c(16,28))

points(x = log2(plotdata2$Concentration),
       y = plotdata2$Cq.Mean,
       cex = 1,
       pch = 21,
       bg = "steelblue",
       col = "black")

points(x = log2(plotdata3$Concentration),
       y = plotdata3$Cq.Mean,
       cex = 1,
       pch = 21,
       bg = "gold",
       col = "black")

points(x = log2(plotdata4$Concentration),
       y = plotdata4$Cq.Mean,
       cex = 1,
       pch = 21,
       bg = "darkgreen",
       col = "black")

points(x = log2(plotdata5$Concentration),
       y = plotdata5$Cq.Mean,
       cex = 1,
       pch = 21,
       bg = "grey",
       col = "black")

linearmodel=lm(as.numeric(plotdata1$Cq.Mean)~log2(plotdata1$Concentration))
abline(a=linearmodel[[1]][[1]],b=linearmodel[[1]][2],lty=2,col="firebrick",lwd=2)
text(x = 2, y = 21,paste0("vap33 slope = ", signif(linearmodel[[1]][2],3)),cex=.8)

linearmodel=lm(as.numeric(plotdata2$Cq.Mean)~log2(plotdata2$Concentration))
abline(a=linearmodel[[1]][[1]],b=linearmodel[[1]][2],lty=2,col="steelblue",lwd=2)
text(x = 2, y = 20.5,paste0("Rp49 slope = ", signif(linearmodel[[1]][2],3)),cex=.8)

linearmodel=lm(as.numeric(plotdata3$Cq.Mean)~log2(plotdata3$Concentration))
abline(a=linearmodel[[1]][[1]],b=linearmodel[[1]][2],lty=2,col="gold",lwd=2)
text(x = 2, y = 20,paste0("xbp1 slope = ", signif(linearmodel[[1]][2],3)),cex=.8)

linearmodel=lm(as.numeric(plotdata4$Cq.Mean)~log2(plotdata4$Concentration))
abline(a=linearmodel[[1]][[1]],b=linearmodel[[1]][2],lty=2,col="darkgreen",lwd=2)
text(x = 2, y = 19.5,paste0("atf4 slope = ", signif(linearmodel[[1]][2],3)),cex=.8)

linearmodel=lm(as.numeric(plotdata5$Cq.Mean)~log2(plotdata5$Concentration))
abline(a=linearmodel[[1]][[1]],b=linearmodel[[1]][2],lty=2,col="grey",lwd=2)
text(x = 2, y = 19,paste0("atf6 slope = ", signif(linearmodel[[1]][2],3)),cex=.8)

legend("topright", legend = c("vap33", "Rp49", 'xbp1', 'atf4', 'atf5'), pch = 21, col = "black", pt.bg = c("firebrick", "steelblue", 'gold', 'darkgreen', 'grey'), bty = "n")

```