### 2024-03-23 Anna qPCR data/AnnaThesisAnalyses.Rmd
update qPCR

---
title: "Anna Analysis with John"
author: "John Santiago"
date: "2024-03-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F}

qPCR = read.csv("https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/Anna%20qPCR%20data/Anna_qPCR_allData.csv")

```

```{r echo = F}

pcb = qPCR[qPCR$Primer == "PCB",]
pcb = pcb[pcb$Censor != 'yes',]
pcb = pcb[pcb$Concentration == 5,]
pcb = pcb[pcb$Efficiency == 'no',]

rp49 = qPCR[qPCR$Primer == "RP49",]
rp49 = rp49[rp49$Censor != 'yes',]
rp49 = rp49[rp49$Concentration == 5,]
rp49 = rp49[rp49$Efficiency == 'no',]

temp = unique(paste(pcb$Genotype, pcb$Biological.replicate))

pcb.mean = rep(0, length(temp))
rp49.mean = rep(0, length(temp)) 
i=1
while(i<=length(temp)){
  pcb.mean[i] = mean(na.omit(as.numeric(pcb[paste(pcb$Genotype, pcb$Biological.replicate) == temp[i], "Cq"])))
  rp49.mean[i] = mean(na.omit(as.numeric(rp49[paste(rp49$Genotype, rp49$Biological.replicate) == temp[i], "Cq"])))
  i=i+1
}

mean.pcb = data.frame(Sample = temp,
                     pcb = pcb.mean,
                     rp49 = rp49.mean)

mean.pcb$dCT = mean.pcb$pcb - mean.pcb$rp49

```

```{r echo = F}

ldh = qPCR[qPCR$Primer == "LDH",]
ldh = ldh[ldh$Censor != 'yes',]
ldh = ldh[ldh$Concentration == 5,]
ldh = ldh[ldh$Efficiency == 'no',]

rp49 = qPCR[qPCR$Primer == "RP49",]
rp49 = rp49[rp49$Censor != 'yes',]
rp49 = rp49[rp49$Concentration == 5,]
rp49 = rp49[rp49$Efficiency == 'no',]

temp = unique(paste(ldh$Genotype, ldh$Biological.replicate))

ldh.mean = rep(0, length(temp))
rp49.mean = rep(0, length(temp)) 
i=1
while(i<=length(temp)){
  ldh.mean[i] = mean(na.omit(as.numeric(ldh[paste(ldh$Genotype, ldh$Biological.replicate) == temp[i], "Cq"])))
  rp49.mean[i] = mean(na.omit(as.numeric(rp49[paste(rp49$Genotype, rp49$Biological.replicate) == temp[i], "Cq"])))
  i=i+1
}

mean.ldh = data.frame(Sample = temp,
                     ldh = ldh.mean,
                     rp49 = rp49.mean)

mean.ldh$dCT = mean.ldh$ldh - mean.ldh$rp49

```

```{r echo = F}

pdh = qPCR[qPCR$Primer == "PDH",]
pdh = pdh[pdh$Censor != 'yes',]
pdh = pdh[pdh$Concentration == 5,]
pdh = pdh[pdh$Efficiency == 'no',]

rp49 = qPCR[qPCR$Primer == "RP49",]
rp49 = rp49[rp49$Censor != 'yes',]
rp49 = rp49[rp49$Concentration == 5,]
rp49 = rp49[rp49$Efficiency == 'no',]

temp = unique(paste(pdh$Genotype, pdh$Biological.replicate))

pdh.mean = rep(0, length(temp))
rp49.mean = rep(0, length(temp)) 
i=1
while(i<=length(temp)){
  pdh.mean[i] = mean(na.omit(as.numeric(pdh[paste(pdh$Genotype, pdh$Biological.replicate) == temp[i], "Cq"])))
  rp49.mean[i] = mean(na.omit(as.numeric(rp49[paste(rp49$Genotype, rp49$Biological.replicate) == temp[i], "Cq"])))
  i=i+1
}

mean.pdh = data.frame(Sample = temp,
                     pdh = pdh.mean,
                     rp49 = rp49.mean)

mean.pdh$dCT = mean.pdh$pdh - mean.pdh$rp49

```

```{r echo = F}

pepck = qPCR[qPCR$Primer == "Pepck-1",]
pepck = pepck[pepck$Censor != 'yes',]
pepck = pepck[pepck$Concentration == 5,]
pepck = pepck[pepck$Efficiency == 'no',]

rp49 = qPCR[qPCR$Primer == "RP49",]
rp49 = rp49[rp49$Censor != 'yes',]
rp49 = rp49[rp49$Concentration == 5,]
rp49 = rp49[rp49$Efficiency == 'no',]

temp = unique(paste(pepck$Genotype, pepck$Biological.replicate))

pepck.mean = rep(0, length(temp))
rp49.mean = rep(0, length(temp)) 
i=1
while(i<=length(temp)){
  pepck.mean[i] = mean(na.omit(as.numeric(pepck[paste(pepck$Genotype, pepck$Biological.replicate) == temp[i], "Cq"])))
  rp49.mean[i] = mean(na.omit(as.numeric(rp49[paste(rp49$Genotype, rp49$Biological.replicate) == temp[i], "Cq"])))
  i=i+1
}

mean.pepck = data.frame(Sample = temp,
                     pepck = pepck.mean,
                     rp49 = rp49.mean)

mean.pepck$dCT = mean.pepck$pepck - mean.pepck$rp49

```

```{r echo = F}

idgf = qPCR[qPCR$Primer == "Idgf6",]
idgf = idgf[idgf$Censor != 'yes',]
idgf = idgf[idgf$Concentration == 5,]
idgf = idgf[idgf$Efficiency == 'no',]

rp49 = qPCR[qPCR$Primer == "RP49",]
rp49 = rp49[rp49$Censor != 'yes',]
rp49 = rp49[rp49$Concentration == 5,]
rp49 = rp49[rp49$Efficiency == 'no',]

temp = unique(paste(idgf$Genotype, idgf$Biological.replicate))

idgf.mean = rep(0, length(temp))
rp49.mean = rep(0, length(temp)) 
i=1
while(i<=length(temp)){
  idgf.mean[i] = mean(na.omit(as.numeric(idgf[paste(idgf$Genotype, idgf$Biological.replicate) == temp[i], "Cq"])))
  rp49.mean[i] = mean(na.omit(as.numeric(rp49[paste(rp49$Genotype, rp49$Biological.replicate) == temp[i], "Cq"])))
  i=i+1
}

mean.idgf = data.frame(Sample = temp,
                     idgf = idgf.mean,
                     rp49 = rp49.mean)

mean.idgf$dCT = mean.idgf$idgf - mean.idgf$rp49

```

```{r echo = F}

p53 = qPCR[qPCR$Primer == "p53",]
p53 = p53[p53$Censor != 'yes',]
p53 = p53[p53$Concentration == 5,]
p53 = p53[p53$Efficiency == 'no',]

rp49 = qPCR[qPCR$Primer == "RP49",]
rp49 = rp49[rp49$Censor != 'yes',]
rp49 = rp49[rp49$Concentration == 5,]
rp49 = rp49[rp49$Efficiency == 'no',]

temp = unique(paste(p53$Genotype, p53$Biological.replicate))

p53.mean = rep(0, length(temp))
rp49.mean = rep(0, length(temp)) 
i=1
while(i<=length(temp)){
  p53.mean[i] = mean(na.omit(as.numeric(p53[paste(p53$Genotype, p53$Biological.replicate) == temp[i], "Cq"])))
  rp49.mean[i] = mean(na.omit(as.numeric(rp49[paste(rp49$Genotype, rp49$Biological.replicate) == temp[i], "Cq"])))
  i=i+1
}

mean.p53 = data.frame(Sample = temp,
                     p53 = p53.mean,
                     rp49 = rp49.mean)

mean.p53$dCT = mean.p53$p53 - mean.p53$rp49

```

```{r echo = F}

PyK = qPCR[qPCR$Primer == "PyK",]
PyK = PyK[PyK$Censor != 'yes',]
PyK = PyK[PyK$Concentration == 5,]
PyK = PyK[PyK$Efficiency == 'no',]

rp49 = qPCR[qPCR$Primer == "RP49",]
rp49 = rp49[rp49$Censor != 'yes',]
rp49 = rp49[rp49$Concentration == 5,]
rp49 = rp49[rp49$Efficiency == 'no',]

temp = unique(paste(PyK$Genotype, PyK$Biological.replicate))

PyK.mean = rep(0, length(temp))
rp49.mean = rep(0, length(temp)) 
i=1
while(i<=length(temp)){
  PyK.mean[i] = mean(na.omit(as.numeric(PyK[paste(PyK$Genotype, PyK$Biological.replicate) == temp[i], "Cq"])))
  rp49.mean[i] = mean(na.omit(as.numeric(rp49[paste(rp49$Genotype, rp49$Biological.replicate) == temp[i], "Cq"])))
  i=i+1
}

mean.PyK = data.frame(Sample = temp,
                     PyK = PyK.mean,
                     rp49 = rp49.mean)

mean.PyK$dCT = mean.PyK$PyK - mean.PyK$rp49

```

```{r echo = F}

plotdata = mean.pcb
keep = plotdata$Sample != "WT-cr 1" & plotdata$Sample != "WT-cr 2"
color = "cornflowerblue"
main = "PCB"
plotdata = plotdata[keep,]
plotdata$Sample[grep("WT",plotdata$Sample)] = "WT"
plotdata$Sample[grep("RNAi",plotdata$Sample)] = "RNAi"
plotdata$Sample[grep("G85R",plotdata$Sample)] = "G85R"
plotdata$Sample = factor(plotdata$Sample,
                         levels = c("WT", "G85R", "RNAi"))

boxplot(plotdata$dCT~plotdata$Sample,
     ylab = "dCT",
     xlab = "",
     main = main,
     las = 2)

points(x = plotdata$Sample,
       y = plotdata$dCT,
       cex = 1.5,
       pch = 21,
       bg = color,
       col = "black")

```

```{r echo = F}

dCT.aov = aov(plotdata$dCT~plotdata$Sample)

TukeyHSD(dCT.aov)

```

```{r echo = F}

plotdata = mean.ldh
keep = T
color = "gold"
main = "LDH"
plotdata = plotdata[keep,]
plotdata$Sample[grep("WT",plotdata$Sample)] = "WT"
plotdata$Sample[grep("RNAi",plotdata$Sample)] = "RNAi"
plotdata$Sample[grep("G85R",plotdata$Sample)] = "G85R"
plotdata$Sample = factor(plotdata$Sample,
                         levels = c("WT", "G85R", "RNAi"))

boxplot(plotdata$dCT~plotdata$Sample,
     ylab = "dCT",
     xlab = "",
     main = main,
     las = 2)

points(x = plotdata$Sample,
       y = plotdata$dCT,
       cex = 1.5,
       pch = 21,
       bg = color,
       col = "black")

```

```{r echo = F}

dCT.aov = aov(plotdata$dCT~plotdata$Sample)

TukeyHSD(dCT.aov)

```

```{r echo = F}

plotdata = mean.pdh
keep = plotdata$Sample != "WT-cr 1" & plotdata$Sample != "WT-cr 2"
color = "brown"
main = "PDH"
plotdata = plotdata[keep,]
plotdata$Sample[grep("WT",plotdata$Sample)] = "WT"
plotdata$Sample[grep("RNAi",plotdata$Sample)] = "RNAi"
plotdata$Sample[grep("G85R",plotdata$Sample)] = "G85R"
plotdata$Sample = factor(plotdata$Sample,
                         levels = c("WT", "G85R", "RNAi"))

boxplot(plotdata$dCT~plotdata$Sample,
     ylab = "dCT",
     xlab = "",
     main = main,
     las = 2)

points(x = plotdata$Sample,
       y = plotdata$dCT,
       cex = 1.5,
       pch = 21,
       bg = color,
       col = "black")

```

```{r echo = F}

dCT.aov = aov(plotdata$dCT~plotdata$Sample)

TukeyHSD(dCT.aov)

```

```{r echo = F}

plotdata = mean.p53
keep = plotdata$Sample != "WT-cr 1" & plotdata$Sample != "WT-cr 2"
color = "black"
main = "p53"
plotdata = plotdata[keep,]
plotdata$Sample[grep("WT",plotdata$Sample)] = "WT"
plotdata$Sample[grep("RNAi",plotdata$Sample)] = "RNAi"
plotdata$Sample[grep("G85R",plotdata$Sample)] = "G85R"
plotdata$Sample = factor(plotdata$Sample,
                         levels = c("WT", "G85R", "RNAi"))

boxplot(plotdata$dCT~plotdata$Sample,
     ylab = "dCT",
     xlab = "",
     main = main,
     las = 2)

points(x = plotdata$Sample,
       y = plotdata$dCT,
       cex = 1.5,
       pch = 21,
       bg = color,
       col = "black")

```

```{r echo = F}

dCT.aov = aov(plotdata$dCT~plotdata$Sample)

TukeyHSD(dCT.aov)

```

```{r echo = F}

plotdata = mean.PyK
keep = plotdata$Sample != "WT-cr 1" & plotdata$Sample != "WT-cr 2"
color = "deeppink"
main = "PyK"
plotdata = plotdata[keep,]
plotdata$Sample[grep("WT",plotdata$Sample)] = "WT"
plotdata$Sample[grep("RNAi",plotdata$Sample)] = "RNAi"
plotdata$Sample[grep("G85R",plotdata$Sample)] = "G85R"
plotdata$Sample = factor(plotdata$Sample,
                         levels = c("WT", "G85R", "RNAi"))

boxplot(plotdata$dCT~plotdata$Sample,
     ylab = "dCT",
     xlab = "",
     main = main,
     las = 2)

points(x = plotdata$Sample,
       y = plotdata$dCT,
       cex = 1.5,
       pch = 21,
       bg = color,
       col = "black")

```

```{r echo = F}

dCT.aov = aov(plotdata$dCT~plotdata$Sample)

TukeyHSD(dCT.aov)

```

```{r echo = F}

plotdata = mean.idgf
keep = T
color = "limegreen"
main = "idgf6"
plotdata = plotdata[keep,]
plotdata$Sample[grep("WT",plotdata$Sample)] = "WT"
plotdata$Sample[grep("RNAi",plotdata$Sample)] = "RNAi"
plotdata$Sample[grep("G85R",plotdata$Sample)] = "G85R"
plotdata$Sample = factor(plotdata$Sample,
                         levels = c("WT", "G85R", "RNAi"))

boxplot(plotdata$dCT~plotdata$Sample,
     ylab = "dCT",
     xlab = "",
     main = main,
     las = 2)

points(x = plotdata$Sample,
       y = plotdata$dCT,
       cex = 1.5,
       pch = 21,
       bg = color,
       col = "black")

```

```{r echo = F}

dCT.aov = aov(plotdata$dCT~plotdata$Sample)

TukeyHSD(dCT.aov)

```

```{r echo = F}

plotdata = mean.pepck
keep = T
color = "orange"
main = "PEPCK-1"
plotdata = plotdata[keep,]
plotdata$Sample[grep("WT",plotdata$Sample)] = "WT"
plotdata$Sample[grep("RNAi",plotdata$Sample)] = "RNAi"
plotdata$Sample[grep("G85R",plotdata$Sample)] = "G85R"
plotdata$Sample = factor(plotdata$Sample,
                         levels = c("WT", "G85R", "RNAi"))

boxplot(plotdata$dCT~plotdata$Sample,
     ylab = "dCT",
     xlab = "",
     main = main,
     las = 2)

points(x = plotdata$Sample,
       y = plotdata$dCT,
       cex = 1.5,
       pch = 21,
       bg = color,
       col = "black")

```

```{r echo = F}

dCT.aov = aov(plotdata$dCT~plotdata$Sample)

TukeyHSD(dCT.aov)

```