### 2024-03-26 Aleah_Thesis_Figures.Rmd
aleah data

---
title: "Aleah Thesis Analyses"
author: "John Santiago"
date: "2024-03-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F}

library(org.Dm.eg.db)
library(goseq)
library(rrvgo)
library(plotly)

git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"

counts = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.countdata.csv", row.names = 1)

cpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.cpmdata.csv", row.names = 1)

meancpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.meancpmdata.csv", row.names = 1)

FDR = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FDRdata.csv", row.names = 1)

FC = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FCdata.csv", row.names = 1)

meta = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.metadata.csv", row.names = 1)

GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)

KEGG.Key = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/RawMetabolomicData.csv", row.names = 1)
KEGG.Key = setNames(KEGG.Key[-c(1:2), 2], row.names(KEGG.Key)[-c(1:2)])

genesingo=as.list(org.Dm.egGO2ALLEGS)

kegg = readRDS(gzcon(url(paste0(git.dir,"GeneralDataFiles/kegg.symbol2path.RData"))))
kegg.names = read.csv(paste0(git.dir, "GeneralDataFiles/KEGG.names.csv"), row.names = 1)

```

```{r echo = F}

total.DEGs = c(length(FDR$AvS3FH[FDR$AvS3FH < .05]),
               length(FDR$AvS9FH[FDR$AvS9FH < .05]),
               length(FDR$AvS40FH[FDR$AvS40FH < .05]),
               length(FDR$AvS3FT[FDR$AvS3FT < .05]),
               length(FDR$AvS9FT[FDR$AvS9FT < .05]),
               length(FDR$AvS40FT[FDR$AvS40FT < .05]),
               length(FDR$AvS3FA[FDR$AvS3FA < .05]),
               length(FDR$AvS9FA[FDR$AvS9FA < .05]),
               length(FDR$AvS40FA[FDR$AvS40FA < .05]),
               NA,
               length(FDR$AvS3MH[FDR$AvS3MH < .05]),
               length(FDR$AvS9MH[FDR$AvS9MH < .05]),
               length(FDR$AvS3MT[FDR$AvS3MT < .05]),
               length(FDR$AvS9MT[FDR$AvS9MT < .05]),
               length(FDR$AvS3MA[FDR$AvS3MA < .05]),
               length(FDR$AvS9MA[FDR$AvS9MA < .05]))

names(total.DEGs) = c('FH3', 'FH9', 'FH40',
                      'FT3', 'FT9', 'FT40',
                      'FA3', 'FA9', 'FA40',
                      NA,
                      'MH3', 'MH9',
                      'MT3', 'MT9',
                      'MA3', 'MA9')
xlabel = names(total.DEGs)

bp.color = c(rep("steelblue", 3), rep("gold2", 3), rep("brown3", 3),
             NA,
             rep("steelblue", 2), rep("gold2", 2), rep("brown3", 2))

```

```{r echo = F}

par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)

barplot(height = total.DEGs,
        names.arg = xlabel,
        beside = T,
        col = bp.color,
        las = 2,
        xlab = "Sample",
        ylab = "Total DEGs",
        main = "Total DEGs in A4V vs Silent",
        ylim = c(0, 4700))
##abline(v = 11.5, lty =2)
legend('topright',
       inset=c(-.35,0),
       bty = 'n',
       legend = c('Head', 'Thorax', 'Abdomen'), 
       fill = c('steelblue', 'gold2', 'brown3'),
       pt.cex = 1,
       cex = 1)
text(x = c(5,16), 
     y = c(4700, 4700),
     labels = c("Female", "Male"))

```

```{r echo = F, warning=FALSE, message=F}
library(VennDiagram)

set1="Male 3 Day"
set2="Female 3 Day"

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1=row.names(FDR[FDR$AvS3MT <.05,])
s2=row.names(FDR[FDR$AvS3FT <.05,])

main = "3 Day Thorax DEGs"

grid.newpage()
draw.pairwise.venn(area1=length(s1),area2=length(s2),
                   cross.area=length(intersect(s1,s2)),
                   fill = c("royalblue1","brown1"),
                   cex=1.5,
                   cat.cex = 1,
                   category = c(set1,set2),
                   main = "main",
                   cat.dist = c(.065,.05),
                   cat.pos = c(-20,20))

```

```{r echo = F, include = F}

Enrichment = function(sigs, bg){
  sigsKEY = GeneIDKey[sigs,]
  sigs = na.omit(sigsKEY$Symbol)
  bgKEY = GeneIDKey[bg,]
  bg = na.omit(bgKEY$Symbol)

  genes = setNames(rep(0, length(bg)), bg)
  genes[sigs] = 1
  pwf=nullp(genes,"dm3","geneSymbol")
  GO.wall=goseq(pwf,"dm3","geneSymbol")
  GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
  row.names(GO.wall) = GO.wall$category
  KEGG=goseq(pwf,gene2cat=kegg)
  KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
  row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
  KEGG$Name=kegg.names[row.names(KEGG),1]

  enrichment.data = list(GO.wall, KEGG)
  names(enrichment.data) = c("GO", "KEGG")
  return(enrichment.data)
}

```

```{r echo = F, include = F}

GO.Sim.PCA = function(use.data){
    pca <- prcomp(t(use.data), scale.=TRUE) 
    pc1=scale(pca$x[,1])
    pc2=scale(pca$x[,2])
    eigs <- pca$sdev^2
    ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(PC1=pc1,
                        PC2=pc2,
                        sdev=pca$sdev,
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group = use.data[row.names(use.data), "parentTerm"],
                        Score = use.data[row.names(use.data), "score"],
                        Term = use.data[row.names(use.data), "term"],
                        TotalDE = GO.data[row.names(use.data), "numDEInCat"],
                        Total = GO.data[row.names(use.data), "numInCat"],
                        SigVal = signif(GO.data[row.names(use.data), "adjp"],3))

    pca.data$Group = factor(pca.data$Group, levels = unique(pca.data$Group))
    pca.data$Score = 20*(1+pca.data$Score/mean(pca.data$Score))

  fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 ##symbol = ~Group,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c("firebrick","gold","deepskyblue"),
                 marker = list(size = pca.data$Score,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Parent Group: ", pca.data$Group,
                                   "\nSample: ", pca.data$Term, 
                                   "\nTotal DE: ", pca.data$TotalDE, 
                                   "\nTotal in Cat.: ", pca.data$Total, 
                                   "\nFDR: ", pca.data$SigVal)) %>%
    layout(xaxis = list(title = paste0("PC1", "(", ve[1], "%)")),
           yaxis = list(title = paste0("PC2", "(", ve[2], "%)")),
           title = "Enriched GO term similarity")

  return(fig)
}

```

```{r echo = F, include =F}

GO20 = function(GO.data, GO.ontology, FC.data, FDR.data){
GO.data = GO.data[GO.data$ontology == GO.ontology,
                  c("category", "term","adjp", "numDEInCat", "numInCat")]

data = GO.data[20:1,c("term","adjp","numDEInCat", "numInCat", "category")]
data$Score = -log10(data$adjp)

temp = nchar(data$term)
data$label = data$term
##data$label[temp > 40] = paste0(substr(data$term[temp>40], 1 ,35), "...")

data$term = factor(data$term, levels = c(unique(data$term)))
data$up = 0
data$down = 0
data$allup = 0
data$alldown = 0

sigs = names(FDR.data[FDR.data <= 0.05])

i = 1
while(i <= nrow(data)){
  go.id = row.names(data)[i]
  go.genes = genesingo[[go.id]]
  go.genes = GeneIDKey[GeneIDKey$ensembl %in% go.genes, "FBgn"]
  go.genes = intersect(go.genes, names(FC.data))
  sig.go.genes = intersect(go.genes, sigs)
  go.FC = FC.data[go.genes]
  sig.go.FC = FC.data[sig.go.genes]

  data$allup[i] = length(go.FC[go.FC>0])
  data$alldown[i] = length(go.FC[go.FC<0])
  data$up[i] = length(sig.go.FC[sig.go.FC>0])
  data$down[i] = length(sig.go.FC[sig.go.FC<0])
  i = i +1
}

data$down = data$down * -1
data$alldown = data$alldown * -1

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

data$label = factor(data$label, levels = data$label)

fig <- plot_ly(data, 
               x = ~up, y = ~label, 
               type = 'bar',
               name = "Sig. Increased FC",
               marker = list(color = 'brown',
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$term,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nDE up: ", data$up,
                      "\nDE down: ", data$down,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    add_trace(x = ~down, 
              y ~term, 
              marker = list(color = 'steelblue', 
                            line = list(color = "black", width = 1.5)), 
              name = "Sig. Decreased FC")%>%

    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = paste0("Top 20 Most Enriched ",GO.ontology, " GO Terms"),
           margin = m, barmode = 'overlay')

  return(fig)
}

```

```{r, echo = F, include = F}

KEGG20 = function(KEGG.data, FC.data, FDR.data){

data = KEGG.data[, c("category", "Name","adjp", "numDEInCat", "numInCat")]

data$term = data$Name
if(length(unique(data$term)) < length(data$term)){
  data$term[duplicated(data$term)] = paste(data$term[duplicated(data$term)], "2")
}  
data = na.omit(data[data$Name != 'Metabolic pathways',])
data = data[20:1,]
data$Score = -log10(data$adjp)
data$term = factor(data$term, levels = c(unique(data$term)))
data$up = 0
data$down = 0
data$allup = 0
data$alldown = 0

sigs = names(FDR.data[FDR.data <= 0.05])

i = 1
while(i <= nrow(data)){
  kegg.id = data$category[i]
  kegg.genes = names(kegg[grep(kegg.id, kegg)])
  kegg.genes = GeneIDKey[GeneIDKey$Symbol %in% kegg.genes, "FBgn"]
  kegg.genes = intersect(kegg.genes, names(FC.data))
  sig.kegg.genes = intersect(kegg.genes, sigs)
  kegg.FC = FC.data[kegg.genes]
  sig.kegg.FC = FC.data[sig.kegg.genes]
  data$allup[i] = length(kegg.FC[kegg.FC>0])
  data$alldown[i] = length(kegg.FC[kegg.FC<0])
  data$up[i] = length(sig.kegg.FC[sig.kegg.FC>0])
  data$down[i] = length(sig.kegg.FC[sig.kegg.FC<0])
  i = i +1
}

data$down = data$down * -1
data$alldown = data$alldown * -1

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

fig <- plot_ly(data, 
               x = ~up, y = ~term, 
               type = 'bar',
               name = "Sig. Increased FC",
               marker = list(color = 'brown',
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$term,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nDE up: ", data$up,
                      "\nDE down: ", data$down,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    add_trace(x = ~down, 
              y ~term, 
              marker = list(color = 'steelblue', 
                            line = list(color = "black", width = 1.5)), 
              name = "Sig. Decreased FC")%>%
    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = paste0("Top 20 Most Enriched KEGG Terms"),
           margin = m, barmode = 'overlay')

return(fig)

}

```

```{r include = F, echo = F}

sigs = row.names(FDR)[FDR$AvS3MT < 0.05]
FC.data = setNames(FC$AvS3MT, row.names(FC))
FDR.data = setNames(FDR$AvS3MT, row.names(FDR))

bg = row.names(FDR)
sig.cats = Enrichment(sigs, bg)
mt3.cats = sig.cats

```

```{r echo =F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "CC"
go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .005 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

CC.simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

CC.reducedTerms <- reduceSimMatrix(CC.simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r echo =F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "BP"
go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .005 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

BP.simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

BP.reducedTerms <- reduceSimMatrix(BP.simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r echo =F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "MF"
go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .005 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

MF.simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

MF.reducedTerms <- reduceSimMatrix(MF.simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r echo = F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "BP"
BP.GO.fig = GO20(GO.data, GO.ontology, FC.data, FDR.data)

GO.ontology = "CC"
CC.GO.fig = GO20(GO.data, GO.ontology, FC.data, FDR.data)

GO.ontology = "MF"
MF.GO.fig = GO20(GO.data, GO.ontology, FC.data, FDR.data)

KEGG.data = sig.cats[["KEGG"]]
KEGG.fig = KEGG20(KEGG.data, FC.data, FDR.data)

```

Male 3 Day Thorax Biological Process GO Terms
```{r, echo = F,  fig.width= 10}

BP.GO.fig

```

Male 3 Day Thorax Biological Process GO Terms
```{r  echo = F, fig.align='left'}

treemapPlot(BP.reducedTerms)

```

Male 3 Day Thorax Cellular Component GO Terms
```{r, echo = F,  fig.width= 10}

CC.GO.fig

```

Male 3 Day Thorax Cellular Component GO Terms
```{r  echo = F, fig.align='left'}

treemapPlot(CC.reducedTerms)

```

Male 3 Day Thorax Molecular Function GO Terms
```{r, echo = F,  fig.width= 10}

MF.GO.fig

```

Male 3 Day Thorax Molecular Function GO Terms
```{r  echo = F, fig.align='left'}

treemapPlot(MF.reducedTerms)

```

Male 3 Day Thorax KEGG Pathways
```{r, echo = F,  fig.width= 10}

KEGG.fig

```

```{r include = F, echo = F}

sigs = row.names(FDR)[FDR$AvS3FT < 0.05]
FC.data = setNames(FC$AvS3FT, row.names(FC))
FDR.data = setNames(FDR$AvS3FT, row.names(FDR))

bg = row.names(FDR)
sig.cats = Enrichment(sigs, bg)
ft3.cats = sig.cats

```

```{r echo =F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "CC"
go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .005 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

CC.simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

CC.reducedTerms <- reduceSimMatrix(CC.simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r echo =F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "BP"
go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .005 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

BP.simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

BP.reducedTerms <- reduceSimMatrix(BP.simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r echo =F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "MF"
go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .005 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

MF.simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

MF.reducedTerms <- reduceSimMatrix(MF.simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r echo = F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "BP"
BP.GO.fig = GO20(GO.data, GO.ontology, FC.data, FDR.data)

GO.ontology = "CC"
CC.GO.fig = GO20(GO.data, GO.ontology, FC.data, FDR.data)

GO.ontology = "MF"
MF.GO.fig = GO20(GO.data, GO.ontology, FC.data, FDR.data)

KEGG.data = sig.cats[["KEGG"]]
KEGG.fig = KEGG20(KEGG.data, FC.data, FDR.data)

```

Female 3 Day Thorax Biological Process GO Terms
```{r, echo = F,  fig.width= 10}

BP.GO.fig

```

Female 3 Day Thorax Biological Process GO Terms
```{r  echo = F, fig.align='left'}

treemapPlot(BP.reducedTerms)

```

Female 3 Day Thorax Cellular Component GO Terms
```{r, echo = F,  fig.width= 10}

CC.GO.fig

```

Female 3 Day Thorax Cellular Component GO Terms
```{r  echo = F, fig.align='left'}

treemapPlot(CC.reducedTerms)

```

Female 3 Day Thorax Molecular Function GO Terms
```{r, echo = F,  fig.width= 10}

MF.GO.fig

```

Female 3 Day Thorax Molecular Function GO Terms
```{r  echo = F, fig.align='left'}

treemapPlot(MF.reducedTerms)

```

Female 3 Day Thorax KEGG Pathways
```{r, echo = F,  fig.width= 10}

KEGG.fig

```

```{r echo = F, warning=FALSE, message=F}
library(VennDiagram)

set1="Male 9 Day"
set2="Female 9 Day"

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1=row.names(FDR[FDR$AvS9MT <.05,])
s2=row.names(FDR[FDR$AvS9FT <.05,])

main = "9 Day Thorax DEGs"

grid.newpage()
draw.pairwise.venn(area1=length(s1),area2=length(s2),
                   cross.area=length(intersect(s1,s2)),
                   fill = c("royalblue1","brown1"),
                   cex=1.5,
                   cat.cex = 1,
                   category = c(set1,set2),
                   main = "main",
                   cat.dist = c(.065,.05),
                   cat.pos = c(-20,20))

```

```{r include = F, echo = F}

sigs = row.names(FDR)[FDR$AvS9MT < 0.05]
FC.data = setNames(FC$AvS9MT, row.names(FC))
FDR.data = setNames(FDR$AvS9MT, row.names(FDR))

bg = row.names(FDR)
sig.cats = Enrichment(sigs, bg)
mt9.cats = sig.cats

```

```{r echo =F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "CC"
go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .005 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

CC.simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

CC.reducedTerms <- reduceSimMatrix(CC.simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r echo =F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "BP"
go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .005 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

BP.simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

BP.reducedTerms <- reduceSimMatrix(BP.simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r echo =F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "MF"
go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .005 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

MF.simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

MF.reducedTerms <- reduceSimMatrix(MF.simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r echo = F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "BP"
BP.GO.fig = GO20(GO.data, GO.ontology, FC.data, FDR.data)

GO.ontology = "CC"
CC.GO.fig = GO20(GO.data, GO.ontology, FC.data, FDR.data)

GO.ontology = "MF"
MF.GO.fig = GO20(GO.data, GO.ontology, FC.data, FDR.data)

KEGG.data = sig.cats[["KEGG"]]
KEGG.fig = KEGG20(KEGG.data, FC.data, FDR.data)

```

Male 9 Day Thorax Biological Process GO Terms
```{r, echo = F,  fig.width= 10}

BP.GO.fig

```

Male 9 Day Thorax Biological Process GO Terms
```{r  echo = F, fig.align='left'}

treemapPlot(BP.reducedTerms)

```

Male 3 Day Thorax Cellular Component GO Terms
```{r, echo = F,  fig.width= 10}

CC.GO.fig

```

Male 9 Day Thorax Cellular Component GO Terms
```{r  echo = F, fig.align='left'}

treemapPlot(CC.reducedTerms)

```

Male 9 Day Thorax Molecular Function GO Terms
```{r, echo = F,  fig.width= 10}

MF.GO.fig

```

Male 9 Day Thorax Molecular Function GO Terms
```{r  echo = F, fig.align='left'}

treemapPlot(MF.reducedTerms)

```

Male 9 Day Thorax KEGG Pathways
```{r, echo = F,  fig.width= 10}

KEGG.fig

```

```{r include = F, echo = F}

sigs = row.names(FDR)[FDR$AvS9FT < 0.05]
FC.data = setNames(FC$AvS9FT, row.names(FC))
FDR.data = setNames(FDR$AvS9FT, row.names(FDR))

bg = row.names(FDR)
sig.cats = Enrichment(sigs, bg)
ft9.cats = sig.cats

```

```{r echo =F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "CC"
go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .005 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

CC.simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

CC.reducedTerms <- reduceSimMatrix(CC.simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r echo =F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "BP"
go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .005 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

BP.simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

BP.reducedTerms <- reduceSimMatrix(BP.simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r echo =F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "MF"
go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .005 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

MF.simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

MF.reducedTerms <- reduceSimMatrix(MF.simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r echo = F, include =F}

GO.data = sig.cats[["GO"]]
GO.ontology = "BP"
BP.GO.fig = GO20(GO.data, GO.ontology, FC.data, FDR.data)

GO.ontology = "CC"
CC.GO.fig = GO20(GO.data, GO.ontology, FC.data, FDR.data)

GO.ontology = "MF"
MF.GO.fig = GO20(GO.data, GO.ontology, FC.data, FDR.data)

KEGG.data = sig.cats[["KEGG"]]
KEGG.fig = KEGG20(KEGG.data, FC.data, FDR.data)

```

Female 9 Day Thorax Biological Process GO Terms
```{r, echo = F,  fig.width= 10}

BP.GO.fig

```

Female 9 Day Thorax Biological Process GO Terms
```{r  echo = F, fig.align='left'}

treemapPlot(BP.reducedTerms)

```

Female 9 Day Thorax Cellular Component GO Terms
```{r, echo = F,  fig.width= 10}

CC.GO.fig

```

Female 9 Day Thorax Cellular Component GO Terms
```{r  echo = F, fig.align='left'}

treemapPlot(CC.reducedTerms)

```

Female 9 Day Thorax Molecular Function GO Terms
```{r, echo = F,  fig.width= 10}

MF.GO.fig

```

Female 9 Day Thorax Molecular Function GO Terms
```{r  echo = F, fig.align='left'}

treemapPlot(MF.reducedTerms)

```

Female 9 Day Thorax KEGG Pathways
```{r, echo = F,  fig.width= 10}

KEGG.fig

```

```{r echo = F}

set1="Male 3 Day"
set2="Female 3 Day"
set3="Male 9 Day"
set4="Female 40 Day"

s1=row.names(FDR[FDR$AvS3MT <.05,])
s2=row.names(FDR[FDR$AvS3FT <.05,])
s3=row.names(FDR[FDR$AvS9MT <.05,])
s4=row.names(FDR[FDR$AvS40FT <.05,])

main = "3 and 9 Day Thorax "

grid.newpage()
tempvenn=draw.quad.venn(area1=length(s1),area2=length(s2),area3=length(s3),area4=length(s4), n12=length(intersect(s1,s2)),n13 = length(intersect(s1,s3)),n14 = length(intersect(s1,s4)),n23 = length(intersect(s2,s3)),n24 = length(intersect(s2,s4)),n34 = length(intersect(s3,s4)),n123 =  length(intersect(s1,intersect(s2,s3))),n124 =  length(intersect(s1,intersect(s2,s4))), n134 = length(intersect(s1,intersect(s3,s4))),n234 = length(intersect(s2,intersect(s3,s4))),n1234 = length(intersect(s1,intersect(s2,intersect(s3,s4)))),category =c(set1,set2,set3,set4),fill = c("royalblue1", "brown1","forestgreen","yellow2"),cex=2,cat.cex = 1, title = "main")

```