### 2024-05-08 Anna qPCR data/AnnaThesisA4VRNAseqFigures.Rmd

---
title: "Anna Thesis A4V RNAseq KEGG Pathway Figures"
author: "John Santiago"
date: "2024-03-27"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<br><br><br>

```{r echo = F, include = F}
library(plotly)
library(org.Dm.eg.db)
library(pathview)
library(visNetwork)
library(org.Dm.eg.db)

##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"
cpmdata = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
groups  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
TKT.EdgeR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)
TKT.EdgeR.FC = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FCTable.csv"),row.names = 1)

GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)
kegg = readRDS(gzcon(url(paste0(git.dir,"GeneralDataFiles/kegg.symbol2path.RData"))))
kegg.names = read.csv(paste0(git.dir, "GeneralDataFiles/KEGG.names.csv"), row.names = 1)

genesingo=as.list(org.Dm.egGO2ALLEGS)

##mean.cpm; Mean cpm for each gene in each condition (groups$Group). Used for filtering
mean.cpm = matrix(0, ncol = length(unique(groups$Group)), nrow = nrow(cpmdata))
colnames(mean.cpm) = unique(groups$Group)
row.names(mean.cpm) = row.names(cpmdata)
i=1
while(i<=ncol(mean.cpm)){
  mean.cpm[,i]= apply(cpmdata[,row.names(groups[groups$Group == colnames(mean.cpm)[i],])], 1, mean)
  i=i+1
}

i=1
for(i in 1:ncol(TKT.EdgeR)){
  TKT.EdgeR[TKT.EdgeR[,i] == 0,i] = 2
  i=i+1
}

rawdata = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/MetabolomicsForMetaboanalyst.txt", sep = "\t", row.names = 1)

KEGG.Key = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/RawMetabolomicData.csv", row.names = 1)
KEGG.Key = setNames(KEGG.Key[-c(1:2), 2], row.names(KEGG.Key)[-c(1:2)])

genotypes = setNames(c("WT", "GR", "TktOEWT", "TktOEGR", "TktDfWT", "TktDfGR"), c("C", "E", "A", "F", "B", "D"))

metadata = data.frame(Sample = colnames(rawdata), 
                      Group = as.character(rawdata[1,]), 
                      TIC = as.numeric(rawdata[2,]), 
                      Genotypes = genotypes[as.character(rawdata[1,])])

row.names(metadata) = metadata$Sample

norm.func = function(data){
  nd = (as.numeric(data)/metadata$TIC)*1000
  return(nd)
}

norm.data = t(apply(rawdata[3:nrow(rawdata),], 1, norm.func))
colnames(norm.data) = colnames(rawdata)

group.mean=function(data){
  gm = mean(na.omit(data))
}

mean.data = matrix(0, nrow=nrow(norm.data), ncol = length(genotypes))
row.names(mean.data) = row.names(norm.data)
colnames(mean.data) = unique(metadata$Group)

i=1
while(i<=ncol(mean.data)){
  mean.data[,i] = apply(norm.data[,metadata[metadata$Group == colnames(mean.data)[i], "Sample"]], 1, group.mean)
  i = i+1
}

mean.cpm = as.data.frame(mean.cpm)

compare.conditions = function(condition1, condition2){
  columns1 = metadata[metadata$Group == condition1, "Sample"]
  columns2 = metadata[metadata$Group == condition2, "Sample"]
  p=setNames(rep(NA, nrow(norm.data)), row.names(norm.data))
  i=1
  while(i<=length(p)){
    if(length(na.omit(norm.data[i,columns1]))>2 & 
       length(na.omit(norm.data[i,columns2]))>2){
      temp = t.test(na.omit(norm.data[i,columns1]),
                    na.omit(norm.data[i,columns2]), paired = F, var.equal = T)
      p[i] = temp[[3]]
      }
    i=i+1
    }

  fdr = p.adjust(p, "BH", length(p))

  fc = mean.data[,condition1]/mean.data[,condition2]

  comparison.table = data.frame(p = p, 
                                FDR = fdr, 
                                FC = fc, 
                                KEGG = KEGG.Key[row.names(mean.data)])

  comparison.table = comparison.table[order(comparison.table$FDR),]
}

GRxWT.C = compare.conditions("E", "C")
GRxWT.Df = compare.conditions("D", "B")
GRxWT.OE = compare.conditions("F", "A")

GR.CxDf = compare.conditions("D", "E")
WT.CxDf = compare.conditions("B", "C")

GR.CxOE = compare.conditions("F", "E")
WT.CxOE = compare.conditions("A", "C")

##Comparison cpm means
ccm = data.frame(GRF.CxDf = apply(mean.cpm[,c("TktDfGR.F","GR.F")], MARGIN = 1, mean), 
                 GRF.CxOE = apply(mean.cpm[,c("TktOEGR.F","GR.F")], MARGIN = 1, mean),
                 WTF.CxDf = apply(mean.cpm[,c("TktDfWT.F","WT.F")], MARGIN = 1, mean),
                 WTF.CxOE = apply(mean.cpm[,c("TktOEWT.F","WT.F")], MARGIN = 1, mean),
                 GRxWT.FC = apply(mean.cpm[,c("GR.F","WT.F")], MARGIN = 1, mean), 
                 GRxWT.FDf = apply(mean.cpm[,c("TktDfGR.F","TktDfWT.F")], MARGIN = 1, mean),
                 GRxWT.FOE = apply(mean.cpm[,c("TktOEGR.F","TktOEWT.F")], MARGIN = 1, mean), 
                 GRM.CxDf = apply(mean.cpm[,c("TktDfGR.M","GR.M")], MARGIN = 1, mean), 
                 GRM.CxOE = apply(mean.cpm[,c("TktOEGR.M","GR.M")], MARGIN = 1, mean),
                 WTM.CxDf = apply(mean.cpm[,c("TktDfWT.M","WT.M")], MARGIN = 1, mean),
                 WTM.CxOE = apply(mean.cpm[,c("TktOEWT.M","WT.M")], MARGIN = 1, mean),
                 GRxWT.MC = apply(mean.cpm[,c("GR.M","WT.M")], MARGIN = 1, mean), 
                 GRxWT.MDf = apply(mean.cpm[,c("TktDfGR.M","TktDfWT.M")], MARGIN = 1, mean),
                 GRxWT.MOE = apply(mean.cpm[,c("TktOEGR.M","TktOEWT.M")], MARGIN = 1, mean),
                 GR.FxM = apply(mean.cpm[,c("GR.F","GR.M")], MARGIN = 1, mean),
                 WT.FxM = apply(mean.cpm[,c("WT.F","WT.M")], MARGIN = 1, mean),
                 row.names = row.names(mean.cpm))

comparison.means = list(GRF.CxDf = c("TktDfGR.F","GR.F"), 
                 GRF.CxOE = c("TktOEGR.F","GR.F"),
                 WTF.CxDf = c("TktDfWT.F","WT.F"),
                 WTF.CxOE = c("TktOEWT.F","WT.F"),
                 GRxWT.FC = c("GR.F","WT.F"), 
                 GRxWT.FDf = c("TktDfGR.F","TktDfWT.F"),
                 GRxWT.FOE = c("TktOEGR.F","TktOEWT.F"), 
                 GRM.CxDf = c("TktDfGR.M","GR.M"), 
                 GRM.CxOE = c("TktOEGR.M","GR.M"),
                 WTM.CxDf = c("TktDfWT.M","WT.M"),
                 WTM.CxOE = c("TktOEWT.M","WT.M"),
                 GRxWT.MC = c("GR.M","WT.M"), 
                 GRxWT.MDf = c("TktDfGR.M","TktDfWT.M"),
                 GRxWT.MOE = c("TktOEGR.M","TktOEWT.M"),
                 GR.FxM = c("GR.F","GR.M"),
                 WT.FxM = c("WT.F","WT.M"))

metab.id = setNames(row.names(GRxWT.C), GRxWT.C$KEGG)
mixed.id = unlist(strsplit(names(metab.id)[grep(" / ", names(metab.id))], " / "))
mixed.id = mixed.id[mixed.id != "n/a"]
mixed.id = setNames(rep("", length(mixed.id)), mixed.id)

i = 1
while(i<=length(mixed.id)){
  mixed.id[i] = metab.id[grep(names(mixed.id)[i], names(metab.id))]
  i = i +1
}

metab.id = metab.id[-grep(" / ", names(metab.id))]

metab.id = c(metab.id, mixed.id)

```

```{r echo = F}
KEGG.diagram = function(PathwayID, out.id, FC.data, FDR.data){
  ##KEGG Mapping Function
  if(length(intersect(list.files(), "KEGG_Image_Files")) == 0){
    dir.create(paste0(dirname(rstudioapi::getSourceEditorContext()$path), "/KEGG_Image_Files/"))
    }
  Home = dirname(rstudioapi::getSourceEditorContext()$path)
  OutputDir = paste0(Home, "/KEGG_Image_Files/")
  setwd(OutputDir)
  sigs = names(FDR.data)
  ##sigs = names(FDR.data[FDR.data <= .05])

  if(class(FC.data)=="numeric"){
    deg.data=setNames(FC.data[sigs],GeneIDKey[sigs, "ensembl"])
  }

  if(class(FC.data)=="data.frame"){
    deg.data = as.matrix(FC.data)
    row.names(deg.data) = GeneIDKey[row.names(FC.data), "ensembl"]
  }

  metab.data = setNames(log2(metab.data[metab.id, 'FC']), names(metab.id))

  pv.out <- pathview(gene.data =  deg.data,
                     ##cpd.data = metab.data,
                   pathway.id = PathwayID,
                   species = "dme",
                   kegg.native = T,
                   limit=list(gene=c(-2,2)),
                   node.sum="max.abs",
                   low="royalblue",mid = "grey",high="firebrick",
                   out.suffix = out.id,
                   bins=list(genes=30),
                   plot.col.key = F,
                   match.data = F)
  setwd(Home)

    if(class(FC.data)=="numeric"){
      img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".png")
    }

    if(class(FC.data)=="data.frame"){
      img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".multi.png")
    }
  knitr::include_graphics(img.file)
}

```

```{r echo = F, message=F, fig.width=7}

PathwayID = "00010"
out.id = 'G85R_F'
FC.data = setNames(TKT.EdgeR.FC$GRxWT.FC, row.names(TKT.EdgeR.FC))
FDR.data = setNames(TKT.EdgeR$GRxWT.FC, row.names(TKT.EdgeR))

FC.data = TKT.EdgeR.FC[,c('GRxWT.FC', 'GRxWT.MC')]
FDR.data = setNames(TKT.EdgeR$GRxWT.MC, row.names(TKT.EdgeR))

metab.data = GRxWT.C
KEGG.diagram(PathwayID, out.id, FC.data, FDR.data)

```

```{r echo = F, message=F, fig.width=7}

PathwayID = "00020"
out.id = 'G85R_F'
FC.data = setNames(TKT.EdgeR.FC$GRxWT.FC, row.names(TKT.EdgeR.FC))
FDR.data = setNames(TKT.EdgeR$GRxWT.FC, row.names(TKT.EdgeR))

FC.data = TKT.EdgeR.FC[,c('GRxWT.FC', 'GRxWT.MC')]
FDR.data = setNames(TKT.EdgeR$GRxWT.MC, row.names(TKT.EdgeR))

metab.data = GRxWT.C
KEGG.diagram(PathwayID, out.id, FC.data, FDR.data)

```

```{r echo = F, message=F, fig.width=7}

PathwayID = "00061"
out.id = 'G85R_F'
FC.data = setNames(TKT.EdgeR.FC$GRxWT.FC, row.names(TKT.EdgeR.FC))
FDR.data = setNames(TKT.EdgeR$GRxWT.FC, row.names(TKT.EdgeR))

FC.data = TKT.EdgeR.FC[,c('GRxWT.FC', 'GRxWT.MC')]
FDR.data = setNames(TKT.EdgeR$GRxWT.MC, row.names(TKT.EdgeR))

metab.data = GRxWT.C
KEGG.diagram(PathwayID, out.id, FC.data, FDR.data)

```

```{r echo = F, message=F, fig.width=7}

PathwayID = "04214"
out.id = 'G85R_F'
FC.data = setNames(TKT.EdgeR.FC$GRxWT.FC, row.names(TKT.EdgeR.FC))
FDR.data = setNames(TKT.EdgeR$GRxWT.FC, row.names(TKT.EdgeR))

FC.data = TKT.EdgeR.FC[,c('GRxWT.FC', 'GRxWT.MC')]
FDR.data = setNames(TKT.EdgeR$GRxWT.MC, row.names(TKT.EdgeR))

metab.data = GRxWT.C
KEGG.diagram(PathwayID, out.id, FC.data, FDR.data)

```

```{r echo = F, message=F}

PathwayID = "04137"
out.id = 'G85R_F'
FC.data = setNames(TKT.EdgeR.FC$GRxWT.FC, row.names(TKT.EdgeR.FC))
FDR.data = setNames(TKT.EdgeR$GRxWT.FC, row.names(TKT.EdgeR))

FC.data = TKT.EdgeR.FC[,c('GRxWT.FC', 'GRxWT.MC')]
FDR.data = setNames(TKT.EdgeR$GRxWT.MC, row.names(TKT.EdgeR))

metab.data = GRxWT.C
KEGG.diagram(PathwayID, out.id, FC.data, FDR.data)

```