### 2024-05-08 RussoPaper/RussoClimbingPlots.Rmd

---
title: "Russo Manuscript Climbing Plot"
author: "John Santiago"
date: "2024-03-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = F, echo = F}

library(plotly)
library(edgeR)

```

```{r include = F, echo = F}

fig4ab = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 4AB Climbing Data ok371 cha gbb climbing cm_sec.csv", header = T)

female = fig4ab[grep('f', fig4ab$vial_ID),]
male = fig4ab[grep('m', fig4ab$vial_ID),]

f.con = na.omit(female[female$geno == 'silent' | female$geno == 'A4V',])
f.cha = na.omit(female[grep('cha', female$geno),])

fallcha = rbind(f.con, f.cha)
fallcha$sample = paste0(fallcha$geno, fallcha$day)

fcha.mean = unique(fallcha[,c('geno', 'day')])
fcha.mean$sample = paste0(fcha.mean$geno, fcha.mean$day)
fcha.mean$mean = 0

i=1
while(i<=nrow(fcha.mean)){
  fcha.mean$mean[i] = mean(fallcha[fallcha$sample == fcha.mean$sample[i], 'corrected.values'])
  i = i+1
}

fcha.mean = fcha.mean[order(fcha.mean$day),]

```

```{r echo = F, include = F}

plot_ly(x = fallcha$day,
        y = fallcha$corrected.values,
        name = fallcha$geno,
        type = 'scatter',
        color = fallcha$color,
        marker = list(colors = fallcha$color)) %>%
  add_lines(x = fcha.mean$day,
            y = fcha.mean$mean,
            name = fcha.mean$geno,
            color = fcha.mean$color,
            line = list(color = fcha.mean$color))

```

```{r echo = F}

female = fig4ab[grep('f', fig4ab$vial_ID),]
female$sample = paste0(female$geno, female$day)

fmean = unique(female[,c('geno', 'day', 'sample')])
fstd = fmean
eb = fmean
fmean$mean = 0
fstd$std = 0

i=1
while(i<=nrow(fmean)){
  fmean$mean[i] = mean(na.omit(female[female$sample == fmean$sample[i], 'corrected.values']))

  fstd$std[i] = (var(na.omit(female[female$sample == fmean$sample[i], 'corrected.values'])))

  i = i+1
}

flm = data.frame(geno = unique(female$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(flm)){
  temp=na.omit(female[female$geno == flm$geno[i], ])
  temp.lm = lm(temp$corrected.values~temp$day)
  flm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  flm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(flm) = flm$geno

fmean = fmean[order(fmean$day),]
fstd = fstd[order(fstd$day),]
eb = eb[order(eb$day),]

eb$y0 = fmean$mean - fstd$std
eb$y1 = fmean$mean + fstd$std

```

```{r echo = F}

norm.female = na.omit(female)
norm.female$normval = 0
i=1
while(i <= nrow(norm.female)){
  norm.female$normval[i] = norm.female$corrected.values[i]/
    fmean$mean[fmean$geno == 'silent' & fmean$day == norm.female$day[i]]
  i = i +1
}

norm.fmean = unique(norm.female[,c('geno', 'day', 'sample')])
norm.fstd = norm.fmean
norm.feb = norm.fmean
norm.fmean$mean = 0
norm.fstd$std = 0

i=1
while(i<=nrow(norm.fmean)){
  norm.fmean$mean[i] = mean(na.omit(norm.female[norm.female$sample == norm.fmean$sample[i], 'normval']))

  norm.fstd$std[i] = (var(na.omit(norm.female[norm.female$sample == norm.fmean$sample[i], 'normval'])))

  i = i+1
}

norm.fmean = norm.fmean[order(norm.fmean$day),]
norm.fstd = norm.fstd[order(norm.fstd$day),]
norm.feb = norm.feb[order(norm.feb$day),]

norm.feb$y0 = norm.fmean$mean - norm.fstd$std
norm.feb$y1 = norm.fmean$mean + norm.fstd$std

norm.flm = data.frame(geno = unique(norm.female$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(norm.flm)){
  temp=na.omit(norm.female[norm.female$geno == norm.flm$geno[i], ])
  temp.lm = lm(temp$normval~temp$day)
  norm.flm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  norm.flm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(norm.flm) = norm.flm$geno

```

```{r echo = F, fig.height=10}

silent = norm.fmean[norm.fmean$geno == "silent", ]
a4v = norm.fmean[norm.fmean$geno == "A4V", ]
silentcha = norm.fmean[norm.fmean$geno == "silentcha", ]
a4vcha = norm.fmean[norm.fmean$geno == "A4Vcha", ]
silentok = norm.fmean[norm.fmean$geno == "silentok", ]
a4vok = norm.fmean[norm.fmean$geno == "A4Vok", ]

ls = norm.fmean[norm.fmean$geno == 'silent',]
la = norm.fmean[norm.fmean$geno == 'A4V',]
lsc = norm.fmean[norm.fmean$geno == 'silentcha',]
lac = norm.fmean[norm.fmean$geno == 'A4Vcha',]
lso = norm.fmean[norm.fmean$geno == 'silentok',]
lao = norm.fmean[norm.fmean$geno == 'A4Vok',]

pcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 9.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0,1.85),
     xlim = c(2,27),
     ylab = 'Relative Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     main = "Females")

axis(1, at=silent$day)

abline(v = 2, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 6, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 9, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 13, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 19, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 22, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 27, lty =2, lwd = 1, col = hex(pcol[7], .5))

lines(x = norm.flm['silent', c("x0", "x1")],
      y = norm.flm['silent', c("y0", "y1")],
      type = 'l',
      col = pcol[1],
      lwd = 2)

lines(x = norm.flm['A4V', c("x0", "x1")],
      y = norm.flm['A4V', c("y0", "y1")],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = norm.flm['silentcha', c("x0", "x1")],
      y = norm.flm['silentcha', c("y0", "y1")],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = norm.flm['A4Vcha', c("x0", "x1")],
      y = norm.flm['A4Vcha', c("y0", "y1")],
      type = 'l',
      col = pcol[4],
      lwd = 2)

lines(x = norm.flm['silentok', c("x0", "x1")],
      y = norm.flm['silentok', c("y0", "y1")],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = norm.flm['A4Vok', c("x0", "x1")],
      y = norm.flm['A4Vok', c("y0", "y1")],
      type = 'l',
      col = pcol[6],
      lwd = 2)

arrows(x0 = norm.feb$day[norm.feb$geno == 'silent'],
       x1 = norm.feb$day[norm.feb$geno == 'silent'],
       y0 = norm.feb$y0[norm.feb$geno == 'silent'],
       y1 = norm.feb$y1[norm.feb$geno == 'silent'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[1], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'A4V'],
       x1 = norm.feb$day[norm.feb$geno == 'A4V'],
       y0 = norm.feb$y0[norm.feb$geno == 'A4V'],
       y1 = norm.feb$y1[norm.feb$geno == 'A4V'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[2], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'silentcha'],
       x1 = norm.feb$day[norm.feb$geno == 'silentcha'],
       y0 = norm.feb$y0[norm.feb$geno == 'silentcha'],
       y1 = norm.feb$y1[norm.feb$geno == 'silentcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[3], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'A4Vcha'],
       x1 = norm.feb$day[norm.feb$geno == 'A4Vcha'],
       y0 = norm.feb$y0[norm.feb$geno == 'A4Vcha'],
       y1 = norm.feb$y1[norm.feb$geno == 'A4Vcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[4], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'silentok'],
       x1 = norm.feb$day[norm.feb$geno == 'silentok'],
       y0 = norm.feb$y0[norm.feb$geno == 'silentok'],
       y1 = norm.feb$y1[norm.feb$geno == 'silentok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[5], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'A4Vok'],
       x1 = norm.feb$day[norm.feb$geno == 'A4Vok'],
       y0 = norm.feb$y0[norm.feb$geno == 'A4Vok'],
       y1 = norm.feb$y1[norm.feb$geno == 'A4Vok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[6], 
       lwd=1)

points(x = silent$day,
       y = silent$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4v$day,
       y = a4v$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)

points(x = silent$day,
       y = silent$mean,
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = a4v$day,
       y = a4v$mean,
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex =2)

legend(x = 29,
       y = 1.8,
       legend = c('Silent', 'A4V',
                  'ChAT>gbb;Silent', 'ChAT>gbb;A4V', 
                  'OK371>gbb;Silent', 'OK371>gbb;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(pcol[1], pcol[2], pcol[3],
                pcol[4], pcol[5], pcol[6]),
       bty = 'n',
       y.intersp = 1,
       pt.cex = 2,
       xpd=TRUE)

```

```{r echo = F}

male$sample = paste0(male$geno, male$day)

norm.male = na.omit(male)
norm.male$normval = 0

mmean = unique(male[,c('geno', 'day', 'sample')])
mstd = mmean
eb = mmean
mmean$mean = 0
mstd$std = 0

i=1
while(i<=nrow(mmean)){
  mmean$mean[i] = mean(na.omit(male[male$sample == mmean$sample[i], 'corrected.values']))

  mstd$std[i] = (var(na.omit(male[male$sample == mmean$sample[i], 'corrected.values'])))

  i = i+1
}

mlm = data.frame(geno = unique(male$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(mlm)){
  temp=na.omit(male[male$geno == mlm$geno[i], ])
  temp.lm = lm(temp$corrected.values~temp$day)
  mlm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  mlm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(mlm) = mlm$geno

mmean = mmean[order(mmean$day),]
mstd = mstd[order(mstd$day),]
eb = eb[order(eb$day),]

eb$y0 = mmean$mean - mstd$std
eb$y1 = mmean$mean + mstd$std

i=1
while(i <= nrow(norm.male)){
  norm.male$normval[i] = norm.male$corrected.values[i]/
    mmean$mean[mmean$geno == 'silent' & mmean$day == norm.male$day[i]]
  i = i +1
}

norm.mmean = unique(norm.male[,c('geno', 'day', 'sample')])
norm.mstd = norm.mmean
norm.eb = norm.mmean
norm.mmean$mean = 0
norm.mstd$std = 0

i=1
while(i<=nrow(norm.mmean)){
  norm.mmean$mean[i] = mean(na.omit(norm.male[norm.male$sample == norm.mmean$sample[i], 'normval']))

  norm.mstd$std[i] = (var(na.omit(norm.male[norm.male$sample == norm.mmean$sample[i], 'normval'])))

  i = i+1
}

norm.mmean = norm.mmean[order(norm.mmean$day),]
norm.mstd = norm.mstd[order(norm.mstd$day),]
norm.eb = norm.eb[order(norm.eb$day),]

norm.eb$y0 = norm.mmean$mean - norm.mstd$std
norm.eb$y1 = norm.mmean$mean + norm.mstd$std

norm.mlm = data.frame(geno = unique(norm.male$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(norm.mlm)){
  temp=na.omit(norm.male[norm.male$geno == norm.mlm$geno[i], ])
  temp.lm = lm(temp$normval~temp$day)
  norm.mlm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  norm.mlm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(norm.mlm) = norm.mlm$geno

```

```{r echo = F, fig.height=10}

silent = norm.mmean[norm.mmean$geno == "silent", ]
a4v = norm.mmean[norm.mmean$geno == "A4V", ]
silentcha = norm.mmean[norm.mmean$geno == "silentcha", ]
a4vcha = norm.mmean[norm.mmean$geno == "A4Vcha", ]
silentok = norm.mmean[norm.mmean$geno == "silentok", ]
a4vok = norm.mmean[norm.mmean$geno == "A4Vok", ]

ls = norm.mmean[norm.mmean$geno == 'silent',]
la = norm.mmean[norm.mmean$geno == 'A4V',]
lsc = norm.mmean[norm.mmean$geno == 'silentcha',]
lac = norm.mmean[norm.mmean$geno == 'A4Vcha',]
lso = norm.mmean[norm.mmean$geno == 'silentok',]
lao = norm.mmean[norm.mmean$geno == 'A4Vok',]

pcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 9.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0,1.85),
     xlim = c(2,27),
     ylab = 'Relative Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     main = "Males")

axis(1, at=silent$day)

abline(v = 2, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 6, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 9, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 13, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 19, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 22, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 27, lty =2, lwd = 1, col = hex(pcol[7], .5))

lines(x = norm.mlm['silent', c("x0", "x1")],
      y = norm.mlm['silent', c("y0", "y1")],
      type = 'l',
      col = pcol[1],
      lwd = 2)

lines(x = norm.mlm['A4V', c("x0", "x1")],
      y = norm.mlm['A4V', c("y0", "y1")],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = norm.mlm['silentcha', c("x0", "x1")],
      y = norm.mlm['silentcha', c("y0", "y1")],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = norm.mlm['A4Vcha', c("x0", "x1")],
      y = norm.mlm['A4Vcha', c("y0", "y1")],
      type = 'l',
      col = pcol[4],
      lwd = 2)

lines(x = norm.mlm['silentok', c("x0", "x1")],
      y = norm.mlm['silentok', c("y0", "y1")],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = norm.mlm['A4Vok', c("x0", "x1")],
      y = norm.mlm['A4Vok', c("y0", "y1")],
      type = 'l',
      col = pcol[6],
      lwd = 2)

arrows(x0 = norm.eb$day[norm.eb$geno == 'silent'],
       x1 = norm.eb$day[norm.eb$geno == 'silent'],
       y0 = norm.eb$y0[norm.eb$geno == 'silent'],
       y1 = norm.eb$y1[norm.eb$geno == 'silent'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[1], 
       lwd=1)

arrows(x0 = norm.eb$day[norm.eb$geno == 'A4V'],
       x1 = norm.eb$day[norm.eb$geno == 'A4V'],
       y0 = norm.eb$y0[norm.eb$geno == 'A4V'],
       y1 = norm.eb$y1[norm.eb$geno == 'A4V'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[2], 
       lwd=1)

arrows(x0 = norm.eb$day[norm.eb$geno == 'silentcha'],
       x1 = norm.eb$day[norm.eb$geno == 'silentcha'],
       y0 = norm.eb$y0[norm.eb$geno == 'silentcha'],
       y1 = norm.eb$y1[norm.eb$geno == 'silentcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[3], 
       lwd=1)

arrows(x0 = norm.eb$day[norm.eb$geno == 'A4Vcha'],
       x1 = norm.eb$day[norm.eb$geno == 'A4Vcha'],
       y0 = norm.eb$y0[norm.eb$geno == 'A4Vcha'],
       y1 = norm.eb$y1[norm.eb$geno == 'A4Vcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[4], 
       lwd=1)

arrows(x0 = norm.eb$day[norm.eb$geno == 'silentok'],
       x1 = norm.eb$day[norm.eb$geno == 'silentok'],
       y0 = norm.eb$y0[norm.eb$geno == 'silentok'],
       y1 = norm.eb$y1[norm.eb$geno == 'silentok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[5], 
       lwd=1)

arrows(x0 = norm.eb$day[norm.eb$geno == 'A4Vok'],
       x1 = norm.eb$day[norm.eb$geno == 'A4Vok'],
       y0 = norm.eb$y0[norm.eb$geno == 'A4Vok'],
       y1 = norm.eb$y1[norm.eb$geno == 'A4Vok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[6], 
       lwd=1)

points(x = silent$day,
       y = silent$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4v$day,
       y = a4v$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)

points(x = silent$day,
       y = silent$mean,
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = a4v$day,
       y = a4v$mean,
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex =2)

legend(x = 29,
       y = 1.8,
       legend = c('Silent', 'A4V',
                  'ChAT>gbb;Silent', 'ChAT>gbb;A4V', 
                  'OK371>gbb;Silent', 'OK371>gbb;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(pcol[1], pcol[2], pcol[3],
                pcol[4], pcol[5], pcol[6]),
       bty = 'n',
       y.intersp = 1,
       pt.cex = 2,
       xpd=TRUE)

```

```{r  echo = F}

fig4cd = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 4CD.csv", header = T)

fig4cd = fig4cd[fig4cd$day != 9,]
fig4cd = fig4cd[fig4cd$day != 32,]

female = fig4cd[grep('f', fig4cd$vial_ID),]
male = fig4cd[grep('m', fig4cd$vial_ID),]

female$sample = paste0(female$geno, female$day)
male$sample = paste0(male$geno, male$day)

norm.female = na.omit(female)
norm.female$normval = 0

fmean = unique(female[,c('geno', 'day', 'sample')])
fstd = fmean
feb = fmean
fmean$mean = 0
fstd$std = 0

norm.male = na.omit(male)
norm.male$normval = 0

mmean = unique(male[,c('geno', 'day', 'sample')])
mstd = mmean
meb = mmean
mmean$mean = 0
mstd$std = 0

i=1
while(i<=nrow(mmean)){
  mmean$mean[i] = mean(na.omit(male[male$sample == mmean$sample[i], 'corrected.values']))

  mstd$std[i] = (var(na.omit(male[male$sample == mmean$sample[i], 'corrected.values'])))

  i = i+1
}

i=1
while(i<=nrow(fmean)){
  fmean$mean[i] = mean(na.omit(female[female$sample == fmean$sample[i], 'corrected.values']))

  fstd$std[i] = (var(na.omit(female[female$sample == fmean$sample[i], 'corrected.values'])))

  i = i+1
}

mlm = data.frame(geno = unique(male$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(mlm)){
  temp=na.omit(male[male$geno == mlm$geno[i], ])
  temp.lm = lm(temp$corrected.values~temp$day)
  mlm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  mlm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(mlm) = mlm$geno

flm = data.frame(geno = unique(female$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(flm)){
  temp=na.omit(female[female$geno == flm$geno[i], ])
  temp.lm = lm(temp$corrected.values~temp$day)
  flm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  flm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(flm) = flm$geno

mmean = mmean[order(mmean$day),]
mstd = mstd[order(mstd$day),]
meb = meb[order(meb$day),]

meb$y0 = mmean$mean - mstd$std
meb$y1 = mmean$mean + mstd$std

fmean = fmean[order(fmean$day),]
fstd = fstd[order(fstd$day),]
feb = feb[order(feb$day),]

feb$y0 = fmean$mean - fstd$std
feb$y1 = fmean$mean + fstd$std

i=1
while(i <= nrow(norm.male)){
  norm.male$normval[i] = norm.male$corrected.values[i]/
    mmean$mean[mmean$geno == 'silent' & mmean$day == norm.male$day[i]]
  i = i +1
}

norm.mmean = unique(norm.male[,c('geno', 'day', 'sample')])
norm.mstd = norm.mmean
norm.meb = norm.mmean
norm.mmean$mean = 0
norm.mstd$std = 0

i=1
while(i <= nrow(norm.female)){
  norm.female$normval[i] = norm.female$corrected.values[i]/
    fmean$mean[fmean$geno == 'silent' & fmean$day == norm.female$day[i]]
  i = i +1
}

norm.fmean = unique(norm.female[,c('geno', 'day', 'sample')])
norm.fstd = norm.fmean
norm.feb = norm.fmean
norm.fmean$mean = 0
norm.fstd$std = 0

i=1
while(i<=nrow(norm.mmean)){
  norm.mmean$mean[i] = mean(na.omit(norm.male[norm.male$sample == norm.mmean$sample[i], 'normval']))

  norm.mstd$std[i] = (var(na.omit(norm.male[norm.male$sample == norm.mmean$sample[i], 'normval'])))

  i = i+1
}

i=1
while(i<=nrow(norm.fmean)){
  norm.fmean$mean[i] = mean(na.omit(norm.female[norm.female$sample == norm.fmean$sample[i], 'normval']))

  norm.fstd$std[i] = (var(na.omit(norm.female[norm.female$sample == norm.fmean$sample[i], 'normval'])))

  i = i+1
}

norm.mmean = norm.mmean[order(norm.mmean$day),]
norm.mstd = norm.mstd[order(norm.mstd$day),]
norm.meb = norm.meb[order(norm.meb$day),]

norm.meb$y0 = norm.mmean$mean - norm.mstd$std
norm.meb$y1 = norm.mmean$mean + norm.mstd$std

norm.mlm = data.frame(geno = unique(norm.male$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(norm.mlm)){
  temp=na.omit(norm.male[norm.male$geno == norm.mlm$geno[i], ])
  temp.lm = lm(temp$normval~temp$day)
  norm.mlm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  norm.mlm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(norm.mlm) = norm.mlm$geno

norm.fmean = norm.fmean[order(norm.fmean$day),]
norm.fstd = norm.fstd[order(norm.fstd$day),]
norm.feb = norm.feb[order(norm.feb$day),]

norm.feb$y0 = norm.fmean$mean - norm.fstd$std
norm.feb$y1 = norm.fmean$mean + norm.fstd$std

norm.flm = data.frame(geno = unique(norm.female$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(norm.flm)){
  temp=na.omit(norm.female[norm.female$geno == norm.flm$geno[i], ])
  temp.lm = lm(temp$normval~temp$day)
  norm.flm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  norm.flm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(norm.flm) = norm.flm$geno

```

```{r echo = F, fig.height=10}

silent = norm.fmean[norm.fmean$geno == "silent", ]
a4v = norm.fmean[norm.fmean$geno == "A4V", ]
silentcha = norm.fmean[norm.fmean$geno == "silentcha", ]
a4vcha = norm.fmean[norm.fmean$geno == "A4Vcha", ]
silentok = norm.fmean[norm.fmean$geno == "silentok", ]
a4vok = norm.fmean[norm.fmean$geno == "A4Vok", ]

ls = norm.fmean[norm.fmean$geno == 'silent',]
la = norm.fmean[norm.fmean$geno == 'A4V',]
lsc = norm.fmean[norm.fmean$geno == 'silentcha',]
lac = norm.fmean[norm.fmean$geno == 'A4Vcha',]
lso = norm.fmean[norm.fmean$geno == 'silentok',]
lao = norm.fmean[norm.fmean$geno == 'A4Vok',]

pcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 9.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0,2.5),
     xlim = c(2,27),
     ylab = 'Relative Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     main = "Females")

axis(1, at=silent$day)

abline(v = 2, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 6, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 9, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 13, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 19, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 22, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 27, lty =2, lwd = 1, col = hex(pcol[7], .5))

lines(x = norm.flm['silent', c("x0", "x1")],
      y = norm.flm['silent', c("y0", "y1")],
      type = 'l',
      col = pcol[1],
      lwd = 2)

lines(x = norm.flm['A4V', c("x0", "x1")],
      y = norm.flm['A4V', c("y0", "y1")],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = norm.flm['silentcha', c("x0", "x1")],
      y = norm.flm['silentcha', c("y0", "y1")],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = norm.flm['A4Vcha', c("x0", "x1")],
      y = norm.flm['A4Vcha', c("y0", "y1")],
      type = 'l',
      col = pcol[4],
      lwd = 2)

lines(x = norm.flm['silentok', c("x0", "x1")],
      y = norm.flm['silentok', c("y0", "y1")],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = norm.flm['A4Vok', c("x0", "x1")],
      y = norm.flm['A4Vok', c("y0", "y1")],
      type = 'l',
      col = pcol[6],
      lwd = 2)

arrows(x0 = norm.feb$day[norm.feb$geno == 'silent'],
       x1 = norm.feb$day[norm.feb$geno == 'silent'],
       y0 = norm.feb$y0[norm.feb$geno == 'silent'],
       y1 = norm.feb$y1[norm.feb$geno == 'silent'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[1], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'A4V'],
       x1 = norm.feb$day[norm.feb$geno == 'A4V'],
       y0 = norm.feb$y0[norm.feb$geno == 'A4V'],
       y1 = norm.feb$y1[norm.feb$geno == 'A4V'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[2], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'silentcha'],
       x1 = norm.feb$day[norm.feb$geno == 'silentcha'],
       y0 = norm.feb$y0[norm.feb$geno == 'silentcha'],
       y1 = norm.feb$y1[norm.feb$geno == 'silentcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[3], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'A4Vcha'],
       x1 = norm.feb$day[norm.feb$geno == 'A4Vcha'],
       y0 = norm.feb$y0[norm.feb$geno == 'A4Vcha'],
       y1 = norm.feb$y1[norm.feb$geno == 'A4Vcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[4], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'silentok'],
       x1 = norm.feb$day[norm.feb$geno == 'silentok'],
       y0 = norm.feb$y0[norm.feb$geno == 'silentok'],
       y1 = norm.feb$y1[norm.feb$geno == 'silentok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[5], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'A4Vok'],
       x1 = norm.feb$day[norm.feb$geno == 'A4Vok'],
       y0 = norm.feb$y0[norm.feb$geno == 'A4Vok'],
       y1 = norm.feb$y1[norm.feb$geno == 'A4Vok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[6], 
       lwd=1)

points(x = silent$day,
       y = silent$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4v$day,
       y = a4v$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)

points(x = silent$day,
       y = silent$mean,
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = a4v$day,
       y = a4v$mean,
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex =2)

legend(x = 29,
       y = 2.5,
       legend = c('Silent', 'A4V',
                  'ChAT>SaxA;Silent', 'ChAT>SaxA;A4V', 
                  'OK371>SaxA;Silent', 'OK371>SaxA;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(pcol[1], pcol[2], pcol[3],
                pcol[4], pcol[5], pcol[6]),
       bty = 'n',
       y.intersp = 1,
       pt.cex = 2,
       xpd=TRUE)

```

```{r echo = F, fig.height=10}

silent = norm.mmean[norm.mmean$geno == "silent", ]
a4v = norm.mmean[norm.mmean$geno == "A4V", ]
silentcha = norm.mmean[norm.mmean$geno == "silentcha", ]
a4vcha = norm.mmean[norm.mmean$geno == "A4Vcha", ]
silentok = norm.mmean[norm.mmean$geno == "silentok", ]
a4vok = norm.mmean[norm.mmean$geno == "A4Vok", ]

ls = norm.mmean[norm.mmean$geno == 'silent',]
la = norm.mmean[norm.mmean$geno == 'A4V',]
lsc = norm.mmean[norm.mmean$geno == 'silentcha',]
lac = norm.mmean[norm.mmean$geno == 'A4Vcha',]
lso = norm.mmean[norm.mmean$geno == 'silentok',]
lao = norm.mmean[norm.mmean$geno == 'A4Vok',]

pcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 9.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0,2.5),
     xlim = c(2,27),
     ylab = 'Relative Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     main = "Males")

axis(1, at=silent$day)

abline(v = 2, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 6, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 9, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 13, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 19, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 22, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 27, lty =2, lwd = 1, col = hex(pcol[7], .5))

lines(x = norm.mlm['silent', c("x0", "x1")],
      y = norm.mlm['silent', c("y0", "y1")],
      type = 'l',
      col = pcol[1],
      lwd = 2)

lines(x = norm.mlm['A4V', c("x0", "x1")],
      y = norm.mlm['A4V', c("y0", "y1")],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = norm.mlm['silentcha', c("x0", "x1")],
      y = norm.mlm['silentcha', c("y0", "y1")],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = norm.mlm['A4Vcha', c("x0", "x1")],
      y = norm.mlm['A4Vcha', c("y0", "y1")],
      type = 'l',
      col = pcol[4],
      lwd = 2)

lines(x = norm.mlm['silentok', c("x0", "x1")],
      y = norm.mlm['silentok', c("y0", "y1")],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = norm.mlm['A4Vok', c("x0", "x1")],
      y = norm.mlm['A4Vok', c("y0", "y1")],
      type = 'l',
      col = pcol[6],
      lwd = 2)

arrows(x0 = norm.meb$day[norm.meb$geno == 'silent'],
       x1 = norm.meb$day[norm.meb$geno == 'silent'],
       y0 = norm.meb$y0[norm.meb$geno == 'silent'],
       y1 = norm.meb$y1[norm.meb$geno == 'silent'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[1], 
       lwd=1)

arrows(x0 = norm.meb$day[norm.meb$geno == 'A4V'],
       x1 = norm.meb$day[norm.meb$geno == 'A4V'],
       y0 = norm.meb$y0[norm.meb$geno == 'A4V'],
       y1 = norm.meb$y1[norm.meb$geno == 'A4V'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[2], 
       lwd=1)

arrows(x0 = norm.meb$day[norm.meb$geno == 'silentcha'],
       x1 = norm.meb$day[norm.meb$geno == 'silentcha'],
       y0 = norm.meb$y0[norm.meb$geno == 'silentcha'],
       y1 = norm.meb$y1[norm.meb$geno == 'silentcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[3], 
       lwd=1)

arrows(x0 = norm.meb$day[norm.meb$geno == 'A4Vcha'],
       x1 = norm.meb$day[norm.meb$geno == 'A4Vcha'],
       y0 = norm.meb$y0[norm.meb$geno == 'A4Vcha'],
       y1 = norm.meb$y1[norm.meb$geno == 'A4Vcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[4], 
       lwd=1)

arrows(x0 = norm.meb$day[norm.meb$geno == 'silentok'],
       x1 = norm.meb$day[norm.meb$geno == 'silentok'],
       y0 = norm.meb$y0[norm.meb$geno == 'silentok'],
       y1 = norm.meb$y1[norm.meb$geno == 'silentok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[5], 
       lwd=1)

arrows(x0 = norm.meb$day[norm.meb$geno == 'A4Vok'],
       x1 = norm.meb$day[norm.meb$geno == 'A4Vok'],
       y0 = norm.meb$y0[norm.meb$geno == 'A4Vok'],
       y1 = norm.meb$y1[norm.meb$geno == 'A4Vok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[6], 
       lwd=1)

points(x = silent$day,
       y = silent$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4v$day,
       y = a4v$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)

points(x = silent$day,
       y = silent$mean,
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = a4v$day,
       y = a4v$mean,
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex =2)

legend(x = 29,
       y = 2.5,
       legend = c('Silent', 'A4V',
                  'ChAT>SaxA;Silent', 'ChAT>SaxA;A4V', 
                  'OK371>SaxA;Silent', 'OK371>SaxA;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(pcol[1], pcol[2], pcol[3],
                pcol[4], pcol[5], pcol[6]),
       bty = 'n',
       y.intersp = 1,
       pt.cex = 2,
       xpd=TRUE)

```