### 2024-09-27 20240514_AxonalMorphologyResults.Rmd

---
title: "20240514_AxonalMorphologyResults"
author: "John Santiago"
date: "2024-05-14"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F}

morph = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/20240513_pharate_axonal_morphology_Raw_Data.csv")

```

```{r echo = F, warning= F}

conditions = unique(morph$Genotype)

sig.data = as.list(6:17)
names(sig.data) = colnames(morph)[6:17]

pvals = setNames(c(0,0,0), c('G85Rcr vs WT', 'G85Rcr vs G85Rhet', 'G85Rhet vs WT'))

j = 6
while(j <= 17){
  morph.data = colnames(morph)[j]
  sig.data[[morph.data]] = pvals
k = 1
while(k<4){

  comp.mat = matrix(c(1,1,2,3,2,3), ncol = 2)

  x = morph[morph$Genotype == conditions[comp.mat[k,1]], morph.data]
  y = morph[morph$Genotype == conditions[comp.mat[k,2]], morph.data]

  if(j != 8 & j != 17){
    cont.table = data.frame(t = c(sum(x), sum(y)),
             f = c((length(x)- sum(x)), (length(y)- sum(y))))
    row.names(cont.table) = conditions[c(comp.mat[k,1],comp.mat[k,2])]
    sig.data[[morph.data]][[k]]= signif(as.numeric(chisq.test(cont.table)[3]),2)
  }

  if(j == 8 | j == 17){
    sig.data[[morph.data]][[k]]= signif(as.numeric(t.test(x, y)[3]),2)
  }
  k = k + 1
}
j=j+1
}

```

```{r echo = F}

j = 6

while(j<=ncol(morph)){
mean.cat = setNames(morph[,j],
                    morph$Genotype)

mean.data = setNames(rep(0, length(unique(names(mean.cat)))), unique(names(mean.cat)))

i=1
while(i<=length(mean.data)){
  mean.data[i] = mean(mean.cat[names(mean.cat) == names(mean.data)[i]])
  i=i+1
}

if(j != 8 & j != 17){
barplot(height = mean.data,
        names.arg = names(mean.data),
        las = 2, 
        main = colnames(morph)[j], 
        cex.names = .8,
        col = c('brown', 'gold', 'steelblue'), 
        density = c(50),
        ylim = c(0,1))
}

if(j == 8 | j == 17){
barplot(height = mean.data,
        names.arg = names(mean.data),
        las = 2, 
        main = colnames(morph)[j], 
        cex.names = .8,
        col = c('brown', 'gold', 'steelblue'), 
        density = c(50))
}

print(sig.data[colnames(morph)[j]])

j=j+1
}

```