### 2024-09-27 Arc1MeetingFigures.Rmd

---
title: "arc1 gene expression"
author: "John Santiago"
date: "2024-08-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = F, echo = F}

library(plotly)
library(edgeR)
library(heatmaply)
library(org.Dm.eg.db)

##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"
cpmdata = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
groups  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
TKT.EdgeR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)
TKT.EdgeR.FC = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FCTable.csv"),row.names = 1)

G85R.meancpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT_meancpmdata.csv", row.names = 1)

groups = groups[colnames(cpmdata),]

GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)

A4V.cpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.cpmdata.csv", row.names = 1)

A4V.meancpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.meancpmdata.csv", row.names = 1)

A4V.FDR = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FDRdata.csv", row.names = 1)

A4V.FC = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FCdata.csv", row.names = 1)

A4V.meta = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.metadata.csv", row.names = 1)

```

```{r}

  FC.data = TKT.EdgeR.FC[, c("GRxWT.FC", "GRxWT.MC")]
  FDR.data = TKT.EdgeR[, c("GRxWT.FC", "GRxWT.MC")]
  mean.data = G85R.meancpm[, c('GRFC', 'WTFC', 'GRMC', 'WTMC')]

  go.genes = row.names(FDR.data[FDR.data$GRxWT.FC < .05 &
                             FDR.data$GRxWT.MC < .05 ,])

data = data.frame(Symbol = GeneIDKey[go.genes, "Symbol"],
                  FBgn = go.genes,
                  FDR1 = -log2(FDR.data[go.genes, 1]),
                  FDR2 = -log2(FDR.data[go.genes, 2]),
                  FC1 = FC.data[go.genes, 1],
                  FC2 = FC.data[go.genes, 2],
                  Color = 'grey',
                  size = 1,
                  Variable1.cpm = mean.data[go.genes,1],
                  Control1.cpm = mean.data[go.genes,2],
                  Variable2.cpm = mean.data[go.genes,3],
                  Control2.cpm = mean.data[go.genes,4])

  main.title = "Female vs Male DEGs"

  ##data$Color[(data$FC1 - 1) > data$FC2] = 'royalblue'
  ##data$Color[(data$FC2 - 1) > data$FC1] = 'firebrick' 
  data$Color[data$Symbol == "Arc1"] = 'gold'
  size = data.frame(G85R = apply(mean.data[go.genes,c(1,3)], 1, mean),
                    WT = apply(mean.data[go.genes,c(2,4)], 1, mean))
  size = size + .001
  data$size = log2(apply(size, 1, max))
  data$size[data$size<3] = 1

  data = na.omit(data)
  ##data = data[data$size > log2(20),]
  data$size = 5*data$size

  fig = plot_ly(data = data,
                x = ~FC1,
                y = ~FC2,
                type = 'scatter',
                mode = 'markers',
                marker = list(color = ~Color,
                              colors = ~Color,
                              line = list(color = "black", width = 1.5),
                              size = ~size),
                hoverinfo = "text",
                hovertext = paste("Symbol:", data$Symbol,
                                  "\nFBgn:", data$FBgn,
                                  "\n ", colnames(FDR.data)[1], "FDR: ", round(2^-data$FDR1, 2),
                                  "\n ", colnames(FDR.data)[2], "FDR: ", round(2^-data$FDR2,2),
                                  "\n ", colnames(FC.data)[1], "FC: ", round(data$FC1, 2),
                                  "\n ", colnames(FC.data)[2], "FC: ", round(data$FC2,2),
                                  "\n ", colnames(mean.data)[1], "mean CPM: ", signif(data$Variable1,2),
                                  "\n ", colnames(mean.data)[2], "mean CPM: ", signif(data$Control1,2),
                                  "\n ", colnames(mean.data)[3], "mean CPM: ", signif(data$Variable2,2),
                                  "\n ", colnames(mean.data)[4], "mean CPM: ", signif(data$Control2,2)))

  fig = fig %>% layout(title = main.title,
                       shapes = list(
                         type = "line",
                         line = list(color = "black", dash = 'dash', width = 2),
                         x0 = min(data$FC1, data$FC2) - 1,
                         x1 = max(data$FC1, data$FC2) + 1,
                         y0 = min(data$FC1, data$FC2) - 1,
                         y1 = max(data$FC1, data$FC2) + 1),
                       xaxis = list(title = paste(colnames(FC.data)[1], "FC")),
                       yaxis = list(title = paste(colnames(FC.data)[2], "FC")),
                       showlegend = F)

    fig = fig %>% add_markers(data = data[data$Symbol == "Arc1",],
                x = ~FC1,
                y = ~FC2,
                type = 'scatter',
                mode = 'markers',
                hoverinfo = "text",
                hovertext = paste("Symbol:", data$Symbol[data$Symbol == "Arc1"],
                                  "\nFBgn:", data$FBgn[data$Symbol == "Arc1"],
                                  "\n ", colnames(FDR.data)[1], "FDR: ", round(2^-data$FDR1[data$Symbol == "Arc1"], 2),
                                  "\n ", colnames(FDR.data)[2], "FDR: ", round(2^-data$FDR2[data$Symbol == "Arc1"],2),
                                  "\n ", colnames(FC.data)[1], "FC: ", round(data$FC1[data$Symbol == "Arc1"], 2),
                                  "\n ", colnames(FC.data)[2], "FC: ", round(data$FC2[data$Symbol == "Arc1"],2),
                                  "\n ", colnames(mean.data)[1], "mean CPM: ", signif(data$Variable1[data$Symbol == "Arc1"],2),
                                  "\n ", colnames(mean.data)[2], "mean CPM: ", signif(data$Control1[data$Symbol == "Arc1"],2),
                                  "\n ", colnames(mean.data)[3], "mean CPM: ", signif(data$Variable2[data$Symbol == "Arc1"],2),
                                  "\n ", colnames(mean.data)[4], "mean CPM: ", signif(data$Control2[data$Symbol == "Arc1"],2)))
  fig
```

```{r echo = F}

gene = data.frame(cpm = as.numeric(cpmdata[FBgn,]), 
                  condition = groups$Group, 
                  group = paste0(groups$Genotype, groups$Sex),
                  TKT = groups$TKT)

gene = gene[gene$TKT != 'OE',]

gene$group = factor(gene$group, levels = c("WTF", "GRF", "WTM", "GRM"))
gene$condition = factor(gene$condition, levels = c('WT.F','GR.F',
                                                   'WT.M','GR.M'))

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar = c(5,7,4,2))
plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(150, 2900),
     xlim = c(.5, 4.5),
     xaxt = "n",
     yaxt = "n",
     ylab = NA,
     xlab = NA)
rect(0, 0, 4.75, 3050, col = hex('whitesmoke', .5))
boxplot(gene$cpm~gene$condition,    
        main = "",
        ylab="",
        xlab = "",
        las = 2,
        cex.axis = 1.5,
        boxlwd = 3,
        lwd = 3,
        xaxt="n",
        ylim = c(10, 3000),
        xlim = c(0,4.5),
        col = c("grey90", hex("hotpink", .25), 
                "grey90", hex("deepskyblue", .2)),
        add = T)

box(lwd = 3)
axis(1, 
     at = c(1:4), 
     cex.axis = 1.5,
     labels = c("Control", "G85R", "Control", "G85R"), 
     las=1)
points(x=as.numeric(factor(gene$condition)),
       y=gene$cpm,
       cex=2,
       pch=21,
       lwd = 2,
       bg=c('grey65',"firebrick","grey65","deepskyblue")[as.numeric(factor(gene$group))])
abline(v=2.5, lty = 2, lwd = 3)
text(x = c(1.5, 3.5), y = c(3100,3100), c("Female", "Male"),cex= 2, xpd = T)
title(ylab="Counts per Million", line=5, cex.lab=2)
title(xlab = "Genotype", line = 3.5, cex.lab=2)

```

```{r echo = F}

gene = data.frame(cpm = as.numeric(A4V.cpm[FBgn,]), 
                  condition = A4V.groups$Group, 
                  group = paste0(groups$Genotype, groups$Sex),
                  TKT = groups$TKT)

gene$group = factor(gene$group, levels = c("WTF", "GRF", "WTM", "GRM"))
gene$condition = factor(gene$condition, levels = c('WT.F', "TktDfWT.F", "TktOEWT.F",
                                                   'GR.F', "TktDfGR.F", "TktOEGR.F",

                                                   'WT.M', "TktDfWT.M", "TktOEWT.M",
                                                   'GR.M', "TktDfGR.M", "TktOEGR.M"))

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar = c(6,7,4,2))
boxplot(gene$cpm~gene$condition,    
        main = "",
        ylab="",
        xlab = "",
        las = 2,
        cex.axis = 1.5,
        boxlwd = 3,
        lwd = 3,
        xaxt="n",
        ylim = c(100, 3500),
        xlim = c(.945, 12.055),
        col = c(rep(hex("lightgrey", .5),3), rep(hex("mistyrose", .5),3), 
                rep(hex("lightgrey", .5),3), rep(hex("lightcyan", .5),3)))
box(lwd = 3)
axis(1, 
     at = c(1:12), 
     cex.axis = 1,
     labels = rep(c("Control", "TKT Df", "TKT OE"),4), 
     las=1)
text(x = c(2, 5, 8, 11), 
     y = c(-300,-300, -300, -300), 
     c("Control", "G85R", "Control", "G85R"),
     cex= 1.25, xpd = T)
points(x=as.numeric(factor(gene$condition)),
       y=gene$cpm,
       cex=2,
       pch=21,
       lwd = 2,
       bg=c('grey25',"firebrick","grey25","deepskyblue")[as.numeric(factor(gene$group))])
abline(v=6.5, lty = 2, lwd = 3)
lines(x=c(3.5,3.5), y = c(-40,-350), lty = 1, lwd = 3, xpd = T)
lines(x=c(.5,.5), y = c(-40,-350), lty = 1, lwd = 3, xpd = T)
lines(x=c(6.5,6.5), y = c(-40,-350), lty = 1, lwd = 3, xpd = T)
lines(x=c(9.5,9.5), y = c(-40,-350), lty = 1, lwd = 3, xpd = T)
lines(x=c(12.5,12.5), y = c(-40,-350), lty = 1, lwd = 3, xpd = T)
text(x = c(3.5, 9.5), y = c(3750,3750), c("Female", "Male"),cex= 2, xpd = T)
title(ylab="Counts per Million", line=5, cex.lab=2)
title(xlab = "Genotype", line = 5, cex.lab=2)

```

```{r echo = F, fig.width=10}

fs = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 1F left_females.csv", header = T)

ms = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 1G left_males.csv", header = T)

lcol = c('brown1', 'firebrick','steelblue', 'navy', 'limegreen', 'darkgreen',  'grey')
pcol = c('brown1', 'firebrick','steelblue', 'navy', 'limegreen', 'darkgreen',  'grey')
hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 6))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, 100),
     xlim = c(0,82),
     ylab = 'Percent Alive',
     xlab = 'Day',
     xaxt = 'n',
     main = "A4V Survival")

axis(1, at= 1:82, lwd.ticks = .5, labels = NA)
axis(1, at= c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90), 
     lwd.ticks = 1, cex.axis = 1,
     col.axis = 'lightgrey')
axis(1, at= c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90), 
     lwd.ticks = 1, cex.axis = 1,
     col.axis = 'black')

lines(x = fs$day,
      y = fs$dsodWT,
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = fs$day,
      y = fs$A4V,
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = ms$day,
      y = ms$dsodWT,
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = ms$day,
      y = ms$A4V,
      type = 'l',
      col = pcol[4],
      lwd = 2)

points(x = fs$day,
       y = fs$dsodWT,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = fs$day,
       y = fs$A4V,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = ms$day,
       y = ms$dsodWT,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = ms$day,
       y = ms$A4V,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)

points(x = fs$day,
       y = fs$dsodWT,
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 21,
       cex =2)

points(x = fs$day,
       y = fs$A4V,
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 23,
       cex =2)
points(x = ms$day,
       y = ms$dsodWT,
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 21,
       cex =2)
points(x = ms$day,
       y = ms$A4V,
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 23,
       cex =2)

legend('topright',
       legend = c('Female Silent', 'Female A4V',
                  'Male Silent', 'Male A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[3], .5),
                hex(pcol[4], .5), hex(pcol[5], .5), hex(pcol[6], .5)),
       bty = 'n',
       y.intersp = 1,
       pt.cex = 2,
       xpd=TRUE)

#abline(v = 3, lty = 2, lwd = 2, col = 'black')
#abline(v = 9, lty = 2, lwd = 2, col = 'black')
#abline(v = 40, lty = 2, lwd = 2, col = 'black')

#points(x = c(3,9,40),
#       y = c(-11.75,-11.75,-11.75),
#       pch = 21,
#       cex = 4,
#       bg = hex('gold', .25),
#       col = 'grey25',
#       xpd = T)

#axis(1, at= c(3, 9, 40), lwd.ticks = 2, labels = c(3,9,40))

```

```{r echo = F}

gene = data.frame(cpm = as.numeric(A4V.cpm[FBgn, row.names(A4V.meta)]),
                  group = A4V.meta$Group,
                  condition = paste(substring(A4V.meta$Genotype,1,1),
                                    substring(A4V.meta$Sex,1,1)))

gene = gene[A4V.meta$Section == 'Head',]

gene$group = factor(gene$group, levels = c("S3FH", "S9FH", "S40FH",
                                           "A3FH", "A9FH", "A40FH",
                                           "S3MH", "S9MH",
                                           "A3MH", "A9MH"))

gene$condition = factor(gene$condition, levels = c("S F", "A F", "S M", "A M"))

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar = c(6,7,4,2))
boxplot(gene$cpm~gene$group,    
        main = "",
        ylab="",
        xlab = "",
        las = 2,
        cex.axis = 1.5,
        boxlwd = 3,
        lwd = 3,
        xaxt="n",
        ylim = c(0, 750),
        xlim = c(.75, 10.25),
        col = c(rep(hex("lightgrey", .5),3), rep(hex("mistyrose", .5),3), 
                rep(hex("lightgrey", .5),3), rep(hex("lightcyan", .5),3)))
box(lwd = 3)
axis(1, 
     at = c(1:10), 
     cex.axis = 1,
     labels = c(rep(c("3 Day", "9 Day", "40 Day"),2),
                rep(c("3 Day", "9 Day"),2)),
     las=1)
text(x = c(2, 5, 7.5, 9.5), 
     y = c(-125,-125, -125, -125), 
     c("Control", "A4V", "Control", "A4V"),
     cex= 1.25, xpd = T)
points(x=as.numeric(factor(gene$group)),
       y=gene$cpm,
       cex=2,
       pch=21,
       lwd = 2,
       bg=c('grey25',"firebrick","grey25","deepskyblue")[as.numeric(factor(gene$condition))])
abline(v=6.5, lty = 2, lwd = 3)
lines(x=c(3.5,3.5), y = c(-33,-125), lty = 1, lwd = 3, xpd = T)
lines(x=c(.37,.37), y = c(-33,-125), lty = 1, lwd = 3, xpd = T)
lines(x=c(6.5,6.5), y = c(-33,-125), lty = 1, lwd = 3, xpd = T)
lines(x=c(8.5,8.5), y = c(-33,-125), lty = 1, lwd = 3, xpd = T)
lines(x=c(10.63,10.63), y = c(-33,-125), lty = 1, lwd = 3, xpd = T)
text(x = c(3.5, 8.5), y = c(825,825), c("Female", "Male"),cex= 2, xpd = T)
title(ylab="Counts per Million", line=5, cex.lab=2)
title(xlab = "Genotype", line = 5, cex.lab=2)

```

```{r echo = F}

gene = data.frame(cpm = as.numeric(A4V.cpm[FBgn, row.names(A4V.meta)]),
                  group = A4V.meta$Group,
                  condition = paste(substring(A4V.meta$Genotype,1,1),
                                    substring(A4V.meta$Sex,1,1)))

gene = gene[A4V.meta$Section == 'Abdomen',]

gene$group = factor(gene$group, levels = c("S3FA", "S9FA", "S40FA",
                                           "A3FA", "A9FA", "A40FA",
                                           "S3MA", "S9MA",
                                           "A3MA", "A9MA"))

gene$condition = factor(gene$condition, levels = c("S F", "A F", "S M", "A M"))

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar = c(6,7,4,2))
boxplot(gene$cpm~gene$group,    
        main = "",
        ylab="",
        xlab = "",
        las = 2,
        cex.axis = 1.5,
        boxlwd = 3,
        lwd = 3,
        xaxt="n",
        ylim = c(0, 1450),
        xlim = c(.75, 10.25),
        col = c(rep(hex("lightgrey", .5),3), rep(hex("mistyrose", .5),3), 
                rep(hex("lightgrey", .5),3), rep(hex("lightcyan", .5),3)))
box(lwd = 3)
axis(1, 
     at = c(1:10), 
     cex.axis = 1,
     labels = c(rep(c("3 Day", "9 Day", "40 Day"),2),
                rep(c("3 Day", "9 Day"),2)),
     las=1)
text(x = c(2, 5, 7.5, 9.5), 
     y = c(-250,-250, -250, -250), 
     c("Control", "A4V", "Control", "A4V"),
     cex= 1.25, xpd = T)
points(x=as.numeric(factor(gene$group)),
       y=gene$cpm,
       cex=2,
       pch=21,
       lwd = 2,
       bg=c('grey25',"firebrick","grey25","deepskyblue")[as.numeric(factor(gene$condition))])
abline(v=6.5, lty = 2, lwd = 3)
lines(x=c(3.5,3.5), y = c(-60,-250), lty = 1, lwd = 3, xpd = T)
lines(x=c(.37,.37), y = c(-60,-250), lty = 1, lwd = 3, xpd = T)
lines(x=c(6.5,6.5), y = c(-60,-250), lty = 1, lwd = 3, xpd = T)
lines(x=c(8.5,8.5), y = c(-60,-250), lty = 1, lwd = 3, xpd = T)
lines(x=c(10.63,10.63), y = c(-60,-250), lty = 1, lwd = 3, xpd = T)
text(x = c(3.5, 8.5), y = c(1600,1600), c("Female", "Male"),cex= 2, xpd = T)
title(ylab="Counts per Million", line=5, cex.lab=2)
title(xlab = "Genotype", line = 5, cex.lab=2)

```

```{r echo = F}

gene = data.frame(cpm = as.numeric(A4V.cpm[FBgn, row.names(A4V.meta)]),
                  group = A4V.meta$Group,
                  condition = paste(substring(A4V.meta$Genotype,1,1),
                                    substring(A4V.meta$Sex,1,1)))

gene = gene[A4V.meta$Section == 'Thorax',]

gene$group = factor(gene$group, levels = c("S3FT", "S9FT", "S40FT",
                                           "A3FT", "A9FT", "A40FT",
                                           "S3MT", "S9MT",
                                           "A3MT", "A9MT"))

gene$condition = factor(gene$condition, levels = c("S F", "A F", "S M", "A M"))

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar = c(6,7,4,2))
boxplot(gene$cpm~gene$group,    
        main = "",
        ylab="",
        xlab = "",
        las = 2,
        cex.axis = 1.5,
        boxlwd = 3,
        lwd = 3,
        xaxt="n",
        ylim = c(0, 8500),
        xlim = c(.75, 10.25),
        col = c(rep(hex("lightgrey", .5),3), rep(hex("mistyrose", .5),3), 
                rep(hex("lightgrey", .5),3), rep(hex("lightcyan", .5),3)))
box(lwd = 3)
axis(1, 
     at = c(1:10), 
     cex.axis = 1,
     labels = c(rep(c("3 Day", "9 Day", "40 Day"),2),
                rep(c("3 Day", "9 Day"),2)),
     las=1)
text(x = c(2, 5, 7.5, 9.5), 
     y = c(-1400,-1400, -1400, -1400), 
     c("Control", "A4V", "Control", "A4V"),
     cex= 1.25, xpd = T)
points(x=as.numeric(factor(gene$group)),
       y=gene$cpm,
       cex=2,
       pch=21,
       lwd = 2,
       bg=c('grey25',"firebrick","grey25","deepskyblue")[as.numeric(factor(gene$condition))])
abline(v=6.5, lty = 2, lwd = 3)
lines(x=c(3.5,3.5), y = c(-370,-1500), lty = 1, lwd = 3, xpd = T)
lines(x=c(.37,.37), y = c(-370,-1500), lty = 1, lwd = 3, xpd = T)
lines(x=c(6.5,6.5), y = c(-370,-1500), lty = 1, lwd = 3, xpd = T)
lines(x=c(8.5,8.5), y = c(-370,-1500), lty = 1, lwd = 3, xpd = T)
lines(x=c(10.63,10.63), y = c(-370,-1500), lty = 1, lwd = 3, xpd = T)
text(x = c(3.5, 8.5), y = c(9500,9500), c("Female", "Male"),cex= 2, xpd = T)
title(ylab="Counts per Million", line=5, cex.lab=2)
title(xlab = "Genotype", line = 5, cex.lab=2)

```

```{r echo = F}

gene = data.frame(cpm = as.numeric(A4V.cpm[FBgn, row.names(A4V.meta)]),
                  group = A4V.meta$Group,
                  condition = paste(substring(A4V.meta$Genotype,1,1),
                                    substring(A4V.meta$Sex,1,1)))

gene = gene[A4V.meta$Section == 'Head',]

gene$group = factor(gene$group, levels = c("S3FH", "S9FH", "S40FH",
                                           "A3FH", "A9FH", "A40FH",
                                           "S3MH", "S9MH",
                                           "A3MH", "A9MH"))

gene$condition = factor(gene$condition, levels = c("S F", "A F", "S M", "A M"))

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar = c(6,7,4,2))
boxplot(gene$cpm~gene$group,    
        main = "",
        ylab="",
        xlab = "",
        las = 2,
        cex.axis = 1.5,
        boxlwd = 3,
        lwd = 3,
        xaxt="n",
        ylim = c(0, 8500),
        xlim = c(.75, 10.25),
        col = c(rep(hex("lightgrey", .5),3), rep(hex("mistyrose", .5),3), 
                rep(hex("lightgrey", .5),3), rep(hex("lightcyan", .5),3)))
box(lwd = 3)
axis(1, 
     at = c(1:10), 
     cex.axis = 1,
     labels = c(rep(c("3 Day", "9 Day", "40 Day"),2),
                rep(c("3 Day", "9 Day"),2)),
     las=1)
text(x = c(2, 5, 7.5, 9.5), 
     y = c(-1000,-1000, -1000, -1000), 
     c("Control", "A4V", "Control", "A4V"),
     cex= 1.25, xpd = T)
points(x=as.numeric(factor(gene$group)),
       y=gene$cpm,
       cex=2,
       pch=21,
       lwd = 2,
       bg=c('grey25',"firebrick","grey25","deepskyblue")[as.numeric(factor(gene$condition))])
abline(v=6.5, lty = 2, lwd = 3)
lines(x=c(3.5,3.5), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(.37,.37), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(6.5,6.5), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(8.5,8.5), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(10.63,10.63), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
text(x = c(3.5, 8.5), y = c(9200,9200), c("Female", "Male"),cex= 2, xpd = T)
title(ylab="Counts per Million", line=5, cex.lab=2)
title(xlab = "Genotype", line = 5, cex.lab=2)

```

```{r echo = F}

gene = data.frame(cpm = as.numeric(A4V.cpm[FBgn, row.names(A4V.meta)]),
                  group = A4V.meta$Group,
                  condition = paste(substring(A4V.meta$Genotype,1,1),
                                    substring(A4V.meta$Sex,1,1)))

gene = gene[A4V.meta$Section == 'Abdomen',]

gene$group = factor(gene$group, levels = c("S3FA", "S9FA", "S40FA",
                                           "A3FA", "A9FA", "A40FA",
                                           "S3MA", "S9MA",
                                           "A3MA", "A9MA"))

gene$condition = factor(gene$condition, levels = c("S F", "A F", "S M", "A M"))

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar = c(6,7,4,2))
boxplot(gene$cpm~gene$group,    
        main = "",
        ylab="",
        xlab = "",
        las = 2,
        cex.axis = 1.5,
        boxlwd = 3,
        lwd = 3,
        xaxt="n",
        ylim = c(0, 8500),
        xlim = c(.75, 10.25),
        col = c(rep(hex("lightgrey", .5),3), rep(hex("mistyrose", .5),3), 
                rep(hex("lightgrey", .5),3), rep(hex("lightcyan", .5),3)))
box(lwd = 3)
axis(1, 
     at = c(1:10), 
     cex.axis = 1,
     labels = c(rep(c("3 Day", "9 Day", "40 Day"),2),
                rep(c("3 Day", "9 Day"),2)),
     las=1)
text(x = c(2, 5, 7.5, 9.5), 
     y = c(-1000,-1000, -1000, -1000), 
     c("Control", "A4V", "Control", "A4V"),
     cex= 1.25, xpd = T)
points(x=as.numeric(factor(gene$group)),
       y=gene$cpm,
       cex=2,
       pch=21,
       lwd = 2,
       bg=c('grey25',"firebrick","grey25","deepskyblue")[as.numeric(factor(gene$condition))])
abline(v=6.5, lty = 2, lwd = 3)
lines(x=c(3.5,3.5), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(.37,.37), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(6.5,6.5), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(8.5,8.5), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(10.63,10.63), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
text(x = c(3.5, 8.5), y = c(9200,9200), c("Female", "Male"),cex= 2, xpd = T)
title(ylab="Counts per Million", line=5, cex.lab=2)
title(xlab = "Genotype", line = 5, cex.lab=2)

```

```{r echo = F}

gene = data.frame(cpm = as.numeric(A4V.cpm[FBgn, row.names(A4V.meta)]),
                  group = A4V.meta$Group,
                  condition = paste(substring(A4V.meta$Genotype,1,1),
                                    substring(A4V.meta$Sex,1,1)))

gene = gene[A4V.meta$Section == 'Thorax',]

gene$group = factor(gene$group, levels = c("S3FT", "S9FT", "S40FT",
                                           "A3FT", "A9FT", "A40FT",
                                           "S3MT", "S9MT",
                                           "A3MT", "A9MT"))

gene$condition = factor(gene$condition, levels = c("S F", "A F", "S M", "A M"))

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar = c(6,7,4,2))
boxplot(gene$cpm~gene$group,    
        main = "",
        ylab="",
        xlab = "",
        las = 2,
        cex.axis = 1.5,
        boxlwd = 3,
        lwd = 3,
        xaxt="n",
        ylim = c(0, 8500),
        xlim = c(.75, 10.25),
        col = c(rep(hex("lightgrey", .5),3), rep(hex("mistyrose", .5),3), 
                rep(hex("lightgrey", .5),3), rep(hex("lightcyan", .5),3)))
box(lwd = 3)
axis(1, 
     at = c(1:10), 
     cex.axis = 1,
     labels = c(rep(c("3 Day", "9 Day", "40 Day"),2),
                rep(c("3 Day", "9 Day"),2)),
     las=1)
text(x = c(2, 5, 7.5, 9.5), 
     y = c(-1000,-1000, -1000, -1000), 
     c("Control", "A4V", "Control", "A4V"),
     cex= 1.25, xpd = T)
points(x=as.numeric(factor(gene$group)),
       y=gene$cpm,
       cex=2,
       pch=21,
       lwd = 2,
       bg=c('grey25',"firebrick","grey25","deepskyblue")[as.numeric(factor(gene$condition))])
abline(v=6.5, lty = 2, lwd = 3)
lines(x=c(3.5,3.5), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(.37,.37), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(6.5,6.5), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(8.5,8.5), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(10.63,10.63), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
text(x = c(3.5, 8.5), y = c(9200,9200), c("Female", "Male"),cex= 2, xpd = T)
title(ylab="Counts per Million", line=5, cex.lab=2)
title(xlab = "Genotype", line = 5, cex.lab=2)

```

```{r}

  FC.data = A4V.FC[, c("AvS9FT", "AvS9MT")]
  FDR.data = A4V.FDR[, c("AvS9FT", "AvS9MT")]
  mean.data = A4V.meancpm[, c('A9FT', 'S9FT', 'A9MT', 'S9MT')]

  go.genes = row.names(FDR.data[FDR.data$AvS9FT < .05 |
                             FDR.data$AvS9MT < .05 ,])

data = data.frame(Symbol = GeneIDKey[go.genes, "Symbol"],
                  FBgn = go.genes,
                  FDR1 = -log2(FDR.data[go.genes, 1]),
                  FDR2 = -log2(FDR.data[go.genes, 2]),
                  FC1 = FC.data[go.genes, 1],
                  FC2 = FC.data[go.genes, 2],
                  Color = 'grey',
                  size = 1,
                  Variable1.cpm = mean.data[go.genes,1],
                  Control1.cpm = mean.data[go.genes,2],
                  Variable2.cpm = mean.data[go.genes,3],
                  Control2.cpm = mean.data[go.genes,4])

  main.title = "Female vs Male DEGs"

  ##data$Color[(data$FC1 - 1) > data$FC2] = 'royalblue'
  ##data$Color[(data$FC2 - 1) > data$FC1] = 'firebrick' 
  data$Color[data$Symbol == "Arc1"] = 'gold'
  size = data.frame(G85R = apply(mean.data[go.genes,c(1,3)], 1, mean),
                    WT = apply(mean.data[go.genes,c(2,4)], 1, mean))
  size = size + .001
  data$size = log2(apply(size, 1, max))
  data$size[data$size<3] = 1

  data = na.omit(data)
  data = data[data$size > log2(10),]
  data$size = 5*data$size

  fig = plot_ly(data = data,
                x = ~FC1,
                y = ~FC2,
                type = 'scatter',
                mode = 'markers',
                marker = list(color = ~Color,
                              colors = ~Color,
                              line = list(color = "black", width = 1.5),
                              size = ~size),
                hoverinfo = "text",
                hovertext = paste("Symbol:", data$Symbol,
                                  "\nFBgn:", data$FBgn,
                                  "\n ", colnames(FDR.data)[1], "FDR: ", round(2^-data$FDR1, 2),
                                  "\n ", colnames(FDR.data)[2], "FDR: ", round(2^-data$FDR2,2),
                                  "\n ", colnames(FC.data)[1], "FC: ", round(data$FC1, 2),
                                  "\n ", colnames(FC.data)[2], "FC: ", round(data$FC2,2),
                                  "\n ", colnames(mean.data)[1], "mean CPM: ", signif(data$Variable1,2),
                                  "\n ", colnames(mean.data)[2], "mean CPM: ", signif(data$Control1,2),
                                  "\n ", colnames(mean.data)[3], "mean CPM: ", signif(data$Variable2,2),
                                  "\n ", colnames(mean.data)[4], "mean CPM: ", signif(data$Control2,2)))

  fig = fig %>% layout(title = main.title,
                       shapes = list(
                         type = "line",
                         line = list(color = "black", dash = 'dash', width = 2),
                         x0 = min(data$FC1, data$FC2) - 1,
                         x1 = max(data$FC1, data$FC2) + 1,
                         y0 = min(data$FC1, data$FC2) - 1,
                         y1 = max(data$FC1, data$FC2) + 1),
                       xaxis = list(title = paste(colnames(FC.data)[1], "FC")),
                       yaxis = list(title = paste(colnames(FC.data)[2], "FC")),
                       showlegend = F)

    fig = fig %>% add_markers(data = data[data$Symbol == "Arc1",],
                x = ~FC1,
                y = ~FC2,
                type = 'scatter',
                mode = 'markers',
                hoverinfo = "text",
                hovertext = paste("Symbol:", data$Symbol[data$Symbol == "Arc1"],
                                  "\nFBgn:", data$FBgn[data$Symbol == "Arc1"],
                                  "\n ", colnames(FDR.data)[1], "FDR: ", round(2^-data$FDR1[data$Symbol == "Arc1"], 2),
                                  "\n ", colnames(FDR.data)[2], "FDR: ", round(2^-data$FDR2[data$Symbol == "Arc1"],2),
                                  "\n ", colnames(FC.data)[1], "FC: ", round(data$FC1[data$Symbol == "Arc1"], 2),
                                  "\n ", colnames(FC.data)[2], "FC: ", round(data$FC2[data$Symbol == "Arc1"],2),
                                  "\n ", colnames(mean.data)[1], "mean CPM: ", signif(data$Variable1[data$Symbol == "Arc1"],2),
                                  "\n ", colnames(mean.data)[2], "mean CPM: ", signif(data$Control1[data$Symbol == "Arc1"],2),
                                  "\n ", colnames(mean.data)[3], "mean CPM: ", signif(data$Variable2[data$Symbol == "Arc1"],2),
                                  "\n ", colnames(mean.data)[4], "mean CPM: ", signif(data$Control2[data$Symbol == "Arc1"],2)))
  fig
```

```{r echo = F}

sun = "FBgn0014391"
mbl = "FBgn0265487"
  ##could not find mblA, just mbl
NimC1 = "FBgn0259896"

gene = data.frame(cpm = as.numeric(A4V.cpm[sun, row.names(A4V.meta)]),
                  group = A4V.meta$Group,
                  condition = paste(substring(A4V.meta$Genotype,1,1),
                                    substring(A4V.meta$Sex,1,1)))

gene = gene[A4V.meta$Section == 'Thorax',]

gene$group = factor(gene$group, levels = c("S3FT", "S9FT", "S40FT",
                                           "A3FT", "A9FT", "A40FT",
                                           "S3MT", "S9MT",
                                           "A3MT", "A9MT"))

gene$condition = factor(gene$condition, levels = c("S F", "A F", "S M", "A M"))

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar = c(6,7,4,2))
boxplot(gene$cpm~gene$group,    
        main = "",
        ylab="",
        xlab = "",
        las = 2,
        cex.axis = 1.5,
        boxlwd = 3,
        lwd = 3,
        xaxt="n",
        ylim = c(0, 8500),
        xlim = c(.75, 10.25),
        col = c(rep(hex("lightgrey", .5),3), rep(hex("mistyrose", .5),3), 
                rep(hex("lightgrey", .5),3), rep(hex("lightcyan", .5),3)))
box(lwd = 3)
axis(1, 
     at = c(1:10), 
     cex.axis = 1,
     labels = c(rep(c("3 Day", "9 Day", "40 Day"),2),
                rep(c("3 Day", "9 Day"),2)),
     las=1)
text(x = c(2, 5, 7.5, 9.5), 
     y = c(-1000,-1000, -1000, -1000), 
     c("Control", "A4V", "Control", "A4V"),
     cex= 1.25, xpd = T)
points(x=as.numeric(factor(gene$group)),
       y=gene$cpm,
       cex=2,
       pch=21,
       lwd = 2,
       bg=c('grey25',"firebrick","grey25","deepskyblue")[as.numeric(factor(gene$condition))])
abline(v=6.5, lty = 2, lwd = 3)
lines(x=c(3.5,3.5), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(.37,.37), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(6.5,6.5), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(8.5,8.5), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
lines(x=c(10.63,10.63), y = c(-370,-1100), lty = 1, lwd = 3, xpd = T)
text(x = c(3.5, 8.5), y = c(9200,9200), c("Female", "Male"),cex= 2, xpd = T)
title(ylab="Counts per Million", line=5, cex.lab=2)
title(xlab = "Genotype", line = 5, cex.lab=2)

```