### 2024-09-27 GeneExpressionBoxPlots.Rmd

---
title: "rac1 gene expression"
author: "John Santiago"
date: "2024-05-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = F, echo = F}

library(plotly)
library(edgeR)
library(heatmaply)
library(org.Dm.eg.db)
library(KEGGREST)

##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"
cpmdata = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
groups  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
TKT.EdgeR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)
TKT.EdgeR.FC = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FCTable.csv"),row.names = 1)

groups = groups[colnames(cpmdata),]

gbbOE = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbOE_CountTable.csv", row.names = 1)
gbbOE.groups = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbOE_Metadata.csv", row.names = 1)
gbbOE = gbbOE[,row.names(gbbOE.groups)]
gbbOE.groups$group = paste0(gbbOE.groups$genotype, "_",gbbOE.groups$stage)

##normalize data
countdata=gbbOE
x <- countdata
gr <- factor(gbbOE.groups$group, levels=c(unique(gbbOE.groups$group)))
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
gbbOE.cpmdata=cpm(z)

gbbOEmean = gbbOE.cpmdata[,1:6]
colnames(gbbOEmean) = unique(gr)
i=1
while(i<=6){
  temp= gbbOE.cpmdata[,row.names(gbbOE.groups[gbbOE.groups$group == colnames(gbbOEmean)[i],])]
  gbbOEmean[,i] = apply(temp, 1, mean)
  i=i+1
}

design<-model.matrix(~0 + gr)
colnames(design) = unique(gr)
z = estimateDisp(z, design, robust = T)

fit <- glmQLFit(z, design)
plotQLDisp(fit)

compare = makeContrasts(G85R_LL3-LoxP_LL3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.LL3=DEout
sGRxWT.LL3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_ML3-LoxP_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.ML3=DEout
sGRxWT.ML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_LL3-G85R_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.LL3xML3=DEout
sGR.LL3xML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(LoxP_LL3-LoxP_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.LL3xML3=DEout
sWT.LL3xML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_gbb_LL3-G85R_gbb_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
G85R_gbb.LL3xML3=DEout
sG85R_gbb.LL3xML3=DEout[DEout$FDR<.05,]

gbbKO = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbKO_CountTable.csv", row.names = 1)
gbbKO.groups = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbKO_Metadata.csv", row.names = 1)

##normalize data
countdata=gbbKO
x <- countdata
gr <- factor(gbbKO.groups$Genotype, levels=c(unique(gbbKO.groups$Genotype)))
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
gbbKO.cpmdata=cpm(z)

GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)

A4V.cpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.cpmdata.csv",row.names = 1)

A4V.FDR = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FDRdata.csv")

nodeID = as.list(org.Dm.egENZYME)

```

```{r echo = F}

##subsetting genes for patterns

```

```{r echo = F}

##relish FBgn0014018
##attacin FBgn0041579
##cecropin FBgn0000276
##diptericin FBgn0004240
##drosocin FBgn0010388
##defensin FBgn0010385
##drosomycin FBgn0283461
##metchnikowin FBgn0014865
##sod1 FBgn0003462
##sod2 FBgn0010213
##catalase FBgn0000261
##gstd1 FBgn0001149
## arc1 FBgn0033926
##rac1
##cgs = c('CG10592','CG10827','CG1462','CG16771','CG1809','CG3264','CG3290','CG3292','CG5150','CG5361','CG5656','CG8105','CG8147')
##mle 'FBgn0002774'
##gstD FBgn0001149
##Tudor-SN FBgn0035121
##atf4 FBgn0000370
##thor FBgn0261560
##sdhbl FBgn0030975
##ND-51L FBgn0034251
##kayak (AP-1) FBgn0001297
## Jra (AP-1) FBgn0001291
#sad FBgn0003312

##Specific enzyme
enzyme.ens = nodeID[grep('3.1.3.5',nodeID)]
enzyme.genes = GeneIDKey[GeneIDKey$ensembl %in% names(enzyme.ens)[c(1,3:5,7)],]

```

> GeneIDKey[intersect(s1,intersect(s2,intersect(s3,s4))),'Symbol']
 [1] "Cyp4p1"  "CG13962" "Cpr51A"  "CG11438" "Sgs5bis" "sad"    
 [7] "Lip2"    "CG4757"  "CG14229" "CG15406" "CG3270"  "CG9877" 
[13] "CG13905" "Zip71B"  "CG17027" "nom"     "CG16727" "CG9702" 
[19] "CG43740"
> GeneIDKey[intersect(s1,intersect(s2,intersect(s3,s4))),'FBgn']
 [1] "FBgn0015037" "FBgn0032824" "FBgn0033942" "FBgn0037164"
 [5] "FBgn0038523" "FBgn0003312" "FBgn0024740" "FBgn0027584"
 [9] "FBgn0031059" "FBgn0031517" "FBgn0033093" "FBgn0034819"
[13] "FBgn0035176" "FBgn0036461" "FBgn0036553" "FBgn0037617"
[17] "FBgn0038719" "FBgn0039787" "FBgn0263997"

```{r echo = F}
FBgn = 'FBgn0011817'
name = 'FBgn0011817: nmo'

gene = data.frame(cpm = as.numeric(cpmdata[FBgn,]), condition = groups$Group, group = paste0(groups$Genotype, groups$Sex))

gene$group = factor(gene$group, levels = c(unique(gene$group)))
gene$condition = factor(gene$condition, levels = c('WT.F', "TktDfWT.F", "TktOEWT.F",
                                                   'GR.F', "TktDfGR.F", "TktOEGR.F",
                                                   'WT.M', "TktDfWT.M", "TktOEWT.M",
                                                   'GR.M', "TktDfGR.M", "TktOEGR.M"))

par(mar = c(5,7,5,2))
plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(min(gene$cpm), max(gene$cpm)),
     xlim = c(.5, 12.5),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = NA)

rect(0, 0-(1.5*max(gene$cpm)), 13, 1.5*max(gene$cpm), col = 'gray95')

boxplot(gene$cpm~gene$condition, 
        xlab="",
        las = 2, 
        cex.axis = .8,
        xaxt = 'none',
        yaxt = 'none',
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        add = T,
        col = c(rep("honeydew", 3),
                rep("thistle1", 3),
                rep("lightcyan", 3),
                rep("lightyellow", 3)))

box(lwd = 2)

points(x=as.numeric(factor(gene$condition)),
       y=gene$cpm,
       cex=1,
       pch=21,
       bg=c('firebrick',"gold","darkgreen","dodgerblue")[as.numeric(factor(gene$group))])

axis(side = 1,
     at = 1:12,
     labels = rep(c("Control", "Df", "OE"), 4),
     line = 0,
     cex.axis = .75,
     lwd.ticks = 2)

axis(side = 2,
     line = 0,
     cex.axis = 1,
     lwd.ticks = 2,
     las = 2)

axis(side = 2,
     at = .5*(max(gene$cpm) + min(gene$cpm)),
     label = "Counts per million reads",
     tick = F,
     padj = -3,
     cex.axis = 1.25)

axis(side = 1,
     at = c(2, 5, 8, 11),
     labels = c("Silent", "G85R","Silent", "G85R"),
     tick = F,
     padj = 2,
     cex.axis = 1)
axis(side = 3,
     at = c(3.5, 9.5),
     labels = c("Female", "Male"),
     tick = F,
     padj = .5,
     cex.axis = 1.5)

axis(side = 3,
     at = 6.5,
     labels = name,
     tick = F,
     padj = -1.5,
     cex.axis = 2)

axis(side = 3,
     at = c(3.5,9.5),
     tick = T,
     outer = T,
     labels = F,
     line = -2.25,
     lwd.ticks = 0)

xline1 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.04
xline2 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.19
lines(x=c(3.5,3.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(6.5,6.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(9.5,9.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
abline(v = 6.5, lty = 2, lwd = 2)

```

```{r echo = F}

gene = data.frame(cpm = as.numeric(cpmdata[FBgn,]), 
                  condition = groups$Group, 
                  group = paste0(groups$Genotype, groups$Sex),
                  TKT = groups$TKT)

gene = gene[gene$TKT != 'OE',]

gene$group = factor(gene$group, levels = c(unique(gene$group)))
gene$condition = factor(gene$condition, levels = c('GR.F', "TktDfGR.F",
                                                 'WT.F', "TktDfWT.F",
                                                 'GR.M', "TktDfGR.M",
                                                 'WT.M', "TktDfWT.M"))

boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"gold","darkgreen","dodgerblue")[as.numeric(factor(gene$group))])

```

```{r echo = F}

A4V.groups = as.vector(colnames(A4V.cpm))

A4V.groups = substr(A4V.groups,1, nchar(A4V.groups)-1)

par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
gene = data.frame(cpm = as.numeric(A4V.cpm[FBgn,]), 
                  condition = A4V.groups, 
                  group = paste0(substr(A4V.groups, 1,1), substr(A4V.groups, nchar(as.vector(A4V.groups))-1,nchar(as.vector(A4V.groups))-1)))

gene$condition = factor(gene$condition, levels = c("A3FH", "A9FH", "A40FH",
                                                   "S3FH", "S9FH", "S40FH",
                                                   "A3FT", "A9FT", "A40FT",
                                                   "S3FT", "S9FT", "S40FT",
                                                   "A3FA", "A9FA", "A40FA",
                                                   "S3FA", "S9FA", "S40FA",
                                                   "A3MH", "A9MH",
                                                   "S3MH", "S9MH",
                                                   "A3MT", "A9MT",
                                                   "S3MT", "S9MT",
                                                   "A3MA", "A9MA",
                                                   "S3MA", "S9MA"))
gene$group = factor(gene$group, levels = c(unique(gene$group)))

boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"gold","darkgreen","dodgerblue")[as.numeric(factor(gene$group))])
legend('topright',
       inset=c(-0.275,0),
       legend = c('A4V F', 'Silent F', 'A4V M', 'Silent F'), 
       fill = c('firebrick',"darkgreen", "gold","dodgerblue"),
       cex = .65,
       bty = 'n',
       pt.cex = .5)

```

```{r}

gene = data.frame(cpm = as.numeric(gbbOE.cpmdata[FBgn,]), condition = gbbOE.groups$group, group = gbbOE.groups$genotype)

gene$group = factor(gene$group, levels = c(unique(gene$group)))
gene$condition = factor(gene$condition, levels = c('LoxP_ML3', "LoxP_LL3",
                                                 'G85R_ML3', "G85R_LL3",
                                                 'G85R_gbb_ML3', "G85R_gbb_LL3"))

boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"dodgerblue","gold")[as.numeric(factor(gene$group))])

```

```{r}

## WT = W1118
## KO = GBB1/GBB2 (both null alleles)
## Aaron and Catherine legend for RNAseq
## Whole larvae

gene = data.frame(cpm = as.numeric(gbbKO.cpmdata[FBgn,]), condition = gbbKO.groups$Genotype)

gene$condition = factor(gene$condition, levels = c('WT', "KO"))

boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"dodgerblue","gold")[as.numeric(factor(gene$condition))])

```