### 2024-09-27 KristiPresentationData.Rmd

---
title: "Kristi Meeting"
author: "John Santiago"
date: "2024-06-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F}

library(plotly)
library(edgeR)
library(heatmaply)
library(pathview)
library(org.Dm.eg.db)

A4V.cpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.cpmdata.csv", row.names = 1)
A4V.meancpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.meancpmdata.csv", row.names = 1)
A4V.FDR = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FDRdata.csv", row.names = 1)
A4V.FC = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FCdata.csv", row.names = 1)
A4V.meta = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.metadata.csv", row.names = 1)

A4V.comparekey = setNames(colnames(A4V.FDR),
                          c('A3FHvS3FH', 'A3MHvS3MH', 'A9FHvS9FH', 'A9MHvS9MH', 'A40FHvS40FH',
                            'A3FHvA9FH', 'S3FHvS9FH', 'A3MHvA9MH', 'S3FHvS9MH',
                            'A9FHvA40FH', 'S9FHvS40FH', 'A3FHvA40FH', 'S3FHvS40FH',

                            'A3FTvS3FT', 'A3MTvS3MT', 'A9FTvS9FT', 'A9MTvS9MT', 'A40FTvS40FT',
                            'A3FTvA9FT', 'S3FTvS9FT', 'A3MTvA9MT', 'S3FTvS9MT',
                            'A9FTvA40FT', 'S9FTvS40FT', 'A3FTvA40FT', 'S3FTvS40FT',

                            'A3FAvS3FA', 'A3MAvS3MA', 'A9FAvS9FA', 'A9MAvS9MA', 'A40FAvS40FA',
                            'A3FAvA9FA', 'S3FAvS9FA', 'A3MAvA9MA', 'S3FAvS9MA',
                            'A9FAvA40FA', 'S9FAvS40FA', 'A3FAvA40FA', 'S3FAvS40FA',

                            'A3FHvA3MH', 'S3FHvS3MH', 'A9FHvA9MH', 'S9FHvS9MH',
                            'A3FTvA3MT', 'S3FTvS3MT', 'A9FTvA9MT', 'S9FTvS9MT',
                            'A3FAvA3MA', 'S3FAvS3MA', 'A9FAvA9MA', 'S9FAvS9MA'))

G85R.cpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT_cpmdata.csv",row.names = 1)
G85R.cpm = G85R.cpm[grep('FBgn', row.names(G85R.cpm)),]
G85R.meancpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT_meancpmdata.csv", row.names = 1)
colnames(G85R.meancpm) = c('GRFC', 'GRMC', 
                           'GRFDf', 'GRMDf', 'WTFDf', 'WTMDf',
                           'GRFOE', 'GRMOE', 'WTFOE', 'WTMOE',
                           'WTFC', 'WTMC')
G85R.FDR = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv",row.names = 1)
G85R.FC = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT.EdgeR.FCTable.csv", row.names = 1)
G85R.meta  = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT.metadata.csv",row.names = 1)

G85R.comparekey = setNames(colnames(G85R.FDR),
                           c('GRFCvGRFDf', 'GRFCvGRFOE', 'WTFCvWTFDf', 'WTFCvWTFOE', 
                             'GRFCvWTFC', 'GRFDfvWTFDf', 'GRFOEvWTFOE',
                             'GRMCvGRMDf', 'GRMCvGRMOE', 'WTMCvWTMDf', 'WTMCvWTMOE', 
                             'GRMCvWTMC', 'GRMDfvWTMFDf', 'GRMOEvWTMOE',
                             'GRFCvGRMC', 'WTFCvWTMC'))

G85R.metab.comparekey = setNames(c("GR.CxDf", "GR.CxOE", "WT.CxDf", "WT.CxOE",
                                   "GRxWT.C", "GRxWT.Df", "GRxWT.OE",
                                   "GR.CxDf", "GR.CxOE", "WT.CxDf", "WT.CxOE",
                                   "GRxWT.C", "GRxWT.Df", "GRxWT.OE"), names(G85R.comparekey)[1:14])

G85R.metabnorm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/NormalizedMetabolomicData.csv", row.names = 1)
G85R.metabmean = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/MeanNormalizedMetabolomicData.csv", row.names = 1)
metab.id = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/MetaboliteMultiKEGGIDKey.csv", row.names = 1)
metab.id = setNames(metab.id$metab, metab.id$KEGG)
G85R.metabFC = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/MetaboliteFCs.csv", row.names = 1)
G85R.metabFDR = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/MetaboliteFDRs.csv", row.names = 1)
G85R.metabp = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/Metabolitepvals.csv", row.names = 1)

GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)
KEGG.MetabKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/RawMetabolomicData.csv", row.names = 1)
KEGG.MetabKey = setNames(KEGG.MetabKey[-c(1:2), 2], row.names(KEGG.MetabKey)[-c(1:2)])
genesingo = readRDS("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/genesingo.RData")
GenesInKegg = readRDS("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/kegg.symbol2path.RData")
KEGG.Names = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/KEGG.names.csv", row.names = 1)
GO.Names = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GO.names.csv", row.names = 1)

git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"
kegg = readRDS(gzcon(url(paste0(git.dir,"GeneralDataFiles/kegg.symbol2path.RData"))))

nodeID = as.list(org.Dm.egENZYME)

```

```{r echo = F, fig.height= 12}

NADmetab = names(kegg[grep('00190', kegg)])
NADmetab = row.names(GeneIDKey[GeneIDKey$Symbol %in% NADmetab, ])

geneset = NADmetab

hmdata = A4V.meancpm[geneset,]

A4V.order = c("S3FH", "S9FH", "S40FH", "A3FH", "A9FH", "A40FH",
              "S3FT", "S9FT", "S40FT", "A3FT", "A9FT", "A40FT",
              "S3FA", "S9FA", "S40FA", "A3FA", "A9FA", "A40FA",
              "S3MH", "S9MH", "A3MH", "A9MH",
              "S3MT", "S9MT", "A3MT", "A9MT",
              "S3MA", "S9MA", "A3MA", "A9MA")

hmdata = (hmdata[!is.na((GeneIDKey[row.names(hmdata), 'Symbol'])),A4V.order])
row.names(hmdata) = GeneIDKey[row.names(hmdata), 'Symbol']

heatmaply(hmdata,
          scale = 'row',
          Colv = F,
          cexRow = .5)

```

```{r echo = F, fig.height= 12}

hmdata = G85R.meancpm[geneset,]
G85R.order = c("WTFC", "WTFDf", "WTFOE",
          "GRFC", "GRFDf", "GRFOE",
          "WTMC", "WTMDf", "WTMOE",
          "GRMC", "GRMDf", "GRMOE")

hmdata = (hmdata[!is.na((GeneIDKey[row.names(hmdata), 'Symbol'])),G85R.order])
row.names(hmdata) = GeneIDKey[row.names(hmdata), 'Symbol']

heatmaply(hmdata,
          scale = 'row',
          Colv = F,
          cexRow = .5)

```

```{r warning=F,  echo = F}

FC.data = G85R.FC[,c("GRF.CxOE", "WTF.CxOE")]
metab.data = setNames(G85R.metabFC[,"GR.CxOE"], G85R.metabFC$KEGG)
PathwayID = "00190"
sigs = row.names(FC.data)

##File name
out.id = "G85RandWT_CxOE_Genes with GR.CxOE_Metab"
OutputDir = "/Users/johncsantiago/Documents/GitHub/WhartonLab/G85R OXPHOS KEGG Pathways/"
setwd(OutputDir)
img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".multi.png")

  if(class(FC.data)=="numeric"){
    deg.data=setNames(FC.data[sigs],GeneIDKey[sigs, "ensembl"])
  }

  if(class(FC.data)=="data.frame"){
    deg.data = as.matrix(FC.data)
    row.names(deg.data) = GeneIDKey[row.names(FC.data), "ensembl"]
  }

  pv.out <- pathview(gene.data =  deg.data,
                     cpd.data = metab.data,
                     pathway.id = PathwayID,
                     species = "dme",
                     kegg.native = T,
                     limit=list(gene=c(-2,2)),
                     node.sum="max.abs",
                     low="royalblue",mid = "grey",high="firebrick",
                     out.suffix = out.id,
                     bins=list(genes=30),
                     plot.col.key = T,
                     match.data = F)

  knitr::include_graphics(img.file)

```

```{r echo = F}

heatmaply(G85R.metabmean[c('nicotinamide', 'nicotinic acid', "NAD+", "NADH"),
                         c('WT', 'TktDfWT', 'TktOEWT', 'GR', 'TktDfGR', 'TktOEGR')],
          scale = 'row',
          Colv = F,
                 custom_hovertext = signif(G85R.metabmean[c('nicotinamide', 'nicotinic acid', "NAD+", "NADH"),
                         c('WT', 'TktDfWT', 'TktOEWT', 'GR', 'TktDfGR', 'TktOEGR')], 3))

```

```{r echo = F}

PPP = names(kegg[grep('00030', kegg)])
PPP = row.names(GeneIDKey[GeneIDKey$Symbol %in% PPP, ])

geneset = PPP

hmdata = A4V.meancpm[geneset,]
row.names(hmdata) = GeneIDKey[geneset, 'Symbol']
A4V.order = c("S3FH", "S9FH", "S40FH", "A3FH", "A9FH", "A40FH",
              "S3FT", "S9FT", "S40FT", "A3FT", "A9FT", "A40FT",
              "S3FA", "S9FA", "S40FA", "A3FA", "A9FA", "A40FA",
              "S3MH", "S9MH", "A3MH", "A9MH",
              "S3MT", "S9MT", "A3MT", "A9MT",
              "S3MA", "S9MA", "A3MA", "A9MA")

hmdata = na.omit(hmdata[,A4V.order])

heatmaply(hmdata,
          scale = 'row',
          Colv = F)

```

```{r echo = F}

hmdata = G85R.meancpm[geneset,]
row.names(hmdata) = GeneIDKey[geneset, 'Symbol']
order = c("WTFC", "WTFDf", "WTFOE",
          "GRFC", "GRFDf", "GRFOE",
          "WTMC", "WTMDf", "WTMOE",
          "GRMC", "GRMDf", "GRMOE")

hmdata = na.omit(hmdata[,order])

heatmaply(hmdata,
          scale = 'row',
          Colv = F)

```

```{r warning=F,  echo = F}

FC = G85R.FC[,c("GRxWT.FC", "GRxWT.MC")]
metab.data = setNames(G85R.metabFC[,"GRxWT.C"], G85R.metabFC$KEGG)
PathwayID = "00030"
use = 'both'

  ##File name prefix
  out.id = paste("G85RvsWT_FemaleandMale")

  ##table with the FC data you want to use for coloring with FBgn for row names 
  ##if only analyzing 1 condition, use setNames to make a named vector
  ##A4V.FC or G85R.FC

  FC.data =  FC

  ##all or a subset of row.names used in FC.data
  usegenes = row.names(FC)

  #Any column name from G85R.metabFC or 'none'
  metab.data = metab.data

  if(use == 'primary'){

    ##File name prefix
    out.id = comparison

    ##table with the FC data you want to use for coloring with FBgn for row names 
    ##if only analyzing 1 condition, use setNames to make a named vector
    ##A4V.FC or G85R.FC

    FC.data =  FC[,1]

    #Any column name from G85R.metabFC or 'none'
    metab.data = metab.data
  }

  ##KEGG Mapping Function
  if(length(intersect(list.files(), "KEGG_Image_Files")) == 0){
    dir.create(paste0(dirname(rstudioapi::getSourceEditorContext()$path), "/KEGG_Image_Files/"))
  }
  Home = dirname(rstudioapi::getSourceEditorContext()$path)
  OutputDir = paste0(Home, "/KEGG_Image_Files/")
  setwd(OutputDir)
  sigs = usegenes
  ##sigs = names(FDR.data[FDR.data <= .05])

  if(class(FC.data)=="numeric"){
    deg.data=setNames(FC.data[sigs],GeneIDKey[sigs, "ensembl"])
  }

  if(class(FC.data)=="data.frame"){
    deg.data = as.matrix(FC.data)
    row.names(deg.data) = GeneIDKey[row.names(FC.data), "ensembl"]
  }

  #if(length(G85R.metabFC[metab.id, names(metab.data)]) > 1 ){
   # metab.data = setNames(log2(G85R.metabFC[metab.id, metab.data]), names(metab.id))
  #}

  pv.out <- pathview(gene.data =  deg.data,
                     cpd.data = metab.data,
                     pathway.id = PathwayID,
                     species = "dme",
                     kegg.native = T,
                     limit=list(gene=c(-2,2)),
                     node.sum="max.abs",
                     low="royalblue",mid = "grey",high="firebrick",
                     out.suffix = out.id,
                     bins=list(genes=30),
                     plot.col.key = T,
                     match.data = F)
  setwd(Home)

  if(class(FC.data)=="numeric"){
    img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".png")
  }

  if(class(FC.data)=="data.frame"){
    img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".multi.png")
  }
  knitr::include_graphics(img.file)

```

```{r warning=F,  echo = F}

FC = G85R.FC[,c("GRF.CxDf", "WTF.CxDf")]
metab.data = setNames(G85R.metabFC[,"GR.CxDf"], G85R.metabFC$KEGG)
PathwayID = "00030"
use = 'both'

  ##File name prefix
  out.id = paste("G85RvsWT_FemaleandMale")

  ##table with the FC data you want to use for coloring with FBgn for row names 
  ##if only analyzing 1 condition, use setNames to make a named vector
  ##A4V.FC or G85R.FC

  FC.data =  FC

  ##all or a subset of row.names used in FC.data
  usegenes = row.names(FC)

  #Any column name from G85R.metabFC or 'none'
  metab.data = metab.data

  if(use == 'primary'){

    ##File name prefix
    out.id = comparison

    ##table with the FC data you want to use for coloring with FBgn for row names 
    ##if only analyzing 1 condition, use setNames to make a named vector
    ##A4V.FC or G85R.FC

    FC.data =  FC[,1]

    #Any column name from G85R.metabFC or 'none'
    metab.data = metab.data
  }

  ##KEGG Mapping Function
  if(length(intersect(list.files(), "KEGG_Image_Files")) == 0){
    dir.create(paste0(dirname(rstudioapi::getSourceEditorContext()$path), "/KEGG_Image_Files/"))
  }
  Home = dirname(rstudioapi::getSourceEditorContext()$path)
  OutputDir = paste0(Home, "/KEGG_Image_Files/")
  setwd(OutputDir)
  sigs = usegenes
  ##sigs = names(FDR.data[FDR.data <= .05])

  if(class(FC.data)=="numeric"){
    deg.data=setNames(FC.data[sigs],GeneIDKey[sigs, "ensembl"])
  }

  if(class(FC.data)=="data.frame"){
    deg.data = as.matrix(FC.data)
    row.names(deg.data) = GeneIDKey[row.names(FC.data), "ensembl"]
  }

  #if(length(G85R.metabFC[metab.id, names(metab.data)]) > 1 ){
   # metab.data = setNames(log2(G85R.metabFC[metab.id, metab.data]), names(metab.id))
  #}

  pv.out <- pathview(gene.data =  deg.data,
                     cpd.data = metab.data,
                     pathway.id = PathwayID,
                     species = "dme",
                     kegg.native = T,
                     limit=list(gene=c(-2,2)),
                     node.sum="max.abs",
                     low="royalblue",mid = "grey",high="firebrick",
                     out.suffix = out.id,
                     bins=list(genes=30),
                     plot.col.key = T,
                     match.data = F)
  setwd(Home)

  if(class(FC.data)=="numeric"){
    img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".png")
  }

  if(class(FC.data)=="data.frame"){
    img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".multi.png")
  }
  knitr::include_graphics(img.file)

```

```{r warning=F,  echo = F}

FC = G85R.FC[,c("GRF.CxOE", "WTF.CxOE")]
metab.data = setNames(G85R.metabFC[,"GR.CxOE"], G85R.metabFC$KEGG)

PathwayID = "00030"
use = 'both'

  ##File name prefix
  out.id = paste("G85RvsWT_FemaleandMale")

  ##table with the FC data you want to use for coloring with FBgn for row names 
  ##if only analyzing 1 condition, use setNames to make a named vector
  ##A4V.FC or G85R.FC

  FC.data =  FC

  ##all or a subset of row.names used in FC.data
  usegenes = row.names(FC)

  #Any column name from G85R.metabFC or 'none'
  metab.data = metab.data

  if(use == 'primary'){

    ##File name prefix
    out.id = comparison

    ##table with the FC data you want to use for coloring with FBgn for row names 
    ##if only analyzing 1 condition, use setNames to make a named vector
    ##A4V.FC or G85R.FC

    FC.data =  FC[,1]

    #Any column name from G85R.metabFC or 'none'
    metab.data = metab.data
  }

  ##KEGG Mapping Function
  if(length(intersect(list.files(), "KEGG_Image_Files")) == 0){
    dir.create(paste0(dirname(rstudioapi::getSourceEditorContext()$path), "/KEGG_Image_Files/"))
  }
  Home = dirname(rstudioapi::getSourceEditorContext()$path)
  OutputDir = paste0(Home, "/KEGG_Image_Files/")
  setwd(OutputDir)
  sigs = usegenes
  ##sigs = names(FDR.data[FDR.data <= .05])

  if(class(FC.data)=="numeric"){
    deg.data=setNames(FC.data[sigs],GeneIDKey[sigs, "ensembl"])
  }

  if(class(FC.data)=="data.frame"){
    deg.data = as.matrix(FC.data)
    row.names(deg.data) = GeneIDKey[row.names(FC.data), "ensembl"]
  }

  #if(length(G85R.metabFC[metab.id, names(metab.data)]) > 1 ){
   # metab.data = setNames(log2(G85R.metabFC[metab.id, metab.data]), names(metab.id))
  #}

  pv.out <- pathview(gene.data =  deg.data,
                     cpd.data = metab.data,
                     pathway.id = PathwayID,
                     species = "dme",
                     kegg.native = T,
                     limit=list(gene=c(-2,2)),
                     node.sum="max.abs",
                     low="royalblue",mid = "grey",high="firebrick",
                     out.suffix = out.id,
                     bins=list(genes=30),
                     plot.col.key = T,
                     match.data = F)
  setwd(Home)

  if(class(FC.data)=="numeric"){
    img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".png")
  }

  if(class(FC.data)=="data.frame"){
    img.file = paste0(OutputDir, "dme",PathwayID, ".", out.id, ".multi.png")
  }
  knitr::include_graphics(img.file)

```