### 2024-09-27 PunnettSquareApp.Rmd

---
title: "Punnett Square"
author: "John Santiago"
date: "2024-09-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F}

##Genotypes
male.genotype = 'w/w;arc1/arc1;+/+'
female.genotype = 'elav-Gal4/Y;+/+;TM6c/UAS-lacz'

##allele or combination of alleles separated by commas
##eg c('x', 'CyO/CyO, 'x/y')
##cells meeting multiple criteria will be colored in order
allele1 = c('Y')
allele2 = c('arc1', 'UAS-lacz', 'elav-Gal4')
allele3 = 'CyO/CyO'
##color for each different allele set listed
color1 = 'lightblue'
color2 = 'gold'
color3 = 'black'

```

```{r echo = F}

library(DT)

male = strsplit(male.genotype, split = ";")[[1]]
male = strsplit(male, split = "/")

female = strsplit(female.genotype, split = ";")[[1]]
female = strsplit(female, split = "/")

nchroms = length(male)

chr1 = male[[1]]
chrdata = chr1
if(nchroms > 1){
  chr1 = c(chr1, chr1)
  chr2 = c(rep(male[[2]][1], 2), rep(male[[2]][2], 2))
  chr12 = paste(paste0(chr1, ";"),chr2)
  chrdata = data.frame(chr1 = chr1,
                       chr2 = chr2,
                       chr12 = chr12)
  if(nchroms > 2){
    chr1 = c(chr1, chr1)
    chr2 = c(chr2, chr2)
    chr12 = paste0(c(chr12, chr12), ";")
    chr3 = c(rep(male[[3]][1], 4), rep(male[[3]][2], 4))
    chr123 = paste(chr12, chr3)
    chrdata = data.frame(chr1 = chr1,
                         chr2 = chr2,
                         chr3 = chr3,
                         chr123 = chr123)
  }
}

male.chr = unique(chrdata)

chr1 = female[[1]]
chrdata = chr1
if(nchroms > 1){
  chr1 = c(chr1, chr1)
  chr2 = c(rep(female[[2]][1], 2), rep(female[[2]][2], 2))
  chr12 = paste(paste0(chr1, ";"),chr2)
  chrdata = data.frame(chr1 = chr1,
                       chr2 = chr2,
                       chr12 = chr12)
  if(nchroms > 2){
    chr1 = c(chr1, chr1)
    chr2 = c(chr2, chr2)
    chr12 = paste0(c(chr12, chr12), ";")
    chr3 = c(rep(female[[3]][1], 4), rep(female[[3]][2], 4))
    chr123 = paste(chr12, chr3)
    chrdata = data.frame(chr1 = chr1,
                         chr2 = chr2,
                         chr3 = chr3,
                         chr123 = chr123)
  }
}

female.chr = unique(chrdata)

ps = matrix("", nrow = nrow(female.chr), ncol = nrow(male.chr))
row.names(ps) = female.chr[, nchroms+1]
colnames(ps) = male.chr[, nchroms+1]
ps.print = ps
i=1
while(i <= ncol(ps)){
  temp.ps = paste0(female.chr[,1], "/", male.chr[i,1])
  if(nchroms > 1){
    temp.ps2 = paste0(female.chr[,2], "/", male.chr[i,2])
    temp.ps = paste0(temp.ps, ";", temp.ps2)
    temp.print = paste0(temp.ps, ";\n", temp.ps2)
  }
  if(nchroms > 2){
    temp.ps3 = paste0(female.chr[,3], "/", male.chr[i,3])
    temp.ps = paste0(temp.ps, ";", temp.ps3)
    temp.print = paste0(temp.ps3, ";\n", temp.ps3)
  }
  ps[,i] = temp.ps
  ps.print[,i] = temp.print
  i=i+1
}

allele.colors = ps
allele.colors[1:nrow(ps),] = ''

i = 1
while(i<=3){
  if(length(allele1)>1)){
    j=1
    temp = ps[grep(allele1[j], ps)]
    j=2
    while(j<=length(allele1)){
      temp = intersect(temp, ps[grep(allele1[j], ps)])
      j=j+1
    }
    if(length(temp) == 1){
      allele.colors[ps == temp] = color1
    }
    if(length(temp) > 1){
      j=1
      while(j <= length(temp)){
        allele.colors[ps == temp[j]] = color1
        j = j+1
      }
    }
  i=i+1
  }

write.csv(ps, "/Users/johncsantiago/Desktop/cross.csv")

```

```{r echo = F}

df = datatable(ps) %>% 
  formatStyle(1:ncol(ps), backgroundColor = styleEqual(ps, c(allele.colors)))

df
```