### 2024-09-27 RussoPaper/RussoClimbingPlots.Rmd

---
title: "Russo Manuscript Climbing Plot"
author: "John Santiago"
date: "2024-03-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = F, echo = F}

library(plotly)
library(edgeR)

```

```{r include = F, echo = F, include = F}

fig2ab = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 2AB.csv", header = T)
fig2ab$Sample = paste(fig2ab$Genotype, paste(fig2ab$Sex, fig2ab$Day))

mean.fig2 = rep(0, length(unique(fig2ab$Sample)))
names(mean.fig2) = unique(fig2ab$Sample)
var.fig2 = mean.fig2

i=1
while(i<=length(mean.fig2)){
  mean.fig2[i] = mean(as.numeric(na.omit(fig2ab[fig2ab$Sample == names(mean.fig2)[i], 'Rate'])))

    var.fig2[i] = var(as.numeric(na.omit(fig2ab[fig2ab$Sample == names(mean.fig2)[i], 'Rate'])))

  i = i+1
}

eb = data.frame(sample = names(mean.fig2),
                y0 = mean.fig2 - var.fig2,
                y1 = mean.fig2 + var.fig2)

```

```{r echo = F, include = F}

f.fig2ab = fig2ab[fig2ab$Sex == 'F' &
                  (fig2ab$Day == 3 |
                  fig2ab$Day == 28 |
                  fig2ab$Day == 35) &
                  fig2ab$Genotype != 'w1118',]

f.fig2ab = f.fig2ab[order(fig2ab$Day),]
f.fig2ab$Sample = paste(f.fig2ab$Genotype, f.fig2ab$Day)
f.fig2ab$Sample = factor(f.fig2ab$Sample, levels = c('dSodWT 3', 'A4V 3',
                                                     'dSodWT 28', 'A4V 28',
                                                     'dSodWT 35', 'A4V 35'))
f.fig2ab = na.omit(f.fig2ab)
f.fig2ab$x = as.numeric(f.fig2ab$Sample)
f.fig2ab$x[f.fig2ab$x > 4] = f.fig2ab$x[f.fig2ab$x > 4] + 1
f.fig2ab$x[f.fig2ab$x > 2] = f.fig2ab$x[f.fig2ab$x > 2] + 1
f.fig2ab$color = 'brown'
f.fig2ab$color[f.fig2ab$Genotype == 'dSodWT'] = 'steelblue'
f.fig2ab = f.fig2ab[order(f.fig2ab$Rate),]
f.fig2ab = f.fig2ab[order(f.fig2ab$Sample),]
f.fig2ab$x.adj = rep(c(-.25, .25 ,0), 12)
f.fig2ab$x.adj = c(0, -.15, .15, -.1, .1, 0,
                   0, -.1, .1, -.1, .1, 0,

                   -.15, .15, 0, 0, 0, 0,
                   -.15, .15, 0, 0, 0, 0,

                   -.15, .15, 0, -.15, .15, 0,
                   -.15, .15, 0, -.15, .15, 0)

f.fig2ab$x = f.fig2ab$x + f.fig2ab$x.adj

par(mar=c(5.1, 4.1, 4.1, 5.6))

boxplot(f.fig2ab$Rate~f.fig2ab$Sample,
        at = c(1,2, 4,5, 7,8),
        xaxt = 'none',
        xlab = "Day",
        ylab = 'Veloxicty (cm/sec)',
        main = "Female Climbing Velocity")
points(x = f.fig2ab$x,
       y = f.fig2ab$Rate,
       pch = 21,
       bg = f.fig2ab$color)
axis(side = 1,
     at = c(1.5, 4.5, 7.5),
     labels = c(3, 28,35))

legend('topright',
       inset=c(-0.175,0),
       legend = c('Silent', 'A4V'), 
       fill = c('steelblue', 'brown'),
       cex = .65,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

```

```{r echo = F, include = F}

m.fig2ab = fig2ab[fig2ab$Sex == 'M' &
                  (fig2ab$Day == 3 |
                  fig2ab$Day == 10 |
                  fig2ab$Day == 16) &
                  fig2ab$Genotype != 'w1118',]

m.fig2ab = m.fig2ab[order(fig2ab$Day),]
m.fig2ab$Sample = paste(m.fig2ab$Genotype, m.fig2ab$Day)
m.fig2ab$Sample = factor(m.fig2ab$Sample, levels = c('dSodWT 3', 'A4V 3',
                                                     'dSodWT 10', 'A4V 10',
                                                     'dSodWT 16', 'A4V 16'))
m.fig2ab = na.omit(m.fig2ab)
m.fig2ab$x = as.numeric(m.fig2ab$Sample)
m.fig2ab$x[m.fig2ab$x > 4] = m.fig2ab$x[m.fig2ab$x > 4] + 1
m.fig2ab$x[m.fig2ab$x > 2] = m.fig2ab$x[m.fig2ab$x > 2] + 1
m.fig2ab$color = 'brown'
m.fig2ab$color[m.fig2ab$Genotype == 'dSodWT'] = 'steelblue'

m.fig2ab = m.fig2ab[order(m.fig2ab$Rate),]
m.fig2ab = m.fig2ab[order(m.fig2ab$Sample),]
m.fig2ab$x.adj = rep(c(-.25, .25 ,0), 12)
m.fig2ab$x.adj = c(0, -.25, 0, .25, .15, 0,
                   0, -.25, 0, .25, -.15, .15,

                   0, 0, 0, -.1, .1, 0,
                   0, 0, -.1, .1, -.1, .1,

                   0, 0, 0, 0, 0, 0,
                   0, -.25, .25, 0, 0, 0)

m.fig2ab$x = m.fig2ab$x + m.fig2ab$x.adj

par(mar=c(5.1, 4.1, 4.1, 5.6))

boxplot(m.fig2ab$Rate~m.fig2ab$Sample,
        at = c(1,2, 4,5, 7,8),
        xaxt = 'none',
        xlab = "Day",
        ylab = 'Veloxicty (cm/sec)',
        main = "Male Climbing Velocity")
points(x = m.fig2ab$x,
       y = m.fig2ab$Rate,
       pch = 21,
       bg = m.fig2ab$color)
axis(side = 1,
     at = c(1.5, 4.5, 7.5),
     labels = c(3, 10, 16))

legend('topright',
       inset=c(-0.175,0),
       legend = c('Silent', 'A4V'), 
       fill = c('steelblue', 'brown'),
       cex = .65,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

```

```{r echo = F, include = F}

f.fig2ab = fig2ab[fig2ab$Sex == 'F' &
                 (fig2ab$Day == 3 |
                  fig2ab$Day == 14 |
                  fig2ab$Day == 24 |
                  fig2ab$Day == 35) &
                  fig2ab$Genotype != 'w1118',]

f.fig2ab = f.fig2ab[order(fig2ab$Day),]
f.fig2ab$Sample = paste(f.fig2ab$Genotype, f.fig2ab$Day)
f.fig2ab$Sample = factor(f.fig2ab$Sample, levels = c('dSodWT 3', 'A4V 3',
                                                     'dSodWT 14', 'A4V 14',
                                                     'dSodWT 24', 'A4V 24',
                                                     'dSodWT 35', 'A4V 35'))
f.fig2ab = na.omit(f.fig2ab)
f.fig2ab$x = as.numeric(f.fig2ab$Sample)
f.fig2ab$x[f.fig2ab$x > 6] = f.fig2ab$x[f.fig2ab$x > 6] + 1
f.fig2ab$x[f.fig2ab$x > 4] = f.fig2ab$x[f.fig2ab$x > 4] + 1
f.fig2ab$x[f.fig2ab$x > 2] = f.fig2ab$x[f.fig2ab$x > 2] + 1
f.fig2ab$color = 'brown'
f.fig2ab$color[f.fig2ab$Genotype == 'dSodWT'] = 'steelblue'
f.fig2ab = f.fig2ab[order(f.fig2ab$Rate),]
f.fig2ab = f.fig2ab[order(f.fig2ab$Sample),]
f.fig2ab$x.adj = rep(c(-.25, .25 ,0), 16)
f.fig2ab$x.adj = c(0, -.25, .25, -.15, .15, 0,
                   0, -.15, .15, -.15, .15, 0,

                   0, 0, .2, -.2, 0, 0,
                   -.15, .15, 0, 0, -.15, .15,

                   0, 0, -.15, .15, 0, 0,
                   0, .2, -.2, 0, -.15, .15,

                   0, -.25, .25, -.15, .15, 0,
                   -.25, .25, 0, -.25, .25, 0)

f.fig2ab$x = f.fig2ab$x + f.fig2ab$x.adj

par(mar=c(5.1, 4.1, 4.1, 5.6))

boxplot(f.fig2ab$Rate~f.fig2ab$Sample,
        at = c(1,2, 4,5, 7,8, 10,11),
        xaxt = 'none',
        xlab = "Day",
        ylab = 'Veloxicty (cm/sec)',
        main = "Female Climbing Velocity")
points(x = f.fig2ab$x,
       y = f.fig2ab$Rate,
       pch = 21,
       bg = f.fig2ab$color)
axis(side = 1,
     at = c(1.5, 4.5, 7.5, 10.5),
     labels = c(3, 14, 24, 35))

legend('topright',
       inset=c(-0.175,0),
       legend = c('Silent', 'A4V'), 
       fill = c('steelblue', 'brown'),
       cex = .65,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

```

```{r echo = F, fig.width=7.5}

f.fig2ab = fig2ab[fig2ab$Sex == 'F' &
                 (fig2ab$Day == 3 |
                  fig2ab$Day == 14 |
                  fig2ab$Day == 20 |
                  fig2ab$Day == 28 |
                  fig2ab$Day == 35) &
                  fig2ab$Genotype != 'w1118',]

f.fig2ab = f.fig2ab[order(fig2ab$Day),]
f.fig2ab$Sample = paste(f.fig2ab$Genotype, f.fig2ab$Day)
f.fig2ab$Sample = factor(f.fig2ab$Sample, levels = c('dSodWT 3', 'A4V 3',
                                                     'dSodWT 14', 'A4V 14',
                                                     'dSodWT 20', 'A4V 20',
                                                     'dSodWT 28', 'A4V 28',
                                                     'dSodWT 35', 'A4V 35'))
f.fig2ab = na.omit(f.fig2ab)
f.fig2ab$x = as.numeric(f.fig2ab$Sample)
f.fig2ab$x[f.fig2ab$x > 8] = f.fig2ab$x[f.fig2ab$x > 8] + 1
f.fig2ab$x[f.fig2ab$x > 6] = f.fig2ab$x[f.fig2ab$x > 6] + 1
f.fig2ab$x[f.fig2ab$x > 4] = f.fig2ab$x[f.fig2ab$x > 4] + 1
f.fig2ab$x[f.fig2ab$x > 2] = f.fig2ab$x[f.fig2ab$x > 2] + 1
f.fig2ab$color = 'brown'
f.fig2ab$color[f.fig2ab$Genotype == 'dSodWT'] = 'steelblue'
f.fig2ab = f.fig2ab[order(f.fig2ab$Rate),]
f.fig2ab = f.fig2ab[order(f.fig2ab$Sample),]
f.fig2ab$x.adj = rep(c(-.25, .25 ,0), 20)
f.fig2ab$x.adj = c(0, -.25, .25, -.15, .15, 0,
                   0, -.15, .15, -.15, .15, 0,

                   0, 0, .2, -.2, 0, 0,
                   -.15, .15, 0, 0, -.15, .15,

                   0, 0, -.15, .15, 0, 0,
                   0, 0, -.25, .25, 0, 0,

                   0, -.25, .25, 0, 0, 0,
                   -.15, .15, 0, 0, 0, 0,

                   0, -.25, .25, -.15, .15, 0,
                   -.25, .25, 0, -.25, .25, 0)

f.fig2ab$x = f.fig2ab$x + f.fig2ab$x.adj

par(mar=c(5.1, 4.1, 4.1, 5.6))

boxplot(f.fig2ab$Rate~f.fig2ab$Sample,
        at = c(1,2, 4,5, 7,8, 10,11, 13,14),
        xaxt = 'none',
        xlab = "Day",
        ylab = 'Veloxicty (cm/sec)',
        main = "Female Climbing Velocity")
points(x = f.fig2ab$x,
       y = f.fig2ab$Rate,
       pch = 21,
       bg = f.fig2ab$color)
axis(side = 1,
     at = c(1.5, 4.5, 7.5, 10.5, 13.5),
     labels = c(3, 14, 20, 28, 35))

legend('topright',
       inset=c(-0.175,0),
       legend = c('Silent', 'A4V'), 
       fill = c('steelblue', 'brown'),
       cex = 1,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

```

```{r echo = F, fig.width=7.5}

m.fig2ab = fig2ab[fig2ab$Sex == 'M' &
                 (fig2ab$Day == 3 |
                  fig2ab$Day == 7 |
                  fig2ab$Day == 12 |
                  fig2ab$Day == 16) &
                  fig2ab$Genotype != 'w1118',]

m.fig2ab = m.fig2ab[order(fig2ab$Day),]
m.fig2ab$Sample = paste(m.fig2ab$Genotype, m.fig2ab$Day)
m.fig2ab$Sample = factor(m.fig2ab$Sample, levels = c('dSodWT 3', 'A4V 3',
                                                     'dSodWT 7', 'A4V 7',
                                                     'dSodWT 12', 'A4V 12',
                                                     'dSodWT 16', 'A4V 16'))
m.fig2ab = na.omit(m.fig2ab)
m.fig2ab$x = as.numeric(m.fig2ab$Sample)
m.fig2ab$x[m.fig2ab$x > 6] = m.fig2ab$x[m.fig2ab$x > 6] + 1
m.fig2ab$x[m.fig2ab$x > 4] = m.fig2ab$x[m.fig2ab$x > 4] + 1
m.fig2ab$x[m.fig2ab$x > 2] = m.fig2ab$x[m.fig2ab$x > 2] + 1
m.fig2ab$color = 'brown'
m.fig2ab$color[m.fig2ab$Genotype == 'dSodWT'] = 'steelblue'

m.fig2ab = m.fig2ab[order(m.fig2ab$Rate),]
m.fig2ab = m.fig2ab[order(m.fig2ab$Sample),]
m.fig2ab$x.adj = rep(c(-.25, .25 ,0), 16)
m.fig2ab$x.adj = c(0, -.25, .25, 0, -.1, .1,
                   0, -.25, .25, .0, -.1, .1,

                   0, 0, -.1, .1, 0, 0,
                   0, -.1, .1, 0, 0, 0,

                   0, -.1, .1, 0, 0, 0,
                   0, -.1, .1, 0, -.1, .1,

                   0, -.25, 0, .25, .15, 0,
                   0, -.25, 0, .25, -.15, .15)

m.fig2ab$x = m.fig2ab$x + m.fig2ab$x.adj

par(mar=c(5.1, 4.1, 4.1, 5.6))

boxplot(m.fig2ab$Rate~m.fig2ab$Sample,
        at = c(1,2, 4,5, 7,8, 10,11),
        xaxt = 'none',
        xlab = "Day",
        ylab = 'Veloxicty (cm/sec)',
        main = "Male Climbing Velocity")
points(x = m.fig2ab$x,
       y = m.fig2ab$Rate,
       pch = 21,
       bg = m.fig2ab$color)
axis(side = 1,
     at = c(1.5, 4.5, 7.5, 10.5),
     labels = c(3, 7, 12, 16))

legend('topright',
       inset=c(-0.175,0),
       legend = c('Silent', 'A4V'), 
       fill = c('steelblue', 'brown'),
       cex = 1,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

```

```{r include = F, echo = F}

fig4ab = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 4AB.csv", header = T)

female = fig4ab[grep('f', fig4ab$vial_ID),]
male = fig4ab[grep('m', fig4ab$vial_ID),]

f.con = na.omit(female[female$geno == 'silent' | female$geno == 'A4V',])
f.cha = na.omit(female[grep('cha', female$geno),])

fallcha = rbind(f.con, f.cha)
fallcha$sample = paste0(fallcha$geno, fallcha$day)

fcha.mean = unique(fallcha[,c('geno', 'day')])
fcha.mean$sample = paste0(fcha.mean$geno, fcha.mean$day)
fcha.mean$mean = 0

i=1
while(i<=nrow(fcha.mean)){
  fcha.mean$mean[i] = mean(fallcha[fallcha$sample == fcha.mean$sample[i], 'corrected.values'])
  i = i+1
}

fcha.mean = fcha.mean[order(fcha.mean$day),]

m.con = na.omit(male[male$geno == 'silent' | male$geno == 'A4V',])
m.cha = na.omit(male[grep('cha', male$geno),])

mallcha = rbind(m.con, m.cha)
mallcha$sample = paste0(mallcha$geno, mallcha$day)

mcha.mean = unique(mallcha[,c('geno', 'day')])
mcha.mean$sample = paste0(mcha.mean$geno, mcha.mean$day)
mcha.mean$mean = 0

i=1
while(i<=nrow(mcha.mean)){
  mcha.mean$mean[i] = mean(mallcha[mallcha$sample == mcha.mean$sample[i], 'corrected.values'])
  i = i+1
}

mcha.mean = mcha.mean[order(mcha.mean$day),]

```

```{r echo = F, include = F}

f.fig4a = fig4ab[grep('f', fig4ab$vial_ID),]
f.fig4a = f.fig4a[f.fig4a$day == 6 |
                  f.fig4a$day == 13 |
                  #f.fig4a$day == 19 |
                  f.fig4a$day == 27,]

f.fig4a$Sample = paste(f.fig4a$geno, f.fig4a$day)
f.fig4a$geno = factor(f.fig4a$geno, levels = c("silent", 'silentcha', 'silentok',
                                               'A4V', 'A4Vcha', 'A4Vok'))

f.fig4a = f.fig4a[order(f.fig4a$day),]
f.fig4a = f.fig4a[order(f.fig4a$geno),]
f.fig4a$Sample = factor(f.fig4a$Sample, levels = unique(f.fig4a$Sample))

f.fig4a = na.omit(f.fig4a)
f.fig4a$x = as.numeric(f.fig4a$Sample)
f.fig4a$x[f.fig4a$x > 15] = f.fig4a$x[f.fig4a$x > 15] + 1
f.fig4a$x[f.fig4a$x > 12] = f.fig4a$x[f.fig4a$x > 12] + 1
f.fig4a$x[f.fig4a$x > 9] = f.fig4a$x[f.fig4a$x > 9] + 2
f.fig4a$x[f.fig4a$x > 6] = f.fig4a$x[f.fig4a$x > 6] + 1
f.fig4a$x[f.fig4a$x > 3] = f.fig4a$x[f.fig4a$x > 3] + 1
#f.fig4a$x[f.fig4a$x > 18] = f.fig4a$x[f.fig4a$x > 18] + 1
#f.fig4a$x[f.fig4a$x > 12] = f.fig4a$x[f.fig4a$x > 12] + 1
#f.fig4a$x[f.fig4a$x > 6] = f.fig4a$x[f.fig4a$x > 6] + 1
f.fig4a$color = 'steelblue'
f.fig4a$color[f.fig4a$geno == 'A4V'] = 'navy'
f.fig4a$color[f.fig4a$geno == 'silentcha'] = 'limegreen'
f.fig4a$color[f.fig4a$geno == 'A4Vcha'] = 'darkgreen'
f.fig4a$color[f.fig4a$geno == 'silentok'] = 'brown1'
f.fig4a$color[f.fig4a$geno == 'A4Vok'] = 'firebrick'

f.fig4a = f.fig4a[order(f.fig4a$corrected.values),]
f.fig4a = f.fig4a[order(f.fig4a$Sample),]
f.fig4a$x.adj = rep(c(-.25, .25 ,0), 36)

f.fig4a$x = f.fig4a$x + f.fig4a$x.adj

par(mar=c(5.1, 4.1, 4.1, 5.6))

boxplot(f.fig4a$corrected.values~f.fig4a$Sample,
        ##at = c(1:6, 8:13, 15:20, 22:27),
        at = c(1:3, 5:7, 9:11, 14:16, 18:20, 22:24),
        xaxt = 'none',
        xlab = "Day",
        ylab = 'Veloxicty (cm/sec)',
        main = "Female Climbing Velocity")
points(x = f.fig4a$x,
       y = f.fig4a$corrected.values,
       pch = 21,
       bg = f.fig4a$color,
       cex = 1)
axis(side = 1,
     #at = c(3.5, 10.5, 17.5, 24.5),
     at = c(1:3, 5:7, 9:11, 14:16, 18:20, 22:24),
     labels = c(rep("", 18)),
     cex.axis = .5,
     line = 0)

axis(side = 1,
     #at = c(3.5, 10.5, 17.5, 24.5),
     at = c(1:3, 5:7, 9:11, 14:16, 18:20, 22:24),
     labels = c(rep(c(6, 13, 27), 6)),
     cex.axis = .5,
     tick = F,
     line = -.5)

axis(side = 3,
     #at = c(3.5, 10.5, 17.5, 24.5),
     at = c(6, 19),
     labels = c('Silent', 'A4V'),
     cex.axis = 1,
     tick = F,
     line = -.75)

legend('topright',
       inset=c(-0.225,0),
       legend = c('Silent', 'A4V',
                  'ChAT>gbb;Silent', 'ChAT>gbb;A4V',
                  'OK371>gbb;Silent', 'OK371>gbb;A4V'), 
       fill = c('steelblue', 'limegreen', 'brown1',
                'navy', 'darkgreen', 'firebrick'),
       cex = .65,
       bty = 'n',
       pt.cex = .75,
       xpd = T)

abline(v = 12.5, lty = 2, lwd =2)

```

```{r echo = F, include = F}

f.fig4a = fig4ab[grep('f', fig4ab$vial_ID),]
f.fig4a = f.fig4a[f.fig4a$day == 6 |
                  f.fig4a$day == 13 |
                  #f.fig4a$day == 19 |
                  f.fig4a$day == 27,]

f.fig4a$Sample = paste(f.fig4a$geno, f.fig4a$day)
f.fig4a$geno = factor(f.fig4a$geno, levels = c("silent", 'A4V',
                                               'silentcha', 'A4Vcha',
                                               'silentok', 'A4Vok'))

f.fig4a = f.fig4a[order(f.fig4a$day),]
f.fig4a = f.fig4a[order(f.fig4a$geno),]
f.fig4a$Sample = factor(f.fig4a$Sample, levels = unique(f.fig4a$Sample))

f.fig4a = na.omit(f.fig4a)
f.fig4a$x = as.numeric(f.fig4a$Sample)
f.fig4a$x[f.fig4a$x > 15] = f.fig4a$x[f.fig4a$x > 15] + 1
f.fig4a$x[f.fig4a$x > 12] = f.fig4a$x[f.fig4a$x > 12] + 2
f.fig4a$x[f.fig4a$x > 9] = f.fig4a$x[f.fig4a$x > 9] + 1
f.fig4a$x[f.fig4a$x > 6] = f.fig4a$x[f.fig4a$x > 6] + 2
f.fig4a$x[f.fig4a$x > 3] = f.fig4a$x[f.fig4a$x > 3] + 1
#f.fig4a$x[f.fig4a$x > 18] = f.fig4a$x[f.fig4a$x > 18] + 1
#f.fig4a$x[f.fig4a$x > 12] = f.fig4a$x[f.fig4a$x > 12] + 1
#f.fig4a$x[f.fig4a$x > 6] = f.fig4a$x[f.fig4a$x > 6] + 1
f.fig4a$color = 'steelblue'
f.fig4a$color[f.fig4a$geno == 'A4V'] = 'navy'
f.fig4a$color[f.fig4a$geno == 'silentcha'] = 'limegreen'
f.fig4a$color[f.fig4a$geno == 'A4Vcha'] = 'darkgreen'
f.fig4a$color[f.fig4a$geno == 'silentok'] = 'brown1'
f.fig4a$color[f.fig4a$geno == 'A4Vok'] = 'firebrick'

f.fig4a = f.fig4a[order(f.fig4a$corrected.values),]
f.fig4a = f.fig4a[order(f.fig4a$Sample),]
f.fig4a$x.adj = rep(c(-.25, .25 ,0), 36)

f.fig4a$x = f.fig4a$x + f.fig4a$x.adj

par(mar=c(5.1, 4.1, 4.1, 5.6))

boxplot(f.fig4a$corrected.values~f.fig4a$Sample,
        ##at = c(1:6, 8:13, 15:20, 22:27),
        at = c(1:3, 5:7, 10:12, 14:16, 19:21, 23:25),
        xaxt = 'none',
        xlab = "Day",
        ylab = 'Veloxicty (cm/sec)',
        main = "Female Climbing Velocity")
points(x = f.fig4a$x,
       y = f.fig4a$corrected.values,
       pch = 21,
       bg = f.fig4a$color,
       cex = 1)
axis(side = 1,
     #at = c(3.5, 10.5, 17.5, 24.5),
     at = c(1:3, 5:7, 10:12, 14:16, 19:21, 23:25),
     labels = c(rep("", 18)),
     cex.axis = .5,
     line = 0)

axis(side = 1,
     #at = c(3.5, 10.5, 17.5, 24.5),
     at = c(1:3, 5:7, 10:12, 14:16, 19:21, 23:25),
     labels = c(rep(c(6, 13, 27), 6)),
     cex.axis = .75,
     tick = F,
     line = -.5)
axis(side = 1,
     #at = c(3.5, 10.5, 17.5, 24.5),
     at = c(2, 6, 11, 15, 20, 24),
     labels = c(rep(c('Silent', 'A4V'), 3)),
     cex.axis = 1,
     tick = F,
     line = .5)

axis(side = 3,
     #at = c(3.5, 10.5, 17.5, 24.5),
     at = c(4, 13, 22),
     labels = c('Control', 'ChAT>gbb', 'OK371>gbb'),
     cex.axis = 1,
     tick = F,
     line = -.75)

legend('topright',
       inset=c(-0.225,0),
       legend = c('Silent', 'A4V',
                  'ChAT>gbb;Silent', 'ChAT>gbb;A4V',
                  'OK371>gbb;Silent', 'OK371>gbb;A4V'), 
       fill = c('steelblue', 'navy',
                'limegreen', 'darkgreen',
                'brown1', 'firebrick'),
       cex = .65,
       bty = 'n',
       pt.cex = .75,
       xpd = T)

abline(v = c(8.5, 17.5), lty = 2, lwd =2)

```

```{r echo = F, include = F}

plot_ly(x = fallcha$day,
        y = fallcha$corrected.values,
        name = fallcha$geno,
        type = 'scatter',
        color = fallcha$color,
        marker = list(colors = fallcha$color)) %>%
  add_lines(x = fcha.mean$day,
            y = fcha.mean$mean,
            name = fcha.mean$geno,
            color = fcha.mean$color,
            line = list(color = fcha.mean$color))

```

```{r echo = F, include = F}

female = fig4ab[grep('f', fig4ab$vial_ID),]
female$sample = paste0(female$geno, female$day)

fmean = unique(female[,c('geno', 'day', 'sample')])
fstd = fmean
eb = fmean
fmean$mean = 0
fstd$std = 0

i=1
while(i<=nrow(fmean)){
  fmean$mean[i] = mean(na.omit(female[female$sample == fmean$sample[i], 'corrected.values']))

  fstd$std[i] = (var(na.omit(female[female$sample == fmean$sample[i], 'corrected.values'])))

  i = i+1
}

flm = data.frame(geno = unique(female$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(flm)){
  temp=na.omit(female[female$geno == flm$geno[i], ])
  temp.lm = lm(temp$corrected.values~temp$day)
  flm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  flm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(flm) = flm$geno

fmean = fmean[order(fmean$day),]
fstd = fstd[order(fstd$day),]
eb = eb[order(eb$day),]

eb$y0 = fmean$mean - fstd$std
eb$y1 = fmean$mean + fstd$std

```

```{r echo = F, include = F}

male = fig4ab[grep('m', fig4ab$vial_ID),]
male$sample = paste0(male$geno, male$day)

mmean = unique(male[,c('geno', 'day', 'sample')])
mstd = mmean
eb = mmean
mmean$mean = 0
mstd$std = 0

i=1
while(i<=nrow(mmean)){
  mmean$mean[i] = mean(na.omit(male[male$sample == mmean$sample[i], 'corrected.values']))

  mstd$std[i] = (var(na.omit(male[male$sample == mmean$sample[i], 'corrected.values'])))

  i = i+1
}

mlm = data.frame(geno = unique(male$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(mlm)){
  temp=na.omit(male[male$geno == mlm$geno[i], ])
  temp.lm = lm(temp$corrected.values~temp$day)
  mlm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  mlm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(mlm) = mlm$geno

mmean = mmean[order(mmean$day),]
mstd = mstd[order(mstd$day),]
eb = eb[order(eb$day),]

eb$y0 = mmean$mean - mstd$std
eb$y1 = mmean$mean + mstd$std

```

```{r echo = F, fig.width=10}

lcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
pcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, max(c(fmean$mean, mmean$mean))),
     xlim = c(2,27),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     main = "Females")

axis(1, at = fmean[fmean$geno == 'silent', 'day'])

lines(x = fmean[fmean$geno == 'silent', 'day'],
      y = fmean[fmean$geno == 'silent', 'mean'],
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = fmean[fmean$geno == 'A4V', 'day'],
      y = fmean[fmean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = fmean[fmean$geno == 'silentcha', 'day'],
      y = fmean[fmean$geno == 'silentcha', 'mean'],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = fmean[fmean$geno == 'A4Vcha', 'day'],
      y = fmean[fmean$geno == 'A4Vcha', 'mean'],
      type = 'l',
      col = pcol[4],
      lwd = 2)

lines(x = fmean[fmean$geno == 'silentok', 'day'],
      y = fmean[fmean$geno == 'silentok', 'mean'],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = fmean[fmean$geno == 'A4Vok', 'day'],
      y = fmean[fmean$geno == 'A4Vok', 'mean'],
      type = 'l',
      col = pcol[6],
      lwd = 2)

points(x = fmean[fmean$geno == 'silent', 'day'],
       y = fmean[fmean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = fmean[fmean$geno == 'A4V', 'day'],
       y = fmean[fmean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = fmean[fmean$geno == 'silentcha', 'day'],
       y = fmean[fmean$geno == 'silentcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = fmean[fmean$geno == 'A4Vcha', 'day'],
       y = fmean[fmean$geno == 'A4Vcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = fmean[fmean$geno == 'silentok', 'day'],
       y = fmean[fmean$geno == 'silentok', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = fmean[fmean$geno == 'A4Vok', 'day'],
       y = fmean[fmean$geno == 'A4Vok', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)

points(x = fmean[fmean$geno == 'silent', 'day'],
       y = fmean[fmean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = fmean[fmean$geno == 'A4V', 'day'],
       y = fmean[fmean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = fmean[fmean$geno == 'silentcha', 'day'],
       y = fmean[fmean$geno == 'silentcha', 'mean'],
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = fmean[fmean$geno == 'A4Vcha', 'day'],
       y = fmean[fmean$geno == 'A4Vcha', 'mean'],
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)
points(x = fmean[fmean$geno == 'silentok', 'day'],
       y = fmean[fmean$geno == 'silentok', 'mean'],
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex =2)
points(x = fmean[fmean$geno == 'A4Vok', 'day'],
       y = fmean[fmean$geno == 'A4Vok', 'mean'],
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex =2)

legend(x = 29,
       y = 1.8,
       legend = c('Silent', 'A4V',
                  'ChAT>gbb;Silent', 'ChAT>gbb;A4V', 
                  'OK371>gbb;Silent', 'OK371>gbb;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[3], .5),
                hex(pcol[4], .5), hex(pcol[5], .5), hex(pcol[6], .5)),
       bty = 'n',
       y.intersp = 1,
       pt.cex = 2,
       xpd=TRUE)

```

```{r echo = F, fig.width=10}

lcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
pcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, max(c(fmean$mean, mmean$mean))),
     xlim = c(2,27),
     ylab = 'Mean Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     main = "Males")

axis(1, at = mmean[mmean$geno == 'silent', 'day'])

lines(x = mmean[mmean$geno == 'silent', 'day'],
      y = mmean[mmean$geno == 'silent', 'mean'],
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = mmean[mmean$geno == 'A4V', 'day'],
      y = mmean[mmean$geno == 'A4V', 'mean'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = mmean[mmean$geno == 'silentcha', 'day'],
      y = mmean[mmean$geno == 'silentcha', 'mean'],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = mmean[mmean$geno == 'A4Vcha', 'day'],
      y = mmean[mmean$geno == 'A4Vcha', 'mean'],
      type = 'l',
      col = pcol[4],
      lwd = 2)

lines(x = mmean[mmean$geno == 'silentok', 'day'],
      y = mmean[mmean$geno == 'silentok', 'mean'],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = mmean[mmean$geno == 'A4Vok', 'day'],
      y = mmean[mmean$geno == 'A4Vok', 'mean'],
      type = 'l',
      col = pcol[6],
      lwd = 2)

points(x = mmean[mmean$geno == 'silent', 'day'],
       y = mmean[mmean$geno == 'silent', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4V', 'day'],
       y = mmean[mmean$geno == 'A4V', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = mmean[mmean$geno == 'silentcha', 'day'],
       y = mmean[mmean$geno == 'silentcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4Vcha', 'day'],
       y = mmean[mmean$geno == 'A4Vcha', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = mmean[mmean$geno == 'silentok', 'day'],
       y = mmean[mmean$geno == 'silentok', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4Vok', 'day'],
       y = mmean[mmean$geno == 'A4Vok', 'mean'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)

points(x = mmean[mmean$geno == 'silent', 'day'],
       y = mmean[mmean$geno == 'silent', 'mean'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = mmean[mmean$geno == 'A4V', 'day'],
       y = mmean[mmean$geno == 'A4V', 'mean'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = mmean[mmean$geno == 'silentcha', 'day'],
       y = mmean[mmean$geno == 'silentcha', 'mean'],
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4Vcha', 'day'],
       y = mmean[mmean$geno == 'A4Vcha', 'mean'],
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)
points(x = mmean[mmean$geno == 'silentok', 'day'],
       y = mmean[mmean$geno == 'silentok', 'mean'],
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex =2)
points(x = mmean[mmean$geno == 'A4Vok', 'day'],
       y = mmean[mmean$geno == 'A4Vok', 'mean'],
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex =2)

legend(x = 29,
       y = 1.8,
       legend = c('Silent', 'A4V',
                  'ChAT>gbb;Silent', 'ChAT>gbb;A4V', 
                  'OK371>gbb;Silent', 'OK371>gbb;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[3], .5),
                hex(pcol[4], .5), hex(pcol[5], .5), hex(pcol[6], .5)),
       bty = 'n',
       y.intersp = 1,
       pt.cex = 2,
       xpd=TRUE)

```

```{r echo = F, fig.width=10}

lcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
pcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

fmean.percent = fmean
fmean.percent$percent = 0

fmean.percent = fmean.percent[fmean.percent$day != 2, ]

fmean.percent[fmean.percent$geno == 'silent', 'percent'] = fmean.percent[fmean.percent$geno == 'silent', 'mean']/fmean.percent[fmean.percent$geno == 'silent' & fmean.percent$day == 6, 'mean']

fmean.percent[fmean.percent$geno == 'A4V', 'percent'] = fmean.percent[fmean.percent$geno == 'A4V', 'mean']/fmean.percent[fmean.percent$geno == 'A4V' & fmean.percent$day == 6, 'mean']

fmean.percent[fmean.percent$geno == 'silentok', 'percent'] = fmean.percent[fmean.percent$geno == 'silentok', 'mean']/fmean.percent[fmean.percent$geno == 'silentok' & fmean.percent$day == 6, 'mean']

fmean.percent[fmean.percent$geno == 'A4Vok', 'percent'] = fmean.percent[fmean.percent$geno == 'A4Vok', 'mean']/fmean.percent[fmean.percent$geno == 'A4Vok' & fmean.percent$day == 6, 'mean']

fmean.percent[fmean.percent$geno == 'A4Vcha', 'percent'] = fmean.percent[fmean.percent$geno == 'A4Vcha', 'mean']/fmean.percent[fmean.percent$geno == 'A4Vcha' & fmean.percent$day == 6, 'mean']

fmean.percent[fmean.percent$geno == 'silentcha', 'percent'] = fmean.percent[fmean.percent$geno == 'silentcha', 'mean']/fmean.percent[fmean.percent$geno == 'silentcha' & fmean.percent$day == 6, 'mean']

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, 1.25),
     xlim = c(6,27),
     ylab = 'Percent of Initial Mean Velocity',
     xlab = 'Day',
     xaxt = 'n',
     main = "Females")

axis(1, at = fmean.percent[fmean.percent$geno == 'silent', 'day'])

lines(x = fmean.percent[fmean.percent$geno == 'silent', 'day'],
      y = fmean.percent[fmean.percent$geno == 'silent', 'percent'],
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = fmean.percent[fmean.percent$geno == 'A4V', 'day'],
      y = fmean.percent[fmean.percent$geno == 'A4V', 'percent'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = fmean.percent[fmean.percent$geno == 'silentcha', 'day'],
      y = fmean.percent[fmean.percent$geno == 'silentcha', 'percent'],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = fmean.percent[fmean.percent$geno == 'A4Vcha', 'day'],
      y = fmean.percent[fmean.percent$geno == 'A4Vcha', 'percent'],
      type = 'l',
      col = pcol[4],
      lwd = 2)

lines(x = fmean.percent[fmean.percent$geno == 'silentok', 'day'],
      y = fmean.percent[fmean.percent$geno == 'silentok', 'percent'],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = fmean.percent[fmean.percent$geno == 'A4Vok', 'day'],
      y = fmean.percent[fmean.percent$geno == 'A4Vok', 'percent'],
      type = 'l',
      col = pcol[6],
      lwd = 2)

points(x = fmean.percent[fmean.percent$geno == 'silent', 'day'],
       y = fmean.percent[fmean.percent$geno == 'silent', 'percent'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = fmean.percent[fmean.percent$geno == 'A4V', 'day'],
       y = fmean.percent[fmean.percent$geno == 'A4V', 'percent'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = fmean.percent[fmean.percent$geno == 'silentcha', 'day'],
       y = fmean.percent[fmean.percent$geno == 'silentcha', 'percent'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = fmean.percent[fmean.percent$geno == 'A4Vcha', 'day'],
       y = fmean.percent[fmean.percent$geno == 'A4Vcha', 'percent'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = fmean.percent[fmean.percent$geno == 'silentok', 'day'],
       y = fmean.percent[fmean.percent$geno == 'silentok', 'percent'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = fmean.percent[fmean.percent$geno == 'A4Vok', 'day'],
       y = fmean.percent[fmean.percent$geno == 'A4Vok', 'percent'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)

points(x = fmean.percent[fmean.percent$geno == 'silent', 'day'],
       y = fmean.percent[fmean.percent$geno == 'silent', 'percent'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = fmean.percent[fmean.percent$geno == 'A4V', 'day'],
       y = fmean.percent[fmean.percent$geno == 'A4V', 'percent'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = fmean.percent[fmean.percent$geno == 'silentcha', 'day'],
       y = fmean.percent[fmean.percent$geno == 'silentcha', 'percent'],
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = fmean.percent[fmean.percent$geno == 'A4Vcha', 'day'],
       y = fmean.percent[fmean.percent$geno == 'A4Vcha', 'percent'],
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)
points(x = fmean.percent[fmean.percent$geno == 'silentok', 'day'],
       y = fmean.percent[fmean.percent$geno == 'silentok', 'percent'],
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex =2)
points(x = fmean.percent[fmean.percent$geno == 'A4Vok', 'day'],
       y = fmean.percent[fmean.percent$geno == 'A4Vok', 'percent'],
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex =2)

legend(x = 29,
       y = 1,
       legend = c('Silent', 'A4V',
                  'ChAT>gbb;Silent', 'ChAT>gbb;A4V', 
                  'OK371>gbb;Silent', 'OK371>gbb;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[3], .5),
                hex(pcol[4], .5), hex(pcol[5], .5), hex(pcol[6], .5)),
       bty = 'n',
       y.intersp = 1,
       pt.cex = 2,
       xpd=TRUE)

```

```{r echo = F, fig.width=10}

lcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
pcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')
hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

mmean.percent = mmean
mmean.percent$percent = 0

mmean.percent[mmean.percent$geno == 'silent', 'percent'] = mmean.percent[mmean.percent$geno == 'silent', 'mean']/mmean.percent[mmean.percent$geno == 'silent' & mmean.percent$day == 2, 'mean']

mmean.percent[mmean.percent$geno == 'A4V', 'percent'] = mmean.percent[mmean.percent$geno == 'A4V', 'mean']/mmean.percent[mmean.percent$geno == 'A4V' & mmean.percent$day == 2, 'mean']

mmean.percent[mmean.percent$geno == 'silentok', 'percent'] = mmean.percent[mmean.percent$geno == 'silentok', 'mean']/mmean.percent[mmean.percent$geno == 'silentok' & mmean.percent$day == 2, 'mean']

mmean.percent[mmean.percent$geno == 'A4Vok', 'percent'] = mmean.percent[mmean.percent$geno == 'A4Vok', 'mean']/mmean.percent[mmean.percent$geno == 'A4Vok' & mmean.percent$day == 2, 'mean']

mmean.percent[mmean.percent$geno == 'A4Vcha', 'percent'] = mmean.percent[mmean.percent$geno == 'A4Vcha', 'mean']/mmean.percent[mmean.percent$geno == 'A4Vcha' & mmean.percent$day == 2, 'mean']

mmean.percent[mmean.percent$geno == 'silentcha', 'percent'] = mmean.percent[mmean.percent$geno == 'silentcha', 'mean']/mmean.percent[mmean.percent$geno == 'silentcha' & mmean.percent$day == 2, 'mean']

par(mar=c(5.1, 4.1, 4.1, 15.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0, 1.25),
     xlim = c(2,27),
     ylab = 'Percent of Initial Mean Velocity',
     xlab = 'Day',
     xaxt = 'n',
     main = "Males")

axis(1, at = mmean.percent[mmean.percent$geno == 'silent', 'day'])

lines(x = mmean.percent[mmean.percent$geno == 'silent', 'day'],
      y = mmean.percent[mmean.percent$geno == 'silent', 'percent'],
      type = 'l',
      col = lcol[1],
      lwd = 2)

lines(x = mmean.percent[mmean.percent$geno == 'A4V', 'day'],
      y = mmean.percent[mmean.percent$geno == 'A4V', 'percent'],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = mmean.percent[mmean.percent$geno == 'silentcha', 'day'],
      y = mmean.percent[mmean.percent$geno == 'silentcha', 'percent'],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = mmean.percent[mmean.percent$geno == 'A4Vcha', 'day'],
      y = mmean.percent[mmean.percent$geno == 'A4Vcha', 'percent'],
      type = 'l',
      col = pcol[4],
      lwd = 2)

lines(x = mmean.percent[mmean.percent$geno == 'silentok', 'day'],
      y = mmean.percent[mmean.percent$geno == 'silentok', 'percent'],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = mmean.percent[mmean.percent$geno == 'A4Vok', 'day'],
      y = mmean.percent[mmean.percent$geno == 'A4Vok', 'percent'],
      type = 'l',
      col = pcol[6],
      lwd = 2)

points(x = mmean.percent[mmean.percent$geno == 'silent', 'day'],
       y = mmean.percent[mmean.percent$geno == 'silent', 'percent'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = mmean.percent[mmean.percent$geno == 'A4V', 'day'],
       y = mmean.percent[mmean.percent$geno == 'A4V', 'percent'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = mmean.percent[mmean.percent$geno == 'silentcha', 'day'],
       y = mmean.percent[mmean.percent$geno == 'silentcha', 'percent'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = mmean.percent[mmean.percent$geno == 'A4Vcha', 'day'],
       y = mmean.percent[mmean.percent$geno == 'A4Vcha', 'percent'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = mmean.percent[mmean.percent$geno == 'silentok', 'day'],
       y = mmean.percent[mmean.percent$geno == 'silentok', 'percent'],
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = mmean.percent[mmean.percent$geno == 'A4Vok', 'day'],
       y = mmean.percent[mmean.percent$geno == 'A4Vok', 'percent'],
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)

points(x = mmean.percent[mmean.percent$geno == 'silent', 'day'],
       y = mmean.percent[mmean.percent$geno == 'silent', 'percent'],
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = mmean.percent[mmean.percent$geno == 'A4V', 'day'],
       y = mmean.percent[mmean.percent$geno == 'A4V', 'percent'],
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = mmean.percent[mmean.percent$geno == 'silentcha', 'day'],
       y = mmean.percent[mmean.percent$geno == 'silentcha', 'percent'],
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = mmean.percent[mmean.percent$geno == 'A4Vcha', 'day'],
       y = mmean.percent[mmean.percent$geno == 'A4Vcha', 'percent'],
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)
points(x = mmean.percent[mmean.percent$geno == 'silentok', 'day'],
       y = mmean.percent[mmean.percent$geno == 'silentok', 'percent'],
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex =2)
points(x = mmean.percent[mmean.percent$geno == 'A4Vok', 'day'],
       y = mmean.percent[mmean.percent$geno == 'A4Vok', 'percent'],
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex =2)

legend(x = 29,
       y = 1,
       legend = c('Silent', 'A4V',
                  'ChAT>gbb;Silent', 'ChAT>gbb;A4V', 
                  'OK371>gbb;Silent', 'OK371>gbb;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(hex(pcol[1], .5), hex(pcol[2], .5), hex(pcol[3], .5),
                hex(pcol[4], .5), hex(pcol[5], .5), hex(pcol[6], .5)),
       bty = 'n',
       y.intersp = 1,
       pt.cex = 2,
       xpd=TRUE)

```

```{r echo = F, include = F}

norm.male = na.omit(male)
norm.male$normval = 0
i=1
while(i <= nrow(norm.male)){
  norm.male$normval[i] = norm.male$corrected.values[i]/
    mmean$mean[mmean$geno == 'silent' & mmean$day == norm.male$day[i]]
  i = i +1
}

norm.mmean = unique(norm.male[,c('geno', 'day', 'sample')])
norm.mstd = norm.mmean
norm.meb = norm.mmean
norm.mmean$mean = 0
norm.mstd$std = 0

i=1
while(i<=nrow(norm.mmean)){
  norm.mmean$mean[i] = mean(na.omit(norm.male[norm.male$sample == norm.mmean$sample[i], 'normval']))

  norm.mstd$std[i] = (var(na.omit(norm.male[norm.male$sample == norm.mmean$sample[i], 'normval'])))

  i = i+1
}

norm.mmean = norm.mmean[order(norm.mmean$day),]
norm.mstd = norm.mstd[order(norm.mstd$day),]
norm.meb = norm.meb[order(norm.meb$day),]

norm.meb$y0 = norm.mmean$mean - norm.mstd$std
norm.meb$y1 = norm.mmean$mean + norm.mstd$std

norm.mlm = data.frame(geno = unique(norm.male$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(norm.mlm)){
  temp=na.omit(norm.male[norm.male$geno == norm.mlm$geno[i], ])
  temp.lm = lm(temp$normval~temp$day)
  norm.mlm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  norm.mlm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(norm.mlm) = norm.mlm$geno

```

```{r echo = F, include = F}

male$sample = paste0(male$geno, male$day)

norm.male = na.omit(male)
norm.male$normval = 0

mmean = unique(male[,c('geno', 'day', 'sample')])
mstd = mmean
eb = mmean
mmean$mean = 0
mstd$std = 0

i=1
while(i<=nrow(mmean)){
  mmean$mean[i] = mean(na.omit(male[male$sample == mmean$sample[i], 'corrected.values']))

  mstd$std[i] = (var(na.omit(male[male$sample == mmean$sample[i], 'corrected.values'])))

  i = i+1
}

mlm = data.frame(geno = unique(male$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(mlm)){
  temp=na.omit(male[male$geno == mlm$geno[i], ])
  temp.lm = lm(temp$corrected.values~temp$day)
  mlm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  mlm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(mlm) = mlm$geno

mmean = mmean[order(mmean$day),]
mstd = mstd[order(mstd$day),]
eb = eb[order(eb$day),]

eb$y0 = mmean$mean - mstd$std
eb$y1 = mmean$mean + mstd$std

i=1
while(i <= nrow(norm.male)){
  norm.male$normval[i] = norm.male$corrected.values[i]/
    mmean$mean[mmean$geno == 'silent' & mmean$day == norm.male$day[i]]
  i = i +1
}

norm.mmean = unique(norm.male[,c('geno', 'day', 'sample')])
norm.mstd = norm.mmean
norm.eb = norm.mmean
norm.mmean$mean = 0
norm.mstd$std = 0

i=1
while(i<=nrow(norm.mmean)){
  norm.mmean$mean[i] = mean(na.omit(norm.male[norm.male$sample == norm.mmean$sample[i], 'normval']))

  norm.mstd$std[i] = (var(na.omit(norm.male[norm.male$sample == norm.mmean$sample[i], 'normval'])))

  i = i+1
}

norm.mmean = norm.mmean[order(norm.mmean$day),]
norm.mstd = norm.mstd[order(norm.mstd$day),]
norm.eb = norm.eb[order(norm.eb$day),]

norm.eb$y0 = norm.mmean$mean - norm.mstd$std
norm.eb$y1 = norm.mmean$mean + norm.mstd$std

norm.mlm = data.frame(geno = unique(norm.male$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(norm.mlm)){
  temp=na.omit(norm.male[norm.male$geno == norm.mlm$geno[i], ])
  temp.lm = lm(temp$normval~temp$day)
  norm.mlm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  norm.mlm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(norm.mlm) = norm.mlm$geno

```

```{r echo = F, fig.height=10, include = F}

silent = norm.mmean[norm.mmean$geno == "silent", ]
a4v = norm.mmean[norm.mmean$geno == "A4V", ]
silentcha = norm.mmean[norm.mmean$geno == "silentcha", ]
a4vcha = norm.mmean[norm.mmean$geno == "A4Vcha", ]
silentok = norm.mmean[norm.mmean$geno == "silentok", ]
a4vok = norm.mmean[norm.mmean$geno == "A4Vok", ]

ls = norm.mmean[norm.mmean$geno == 'silent',]
la = norm.mmean[norm.mmean$geno == 'A4V',]
lsc = norm.mmean[norm.mmean$geno == 'silentcha',]
lac = norm.mmean[norm.mmean$geno == 'A4Vcha',]
lso = norm.mmean[norm.mmean$geno == 'silentok',]
lao = norm.mmean[norm.mmean$geno == 'A4Vok',]

pcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 9.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0,1.85),
     xlim = c(2,27),
     ylab = 'Relative Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     main = "Males")

axis(1, at=silent$day)

abline(v = 2, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 6, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 9, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 13, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 19, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 22, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 27, lty =2, lwd = 1, col = hex(pcol[7], .5))

lines(x = norm.mlm['silent', c("x0", "x1")],
      y = norm.mlm['silent', c("y0", "y1")],
      type = 'l',
      col = pcol[1],
      lwd = 2)

lines(x = norm.mlm['A4V', c("x0", "x1")],
      y = norm.mlm['A4V', c("y0", "y1")],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = norm.mlm['silentcha', c("x0", "x1")],
      y = norm.mlm['silentcha', c("y0", "y1")],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = norm.mlm['A4Vcha', c("x0", "x1")],
      y = norm.mlm['A4Vcha', c("y0", "y1")],
      type = 'l',
      col = pcol[4],
      lwd = 2)

lines(x = norm.mlm['silentok', c("x0", "x1")],
      y = norm.mlm['silentok', c("y0", "y1")],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = norm.mlm['A4Vok', c("x0", "x1")],
      y = norm.mlm['A4Vok', c("y0", "y1")],
      type = 'l',
      col = pcol[6],
      lwd = 2)

arrows(x0 = norm.eb$day[norm.eb$geno == 'silent'],
       x1 = norm.eb$day[norm.eb$geno == 'silent'],
       y0 = norm.eb$y0[norm.eb$geno == 'silent'],
       y1 = norm.eb$y1[norm.eb$geno == 'silent'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[1], 
       lwd=1)

arrows(x0 = norm.eb$day[norm.eb$geno == 'A4V'],
       x1 = norm.eb$day[norm.eb$geno == 'A4V'],
       y0 = norm.eb$y0[norm.eb$geno == 'A4V'],
       y1 = norm.eb$y1[norm.eb$geno == 'A4V'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[2], 
       lwd=1)

arrows(x0 = norm.eb$day[norm.eb$geno == 'silentcha'],
       x1 = norm.eb$day[norm.eb$geno == 'silentcha'],
       y0 = norm.eb$y0[norm.eb$geno == 'silentcha'],
       y1 = norm.eb$y1[norm.eb$geno == 'silentcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[3], 
       lwd=1)

arrows(x0 = norm.eb$day[norm.eb$geno == 'A4Vcha'],
       x1 = norm.eb$day[norm.eb$geno == 'A4Vcha'],
       y0 = norm.eb$y0[norm.eb$geno == 'A4Vcha'],
       y1 = norm.eb$y1[norm.eb$geno == 'A4Vcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[4], 
       lwd=1)

arrows(x0 = norm.eb$day[norm.eb$geno == 'silentok'],
       x1 = norm.eb$day[norm.eb$geno == 'silentok'],
       y0 = norm.eb$y0[norm.eb$geno == 'silentok'],
       y1 = norm.eb$y1[norm.eb$geno == 'silentok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[5], 
       lwd=1)

arrows(x0 = norm.eb$day[norm.eb$geno == 'A4Vok'],
       x1 = norm.eb$day[norm.eb$geno == 'A4Vok'],
       y0 = norm.eb$y0[norm.eb$geno == 'A4Vok'],
       y1 = norm.eb$y1[norm.eb$geno == 'A4Vok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[6], 
       lwd=1)

points(x = silent$day,
       y = silent$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4v$day,
       y = a4v$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)

points(x = silent$day,
       y = silent$mean,
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = a4v$day,
       y = a4v$mean,
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex =2)

legend(x = 29,
       y = 1.8,
       legend = c('Silent', 'A4V',
                  'ChAT>gbb;Silent', 'ChAT>gbb;A4V', 
                  'OK371>gbb;Silent', 'OK371>gbb;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(pcol[1], pcol[2], pcol[3],
                pcol[4], pcol[5], pcol[6]),
       bty = 'n',
       y.intersp = 1,
       pt.cex = 2,
       xpd=TRUE)

```

```{r  echo = F, include = F}

fig4cd = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 4CD.csv", header = T)

fig4cd = fig4cd[fig4cd$day != 9,]
fig4cd = fig4cd[fig4cd$day != 32,]

female = fig4cd[grep('f', fig4cd$vial_ID),]
male = fig4cd[grep('m', fig4cd$vial_ID),]

female$sample = paste0(female$geno, female$day)
male$sample = paste0(male$geno, male$day)

norm.female = na.omit(female)
norm.female$normval = 0

fmean = unique(female[,c('geno', 'day', 'sample')])
fstd = fmean
feb = fmean
fmean$mean = 0
fstd$std = 0

norm.male = na.omit(male)
norm.male$normval = 0

mmean = unique(male[,c('geno', 'day', 'sample')])
mstd = mmean
meb = mmean
mmean$mean = 0
mstd$std = 0

i=1
while(i<=nrow(mmean)){
  mmean$mean[i] = mean(na.omit(male[male$sample == mmean$sample[i], 'corrected.values']))

  mstd$std[i] = (var(na.omit(male[male$sample == mmean$sample[i], 'corrected.values'])))

  i = i+1
}

i=1
while(i<=nrow(fmean)){
  fmean$mean[i] = mean(na.omit(female[female$sample == fmean$sample[i], 'corrected.values']))

  fstd$std[i] = (var(na.omit(female[female$sample == fmean$sample[i], 'corrected.values'])))

  i = i+1
}

mlm = data.frame(geno = unique(male$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(mlm)){
  temp=na.omit(male[male$geno == mlm$geno[i], ])
  temp.lm = lm(temp$corrected.values~temp$day)
  mlm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  mlm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(mlm) = mlm$geno

flm = data.frame(geno = unique(female$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(flm)){
  temp=na.omit(female[female$geno == flm$geno[i], ])
  temp.lm = lm(temp$corrected.values~temp$day)
  flm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  flm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(flm) = flm$geno

mmean = mmean[order(mmean$day),]
mstd = mstd[order(mstd$day),]
meb = meb[order(meb$day),]

meb$y0 = mmean$mean - mstd$std
meb$y1 = mmean$mean + mstd$std

fmean = fmean[order(fmean$day),]
fstd = fstd[order(fstd$day),]
feb = feb[order(feb$day),]

feb$y0 = fmean$mean - fstd$std
feb$y1 = fmean$mean + fstd$std

i=1
while(i <= nrow(norm.male)){
  norm.male$normval[i] = norm.male$corrected.values[i]/
    mmean$mean[mmean$geno == 'silent' & mmean$day == norm.male$day[i]]
  i = i +1
}

norm.mmean = unique(norm.male[,c('geno', 'day', 'sample')])
norm.mstd = norm.mmean
norm.meb = norm.mmean
norm.mmean$mean = 0
norm.mstd$std = 0

i=1
while(i <= nrow(norm.female)){
  norm.female$normval[i] = norm.female$corrected.values[i]/
    fmean$mean[fmean$geno == 'silent' & fmean$day == norm.female$day[i]]
  i = i +1
}

norm.fmean = unique(norm.female[,c('geno', 'day', 'sample')])
norm.fstd = norm.fmean
norm.feb = norm.fmean
norm.fmean$mean = 0
norm.fstd$std = 0

i=1
while(i<=nrow(norm.mmean)){
  norm.mmean$mean[i] = mean(na.omit(norm.male[norm.male$sample == norm.mmean$sample[i], 'normval']))

  norm.mstd$std[i] = (var(na.omit(norm.male[norm.male$sample == norm.mmean$sample[i], 'normval'])))

  i = i+1
}

i=1
while(i<=nrow(norm.fmean)){
  norm.fmean$mean[i] = mean(na.omit(norm.female[norm.female$sample == norm.fmean$sample[i], 'normval']))

  norm.fstd$std[i] = (var(na.omit(norm.female[norm.female$sample == norm.fmean$sample[i], 'normval'])))

  i = i+1
}

norm.mmean = norm.mmean[order(norm.mmean$day),]
norm.mstd = norm.mstd[order(norm.mstd$day),]
norm.meb = norm.meb[order(norm.meb$day),]

norm.meb$y0 = norm.mmean$mean - norm.mstd$std
norm.meb$y1 = norm.mmean$mean + norm.mstd$std

norm.mlm = data.frame(geno = unique(norm.male$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(norm.mlm)){
  temp=na.omit(norm.male[norm.male$geno == norm.mlm$geno[i], ])
  temp.lm = lm(temp$normval~temp$day)
  norm.mlm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  norm.mlm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(norm.mlm) = norm.mlm$geno

norm.fmean = norm.fmean[order(norm.fmean$day),]
norm.fstd = norm.fstd[order(norm.fstd$day),]
norm.feb = norm.feb[order(norm.feb$day),]

norm.feb$y0 = norm.fmean$mean - norm.fstd$std
norm.feb$y1 = norm.fmean$mean + norm.fstd$std

norm.flm = data.frame(geno = unique(norm.female$geno),
                x0 = 2,
                x1 = 27,
                y0 = 0,
                y1 = 0)

i=1
while(i<=nrow(norm.flm)){
  temp=na.omit(norm.female[norm.female$geno == norm.flm$geno[i], ])
  temp.lm = lm(temp$normval~temp$day)
  norm.flm$y0[i] = temp.lm[[1]][2]*2 + temp.lm[[1]][1]
  norm.flm$y1[i] = temp.lm[[1]][2]*27 + temp.lm[[1]][1]
  i = i +1
}

row.names(norm.flm) = norm.flm$geno

```

```{r echo = F, fig.height=10, include = F}

silent = norm.fmean[norm.fmean$geno == "silent", ]
a4v = norm.fmean[norm.fmean$geno == "A4V", ]
silentcha = norm.fmean[norm.fmean$geno == "silentcha", ]
a4vcha = norm.fmean[norm.fmean$geno == "A4Vcha", ]
silentok = norm.fmean[norm.fmean$geno == "silentok", ]
a4vok = norm.fmean[norm.fmean$geno == "A4Vok", ]

ls = norm.fmean[norm.fmean$geno == 'silent',]
la = norm.fmean[norm.fmean$geno == 'A4V',]
lsc = norm.fmean[norm.fmean$geno == 'silentcha',]
lac = norm.fmean[norm.fmean$geno == 'A4Vcha',]
lso = norm.fmean[norm.fmean$geno == 'silentok',]
lao = norm.fmean[norm.fmean$geno == 'A4Vok',]

pcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 9.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0,1.85),
     xlim = c(2,27),
     ylab = 'Relative Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     main = "Females")

axis(1, at=silent$day)

abline(v = 2, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 6, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 9, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 13, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 19, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 22, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 27, lty =2, lwd = 1, col = hex(pcol[7], .5))

lines(x = norm.flm['silent', c("x0", "x1")],
      y = norm.flm['silent', c("y0", "y1")],
      type = 'l',
      col = pcol[1],
      lwd = 2)

lines(x = norm.flm['A4V', c("x0", "x1")],
      y = norm.flm['A4V', c("y0", "y1")],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = norm.flm['silentcha', c("x0", "x1")],
      y = norm.flm['silentcha', c("y0", "y1")],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = norm.flm['A4Vcha', c("x0", "x1")],
      y = norm.flm['A4Vcha', c("y0", "y1")],
      type = 'l',
      col = pcol[4],
      lwd = 2)

lines(x = norm.flm['silentok', c("x0", "x1")],
      y = norm.flm['silentok', c("y0", "y1")],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = norm.flm['A4Vok', c("x0", "x1")],
      y = norm.flm['A4Vok', c("y0", "y1")],
      type = 'l',
      col = pcol[6],
      lwd = 2)

arrows(x0 = norm.feb$day[norm.feb$geno == 'silent'],
       x1 = norm.feb$day[norm.feb$geno == 'silent'],
       y0 = norm.feb$y0[norm.feb$geno == 'silent'],
       y1 = norm.feb$y1[norm.feb$geno == 'silent'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[1], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'A4V'],
       x1 = norm.feb$day[norm.feb$geno == 'A4V'],
       y0 = norm.feb$y0[norm.feb$geno == 'A4V'],
       y1 = norm.feb$y1[norm.feb$geno == 'A4V'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[2], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'silentcha'],
       x1 = norm.feb$day[norm.feb$geno == 'silentcha'],
       y0 = norm.feb$y0[norm.feb$geno == 'silentcha'],
       y1 = norm.feb$y1[norm.feb$geno == 'silentcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[3], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'A4Vcha'],
       x1 = norm.feb$day[norm.feb$geno == 'A4Vcha'],
       y0 = norm.feb$y0[norm.feb$geno == 'A4Vcha'],
       y1 = norm.feb$y1[norm.feb$geno == 'A4Vcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[4], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'silentok'],
       x1 = norm.feb$day[norm.feb$geno == 'silentok'],
       y0 = norm.feb$y0[norm.feb$geno == 'silentok'],
       y1 = norm.feb$y1[norm.feb$geno == 'silentok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[5], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'A4Vok'],
       x1 = norm.feb$day[norm.feb$geno == 'A4Vok'],
       y0 = norm.feb$y0[norm.feb$geno == 'A4Vok'],
       y1 = norm.feb$y1[norm.feb$geno == 'A4Vok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[6], 
       lwd=1)

points(x = silent$day,
       y = silent$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4v$day,
       y = a4v$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)

points(x = silent$day,
       y = silent$mean,
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = a4v$day,
       y = a4v$mean,
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex =2)

legend(x = 29,
       y = 1.8,
       legend = c('Silent', 'A4V',
                  'ChAT>gbb;Silent', 'ChAT>gbb;A4V', 
                  'OK371>gbb;Silent', 'OK371>gbb;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(pcol[1], pcol[2], pcol[3],
                pcol[4], pcol[5], pcol[6]),
       bty = 'n',
       y.intersp = 1,
       pt.cex = 2,
       xpd=TRUE)

```

```{r echo = F, fig.height=10, include = F}

silent = norm.fmean[norm.fmean$geno == "silent", ]
a4v = norm.fmean[norm.fmean$geno == "A4V", ]
silentcha = norm.fmean[norm.fmean$geno == "silentcha", ]
a4vcha = norm.fmean[norm.fmean$geno == "A4Vcha", ]
silentok = norm.fmean[norm.fmean$geno == "silentok", ]
a4vok = norm.fmean[norm.fmean$geno == "A4Vok", ]

ls = norm.fmean[norm.fmean$geno == 'silent',]
la = norm.fmean[norm.fmean$geno == 'A4V',]
lsc = norm.fmean[norm.fmean$geno == 'silentcha',]
lac = norm.fmean[norm.fmean$geno == 'A4Vcha',]
lso = norm.fmean[norm.fmean$geno == 'silentok',]
lao = norm.fmean[norm.fmean$geno == 'A4Vok',]

pcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 9.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0,2.5),
     xlim = c(2,27),
     ylab = 'Relative Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     main = "Females")

axis(1, at=silent$day)

abline(v = 2, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 6, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 9, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 13, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 19, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 22, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 27, lty =2, lwd = 1, col = hex(pcol[7], .5))

lines(x = norm.flm['silent', c("x0", "x1")],
      y = norm.flm['silent', c("y0", "y1")],
      type = 'l',
      col = pcol[1],
      lwd = 2)

lines(x = norm.flm['A4V', c("x0", "x1")],
      y = norm.flm['A4V', c("y0", "y1")],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = norm.flm['silentcha', c("x0", "x1")],
      y = norm.flm['silentcha', c("y0", "y1")],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = norm.flm['A4Vcha', c("x0", "x1")],
      y = norm.flm['A4Vcha', c("y0", "y1")],
      type = 'l',
      col = pcol[4],
      lwd = 2)

lines(x = norm.flm['silentok', c("x0", "x1")],
      y = norm.flm['silentok', c("y0", "y1")],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = norm.flm['A4Vok', c("x0", "x1")],
      y = norm.flm['A4Vok', c("y0", "y1")],
      type = 'l',
      col = pcol[6],
      lwd = 2)

arrows(x0 = norm.feb$day[norm.feb$geno == 'silent'],
       x1 = norm.feb$day[norm.feb$geno == 'silent'],
       y0 = norm.feb$y0[norm.feb$geno == 'silent'],
       y1 = norm.feb$y1[norm.feb$geno == 'silent'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[1], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'A4V'],
       x1 = norm.feb$day[norm.feb$geno == 'A4V'],
       y0 = norm.feb$y0[norm.feb$geno == 'A4V'],
       y1 = norm.feb$y1[norm.feb$geno == 'A4V'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[2], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'silentcha'],
       x1 = norm.feb$day[norm.feb$geno == 'silentcha'],
       y0 = norm.feb$y0[norm.feb$geno == 'silentcha'],
       y1 = norm.feb$y1[norm.feb$geno == 'silentcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[3], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'A4Vcha'],
       x1 = norm.feb$day[norm.feb$geno == 'A4Vcha'],
       y0 = norm.feb$y0[norm.feb$geno == 'A4Vcha'],
       y1 = norm.feb$y1[norm.feb$geno == 'A4Vcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[4], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'silentok'],
       x1 = norm.feb$day[norm.feb$geno == 'silentok'],
       y0 = norm.feb$y0[norm.feb$geno == 'silentok'],
       y1 = norm.feb$y1[norm.feb$geno == 'silentok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[5], 
       lwd=1)

arrows(x0 = norm.feb$day[norm.feb$geno == 'A4Vok'],
       x1 = norm.feb$day[norm.feb$geno == 'A4Vok'],
       y0 = norm.feb$y0[norm.feb$geno == 'A4Vok'],
       y1 = norm.feb$y1[norm.feb$geno == 'A4Vok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[6], 
       lwd=1)

points(x = silent$day,
       y = silent$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4v$day,
       y = a4v$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)

points(x = silent$day,
       y = silent$mean,
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = a4v$day,
       y = a4v$mean,
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex =2)

legend(x = 29,
       y = 2.5,
       legend = c('Silent', 'A4V',
                  'ChAT>SaxA;Silent', 'ChAT>SaxA;A4V', 
                  'OK371>SaxA;Silent', 'OK371>SaxA;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(pcol[1], pcol[2], pcol[3],
                pcol[4], pcol[5], pcol[6]),
       bty = 'n',
       y.intersp = 1,
       pt.cex = 2,
       xpd=TRUE)

```

```{r echo = F, fig.height=10, include = F}

silent = norm.mmean[norm.mmean$geno == "silent", ]
a4v = norm.mmean[norm.mmean$geno == "A4V", ]
silentcha = norm.mmean[norm.mmean$geno == "silentcha", ]
a4vcha = norm.mmean[norm.mmean$geno == "A4Vcha", ]
silentok = norm.mmean[norm.mmean$geno == "silentok", ]
a4vok = norm.mmean[norm.mmean$geno == "A4Vok", ]

ls = norm.mmean[norm.mmean$geno == 'silent',]
la = norm.mmean[norm.mmean$geno == 'A4V',]
lsc = norm.mmean[norm.mmean$geno == 'silentcha',]
lac = norm.mmean[norm.mmean$geno == 'A4Vcha',]
lso = norm.mmean[norm.mmean$geno == 'silentok',]
lao = norm.mmean[norm.mmean$geno == 'A4Vok',]

pcol = c('steelblue', 'navy', 'limegreen', 'darkgreen', 'brown1', 'firebrick', 'grey')

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

par(mar=c(5.1, 4.1, 4.1, 9.1))

plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(0,2.5),
     xlim = c(2,27),
     ylab = 'Relative Velocity (cm/s)',
     xlab = 'Day',
     xaxt = 'n',
     main = "Males")

axis(1, at=silent$day)

abline(v = 2, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 6, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 9, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 13, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 19, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 22, lty =2, lwd = 1, col = hex(pcol[7], .5))
abline(v = 27, lty =2, lwd = 1, col = hex(pcol[7], .5))

lines(x = norm.mlm['silent', c("x0", "x1")],
      y = norm.mlm['silent', c("y0", "y1")],
      type = 'l',
      col = pcol[1],
      lwd = 2)

lines(x = norm.mlm['A4V', c("x0", "x1")],
      y = norm.mlm['A4V', c("y0", "y1")],
      type = 'l',
      col = pcol[2],
      lwd = 2)

lines(x = norm.mlm['silentcha', c("x0", "x1")],
      y = norm.mlm['silentcha', c("y0", "y1")],
      type = 'l',
      col = pcol[3],
      lwd = 2)

lines(x = norm.mlm['A4Vcha', c("x0", "x1")],
      y = norm.mlm['A4Vcha', c("y0", "y1")],
      type = 'l',
      col = pcol[4],
      lwd = 2)

lines(x = norm.mlm['silentok', c("x0", "x1")],
      y = norm.mlm['silentok', c("y0", "y1")],
      type = 'l',
      col = pcol[5],
      lwd = 2)

lines(x = norm.mlm['A4Vok', c("x0", "x1")],
      y = norm.mlm['A4Vok', c("y0", "y1")],
      type = 'l',
      col = pcol[6],
      lwd = 2)

arrows(x0 = norm.meb$day[norm.meb$geno == 'silent'],
       x1 = norm.meb$day[norm.meb$geno == 'silent'],
       y0 = norm.meb$y0[norm.meb$geno == 'silent'],
       y1 = norm.meb$y1[norm.meb$geno == 'silent'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[1], 
       lwd=1)

arrows(x0 = norm.meb$day[norm.meb$geno == 'A4V'],
       x1 = norm.meb$day[norm.meb$geno == 'A4V'],
       y0 = norm.meb$y0[norm.meb$geno == 'A4V'],
       y1 = norm.meb$y1[norm.meb$geno == 'A4V'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[2], 
       lwd=1)

arrows(x0 = norm.meb$day[norm.meb$geno == 'silentcha'],
       x1 = norm.meb$day[norm.meb$geno == 'silentcha'],
       y0 = norm.meb$y0[norm.meb$geno == 'silentcha'],
       y1 = norm.meb$y1[norm.meb$geno == 'silentcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[3], 
       lwd=1)

arrows(x0 = norm.meb$day[norm.meb$geno == 'A4Vcha'],
       x1 = norm.meb$day[norm.meb$geno == 'A4Vcha'],
       y0 = norm.meb$y0[norm.meb$geno == 'A4Vcha'],
       y1 = norm.meb$y1[norm.meb$geno == 'A4Vcha'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[4], 
       lwd=1)

arrows(x0 = norm.meb$day[norm.meb$geno == 'silentok'],
       x1 = norm.meb$day[norm.meb$geno == 'silentok'],
       y0 = norm.meb$y0[norm.meb$geno == 'silentok'],
       y1 = norm.meb$y1[norm.meb$geno == 'silentok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[5], 
       lwd=1)

arrows(x0 = norm.meb$day[norm.meb$geno == 'A4Vok'],
       x1 = norm.meb$day[norm.meb$geno == 'A4Vok'],
       y0 = norm.meb$y0[norm.meb$geno == 'A4Vok'],
       y1 = norm.meb$y1[norm.meb$geno == 'A4Vok'], 
       code = 3, 
       angle = 90, 
       length = 0.1, 
       col = pcol[6], 
       lwd=1)

points(x = silent$day,
       y = silent$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4v$day,
       y = a4v$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = 'white',
       col = 'white',
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = 'white',
       col = 'white',
       pch = 21,
       cex =2)

points(x = silent$day,
       y = silent$mean,
       bg = hex(pcol[1], .5),
       col = pcol[1],
       pch = 23,
       cex =2)

points(x = a4v$day,
       y = a4v$mean,
       bg = hex(pcol[2], .5),
       col = pcol[2],
       pch = 21,
       cex =2)
points(x = silentcha$day,
       y = silentcha$mean,
       bg = hex(pcol[3], .5),
       col = pcol[3],
       pch = 23,
       cex =2)
points(x = a4vcha$day,
       y = a4vcha$mean,
       bg = hex(pcol[4], .5),
       col = pcol[4],
       pch = 21,
       cex =2)
points(x = silentok$day,
       y = silentok$mean,
       bg = hex(pcol[5], .5),
       col = pcol[5],
       pch = 23,
       cex =2)
points(x = a4vok$day,
       y = a4vok$mean,
       bg = hex(pcol[6], .5),
       col = pcol[6],
       pch = 21,
       cex =2)

legend(x = 29,
       y = 2.5,
       legend = c('Silent', 'A4V',
                  'ChAT>SaxA;Silent', 'ChAT>SaxA;A4V', 
                  'OK371>SaxA;Silent', 'OK371>SaxA;A4V'),
       pch = c(23,21,23,21,23,21),
       pt.bg = c(pcol[1], pcol[2], pcol[3],
                pcol[4], pcol[5], pcol[6]),
       bty = 'n',
       y.intersp = 1,
       pt.cex = 2,
       xpd=TRUE)
```