### 2024-09-27 TKT_RNAseq/R Code/TKTRNAseqRescueResponse.Rmd

---
title: "G85R Analysis: Rescue Response"
author: "John Santiago"
date: "2024-09-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = F, echo = F}

library(plotly)
library(edgeR)
library(heatmaply)
library(org.Dm.eg.db)
library(KEGGREST)
library(RcisTarget)
library(DT)
##source('/Users/johncsantiago/Documents/GitHub/WhartonLab/GOandKEGGPlottingFunctions.R')
library(VennDiagram)

```

```{r include = F, echo = F}
##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"
cpmdata = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
groups  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
TKT.EdgeR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)
TKT.EdgeR.FC = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FCTable.csv"),row.names = 1)

G85R.meancpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT_meancpmdata.csv", row.names = 1)

groups = groups[colnames(cpmdata),]

GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)

##Clean Trait Data and calculate CPMs
countdata=read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT.counttable.csv",row.names=1)

##normalize data
countdata=countdata[, row.names(groups)]
x <- countdata
gr <- factor(groups$Group)
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
design<-model.matrix(~0 + gr)
colnames(design) = unique(gr)
z = estimateDisp(z, design, robust = T)
fit <- glmQLFit(z, design)

compare = makeContrasts(TktDfGR.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.FDfxWT.FC=DEout
sGR.FDfxWT.FC=DEout[DEout$FDR<.05,]

TKT.EdgeR = cbind(TKT.EdgeR, GR.FDfxWT.FC[row.names(TKT.EdgeR), "FDR"])
colnames(TKT.EdgeR)[17] = 'GR.FDfxWT.FC'

TKT.EdgeR.FC = cbind(TKT.EdgeR.FC, GR.FDfxWT.FC[row.names(TKT.EdgeR), "logFC"])
colnames(TKT.EdgeR.FC)[17] = 'GR.FDfxWT.FC'

compare = makeContrasts(TktDfGR.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.MDfxWT.MC=DEout
sGR.MDfxWT.MC=DEout[DEout$FDR<.05,]

TKT.EdgeR = cbind(TKT.EdgeR, GR.MDfxWT.MC[row.names(TKT.EdgeR), "FDR"])
colnames(TKT.EdgeR)[18] = 'GR.MDfxWT.MC'

TKT.EdgeR.FC = cbind(TKT.EdgeR.FC, GR.MDfxWT.MC[row.names(TKT.EdgeR), "logFC"])
colnames(TKT.EdgeR.FC)[18] = 'GR.MDfxWT.MC'

compare = makeContrasts(TktOEGR.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.FOExWT.FC=DEout
sGR.FOExWT.FC=DEout[DEout$FDR<.05,]

TKT.EdgeR = cbind(TKT.EdgeR, GR.FOExWT.FC[row.names(TKT.EdgeR), "FDR"])
colnames(TKT.EdgeR)[19] = 'GR.FOExWT.FC'

TKT.EdgeR.FC = cbind(TKT.EdgeR.FC, GR.FOExWT.FC[row.names(TKT.EdgeR), "logFC"])
colnames(TKT.EdgeR.FC)[19] = 'GR.FOExWT.FC'

compare = makeContrasts(TktOEGR.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.MOExWT.MC=DEout
sGR.MOExWT.MC=DEout[DEout$FDR<.05,]

TKT.EdgeR = cbind(TKT.EdgeR, GR.MOExWT.MC[row.names(TKT.EdgeR), "FDR"])
colnames(TKT.EdgeR)[20] = 'GR.MOExWT.MC'

TKT.EdgeR.FC = cbind(TKT.EdgeR.FC, GR.MOExWT.MC[row.names(TKT.EdgeR), "logFC"])
colnames(TKT.EdgeR.FC)[20] = 'GR.MOExWT.MC'

rawdata = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/MetabolomicsForMetaboanalyst.txt", sep = "\t", row.names = 1)
KEGG.Key = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/RawMetabolomicData.csv", row.names = 1)
KEGG.Key = setNames(KEGG.Key[-c(1:2), 2], row.names(KEGG.Key)[-c(1:2)])

genotypes = setNames(c("WT", "GR", "TktOEWT", "TktOEGR", "TktDfWT", "TktDfGR"), c("C", "E", "A", "F", "B", "D"))

metadata = data.frame(Sample = colnames(rawdata), 
                      Group = as.character(rawdata[1,]), 
                      TIC = as.numeric(rawdata[2,]), 
                      Genotypes = genotypes[as.character(rawdata[1,])])

row.names(metadata) = metadata$Sample

norm.func = function(data){
  nd = (as.numeric(data)/metadata$TIC)*1000
  return(nd)
}

norm.data = t(apply(rawdata[3:nrow(rawdata),], 1, norm.func))
colnames(norm.data) = colnames(rawdata)

group.mean=function(data){
  gm = mean(na.omit(data))
}

mean.data = matrix(0, nrow=nrow(norm.data), ncol = length(genotypes))
row.names(mean.data) = row.names(norm.data)
colnames(mean.data) = unique(metadata$Group)

i=1
while(i<=ncol(mean.data)){
  mean.data[,i] = apply(norm.data[,metadata[metadata$Group == colnames(mean.data)[i], "Sample"]], 1, group.mean)
  i = i+1
}

mean.metab = mean.data

compare.conditions = function(condition1, condition2){
  columns1 = metadata[metadata$Group == condition1, "Sample"]
  columns2 = metadata[metadata$Group == condition2, "Sample"]
  p=setNames(rep(NA, nrow(norm.data)), row.names(norm.data))
  i=1
  while(i<=length(p)){
    if(length(na.omit(norm.data[i,columns1]))>2 & 
       length(na.omit(norm.data[i,columns2]))>2){
      temp = t.test(na.omit(norm.data[i,columns1]),
                    na.omit(norm.data[i,columns2]), paired = F, var.equal = T)
      p[i] = temp[[3]]
      }
    i=i+1
    }

  fdr = p.adjust(p, "BH", length(p))

  fc = mean.metab[,condition1]/mean.metab[,condition2]

  comparison.table = data.frame(p = p, 
                                FDR = fdr, 
                                FC = fc, 
                                KEGG = KEGG.Key[row.names(mean.metab)])

  comparison.table = comparison.table[order(comparison.table$FDR),]
}

GRxWT.C = compare.conditions("E", "C")
GRxWT.Df = compare.conditions("D", "B")[row.names(GRxWT.C),]
GRxWT.OE = compare.conditions("F", "A")[row.names(GRxWT.C),]

GR.CxDf = compare.conditions("D", "E")[row.names(GRxWT.C),]
WT.CxDf = compare.conditions("B", "C")[row.names(GRxWT.C),]

GR.CxOE = compare.conditions("F", "E")[row.names(GRxWT.C),]
WT.CxOE = compare.conditions("A", "C")[row.names(GRxWT.C),]

GR.DfxWT.C = compare.conditions("D", "C")[row.names(GRxWT.C),]
GR.OExWT.C = compare.conditions("F", "C")[row.names(GRxWT.C),]

comparison.means = list(GRF.CxDf = c("TktDfGR.F","GR.F"), 
                 GRF.CxOE = c("TktOEGR.F","GR.F"),
                 WTF.CxDf = c("TktDfWT.F","WT.F"),
                 WTF.CxOE = c("TktOEWT.F","WT.F"),
                 GRxWT.FC = c("GR.F","WT.F"), 
                 GRxWT.FDf = c("TktDfGR.F","TktDfWT.F"),
                 GRxWT.FOE = c("TktOEGR.F","TktOEWT.F"), 
                 GRM.CxDf = c("TktDfGR.M","GR.M"), 
                 GRM.CxOE = c("TktOEGR.M","GR.M"),
                 WTM.CxDf = c("TktDfWT.M","WT.M"),
                 WTM.CxOE = c("TktOEWT.M","WT.M"),
                 GRxWT.MC = c("GR.M","WT.M"), 
                 GRxWT.MDf = c("TktDfGR.M","TktDfWT.M"),
                 GRxWT.MOE = c("TktOEGR.M","TktOEWT.M"),
                 GR.FxM = c("GR.F","GR.M"),
                 WT.FxM = c("WT.F","WT.M"))

metab.FC = data.frame('GRxWT.C' = GRxWT.C$FC,
                      'GRxWT.Df' = GRxWT.Df$FC,
                      'GRxWT.OE' = GRxWT.OE$FC,
                      'GR.CxDf' = GR.CxDf$FC,
                      'WT.CxDf' = WT.CxDf$FC,
                      'GR.CxOE' = GR.CxOE$FC,
                      'WT.CxOE' = WT.CxOE$FC,
                      'GR.DfxWT.C' = GR.DfxWT.C$FC,
                      'GR.OExWT.C' = GR.OExWT.C$FC)
row.names(metab.FC) = row.names(GRxWT.C)
metab.FC = log2(metab.FC)

metab.FDR = data.frame(GRxWT.C = GRxWT.C$p,
                       GRxWT.Df = GRxWT.Df$p,
                       GRxWT.OE = GRxWT.OE$p,
                       GR.CxDf = GR.CxDf$p,
                       WT.CxDf = WT.CxDf$p,
                       GR.CxOE = GR.CxOE$p,
                       WT.CxOE = WT.CxOE$p,
                       GR.DfxWT.C = GR.DfxWT.C$p,
                       GR.OExWT.C = GR.OExWT.C$p)
row.names(metab.FDR) = row.names(GRxWT.C)

colnames(mean.metab) = as.character(genotypes[colnames(mean.metab)])

enzyme2KEGG = as.list(org.Dm.egENZYME2EG)

```

```{r echo = F}

TKT.biplot = function(x, y, mean.data, main.title){
min.expression = 10

FDR.data = TKT.EdgeR[, c(x, y)]
x.genes = row.names(FDR.data[FDR.data[,x] < 0.05,])
y.genes = row.names(FDR.data[FDR.data[,y] < 0.05,])
xy.genes = intersect(x.genes, y.genes)
DEGs = unique(c(x.genes, y.genes))

FC.data = TKT.EdgeR.FC[DEGs, c(x, y)]
FDR.data = TKT.EdgeR[DEGs, c(x, y)]
mean.data = mean.data[DEGs, ]

data = data.frame(Symbol = GeneIDKey[DEGs, "Symbol"],
                  FBgn = DEGs,
                  FDR1 = FDR.data[, x],
                  FDR2 = FDR.data[, y],
                  FC1 = FC.data[, x],
                  FC2 = FC.data[, y],
                  Color = 'grey',
                  size = 1,
                  Variable1.cpm = mean.data[,1],
                  Control1.cpm = mean.data[,2],
                  Variable2.cpm = mean.data[,3],
                  Control2.cpm = mean.data[,4])
  row.names(data) = data$FBgn

  data[x.genes, 'Color'] = 'royalblue'
  data[y.genes, 'Color'] = 'firebrick' 
  data[xy.genes, 'Color'] = 'gold'

  data$size = log2(apply(mean.data, 1, max))
  data$size[data$size<3] = 1
  data = na.omit(data)
  data = data[data$size > log2(min.expression),]
  data$size = 2.5*data$size

  fig = plot_ly(data = data,
                x = ~FC1,
                y = ~FC2,
                type = 'scatter',
                mode = 'markers',
                marker = list(color = ~Color,
                              colors = ~Color,
                              line = list(color = "black", width = 1.5),
                              size = ~size),
                hoverinfo = "text",
                hovertext = paste("Symbol:", data$Symbol,
                                  "\nFBgn:", data$FBgn,
                                  "\n ", colnames(FDR.data)[1], "FDR: ", round(data$FDR1, 2),
                                  "\n ", colnames(FDR.data)[2], "FDR: ", round(data$FDR2,2),
                                  "\n ", colnames(FC.data)[1], "FC: ", round(data$FC1, 2),
                                  "\n ", colnames(FC.data)[2], "FC: ", round(data$FC2,2),
                                  "\n ", colnames(mean.data)[1], "mean CPM: ", signif(data$Variable1,2),
                                  "\n ", colnames(mean.data)[2], "mean CPM: ", signif(data$Control1,2),
                                  "\n ", colnames(mean.data)[3], "mean CPM: ", signif(data$Variable2,2),
                                  "\n ", colnames(mean.data)[4], "mean CPM: ", signif(data$Control2,2)))

  fig = fig %>% layout(title = main.title,
                       shapes = list(
                         type = "line",
                         line = list(color = "black", dash = 'dash', width = 2),
                         x0 = min(data$FC1, data$FC2) - 1,
                         x1 = max(data$FC1, data$FC2) + 1,
                         y0 = min(data$FC1, data$FC2) - 1,
                         y1 = max(data$FC1, data$FC2) + 1),
                       xaxis = list(title = paste(colnames(FC.data)[1], "FC")),
                       yaxis = list(title = paste(colnames(FC.data)[2], "FC")),
                       showlegend = F)

  return(list(fig, data))
}
```

```{r include = F, echo = F}
Subset.Enrichment = function(bg, geneset){

  sigs = geneset

  sigsKEY = GeneIDKey[sigs,]
  sigs = na.omit(sigsKEY$Symbol)
  bgKEY = GeneIDKey[bg,]
  bg = na.omit(bgKEY$Symbol)

  genes = setNames(rep(0, length(bg)), bg)
  genes[sigs] = 1
  pwf=nullp(genes,"dm3","geneSymbol")
  GO.wall=goseq(pwf,"dm3","geneSymbol")
  GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
  row.names(GO.wall) = GO.wall$category
  KEGG=goseq(pwf,gene2cat=GenesInKegg)
  KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
  row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
  KEGG$Name=KEGG.Names[row.names(KEGG),1]

  enrichment.data = list(GO.wall, KEGG)
  names(enrichment.data) = c("GO", "KEGG")
  return(enrichment.data)
}

```

```{r echo = F}

compoundandgenes = function(CompoundData, x, y, mean.data, Fx, My, Fmean, Mmean){
KEGG.ID = unlist(strsplit(CompoundData$compoundID, " / "))
KEGG.ID = unlist(strsplit(KEGG.ID, "-"))

symbols = schema(F)$traces$scatter$attributes$marker$symbol$values
symbols = symbols[c(3, 15, 27, 63, 39, 51, 159, 207, 75, 87, 99)]
i = 1
enzymes = keggGet(KEGG.ID[i])[[1]]$ENZYME
if(length(enzymes > 0)){
  enzymes = unlist(enzyme2KEGG[intersect(enzymes, names(enzyme2KEGG))])
  if(length(enzymes > 0)){
    enzymes = GeneIDKey[GeneIDKey$Name %in% enzymes,]
    enzymes$Compound = CompoundData[grep(KEGG.ID[i], CompoundData$compoundID),'Metab']
  }
}

if(length(KEGG.ID)>1){
  i = 2
  while(i <= length(KEGG.ID)){
    temp.enzymes = keggGet(KEGG.ID[i])[[1]]$ENZYME
    if(length(temp.enzymes > 0)){
      temp.enzymes = unlist(enzyme2KEGG[intersect(temp.enzymes, names(enzyme2KEGG))])
      if(length(temp.enzymes > 0)){
        temp.enzymes = GeneIDKey[GeneIDKey$Name %in% temp.enzymes,]
        temp.enzymes$Compound = CompoundData[grep(KEGG.ID[i], CompoundData$compoundID),'Metab']
        enzymes = rbind(enzymes, temp.enzymes)
      }
    }
    i = i + 1
  }
}

enzymes$symbol = symbols[as.numeric(factor(enzymes$Symbol))]

colnames(CompoundData)[1:2] = c('Molecule', 'ID')
bg = setNames(c('firebrick',"gold","darkgreen","dodgerblue", 'deeppink', 'grey', 'darkorange')[as.numeric(factor(CompoundData$Molecule))], factor(CompoundData$Molecule))
CompoundData$shape = 'hexagon'
#CompoundData$Color = bg[CompoundData$Molecule]
CompoundData$Color = 'grey'
CompoundData$size = 35
CompoundData$group = 'Metabolite'

enzymes = data.frame(Molecule = enzymes$Symbol,
                     ID = enzymes$FBgn,
                     FDR1 = TKT.EdgeR[enzymes$FBgn, Fx],
                     FDR2 = TKT.EdgeR[enzymes$FBgn, My],
                     FC1 = TKT.EdgeR.FC[enzymes$FBgn, Fx],
                     FC2 = TKT.EdgeR.FC[enzymes$FBgn, My],
                     Color = 'gold',
                     size = 20,
                     Variable1.mean = Fmean[enzymes$FBgn,1],
                     Control1.mean  = Fmean[enzymes$FBgn,2],
                     Variable2.mean = Mmean[enzymes$FBgn,1],
                     Control2.mean  = Mmean[enzymes$FBgn,2],
                     shape = enzymes$symbol,
                     group = 'Enzyme')

GeneData = na.omit(enzymes)

metabs = data.frame(Molecule = CompoundData$Molecule,
                     ID = CompoundData$ID,
                     FDR1 = CompoundData$FDR,
                     FDR2 = CompoundData$FDR,
                     FC1 = CompoundData$FC,
                     FC2 = CompoundData$FC,
                     Color = CompoundData$Color,
                     size = CompoundData$size,
                     Variable1.mean = CompoundData$Variable1.mean,
                     Control1.mean  = CompoundData$Control1.mean,
                     Variable2.mean = CompoundData$Variable1.mean,
                     Control2.mean  = CompoundData$Variable1.mean,
                     shape = 'hexagon',
                     group = 'Metabolite')

CompoundData = na.omit(metabs)

data = rbind(CompoundData, GeneData)

  fig = plot_ly(type = 'scatter',
                mode = 'markers')
  fig = fig %>% add_trace(x = enzymes$FC1,
                          y = enzymes$FC2,
                          name = 'Genes',
                          marker = list(
                            color = enzymes$Color,
                            size = enzymes$size,
                            symbol = enzymes$shape,
                            line = list(
                              color = 'black',
                              width = 2
                            )
                          ),
                hoverinfo = "text",
                hovertext = paste("Gene Symbol:", enzymes$Molecule,
                                  "\nCompound ID:", enzymes$ID,
                                  "\n", Fx, "p: ", round(2^-enzymes$FDR1, 2),
                                  "\n", My, "p: ", round(2^-enzymes$FDR2,2),
                                  "\n", Fx, "FC: ", round(enzymes$FC1, 2),
                                  "\n", My, "FC: ", round(enzymes$FC2,2),
                                  "\n", Fx, "mean: ", signif(enzymes$Variable1,2),
                                  "\n ", My, "mean: ", signif(enzymes$Control1,2)))

      fig = fig %>% add_trace(x = CompoundData$FC1,
                          y = CompoundData$FC2,
                          name = 'Metabolites',
                          marker = list(
                            color = CompoundData$Color,
                            size = CompoundData$size,
                            symbol = CompoundData$shape,
                            line = list(
                              color = 'black',
                              width = 2
                            )
                          ),
                hoverinfo = "text",
                hovertext = paste("Gene Symbol:", CompoundData$Molecule,
                                  "\nCompound ID:", CompoundData$ID,
                                  "\nSex: Male",
                                  "\n", x, "p: ", round(2^-CompoundData$FDR1, 2),
                                  "\n", y, "p: ", round(2^-CompoundData$FDR2,2),
                                  "\n", x, "FC: ", round(CompoundData$FC1, 2),
                                  "\n", y, "FC: ", round(CompoundData$FC2,2),
                                  "\n", colnames(mean.data)[1], "mean: ", signif(CompoundData$Variable1,2),
                                  "\n ", colnames(mean.data)[2], "mean: ", signif(CompoundData$Control1,2),
                                  "\n ", colnames(mean.data)[3], "mean: ", signif(CompoundData$Variable2,2),
                                  "\n ", colnames(mean.data)[4], "mean: ", signif(CompoundData$Control2,2)))

        fig = fig %>% layout(title = data$Molecule[1],
                       shapes = list(
                         type = "line",
                         line = list(color = "black", dash = 'dash', width = 2),
                         x0 = min(data$FC1, data$FC2) - 1,
                         x1 = max(data$FC1, data$FC2) + 1,
                         y0 = min(data$FC1, data$FC2) - 1,
                         y1 = max(data$FC1, data$FC2) + 1),
                       xaxis = list(title = paste(Fx, "FC")),
                       yaxis = list(title = paste(My, "FC")),
                       showlegend = T)

   fig

}
```

## Response Genes in G85R relative to WT

```{r echo = F}

x = "GRxWT.FC"
y = "GRxWT.MC"

mean.data = G85R.meancpm[, c('WT.F', 'WT.M',
                             'GR.F', 'GR.M')]
main.title = "G85R vs WT: Female and Male DE Genes"

##RG = Response Genes
GRxWT.RG = TKT.biplot(x, y, mean.data, main.title)

GRxWT.RG[[1]]

```

```{r echo = F}

##Sample titles as strings. Only fill in up to your number of selected categories
set1 = "Female Increased"
set2 = "Female Decreased"
set3 = "Male Increased"
set4 = "Male Decreased"

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1 = GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR1'] <= 0.05 &
                   GRxWT.RG[[2]][,'FC1'] > 0, "FBgn"]
s2 = GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR1'] <= 0.05 &
                   GRxWT.RG[[2]][,'FC1'] < 0, "FBgn"]
s3 = GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR2'] <= 0.05 &
                   GRxWT.RG[[2]][,'FC2'] > 0, "FBgn"]
s4 = GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR2'] <= 0.05 &
                   GRxWT.RG[[2]][,'FC2'] < 0, "FBgn"]

NormalResponse = data.frame(Female = c(length(s2), length(s1)),
                            Male =  c(length(s4), length(s3)))
row.names(NormalResponse) = c("Decreased", "Increased")
NormalResponse = as.matrix(NormalResponse)

par(mar = c(2,4,2,7.5))
barplot(NormalResponse,
        beside = T,
        ylab = "Total DEGs",
        main = "G85R vs Silent",
        col = c('navy', 'firebrick'),
        angle = c(45, 45, 315, 315), density = 50,
        legend.text = c("Decreased", "Increased"),
        args.legend = list(x = 8,
                           y = 700,
                           bty = 'n'))
#legend(x = 6.25, 
#       y = 700, 
#       legend = c("Increased", "Decreased"),
#       pt.bg = c('grey35', 'lightgrey'),
#       xpd = T, 
#       bty = "o",
#       pch = 22,
#       pt.cex = 1.5)

```

```{r echo = F}

##Sample titles as strings. Only fill in up to your number of selected categories
set3 = "Female Increased"
set1 = "Female Decreased"
set4 = "Male Increased"
set2 = "Male Decreased"

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s3 = GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR1'] <= 0.05 &
                   GRxWT.RG[[2]][,'FC1'] > 0, "FBgn"]
s1 = GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR1'] <= 0.05 &
                   GRxWT.RG[[2]][,'FC1'] < 0, "FBgn"]
s4 = GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR2'] <= 0.05 &
                   GRxWT.RG[[2]][,'FC2'] > 0, "FBgn"]
s2 = GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR2'] <= 0.05 &
                   GRxWT.RG[[2]][,'FC2'] < 0, "FBgn"]

main = ""

grid.newpage()
tempvenn=draw.quad.venn(area1=length(s1),area2=length(s2),area3=length(s3),area4=length(s4), n12=length(intersect(s1,s2)),n13 = length(intersect(s1,s3)),n14 = length(intersect(s1,s4)),n23 = length(intersect(s2,s3)),n24 = length(intersect(s2,s4)),n34 = length(intersect(s3,s4)),n123 =  length(intersect(s1,intersect(s2,s3))),n124 =  length(intersect(s1,intersect(s2,s4))), n134 = length(intersect(s1,intersect(s3,s4))),n234 = length(intersect(s2,intersect(s3,s4))),n1234 = length(intersect(s1,intersect(s2,intersect(s3,s4)))),category =c(set1,set2,set3,set4),fill = c("navy", "darkgreen","gold","firebrick"),cex=2,cat.cex = 1, title = "main")

```

##Female and Male Shared DEGs
```{r echo = F, include = F}

s1 = GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR1'] <= 0.05 &
                   GRxWT.RG[[2]][,'FC1'] > 0, "FBgn"]
s2 = GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR1'] <= 0.05 &
                   GRxWT.RG[[2]][,'FC1'] < 0, "FBgn"]
s3 = GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR2'] <= 0.05 &
                   GRxWT.RG[[2]][,'FC2'] > 0, "FBgn"]
s4 = GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR2'] <= 0.05 &
                   GRxWT.RG[[2]][,'FC2'] < 0, "FBgn"]

MF.shared.DEGs = c(intersect(s1,s3), intersect(s2,s4))

bg = row.names(TKT.EdgeR)
GRWTMF.Enr = Subset.Enrichment(bg, MF.shared.DEGs)

```

```{r echo = F}

RTCC.GRWTMF = tree(GRWTMF.Enr, "CC")
treemapPlot(RTCC.GRWTMF)

```

```{r echo = F}

RTCC.GRWTMF = tree(GRWTMF.Enr, "BP")
treemapPlot(RTCC.GRWTMF)

```

```{r echo = F}

RTCC.GRWTMF = tree(GRWTMF.Enr, "MF")
treemapPlot(RTCC.GRWTMF)

```

```{r echo = F}

cat.data = GRWTMF.Enr
GO.ontology = "CC"
FC = setNames(GRxWT.RG[[2]][, "FC1"], row.names(GRxWT.RG[[2]]))
FDR = setNames(GRxWT.RG[[2]][, "FDR1"], row.names(GRxWT.RG[[2]]))
usecats = "sigs.only"
sigs = MF.shared.DEGs

reversal.GO20(cat.data, GO.ontology, FC, FDR, usecats, sigs)

```

```{r echo = F}

cat.data = GRWTMF.Enr
GO.ontology = "BP"
FC = setNames(GRxWT.RG[[2]][, "FC1"], row.names(GRxWT.RG[[2]]))
FDR = setNames(GRxWT.RG[[2]][, "FDR1"], row.names(GRxWT.RG[[2]]))
usecats = c(1:39)
sigs = MF.shared.DEGs

reversal.GO20(cat.data, GO.ontology, FC, FDR, usecats, sigs)

```

```{r echo = F}

cat.data = GRWTMF.Enr
GO.ontology = "MF"
FC = setNames(GRxWT.RG[[2]][, "FC1"], row.names(GRxWT.RG[[2]]))
FDR = setNames(GRxWT.RG[[2]][, "FDR1"], row.names(GRxWT.RG[[2]]))
usecats = c(1:30)
sigs = MF.shared.DEGs

reversal.GO20(cat.data, GO.ontology, FC, FDR, usecats, sigs)

```

##Female Only DEGs
```{r echo = F, include = F}

F.only.DEGs = c(setdiff(GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR1'] <= 0.05, "FBgn"],
                        GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR2'] <= 0.05, "FBgn"]))

bg = row.names(TKT.EdgeR)
GRWTFonly.Enr = Subset.Enrichment(bg, F.only.DEGs)

```

```{r echo = F}

RTCC.GRWTFonly = tree(GRWTFonly.Enr, "CC")
treemapPlot(RTCC.GRWTFonly)

```

```{r echo = F}

RTCC.GRWTFonly = tree(GRWTFonly.Enr, "BP")
treemapPlot(RTCC.GRWTFonly)

```

```{r echo = F}

RTCC.GRWTFonly = tree(GRWTFonly.Enr, "MF")
treemapPlot(RTCC.GRWTFonly)
```

```{r echo = F}

cat.data = GRWTFonly.Enr
GO.ontology = "CC"
FC = setNames(GRxWT.RG[[2]][, "FC1"], row.names(GRxWT.RG[[2]]))
FDR = setNames(GRxWT.RG[[2]][, "FDR1"], row.names(GRxWT.RG[[2]]))
usecats = "sigs.only"
sigs = F.only.DEGs

reversal.GO20(cat.data, GO.ontology, FC, FDR, usecats, sigs)

```

```{r echo = F}

cat.data = GRWTFonly.Enr
GO.ontology = "BP"
FC = setNames(GRxWT.RG[[2]][, "FC1"], row.names(GRxWT.RG[[2]]))
FDR = setNames(GRxWT.RG[[2]][, "FDR1"], row.names(GRxWT.RG[[2]]))
usecats = "sigs.only"
sigs = F.only.DEGs

reversal.GO20(cat.data, GO.ontology, FC, FDR, usecats, sigs)

```

```{r echo = F}

cat.data = GRWTFonly.Enr
GO.ontology = "MF"
FC = setNames(GRxWT.RG[[2]][, "FC1"], row.names(GRxWT.RG[[2]]))
FDR = setNames(GRxWT.RG[[2]][, "FDR1"], row.names(GRxWT.RG[[2]]))
usecats = "sigs.only"
sigs = F.only.DEGs

reversal.GO20(cat.data, GO.ontology, FC, FDR, usecats, sigs)

```

##Male Only DEGs
```{r echo = F, include = F}

M.only.DEGs = c(setdiff(GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR2'] <= 0.05, "FBgn"],
                        GRxWT.RG[[2]][GRxWT.RG[[2]][,'FDR1'] <= 0.05, "FBgn"]))

bg = row.names(TKT.EdgeR)
GRWTMonly.Enr = Subset.Enrichment(bg, M.only.DEGs)
```

```{r echo = F}

RTCC.GRWTMonly = tree(GRWTMonly.Enr, "CC")
treemapPlot(RTCC.GRWTMonly)

```

```{r echo = F}

RTCC.GRWTMonly = tree(GRWTMonly.Enr, "BP")
treemapPlot(RTCC.GRWTMonly)

```

```{r echo = F}

RTCC.GRWTMonly = tree(GRWTMonly.Enr, "MF")
treemapPlot(RTCC.GRWTMonly)

```

```{r echo = F}

cat.data = GRWTMonly.Enr
GO.ontology = "CC"
FC = setNames(GRxWT.RG[[2]][, "FC2"], row.names(GRxWT.RG[[2]]))
FDR = setNames(GRxWT.RG[[2]][, "FDR2"], row.names(GRxWT.RG[[2]]))
usecats = c(1:28)
sigs = M.only.DEGs

reversal.GO20(cat.data, GO.ontology, FC, FDR, usecats, sigs)

```

```{r echo = F}

cat.data = GRWTMonly.Enr
GO.ontology = "BP"
FC = setNames(GRxWT.RG[[2]][, "FC2"], row.names(GRxWT.RG[[2]]))
FDR = setNames(GRxWT.RG[[2]][, "FDR2"], row.names(GRxWT.RG[[2]]))
usecats = c(1:50)
sigs = M.only.DEGs

reversal.GO20(cat.data, GO.ontology, FC, FDR, usecats, sigs)

```

```{r echo = F}

cat.data = GRWTMonly.Enr
GO.ontology = "MF"
FC = setNames(GRxWT.RG[[2]][, "FC2"], row.names(GRxWT.RG[[2]]))
FDR = setNames(GRxWT.RG[[2]][, "FDR2"], row.names(GRxWT.RG[[2]]))
usecats = c(1:50)
sigs = M.only.DEGs

reversal.GO20(cat.data, GO.ontology, FC, FDR, usecats, sigs)

```

```{r echo = F}

##Sample titles as strings. Only fill in up to your number of selected categories
set1 = "Female"
set2 = "Male"

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1 = GRWTFonly.Enr$GO[GRWTFonly.Enr$GO$adjp <= 0.05, "term"]
s2 = GRWTMonly.Enr$GO[GRWTMonly.Enr$GO$adjp <= 0.05, "term"]

main = ""

grid.newpage()
tempvenn=draw.pairwise.venn(area1 = length(s1),
                            area2 = length(s2),
                            cross.area = length(intersect(s1,s2)),
                            fill = c("blue", "red"),
                            cex = 2,
                            cat.cex = 1,
                            category = c(set1, set2),
                            title = "main")

```

```{r echo = F, include = F}

data("motifAnnotations_dmel_v9")
motifRankings = importRankings("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/DataFiles/dm6_v10_clust.genes_vs_motifs.rankings.feather")
motifAnnotations = motifAnnotations_dmel_v9

##motif rankings table for every TF
motif.data = getRanking(motifRankings)
##motifs associated with a TF
##TF.data = motifAnnotations[motifAnnotations$TF == 'Tudor-SN',]
###TF.data = motifAnnotations[motifAnnotations$TF == 'Mad',]

##TF.motif.data = motif.data[motif.data$motifs %in% TF.data$motif,]

##Motif collection data directory
motif.collection.dir = "/Users/johncsantiago/Documents/GitHub/WhartonLab/DrosophilaMotifsSingletonData/"

##geneLists = GeneIDKey[MF.shared.DEGs, 'Symbol']
geneLists = GeneIDKey[F.only.DEGs, 'Symbol']
geneLists = GeneIDKey[M.only.DEGs, 'Symbol']

motifEnrichmentTable_wGenes <- cisTarget(geneLists, 
         motifRankings,
         motifAnnot=motifAnnotations)

motifs_AUC <- calcAUC(geneLists, motifRankings, nCores=1)

auc <- getAUC(motifs_AUC)
hist(auc, xlab="dAUC histogram",
     breaks=100, col="#ff000050", border="darkred")
nes3 <- (3*sd(auc)) + mean(auc)
abline(v=nes3, col="red")

motifEnrichmentTable <- addMotifAnnotation(motifs_AUC, nesThreshold=3,
     motifAnnot=motifAnnotations)

motifEnrichmentTable_wGenes <- addSignificantGenes(motifEnrichmentTable,
                      rankings=motifRankings, 
                      geneSets=geneLists)
dim(motifEnrichmentTable_wGenes)

resultsSubset <- motifEnrichmentTable_wGenes

```

```{r echo = F}

showLogo(resultsSubset)

```

```{r echo = F, include = F}

anotatedTfs <- lapply(split(motifEnrichmentTable_wGenes$TF_highConf,
                            motifEnrichmentTable$geneSet),
                      function(x) {
                        genes <- gsub(" \\(.*\\). ", "; ", x, fixed=FALSE)
                        genesSplit <- unique(unlist(strsplit(genes, "; ")))
                        return(genesSplit)
                        })

anotatedTfs$hypoxia

```

```{r echo = F}

##comparisons
primary = 'GRxWT.C'

x = 'GRxWT.C'

mean.data = mean.metab[, c('GR', 'WT')]

min.expression = 0

##Reversal pattern
FC.data = metab.FC[, c(primary)]
FC.data = na.omit(FC.data)

##Sig cutoff. Currently, at least 1 significant change
sig.metab = metab.FDR[metab.FDR[,primary]  < .05, ]
sig.metab = na.omit(sig.metab)
sig.metab = row.names(sig.metab)

FC.data = metab.FC[sig.metab, c(x)]
FDR.data = metab.FDR[sig.metab, c(x)]
mean.data = mean.data[sig.metab, ]

data = data.frame(Metab = sig.metab,
                  compoundID = KEGG.Key[sig.metab],
                  FDR = -log2(FDR.data),
                  FC = FC.data,
                  Color = 'grey',
                  size = 1,
                  Variable1.mean = mean.data[,1],
                  Control1.mean = mean.data[,2])

sig.metabs = data

main.title = "G85R Metabolites"

  data$Color[data$FC < 0] = 'royalblue'
  data$Color[data$FC > 0] = 'firebrick' 
  #data$Color[data$Symbol == "Arc1"] = 'gold'
  data$size = log10(apply(mean.data[,c(1,2)], 1, max))
  data$size[data$size<3] = 1

  data = na.omit(data)
  data = data[data$size > log2(min.expression),]
  data$size = data$size

  fig = plot_ly(data = data,
                x = ~FC,
                y = ~FDR,
                type = 'scatter',
                mode = 'markers',
                marker = list(color = ~Color,
                              colors = ~Color,
                              line = list(color = "black", width = 1.5),
                              size = ~size),
                hoverinfo = "text",
                hovertext = paste("Metabolite:", data$Metab,
                                  "\n Compound ID:", data$compoundID,
                                  "\n ", x, "p: ", round(2^-data$FDR, 2),
                                  "\n ", x, "FC: ", round(data$FC, 2),
                                  "\n ", colnames(mean.data)[1], "mean: ", signif(data$Variable1,2),
                                  "\n ", colnames(mean.data)[2], "mean: ", signif(data$Control1,2)))

  fig = fig %>% layout(title = main.title,
                       xaxis = list(title = paste(x, "FC")),
                       yaxis = list(title = paste(x, "-log2(p-value)")),
                       showlegend = F)

  fig

```

```{r echo = F}

CompoundData = sig.metabs[1,]

x = 'GRxWT.C'
y = 'GRxWT.C'

mean.data = mean.metab[, c('GR', 'WT')]

Fx = 'GRxWT.FC'
My = 'GRxWT.MC'

Fmean = G85R.meancpm[, c('GR.F', 'WT.F')]
Mmean = G85R.meancpm[, c('GR.M', 'WT.M')]

compoundandgenes(CompoundData, x, y, mean.data, Fx, My, Fmean, Mmean)

```

```{r echo = F}

CompoundData = sig.metabs[2,]

x = 'GRxWT.C'
y = 'GRxWT.C'

mean.data = mean.metab[, c('GR', 'WT')]

Fx = 'GRxWT.FC'
My = 'GRxWT.MC'

Fmean = G85R.meancpm[, c('GR.F', 'WT.F')]
Mmean = G85R.meancpm[, c('GR.M', 'WT.M')]

compoundandgenes(CompoundData, x, y, mean.data, Fx, My, Fmean, Mmean)

```

```{r echo = F}

CompoundData = sig.metabs[3,]

x = 'GRxWT.C'
y = 'GRxWT.C'

mean.data = mean.metab[, c('GR', 'WT')]

Fx = 'GRxWT.FC'
My = 'GRxWT.MC'

Fmean = G85R.meancpm[, c('GR.F', 'WT.F')]
Mmean = G85R.meancpm[, c('GR.M', 'WT.M')]

compoundandgenes(CompoundData, x, y, mean.data, Fx, My, Fmean, Mmean)

```

```{r echo = F}

CompoundData = sig.metabs[4,]

x = 'GRxWT.C'
y = 'GRxWT.C'

mean.data = mean.metab[, c('GR', 'WT')]

Fx = 'GRxWT.FC'
My = 'GRxWT.MC'

Fmean = G85R.meancpm[, c('GR.F', 'WT.F')]
Mmean = G85R.meancpm[, c('GR.M', 'WT.M')]

compoundandgenes(CompoundData, x, y, mean.data, Fx, My, Fmean, Mmean)

```

```{r echo = F}

CompoundData = sig.metabs[5,]

x = 'GRxWT.C'
y = 'GRxWT.C'

mean.data = mean.metab[, c('GR', 'WT')]

Fx = 'GRxWT.FC'
My = 'GRxWT.MC'

Fmean = G85R.meancpm[, c('GR.F', 'WT.F')]
Mmean = G85R.meancpm[, c('GR.M', 'WT.M')]

compoundandgenes(CompoundData, x, y, mean.data, Fx, My, Fmean, Mmean)

```

```{r echo = F}

CompoundData = sig.metabs[41,]

x = 'GRxWT.C'
y = 'GRxWT.C'

mean.data = mean.metab[, c('GR', 'WT')]

Fx = 'GRxWT.FC'
My = 'GRxWT.MC'

Fmean = G85R.meancpm[, c('GR.F', 'WT.F')]
Mmean = G85R.meancpm[, c('GR.M', 'WT.M')]

compoundandgenes(CompoundData, x, y, mean.data, Fx, My, Fmean, Mmean)

```

##  Are there genes that respond to rescue in Silent and G85R for both sexes?

##  TKT-Df Response
```{r echo = F}

x = "GRF.CxDf"
y = "WTF.CxDf"

mean.data = G85R.meancpm[, c('GR.F', 'WT.F',
                             'TktDfGR.F', 'TktDfWT.F')]
main.title = "Female TKT-Df DE Genes"

##RG = Response Genes
FDf.RG = TKT.biplot(x, y, mean.data, main.title)

FDf.RG[[1]]

```

```{r echo = F}

x = "GRM.CxDf"
y = "WTM.CxDf"

mean.data = G85R.meancpm[, c('GR.M', 'WT.M',
                             'TktDfGR.M', 'TktDfWT.M')]
main.title = "Male TKT-Df DE Genes"

##RG = Response Genes
MDf.RG = TKT.biplot(x, y, mean.data, main.title)

MDf.RG[[1]]

```

```{r echo = F}

x = "GRF.CxDf"
y = "GRM.CxDf"

mean.data = G85R.meancpm[, c('GR.F', 'GR.M',
                             'TktDfGR.F', 'TktDfGR.M')]
main.title = "G85R TKT-Df DE Genes"

##RG = Response Genes
GRDf.RG = TKT.biplot(x, y, mean.data, main.title)

GRDf.RG[[1]]

```

```{r echo = F}

x = "WTF.CxDf"
y = "WTM.CxDf"

mean.data = G85R.meancpm[, c('WT.F', 'WT.M',
                             'TktDfWT.F', 'TktDfWT.M')]
main.title = "Silent TKT-Df DE Genes"

##RG = Response Genes
WTDf.RG = TKT.biplot(x, y, mean.data, main.title)

WTDf.RG[[1]]

```

```{r echo = F}

##Sample titles as strings. Only fill in up to your number of selected categories
set1 = "Silent Female Df"
set2 = "Silent Male Df"
set3 = "G85R Female Df"
set4 = "G85R Male Df"

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1 = FDf.RG[[2]][FDf.RG[[2]][,'FDR2'] <= 0.05, "FBgn"]
s2 = MDf.RG[[2]][MDf.RG[[2]][,'FDR2'] <= 0.05, "FBgn"]
s3 = FDf.RG[[2]][FDf.RG[[2]][,'FDR1'] <= 0.05, "FBgn"]
s4 = MDf.RG[[2]][MDf.RG[[2]][,'FDR1'] <= 0.05, "FBgn"]

NormalResponse = data.frame(Rescue = c(set1, set2, set3, set4),
                            TotalGenes =  c(length(s1), length(s2), length(s3), length(s4)))

fig <- plot_ly(data = NormalResponse, x = ~Rescue, y = ~TotalGenes, type = 'bar', name = 'Normal Response',
               marker = list(color = 'lightgrey',
                      line = list(color = 'rgb(8,48,107)',
                                  width = 3)))
fig <- fig %>% layout(yaxis = list(title = 'Total Response Genes (Control vs TKT-Df'))

fig

```

```{r echo = F}

##Sample titles as strings. Only fill in up to your number of selected categories
set1 = "Silent Female Df"
set2 = "Silent Male Df"
set3 = "G85R Female Df"
set4 = "G85R Male Df"

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1 = FDf.RG[[2]][FDf.RG[[2]][,'FDR2'] <= 0.05, "FBgn"]
s2 = MDf.RG[[2]][MDf.RG[[2]][,'FDR2'] <= 0.05, "FBgn"]
s3 = FDf.RG[[2]][FDf.RG[[2]][,'FDR1'] <= 0.05, "FBgn"]
s4 = MDf.RG[[2]][MDf.RG[[2]][,'FDR1'] <= 0.05, "FBgn"]

main = ""

grid.newpage()
tempvenn=draw.quad.venn(area1=length(s1),area2=length(s2),area3=length(s3),area4=length(s4), n12=length(intersect(s1,s2)),n13 = length(intersect(s1,s3)),n14 = length(intersect(s1,s4)),n23 = length(intersect(s2,s3)),n24 = length(intersect(s2,s4)),n34 = length(intersect(s3,s4)),n123 =  length(intersect(s1,intersect(s2,s3))),n124 =  length(intersect(s1,intersect(s2,s4))), n134 = length(intersect(s1,intersect(s3,s4))),n234 = length(intersect(s2,intersect(s3,s4))),n1234 = length(intersect(s1,intersect(s2,intersect(s3,s4)))),category =c(set1,set2,set3,set4),fill = c("blue", "red","green","yellow"),cex=2,cat.cex = 1, title = "main")

```

##WTF.C vs WTF.TKT-Df
```{r echo = F, include = F}

bg = row.names(TKT.EdgeR)
WTF.Df.Enr = Subset.Enrichment(bg, FDf.RG[[2]][FDf.RG[[2]][,'FDR2'] <= 0.05, "FBgn"])

```

```{r echo = F}

RTCC.WTF.Df = tree(WTF.Df.Enr, "CC")
treemapPlot(RTCC.WTF.Df)

```

```{r echo = F}

RTBP.WTF.Df = tree(WTF.Df.Enr, "BP")
treemapPlot(RTBP.WTF.Df)

```

```{r echo = F}

RTMF.WTF.Df = tree(WTF.Df.Enr, "MF")
treemapPlot(RTMF.WTF.Df)

```

##WTM.C vs WTM.TKT-Df
```{r echo = F, include = F}

bg = row.names(TKT.EdgeR)
WTM.Df.Enr = Subset.Enrichment(bg, MDf.RG[[2]][MDf.RG[[2]][,'FDR2'] <= 0.05, "FBgn"])

```

```{r echo = F}

RTCC.WTM.Df = tree(WTM.Df.Enr, "CC")
treemapPlot(RTCC.WTM.Df)

```

```{r echo = F}

RTBP.WTF.Df = tree(WTF.Df.Enr, "BP")
treemapPlot(RTBP.WTF.Df)

```

```{r echo = F}

RTMF.WTF.Df = tree(WTF.Df.Enr, "MF")
treemapPlot(RTMF.WTF.Df)

```

##GRF.C vs GRF.TKT-Df
```{r echo = F, include = F}

bg = row.names(TKT.EdgeR)
GRF.Df.Enr = Subset.Enrichment(bg, FDf.RG[[2]][FDf.RG[[2]][,'FDR1'] <= 0.05, "FBgn"])

```

```{r echo = F}

RTCC.WTF.Df = tree(WTF.Df.Enr, "CC")
treemapPlot(RTCC.WTF.Df)

```

```{r echo = F}

RTBP.WTF.Df = tree(WTF.Df.Enr, "BP")
treemapPlot(RTBP.WTF.Df)

```

```{r echo = F}

RTMF.WTF.Df = tree(WTF.Df.Enr, "MF")
treemapPlot(RTMF.WTF.Df)

```

##GRM.C vs GRM.TKT-Df
```{r echo = F, include = F}

bg = row.names(TKT.EdgeR)
GRM.Df.Enr = Subset.Enrichment(bg, MDf.RG[[2]][MDf.RG[[2]][,'FDR1'] <= 0.05, "FBgn"])

```

```{r echo = F}

RTCC.WTF.Df = tree(WTF.Df.Enr, "CC")
treemapPlot(RTCC.WTF.Df)

```

```{r echo = F}

RTBP.WTF.Df = tree(WTF.Df.Enr, "BP")
treemapPlot(RTBP.WTF.Df)

```

```{r echo = F}

RTMF.WTF.Df = tree(WTF.Df.Enr, "MF")
treemapPlot(RTMF.WTF.Df)

```

```{r echo = F}

##Sample titles as strings. Only fill in up to your number of selected categories
set1 = "Silent Female Df"
set2 = "Silent Male Df"
set3 = "G85R Female Df"
set4 = "G85R Male Df"

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1 = WTF.Df.Enr$GO$term[WTF.Df.Enr$GO$adjp <= .05]
s2 = WTM.Df.Enr$GO$term[WTM.Df.Enr$GO$adjp <= .05]
s3 = GRF.Df.Enr$GO$term[GRF.Df.Enr$GO$adjp <= .05]
s4 = GRM.Df.Enr$GO$term[GRM.Df.Enr$GO$adjp <= .05]

main = ""

grid.newpage()
tempvenn=draw.quad.venn(area1=length(s1),area2=length(s2),area3=length(s3),area4=length(s4), n12=length(intersect(s1,s2)),n13 = length(intersect(s1,s3)),n14 = length(intersect(s1,s4)),n23 = length(intersect(s2,s3)),n24 = length(intersect(s2,s4)),n34 = length(intersect(s3,s4)),n123 =  length(intersect(s1,intersect(s2,s3))),n124 =  length(intersect(s1,intersect(s2,s4))), n134 = length(intersect(s1,intersect(s3,s4))),n234 = length(intersect(s2,intersect(s3,s4))),n1234 = length(intersect(s1,intersect(s2,intersect(s3,s4)))),category =c(set1,set2,set3,set4),fill = c("blue", "red","green","yellow"),cex=2,cat.cex = 1, title = "main")

```

###################################

###################################

###################################

## TKT-OE Response
```{r echo = F}

x = "GRF.CxOE"
y = "WTF.CxOE"

mean.data = G85R.meancpm[, c('GR.F', 'WT.F',
                             'TktOEGR.F', 'TktOEWT.F')]
main.title = "Female TKT-OE DE Genes"

##RG = Response Genes
FOE.RG = TKT.biplot(x, y, mean.data, main.title)

FOE.RG[[1]]

```

```{r echo = F}

x = "GRM.CxOE"
y = "WTM.CxOE"

mean.data = G85R.meancpm[, c('GR.M', 'WT.M',
                             'TktOEGR.M', 'TktOEWT.M')]
main.title = "Male TKT-OE DE Genes"

##RG = Response Genes
MOE.RG = TKT.biplot(x, y, mean.data, main.title)

MOE.RG[[1]]

```

```{r echo = F}

x = "GRF.CxOE"
y = "GRM.CxOE"

mean.data = G85R.meancpm[, c('GR.F', 'GR.M',
                             'TktOEGR.F', 'TktOEGR.M')]
main.title = "G85R TKT-OE DE Genes"

##RG = Response Genes
GROE.RG = TKT.biplot(x, y, mean.data, main.title)

GROE.RG[[1]]

```

```{r echo = F}

x = "WTF.CxOE"
y = "WTM.CxOE"

mean.data = G85R.meancpm[, c('WT.F', 'WT.M',
                             'TktOEWT.F', 'TktOEWT.M')]
main.title = "Silent TKT-OE DE Genes"

##RG = Response Genes
WTOE.RG = TKT.biplot(x, y, mean.data, main.title)

WTOE.RG[[1]]

```

```{r echo = F}

##Sample titles as strings. Only fill in up to your number of selected categories
set1 = "Silent Female OE"
set2 = "Silent Male OE"
set3 = "G85R Female OE"
set4 = "G85R Male OE"

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1 = FOE.RG[[2]][FOE.RG[[2]][,'FDR2'] <= 0.05, "FBgn"]
s2 = MOE.RG[[2]][MOE.RG[[2]][,'FDR2'] <= 0.05, "FBgn"]
s3 = FOE.RG[[2]][FOE.RG[[2]][,'FDR1'] <= 0.05, "FBgn"]
s4 = MOE.RG[[2]][MOE.RG[[2]][,'FDR1'] <= 0.05, "FBgn"]

NormalResponse = data.frame(Rescue = c(set1, set2, set3, set4),
                            TotalGenes =  c(length(s1), length(s2), length(s3), length(s4)))

fig <- plot_ly(data = NormalResponse, x = ~Rescue, y = ~TotalGenes, type = 'bar', name = 'Normal Response',
               marker = list(color = 'lightgrey',
                      line = list(color = 'rgb(8,48,107)',
                                  width = 3)))
fig <- fig %>% layout(yaxis = list(title = 'Total Response Genes (Control vs TKT-Df'))

fig

```

```{r echo = F}

##Sample titles as strings. Only fill in up to your number of selected categories
set1 = "Silent Female OE"
set2 = "Silent Male OE"
set3 = "G85R Female OE"
set4 = "G85R Male OE"

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1 = FOE.RG[[2]][FOE.RG[[2]][,'FDR2'] <= 0.05, "FBgn"]
s2 = MOE.RG[[2]][MOE.RG[[2]][,'FDR2'] <= 0.05, "FBgn"]
s3 = FOE.RG[[2]][FOE.RG[[2]][,'FDR1'] <= 0.05, "FBgn"]
s4 = MOE.RG[[2]][MOE.RG[[2]][,'FDR1'] <= 0.05, "FBgn"]

main = ""

grid.newpage()
tempvenn=draw.quad.venn(area1=length(s1),area2=length(s2),area3=length(s3),area4=length(s4), n12=length(intersect(s1,s2)),n13 = length(intersect(s1,s3)),n14 = length(intersect(s1,s4)),n23 = length(intersect(s2,s3)),n24 = length(intersect(s2,s4)),n34 = length(intersect(s3,s4)),n123 =  length(intersect(s1,intersect(s2,s3))),n124 =  length(intersect(s1,intersect(s2,s4))), n134 = length(intersect(s1,intersect(s3,s4))),n234 = length(intersect(s2,intersect(s3,s4))),n1234 = length(intersect(s1,intersect(s2,intersect(s3,s4)))),category =c(set1,set2,set3,set4),fill = c("blue", "red","green","yellow"),cex=2,cat.cex = 1, title = "main")

```