### 2024-09-27 gbbmeetingfigures.Rmd

---
title: "gbb conference analyses"
author: "John Santiago"
date: "2024-07-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = F, echo = F}

library(plotly)
library(edgeR)
library(org.Dm.eg.db)
library(VennDiagram)

gbb.genes = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/BMP Signaling Genes.csv",row.names=1)

##Clean Trait Data and calculate CPMs
countdata=read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT.counttable.csv",row.names=1)
groups=read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/All.TKT.metadata.csv",row.names=1)

groups = groups[setdiff(row.names(groups), c("GR.F3", "WT.F3", "TktDfWT.F3")),]
##normalize data
countdata=countdata[, row.names(groups)]
countdata = countdata[-(grep("ti",row.names(countdata))),]
countdata = countdata[-(grep("trans",row.names(countdata))),]

x <- countdata
gr <- factor(groups$Group)
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
GR.cpmdata=cpm(z)

design<-model.matrix(~0 + gr)
colnames(design) = unique(gr)
z = estimateDisp(z, design, robust = T)
fit <- glmQLFit(z, design)
plotQLDisp(fit)

##Will calculate fold-change with the first over the second
compare = makeContrasts(TktDfGR.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.FCxGR.FDf  = DEout
sWT.FCxGR.FDf = DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfGR.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.MCxGR.MDf  = DEout
sWT.MCxGR.MDf = DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.F-GR.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.FOExGR.F  = DEout
sGR.FOExGR.F = DEout[DEout$FDR<.05,]

compare = makeContrasts(TktOEGR.M-GR.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.MOExGR.M  = DEout
sGR.MOExGR.M = DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfWT.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.FDfxWT.F  = DEout
sWT.FDfxWT.F = DEout[DEout$FDR<.05,]

compare = makeContrasts(TktDfWT.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.MDfxWT.M  = DEout
sWT.MDfxWT.M = DEout[DEout$FDR<.05,]

compare = makeContrasts(GR.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.F = DEout

compare = makeContrasts(GR.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.M  = DEout

compare = makeContrasts(TktOEGR.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRDf.F  = DEout

compare = makeContrasts(TktOEGR.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRDf.M  = DEout

compare = makeContrasts(TktOEWT.F-WT.F, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WTDf.F  = DEout

compare = makeContrasts(TktOEWT.M-WT.M, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WTDf.M  = DEout

gbbOE = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbOE_CountTable.csv", row.names = 1)
gbbOE.groups = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbOE_Metadata.csv", row.names = 1)
gbbOE = gbbOE[,row.names(gbbOE.groups)]
gbbOE.groups$group = paste0(gbbOE.groups$genotype, "_",gbbOE.groups$stage)

#anqi.sampletrim = c("AHH0001",	"AHH0002",	"AHH0003",	"AHH0004",	"AHH0007",	#"AHH0008",	"AHH0009",	"AHH0011",	"AHH0012",	"AHH0015",	"AHH0016",	#"AHH0018",	"AHH0019",	"AHH0026")
#gbbOE= gbbOE[,anqi.sampletrim]
#gbbOE.groups = gbbOE.groups[anqi.sampletrim,]

#anqi.genetrim = row.names(read.csv('/Users/johncsantiago/Documents/gbbOE.CPM_c#ounts_all samples.csv', row.names =1))

#gbbOE= na.omit(gbbOE[anqi.genetrim,anqi.sampletrim])
#gbbOE.groups = gbbOE.groups[anqi.sampletrim,]

##normalize data
countdata=gbbOE
x <- countdata
gr <- factor(gbbOE.groups$group, levels=c(unique(gbbOE.groups$group)))
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
gbbOE.cpmdata=cpm(z)

design<-model.matrix(~0 + gr)
colnames(design) = unique(gr)
z = estimateDisp(z, design, robust = T)
fit <- glmQLFit(z, design)
plotQLDisp(fit)

compare = makeContrasts(G85R_LL3-LoxP_LL3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.LL3=DEout
sGRxWT.LL3=DEout[DEout$FDR<.05,]

compare = makeContrasts(LoxP_LL3-G85R_LL3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WTxGR.LL3=DEout
sWTxGR.LL3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_ML3-LoxP_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GRxWT.ML3=DEout
sGRxWT.ML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_LL3-G85R_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
GR.LL3xML3=DEout
sGR.LL3xML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(LoxP_LL3-LoxP_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
WT.LL3xML3=DEout
sWT.LL3xML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_gbb_LL3-G85R_gbb_ML3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
G85R_gbb.LL3xML3=DEout
sG85R_gbb.LL3xML3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_gbb_LL3-G85R_LL3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
G85R_gbbxGR.LL3=DEout
sG85R_gbbxGR.LL3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_gbb_LL3-LoxP_LL3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
G85R_gbbxWT.LL3=DEout
sG85R_gbbxWT.LL3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_gbb_LL3-G85R_LL3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
G85R_gbbxG85R.LL3=DEout
sG85R_gbbxG85R.LL3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_LL3-LoxP_LL3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
G85RxLoxP.LL3=DEout
sG85RxLoxP.LL3=DEout[DEout$FDR<.05,]

compare = makeContrasts(G85R_gbb_LL3-LoxP_LL3, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
G85R_gbbxLoxP.LL3=DEout
sG85R_gbbxLoxP.LL3=DEout[DEout$FDR<.05,]

gbbKO = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbKO_CountTable.csv", row.names = 1)
gbbKO.groups = read.csv("/Users/johncsantiago/Documents/GitHub/Datasets/gbbKO_Metadata.csv", row.names = 1)

gbbKO = gbbKO[,colnames(gbbKO)!= 'AHH0025']
##gbbKO.groups = gbbKO.groups[row.names(gbbKO.groups)!= 'AHH0025',]

anqi.genetrim = row.names(read.csv('/Users/johncsantiago/Documents/gbbKO.CPM_170629_gbbKnockout.csv', row.names =1))

gbbKO = (gbbKO[anqi.genetrim,])

##normalize data
countdata=gbbKO
x <- countdata
gr <- factor(gbbKO.groups$Genotype[1:5], levels=c(unique(gbbKO.groups$Genotype[1:5])))
y <- DGEList(counts=x,group=gr)
keep <- filterByExpr(y)
y <- y[keep,,keep.lib.sizes=FALSE]
z <- calcNormFactors(y, method = "TMM")
##normalized cpm
gbbKO.cpmdata=cpm(z)

design<-model.matrix(~0 + gr)
colnames(design) = unique(gr)
z = estimateDisp(z, design, robust = T)
fit <- glmQLFit(z, design)
plotQLDisp(fit)

compare = makeContrasts(KO-WT, levels=design)
lrt = glmQLFTest(fit,contrast=as.vector(compare))
G_X_E<-topTags(lrt,adjust.method="BH",n = nrow(z$counts), sort.by="PValue")
DEout=G_X_E$table
gbbKO.DEG=DEout
sgbbKO.DEG=DEout[DEout$FDR<.05,]

gbbKO.mean = data.frame(gbbKO.DEG[,1:2])
gbbKO.mean[,1] = apply(gbbKO.cpmdata[row.names(gbbKO.mean),
                                     c('AHH0020', 'AHH0021', 'AHH0027')], MARGIN = 1, mean)
gbbKO.mean[,2] = apply(gbbKO.cpmdata[row.names(gbbKO.mean),
                                     c('AHH0023', 'AHH0024')], MARGIN = 1, mean)
                                     ##c('AHH0023', 'AHH0024', 'AHH0025')], MARGIN = 1, mean)
colnames(gbbKO.mean) = c("WT", "KO")

GeneIDKey = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GeneIDKey.csv", row.names = 1)

```

```{r echo = F}

gbbKOGRxWT = row.names(gbbKO.DEG[gbbKO.DEG$FDR <.1,])
tktDfWTxWT.F = row.names(WT.FDfxWT.F[WT.FDfxWT.F$FDR <.05,])
tktDfWTxWT.M = row.names(WT.MDfxWT.M[WT.MDfxWT.M$FDR <.05,])

##Sample titles as strings. Only fill in up to your number of selected categories
set1 = "G85R-gbbKO vs WT (pvalue < .05)"
set2 = "Female WT-TKT.Df vs WT (FDR < .05)"
set3 = "Male WT-TKT.Df vs WT (FDR < .05)"
set4 = "BMP Pathway"

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1 = gbbKOGRxWT
s2 = tktDfWTxWT.F
s3 = tktDfWTxWT.M
s4 = row.names(gbb.genes)

grid.newpage()
tempvenn=draw.quad.venn(area1=length(s1),area2=length(s2),area3=length(s3),area4=length(s4), n12=length(intersect(s1,s2)),n13 = length(intersect(s1,s3)),n14 = length(intersect(s1,s4)),n23 = length(intersect(s2,s3)),n24 = length(intersect(s2,s4)),n34 = length(intersect(s3,s4)),n123 =  length(intersect(s1,intersect(s2,s3))),n124 =  length(intersect(s1,intersect(s2,s4))), n134 = length(intersect(s1,intersect(s3,s4))),n234 = length(intersect(s2,intersect(s3,s4))),n1234 = length(intersect(s1,intersect(s2,intersect(s3,s4)))),category =c(set1,set2,set3,set4),fill = c("blue", "red","green","yellow"),cex=2,cat.cex = 1)

```

```{r echo = F}

gbbKOGRxWT = row.names(G85R_gbbxG85R.LL3[G85R_gbbxG85R.LL3$PValue <.05,])
tktOEGRxGR.F = row.names(GR.FOExGR.F[GR.FOExGR.F$FDR <.05,])
tktOEGRxGR.M = row.names(GR.MOExGR.M[GR.MOExGR.M$FDR <.05,])

##Sample titles as strings. Only fill in up to your number of selected categories
set1 = "G85R-gbbKO vs WT (p-value < .05)"
set2 = "Female G85R-TKT.OE vs WT (FDR < .05)"
set3 = "Male G85R-TKT.OE vs WT (FDR < .05)"
set4 = "BMP Pathway"

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1 = gbbKOGRxWT
s2 = tktOEGRxGR.F
s3 = tktOEGRxGR.M
s4 = row.names(gbb.genes)

grid.newpage()
tempvenn=draw.quad.venn(area1=length(s1),area2=length(s2),area3=length(s3),area4=length(s4), n12=length(intersect(s1,s2)),n13 = length(intersect(s1,s3)),n14 = length(intersect(s1,s4)),n23 = length(intersect(s2,s3)),n24 = length(intersect(s2,s4)),n34 = length(intersect(s3,s4)),n123 =  length(intersect(s1,intersect(s2,s3))),n124 =  length(intersect(s1,intersect(s2,s4))), n134 = length(intersect(s1,intersect(s3,s4))),n234 = length(intersect(s2,intersect(s3,s4))),n1234 = length(intersect(s1,intersect(s2,intersect(s3,s4)))),category =c(set1,set2,set3,set4),fill = c("blue", "red","green","yellow"),cex=2,cat.cex = 1)

```

```{r echo = F}

gbbOExWT = row.names(G85R_gbbxLoxP.LL3[G85R_gbbxLoxP.LL3$PValue <.05,])
GRxWT = row.names(G85RxLoxP.LL3[G85RxLoxP.LL3$PValue <.05,])
gbbOExG85R = row.names(G85R_gbbxG85R.LL3[G85R_gbbxG85R.LL3$PValue <.05,])

##Sample titles as strings. Only fill in up to your number of selected categories
set1 = "G85R-gbbOE vs WT (pvalue < .05)"
set2 = "G85R vs WT (FDR < .05)"
set3 = "G85R-gbbOE vs G85R (FDR < .05)"
set4 = "BMP Pathway"

##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
s1 = gbbOExWT
s2 = GRxWT
s3 = gbbOExG85R
s4 = row.names(gbb.genes)

grid.newpage()
tempvenn=draw.quad.venn(area1=length(s1),area2=length(s2),area3=length(s3),area4=length(s4), n12=length(intersect(s1,s2)),n13 = length(intersect(s1,s3)),n14 = length(intersect(s1,s4)),n23 = length(intersect(s2,s3)),n24 = length(intersect(s2,s4)),n34 = length(intersect(s3,s4)),n123 =  length(intersect(s1,intersect(s2,s3))),n124 =  length(intersect(s1,intersect(s2,s4))), n134 = length(intersect(s1,intersect(s3,s4))),n234 = length(intersect(s2,intersect(s3,s4))),n1234 = length(intersect(s1,intersect(s2,intersect(s3,s4)))),category =c(set1,set2,set3,set4),fill = c("blue", "red","green","yellow"),cex=2,cat.cex = 1)

```

```{r echo = F}

grid.newpage()
draw.triple.venn(area1=length(s1),area2=length(s2),area3=length(s3),n12=length(intersect(s1,s2)),n23 = length(intersect(s2,s3)),n13 = length(intersect(s1,s3)),n123 = length(intersect(intersect(s1,s2),s3)),category=c(set1,set2,set3),fill = c("green", "blue","red"),cex=1.5,cat.cex = 1)

#Quad Venn
grid.newpage()
tempvenn=draw.quad.venn(area1=length(s1),area2=length(s2),area3=length(s3),area4=length(s4), n12=length(intersect(s1,s2)),n13 = length(intersect(s1,s3)),n14 = length(intersect(s1,s4)),n23 = length(intersect(s2,s3)),n24 = length(intersect(s2,s4)),n34 = length(intersect(s3,s4)),n123 =  length(intersect(s1,intersect(s2,s3))),n124 =  length(intersect(s1,intersect(s2,s4))), n134 = length(intersect(s1,intersect(s3,s4))),n234 = length(intersect(s2,intersect(s3,s4))),n1234 = length(intersect(s1,intersect(s2,intersect(s3,s4)))),category =c(set1,set2,set3,set4),fill = c("blue", "red","green","yellow"),cex=2,cat.cex = 1)

```

```{r}

  FC.data = setNames(gbbKO.DEG[,1], row.names(gbbKO.DEG))
  FDR.data = setNames(gbbKO.DEG[,4], row.names(gbbKO.DEG))
  mean.data = gbbKO.mean[,1:2]

  ##select F if you want point size to scale with mean cpm of variable data
  set.size = F
  ##select T if you only want the genes in category to show
  catgenes.only = T

  volcano.data = data.frame(Symbol = GeneIDKey[names(FDR.data), "Symbol"], 
                            FDR = -log2(FDR.data),
                            FC = FC.data,
                            Color = 'grey',
                            size = 5*log(mean.data[,1]+2),
                            Variable.cpm = mean.data[,1],
                            Control.cpm = mean.data[,2])

  if(set.size == T){
    volcano.data$size = 15
    volcano.data$size[volcano.data$FDR <= -log2(.05)] = 5
  }

  volcano.data$Color[volcano.data$FDR >= -log2(.05) & volcano.data$FC < 0] = 'lightblue'
  volcano.data$Color[volcano.data$FDR >= -log2(.0001) & volcano.data$FC < 0] = 'steelblue'
  volcano.data$Color[volcano.data$FDR >= -log2(.000001) & volcano.data$FC < 0] = 'dodgerblue'

  volcano.data$Color[volcano.data$FDR >= -log2(.05) & volcano.data$FC > 0] = 'lightcoral'
  volcano.data$Color[volcano.data$FDR >= -log2(.0001) & volcano.data$FC > 0] = 'tomato'
  volcano.data$Color[volcano.data$FDR >= -log2(.000001) & volcano.data$FC > 0] = 'firebrick'

  volcano.data$size[volcano.data$FDR <= -log2(.05)] = 3

  main.title = paste(colnames(mean.data)[1], "vs. ", colnames(mean.data)[2])

if(set.size == T){
  volcano.data$size = 15
  volcano.data$size[volcano.data$FDR <= -log2(.05)] = 5
}

volcano.data = volcano.data[volcano.data$FDR >= -log2(.1),]

fig = plot_ly(data = volcano.data,
              x = ~FC,
              y = ~FDR,
              type = 'scatter',
              mode = 'markers',
              marker = list(color = ~Color, colors = ~Color, size = volcano.data$size,
                            line = list(color = 'black', width = .5)),
              hoverinfo = "text",
              hovertext = paste("Gene:", volcano.data$Symbol,
                                "\n-log2(FDR): ", round(volcano.data$FDR,2),
                                "\nFC: ", round(volcano.data$FC,2),
                                "\n ", colnames(mean.data)[1], "mean cpm: ",
                                round(volcano.data$Variable.cpm, 1),
                                "\n ", colnames(mean.data)[2], "mean cpm: ",
                                round(volcano.data$Control.cpm, 1)))
fig = fig %>% layout(title = main.title,
                     yaxis = list(ymin = 0, ymax = 9))

fig

```

```{r echo = F, warning = F}

library(goseq)

FDR = gbbKO.DEG
  sigs = row.names(FDR)[FDR[,'FDR'] < .05]
  bg = row.names(FDR)

  sigsKEY = GeneIDKey[sigs,]
  sigs = na.omit(sigsKEY$Symbol)
  bgKEY = GeneIDKey[bg,]
  bg = na.omit(bgKEY$Symbol)

  genes = setNames(rep(0, length(bg)), bg)
  genes[sigs] = 1
  pwf=nullp(genes,"dm3","geneSymbol")
  GO.wall=goseq(pwf,"dm3","geneSymbol")
  GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
  row.names(GO.wall) = GO.wall$category
  KEGG=goseq(pwf,gene2cat=GenesInKegg)
  KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
  row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
  KEGG$Name=KEGG.Names[row.names(KEGG),1]

  enrichment.data = list(GO.wall, KEGG)
  names(enrichment.data) = c("GO", "KEGG")

```

```{r}

  GO.data = enrichment.data$GO
  GO.ontology = "BP"

  go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .05 & 
                                   GO.data$ontology == GO.ontology]
  scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

  simMatrix <- calculateSimMatrix(go_analysis,
                                  orgdb="org.Dm.eg.db",
                                  ont=GO.ontology,
                                  method="Rel")

  reducedTerms <- reduceSimMatrix(simMatrix,
                                  scores,
                                  threshold=0.7,
                                  orgdb="org.Dm.eg.db")

```

```{r}

treemapPlot(reducedTerms)

```

```{r echo = F, warning = F}

library(goseq)

FDR = G85R_gbbxG85R.LL3
  sigs = row.names(FDR)[FDR[,'FDR'] < .1]
  bg = row.names(FDR)

  sigsKEY = GeneIDKey[sigs,]
  sigs = na.omit(sigsKEY$Symbol)
  bgKEY = GeneIDKey[bg,]
  bg = na.omit(bgKEY$Symbol)

  genes = setNames(rep(0, length(bg)), bg)
  genes[sigs] = 1
  pwf=nullp(genes,"dm3","geneSymbol")
  GO.wall=goseq(pwf,"dm3","geneSymbol")
  GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
  row.names(GO.wall) = GO.wall$category
  KEGG=goseq(pwf,gene2cat=GenesInKegg)
  KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
  row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
  KEGG$Name=KEGG.Names[row.names(KEGG),1]

  enrichment.data = list(GO.wall, KEGG)
  names(enrichment.data) = c("GO", "KEGG")

```

```{r}

  GO.data = enrichment.data$GO
  GO.ontology = "BP"

  go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .05 & 
                                   GO.data$ontology == GO.ontology]
  scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

  simMatrix <- calculateSimMatrix(go_analysis,
                                  orgdb="org.Dm.eg.db",
                                  ont=GO.ontology,
                                  method="Rel")

  reducedTerms <- reduceSimMatrix(simMatrix,
                                  scores,
                                  threshold=0.7,
                                  orgdb="org.Dm.eg.db")

```

```{r}

treemapPlot(reducedTerms)

```

```{r echo = F, warning = F}

library(goseq)
temp = read.csv('/Users/johncsantiago/Downloads/LoxP-G85R_LL3.csv',
                row.names = 1, header = T)

FDR = G85RxLoxP.LL3
  sigs = row.names(FDR)[FDR[,'FDR'] < .1]
  bg = row.names(FDR)

  sigsKEY = GeneIDKey[sigs,]
  sigs = na.omit(sigsKEY$Symbol)
  bgKEY = GeneIDKey[bg,]
  bg = na.omit(bgKEY$Symbol)

  genes = setNames(rep(0, length(bg)), bg)
  genes[sigs] = 1
  pwf=nullp(genes,"dm3","geneSymbol")
  GO.wall=goseq(pwf,"dm3","geneSymbol")
  GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
  row.names(GO.wall) = GO.wall$category
  KEGG=goseq(pwf,gene2cat=GenesInKegg)
  KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
  row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
  KEGG$Name=KEGG.Names[row.names(KEGG),1]

  enrichment.data = list(GO.wall, KEGG)
  names(enrichment.data) = c("GO", "KEGG")

```

```{r}

  GO.data = enrichment.data$GO
  GO.ontology = "BP"

  go_analysis = GO.data$category[GO.data$over_represented_pvalue <= .05 & 
                                   GO.data$ontology == GO.ontology]
  scores = setNames(-log10(GO.data$over_represented_pvalue), go_analysis)

  simMatrix <- calculateSimMatrix(go_analysis,
                                  orgdb="org.Dm.eg.db",
                                  ont=GO.ontology,
                                  method="Rel")

  reducedTerms <- reduceSimMatrix(simMatrix,
                                  scores,
                                  threshold=0.7,
                                  orgdb="org.Dm.eg.db")

```

```{r}

treemapPlot(reducedTerms)

```

```{r echo = F, message = F}

library(heatmaply)
library(dendextend)

counts = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.countdata.csv", row.names = 1)

cpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.cpmdata.csv", row.names = 1)

meancpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.meancpmdata.csv", row.names = 1)

FDR = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FDRdata.csv", row.names = 1)

FC = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.FCdata.csv", row.names = 1)

meta = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/A4V_BodySections_RNAseq/A4V.metadata.csv", row.names = 1)

git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"
GR.cpm = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT_cpmdata.csv"),row.names = 1)
GR.meta  = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.metadata.csv"),row.names = 1)
GR.FDR = read.csv(paste0(git.dir,"TKT_RNAseq/CountTables/TKT.EdgeR.FDRTable.csv"),row.names = 1)

bmp = read.csv('/Users/johncsantiago/Documents/GitHub/WhartonLab/BMP Signaling Genes annotated.csv', row.names = 1, header = T)
#bmp = bmp[intersect(row.names(bmp), row.names(GR.cpm)),]

bmp$fb = row.names(bmp)

bmp = bmp[,1:4]

bmp.set = GeneIDKey[GeneIDKey$Symbol %in% setdiff(c("cmpy", "nmo", "Neto", "LIMK1", 'trio',
                                            "spartin", "GluRIIA", 'GluRIIB', 'Rac1',
                                            'spin', 'hiw', 'faf'), bmp$X.1),]
bmp.set = bmp.set[,2:5]
colnames(bmp.set) =  colnames(bmp)
bmp.set[,2] = "Others"
bmp.set[,3] = ""
bmp.set[,4] = ""

bmp=rbind(bmp, bmp.set)

bmpcpm = cpm[intersect(row.names(bmp), row.names(cpm)),]

GR.bmp = GR.cpm[intersect(row.names(GR.cpm), row.names(bmp)),]

GR.mean = matrix(rep(0, 12*nrow(GR.bmp)), ncol = 12)
colnames(GR.mean) = unique(GR.meta$Group)
row.names(GR.mean) = row.names(GR.bmp)

i = 1
while(i <= ncol(GR.mean)){
  GR.mean[,i] = apply(GR.bmp[,row.names(GR.meta)[GR.meta$Group == colnames(GR.mean)[i]]], 1, mean)
  i = i +1
}

GR.mean.meta = GR.meta[GR.meta$Replicate == 1, c(1:3,5)]
row.names(GR.mean.meta) = GR.mean.meta$Group

```

```{r}

GR.FC = data.frame(GR.F = GR.F[row.names(bmp), 'logFC'],
                   GRDf.F = GRDf.F[row.names(bmp), 'logFC'],
                   WTDf.F = WTDf.F[row.names(bmp), 'logFC'],
                   GR.M = GR.M[row.names(bmp), 'logFC'],
                   GRDf.M = GRDf.M[row.names(bmp), 'logFC'],
                   WTDf.M = WTDf.M[row.names(bmp), 'logFC'])
row.names(GR.FC) = bmp[,1]
GR.FC = na.omit(GR.FC)

##scaled = (2^(abs(GR.FC)))*(GR.FC/abs(GR.FC))
scaled = GR.FC
Title = " "
branches=3

allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdBu(5)[5:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = .9,
          main = Title)

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M', 'TktOEWT.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))
#Mhmdata=hmdata

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F', 'TktOEWT.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))
#Fhmdata=hmdata

allscaled = cbind(Fhmdata,Mhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "both",
          Rowv=colordend,
          ##Colv="NA",
          cexRow = .9,
          main = Title)

```

```{r}

A4V.groups = A4V.meta[A4V.meta$Replicate == 1,]
row.names(A4V.groups) = A4V.groups$Group
A4V.groups = A4V.groups[colnames(A4V.meancpm),]

hmgroups = A4V.groups[A4V.groups$Sex == "M",]
hmgroups = hmgroups[c('S3MT', 'S9MT', 'A3MT', 'A9MT'),]

hmdata = A4V.meancpm[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = A4V.groups[A4V.groups$Sex == "F",]
hmgroups = hmgroups[c('S3FT', 'S9FT', 'S40FT', 'A3FT', 'A9FT', 'A40FT'),]

hmdata = A4V.meancpm[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))
#Fhmdata=hmdata

allscaled = cbind(Fhmdata,Mhmdata)

scaled = allscaled[intersect(row.names(bmp), row.names(allscaled)),]
Title = " "
branches=3

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = .9,
          main = Title)

```

```{r}

A4V.groups = A4V.meta[A4V.meta$Replicate == 1,]
row.names(A4V.groups) = A4V.groups$Group
A4V.groups = A4V.groups[colnames(A4V.meancpm),]

hmgroups = A4V.groups[A4V.groups$Sex == "M",]
hmgroups = hmgroups[c('S3MT', 'S9MT', 'A3MT', 'A9MT'),]

hmdata = A4V.meancpm[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = A4V.groups[A4V.groups$Sex == "F",]
hmgroups = hmgroups[c('S3FT', 'S9FT', 'S40FT', 'A3FT', 'A9FT', 'A40FT'),]

hmdata = A4V.meancpm[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))
#Fhmdata=hmdata

allscaled = cbind(Fhmdata)

scaled = allscaled[intersect(row.names(bmp), row.names(allscaled)),]
Title = " "
branches=3

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = .9,
          main = Title)

```

```{r}

A4V.groups = A4V.meta[A4V.meta$Replicate == 1,]
row.names(A4V.groups) = A4V.groups$Group
A4V.groups = A4V.groups[colnames(A4V.meancpm),]

hmgroups = A4V.groups[A4V.groups$Sex == "M",]
hmgroups = hmgroups[c('S3MT', 'S9MT', 'A3MT', 'A9MT'),]

hmdata = A4V.meancpm[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = A4V.groups[A4V.groups$Sex == "F",]
hmgroups = hmgroups[c('S3FT', 'S9FT', 'S40FT', 'A3FT', 'A9FT', 'A40FT'),]

hmdata = A4V.meancpm[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))
#Fhmdata=hmdata

allscaled = cbind(Mhmdata)

scaled = allscaled[intersect(row.names(bmp), row.names(allscaled)),]
Title = " "
branches=3

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = .9,
          main = Title)

```

```{r}

hmgroups = gbbOE.groups[gbbOE.groups$stage == "LL3",]
hmgroups = setNames(unique(hmgroups$group), unique(hmgroups$genotype))

hmdata = gbbOEmean[, hmgroups]
colnames(hmdata) = names(hmgroups)
hmdata=t(scale(t(hmdata)))
hmdata = hmdata[intersect(row.names(bmp), row.names(hmdata)),]
row.names(hmdata) = bmp[row.names(hmdata), 1]

scaled = hmdata
Title = " "
branches=3

allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = .9,
          main = Title)

```

```{r}

gbb.plotdata = data.frame(genotype = gbbOE.groups[colnames(gbbOE.cpmdata), 'group'],
                          gbb = gbbOE.cpmdata["FBgn0037607",])
                          ##gbb = gbbOE.cpmdata["FBgn0024234",])

gbb.plotdata$genotype = factor(gbb.plotdata$genotype,
                               levels = c("LoxP_ML3", "LoxP_LL3",
                                          "G85R_ML3", "G85R_LL3",
                                          "G85R_gbb_ML3", "G85R_gbb_LL3"))

boxplot(gbb.plotdata$gbb ~ gbb.plotdata$genotype, 
        xlab="",     
        main= "",
        ylab="gbb cpm", 
        las = 1, 
        cex.axis = .8)

points(x=as.numeric(factor(gbb.plotdata$genotype)),
       y=gbb.plotdata$gbb,
       cex=2,
       pch=21,
       bg=c('red', 'firebrick',
            'dodgerblue', "steelblue", 
            "limegreen", 'darkgreen')
       [as.numeric(factor(gbb.plotdata$genotype))])

```

```{r}

gbbOE.plotdata = data.frame(GR.FC = G85RxLoxP.LL3[intersect(row.names(bmp), row.names(G85RxLoxP.LL3)), "logFC"],
                            OE.FC = G85R_gbbxLoxP.LL3[intersect(row.names(bmp), row.names(G85RxLoxP.LL3)), "logFC"])
row.names(gbbOE.plotdata) = bmp[intersect(row.names(bmp), row.names(G85RxLoxP.LL3)), 1]

data = data.frame(Symbol = row.names(gbbOE.plotdata),
                  FC1 = gbbOE.plotdata$GR.FC,
                  FC2 = gbbOE.plotdata$OE.FC,
                  Color = 'grey',
                  size = gbbOEmean[intersect(row.names(bmp), row.names(G85RxLoxP.LL3)), 'LoxP_LL3'],
                  LoxP = gbbOEmean[intersect(row.names(bmp), row.names(G85RxLoxP.LL3)), 'LoxP_LL3'])

  main.title = 'FC relative to LoxP'

  data$Color[(data$FC1 - log2(1.5)) > data$FC2] = 'royalblue'
  data$Color[(data$FC2 - log2(1.5)) > data$FC1] = 'firebrick' 
  data$size = log2(data$size)
  data$size[data$size < 2] = 2

fig = plot_ly(data = data,
                x = ~FC1,
                y = ~FC2,
                type = 'scatter',
                mode = 'markers',
                marker = list(color = ~Color,
                              colors = ~Color,
                              line = list(color = "black", width = 1.5),
                              size = ~size),
                hoverinfo = "text",
                hovertext = paste("Symbol:", data$Symbol,
                                  "\n ", colnames(FC.data)[1], "FC: ", round(data$FC1, 2),
                                  "\n ", colnames(FC.data)[2], "FC: ", round(data$FC2, 2),
              "\n LoxP mean cpm: ", round(data$LoxP, 2)))
  fig = fig %>% layout(title = main.title,
                       shapes = list(
                         type = "line",
                         line = list(color = "black", dash = 'dash', width = 2),
                         x0 = -1,
                         x1 = 1,
                         y0 = -1.2,
                         y1 = 1.2),
                       xaxis = list(title = "G85R FC"),
                       yaxis = list(title = "gbbOE FC"))

  fig

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))
#Mhmdata=hmdata

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))
#Fhmdata=hmdata

allscaled = cbind(Mhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = .9,
          main = Title)

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

row.names(Mhmdata) = bmp[row.names(Mhmdata),1]
Mhmdata = Mhmdata[row.names(scaled),]
Mhmdata = Mhmdata[hc$order,]

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

row.names(Fhmdata) = bmp[row.names(Fhmdata),1]
Fhmdata = Fhmdata[row.names(scaled),]
Fhmdata = Fhmdata[hc$order,]

heatmaply(Fhmdata,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "none",
          Rowv= "NA",
          Colv="NA",
          cexRow = .9,
          main = Title)

```

```{r}

hmgroups = gbbOE.groups[gbbOE.groups$stage == "LL3",]
hmgroups = setNames(unique(hmgroups$group), unique(hmgroups$genotype))

hmdata = gbbOEmean[, hmgroups]
colnames(hmdata) = names(hmgroups)
hmdata=t(scale(t(hmdata)))
hmdata = hmdata[intersect(row.names(bmp), row.names(hmdata)),]
row.names(hmdata) = bmp[row.names(hmdata), 1]
hmdata = hmdata[intersect(row.names(Fhmdata),row.names(hmdata)),]

heatmaply(hmdata,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "none",
          Rowv= "NA",
          Colv="NA",
          cexRow = .9,
          main = Title)

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

allscaled = cbind(Fhmdata, Mhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

bmp.set = bmp[row.names(scaled),]

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

bmp.set = bmp.set[bmp.set$X.2 == 'BMP binding',]
scaled = scaled[bmp.set$X.1,]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = 'BMP binding')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

row.names(Mhmdata) = bmp[row.names(Mhmdata),1]
Mhmdata = Mhmdata[row.names(scaled),]
Mhmdata = Mhmdata[hc$order,]

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

row.names(Fhmdata) = bmp[row.names(Fhmdata),1]
Fhmdata = Fhmdata[row.names(scaled),]
Fhmdata = Fhmdata[hc$order,]

heatmaply(Fhmdata[row.names(scaled),],
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "none",
          Rowv= "NA",
          Colv="NA",
          cexRow = .9,
          main = "BMP Binding")

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

row.names(Mhmdata) = bmp[row.names(Mhmdata),1]
Mhmdata = Mhmdata[row.names(scaled),]
Mhmdata = Mhmdata[hc$order,]

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

row.names(Fhmdata) = bmp[row.names(Fhmdata),1]
Fhmdata = Fhmdata[row.names(scaled),]
Fhmdata = Fhmdata[hc$order,]

heatmaply(Mhmdata[row.names(scaled),],
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "none",
          Rowv= "NA",
          Colv="NA",
          cexRow = .9,
          main = 'BMP Binding')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
#Mhmdata=t(scale(t(hmdata)))
Mhmdata=hmdata

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
#Fhmdata=t(scale(t(hmdata)))
Fhmdata=hmdata

allscaled = cbind(Fhmdata, Mhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

bmp.set = bmp[row.names(scaled),]

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

bmp.set = bmp.set[bmp.set$X.2 == 'ligand',]
scaled = scaled[bmp.set$X.1,]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = 'Ligand')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

row.names(Mhmdata) = bmp[row.names(Mhmdata),1]
Mhmdata = Mhmdata[row.names(scaled),]
Mhmdata = Mhmdata[hc$order,]

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

row.names(Fhmdata) = bmp[row.names(Fhmdata),1]
Fhmdata = Fhmdata[row.names(scaled),]
Fhmdata = Fhmdata[hc$order,]

heatmaply(Fhmdata[row.names(scaled),],
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "none",
          Rowv= "NA",
          Colv="NA",
          cexRow = .9,
          main = 'Ligands')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

row.names(Mhmdata) = bmp[row.names(Mhmdata),1]
Mhmdata = Mhmdata[row.names(scaled),]
Mhmdata = Mhmdata[hc$order,]

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

row.names(Fhmdata) = bmp[row.names(Fhmdata),1]
Fhmdata = Fhmdata[row.names(scaled),]
Fhmdata = Fhmdata[hc$order,]

heatmaply(Mhmdata[row.names(scaled),],
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "none",
          Rowv= "NA",
          Colv="NA",
          cexRow = .9,
          main = 'Ligands')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

allscaled = cbind(Fhmdata, Mhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

bmp.set = bmp[row.names(scaled),]

row.names(scaled) = bmp[row.names(scaled),1]

allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

bmp.set = bmp.set[bmp.set$X.2 == 'receptor',]
scaled = scaled[bmp.set$X.1,]
scaled = scaled[row.names(scaled)!='lili',]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = 'Receptor')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

row.names(Mhmdata) = bmp[row.names(Mhmdata),1]
Mhmdata = Mhmdata[row.names(scaled),]
Mhmdata = Mhmdata[hc$order,]

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

row.names(Fhmdata) = bmp[row.names(Fhmdata),1]
Fhmdata = Fhmdata[row.names(scaled),]
Fhmdata = Fhmdata[hc$order,]

heatmaply(Fhmdata[row.names(scaled),],
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "none",
          Rowv= "NA",
          Colv="NA",
          cexRow = .9,
          main = 'Receptors')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

row.names(Mhmdata) = bmp[row.names(Mhmdata),1]
Mhmdata = Mhmdata[row.names(scaled),]
Mhmdata = Mhmdata[hc$order,]

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

row.names(Fhmdata) = bmp[row.names(Fhmdata),1]
Fhmdata = Fhmdata[row.names(scaled),]
Fhmdata = Fhmdata[hc$order,]

heatmaply(Mhmdata[row.names(scaled),],
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "none",
          Rowv= "NA",
          Colv="NA",
          cexRow = .9,
          main = 'Receptors')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

allscaled = cbind(Fhmdata, Mhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

bmp.set = bmp[row.names(scaled),]

row.names(scaled) = bmp[row.names(scaled),1]

allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

bmp.set = c("Mad", "Med", 'smox', 'Dad')
scaled = scaled[bmp.set,]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = 'Smads')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

row.names(Mhmdata) = bmp[row.names(Mhmdata),1]
Mhmdata = Mhmdata[row.names(scaled),]
Mhmdata = Mhmdata[hc$order,]

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

row.names(Fhmdata) = bmp[row.names(Fhmdata),1]
Fhmdata = Fhmdata[row.names(scaled),]
Fhmdata = Fhmdata[hc$order,]

heatmaply(Fhmdata[row.names(scaled),],
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "none",
          Rowv= "NA",
          Colv="NA",
          cexRow = .9,
          main = 'Smads')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

row.names(Mhmdata) = bmp[row.names(Mhmdata),1]
Mhmdata = Mhmdata[row.names(scaled),]
Mhmdata = Mhmdata[hc$order,]

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

row.names(Fhmdata) = bmp[row.names(Fhmdata),1]
Fhmdata = Fhmdata[row.names(scaled),]
Fhmdata = Fhmdata[hc$order,]

heatmaply(Mhmdata[row.names(scaled),],
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "none",
          Rowv= "NA",
          Colv="NA",
          cexRow = .9,
          main = 'Smads')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
#Mhmdata=t(scale(t(hmdata)))
Mhmdata=hmdata

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
#Fhmdata=t(scale(t(hmdata)))
Fhmdata=hmdata

allscaled = cbind(Fhmdata, Mhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

bmp.set = bmp[row.names(scaled),]

row.names(scaled) = bmp[row.names(scaled),1]

allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

bmp.set = c("cmpy", "nmo", "Neto", "LIMK1", 'trio',
            "spartin", "GluRIIA", 'GluRIIB', 'Rac1',
            'spin', 'hiw', 'faf')

scaled = scaled[bmp.set,]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = 'Others')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

row.names(Mhmdata) = bmp[row.names(Mhmdata),1]
Mhmdata = Mhmdata[row.names(scaled),]
Mhmdata = Mhmdata[hc$order,]

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

row.names(Fhmdata) = bmp[row.names(Fhmdata),1]
Fhmdata = Fhmdata[row.names(scaled),]
Fhmdata = Fhmdata[hc$order,]

heatmaply(Fhmdata[row.names(scaled),],
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "none",
          Rowv= "NA",
          Colv="NA",
          cexRow = .9,
          main = 'Others')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

row.names(Mhmdata) = bmp[row.names(Mhmdata),1]
Mhmdata = Mhmdata[row.names(scaled),]
Mhmdata = Mhmdata[hc$order,]

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

row.names(Fhmdata) = bmp[row.names(Fhmdata),1]
Fhmdata = Fhmdata[row.names(scaled),]
Fhmdata = Fhmdata[hc$order,]

heatmaply(Mhmdata[row.names(scaled),],
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "none",
          Rowv= "NA",
          Colv="NA",
          cexRow = .9,
          main = 'Others')

```
```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

allscaled = cbind(Fhmdata, Mhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

bmp.set = bmp[row.names(scaled),]

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

bmp.set = bmp.set[bmp.set$X.2 == 'TF',]
scaled = scaled[bmp.set$X.1,]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = 'Transcription Factor')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

allscaled = cbind(Fhmdata, Mhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

bmp.set = bmp[row.names(scaled),]

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

bmp.set = bmp.set[bmp.set$pathway.regulator == 'negative',]
scaled = scaled[bmp.set$X.1,]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = 'Negative Regulators')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

allscaled = cbind(Fhmdata, Mhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

bmp.set = bmp[row.names(scaled),]

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

bmp.set = bmp.set[bmp.set$pathway.regulator == 'positive',]
scaled = scaled[bmp.set$X.1,]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = 'Positive Regulators')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M', 'TktOEWT.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

allscaled = cbind(Fhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

bmp.set = bmp[row.names(scaled),]

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

bmp.set = bmp.set[bmp.set$pathway.regulator == 'negative',]
scaled = scaled[bmp.set$X.1,]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = 'Negative Regulators')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M', 'TktOEWT.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

allscaled = cbind(Fhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

bmp.set = bmp[row.names(scaled),]

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

bmp.set = bmp.set[bmp.set$pathway.regulator == 'positive',]
scaled = scaled[bmp.set$X.1,]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = 'Positive Regulators')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

allscaled = cbind(Mhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

bmp.set = bmp[row.names(scaled),]

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

bmp.set = bmp.set[bmp.set$pathway.regulator == 'negative',]
scaled = scaled[bmp.set$X.1,]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = 'Negative Regulators')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

allscaled = cbind(Mhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

bmp.set = bmp[row.names(scaled),]

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

bmp.set = bmp.set[bmp.set$pathway.regulator == 'positive',]
scaled = scaled[bmp.set$X.1,]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = 'Positive Regulators')

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M', 'TktOEWT.M'),]
Mhmgroups = hmgroups

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F', 'TktOEWT.F'),]
Fhmgroups = hmgroups

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

allscaled = cbind(Fhmdata,Mhmdata)
allgroups = rbind(Fhmgroups, Mhmgroups)
allgroups$Group = paste(allgroups$Genotype, allgroups$TKT)
allgroups$Labels = paste(row.names(allgroups), allgroups$TKT)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

    use.data   = allscaled
    use.groups = allgroups
    pca <- prcomp(t(use.data), scale.=TRUE) 
    gr <- as.data.frame(use.groups)
    PC1=scale(pca$x[,1])
    PC2=scale(pca$x[,2])
    eigs <- pca$sdev^2
    ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(Sample=row.names(gr),
                        PC1=scale(pca$x[,1]),
                        PC2=scale(pca$x[,2]),
                        sdev=pca$sdev,
                        Genotype=gr$Genotype, 
                        Sex=gr$Sex, 
                        TKT=gr$TKT,
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group=gr$Group,
                        Label = gr$Label)

  fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Label,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Sex,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 colors = c("firebrick","gold","royalblue", 'darkgreen'),
                 marker = list(size = 25,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Sample)) %>%
    layout(xaxis = list(title = paste0("PC1", "(", ve[1], "%)")),
           yaxis = list(title = paste0("PC2", "(", ve[2], "%)")))

fig

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M', 'TktOEWT.M'),]
hmgroups = GR.meta[GR.meta$Group %in% row.names(hmgroups),]

hmdata = GR.bmp[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))
Mhmgroups = hmgroups

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F', 'TktOEWT.F'),]
hmgroups = GR.meta[GR.meta$Group %in% row.names(hmgroups),]

hmdata = GR.bmp[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))
Fhmgroups = hmgroups

allscaled = cbind(Fhmdata,Mhmdata)
allgroups = rbind(Fhmgroups, Mhmgroups)
allgroups$Group = paste(allgroups$Genotype, allgroups$TKT)
allgroups$Labels = paste(allgroups$Genotype, paste(allgroups$Sex, allgroups$TKT))
alldata = GR.bmp[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

    use.data   = allscaled
    use.groups = allgroups
    pca <- prcomp(t(use.data), scale.=TRUE) 
    gr <- as.data.frame(use.groups)
    PC1=scale(pca$x[,1])
    PC2=scale(pca$x[,2])
    eigs <- pca$sdev^2
    ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(Sample=row.names(gr),
                        PC1=scale(pca$x[,1]),
                        PC2=scale(pca$x[,2]),
                        sdev=pca$sdev,
                        Genotype=gr$Genotype, 
                        Sex=gr$Sex, 
                        TKT=gr$TKT,
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group=gr$Group,
                        Label = gr$Label)

  fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Label,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 symbol = ~Sex,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 colors = c("firebrick","gold","royalblue", 'darkgreen'),
                 marker = list(size = 25,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Sample)) %>%
    layout(xaxis = list(title = paste0("PC1", "(", ve[1], "%)")),
           yaxis = list(title = paste0("PC2", "(", ve[2], "%)")))

fig

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

allscaled = cbind(Fhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = Title)

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M', 'TktOEWT.M'),]
hmgroups = GR.meta[GR.meta$Group %in% row.names(hmgroups),]

hmdata = GR.bmp[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F', 'TktOEWT.F'),]
hmgroups = GR.meta[GR.meta$Group %in% row.names(hmgroups),]

hmdata = GR.bmp[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

allscaled = cbind(Fhmdata, Mhmdata)
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = Title)

```

```{r}

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "M",]
hmgroups = hmgroups[c('WT.M', 'GR.M', 'TktOEGR.M'),]

hmdata = GR.mean[, row.names(hmgroups)]
Mhmdata=t(scale(t(hmdata)))

hmgroups = GR.mean.meta[GR.mean.meta$Sex == "F",]
hmgroups = hmgroups[c('WT.F', 'GR.F', 'TktOEGR.F'),]

hmdata = GR.mean[, row.names(hmgroups)]
Fhmdata=t(scale(t(hmdata)))

allscaled = cbind(Mhmdata)
allmean = GR.mean[,colnames(allscaled)]
row.names(allmean) = bmp[row.names(allmean),1]

scaled = allscaled
Title = " "
branches=3

row.names(scaled) = bmp[row.names(scaled),1]
allcolors=c('brown', 'gold3', 'darkgreen', 'cornflowerblue', 'mediumorchid', 'orangered')
colors = allcolors[1:branches]

hc <- hclust(dist(scaled), "complete")
heights=sort(hc$height,decreasing = T)

if(branches>1){
  cutheight=heights[branches]
}

if(branches==1){
  cutheight=max(heights)+1
}

dnd=cut(as.dendrogram(hc),h=cutheight)

if(branches>1){
  branchcolors=colors[1:length(dnd$lower)]
}
if(branches==1){
  branchcolors="black"
}

colordend=(color_branches(hc, h=cutheight, col=branchcolors))

heatmaply(scaled,
          trace="none",
          col=RdYlBu(100)[100:1],
          dendrogram = "row",
          Rowv=colordend,
          Colv="NA",
          cexRow = 1,
          main = Title)

```

```{r}

GeneIDKey = read.csv(paste0(git.dir,"GeneralDataFiles/GeneIDKey.csv"),row.names = 1)
Df.data = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/DeficiencyModifierGenes.csv"),row.names = 1)
Df.key = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/DfKey.csv", row.names = 1)

bmp.df = Df.data[row.names(Df.data) %in% row.names(bmp),]

bmp.df = bmp.df[bmp.df$Repress == 1 | bmp.df$Enhancer == 1,]

bmp.df.info = GeneIDKey[row.names(GeneIDKey) %in% row.names(bmp.df),]

bmp.df.info$MLE = 0
i = 1
while(i<=nrow(bmp.df.info)){
  temp = Df.key[Df.key$Chr == bmp.df.info$CHR[i],]
  temp = temp[(temp$Start < bmp.df.info$Start[i] & temp$End > bmp.df.info$Start[i]) | 
                (temp$Start < bmp.df.info$End[i] & temp$End > bmp.df.info$End[i]) |
                (temp$Start > bmp.df.info$Start[i] & temp$End < bmp.df.info$End[i]),]
  bmp.df.info[i, 'MLE'] = max(as.numeric(temp$MLE))
  i = i + 1
}

bmp.df.info = na.omit(bmp.df.info)
bmp.df.info = bmp.df.info[order(bmp.df.info$MLE),]
bmp.df.info = bmp.df.info[abs(bmp.df.info$MLE)>=2 & abs(bmp.df.info$MLE)<=100, ]
bmp.df.info$color = 'brown'
bmp.df.info$color[bmp.df.info$MLE <= 5] = "orangered3"
bmp.df.info$color[bmp.df.info$MLE <= 3] = "orange3"
bmp.df.info$color[bmp.df.info$MLE <= 2.5] = "orange"
bmp.df.info$color[bmp.df.info$MLE <= 2] = "goldenrod1"
bmp.df.info$color[bmp.df.info$MLE < 0] = "steelblue"

barplot(height = bmp.df.info$MLE,
        names.arg = bmp.df.info$Symbol,
        beside =T,
        horiz = T,
        las = 1,
        cex.names = .55,
        cex.axis = .55,
        col = bmp.df.info$color,
        xlim = c(-4,8))

```

```{r warning=FALSE}

bmp.df.info = GeneIDKey[row.names(GeneIDKey) %in% row.names(bmp),]

bmp.df.info$MLE = 0
i = 1
while(i<=nrow(bmp.df.info)){
  temp = Df.key[Df.key$Chr == bmp.df.info$CHR[i],]
  temp = temp[(temp$Start < bmp.df.info$Start[i] & temp$End > bmp.df.info$Start[i]) | 
                (temp$Start < bmp.df.info$End[i] & temp$End > bmp.df.info$End[i]) |
                (temp$Start > bmp.df.info$Start[i] & temp$End < bmp.df.info$End[i]),]
  bmp.df.info[i, 'MLE'] = max(as.numeric(temp$MLE))
  i = i + 1
}

bmp.df.info = na.omit(bmp.df.info)
bmp.df.info = bmp.df.info[order(bmp.df.info$MLE),]
bmp.df.info = bmp.df.info[abs(bmp.df.info$MLE)<=100, ]
bmp.df.info$color = 'green4'
bmp.df.info$color[bmp.df.info$MLE <= 5] = "limegreen"
bmp.df.info$color[bmp.df.info$MLE <= 3] = "yellowgreen"
bmp.df.info$color[bmp.df.info$MLE <= 2.5] = "gold"
bmp.df.info$color[bmp.df.info$MLE <= 2] = "yellow2"
bmp.df.info$color[bmp.df.info$MLE <= 1.5] = "antiquewhite2"
bmp.df.info$color[bmp.df.info$MLE <= 1] = "antiquewhite"
bmp.df.info$color[bmp.df.info$MLE <= .5] = "white"
bmp.df.info$color[bmp.df.info$MLE < -1] = "antiquewhite"
bmp.df.info$color[bmp.df.info$MLE <= -1.5] = "antiquewhite2"
bmp.df.info$color[bmp.df.info$MLE < -2] = "firebrick"

barplot(height = bmp.df.info$MLE,
        names.arg = bmp.df.info$Symbol,
        beside =T,
        horiz = T,
        las = 1,
        cex.names = .75,
        cex.axis = .75,
        col = bmp.df.info$color,
        xlim = c(-4,8),
        xlab = "MLE")

```