### 2025-01-13 mt.nuc PCA for grant.Rmd

---
title: "mt:nuc PCA for grant"
author: "John Santiago"
date: "2024-11-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, warning = F, include = F}

if(intersect(row.names(installed.packages()), "factoextra") != 'factoextra'){
  # Install
  if(!require(devtools)) install.packages("devtools")
  devtools::install_github("kassambara/survminer")
}

library(devtools)
library(factoextra)

##CPM and other gene info for a subset of OXPHOS genes that met the cpm cutoff
complex.data = read.csv("https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/mt.nuc.OXPHOS_PCAdata.csv", row.names = 1)

##just the cpm data
complex.cpm = complex.data[,5:34]
##scaled the cpm data
complex.cpm = t(scale(t(complex.cpm)))
##pca of the scaled cpm data
pca <- prcomp((complex.cpm), scale.=TRUE)

##other information for each gene besides cpm
gene.info = complex.data[,1:4]
##change non-mt chromosomes to 'nuc' for coloring
gene.info$CHR[gene.info$CHR != "mt"] = 'nuc'

```

```{r echo = F}

fviz_pca_ind(pca, 
             col.ind = gene.info$CHR,
             addEllipses = T, 
             label = "none" )

```