### 2025-01-14 Shortcuts_OmicsQuickPlotFunctions.Rmd
finalized omics functions

---
title: "Specifgic Gene or Metabolite Levels"
author: "John Santiago"
date: "2024-05-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = F, echo = F}
##Load Libraries and Data

##library(plotly)
library(edgeR)

##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/refs/heads/master/GeneralDataFiles/"

GeneIDKey = read.csv(paste0(git.dir, "GeneIDKey.csv"), row.names = 1)

TKT.cpm = read.csv(paste0(git.dir,"TKT_cpmdata.csv"), row.names = 1)

TKT.groups  = read.csv(paste0(git.dir, "TKT.metadata.csv"), row.names = 1)

TKT.groups = TKT.groups[colnames(TKT.cpm),]

A4V.cpm = read.csv(paste0(git.dir, "A4V.cpmdata.csv"), row.names = 1)

gbbOE = read.csv(paste0(git.dir, "gbbOE_CountTable.csv"), row.names = 1)

gbbOE.groups = read.csv(paste0(git.dir, "gbbOE_Metadata.csv"), row.names = 1)

gbbOE = gbbOE[,row.names(gbbOE.groups)]
gbbOE.groups$group = paste0(gbbOE.groups$genotype, "_",gbbOE.groups$stage)

Metab.data = read.csv(paste0(git.dir, "RawMetabolomicData.csv"), row.names = 1)
Metab.Meta = read.csv(paste0(git.dir, "MetabolomicMetadata.csv"), row.names = 1)

KEGG.Key = setNames(Metab.data[-c(1:2), 2], row.names(Metab.data)[-c(1:2)])

norm.func = function(data){
  nd = (as.numeric(data)/Metab.Meta$TIC)*1000
  return(nd)
}

norm.data = t(apply(Metab.data[3:nrow(Metab.data),3:ncol(Metab.data)], 1, norm.func))

colnames(norm.data) = colnames(Metab.data)[3:ncol(Metab.data)]
colnames(norm.data) = Metab.Meta[colnames(norm.data), "Genotypes"]

SCdata = read.csv(paste0(git.dir, "Nguyen%20Serpe%202024%20scRNA-seq%20VNC.csv"))

```

```{r echo = F}
##Function to plot all G85R groups
plot.G85R = function(ID){
ID.row = GeneIDKey[c(grep(ID, GeneIDKey$FBgn),
                     grep(ID, GeneIDKey$Symbol)),]
ID.row = na.omit(ID.row)

if(nrow(ID.row) < 50 & nrow(ID.row) > 0){
if(nrow(ID.row) == 1
   | length(intersect(GeneIDKey$FBgn, ID)) == 1
   | length(intersect(GeneIDKey$Symbol, ID)) == 1){
  if(nrow(ID.row) == 1){
    FBgn = ID.row$FBgn
    name = ID.row$Symbol
  }
  if(length(intersect(GeneIDKey$FBgn, ID)) == 1){
    FBgn = ID.row$FBgn[ID.row$FBgn == ID]
    name = ID.row$Symbol[ID.row$FBgn == ID]
  }
  if(length(intersect(GeneIDKey$Symbol, ID)) == 1){
    FBgn = ID.row$FBgn[ID.row$Symbol == ID]
    name = ID.row$Symbol[ID.row$Symbol == ID]
  }

gene = data.frame(cpm = as.numeric(TKT.cpm[FBgn,]), condition = TKT.groups$Group, group = paste0(TKT.groups$Genotype, TKT.groups$Sex))

gene$group = factor(gene$group, levels = c(unique(gene$group)))
gene$condition = factor(gene$condition, levels = c('WT.F', "TktDfWT.F", "TktOEWT.F",
                                                   'GR.F', "TktDfGR.F", "TktOEGR.F",
                                                   'WT.M', "TktDfWT.M", "TktOEWT.M",
                                                   'GR.M', "TktDfGR.M", "TktOEGR.M"))

if(nrow(na.omit(gene)) > 0){
par(mar = c(5,7,5,2))
plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(min(gene$cpm), max(gene$cpm)),
     xlim = c(.5, 12.5),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = NA)

rect(0, 0-(1.5*max(gene$cpm)), 13, 1.5*max(gene$cpm), col = 'gray95')

boxplot(gene$cpm~gene$condition, 
        xlab="",
        las = 2, 
        cex.axis = .8,
        xaxt = 'none',
        yaxt = 'none',
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        add = T,
        col = c(rep("honeydew", 3),
                rep("thistle1", 3),
                rep("lightcyan", 3),
                rep("lightyellow", 3)))

box(lwd = 2)

points(x=as.numeric(factor(gene$condition)),
       y=gene$cpm,
       cex=1,
       pch=21,
       bg=c('firebrick',"gold","darkgreen","dodgerblue")[as.numeric(factor(gene$group))])

axis(side = 1,
     at = 1:12,
     labels = rep(c("Control", "Df", "OE"), 4),
     line = 0,
     cex.axis = .75,
     lwd.ticks = 2)

axis(side = 2,
     line = 0,
     cex.axis = 1,
     lwd.ticks = 2,
     las = 2)

axis(side = 2,
     at = .5*(max(gene$cpm) + min(gene$cpm)),
     label = "Counts per million reads",
     tick = F,
     padj = -3,
     cex.axis = 1.25)

axis(side = 1,
     at = c(2, 5, 8, 11),
     labels = c("Silent", "G85R","Silent", "G85R"),
     tick = F,
     padj = 2,
     cex.axis = 1)
axis(side = 3,
     at = c(3.5, 9.5),
     labels = c("Female", "Male"),
     tick = F,
     padj = .5,
     cex.axis = 1.5)

axis(side = 3,
     at = 6.5,
     labels = name,
     tick = F,
     padj = -1.5,
     cex.axis = 2)

axis(side = 3,
     at = c(3.5,9.5),
     tick = T,
     outer = T,
     labels = F,
     line = -2.25,
     lwd.ticks = 0)

xline1 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.04
xline2 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.19
lines(x=c(3.5,3.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(6.5,6.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(9.5,9.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
abline(v = 6.5, lty = 2, lwd = 2)
}
}
}

if(nrow(ID.row) != 1
   | length(intersect(GeneIDKey$FBgn, ID)) > 1
   | length(intersect(GeneIDKey$Symbol, ID)) > 1){
  ID.row
}
}

##Function to plot a specific gene expression levels in all A4V groups
plot.A4V = function(ID){
ID.row = GeneIDKey[c(grep(ID, GeneIDKey$FBgn),
                     grep(ID, GeneIDKey$Symbol)),]
ID.row = na.omit(ID.row)

if(nrow(ID.row) < 50 & nrow(ID.row) > 0){
if(nrow(ID.row) == 1
   | length(intersect(GeneIDKey$FBgn, ID)) == 1
   | length(intersect(GeneIDKey$Symbol, ID)) == 1){
  if(nrow(ID.row) == 1){
    FBgn = ID.row$FBgn
    name = ID.row$Symbol
  }
  if(length(intersect(GeneIDKey$FBgn, ID)) == 1){
    FBgn = ID.row$FBgn[ID.row$FBgn == ID]
    name = ID.row$Symbol[ID.row$FBgn == ID]
  }
  if(length(intersect(GeneIDKey$Symbol, ID)) == 1){
    FBgn = ID.row$FBgn[ID.row$Symbol == ID]
    name = ID.row$Symbol[ID.row$Symbol == ID]
  }
  A4V.groups = as.vector(colnames(A4V.cpm))

  A4V.groups = substr(A4V.groups,1, nchar(A4V.groups)-1)

  if(nrow(na.omit(A4V.cpm[FBgn,]))){
  par(mar=c(5.1, 4.1, 4.1, 8.1), xpd=TRUE)
  gene = data.frame(cpm = as.numeric(A4V.cpm[FBgn,]),
                    condition = A4V.groups,
                    group = paste0(substr(A4V.groups, 1,1), substr(A4V.groups, start = nchar(as.vector(A4V.groups))-1,nchar(as.vector(A4V.groups))-1)))

  gene$condition = factor(gene$condition, levels = c("S3FH", "S9FH", "S40FH",
                                                     "S3FT", "S9FT", "S40FT",
                                                     "S3FA", "S9FA", "S40FA",
                                                     "A3FH", "A9FH", "A40FH",
                                                     "A3FT", "A9FT", "A40FT",
                                                     "A3FA", "A9FA", "A40FA",
                                                     "S3MH", "S9MH",
                                                     "A3MT", "A9MT",
                                                     "A3MA", "A9MA",
                                                     "A3MH", "A9MH",
                                                     "S3MT", "S9MT",
                                                     "S3MA", "S9MA"))
  gene$group = factor(gene$group, levels = c("SF", "AF", "SM", "AM"))

  boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
  points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"gold","darkgreen","dodgerblue")[as.numeric(factor(gene$group))])
  legend('topright',
       inset=c(-0.275,0),
       legend = c('A4V F', 'Silent F', 'A4V M', 'Silent M'), 
       fill = c('firebrick',"darkgreen", "gold","dodgerblue"),
       cex = .65,
       bty = 'n',
       pt.cex = .5)
  }
}

if(nrow(ID.row) != 1
   | length(intersect(GeneIDKey$FBgn, ID)) > 1
   | length(intersect(GeneIDKey$Symbol, ID)) > 1){
  ID.row
}
}
}

##Plot levels of a specific metabolite
plot.metab = function(metabID){

ID.row = grep(metabID, row.names(norm.data))

if(length(ID.row) < 50 & length(ID.row) > 0){
if(length(ID.row) == 1 | length(intersect(row.names(norm.data), metabID)) == 1){

  if(length(ID.row) == 1){
    metab = row.names(norm.data)[ID.row]
    name = row.names(norm.data)[ID.row]
  }

  if(length(intersect(row.names(norm.data), metabID)) == 1){
    metab = metabID
    name = metabID
  }

metab = data.frame(data = as.numeric(norm.data[metab,]), 
                   group = colnames(norm.data), 
                   geno = colnames(norm.data))

metab$group = factor(metab$group, levels = c("WT", "TktDfWT", "TktOEWT",
                                             "GR", "TktDfGR", "TktOEGR"))

metab$geno = factor(substr(metab$group, (nchar(as.character(metab$group)) -1), nchar(as.character(metab$group))), 
                     levels = c("WT", "GR"))

par(mar = c(5,7,5,2))
plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(min(na.omit(metab$data)), max(na.omit(metab$data))),
     xlim = c(.5, 6.5),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = NA)

rect(0, 0-(1.5*max(na.omit(metab$data))), 13, 1.5*max(na.omit(metab$data)), col = 'gray95')

boxplot(metab$data~metab$group, 
        xlab="",
        las = 2, 
        cex.axis = .8,
        xaxt = 'none',
        yaxt = 'none',
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        add = T,
        col = c(rep("lightyellow", 3),
                rep("lightgreen", 3)))

box(lwd = 2)

points(x=as.numeric(factor(metab$group)),
       y=metab$data,
       cex=1.25,
       pch=21,
       bg=c('gold3',"darkgreen")[as.numeric(factor(metab$geno))])

axis(side = 1,
     at = 1:12,
     labels = rep(c("Control", "Df", "OE"), 4),
     line = 0,
     cex.axis = .75,
     lwd.ticks = 2)

axis(side = 2,
     line = 0,
     cex.axis = 1,
     lwd.ticks = 2,
     las = 2)

axis(side = 1,
     at = c(2, 5, 8, 11),
     labels = c("Silent", "G85R","Silent", "G85R"),
     tick = F,
     padj = 2,
     cex.axis = 1)

axis(side = 3,
     at = 3.5,
     labels = name,
     tick = F,
     padj = -.5,
     cex.axis = 2)

}

if(length(ID.row) != 1 | length(intersect(row.names(norm.data), metabID)) > 1){
  norm.data[ID.row,]
}

}
}

##Function to plot a specific gene expression levels in all cell types for the SC data
plot.SC = function(gene.name){
ID.row = SCdata[grep(gene.name, SCdata$Gene.name),]

if(nrow(ID.row) < 50 & nrow(ID.row) > 0){
if(nrow(ID.row) == 1 | length(intersect(gene.name, SCdata$Gene.name)) == 1){

  if(length(intersect(gene.name, SCdata$Gene.name)) == 1){
    gene.levels = setNames(as.numeric(ID.row[ID.row$Gene.name == gene.name, colnames(ID.row) != "Gene.name"]), colnames(ID.row)[colnames(ID.row) != "Gene.name"]) 
  }

  if(nrow(ID.row) == 1){
    gene.name = ID.row$Gene.name
    gene.levels = setNames(as.numeric(ID.row[,colnames(ID.row) != "Gene.name"]), colnames(ID.row)[colnames(ID.row) != "Gene.name"]) 
  }

  gene.levels = gene.levels[18:1]
  highlight = rep(0, 18)
  highlight[(gene.levels > 2*(mean(gene.levels)))] = 1
  highlight = gene.levels * highlight

  nochange = gene.levels
  nochange[gene.levels > 1.5] = 0
  nochange = gene.levels * nochange

  label.colors = highlight

  label.colors[label.colors > 0] = "deeppink"
  label.colors[label.colors == 0] = "black"
  label.colors[nochange > 0] = 'steelblue'

  if(1.25*(max(gene.levels)) < 5){
    xmax = 5
  }

  if(1.25*(max(gene.levels)) >= 5){
    xmax = 1.25*(max(gene.levels))
  }

  par(mar=c(5, 12, 4, 6), xpd=TRUE)

  plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(.95, 21),
     xlim = c(0, xmax),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = "Mean UMI Counts",
     bty = "n",
     main = gene.name)

  barplot(height = gene.levels, 
          horiz = T, 
          #las = 2,
          xlim = c(0, xmax),
          names.arg = NA,
          border = T,
          add = T)

    barplot(height = highlight, 
          horiz = T, 
          #las = 2, 
          names.arg = NA,
          border = T,
          col = "deeppink",
          add = T)

    barplot(height = nochange, 
          horiz = T, 
          #las = 2, 
          names.arg = NA,
          border = T,
          col = "lightsteelblue",
          add = T)

  mtext(side = 2,
        las = 2,
        at = .7+(1.2*(0:17)),
        text = names(gene.levels),
        col = label.colors,
        line = 0)

  abline(v = 1,
         lty = 2,
         lwd = 2,
         xpd = F)

  legend(x = .9*xmax,
         y = 23,
         legend = c("high", "mid", "low"),
         fill = c("deeppink", "grey", "lightsteelblue"),
         bty = "n",
         xpd = T)

}

if(nrow(ID.row) != 1 | length(intersect(gene.name, SCdata$Gene.name)) > 1){
  ID.row
}
}
}

##Function to plot a specific gene expression levels in the gbbOE transcriptomic data set
plot.gbbOE = function(ID){

ID.row = GeneIDKey[grep(ID, GeneIDKey$FBgn),]

if(nrow(ID.row) < 500 & nrow(ID.row) > 0){

if(nrow(ID.row) == 1){
FBgn = ID.row$FBgn
name = ID.row$Symbol

gene = data.frame(cpm = as.numeric(gbbOE.cpmdata[FBgn,]), condition = gbbOE.groups$group, group = gbbOE.groups$genotype)

gene$group = factor(gene$group, levels = c(unique(gene$group)))
gene$condition = factor(gene$condition, levels = c('LoxP_ML3', "LoxP_LL3",
                                                 'G85R_ML3', "G85R_LL3",
                                                 'G85R_gbb_ML3', "G85R_gbb_LL3"))

boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"dodgerblue","gold")[as.numeric(factor(gene$group))])

}

if(nrow(ID.row) != 1){
  ID.row
}
}
}

##Function to plot a specific gene expression levels in the gbbKO transcriptomic data set
plot.gbbKO = function(ID){
ID.row = GeneIDKey[grep(ID, GeneIDKey$FBgn),]

if(nrow(ID.row) < 500 & nrow(ID.row) > 0){
if(nrow(ID.row) == 1){
FBgn = ID.row$FBgn
name = ID.row$Symbol

gene = data.frame(cpm = as.numeric(gbbKO.cpmdata[FBgn,]), condition = gbbKO.groups$Genotype)

gene$condition = factor(gene$condition, levels = c('WT', "KO"))

boxplot(gene$cpm~gene$condition, xlab="",     
        main= name,ylab="Counts per Million Reads", las = 2, cex.axis = .8)
points(x=as.numeric(factor(gene$condition)),y=gene$cpm,cex=1,pch=21,bg=c('firebrick',"dodgerblue","gold")[as.numeric(factor(gene$condition))])

}

if(nrow(ID.row) != 1){
  ID.row
}
}
}

```

```{r echo = F}

## Plot gene expression from G85R TKT RNAseq experiment
## ID needs to be the full or partial gene symbol or FBgn (case sensitive)

ID = "ETH"

plot.G85R(ID)

```

```{r echo = F}

## Plot gene expression from A4V RNAseq experiment
## ID needs to be the full or partial gene symbol or FBgn (case sensitive)

ID = "ETH"

plot.A4V(ID)
```

```{r echo = F}

## Plot metabolite levels from the G85R TKT metabolomics experiment
## metabID needs to be the full or partial metabolite name as listed in the table (case sensitive)

metabID = 'NAD'

plot.metab(metabID)

```

```{r echo = F}

## Plot gene expression from Single Cell RNAseq experiment
## gene name needs to be the full or partial gene symbol as listed in the table (case sensitive)

gene.name = 'CCAP'

plot.SC(gene.name)

```

```{r echo = F}
##Plot expression levels for a specific gene in the gbbOE dataset

##ID = "0000018"

plot.gbbOE(ID)

```

```{r echo = F}
##Plot expression levels for a specific gene in the gbb-null dataset

## WT = W1118
## KO = GBB1/GBB2 (both null alleles)
## Aaron and Catherine legend for RNAseq
## Whole larvae

##ID = "0000018"

plot.gbbKO(ID)

```