### 2025-01-28 G85R GOandKEGGTermEnrichment.Rmd
serpe data

---
title: "GO and KEGG Term Enrichment"
author: "John Santiago"
date: "2024-05-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, message = F}

source('/Users/johncsantiago/Documents/GitHub/WhartonLab/GOandKEGGPlottingFunctions.R')

```

```{r echo = F, message = F}

# Conditions
#Geno is WT or GR
#TKT is either "C", "OE", or "Df"
#Sex is "F" or "M"

Geno1 = "GR"
TKT1 = "C"
Sex1 = "F"

Geno2 = "WT"
TKT2 = "C"
Sex2 = "F"

temp.FC = colnames(G85R.FC)
temp1 = temp.FC[grep(Geno1, temp.FC)]

if(Sex1 != Sex2){
  temp1 = temp1[intersect(grep(Sex1, temp1), grep(Sex2, temp1))]
  temp1 = temp1[grep(Geno1)]
}

if(Sex1 == Sex2){
  if(Geno1 == Geno2){
  temp1 = temp1[grep(setdiff(c("WT", "GR"), c(Geno1, Geno2)), temp1, invert = T)]
  }

  if(Geno1 != Geno2){
  temp1 = temp1[intersect(grep(Geno1, temp1), grep(Geno2, temp1))]
  }

  temp1 = temp1[intersect(grep(TKT1, temp1), grep(TKT2, temp1))]

  temp1 = temp1[grep(Sex1, temp1)]
}

FC = G85R.FC[temp1]
FDR = G85R.FC[temp1]

Variable.Condition = "GRFC" 
Control.Condition = "WTFC"
secondary.Variable.Condition = "GRMC"
secondary.Control.Condition = "WTMC"

comparison = paste0(Variable.Condition, 'v', Control.Condition)
secondary.comparison = paste0(secondary.Variable.Condition, 'v', secondary.Control.Condition)

if(length(intersect(comparison, names(A4V.comparekey))) == 1){
  FC = A4V.FC[,c(A4V.comparekey[comparison], A4V.comparekey[secondary.comparison])]
  FDR = A4V.FDR[,c(A4V.comparekey[comparison], A4V.comparekey[secondary.comparison])]
  mean.data = A4V.meancpm[,c(Variable.Condition, Control.Condition, 
                             secondary.Variable.Condition, secondary.Control.Condition)]
  metab.data = 'none'
}

if(length(intersect(comparison, names(G85R.comparekey))) == 1){
  FC = G85R.FC[,c(G85R.comparekey[comparison], G85R.comparekey[secondary.comparison])]
  FDR = G85R.FDR[,c(G85R.comparekey[comparison], G85R.comparekey[secondary.comparison])]
  mean.data = G85R.meancpm[,c(Variable.Condition, Control.Condition, 
                              secondary.Variable.Condition, secondary.Control.Condition)]
  metab.data = G85R.metab.comparekey[c(comparison, secondary.comparison)]
  if(G85R.metab.comparekey[c(comparison)] == G85R.metab.comparekey[c(secondary.comparison)]){
    metab.data = G85R.metab.comparekey[c(comparison)]
  }
}

subset = F

if(subset == T){
  geneset = biomarkers
  FC[setdiff(row.names(FC),geneset),] = 1 
  FDR[setdiff(row.names(FDR),geneset),] = 1 
}

```

```{r echo = F}

##Uses data from primary comparison
##point size is relative to mean expression in variable condition
volcano(FC, FDR, mean.data)

```

```{r echo = F}

##Uses data from primary comparison
##point size is relative to mean expression in variable condition
compare.genes(FC, FDR, mean.data)

```

```{r echo = F}
#library(heatmaply)
#hmdata = mean.data[geneset,]
#row.names(hmdata) = GeneIDKey[geneset, 'Symbol']
#hmdata = na.omit(hmdata)

#heatmaply(hmdata,
#          scale = 'row',
#          Colv = F)

```

```{r echo = F}

##Uses data from primary comparison
##point size is relative to mean expression in variable condition
DEG.venn(FDR, FC)

```

```{r echo = F, include = F}

####Uses sigDEGs from primary comparison
cat.data = Enrichment(FDR)

```

```{r echo = F}

##which ranks of categories to use
##a vector of numbers specifying which ranks to use
##eg. c(1, 5:6) will use the first, fifth and sixth most sig. cats.
##can also just type 'sigs.only' 
usecats=c(1:20)
#usecats = 'sigs.only'

KEGG20(cat.data, FC, FDR, usecats)

```

```{r echo = F, include = T}

##can be the exact name or 5 digit code in quotes
specific.kegg = "00980"

plotkegg.genes(specific.kegg, FC, FDR, mean.data)

```

```{r echo = F, message=F, fig.width=7}

##5 digits from KEGG ID
PathwayID = "00010"

#options include 'primary' or 'both'
use = 'both'

KEGG.diagram(PathwayID, FC, use, metab.data)

```

```{r echo = F, include = T, warnings = F}

##can be the exact name or 5 digit code in quotes
specific.kegg = "00010"

comparekegg.genes(specific.kegg, FC, FDR, mean.data)

```

```{r echo = F}

## BP, CC, or MF
GO.ontology = 'BP'

##can select which categories ot include or say 'sigs.only'
usecats=c(1:20)
#usecats = 'sigs.only'

GO20(cat.data, GO.ontology, FC, FDR, usecats)

```

```{r echo = F}

##the 7 digits of a GOterm id or the exact term name
specific.go = '0005811'

plotgo.genes(specific.go, FC, FDR, mean.data)

```

```{r echo = F, warning = F}

##the 7 digits of a GOterm id or the exact term name
specific.go = '0005811'

comparego.genes(specific.go, FC, FDR, mean.data)

```

```{r echo = F, warning = F, message = F}

## BP, CC, or MF
GO.ontology = 'BP'

rt = tree(cat.data, GO.ontology)

treemapPlot(rt)

```

```{r echo = F, warning = F}

##a parent category generated in the treemapPlot function
parent.category = 'membrane lipid biosynthetic process'

plotparentcat.genes(parent.category, FC, FDR, mean.data)

```

```{r echo = F}

##a parent category generated in the treemapPlot function
##parent.category = 'small molecule biosynthetic process'

compareparent.genes(parent.category, FC, FDR, mean.data)
##dot size is relative to the max FDR of either condition

```