### 2025-01-28 SerpeSCdata.Rmd

---
title: "Serpe SC data"
author: "John Santiago"
date: "2025-01-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F}

#library("hdf5r")
#library(SCopeLoomR)

#loom <- open_loom("/Users/johncsantiago/Downloads/5a3ba000_20180809__Davie_Janssens_Koldere_et_al_2018_Adult_Brain.HARMONY_SCENIC.loom", mode = "r")

##loom = open_loom("/Users/johncsantiago/Downloads/Goodwin_Fly_AdultVNC_elife54074.loom", mode = "r")

#dgem <- get_dgem(loom)
#close_loom(loom)
#dgem[1:5,100:105]

#clustering_annotations_df <- get_clustering_annotations(loom = loom)

```

```{r echo = F}

##manually currated
serpe = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Nguyen Serpe 2024 scRNA-seq VNC.csv", row.names = 1)

serpe.glia = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Serpe 2024 GliaSubtypes.csv", row.names = 1)

serpe.mn = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Serpe 2024 MNClusters.csv", row.names = 1)

serpe.inter = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Serpe 2024 InterneuronClusters.csv", row.names = 1)

serpe.annotated.names = as.character(read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Nguyen Serpe 2024 scRNA-seq VNC.csv", row.names = 1, header = F)[1,])

serpe.glia.names = as.character(read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Serpe 2024 GliaSubtypes.csv", row.names = 1, header = F)[1,])

```

```{r echo = F}

plot.serpe = function(gene.name){
serpe.annotated.gene = data.frame(level = as.numeric(serpe[gene.name,]),
                                  cluster = serpe.annotated.names,
                                  color = "darkgreen")

serpe.mn.gene = data.frame(level = as.numeric(serpe.mn[gene.name,]),
                           cluster = colnames(serpe.mn),
                           color = "brown")

serpe.inter.gene = data.frame(level = as.numeric(serpe.inter[gene.name,]),
                              cluster = colnames(serpe.inter),
                              color = "gold")

serpe.glia.gene = data.frame(level = as.numeric(serpe.glia[gene.name,]),
                             cluster = serpe.glia.names,
                             color = "steelblue")

serpe.all = rbind(serpe.annotated.gene,
                  rbind(serpe.mn.gene,
                        rbind(serpe.inter.gene,
                              serpe.glia.gene)))

par(mar = c(4, 13, 4, 8))
barplot(as.numeric(serpe.all[,"level"])-1,
        names.arg = serpe.all$cluster,
        las = 2, 
        beside = T,
        col = serpe.all$color,
        cex.names = .75,
        main = gene.name,
        horiz = T,
        xlab = "Mean Expression Level per Cell")
      legend(x = max(as.numeric(serpe.all[,"level"])-1),
             y = 125,
             pch = 21,
             legend = c("Glia", "Interneurons", "Motor Neurons", "Manually Curated"),
             pt.bg = c("steelblue", "gold", "brown", "darkgreen"),
             bty = "n",
             xpd = T,
             pt.cex = 1.5,
             y.intersp = 1)
}
```

```{r echo = F, fig.width= 10, fig.height=15}

##gene.name = "gbb"

##plot.serpe(gene.name)

```

```{r echo = F}

##Function to plot a specific gene expression levels in all cell types for the SC data
plot.serpe.allclusters = function(gene.name){

  ID.row = serpe[grep(gene.name, row.names(serpe)),]

  if(nrow(ID.row) < 50 & nrow(ID.row) > 0){
    if(nrow(ID.row) == 1 | length(intersect(gene.name, row.names(serpe))) == 1){

  serpe.annotated.gene = data.frame(level = as.numeric(serpe[gene.name,]),
                                    cluster = serpe.annotated.names,
                                    color = "darkgreen")

  serpe.mn.gene = data.frame(level = as.numeric(serpe.mn[gene.name,]),
                            cluster = colnames(serpe.mn),
                            color = "brown")

  serpe.inter.gene = data.frame(level = as.numeric(serpe.inter[gene.name,]),
                                cluster = colnames(serpe.inter),
                                color = "gold")

  serpe.glia.gene = data.frame(level = as.numeric(serpe.glia[gene.name,]),
                              cluster = serpe.glia.names,
                              color = "steelblue")

  serpe.all = rbind(serpe.annotated.gene,
                    rbind(serpe.mn.gene,
                          rbind(serpe.inter.gene,
                                serpe.glia.gene)))

  #serpe.all$level = serpe.all$level - 1

  #highlight = rep(0, nrow(serpe.all))
  #highlight[(serpe.all$level > highlight.cutoff*(mean(serpe.all$level)))] = 1
  #serpe.all$highlight = serpe.all$level * highlight

  #serpe.all$color[highlight > 0 & serpe.all$color == "lightgreen"] = "darkgreen"
  #serpe.all$color[highlight > 0 & serpe.all$color == "thistle1"] = "brown"
  #serpe.all$color[highlight > 0 & serpe.all$color == "lightyellow"] = "gold"
  #serpe.all$color[highlight > 0 & serpe.all$color == "lightcyan"] = "royalblue"
  serpe.all$color[serpe.all$level < 1.5] = "darkgrey"

  #label.colors = serpe.all$color

  #label.colors[serpe.all$color == "darkgreen"] = "darkgreen"
  #label.colors[serpe.all$color == "brown"] = "brown"
  #label.colors[serpe.all$color == "gold"] = "gold"
  #label.colors[serpe.all$color == "royalblue"] = "royalblue"
  #serpe.all$label.colors = label.colors

  xmax = max(serpe.all$level)
  if(xmax < 10){
    xmax = 10
  }

      par(mar = c(4, 13, 4, 8), xpd=TRUE)

      plot(x = NA,
           y = NA,
           type = 'n',
           ylim = c(4.5, 120),
           xlim = c(0, xmax),
           xaxt = "n",
           ylab="",
           yaxt = "n",
           xlab = "Mean Expression Levels",
           bty = "n",
           main = gene.name)

      barplot(height = serpe.all$level, 
              horiz = T, 
              #las = 2,
              xlim = c(0, xmax),
              names.arg = NA,
              border = T,
              col = serpe.all$color,
              add = T)

      mtext(side = 2,
            las = 2,
            at = .7+(1.2*(0:104)),
            text = serpe.all$cluster,
            col = serpe.all$color,
            line = 0)

      legend(x = xmax,
             y = 125,
             pch = 21,
             legend = c("Glia", "Interneurons", "Motor Neurons", "Manually Curated"),
             pt.bg = c("steelblue", "gold", "brown", "darkgreen"),
             bty = "n",
             xpd = T,
             pt.cex = 1.5,
             y.intersp = 1)

      rect(0 - (xmax/40), 0, 0 - (xmax/100), 18*1.2, col = 'darkgreen')
      rect(0 - (xmax/40), 18*1.2, 0 - (xmax/100), 46*1.2, col = 'brown')
      rect(0 - (xmax/40), 46*1.2, 0 - (xmax/100), 95*1.2, col = 'gold')
      rect(0 - (xmax/40), 95*1.2, 0 - (xmax/100), 105*1.2, col = 'steelblue')

    }

    #if(nrow(ID.row) != 1 | length(intersect(gene.name, row.names(serpe))) > 1){
    #  ID.row
    #}
  }
}

```

```{r echo = F, fig.width= 10, fig.height=16}

gene.name = "gbb"

plot.serpe.allclusters(gene.name)

```

```{r echo = F, fig.width= 10, fig.height=16}

gene.name = "dpp"

plot.serpe.allclusters(gene.name)

```

```{r echo = F, fig.width= 10, fig.height=16}

gene.name = "wit"

plot.serpe.allclusters(gene.name)

```

```{r echo = F, fig.width= 10, fig.height=16}

gene.name = "mav"

plot.serpe.allclusters(gene.name)

```

```{r echo = F, fig.width= 10, fig.height=16}

gene.name = "put"

plot.serpe.allclusters(gene.name)

```