### 2025-01-28 TKT_RNAseq/R Code/TKTRNAseqPNPAssessment.Rmd

---
title: 'TKT RNAseq: PNP analysis'
author: "John Santiago"
date: "2024-11-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include = F}

library(visNetwork)
library(org.Dm.eg.db)
library(KEGGREST)
keggpath = as.list(org.Dm.egPATH2EG)
nodeID = as.list(org.Dm.egENZYME)
nodeID['40528'] = '3.5.4.3'
nodeID['41318'] = '2.7.4.10'
nodeID['42273'] = '2.7.1.145'
nodeID['33924'] = '3.1.3.89'
nodeID[['31309']] = c("3.1.4.17", '3.1.4.53')
nodeID['37741'] = '3.1.4.53'

source('/Users/johncsantiago/Documents/GitHub/WhartonLab/GOandKEGGPlottingFunctions.R')

mean.cpm = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/CountTables/TKT_meancpmdata.csv", row.names = 1)
cpmdata = G85R.cpm
groups = G85R.meta

purine.path = data.frame(enzyme = c('1.17.1.4', '1.17.3.2'),
                         from = c('xanthine', 'xanthine'),
                         to = c(rep('uric acid',2)))

xanthine = data.frame(enzyme = c('1.17.1.4', '1.17.3.2', '2.4.2.1', '3.5.4.3'),
                      from = c('hypoxanthine', 'hypoxanthine', 'xanthosine', 'guanine'),
                      to = c(rep('xanthine',4)))
purine.path = rbind(purine.path, xanthine)

hypoxanthine = data.frame(enzyme = c('2.4.2.1', '2.4.2.1'),
                          from = c('deoxyinosine', 'inosine'),
                          to = c(rep('hypoxanthine', 2)))
purine.path = rbind(purine.path, hypoxanthine)

xanthosine = data.frame(enzyme = c('3.1.3.5'),
                        from = c('XMP'),
                        to = 'xanthosine')
purine.path = rbind(purine.path, xanthosine)

XMP = data.frame(enzyme = c('1.1.1.205'),
                 from = c('IMP'),
                 to = 'XMP')
purine.path = rbind(purine.path, XMP)

IMP = data.frame(enzyme = c('3.6.1.6', '3.5.4.6'),
                 from = c('IDP', 'AMP'),
                 to = c(rep('IMP', 2)))
purine.path = rbind(purine.path, IMP)

adenylosuccinate = data.frame(enzyme = c('6.3.4.4'),
                              from = c('IMP'),
                              to = 'adenylosuccinate')
purine.path = rbind(purine.path, adenylosuccinate)

inosine = data.frame(enzyme = c('2.4.2.1', '3.5.4.4', '3.1.3.5'),
                     from = c('hypoxanthine', 'adenosine', 'IMP'),
                     to = c(rep('inosine', 3)))
purine.path = rbind(purine.path, inosine)

deoxyinosine = data.frame(enzyme = c('2.4.2.1', '3.5.4.4'),
                          from = c('hypoxanthine', 'deoxyadenosine'),
                          to = c(rep('deoxyinosine', 2)))
purine.path = rbind(purine.path, deoxyinosine)

adenosine = data.frame(enzyme = c('2.4.2.1', '3.1.3.5', '2.7.1.20'),
                       from = c('adenine', 'AMP', 'AMP'),
                       to = c(rep('adenosine', 3)))
purine.path = rbind(purine.path, adenosine)

AMP = data.frame(enzyme = c('4.3.2.2', '2.7.1.20', '2.4.2.7', '2.7.4.3', '2.7.4.10', '3.1.4.17', '3.1.4.53'),
                 from = c('adenylosuccinate', 'adenosine', 'adenine', 'ADP', 'ADP', 'cAMP', 'cAMP'),
                 to = c(rep('AMP', 7)))
purine.path = rbind(purine.path, AMP)

adenine = data.frame(enzyme = c('2.4.2.1', '2.4.2.7', '2.4.2.1'),
                     from = c('adenosine', 'AMP', 'deoxyadenosine'),
                     to = c(rep('adenine', 3)))
purine.path = rbind(purine.path, adenine)

ATP = data.frame(enzyme = c('2.7.4.6'),
                 from = c('ADP'),
                 to = c(rep('ATP', 1)))
purine.path = rbind(purine.path, ATP)

dATP = data.frame(enzyme = c('2.7.4.6'),
                  from = c('dADP'),
                  to = c(rep('dATP', 1)))
purine.path = rbind(purine.path, dATP)

ADP = data.frame(enzyme = c('2.7.4.3', '2.7.4.10', '2.7.4.6', '3.6.1.15'),
                 from = c('AMP', 'AMP', 'ATP', 'ATP'),
                 to = c(rep('ADP', 4)))
purine.path = rbind(purine.path, ADP)

dADP = data.frame(enzyme = c('2.7.4.3', '1.17.4.1', '2.7.4.6'),
                  from = c('dAMP', 'ADP', 'dATP'),
                  to = c(rep('dADP', 3)))
purine.path = rbind(purine.path, dADP)

dAMP = data.frame(enzyme = c('2.7.1.145', '2.7.4.3'),
                  from = c('deoxyadenosine', 'dADP'),
                  to = c(rep('dAMP', 2)))
purine.path = rbind(purine.path, dAMP)

deoxyadenosine = data.frame(enzyme = c('2.4.2.1', '3.1.3.5', '3.1.3.89'),
                            from = c('adenine', 'dAMP', 'dAMP'),
                            to = c(rep('deoxyadenosine', 3)))
purine.path = rbind(purine.path, deoxyadenosine)

dGTP = data.frame(enzyme = c('2.7.4.6'),
                  from = c('dGDP'),
                  to = c('dGTP'))
purine.path = rbind(purine.path, dGTP)

guanine = data.frame(enzyme = c('2.4.2.1', '2.4.2.1'),
                     from = c('guanosine', 'deoxyguanosine'),
                     to = c(rep('guanine', 2)))
purine.path = rbind(purine.path, guanine)

dGDP = data.frame(enzyme = c('2.7.4.6', '1.17.4.1', '2.7.4.8'),
                  from = c('dGTP', 'GDP', 'dGMP'),
                  to = c(rep('dGDP', 3)))
purine.path = rbind(purine.path, dGDP)

guanosine = data.frame(enzyme = c('2.4.2.1', '3.1.3.5'),
                       from = c('guanine', 'GMP'),
                       to = c(rep('guanosine', 2)))
purine.path = rbind(purine.path, guanosine)

GTP = data.frame(enzyme = c('2.7.4.6'),
                 from = c('GDP'),
                 to = c(rep('GTP', 1)))
purine.path = rbind(purine.path, GTP)

GDP = data.frame(enzyme = c('2.7.4.6', '2.7.4.8'),
                 from = c('GTP', 'GMP'),
                 to = c(rep('GDP', 2)))
purine.path = rbind(purine.path, GDP)

GMP = data.frame(enzyme = c('6.3.5.2','3.6.1.6', '2.7.4.8'),
                 from = c('XMP', 'GDP', 'GDP'),
                 to = c(rep('GMP', 3)))
purine.path = rbind(purine.path, GMP)

dGMP = data.frame(enzyme = c('2.7.4.8', '2.7.1.145'),
                  from = c('dGDP', 'deoxyguanosine'),
                  to = c(rep('dGMP', 2)))
purine.path = rbind(purine.path, dGMP)

cAMP = data.frame(enzyme = '4.6.1.1',
                  from = 'ATP',
                  to = 'cAMP')
purine.path = rbind(purine.path, cAMP)

FGAM = data.frame(c(),
               c())

purine.path$arrows = 'to'
temp = purine.path
temp[,5] = paste0(purine.path[,1], purine.path[,2], purine.path[,3])
temp[,6] = paste0(purine.path[,1], purine.path[,3], purine.path[,2])
temp$keep = TRUE

i=1
while(i<=nrow(temp)){
  if(length(intersect(temp[i,6], temp[temp$keep==TRUE,5]))==1){
        temp$keep[i] = FALSE
        temp[temp[,5] == temp[i,6], 'arrows'] = 'to;from'
  }
  i=i+1
}

purine.path = temp[temp$keep, 1:4]
purine.path$FBgn = ''

temp.purine.path = purine.path

i=1
while(i<=nrow(purine.path)){  
  enzymeID = purine.path$enzyme[i]
  temp = names(nodeID[grep(enzymeID, nodeID)])
  temp = setNames(unlist(nodeID[temp], use.names = F), rep(names(nodeID[temp]), lengths(nodeID[temp])))
  temp = names(temp[temp == enzymeID])
  temp = row.names(GeneIDKey[GeneIDKey$ensembl %in% temp, ])
  temp2 = purine.path[rep(i, length(temp)),]
  temp2$FBgn = temp
  temp.purine.path = rbind(temp.purine.path,temp2)
  i = i +1
}

purine.path = temp.purine.path[(nrow(purine.path)+1):nrow(temp.purine.path),]
purine.path = purine.path[!duplicated(purine.path[,2:5]),]

TerminalPoints = data.frame(enzyme = "Excretion",
                            from = 'uric acid',
                            to = 'Excretion')

DNA.A = data.frame(enzyme = c('DNA.A'),
                 from = c('dATP'),
                 to = c('DNA.A'))
TerminalPoints = rbind(TerminalPoints, DNA.A)

DNA.G = data.frame(enzyme = c('DNA.G'),
                 from = c('dGTP'),
                 to = c('DNA.G'))
TerminalPoints = rbind(TerminalPoints, DNA.G)

RNA.A = data.frame(enzyme = c('RNA.A'),
                 from = c('ATP'),
                 to = c('RNA.A'))
TerminalPoints = rbind(TerminalPoints, RNA.A)

RNA.G = data.frame(enzyme = c('RNA.G'),
                 from = c('GTP'),
                 to = c('RNA.G'))
TerminalPoints = rbind(TerminalPoints, RNA.G)

Riboflavin.Metabolism = data.frame(enzyme = c('Riboflavin.Metabolism'),
                 from = c('GTP'),
                 to = c('Riboflavin.Metabolism'))
TerminalPoints = rbind(TerminalPoints, Riboflavin.Metabolism)

Folate.Biosynthesis = data.frame(enzyme = c('Folate.Biosynthesis'),
                 from = c('GTP'),
                 to = c('Folate.Biosynthesis'))
TerminalPoints = rbind(TerminalPoints, Folate.Biosynthesis)

PPP = data.frame(enzyme = c('Pentose Phosphate'),
                 from = c('Pentose Phosphate'),
                 to = c('IMP'))
TerminalPoints = rbind(TerminalPoints, PPP)

TerminalPoints = data.frame(enzyme = TerminalPoints$enzyme,
                            from = TerminalPoints$from,
                            to = TerminalPoints$to,
                            arrows = 'to',
                            FBgn = NA)

purine.path = rbind(purine.path, TerminalPoints)
row.names(purine.path) = 1:nrow(purine.path)
#write.csv(purine.path, "/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.path.csv")

temp = keggGet("dme:CG16758" )[[1]]

temp = keggGet(names(temp$ORTHOLOGY))[[1]]

temp = names(temp$REACTION)

equation = as.character()
reaction = as.character()

i=1
while(i<= length(temp)){
  equation[i] = keggGet(temp[i])[[1]]["EQUATION"][[1]]
  reaction[i] = keggGet(temp[i])[[1]]["DEFINITION"][[1]]
  i=i+1
}

equation.length = lengths(strsplit(equation, " "))

interactions = matrix("",
                      nrow = length(equation),
                      ncol = max(equation.length))

i=1
while(i<= length(equation)){
  interactions[i, 1:equation.length[i]] = strsplit(equation[i], " ")[[1]]
  i=i+1
}

unique.metabs = unique(as.character(interactions))
unique.metabs = unique.metabs[grep("C", unique.metabs)]

i = 1
while(i <= length(unique.metabs)){
  if(length(metab.id[grep(unique.metabs[i], names(metab.id))])>0){
    interactions[interactions == unique.metabs[i]] = metab.id[grep(unique.metabs[i], names(metab.id))]
    }
  i=i+1
}

interactions = interactions[c(1:5, 7:9), c(1,4,5)]

purine.path = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.path.csv", row.names = 1)

temp.edges = data.frame(from = purine.path$from,
                        to = purine.path$to,
                        arrows = purine.path$arrows,
                        color = "black",
                        width = 0,
                        title = "",
                        FBgn = purine.path$FBgn)

add.edges = data.frame(enzyme = c("3.5.4.10", 
                                "2.1.2.3",
                                "4.3.2.2",
                                "6.3.2.6",
                                "4.1.1.21",
                                "6.3.3.1",
                                "6.3.5.3",
                                "2.1.2.2",
                                "6.3.4.13",
                                "2.4.2.14",
                                "2.7.6.1"),
                       from = c("FAICAR", 
                                "AICA ribonucleotide", 
                                "SAICAR",
                                "CAIR",
                                "AIR",
                                "2-(Formamido)-N1-(5-phospho-D-ribosyl)acetamidine",
                                "FGAR",
                                "GAR",
                                "Ribosylamine-5p",
                                "PRPP",
                                "D-ribose-1/5-phosphate / D-ribulose 5-phosphate"),

                       to = c("IMP", 
                              "FAICAR", 
                              "AICA ribonucleotide",
                              "SAICAR",
                              "CAIR",
                              "AIR",
                              "2-(Formamido)-N1-(5-phospho-D-ribosyl)acetamidine",
                              "FGAR",
                              "GAR",
                              "Ribosylamine-5p",
                              "PRPP"),
                       arrows = c("to;from", 
                                  "to", 
                                  "to;from",
                                  "to;from",
                                  "to;from",
                                  "to",
                                  "to",
                                  "to",
                                  "to",
                                  "to",
                                  "to"),
                       FBgn = "")

#write.csv(temp.edges, "/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.edges.csv")

i = 1
new.edges = data.frame(enzyme = as.character(),
                       from = as.character(),
                       to = as.character(),
                       arrows = as.character(),
                       FBgn = as.character())

while(i <= nrow(add.edges)){
  temp.fbgn = names(nodeID[grep(add.edges[i, "enzyme"], nodeID)])
  temp.fbgn = GeneIDKey[GeneIDKey$ensembl %in% temp.fbgn, "FBgn"]
  temp.new.edges = add.edges[rep(i, length(temp.fbgn)), ]
  temp.new.edges$FBgn = temp.fbgn
  new.edges = rbind(new.edges, temp.new.edges)
  i = i +1
}

purine.path = rbind(purine.path, new.edges)
row.names(purine.path) = 1:nrow(purine.path)
#write.csv(purine.path, "/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.path.csv")

temp.edges = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.edges.csv", row.names = 1)

new.edges$color = "black"
new.edges$width = 0
new.edges$title = ""

new.edges = new.edges[, colnames(temp.edges)]

temp.edges = rbind(temp.edges, new.edges)

#write.csv(temp.edges, "/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.edges.csv")

```

```{r echo = F}

plot(x = G85R.FC$GRxWT.FC,
     y= -log2(G85R.FDR$GRxWT.FC),
     pch = 21,
     bg = 'lightgrey',
     cex = 1.5,
     ylab = "-log2(FDR)",
     xlab = "log2(Fold Change)",
     main = "G85R Females/Silent Females")

points(x = G85R.FC[GeneIDKey[GeneIDKey$Symbol == "CG16758", "FBgn"], 'GRxWT.FC'],
       y= -log2(G85R.FDR[GeneIDKey[GeneIDKey$Symbol == "CG16758", "FBgn"], 'GRxWT.FC']),
       pch = 21,
       bg = 'deeppink',
       cex = 1.5)

```

```{r echo = F}

plot(x = G85R.FC$GRxWT.MC,
     y= -log2(G85R.FDR$GRxWT.MC),
     pch = 21,
     bg = 'lightgrey',
     cex = 1.5,
     ylab = "-log2(FDR)",
     xlab = "log2(Fold Change)",
     main = "G85R Males/Silent Males")

points(x = G85R.FC[GeneIDKey[GeneIDKey$Symbol == "CG16758", "FBgn"], 'GRxWT.MC'],
       y= -log2(G85R.FDR[GeneIDKey[GeneIDKey$Symbol == "CG16758", "FBgn"], 'GRxWT.MC']),
       pch = 21,
       bg = 'cyan',
       cex = 1.5)

```

```{r echo = F}

plot(x = log2(G85R.metabFC$GRxWT.C),
     y= -log2(G85R.metabFDR$GRxWT.C),
     pch = 21,
     bg = 'lightgrey',
     cex = 1.5,
     ylab = "-log2(FDR)",
     xlab = "log2(Fold Change)",
     main = "G85R/Silent Metabolites")

points(x = log2(G85R.metabFC['uric acid', 'GRxWT.C']),
       y= -log2(G85R.metabFDR['uric acid', 'GRxWT.C']),
       pch = 21,
       bg = 'gold',
       cex = 1.5)

points(x = log2(G85R.metabFC['uric acid', 'GRxWT.C']),
       y= -log2(G85R.metabFDR['xanthine', 'GRxWT.C']),
       pch = 21,
       bg = 'orange',
       cex = 1.5)

```

```{r echo = F}

FBgn = 'FBgn0035348'
name = 'PNP (CG16758)'

gene = data.frame(cpm = as.numeric(cpmdata[FBgn,]), condition = groups$Group, group = paste0(groups$Genotype, groups$Sex))

gene$group = factor(gene$group, levels = c(unique(gene$group)))
gene$condition = factor(gene$condition, levels = c('WT.F', "TktDfWT.F", "TktOEWT.F",
                                                   'GR.F', "TktDfGR.F", "TktOEGR.F",
                                                   'WT.M', "TktDfWT.M", "TktOEWT.M",
                                                   'GR.M', "TktDfGR.M", "TktOEGR.M"))

par(mar = c(5,7,5,2))
plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(min(gene$cpm), max(gene$cpm)),
     xlim = c(.5, 12.5),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = NA)

rect(0, 0-(1.5*max(gene$cpm)), 13, 1.5*max(gene$cpm), col = 'gray95')

boxplot(gene$cpm~gene$condition, 
        xlab="",
        las = 2, 
        cex.axis = .8,
        xaxt = 'none',
        yaxt = 'none',
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        add = T,
        col = c(rep("honeydew", 3),
                rep("thistle1", 3),
                rep("lightcyan", 3),
                rep("lightyellow", 3)))

box(lwd = 2)

points(x=as.numeric(factor(gene$condition)),
       y=gene$cpm,
       cex=1,
       pch=21,
       bg=c('firebrick',"gold","darkgreen","dodgerblue")[as.numeric(factor(gene$group))])

axis(side = 1,
     at = 1:12,
     labels = rep(c("Control", "Df", "OE"), 4),
     line = 0,
     cex.axis = .75,
     lwd.ticks = 2)

axis(side = 2,
     line = 0,
     cex.axis = 1,
     lwd.ticks = 2,
     las = 2)

axis(side = 2,
     at = .5*(max(gene$cpm) + min(gene$cpm)),
     label = "Counts per million reads",
     tick = F,
     padj = -3,
     cex.axis = 1.25)

axis(side = 1,
     at = c(2, 5, 8, 11),
     labels = c("Silent", "G85R","Silent", "G85R"),
     tick = F,
     padj = 2,
     cex.axis = 1)
axis(side = 3,
     at = c(3.5, 9.5),
     labels = c("Female", "Male"),
     tick = F,
     padj = .5,
     cex.axis = 1.5)

axis(side = 3,
     at = 6.5,
     labels = name,
     tick = F,
     padj = -1.5,
     cex.axis = 2)

axis(side = 3,
     at = c(3.5,9.5),
     tick = T,
     outer = T,
     labels = F,
     line = -2.25,
     lwd.ticks = 0)

xline1 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.04
xline2 = min(gene$cpm) - (max(gene$cpm)-min(gene$cpm))*.19
lines(x=c(3.5,3.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(6.5,6.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
lines(x=c(9.5,9.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)
abline(v = 6.5, lty = 2, lwd = 2)

```

```{r echo = F}

name = 'uric acid'

gene = data.frame(level = as.numeric(G85R.metabnorm[name,]),
                  group = c(substr(colnames(G85R.metabnorm), 1, nchar(colnames(G85R.metabnorm))-1)))

gene$group = substr(colnames(G85R.metabnorm), 1, nchar(colnames(G85R.metabnorm))-1)

gene$group = factor(gene$group, levels = c("WT", "TktDfWT", "TktOEWT",
                                           "GR", "TktDfGR", "TktOEGR"))

par(mar = c(5,7,5,2))
plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(min(gene$level), max(gene$level)),
     xlim = c(.5, 6.5),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = NA)

rect(0, 0-(1.5*max(gene$level)), 13, 1.5*max(gene$level), col = 'gray95')

boxplot(gene$level~gene$group, 
        xlab="",
        las = 2, 
        cex.axis = .8,
        xaxt = 'none',
        yaxt = 'none',
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        add = T,
        col = c(rep("lightcyan", 3),
                rep("thistle1", 3)))

box(lwd = 2)

points(x=as.numeric(factor(gene$group)),
       y=gene$level,
       cex=1,
       pch=21,
       bg=c(rep("dodgerblue", 3), rep('firebrick', 3))[as.numeric(factor(gene$group))])

axis(side = 1,
     at = 1:6,
     labels = rep(c("Control", "Df", "OE"), 2),
     line = 0,
     cex.axis = .75,
     lwd.ticks = 2)

axis(side = 2,
     line = 0,
     cex.axis = 1,
     lwd.ticks = 2,
     las = 2)

axis(side = 2,
     at = .5*(max(gene$level) + min(gene$level)),
     label = "Normalized Metabolite Level",
     tick = F,
     padj = -4,
     cex.axis = 1.25)

axis(side = 1,
     at = c(2, 5),
     labels = c("Silent", "G85R"),
     tick = F,
     padj = 2,
     cex.axis = 1)

axis(side = 3,
     at = 3.5,
     labels = name,
     tick = F,
     padj = -1.5,
     cex.axis = 2)

xline1 = min(gene$level) - (max(gene$level)-min(gene$level))*.04
xline2 = min(gene$level) - (max(gene$level)-min(gene$level))*.19
lines(x=c(3.5,3.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)

```

```{r echo = F}

name = 'uric acid'
name = 'NAD+'
name = 'NADP+'
name = 'nicotinamide riboside+'
name = 'NMN'

gene = data.frame(level = as.numeric(G85R.metabnorm[name,]),
                  group = c(substr(colnames(G85R.metabnorm), 1, nchar(colnames(G85R.metabnorm))-1)))

if(name == "NADP+"){
  gene = data.frame(level = as.numeric(G85R.metabnorm[name,]),
                    group = c(colnames(G85R.metabnorm)))
  gene = gene[gene$group != "WT1",]
  gene$group = substr(gene$group, 1, nchar(gene$group)-1)
}

#gene$group = substr(colnames(G85R.metabnorm), 1, nchar(colnames(G85R.metabnorm))-1)

gene$group = factor(gene$group, levels = c("WT", "TktDfWT", "TktOEWT",
                                           "GR", "TktDfGR", "TktOEGR"))

par(mar = c(5,7,5,2))
plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(min(gene$level), max(gene$level)),
     xlim = c(.5, 6.5),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = NA)

rect(0, 0-(1.5*max(gene$level)), 13, 1.5*max(gene$level), col = 'gray95')

boxplot(gene$level~gene$group, 
        xlab="",
        las = 2, 
        cex.axis = .8,
        xaxt = 'none',
        yaxt = 'none',
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        add = T,
        col = c(rep("lightcyan", 3),
                rep("thistle1", 3)))

box(lwd = 2)

points(x=as.numeric(factor(gene$group)),
       y=gene$level,
       cex=1,
       pch=21,
       bg=c(rep("dodgerblue", 3), rep('firebrick', 3))[as.numeric(factor(gene$group))])

axis(side = 1,
     at = 1:6,
     labels = rep(c("Control", "Df", "OE"), 2),
     line = 0,
     cex.axis = .75,
     lwd.ticks = 2)

axis(side = 2,
     line = 0,
     cex.axis = 1,
     lwd.ticks = 2,
     las = 2)

axis(side = 2,
     at = .5*(max(gene$level) + min(gene$level)),
     label = "Normalized Metabolite Level",
     tick = F,
     padj = -4,
     cex.axis = 1.25)

axis(side = 1,
     at = c(2, 5),
     labels = c("Silent", "G85R"),
     tick = F,
     padj = 2,
     cex.axis = 1)

axis(side = 3,
     at = 3.5,
     labels = name,
     tick = F,
     padj = -1.5,
     cex.axis = 2)

xline1 = min(gene$level) - (max(gene$level)-min(gene$level))*.04
xline2 = min(gene$level) - (max(gene$level)-min(gene$level))*.19
lines(x=c(3.5,3.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)

```

```{r echo = F}

name = 'uric acid'

name = 'NADP+'
name = 'nicotinamide riboside+'
name = 'NMN'
name = 'NAD+'

gene = data.frame(level = as.numeric(G85R.metabnorm[name,]),
                  group = c(substr(colnames(G85R.metabnorm), 1, nchar(colnames(G85R.metabnorm))-1)))

if(name == "NADP+"){
  gene = data.frame(level = as.numeric(G85R.metabnorm[name,]),
                    group = c(colnames(G85R.metabnorm)))
  gene = gene[gene$group != "WT1",]
  gene$group = substr(gene$group, 1, nchar(gene$group)-1)
}

#gene$group = substr(colnames(G85R.metabnorm), 1, nchar(colnames(G85R.metabnorm))-1)

gene$group = factor(gene$group, levels = c("WT", "TktDfWT", "TktOEWT",
                                           "GR", "TktDfGR", "TktOEGR"))

par(mar = c(5,7,5,2))
plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(min(gene$level), max(gene$level)),
     xlim = c(.5, 6.5),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = NA)

rect(0, 0-(1.5*max(gene$level)), 13, 1.5*max(gene$level), col = 'gray95')

boxplot(gene$level~gene$group, 
        xlab="",
        las = 2, 
        cex.axis = .8,
        xaxt = 'none',
        yaxt = 'none',
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        add = T,
        col = c(rep("lightcyan", 3),
                rep("thistle1", 3)))

box(lwd = 2)

points(x=as.numeric(factor(gene$group)),
       y=gene$level,
       cex=1,
       pch=21,
       bg=c(rep("dodgerblue", 3), rep('firebrick', 3))[as.numeric(factor(gene$group))])

axis(side = 1,
     at = 1:6,
     labels = rep(c("Control", "Df", "OE"), 2),
     line = 0,
     cex.axis = .75,
     lwd.ticks = 2)

axis(side = 2,
     line = 0,
     cex.axis = 1,
     lwd.ticks = 2,
     las = 2)

axis(side = 2,
     at = .5*(max(gene$level) + min(gene$level)),
     label = "Normalized Metabolite Level",
     tick = F,
     padj = -4,
     cex.axis = 1.25)

axis(side = 1,
     at = c(2, 5),
     labels = c("Silent", "G85R"),
     tick = F,
     padj = 2,
     cex.axis = 1)

axis(side = 3,
     at = 3.5,
     labels = name,
     tick = F,
     padj = -1.5,
     cex.axis = 2)

xline1 = min(gene$level) - (max(gene$level)-min(gene$level))*.04
xline2 = min(gene$level) - (max(gene$level)-min(gene$level))*.19
lines(x=c(3.5,3.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)

```

```{r echo = F}

name = 'uric acid'

name = 'NADP+'
name = 'nicotinamide riboside+'
name = 'NMN'
name = 'NADH'

gene = data.frame(level = as.numeric(G85R.metabnorm[name,]),
                  group = c(substr(colnames(G85R.metabnorm), 1, nchar(colnames(G85R.metabnorm))-1)))

if(name == "NADP+"){
  gene = data.frame(level = as.numeric(G85R.metabnorm[name,]),
                    group = c(colnames(G85R.metabnorm)))
  gene = gene[gene$group != "WT1",]
  gene$group = substr(gene$group, 1, nchar(gene$group)-1)
}

#gene$group = substr(colnames(G85R.metabnorm), 1, nchar(colnames(G85R.metabnorm))-1)

gene$group = factor(gene$group, levels = c("WT", "TktDfWT", "TktOEWT",
                                           "GR", "TktDfGR", "TktOEGR"))

par(mar = c(5,7,5,2))
plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(min(gene$level), max(gene$level)),
     xlim = c(.5, 6.5),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = NA)

rect(0, 0-(1.5*max(gene$level)), 13, 1.5*max(gene$level), col = 'gray95')

boxplot(gene$level~gene$group, 
        xlab="",
        las = 2, 
        cex.axis = .8,
        xaxt = 'none',
        yaxt = 'none',
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        add = T,
        col = c(rep("lightcyan", 3),
                rep("thistle1", 3)))

box(lwd = 2)

points(x=as.numeric(factor(gene$group)),
       y=gene$level,
       cex=1,
       pch=21,
       bg=c(rep("dodgerblue", 3), rep('firebrick', 3))[as.numeric(factor(gene$group))])

axis(side = 1,
     at = 1:6,
     labels = rep(c("Control", "Df", "OE"), 2),
     line = 0,
     cex.axis = .75,
     lwd.ticks = 2)

axis(side = 2,
     line = 0,
     cex.axis = 1,
     lwd.ticks = 2,
     las = 2)

axis(side = 2,
     at = .5*(max(gene$level) + min(gene$level)),
     label = "Normalized Metabolite Level",
     tick = F,
     padj = -4,
     cex.axis = 1.25)

axis(side = 1,
     at = c(2, 5),
     labels = c("Silent", "G85R"),
     tick = F,
     padj = 2,
     cex.axis = 1)

axis(side = 3,
     at = 3.5,
     labels = name,
     tick = F,
     padj = -1.5,
     cex.axis = 2)

xline1 = min(gene$level) - (max(gene$level)-min(gene$level))*.04
xline2 = min(gene$level) - (max(gene$level)-min(gene$level))*.19
lines(x=c(3.5,3.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)

```

```{r echo = F}

name = 'uric acid'

name = 'NADP+'
name = 'nicotinamide riboside+'
name = 'NMN'
name = 'NADH'
name = 'NADP+'

gene = data.frame(level = as.numeric(G85R.metabnorm[name,]),
                  group = c(substr(colnames(G85R.metabnorm), 1, nchar(colnames(G85R.metabnorm))-1)))

if(name == "NADP+"){
  gene = data.frame(level = as.numeric(G85R.metabnorm[name,]),
                    group = c(colnames(G85R.metabnorm)))
  gene = gene[gene$group != "WT1",]
  gene$group = substr(gene$group, 1, nchar(gene$group)-1)
}

#gene$group = substr(colnames(G85R.metabnorm), 1, nchar(colnames(G85R.metabnorm))-1)

gene$group = factor(gene$group, levels = c("WT", "TktDfWT", "TktOEWT",
                                           "GR", "TktDfGR", "TktOEGR"))

par(mar = c(5,7,5,2))
plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(min(gene$level), max(gene$level)),
     xlim = c(.5, 6.5),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = NA)

rect(0, 0-(1.5*max(gene$level)), 13, 1.5*max(gene$level), col = 'gray95')

boxplot(gene$level~gene$group, 
        xlab="",
        las = 2, 
        cex.axis = .8,
        xaxt = 'none',
        yaxt = 'none',
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        add = T,
        col = c(rep("lightcyan", 3),
                rep("thistle1", 3)))

box(lwd = 2)

points(x=as.numeric(factor(gene$group)),
       y=gene$level,
       cex=1,
       pch=21,
       bg=c(rep("dodgerblue", 3), rep('firebrick', 3))[as.numeric(factor(gene$group))])

axis(side = 1,
     at = 1:6,
     labels = rep(c("Control", "Df", "OE"), 2),
     line = 0,
     cex.axis = .75,
     lwd.ticks = 2)

axis(side = 2,
     line = 0,
     cex.axis = 1,
     lwd.ticks = 2,
     las = 2)

axis(side = 2,
     at = .5*(max(gene$level) + min(gene$level)),
     label = "Normalized Metabolite Level",
     tick = F,
     padj = -4,
     cex.axis = 1.25)

axis(side = 1,
     at = c(2, 5),
     labels = c("Silent", "G85R"),
     tick = F,
     padj = 2,
     cex.axis = 1)

axis(side = 3,
     at = 3.5,
     labels = name,
     tick = F,
     padj = -1.5,
     cex.axis = 2)

xline1 = min(gene$level) - (max(gene$level)-min(gene$level))*.04
xline2 = min(gene$level) - (max(gene$level)-min(gene$level))*.19
lines(x=c(3.5,3.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)

```

```{r echo = F}

name = 'uric acid'

name = 'NADP+'
name = 'nicotinamide riboside+'
name = 'NMN'
name = 'NAD+'
name = 'NADPH'
gene = data.frame(level = as.numeric(G85R.metabnorm[name,]),
                  group = c(substr(colnames(G85R.metabnorm), 1, nchar(colnames(G85R.metabnorm))-1)))

if(name == "NADP+"){
  gene = data.frame(level = as.numeric(G85R.metabnorm[name,]),
                    group = c(colnames(G85R.metabnorm)))
  gene = gene[gene$group != "WT1",]
  gene$group = substr(gene$group, 1, nchar(gene$group)-1)
}

if(name == "NADPH"){
  gene = na.omit(gene)
}

#gene$group = substr(colnames(G85R.metabnorm), 1, nchar(colnames(G85R.metabnorm))-1)

gene$group = factor(gene$group, levels = c("WT", "TktDfWT", "TktOEWT",
                                           "GR", "TktDfGR", "TktOEGR"))

par(mar = c(5,7,5,2))
plot(x = NA,
      y = NA,
     type = 'n',
     ylim = c(min(gene$level), max(gene$level)),
     xlim = c(.5, 6.5),
     xaxt = "n",
     ylab="",
     yaxt = "n",
     xlab = NA)

rect(0, 0-(1.5*max(gene$level)), 13, 1.5*max(gene$level), col = 'gray95')

boxplot(gene$level~gene$group, 
        xlab="",
        las = 2, 
        cex.axis = .8,
        xaxt = 'none',
        yaxt = 'none',
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        add = T,
        col = c(rep("lightcyan", 3),
                rep("thistle1", 3)))

box(lwd = 2)

points(x=as.numeric(factor(gene$group)),
       y=gene$level,
       cex=1,
       pch=21,
       bg=c(rep("dodgerblue", 3), rep('firebrick', 3))[as.numeric(factor(gene$group))])

axis(side = 1,
     at = 1:6,
     labels = rep(c("Control", "Df", "OE"), 2),
     line = 0,
     cex.axis = .75,
     lwd.ticks = 2)

axis(side = 2,
     line = 0,
     cex.axis = 1,
     lwd.ticks = 2,
     las = 2)

axis(side = 2,
     at = .5*(max(gene$level) + min(gene$level)),
     label = "Normalized Metabolite Level",
     tick = F,
     padj = -4,
     cex.axis = 1.25)

axis(side = 1,
     at = c(2, 5),
     labels = c("Silent", "G85R"),
     tick = F,
     padj = 2,
     cex.axis = 1)

axis(side = 3,
     at = 3.5,
     labels = name,
     tick = F,
     padj = -1.5,
     cex.axis = 2)

xline1 = min(gene$level) - (max(gene$level)-min(gene$level))*.04
xline2 = min(gene$level) - (max(gene$level)-min(gene$level))*.19
lines(x=c(3.5,3.5), y = c(xline1, xline2), lty = 1, lwd = 2, xpd = T)

```

```{r include = F}

rawdata = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/MetabolomicsForMetaboanalyst.txt", sep = "\t", row.names = 1)

KEGG.Key = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Metabolomics/RawMetabolomicData.csv", row.names = 1)
KEGG.Key = setNames(KEGG.Key[-c(1:2), 2], row.names(KEGG.Key)[-c(1:2)])

genotypes = setNames(c("WT", "GR", "TktOEWT", "TktOEGR", "TktDfWT", "TktDfGR"), c("C", "E", "A", "F", "B", "D"))

metadata = data.frame(Sample = colnames(rawdata), 
                      Group = as.character(rawdata[1,]), 
                      TIC = as.numeric(rawdata[2,]), 
                      Genotypes = genotypes[as.character(rawdata[1,])])

row.names(metadata) = metadata$Sample

norm.func = function(data){
  nd = (as.numeric(data)/metadata$TIC)*1000
  return(nd)
}

norm.data = t(apply(rawdata[3:nrow(rawdata),], 1, norm.func))
colnames(norm.data) = colnames(rawdata)

group.mean=function(data){
  gm = mean(na.omit(data))
}

mean.data = matrix(0, nrow=nrow(norm.data), ncol = length(genotypes))
row.names(mean.data) = row.names(norm.data)
colnames(mean.data) = unique(metadata$Group)

i=1
while(i<=ncol(mean.data)){
  mean.data[,i] = apply(norm.data[,metadata[metadata$Group == colnames(mean.data)[i], "Sample"]], 1, group.mean)
  i = i+1
}

compare.conditions = function(condition1, condition2){
  columns1 = metadata[metadata$Group == condition1, "Sample"]
  columns2 = metadata[metadata$Group == condition2, "Sample"]
  p=setNames(rep(NA, nrow(norm.data)), row.names(norm.data))
  i=1
  while(i<=length(p)){
    if(length(na.omit(norm.data[i,columns1]))>2 & 
       length(na.omit(norm.data[i,columns2]))>2){
      temp = t.test(na.omit(norm.data[i,columns1]),
                    na.omit(norm.data[i,columns2]), paired = F, var.equal = T)
      p[i] = temp[[3]]
      }
    i=i+1
    }

  fdr = p.adjust(p, "BH", length(p))

  fc = mean.data[,condition1]/mean.data[,condition2]

  comparison.table = data.frame(p = p, 
                                FDR = fdr, 
                                FC = fc, 
                                KEGG = KEGG.Key[row.names(mean.data)])

  comparison.table = comparison.table[order(comparison.table$FDR),]
}

GRxWT.C = compare.conditions("E", "C")
GRxWT.Df = compare.conditions("D", "B")
GRxWT.OE = compare.conditions("F", "A")

GR.CxDf = compare.conditions("D", "E")
WT.CxDf = compare.conditions("B", "C")

GR.CxOE = compare.conditions("F", "E")
WT.CxOE = compare.conditions("A", "C")

TKT.EdgeR.FC = G85R.FC
TKT.EdgeR = G85R.FDR

```

```{r echo = F}

temp.edges = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.edges.csv", row.names = 1)

purine.path = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.path.csv", row.names = 1)

temp.nodes = data.frame(id = unique(c(purine.path$from, purine.path$to)),
                        group = "metabolite",
                        color.background = "white",
                        color.border = "black",
                        x = 0,
                        y = 0,
                        size = 30,
                        FC = NA,
                        pval = NA,
                        title = "")
temp.nodes$title = temp.nodes$id

row.names(temp.nodes) = temp.nodes$id

temp.nodes[c("Pentose Phosphate", "Excretion", "Riboflavin.Metabolism",
            "Folate.Biosynthesis", "DNA.A", "DNA.G", "RNA.A", "RNA.G"), "group"] = "Outside Path"

temp.nodes["RNA.G", 'x'] = 0
temp.nodes["RNA.G", 'y'] = 1.5

temp.nodes["Folate.Biosynthesis", 'x'] = 0
temp.nodes["Folate.Biosynthesis", 'y'] = 2

temp.nodes["Riboflavin.Metabolism", 'x'] = 0
temp.nodes["Riboflavin.Metabolism", 'y'] = 2.5

temp.nodes["DNA.G", 'x'] = 0
temp.nodes["DNA.G", 'y'] = 4

temp.nodes["GTP", 'x'] = 1
temp.nodes["GTP", 'y'] = 2

temp.nodes["dGTP", 'x'] = 1
temp.nodes["dGTP", 'y'] = 4

temp.nodes["GDP", 'x'] = 2
temp.nodes["GDP", 'y'] = 3

temp.nodes["dGDP", 'x'] = 2
temp.nodes["dGDP", 'y'] = 3

temp.nodes["dGMP", 'x'] = 2
temp.nodes["dGMP", 'y'] = 4

temp.nodes["deoxyguanosine", 'x'] = 2
temp.nodes["deoxyguanosine", 'y'] = 5

temp.nodes["GMP", 'x'] = 3
temp.nodes["GMP", 'y'] = 2

temp.nodes["guanosine", 'x'] = 3
temp.nodes["guanosine", 'y'] = 3

temp.nodes["guanine", 'x'] = 3
temp.nodes["guanine", 'y'] = 4

temp.nodes["IDP", 'x'] = 4
temp.nodes["IDP", 'y'] = 1

temp.nodes["XMP", 'x'] = 4
temp.nodes["XMP", 'y'] = 2

temp.nodes["xanthosine", 'x'] = 4
temp.nodes["xanthosine", 'y'] = 3

temp.nodes["xanthine", 'x'] = 4
temp.nodes["xanthine", 'y'] = 4

temp.nodes["uric acid", 'x'] = 4
temp.nodes["uric acid", 'y'] = 5

temp.nodes["Excretion", 'x'] = 4
temp.nodes["Excretion", 'y'] = 6

temp.nodes = temp.nodes[temp.nodes$id != "Pentose Phosphate" ,]
temp.edges = temp.edges[temp.edges$from != "Pentose Phosphate" & temp.edges$to != "Pentose Phosphate" ,]

temp.nodes["FAICAR", "x"] = 5
temp.nodes["FAICAR", "y"] = 1

temp.nodes["AICA ribonucleotide", "x"] = 5
temp.nodes["AICA ribonucleotide", "y"] = 0

temp.nodes["IMP", 'x'] = 5
temp.nodes["IMP", 'y'] = 2

temp.nodes["inosine", 'x'] = 5
temp.nodes["inosine", 'y'] = 3

temp.nodes["hypoxanthine", 'x'] = 5
temp.nodes["hypoxanthine", 'y'] = 4

temp.nodes["deoxyinosine", 'x'] = 5
temp.nodes["deoxyinosine", 'y'] = 5

temp.nodes["adenylosuccinate", 'x'] = 6
temp.nodes["adenylosuccinate", 'y'] = 1

temp.nodes["AMP", 'x'] = 6
temp.nodes["AMP", 'y'] = 2

temp.nodes["adenosine", 'x'] = 6
temp.nodes["adenosine", 'y'] = 3

temp.nodes["adenine", 'x'] = 6
temp.nodes["adenine", 'y'] = 4

temp.nodes["deoxyadenosine", 'x'] = 6
temp.nodes["deoxyadenosine", 'y'] = 5

temp.nodes["cAMP", 'x'] = 7
temp.nodes["cAMP", 'y'] = 1

temp.nodes["ADP", 'x'] = 7
temp.nodes["ADP", 'y'] = 3

temp.nodes["dADP", 'x'] = 7
temp.nodes["dADP", 'y'] = 4

temp.nodes["dAMP", 'x'] = 7
temp.nodes["dAMP", 'y'] = 5

temp.nodes["ATP", 'x'] = 8
temp.nodes["ATP", 'y'] = 2

temp.nodes["dATP", 'x'] = 8
temp.nodes["dATP", 'y'] = 4

temp.nodes["RNA.A", 'x'] = 9
temp.nodes["RNA.A", 'y'] = 2

temp.nodes["DNA.A", 'x'] = 9
temp.nodes["DNA.A", 'y'] = 4

temp.nodes$x = temp.nodes$x*150
temp.nodes$y = temp.nodes$y*150

visNetwork(temp.nodes, temp.edges)%>%
  visNodes(physics = F)

#write.csv(temp.nodes, "/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.nodes.csv")

```

```{r}
temp.edges = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.edges.csv", row.names = 1)

temp.edges[temp.edges$from == "Pentose Phosphate", "to"] = "PRPP"

purine.path = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.path.csv", row.names = 1)

add.edges = data.frame(enzyme = c("2.2.1.1",
                                  "5.1.3.1",
                                  "5.3.1.6",
                                  "5.4.2.7",
                                  "5.4.2.2"),
                       from = c("D-ribose-1/5-phosphate / D-ribulose 5-phosphate",
                                "D-ribose-1/5-phosphate / D-ribulose 5-phosphate",
                                "D-ribose-1/5-phosphate / D-ribulose 5-phosphate",
                                "D-ribose-1/5-phosphate / D-ribulose 5-phosphate"),
                       to = c("D-sedoheptulose-7-phosphate",
                              "D-Xylulose 5-phosphate",
                              "D-ribose-1/5-phosphate / D-ribulose 5-phosphate",
                              "D-ribose-1/5-phosphate / D-ribulose 5-phosphate"),
                       arrows = c("to;from",
                                  "to;from",
                                  "to;from",
                                  "to;from",
                                  "to;from"),
                       FBgn = "")

```

```{r echo = F}

temp.edges = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.edges.csv", row.names = 1)

temp.edges[temp.edges$from == "Pentose Phosphate", "to"] = "PRPP"

purine.path = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.path.csv", row.names = 1)

temp.nodes = data.frame(id = unique(c(purine.path$from, purine.path$to)),
                        group = "metabolite",
                        color.background = "white",
                        color.border = "black",
                        x = 0,
                        y = 0,
                        size = 0,
                        FC = NA,
                        pval = NA,
                        title = "",
                        label = "")
temp.nodes$label = temp.nodes$id

row.names(temp.nodes) = temp.nodes$id

temp.nodes[c("Pentose Phosphate", "Excretion", "Riboflavin.Metabolism",
            "Folate.Biosynthesis", "DNA.A", "DNA.G", "RNA.A", "RNA.G"), "group"] = "Outside Path"

temp.nodes["Pentose Phosphate" , 'x'] = -2.5
temp.nodes["Pentose Phosphate" , 'y'] = -1.5

temp.nodes["D-ribose-1/5-phosphate / D-ribulose 5-phosphate", 'x'] = -3.5
temp.nodes["D-ribose-1/5-phosphate / D-ribulose 5-phosphate", 'y'] = 0.25

temp.nodes["PRPP", 'x'] = -2.5
temp.nodes["PRPP", 'y'] = 0

temp.nodes["Ribosylamine-5p", 'x'] = -1.5
temp.nodes["Ribosylamine-5p", 'y'] = 0.25

temp.nodes["GAR", 'x'] = -.5
temp.nodes["GAR", 'y'] = 0

temp.nodes["RNA.G", 'x'] = -.25
temp.nodes["RNA.G", 'y'] = 1

temp.nodes["Folate.Biosynthesis", 'x'] = -.25
temp.nodes["Folate.Biosynthesis", 'y'] = 1.5

temp.nodes["Riboflavin.Metabolism", 'x'] = -.25
temp.nodes["Riboflavin.Metabolism", 'y'] = 2

temp.nodes["DNA.G", 'x'] = 0
temp.nodes["DNA.G", 'y'] = 4.5

temp.nodes["FGAR", 'x'] = .5
temp.nodes["FGAR", 'y'] = 0.25

temp.nodes["GTP", 'x'] = .75
temp.nodes["GTP", 'y'] = 1.5

temp.nodes["dGTP", 'x'] = .75
temp.nodes["dGTP", 'y'] = 3.75

temp.nodes["2-(Formamido)-N1-(5-phospho-D-ribosyl)acetamidine", 'x'] = 1.5
temp.nodes["2-(Formamido)-N1-(5-phospho-D-ribosyl)acetamidine", 'y'] = 0

temp.nodes["GDP", 'x'] = 1.75
temp.nodes["GDP", 'y'] = 2.25

temp.nodes["dGDP", 'x'] = 1.5
temp.nodes["dGDP", 'y'] = 3

temp.nodes["dGMP", 'x'] = 1.75
temp.nodes["dGMP", 'y'] = 3.75

temp.nodes["AIR", 'x'] = 2.5
temp.nodes["AIR", 'y'] = 0.25

temp.nodes["deoxyguanosine", 'x'] = 2.75
temp.nodes["deoxyguanosine", 'y'] = 4.25

temp.nodes["GMP", 'x'] = 2.75
temp.nodes["GMP", 'y'] = 1.75

temp.nodes["CAIR", 'x'] = 3.5
temp.nodes["CAIR", 'y'] = 0

temp.nodes["guanosine", 'x'] = 3.5
temp.nodes["guanosine", 'y'] = 2.5

temp.nodes["guanine", 'x'] = 3.5
temp.nodes["guanine", 'y'] = 3.5

temp.nodes["IDP", 'x'] = 4.5
temp.nodes["IDP", 'y'] = 1.25

temp.nodes["SAICAR", 'x'] = 4.5
temp.nodes["SAICAR", 'y'] = 0.25

temp.nodes["XMP", 'x'] = 4.5
temp.nodes["XMP", 'y'] = 2

temp.nodes["xanthosine", 'x'] = 4.5
temp.nodes["xanthosine", 'y'] = 3

temp.nodes["xanthine", 'x'] = 4.5
temp.nodes["xanthine", 'y'] = 4

temp.nodes["uric acid", 'x'] = 4.5
temp.nodes["uric acid", 'y'] = 5

temp.nodes["Excretion", 'x'] = 4.5
temp.nodes["Excretion", 'y'] = 6

temp.nodes["AICA ribonucleotide", "x"] = 5.5
temp.nodes["AICA ribonucleotide", "y"] = 0

temp.nodes["FAICAR", "x"] = 5.5
temp.nodes["FAICAR", "y"] = 1

temp.nodes["IMP", 'x'] = 5.5
temp.nodes["IMP", 'y'] = 2

temp.nodes["inosine", 'x'] = 5.5
temp.nodes["inosine", 'y'] = 3

temp.nodes["hypoxanthine", 'x'] = 5.5
temp.nodes["hypoxanthine", 'y'] = 4

temp.nodes["deoxyinosine", 'x'] = 6.5
temp.nodes["deoxyinosine", 'y'] = 4.5

temp.nodes["adenylosuccinate", 'x'] = 6.5
temp.nodes["adenylosuccinate", 'y'] = 1.25

temp.nodes["AMP", 'x'] = 7.5
temp.nodes["AMP", 'y'] = 2

temp.nodes["adenosine", 'x'] = 6.5
temp.nodes["adenosine", 'y'] = 2.5

temp.nodes["adenine", 'x'] = 7.5
temp.nodes["adenine", 'y'] = 3

temp.nodes["deoxyadenosine", 'x'] = 7.5
temp.nodes["deoxyadenosine", 'y'] = 4

temp.nodes["cAMP", 'x'] = 8.25
temp.nodes["cAMP", 'y'] = 1.25

temp.nodes["ADP", 'x'] = 8.25
temp.nodes["ADP", 'y'] = 2.5

temp.nodes["dADP", 'x'] = 8.5
temp.nodes["dADP", 'y'] = 3

temp.nodes["dAMP", 'x'] = 8.5
temp.nodes["dAMP", 'y'] = 4

temp.nodes["ATP", 'x'] = 9
temp.nodes["ATP", 'y'] = 1.825

temp.nodes["dATP", 'x'] = 9.25
temp.nodes["dATP", 'y'] = 3.5

temp.nodes["RNA.A", 'x'] = 10
temp.nodes["RNA.A", 'y'] = 1.825

temp.nodes["DNA.A", 'x'] = 10
temp.nodes["DNA.A", 'y'] = 4

temp.nodes$x = temp.nodes$x*150
temp.nodes$y = temp.nodes$y*150

visNetwork(temp.nodes, temp.edges)%>%
  visNodes(physics = F, shape = "box")

#write.csv(temp.nodes, "/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.nodes.csv")

```

```{r}

##genes in the kegg pathway and their enzyme IDs
glycolysis.genes = read.csv('/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GlycolysisGenesKEGGinfo.csv')
##Edge data for genes connecting metabolites
glycolysis.edges = read.csv('/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GlycolysisEnzymeEdgesKEGGinfo.csv')
##Metabolites in the KEGG pathway with their ID
glycolysis.metabs = read.csv('/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/GlycolysisMetaboliteKEGGinfo.csv')

node.data = node.info('/Users/johncsantiago/Documents/GitHub/WhartonLab/TKT_RNAseq/R Code/dme00010.xml')

##extracting position data for nodes
node.info = data.frame(ID = unlist(node.data$kegg.names, use.names = F),
                       number = rep(names(node.data$kegg.names),
                                    lengths(node.data$kegg.names)),
                       x = 0,
                       y = 0,
                       color = 0,
                       names = "")
node.info$x = node.data$x[node.info$number]
node.info$y = node.data$y[node.info$number]

##make new columns for transcriptomic data
glycolysis.genes$FBgn = ""
glycolysis.genes$Symbol = ""
glycolysis.genes$FC = 0
glycolysis.genes$pval = 1

##Isolate gene ID info for goi
temp = GeneIDKey[GeneIDKey$CG %in% glycolysis.genes$CG,]
row.names(temp) = temp$CG

##fill in new columns with transcriptomic data
glycolysis.genes$FBgn = temp[glycolysis.genes$CG,'FBgn']
glycolysis.genes$Symbol = temp[glycolysis.genes$CG,'Symbol']
glycolysis.genes$FC = 0
glycolysis.genes$FC = 0

##make new columns for metabolomic data
row.names(glycolysis.metabs) = glycolysis.metabs$KEGG.ID
glycolysis.metabs$metabolomicsID = ""
glycolysis.metabs$FC = 0
glycolysis.metabs$pval = 1

##retrieve name used in metabolomic data for metabolite IDs in path
##including IDs associated with multiple metabolomic names
i = 1
while(i<=nrow(glycolysis.metabs)){
  if(length(grep(glycolysis.metabs$KEGG.ID[i], KEGG.Key)) > 0){
    glycolysis.metabs$metabolomicsID[i]= names(KEGG.Key[grep(glycolysis.metabs$KEGG.ID[i], KEGG.Key)])
  }
  i = i +1
}

##retrieve metabolite fold change and pvalue data
glycolysis.metabs$FC = 0
glycolysis.metabs$pval = 0

##using the positions for nodes (metabolites) only
metab.info = node.info[node.info$ID %in% glycolysis.metabs$KEGG.ID,]

metab.info$FC = glycolysis.metabs[metab.info$ID, 'FC']
metab.info$pval = glycolysis.metabs[metab.info$ID, 'pval']
metab.info$names = glycolysis.metabs[metab.info$ID, 'Metabolite']
metab.info$color = glycolysis.metabs[metab.info$ID, 'FC']
metab.info$size = glycolysis.metabs[metab.info$ID, 'pval']

color.pallette = colorRampPalette(c('royalblue', 'grey', "firebrick"))(401)

metab.info$color[metab.info$color > 1] = 1
metab.info$color[metab.info$color < -1] = -1
metab.info$color = (metab.info$color * 200) +201
metab.info$color = color.pallette[metab.info$color]
metab.info$color[is.na(metab.info$color)] = 'dimgrey'

metab.info$size = 10 + 10*abs(metab.info$size)
metab.info$size[is.na(metab.info$size)] = 10

metab.info$title = paste0(glycolysis.metabs[metab.info$ID, "metabolomicsID"],
                          "<br>log2(FC): ", signif(metab.info$FC,3),
                          "<br>pval: ", signif(10^-metab.info$pval,3))

nodes = data.frame(id = paste0(metab.info$names),
                   color = metab.info$color,
                   size = metab.info$size,
                   x = 2*metab.info$x,
                   y = 2*metab.info$y,
                   physics = F,
                   group = "",
                   title = metab.info$title)

glycolysis.edges$FBgn = ''
temp.glycolysis.path = glycolysis.edges

i = 1
while(i<=nrow(glycolysis.edges)){
  enzymeID = paste0(glycolysis.edges$enzyme[i], "$")
  if(nrow(glycolysis.genes[grep(enzymeID, glycolysis.genes$Enzyme),])>0){
    temp = glycolysis.genes[grep(enzymeID, glycolysis.genes$Enzyme), "CG"]
    temp = row.names(GeneIDKey[GeneIDKey$CG %in% temp,])
    temp2 = glycolysis.edges[rep(i, length(temp)),]
    temp2$FBgn = temp
  }else{
    temp2 = glycolysis.edges[i,]
  }
  temp.glycolysis.path = rbind(temp.glycolysis.path,temp2)

  i = i +1
}

glycolysis.edges = temp.glycolysis.path[(nrow(glycolysis.edges)+1):nrow(temp.glycolysis.path),]

glycolysis.edges$size = 0
glycolysis.edges$size[glycolysis.edges$size <1] = 1
glycolysis.edges$size[is.na(glycolysis.edges$size)] = 1
glycolysis.edges$size = 2*glycolysis.edges$size

glycolysis.edges$color = "white"
## max color intensity is at FC of 2
#glycolysis.edges$color[glycolysis.edges$color>2] = 2
#glycolysis.edges$color[glycolysis.edges$color<(-2)] = -2
#glycolysis.edges$color = 201+(glycolysis.edges$color *100)
#glycolysis.edges$color = color.pallette[glycolysis.edges$color]
#glycolysis.edges$color[is.na(glycolysis.edges$color)] = 'dimgrey'

glycolysis.edges$title = paste0(GeneIDKey[glycolysis.edges$FBgn, "Symbol"],
                          "<br>log2(FC): ", signif(TKT.EdgeR.FC[glycolysis.edges[,'FBgn'], comparison.T],3),
                          "<br>FDR: ", signif(TKT.EdgeR[glycolysis.edges[,'FBgn'], comparison.T],3),
                          "<br>", comparison.means[[comparison.T]][1], " cpm: ", signif(mean.cpm[glycolysis.edges[,'FBgn'],comparison.means[[comparison.T]][1]],3),
                          "<br>", comparison.means[[comparison.T]][2], " cpm: ", signif(mean.cpm[glycolysis.edges[,'FBgn'],comparison.means[[comparison.T]][2]],3))

glycolysis.edges = data.frame(from = glycolysis.edges$from.name,
                              to = glycolysis.edges$to.name)
,
                              arrows = glycolysis.edges$arrow)
,
                              color = glycolysis.edges$color)
,
                              width = glycolysis.edges$size)
,
                              title = glycolysis.edges$title,
                              physics = T)

nodes = data.frame(id = paste0(nodes$id),
                   color = nodes$color,
                   size = nodes$size,
                   x = nodes$x,
                   y = nodes$y,
                   physics = F,
                   group = "",
                   title = nodes$title)

OutsidePath = data.frame(id = c('Pentose Phosphate Pathway',
                               'Propanoate Metabolism',
                               'Starch and Sucrose Metabolism',
                               'Citrate Cycle'),
                        color = "black",
                        size = 15,
                        x= c(1350,1254,966,175),
                        y=c(1112,1822,110,1700),
                        physics = F,
                        group = 'Outside Path',
                        title = "")

nodes = rbind(nodes, OutsidePath)

```

```{r echo = F}

network.fun = function(comparison.T, comparison.M, cond1, cond2, pnp){
temp = colorRampPalette(c("dodgerblue", 'royalblue', "firebrick",'red'))(401)

visedges = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.edges.csv", row.names = 1)
visedges$color = round(TKT.EdgeR.FC[purine.path$FBgn, comparison.T],2)
visedges$color[visedges$color < (-2)] = -2
visedges$color[visedges$color > (2)] = 2
visedges$color = (visedges$color * 100) +201
visedges$color = temp[visedges$color]
visedges$symbol = GeneIDKey[visedges$FBgn, 'Symbol']
visedges$fdr = signif(TKT.EdgeR[visedges$FBgn, comparison.T], 2)
visedges$GR = round(mean.cpm[visedges$FBgn, cond1],1)
visedges$WT = round(mean.cpm[visedges$FBgn, cond2],1)
visedges$width = .01 + (1*abs(log10(visedges$WT)))
visedges$title = paste0(visedges$symbol,
                        "<br>FDR: ", visedges$fdr, 
                        "<br>",cond1," CPM: ", visedges$GR, 
                        "<br>",cond2," CPM: ", visedges$WT)

row.names(comparison.M)[row.names(comparison.M) == "cAMP(2',3') / cAMP(3',5')"] = 'cAMP'

node.data = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/GeneralDataFiles/purine.nodes.csv", row.names = 1)
node.colors = na.omit(comparison.M[intersect(row.names(comparison.M), node.names),])
node.pval = setNames(signif((node.colors$p),2), row.names(node.colors))
node.data[names(node.pval), 'pval'] = node.pval
node.colors = setNames(round(log2(node.colors$FC),2), row.names(node.colors))
node.data[names(node.colors), "FC"] = node.colors
node.data[names(node.colors),'size'] = node.colors
node.data[,'size'] = (node.data[,'size'] + 4.5)*7.5
node.colors[node.colors < (-2)] = -2
node.colors[node.colors>2] = 2
node.colors = (node.colors*100)+201
node.colors = setNames(temp[node.colors], names(node.colors)) 
node.data[names(node.colors), 'color.background'] = node.colors
node.data$title = paste0(node.data$id,
                         "<br>FC: ", node.data$FC,
                         "<br>pval: ", node.data$pval)
node.data[node.data$group == "Outside Path", 'color.background'] = 'black'
node.data[node.data$color.background == 'white', 'group'] = "no data"

nodes = node.data
nodes[names(node.pval[node.pval > 0.05]), "color"] = 'grey'

visedges$color[visedges$from %in% row.names(nodes)[nodes$group == "Outside Path"]] = 'black'
visedges$color[visedges$to %in% row.names(nodes)[nodes$group == "Outside Path"]] = 'black'
visedges = visedges[,c('from', 'to', 'arrows', 'color', 'width', 'title')]
row.names(visedges) = c(1:nrow(visedges))
visedges = visedges[!is.na(visedges$color),]

lnodes <- data.frame(label = c("Sig. Increased", "Sig. Decreased", "Not. Sig.", "No Data"), 
shape = 'dot', color = list(background = c("red", "dodgerblue",'dimgrey', 'white'), border = "black"),
size = c(40, 15, 25, 25),
 title = "")

visNetwork(nodes,visedges,
                  main = paste0(cond1, " vs ", cond2))%>%
  visNodes(physics = F)%>%
  visLayout(hierarchical=F, improvedLayout=T)%>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE)%>%
  visEdges(arrows = visedges$arrows)%>%
  visGroups(groupname = "Outside Path", shape = 'box', physics=F, size = 25, font=list(color="white"))%>%
    visLegend(addNodes = lnodes, useGroups = FALSE, width = .115)

}

```

```{r echo = F}

comparison.T = "GRxWT.FC"
cond1 = 'GR.F'
cond2 = "WT.F"
comparison.M = GRxWT.C
pnp = F

network.fun(comparison.T, comparison.M, cond1, cond2, pnp)

```

```{r echo = F}

comparison.T = "GRxWT.FDf"
cond1 = 'TktDfGR.F'
cond2 = "TktDfWT.F"
comparison.M = GRxWT.Df
pnp = F

network.fun(comparison.T, comparison.M, cond1, cond2, pnp)

```

```{r echo = F}

comparison.T = "GRxWT.FOE"
cond1 = 'TktOEGR.F'
cond2 = "TktOEWT.F"
comparison.M = GRxWT.OE
pnp = F

network.fun(comparison.T, comparison.M, cond1, cond2, pnp)

```

```{r echo = F}

comparison.T = "GRxWT.FC"
cond1 = 'GR.F'
cond2 = "WT.F"
comparison.M = GRxWT.C
pnp = T

network.fun(comparison.T, comparison.M, cond1, cond2, pnp)

```

```{r echo = F}

comparison.T = "GRxWT.FDf"
cond1 = 'TktDfGR.F'
cond2 = "TktDfWT.F"
comparison.M = GRxWT.Df
pnp = T

network.fun(comparison.T, comparison.M, cond1, cond2, pnp)

```

```{r echo = F}

comparison.T = "GRxWT.FOE"
cond1 = 'TktOEGR.F'
cond2 = "TktOEWT.F"
comparison.M = GRxWT.OE
pnp = T

network.fun(comparison.T, comparison.M, cond1, cond2, pnp)

```

```{r include = F}

NAD.nodes = data.frame(ID = character(),
                       x = numeric(),
                       y = numeric(),
                       names = character())

NAD.nodes[1,] = c('C03150', 0,  0,    "nicotinamide riboside+")
NAD.nodes[2,] = c('C00153', 0,  5,    "nicotinamide")
NAD.nodes[3,] = c('C00253', 0,  7.5,  "nicotinic acid")
NAD.nodes[4,] = c('C00455', 5,  2.5,  "NMN")
NAD.nodes[5,] = c('C01185', 5,  7.5,  "NAMN")
NAD.nodes[6,] = c('C00006', 10, -2.5, "NADP+")
NAD.nodes[7,] = c('C00003', 10, 2.5,  "NAD+")
NAD.nodes[8,] = c('C00857', 10, 7.5,  "NAAD")
NAD.nodes[9,] = c('C00005', 15, -2.5, "NADPH")
NAD.nodes[10,] = c('C00004', 15, 2.5,  "NADH")

NAD.nodes[11,] = c('C00806', 20, 10,  "tryptophan")
NAD.nodes[12,] = c('C02794', 5, 10,  "3-hydroxykynurenine")
NAD.nodes[13,] = c('C01718', 10, 10,  "kynurenine")
NAD.nodes[14,] = c('C01717', 5, 12.5,  "kynurenic acid")
NAD.nodes[15,] = c('C02700', 15, 10,  "N-Formylkynurenine")

temp = unlist(strsplit(keggGet("C00806")[[1]]["REACTION"][[1]], split = " "))

#NR <=> NMN: "3.1.3.5", "2.7.1.22"
#NR <=> NAM: "2.4.2.1", "3.2.2.3"
#NMN <=> NAM: "3.2.2.14" "2.4.2.12"
#NA <=> NAMN: "2.4.2.55" "6.3.4.21" "2.4.2.21"
#NAMN <=> NAAD: "3.6.1.9"  "3.6.1.22" "2.7.7.1"  "2.7.7.18" 
#NAMN <=> NMN: "3.5.1.42"

#temp = unlist(strsplit(keggGet("C00003")[[1]]["REACTION"][[1]], split = " "))
#i = 1

#NADH = character()
#NADP = character()
#NMN = character()
#NAAD = character()

#while(i <= length(temp)){
#  temp.eq = keggGet(temp[i])[[1]][c("EQUATION")][[1]]
#  temp.enz = keggGet(temp[i])[[1]][c("ENZYME")][[1]]
#  if(length(grep("C00004", temp.eq)) == 1){
#    NADH = unique(c(NADH, temp.enz))
#  }
#  if(length(grep("C00006", temp.eq)) == 1){
#    NADP = unique(c(NADP, temp.enz))
#  }
#  if(length(grep("C00857", temp.eq)) == 1){
#    NAAD = unique(c(NAAD, temp.enz))
#  }
#  if(length(grep("C00455", temp.eq)) == 1){
#    NMN = unique(c(NMN, temp.enz))
#  }
#  i = i +1
#}

#NAD+ <=> NMN: NMN
#NAD+ <=> NADP: NADP
#NAD+ <=> NAAD: NAAD
#NAD+ <=> NADH: NADH

#temp = unlist(strsplit(keggGet("C00006")[[1]]["REACTION"][[1]], split = " "))
#i = 1

#NADPH = character()

#while(i <= length(temp)){
#  temp.eq = keggGet(temp[i])[[1]][c("EQUATION")][[1]]
#  temp.enz = keggGet(temp[i])[[1]][c("ENZYME")][[1]]
#  if(length(grep("C00005", temp.eq)) == 1){
#    NADPH = c(NADPH, temp.enz)
#  }
#  i = i +1
#  temp.eq
#  NADPH
#}

#NADPH = unique(NADPH)

#enzymes = list(NR2NAM  = c("2.4.2.1", "3.2.2.3"),
#               NR2NMN  = c("3.1.3.5", "2.7.1.22"),
#               NAM2NMN = c("3.2.2.14", "2.4.2.12"),
#               NMN2NAD = NMN,
#               NAD2NADP = NADP,
#               NAD2NADH = NADH,
#               NA2NAMN = c("2.4.2.55", "6.3.4.21", "2.4.2.21"),
#               NAMN2NAAD = c("3.6.1.9", "3.6.1.22", "2.7.7.1", "2.7.7.18" ),
#               NAMN2NMN = c("3.5.1.42"),
#               NAAD2NAD = NAAD,
#               NADP2NADPH = NADPH)

#NR2NAM = unique(c(names(nodeID[grep("\\b2.4.2.1\\b", nodeID)]),
#                  names(nodeID[grep("\\b3.2.2.3\\b", nodeID)])))

#NR2NMN = unique(c(names(nodeID[grep("\\b3.1.3.5\\b", nodeID)]),
#                  names(nodeID[grep("\\b2.7.1.22\\b", nodeID)])))

#NAM2NMN = unique(c(names(nodeID[grep("\\b3.2.2.14\\b", nodeID)]),
#                   names(nodeID[grep("\\b2.4.2.12\\b", nodeID)])))

#NA2NAMN = unique(c(names(nodeID[grep("\\b2.4.2.55\\b", nodeID)]),
#                   names(nodeID[grep("\\b6.3.4.21\\b", nodeID)]),
#                   names(nodeID[grep("\\b2.4.2.21\\b", nodeID)])))

#NAMN2NAAD = unique(c(names(nodeID[grep("\\b3.6.1.9\\b", nodeID)]),
#                     names(nodeID[grep("\\b3.6.1.22\\b", nodeID)]),
#                     names(nodeID[grep("\\b2.7.7.1\\b", nodeID)]),
#                     names(nodeID[grep("\\b2.7.7.18\\b", nodeID)])))

#NAMN2NMN = unique(c(names(nodeID[grep("\\b3.5.1.42\\b", nodeID)])))

#NAAD2NAD = unique(c(names(nodeID[grep("\\b6.3.1.5\\b", nodeID)]),
#                    names(nodeID[grep("\\b6.3.5.1\\b", nodeID)])))

#NMN2NAD = character()
#i=1
#while(i <= length(NMN)){
#  NMN2NAD = unique(c(NMN2NAD, names(nodeID[grep(paste0("\\b",NMN[i],"\\b"), nodeID)])))
#  i=i+1
#}

#NAD2NADP = character()
#i=1
#while(i <= length(NADP)){
#  NAD2NADP = unique(c(NAD2NADP, names(nodeID[grep(paste0("\\b",NADP[i],"\\b"), #nodeID)])))
#  i=i+1
#}

#NAD2NADH = character()
#i=1
#while(i <= length(NADH)){
#  NAD2NADH = unique(c(NAD2NADH, names(nodeID[grep(paste0("\\b",NADH[i],"\\b"), nodeID)])))
#  i=i+1
#}

#NADP2NADPH = character()
#i=1
#while(i <= length(NADPH)){
#  NADP2NADPH = unique(c(NADP2NADPH, names(nodeID[grep(paste0("\\b",NADPH[i],"\\b"), #nodeID)])))
#  i=i+1
#}

#NAD.edges = matrix(rep("", (11*length(NAD2NADH))), ncol = 11)

#colnames(NAD.edges) = c("NR2NAM", "NR2NMN", "NAM2NMN", "NMN2NAD", "NAD2NADP", "NAD2NADH",
#                        "NA2NAMN", "NAMN2NAAD", "NAMN2NMN", "NAAD2NAD", "NADP2NADPH")

#NAD.edges[1:length(NR2NAM),    "NR2NAM"]     = NR2NAM
#NAD.edges[1:length(NR2NMN),    "NR2NMN"]     = NR2NMN
#NAD.edges[1:length(NAM2NMN),   "NAM2NMN"]    = NAM2NMN
#NAD.edges[1:length(NMN2NAD),   "NMN2NAD"]    = NMN2NAD
#NAD.edges[1:length(NAD2NADP),  "NAD2NADP"]   = NAD2NADP
#NAD.edges[1:length(NAD2NADH),  "NAD2NADH"]   = NAD2NADH
#NAD.edges[1:length(NA2NAMN),   "NA2NAMN"]    = NA2NAMN
#NAD.edges[1:length(NAMN2NAAD), "NAMN2NAAD"]  = NAMN2NAAD
#NAD.edges[1:length(NAMN2NMN),  "NAMN2NMN"]   = NAMN2NMN
#NAD.edges[1:length(NAAD2NAD),  "NAAD2NAD"]   = NAAD2NAD
#NAD.edges[1:length(NADP2NADPH),"NADP2NADPH"] = NADP2NADPH

#write.csv(NAD.edges, "/Users/johncsantiago/Documents/GitHub/WhartonLab/NAD.edges.csv")

```

```{r echo = F}

NAD.edges = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/NAD.edges.csv", row.names = 1)

NAD.nodes = data.frame(ID = character(),
                       x = numeric(),
                       y = numeric(),
                       names = character())

NAD.nodes[1,] = c('C03150', 0,  0,    "nicotinamide riboside+")
NAD.nodes[2,] = c('C00153', 0,  5,    "nicotinamide")
NAD.nodes[3,] = c('C00253', 0,  7.5,  "nicotinic acid")
NAD.nodes[4,] = c('C00455', 5,  2.5,  "NMN")
NAD.nodes[5,] = c('C01185', 5,  7.5,  "NAMN")
NAD.nodes[6,] = c('C00006', 10, -2.5, "NADP+")
NAD.nodes[7,] = c('C00003', 10, 2.5,  "NAD+")
NAD.nodes[8,] = c('C00857', 10, 7.5,  "NAAD")
NAD.nodes[9,] = c('C00005', 15, -2.5, "NADPH")
NAD.nodes[10,] = c('C00004', 15, 2.5,  "NADH")

NAD.nodes[11,] = c('C00806', 20, 10,  "tryptophan")
NAD.nodes[12,] = c('C02794', 5, 10,  "3-hydroxykynurenine")
NAD.nodes[13,] = c('C01718', 10, 10,  "kynurenine")
NAD.nodes[14,] = c('C01717', 10, 12.5,  "kynurenic acid")
NAD.nodes[15,] = c('C02700', 15, 10,  "N-Formylkynurenine")

row.names(NAD.nodes) = NAD.nodes$names

NAD.genes = data.frame( FBgn = character(),
                        Symbol = character(),
                        FC = numeric(),
                        FDR = numeric())

edges = data.frame(from = character(),
                   to = character(),
                   arrows = character(),
                   FBgn = character(),
                   Symbol = character())

#color = character(),
#                   FDR = numeric(),
#                   GR = numeric(),
#                   WT = numeric(),
#                   FC = numeric())

temp = GeneIDKey[GeneIDKey$ensembl %in% na.omit(NAD.edges$NR2NAM), c("FBgn", "Symbol")]
temp$from = "nicotinamide riboside+"
temp$to = "nicotinamide"
temp$arrows = "to"
temp = temp[,colnames(edges)]
edges = rbind(edges, temp)

temp = GeneIDKey[GeneIDKey$ensembl %in% na.omit(NAD.edges$NR2NMN), c("FBgn", "Symbol")]
temp$from = "nicotinamide riboside+"
temp$to = "NMN"
temp$arrows = "to"
temp = temp[,colnames(edges)]
edges = rbind(edges, temp)

temp = data.frame(FBgn = "",
                  Symbol = "")
temp$from = "nicotinamide"
temp$to = "NMN"
temp$arrows = "to"
temp = temp[,colnames(edges)]
edges = rbind(edges, temp)

temp = GeneIDKey[GeneIDKey$ensembl %in% na.omit(NAD.edges$NMN2NAD), c("FBgn", "Symbol")]
temp$from = "NMN"
temp$to = "NAD+"
temp$arrows = "to"
temp = temp[,colnames(edges)]
edges = rbind(edges, temp)

temp = GeneIDKey[GeneIDKey$ensembl %in% na.omit(NAD.edges$NAD2NADP), c("FBgn", "Symbol")]
temp$from = "NAD+"
temp$to = "NADP+"
temp$arrows = "to"
temp = temp[,colnames(edges)]
edges = rbind(edges, temp)

temp = data.frame(FBgn = "",
                  Symbol = "")
temp$from = "NAD+"
temp$to = "NADH"
temp$arrows = "to;from"
temp = temp[,colnames(edges)]
edges = rbind(edges, temp)

temp = data.frame(FBgn = "",
                  Symbol = "")
temp$from = "NADP+"
temp$to = "NADPH"
temp$arrows = "to;from"
temp = temp[,colnames(edges)]
edges = rbind(edges, temp)

temp = data.frame(FBgn = "",
                  Symbol = "")
temp$from = "nicotinic acid"
temp$to = "NAMN"
temp$arrows = "to"
temp = temp[,colnames(edges)]
edges = rbind(edges, temp)

temp = GeneIDKey[GeneIDKey$ensembl %in% na.omit(NAD.edges$NAMN2NAAD), c("FBgn", "Symbol")]
temp$from = "NAMN"
temp$to = "NAAD"
temp$arrows = "to"
temp = temp[,colnames(edges)]
edges = rbind(edges, temp)

temp = GeneIDKey[GeneIDKey$ensembl %in% na.omit(NAD.edges$NAAD2NAD), c("FBgn", "Symbol")]
temp$from = "NAAD"
temp$to = "NAD+"
temp$arrows = "to"
temp = temp[,colnames(edges)]
edges = rbind(edges, temp)

FBgn0003965 = data.frame(from = "tryptophan",
                         to = "N-Formylkynurenine",
                         arrows = "to",
                         FBgn = "FBgn0003965",
                         Symbol = "v")

FBgn0031821 = data.frame(from = "N-Formylkynurenine",
                         to = "kynurenine",
                         arrows = "to",
                         FBgn = "FBgn0031821",
                         Symbol = "KFase")

FBgn0000337 = data.frame(from = "kynurenine",
                         to = "3-hydroxykynurenine",
                         arrows = "to",
                         FBgn = "FBgn0000337",
                         Symbol = "cn")

FBgn0037955 = data.frame(from = "kynurenine",
                         to = "kynurenic acid",
                         arrows = "to",
                         FBgn = "FBgn0037955",
                         Symbol = "Kyat")

HKY2NAMN = data.frame(from = "3-hydroxykynurenine",
                         to = "NAMN",
                         arrows = "to",
                         FBgn = "",
                         Symbol = "")

edges = rbind(edges, 
              rbind(FBgn0003965, 
                    rbind(FBgn0031821, 
                          rbind(FBgn0000337, 
                                rbind(FBgn0037955, HKY2NAMN)))))

NAD.path = edges

```

```{r echo = F}
NAD.network.fun = function(comparison.T, comparison.M, cond1, cond2){
temp = colorRampPalette(c('royalblue',"dodgerblue",'red', "firebrick"))(401)

visedges = NAD.path
visedges$color = round(TKT.EdgeR.FC[NAD.path$FBgn, comparison.T],2)
##visedges$width = .5 + abs(visedges$color)
visedges$color[visedges$color < (-2)] = -2
visedges$color[visedges$color > (2)] = 2
visedges$color = (visedges$color * 100) +201
visedges$color = temp[visedges$color]
visedges$color[is.na(visedges$color)] = "black"

#visedges$symbol = GeneIDKey[visedges$FBgn, 'Symbol']
visedges$fdr = signif(TKT.EdgeR[visedges$FBgn, comparison.T], 2)
visedges$GR = round(mean.cpm[visedges$FBgn, cond1],1)
visedges$WT = round(mean.cpm[visedges$FBgn, cond2],1)
temp.width = data.frame(GR = visedges$GR,
                        WT = visedges$WT)
temp.width$max = apply(temp.width, 1, max)/50
temp.width$max[is.na(temp.width$max)] = .1
visedges$width = temp.width$max
visedges$width[visedges$width <= 1] = 1

visedges$title = paste0(visedges$Symbol,
                        "<br>FDR: ", visedges$fdr, 
                        "<br>",cond1," CPM: ", visedges$GR, 
                        "<br>",cond2," CPM: ", visedges$WT)
#visedges = na.omit(visedges)

node.names = unique(c(visedges$from, visedges$to))
node.data = data.frame(id = node.names,
                       color = 'dimgrey',
                       size = 10.5,
                       FC = NA,
                       pval = NA,
                       row.names = node.names)

#row.names(comparison.M)[row.names(comparison.M) == "cAMP(2',3') / cAMP(3',5')"] = 'cAMP'

node.colors = na.omit(comparison.M[intersect(row.names(comparison.M), node.names),])
node.pval = setNames(signif((node.colors$p),2), row.names(node.colors))
node.data[names(node.pval), 'pval'] = node.pval
node.colors = setNames(round(log2(node.colors$FC),2), row.names(node.colors))
node.data[names(node.colors), "FC"] = node.colors
node.data[names(node.colors),'size'] = node.colors
#node.data[names(node.colors),'size'] = 
#  (node.data[names(node.colors),'size'] + .5 + abs(min(node.colors)))*7.5
node.data[names(node.colors),'size'] = 
  (node.data[names(node.colors),'size'] + 3)*3.5

node.colors[node.colors < (-2)] = -2
node.colors[node.colors>2] = 2

node.colors = (node.colors*100)+201

node.colors = setNames(temp[node.colors], names(node.colors)) 
node.data[names(node.colors), 'color'] = node.colors

node.data$title = paste0(node.data$id,
                         "<br>FC: ", node.data$FC,
                         "<br>pval: ", node.data$pval)

node.data$group = 'metabolite'
#node.data[node.data$color == 'black', 'group'] = "Outside Path"
#node.data[node.data$color == 'dimgrey', 'group'] = "no data"

row.names(NAD.nodes) = NAD.nodes$names

# passing custom nodes and/or edges
lnodes <- data.frame(label = c("Increased", "Decreased", "No Data"), 
shape = 'dot', color = c("firebrick", "royalblue",'dimgrey'),
 title = "")

nodes = node.data

##nodes$color[nodes$color != "black"] = 'dimgrey'

colored.nodes = unique(c(interactions[,1], interactions[,3]))
colored.nodes = c(setdiff(row.names(nodes), c(colored.nodes, row.names(nodes)[nodes$color == "black"])))

#nodes[row.names(nodes) %in% colored.nodes, 'color'] = "gold"

visedges = visedges[,c('from', 'to', 'arrows', 'color', 'width', 'title')]
row.names(visedges) = c(1:nrow(visedges))

colored.edges = row.names(visedges[grep("CG16758", visedges$title),])
colored.edges = setdiff(row.names(visedges), colored.edges)

nodes$x = as.numeric(NAD.nodes[nodes$id, "x"])*50
nodes$y = as.numeric(NAD.nodes[nodes$id, "y"])*50

visNetwork(nodes,visedges,
                  main = paste0(cond1, " vs ", cond2))%>%
  visNodes(physics = F)%>%
  #visLayout(hierarchical=F, improvedLayout=T)%>%
  visPhysics(solver="forceAtlas2Based", forceAtlas2Based = list(gravitationalConstant=-50),enabled=T, minVelocity = .5)%>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE)%>%
  visEdges(arrows = visedges$arrows, physics = T)%>%
  #visGroups(groupname = "Outside Path", shape = 'box', physics=F, size = 25, font=list(color="white"))%>%
    visLegend(addNodes = lnodes, useGroups = FALSE, width = .115)

}

```

```{r echo = F}

comparison.T = "GRxWT.FC"
cond1 = 'GR.F'
cond2 = "WT.F"
comparison.M = GRxWT.C

NAD.network.fun(comparison.T, comparison.M, cond1, cond2)

```

```{r echo = F}

comparison.T = "GRxWT.MC"
cond1 = 'GR.M'
cond2 = "WT.M"
comparison.M = GRxWT.C

NAD.network.fun(comparison.T, comparison.M, cond1, cond2)

```

```{r echo = F}

comparison.T = "GRxWT.FDf"
cond1 = 'TktDfGR.F'
cond2 = "TktDfWT.F"
comparison.M = GRxWT.Df

NAD.network.fun(comparison.T, comparison.M, cond1, cond2)

```

```{r echo = F}

comparison.T = "GRxWT.MDf"
cond1 = 'TktDfGR.M'
cond2 = "TktDfWT.M"
comparison.M = GRxWT.Df

NAD.network.fun(comparison.T, comparison.M, cond1, cond2)

```

```{r echo = F}

comparison.T = "GRxWT.FOE"
cond1 = 'TktOEGR.F'
cond2 = "TktOEWT.F"
comparison.M = GRxWT.OE

NAD.network.fun(comparison.T, comparison.M, cond1, cond2)

```

```{r echo = F}

comparison.T = "GRxWT.MOE"
cond1 = 'TktOEGR.M'
cond2 = "TktOEWT.M"
comparison.M = GRxWT.OE

NAD.network.fun(comparison.T, comparison.M, cond1, cond2)

```