### 2025-01-30 SerpeSCdata.Rmd
final SC function addition

---
title: "Serpe SC data"
author: "John Santiago"
date: "2025-01-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F}

#library("hdf5r")
#library(SCopeLoomR)

#loom <- open_loom("/Users/johncsantiago/Downloads/5a3ba000_20180809__Davie_Janssens_Koldere_et_al_2018_Adult_Brain.HARMONY_SCENIC.loom", mode = "r")

##loom = open_loom("/Users/johncsantiago/Downloads/Goodwin_Fly_AdultVNC_elife54074.loom", mode = "r")

#dgem <- get_dgem(loom)
#close_loom(loom)
#dgem[1:5,100:105]

#clustering_annotations_df <- get_clustering_annotations(loom = loom)

```

```{r echo = F}

library(gplots)

##manually currated
serpe = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Nguyen Serpe 2024 scRNA-seq VNC.csv", row.names = 1)

serpe.glia = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Serpe 2024 GliaSubtypes.csv", row.names = 1)

serpe.mn = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Serpe 2024 MNClusters.csv", row.names = 1)

serpe.inter = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Serpe 2024 InterneuronClusters.csv", row.names = 1)

serpe.annotated.names = as.character(read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Nguyen Serpe 2024 scRNA-seq VNC.csv", row.names = 1, header = F)[1,])

serpe.glia.names = as.character(read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/Serpe 2024 GliaSubtypes.csv", row.names = 1, header = F)[1,])

```

```{r echo = F}

##Function to plot a specific gene expression levels in all cell types for the SC data
plot.serpe.allclusters = function(gene.name){

  ID.row = serpe[grep(gene.name, row.names(serpe)),]

  if(nrow(ID.row) < 50 & nrow(ID.row) > 0){
    if(nrow(ID.row) == 1 | length(intersect(gene.name, row.names(serpe))) == 1){

  serpe.annotated.gene = data.frame(level = as.numeric(serpe[gene.name,]),
                                    cluster = serpe.annotated.names,
                                    color = "darkgreen")

  serpe.mn.gene = data.frame(level = as.numeric(serpe.mn[gene.name,]),
                            cluster = colnames(serpe.mn),
                            color = "brown")

  serpe.inter.gene = data.frame(level = as.numeric(serpe.inter[gene.name,]),
                                cluster = colnames(serpe.inter),
                                color = "gold")

  serpe.glia.gene = data.frame(level = as.numeric(serpe.glia[gene.name,]),
                              cluster = serpe.glia.names,
                              color = "steelblue")

  serpe.all = rbind(serpe.annotated.gene,
                    rbind(serpe.mn.gene,
                          rbind(serpe.inter.gene,
                                serpe.glia.gene)))

  serpe.all$color[serpe.all$level < 1.5] = "darkgrey"

  xmax = max(serpe.all$level)
  if(xmax < 10){
    xmax = 10
  }

      par(mar = c(4, 13, 4, 8), xpd=TRUE)

      plot(x = NA,
           y = NA,
           type = 'n',
           ylim = c(4.5, 120),
           xlim = c(0, xmax),
           xaxt = "n",
           ylab="",
           yaxt = "n",
           xlab = "Mean Expression Levels",
           bty = "n",
           main = gene.name)

      barplot(height = serpe.all$level, 
              horiz = T, 
              #las = 2,
              xlim = c(0, xmax),
              names.arg = NA,
              border = T,
              col = serpe.all$color,
              add = T)

      mtext(side = 2,
            las = 2,
            at = .7+(1.2*(0:104)),
            text = serpe.all$cluster,
            col = serpe.all$color,
            line = 0)

      legend(x = xmax,
             y = 125,
             pch = 21,
             legend = c("Glia", "Interneurons", "Motor Neurons", "Manually Curated"),
             pt.bg = c("steelblue", "gold", "brown", "darkgreen"),
             bty = "n",
             xpd = T,
             pt.cex = 1.5,
             y.intersp = 1)

      rect(0 - (xmax/40), 0, 0 - (xmax/100), 18*1.2, col = 'darkgreen')
      rect(0 - (xmax/40), 18*1.2, 0 - (xmax/100), 46*1.2, col = 'brown')
      rect(0 - (xmax/40), 46*1.2, 0 - (xmax/100), 95*1.2, col = 'gold')
      rect(0 - (xmax/40), 95*1.2, 0 - (xmax/100), 105*1.2, col = 'steelblue')

    }

    #if(nrow(ID.row) != 1 | length(intersect(gene.name, row.names(serpe))) > 1){
    #  ID.row
    #}
  }
}

```

```{r echo = F, fig.width= 10, fig.height=16}

gene.name = "gbb"

plot.serpe.allclusters(gene.name)

```

```{r echo = F, fig.width= 10, fig.height=16}

gene.name = "dpp"

plot.serpe.allclusters(gene.name)

```

```{r echo = F, fig.width= 10, fig.height=16}

gene.name = "wit"

plot.serpe.allclusters(gene.name)

```

```{r echo = F, fig.width= 10, fig.height=16}

gene.name = "mav"

plot.serpe.allclusters(gene.name)

```

```{r echo = F, fig.width= 10, fig.height=16}

gene.name = "burs"

plot.serpe.allclusters(gene.name)

```

```{r echo = F, fig.width=16, fig.height=10}

heat.genes = c("dpp", "gbb", "put", "wit", "mav", "sax", "tkv", "repo", "ACC", "Ack-like")

serpe.heatmap = function(heat.genes){
heat.data = na.omit(cbind(serpe[heat.genes,],
                          cbind(serpe.mn[heat.genes,],
                                cbind(serpe.inter[heat.genes,], serpe.glia[heat.genes,]))))

heat.data = apply(heat.data, MARGIN = 2, as.numeric)
#heat.data = log2(heat.data)
row.names(heat.data) = heat.genes

#heat.data[heat.data == 0] = NA

Cluster.id = c(serpe.annotated.names,
               colnames(serpe.mn),
               colnames(serpe.inter),
               serpe.glia.names)

Cluster.labelcolor = c(rep("darkgreen", length(serpe.annotated.names)),
                       rep("brown", ncol(serpe.mn)),
                       rep("orange", ncol(serpe.inter)),
                       rep("steelblue", length(serpe.glia.names)))

heat.colors = colorRampPalette(c("grey95", "steelblue", "gold", "red3"))

numcols = 100

max.break = log2(max(heat.data))

if(max.break > 7){
  max.break = 7
}

color.breaks = 2^(c(0:(numcols))*(max.break/(numcols + 1)))

heatmap.2(heat.data,
        Colv = NA,
        #Rowv = NA,
        labCol = Cluster.id,
        scale = 'none',
        ColSideColors = Cluster.labelcolor,
        trace = "none",
        col = heat.colors(numcols),
        dendrogram = "none",
        na.color = "white",
        key.title = "Mean Expression",
        colCol = Cluster.labelcolor,
        margins = c(10, 5), 
        colsep = c(18, 46, 95),
        sepcolor = "black",
        breaks = color.breaks)
legend("left",
       title = "Cell Types",
       legend = c("Manually Currated", "Motor Neurons", "Interneurons", "Glia"),
       fill = c("darkgreen", "brown", "orange", "steelblue"),
       xpd = T)

}

#library(corrplot)
#corrplot(heat.data, is.corr = F, col =COL2("RdBu", 200))

```

```{r echo = F, fig.width=16, fig.height=10}

heat.genes = c("dpp", "gbb", "put", "wit", "mav", "sax", "tkv", "repo", "ACC", "Ack-like")##, "trio")#, "Rac1")

heat.data = na.omit(cbind(serpe.mn[heat.genes,],
                          cbind(serpe.inter[heat.genes,], serpe.glia[heat.genes,])))

heat.data = apply(heat.data, MARGIN = 2, as.numeric)
#heat.data = log2(heat.data)
row.names(heat.data) = heat.genes

#heat.data[heat.data == 0] = NA

Cluster.id = c(colnames(serpe.mn),
               colnames(serpe.inter),
               serpe.glia.names)

Cluster.labelcolor = c(rep("brown", ncol(serpe.mn)),
                       rep("orange", ncol(serpe.inter)),
                       rep("steelblue", length(serpe.glia.names)))

heat.colors = colorRampPalette(c("grey95", "steelblue", "gold", "red3"))

numcols = 100

max.break = log2(max(heat.data))

if(max.break > 7){
  max.break = 7
}

color.breaks = 2^(c(0:(numcols))*(max.break/(numcols + 1)))

heatmap.2(heat.data,
        Colv = NA,
        #Rowv = NA,
        labCol = Cluster.id,
        scale = 'none',
        ColSideColors = Cluster.labelcolor,
        trace = "none",
        col = heat.colors(numcols),
        dendrogram = "none",
        na.color = "white",
        key.title = "Mean Expression",
        colCol = Cluster.labelcolor,
        margins = c(10, 5), 
        colsep = c(28, 77),
        sepcolor = "black",
        breaks = color.breaks)
legend("left",
       title = "Cell Types",
       legend = c("Motor Neurons", "Interneurons", "Glia"),
       fill = c("brown", "orange", "steelblue"),
       xpd = T)
#library(corrplot)
#corrplot(heat.data, is.corr = F, col =COL2("RdBu", 200))

```

```{r echo = F, fig.width=16, fig.height=10}

heat.genes = c("dpp", "gbb", "put", "wit", "mav", "sax", "tkv", "repo", "ACC", "Ack-like")

heat.data = na.omit(serpe[heat.genes,])

heat.data = apply(heat.data, MARGIN = 2, as.numeric)
#heat.data = log2(heat.data)
row.names(heat.data) = heat.genes

#heat.data[heat.data == 0] = NA

Cluster.id = c(serpe.annotated.names)

Cluster.labelcolor = c(rep("darkgreen", length(serpe.annotated.names)))

heat.colors = colorRampPalette(c("grey95", "steelblue", "gold", "red3"))

numcols = 100

max.break = log2(max(heat.data))

if(max.break > 7){
  max.break = 7
}

color.breaks = 2^(c(0:(numcols))*(max.break/(numcols + 1)))

heatmap.2(heat.data,
        Colv = NA,
        #Rowv = NA,
        labCol = Cluster.id,
        scale = 'none',
        ColSideColors = Cluster.labelcolor,
        trace = "none",
        col = heat.colors(numcols),
        dendrogram = "none",
        na.color = "white",
        key.title = "Mean Expression",
        colCol = Cluster.labelcolor,
        margins = c(10, 5), 
        #colsep = c(18, 46, 95),
        #sepcolor = "black",
        breaks = color.breaks)
legend("left",
       title = "Cell Types",
       legend = c("Manually Currated"),
       fill = c("darkgreen"),
       xpd = T)
#library(corrplot)
#corrplot(heat.data, is.corr = F, col =COL2("RdBu", 200))

```