### 2025-02-03 myfile.Rmd
fixed link

<!DOCTYPE html>
<html
  lang="en"

  data-color-mode="auto" data-light-theme="light" data-dark-theme="dark"
  data-a11y-animated-images="system" data-a11y-link-underlines="true"

  >

  <head>
    <meta charset="utf-8">
  <link rel="dns-prefetch" href="https://github.githubassets.com">
  <link rel="dns-prefetch" href="https://avatars.githubusercontent.com">
  <link rel="dns-prefetch" href="https://github-cloud.s3.amazonaws.com">
  <link rel="dns-prefetch" href="https://user-images.githubusercontent.com/">
  <link rel="preconnect" href="https://github.githubassets.com" crossorigin>
  <link rel="preconnect" href="https://avatars.githubusercontent.com">

  <link crossorigin="anonymous" media="all" rel="stylesheet" href="https://github.githubassets.com/assets/light-7aa84bb7e11e.css" /><link crossorigin="anonymous" media="all" rel="stylesheet" href="https://github.githubassets.com/assets/dark-f65db3e8d171.css" /><link data-color-theme="dark_dimmed" crossorigin="anonymous" media="all" rel="stylesheet" data-href="https://github.githubassets.com/assets/dark_dimmed-a8258e3c6dda.css" /><link data-color-theme="dark_high_contrast" crossorigin="anonymous" media="all" rel="stylesheet" data-href="https://github.githubassets.com/assets/dark_high_contrast-7e97d834719c.css" /><link data-color-theme="dark_colorblind" crossorigin="anonymous" media="all" rel="stylesheet" data-href="https://github.githubassets.com/assets/dark_colorblind-01d869f460be.css" /><link data-color-theme="light_colorblind" crossorigin="anonymous" media="all" rel="stylesheet" data-href="https://github.githubassets.com/assets/light_colorblind-534f3e971240.css" /><link data-color-theme="light_high_contrast" crossorigin="anonymous" media="all" rel="stylesheet" data-href="https://github.githubassets.com/assets/light_high_contrast-a8cc7d138001.css" /><link data-color-theme="light_tritanopia" crossorigin="anonymous" media="all" rel="stylesheet" data-href="https://github.githubassets.com/assets/light_tritanopia-35e9dfdc4f9f.css" /><link data-color-theme="dark_tritanopia" crossorigin="anonymous" media="all" rel="stylesheet" data-href="https://github.githubassets.com/assets/dark_tritanopia-cf4cc5f62dfe.css" />

    <link crossorigin="anonymous" media="all" rel="stylesheet" href="https://github.githubassets.com/assets/primer-primitives-d9abecd14f1e.css" />
    <link crossorigin="anonymous" media="all" rel="stylesheet" href="https://github.githubassets.com/assets/primer-93aded0ee8a1.css" />
    <link crossorigin="anonymous" media="all" rel="stylesheet" href="https://github.githubassets.com/assets/global-d579f4a5b443.css" />
    <link crossorigin="anonymous" media="all" rel="stylesheet" href="https://github.githubassets.com/assets/github-8049f990d299.css" />
  <link crossorigin="anonymous" media="all" rel="stylesheet" href="https://github.githubassets.com/assets/repository-4fce88777fa8.css" />
<link crossorigin="anonymous" media="all" rel="stylesheet" href="https://github.githubassets.com/assets/code-0210be90f4d3.css" />

  <script type="application/json" id="client-env">{"locale":"en","featureFlags":["bypass_copilot_indexing_quota","copilot_new_references_ui","copilot_beta_features_opt_in","copilot_chat_repo_custom_instructions_preview","copilot_chat_retry_on_error","copilot_chat_persist_submitted_input","copilot_conversational_ux_history_refs","copilot_chat_shared_chat_input","copilot_chat_shared_topic_indicator","copilot_chat_shared_repo_sso_banner","copilot_editor_upsells","copilot_free_limited_user","copilot_implicit_context","copilot_no_floating_button","copilot_smell_icebreaker_ux","experimentation_azure_variant_endpoint","failbot_handle_non_errors","geojson_azure_maps","ghost_pilot_confidence_truncation_25","ghost_pilot_confidence_truncation_40","github_models_o3_mini_streaming","hovercard_accessibility","issues_react_remove_placeholders","issues_react_blur_item_picker_on_close","issues_react_include_bots_in_pickers","marketing_pages_search_explore_provider","react_keyboard_shortcuts_dialog","remove_child_patch","report_hydro_web_vitals","sample_network_conn_type","issues_react_create_milestone","issues_react_cache_fix_workaround","lifecycle_label_name_updates"]}</script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/wp-runtime-ea3cb3d11020.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_oddbird_popover-polyfill_dist_popover_js-9da652f58479.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_github_arianotify-polyfill_ariaNotify-polyfill_js-node_modules_github_mi-3abb8f-d7e6bc799724.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/ui_packages_failbot_failbot_ts-25697e0f4c47.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/environment-04ca94cb6e8a.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_primer_behaviors_dist_esm_index_mjs-0dbb79f97f8f.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_github_selector-observer_dist_index_esm_js-f690fd9ae3d5.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_github_relative-time-element_dist_index_js-f6da4b3fa34c.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_github_auto-complete-element_dist_index_js-node_modules_github_catalyst_-8e9f78-a74b4e0a8a6b.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_github_text-expander-element_dist_index_js-78748950cb0c.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_github_filter-input-element_dist_index_js-node_modules_github_remote-inp-b5f1d7-a1760ffda83d.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_github_markdown-toolbar-element_dist_index_js-ceef33f593fa.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_github_file-attachment-element_dist_index_js-node_modules_primer_view-co-c44a69-f0c8a795d1fd.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/github-elements-b487d8db3717.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/element-registry-0237e7f19fec.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_braintree_browser-detection_dist_browser-detection_js-node_modules_githu-bb80ec-72267f4e3ff9.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_lit-html_lit-html_js-be8cb88f481b.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_github_mini-throttle_dist_index_js-node_modules_morphdom_dist_morphdom-e-7c534c-a4a1922eb55f.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_github_turbo_dist_turbo_es2017-esm_js-e3cbe28f1638.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_github_remote-form_dist_index_js-node_modules_delegated-events_dist_inde-893f9f-6cf3320416b8.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_color-convert_index_js-e3180fe3bcb3.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_github_quote-selection_dist_index_js-node_modules_github_session-resume_-69cfcc-833249ee3034.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/ui_packages_updatable-content_updatable-content_ts-439f48470426.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/app_assets_modules_github_behaviors_task-list_ts-app_assets_modules_github_sso_ts-ui_packages-900dde-e8bb0b049301.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/app_assets_modules_github_sticky-scroll-into-view_ts-5316a27f9573.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/app_assets_modules_github_behaviors_ajax-error_ts-app_assets_modules_github_behaviors_include-87a4ae-0a6bb0ce2586.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/app_assets_modules_github_behaviors_commenting_edit_ts-app_assets_modules_github_behaviors_ht-83c235-42e06545c1fa.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/behaviors-6af14f7fc2ee.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_delegated-events_dist_index_js-node_modules_github_catalyst_lib_index_js-f6223d90c7ba.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/notifications-global-bd5619ba39ac.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_github_mini-throttle_dist_index_js-node_modules_github_catalyst_lib_inde-dbbea9-26cce2010167.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/code-menu-bfd00ac7c892.js"></script>

  <script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/primer-react-4f904282b093.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/react-core-5257d05d161f.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/react-lib-f09868a8643f.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/octicons-react-611691cca2f6.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_emotion_is-prop-valid_dist_emotion-is-prop-valid_esm_js-node_modules_emo-62da9f-2df2f32ec596.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_github_mini-throttle_dist_index_js-node_modules_stacktrace-parser_dist_s-e7dcdd-f7cc96ebae76.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_oddbird_popover-polyfill_dist_popover-fn_js-55fea94174bf.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_dompurify_dist_purify_js-b89b98661809.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_lodash-es__Stack_js-node_modules_lodash-es__Uint8Array_js-node_modules_l-4faaa6-4a736fde5c2f.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_lodash-es__baseIsEqual_js-8929eb9718d5.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_react-reverse-portal_dist_web_index_js-node_modules_primer_live-region-e-18c5ad-0b67a5a18b33.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/ui_packages_aria-live_aria-live_ts-ui_packages_promise-with-resolvers-polyfill_promise-with-r-014121-adb7f840e9e5.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/ui_packages_paths_index_ts-a30153db44f6.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/ui_packages_ref-selector_RefSelector_tsx-9f1e57bbb696.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/ui_packages_commit-attribution_index_ts-ui_packages_commit-checks-status_index_ts-ui_packages-530254-a7bdad87b7b4.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/ui_packages_code-view-shared_hooks_use-canonical-object_ts-ui_packages_code-view-shared_hooks-495298-d9d5efcddd99.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/ui_packages_copilot-chat_utils_copilot-chat-types_ts-ui_packages_repos-file-tree-view_repos-f-894be6-2f648fa9490a.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/app_assets_modules_github_blob-anchor_ts-ui_packages_code-nav_code-nav_ts-ui_packages_filter--8253c1-1320af121fa2.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/ui_packages_diffs_diff-parts_ts-b01c1ac6f203.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/react-code-view-e702702f4734.js"></script>
<link crossorigin="anonymous" media="all" rel="stylesheet" href="https://github.githubassets.com/assets/primer-react.1275b2aabc5faff7be57.module.css" />
<link crossorigin="anonymous" media="all" rel="stylesheet" href="https://github.githubassets.com/assets/react-code-view.ab7d8fac328c00e5e0cc.module.css" />

  <script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/notifications-subscriptions-menu-7eba7d01e0ba.js"></script>
<link crossorigin="anonymous" media="all" rel="stylesheet" href="https://github.githubassets.com/assets/primer-react.1275b2aabc5faff7be57.module.css" />
<link crossorigin="anonymous" media="all" rel="stylesheet" href="https://github.githubassets.com/assets/notifications-subscriptions-menu.1bcff9205c241e99cff2.module.css" />

  <title>WhartonLab/G85R GOandKEGGTermEnrichment.Rmd at master · johncsantiago/WhartonLab · GitHub</title>

  <meta name="route-pattern" content="/:user_id/:repository/blob/*name(/*path)" data-turbo-transient>
  <meta name="route-controller" content="blob" data-turbo-transient>
  <meta name="route-action" content="show" data-turbo-transient>

  <meta name="current-catalog-service-hash" content="f3abb0cc802f3d7b95fc8762b94bdcb13bf39634c40c357301c4aa1d67a256fb">

  <meta name="request-id" content="527A:2D1890:763472:A6139F:67A10520" data-pjax-transient="true"/><meta name="html-safe-nonce" content="5d4717541431eeb35e061369830ebefe101e79bcb17f34f9a298ee5ea100ed3c" data-pjax-transient="true"/><meta name="visitor-payload" content="eyJyZWZlcnJlciI6IiIsInJlcXVlc3RfaWQiOiI1MjdBOjJEMTg5MDo3NjM0NzI6QTYxMzlGOjY3QTEwNTIwIiwidmlzaXRvcl9pZCI6IjUwMjY1NDYzNjQ1MDk2NTIyNTYiLCJyZWdpb25fZWRnZSI6ImlhZCIsInJlZ2lvbl9yZW5kZXIiOiJpYWQifQ==" data-pjax-transient="true"/><meta name="visitor-hmac" content="72e28efc0fb87ca24293b07fa3e89a31531256fe0783789bbff5f0a89f7d6f09" data-pjax-transient="true"/>

    <meta name="hovercard-subject-tag" content="repository:652065041" data-turbo-transient>

  <meta name="github-keyboard-shortcuts" content="repository,source-code,file-tree,copilot" data-turbo-transient="true" />

  <meta name="selected-link" value="repo_source" data-turbo-transient>
  <link rel="assets" href="https://github.githubassets.com/">

    <meta name="google-site-verification" content="Apib7-x98H0j5cPqHWwSMm6dNU4GmODRoqxLiDzdx9I">

<meta name="octolytics-url" content="https://collector.github.com/github/collect" />

  <meta name="analytics-location" content="/&lt;user-name&gt;/&lt;repo-name&gt;/blob/show" data-turbo-transient="true" />

    <meta name="user-login" content="">

    <meta name="viewport" content="width=device-width">

      <meta name="description" content="Contribute to johncsantiago/WhartonLab development by creating an account on GitHub.">

      <link rel="search" type="application/opensearchdescription+xml" href="/opensearch.xml" title="GitHub">

    <link rel="fluid-icon" href="https://github.com/fluidicon.png" title="GitHub">
    <meta property="fb:app_id" content="1401488693436528">
    <meta name="apple-itunes-app" content="app-id=1477376905, app-argument=https://github.com/johncsantiago/WhartonLab/blob/master/G85R%20GOandKEGGTermEnrichment.Rmd" />

      <meta name="twitter:image" content="https://opengraph.githubassets.com/34050d6b9b05b9683112955dd9a2d8ff54e5d7545bfa645f3d53317d007cae20/johncsantiago/WhartonLab" /><meta name="twitter:site" content="@github" /><meta name="twitter:card" content="summary_large_image" /><meta name="twitter:title" content="WhartonLab/G85R GOandKEGGTermEnrichment.Rmd at master · johncsantiago/WhartonLab" /><meta name="twitter:description" content="Contribute to johncsantiago/WhartonLab development by creating an account on GitHub." />
  <meta property="og:image" content="https://opengraph.githubassets.com/34050d6b9b05b9683112955dd9a2d8ff54e5d7545bfa645f3d53317d007cae20/johncsantiago/WhartonLab" /><meta property="og:image:alt" content="Contribute to johncsantiago/WhartonLab development by creating an account on GitHub." /><meta property="og:image:width" content="1200" /><meta property="og:image:height" content="600" /><meta property="og:site_name" content="GitHub" /><meta property="og:type" content="object" /><meta property="og:title" content="WhartonLab/G85R GOandKEGGTermEnrichment.Rmd at master · johncsantiago/WhartonLab" /><meta property="og:url" content="https://github.com/johncsantiago/WhartonLab/blob/master/G85R%20GOandKEGGTermEnrichment.Rmd" /><meta property="og:description" content="Contribute to johncsantiago/WhartonLab development by creating an account on GitHub." />

      <meta name="hostname" content="github.com">

        <meta name="expected-hostname" content="github.com">

  <meta http-equiv="x-pjax-version" content="c6215f87fd63ac21525a969dd87c6591539570da1056d475d87280da437a52a9" data-turbo-track="reload">
  <meta http-equiv="x-pjax-csp-version" content="ace39c3b6632770952207593607e6e0be0db363435a8b877b1f96abe6430f345" data-turbo-track="reload">
  <meta http-equiv="x-pjax-css-version" content="1176077845c9479908e07e5cf5cf240c66928ac47340d56f202a3cfb9fc04dd3" data-turbo-track="reload">
  <meta http-equiv="x-pjax-js-version" content="923f4ffd6d2be276e59d6be4a93c84674e92d424b9c4051afc7a3806c89c6780" data-turbo-track="reload">

  <meta name="turbo-cache-control" content="no-preview" data-turbo-transient="">

      <meta name="turbo-cache-control" content="no-cache" data-turbo-transient>

    <meta data-hydrostats="publish">
  <meta name="go-import" content="github.com/johncsantiago/WhartonLab git https://github.com/johncsantiago/WhartonLab.git">

  <meta name="octolytics-dimension-user_id" content="61519162" /><meta name="octolytics-dimension-user_login" content="johncsantiago" /><meta name="octolytics-dimension-repository_id" content="652065041" /><meta name="octolytics-dimension-repository_nwo" content="johncsantiago/WhartonLab" /><meta name="octolytics-dimension-repository_public" content="true" /><meta name="octolytics-dimension-repository_is_fork" content="false" /><meta name="octolytics-dimension-repository_network_root_id" content="652065041" /><meta name="octolytics-dimension-repository_network_root_nwo" content="johncsantiago/WhartonLab" />

    <meta name="turbo-body-classes" content="logged-out env-production page-responsive">

  <meta name="browser-stats-url" content="https://api.github.com/_private/browser/stats">

  <meta name="browser-errors-url" content="https://api.github.com/_private/browser/errors">

  <link rel="mask-icon" href="https://github.githubassets.com/assets/pinned-octocat-093da3e6fa40.svg" color="#000000">
  <link rel="alternate icon" class="js-site-favicon" type="image/png" href="https://github.githubassets.com/favicons/favicon.png">
  <link rel="icon" class="js-site-favicon" type="image/svg+xml" href="https://github.githubassets.com/favicons/favicon.svg" data-base-href="https://github.githubassets.com/favicons/favicon">

<meta name="theme-color" content="#1e2327">
<meta name="color-scheme" content="light dark" />

  <link rel="manifest" href="/manifest.json" crossOrigin="use-credentials">

  </head>

  <body class="logged-out env-production page-responsive" style="word-wrap: break-word;">
    <div data-turbo-body class="logged-out env-production page-responsive" style="word-wrap: break-word;">

    <div class="position-relative header-wrapper js-header-wrapper ">
      <a href="#start-of-content" data-skip-target-assigned="false" class="px-2 py-4 color-bg-accent-emphasis color-fg-on-emphasis show-on-focus js-skip-to-content">Skip to content</a>

      <span data-view-component="true" class="progress-pjax-loader Progress position-fixed width-full">
    <span style="width: 0%;" data-view-component="true" class="Progress-item progress-pjax-loader-bar left-0 top-0 color-bg-accent-emphasis"></span>
</span>

      <script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/ui_packages_ui-commands_ui-commands_ts-ac3420ecd15f.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/keyboard-shortcuts-dialog-4ac6f562b41b.js"></script>
<link crossorigin="anonymous" media="all" rel="stylesheet" href="https://github.githubassets.com/assets/primer-react.1275b2aabc5faff7be57.module.css" />

<react-partial
  partial-name="keyboard-shortcuts-dialog"
  data-ssr="false"
  data-attempted-ssr="false"
>

  <script type="application/json" data-target="react-partial.embeddedData">{"props":{"docsUrl":"https://docs.github.com/get-started/accessibility/keyboard-shortcuts"}}</script>
  <div data-target="react-partial.reactRoot"></div>
</react-partial>

<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/vendors-node_modules_github_remote-form_dist_index_js-node_modules_delegated-events_dist_inde-94fd67-73b675cf164a.js"></script>
<script crossorigin="anonymous" defer="defer" type="application/javascript" src="https://github.githubassets.com/assets/sessions-fc22c41247ca.js"></script>
<header class="HeaderMktg header-logged-out js-details-container js-header Details f4 py-3" role="banner" data-is-top="true" data-color-mode=light data-light-theme=light data-dark-theme=dark>
  <h2 class="sr-only">Navigation Menu</h2>

  <button type="button" class="HeaderMktg-backdrop d-lg-none border-0 position-fixed top-0 left-0 width-full height-full js-details-target" aria-label="Toggle navigation">
    <span class="d-none">Toggle navigation</span>
  </button>

  <div class="d-flex flex-column flex-lg-row flex-items-center px-3 px-md-4 px-lg-5 height-full position-relative z-1">
    <div class="d-flex flex-justify-between flex-items-center width-full width-lg-auto">
      <div class="flex-1">
        <button aria-label="Toggle navigation" aria-expanded="false" type="button" data-view-component="true" class="js-details-target js-nav-padding-recalculate js-header-menu-toggle Button--link Button--medium Button d-lg-none color-fg-inherit p-1">  <span class="Button-content">
    <span class="Button-label"><div class="HeaderMenu-toggle-bar rounded my-1"></div>
            <div class="HeaderMenu-toggle-bar rounded my-1"></div>
            <div class="HeaderMenu-toggle-bar rounded my-1"></div></span>
  </span>
</button>
      </div>

      <a class="mr-lg-3 color-fg-inherit flex-order-2 js-prevent-focus-on-mobile-nav"
        href="/"
        aria-label="Homepage"
        data-analytics-event="{&quot;category&quot;:&quot;Marketing nav&quot;,&quot;action&quot;:&quot;click to go to homepage&quot;,&quot;label&quot;:&quot;ref_page:Marketing;ref_cta:Logomark;ref_loc:Header&quot;}">
        <svg height="32" aria-hidden="true" viewBox="0 0 24 24" version="1.1" width="32" data-view-component="true" class="octicon octicon-mark-github">
    <path d="M12.5.75C6.146.75 1 5.896 1 12.25c0 5.089 3.292 9.387 7.863 10.91.575.101.79-.244.79-.546 0-.273-.014-1.178-.014-2.142-2.889.532-3.636-.704-3.866-1.35-.13-.331-.69-1.352-1.18-1.625-.402-.216-.977-.748-.014-.762.906-.014 1.553.834 1.769 1.179 1.035 1.74 2.688 1.25 3.349.948.1-.747.402-1.25.733-1.538-2.559-.287-5.232-1.279-5.232-5.678 0-1.25.445-2.285 1.178-3.09-.115-.288-.517-1.467.115-3.048 0 0 .963-.302 3.163 1.179.92-.259 1.897-.388 2.875-.388.977 0 1.955.13 2.875.388 2.2-1.495 3.162-1.179 3.162-1.179.633 1.581.23 2.76.115 3.048.733.805 1.179 1.825 1.179 3.09 0 4.413-2.688 5.39-5.247 5.678.417.36.776 1.05.776 2.128 0 1.538-.014 2.774-.014 3.162 0 .302.216.662.79.547C20.709 21.637 24 17.324 24 12.25 24 5.896 18.854.75 12.5.75Z"></path>
</svg>
      </a>

      <div class="flex-1 flex-order-2 text-right">
          <a
            href="/login?return_to=https%3A%2F%2Fgithub.com%2Fjohncsantiago%2FWhartonLab%2Fblob%2Fmaster%2FG85R%2520GOandKEGGTermEnrichment.Rmd"
            class="HeaderMenu-link HeaderMenu-button d-inline-flex d-lg-none flex-order-1 f5 no-underline border color-border-default rounded-2 px-2 py-1 color-fg-inherit js-prevent-focus-on-mobile-nav"
            data-hydro-click="{&quot;event_type&quot;:&quot;authentication.click&quot;,&quot;payload&quot;:{&quot;location_in_page&quot;:&quot;site header menu&quot;,&quot;repository_id&quot;:null,&quot;auth_type&quot;:&quot;SIGN_UP&quot;,&quot;originating_url&quot;:&quot;https://github.com/johncsantiago/WhartonLab/blob/master/G85R%20GOandKEGGTermEnrichment.Rmd&quot;,&quot;user_id&quot;:null}}" data-hydro-click-hmac="8157e09adb62d5d215219a237bffb5c7f645d21a9e7ca38ff2b80dd99879c182"
            data-analytics-event="{&quot;category&quot;:&quot;Marketing nav&quot;,&quot;action&quot;:&quot;click to Sign in&quot;,&quot;label&quot;:&quot;ref_page:Marketing;ref_cta:Sign in;ref_loc:Header&quot;}"
          >
            Sign in
          </a>
      </div>
    </div>

    <div class="HeaderMenu js-header-menu height-fit position-lg-relative d-lg-flex flex-column flex-auto top-0">
      <div class="HeaderMenu-wrapper d-flex flex-column flex-self-start flex-lg-row flex-auto rounded rounded-lg-0">
          <nav class="HeaderMenu-nav" aria-label="Global">
            <ul class="d-lg-flex list-style-none">
                <li class="HeaderMenu-item position-relative flex-wrap flex-justify-between flex-items-center d-block d-lg-flex flex-lg-nowrap flex-lg-items-center js-details-container js-header-menu-item">
      <button type="button" class="HeaderMenu-link border-0 width-full width-lg-auto px-0 px-lg-2 py-lg-2 no-wrap d-flex flex-items-center flex-justify-between js-details-target" aria-expanded="false">
        Product
        <svg opacity="0.5" aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-chevron-down HeaderMenu-icon ml-1">
    <path d="M12.78 5.22a.749.749 0 0 1 0 1.06l-4.25 4.25a.749.749 0 0 1-1.06 0L3.22 6.28a.749.749 0 1 1 1.06-1.06L8 8.939l3.72-3.719a.749.749 0 0 1 1.06 0Z"></path>
</svg>
      </button>
      <div class="HeaderMenu-dropdown dropdown-menu rounded m-0 p-0 pt-2 pt-lg-4 position-relative position-lg-absolute left-0 left-lg-n3 pb-2 pb-lg-4 d-lg-flex flex-wrap dropdown-menu-wide">
          <div class="HeaderMenu-column px-lg-4 border-lg-right mb-4 mb-lg-0 pr-lg-7">
              <div class="border-bottom pb-3 pb-lg-0 border-lg-bottom-0">
                <ul class="list-style-none f5" >
                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary d-flex flex-items-center Link--has-description pb-lg-3" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;github_copilot&quot;,&quot;context&quot;:&quot;product&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;github_copilot_link_product_navbar&quot;}" href="https://github.com/features/copilot">
      <svg aria-hidden="true" height="24" viewBox="0 0 24 24" version="1.1" width="24" data-view-component="true" class="octicon octicon-copilot color-fg-subtle mr-3">
    <path d="M23.922 16.992c-.861 1.495-5.859 5.023-11.922 5.023-6.063 0-11.061-3.528-11.922-5.023A.641.641 0 0 1 0 16.736v-2.869a.841.841 0 0 1 .053-.22c.372-.935 1.347-2.292 2.605-2.656.167-.429.414-1.055.644-1.517a10.195 10.195 0 0 1-.052-1.086c0-1.331.282-2.499 1.132-3.368.397-.406.89-.717 1.474-.952 1.399-1.136 3.392-2.093 6.122-2.093 2.731 0 4.767.957 6.166 2.093.584.235 1.077.546 1.474.952.85.869 1.132 2.037 1.132 3.368 0 .368-.014.733-.052 1.086.23.462.477 1.088.644 1.517 1.258.364 2.233 1.721 2.605 2.656a.832.832 0 0 1 .053.22v2.869a.641.641 0 0 1-.078.256ZM12.172 11h-.344a4.323 4.323 0 0 1-.355.508C10.703 12.455 9.555 13 7.965 13c-1.725 0-2.989-.359-3.782-1.259a2.005 2.005 0 0 1-.085-.104L4 11.741v6.585c1.435.779 4.514 2.179 8 2.179 3.486 0 6.565-1.4 8-2.179v-6.585l-.098-.104s-.033.045-.085.104c-.793.9-2.057 1.259-3.782 1.259-1.59 0-2.738-.545-3.508-1.492a4.323 4.323 0 0 1-.355-.508h-.016.016Zm.641-2.935c.136 1.057.403 1.913.878 2.497.442.544 1.134.938 2.344.938 1.573 0 2.292-.337 2.657-.751.384-.435.558-1.15.558-2.361 0-1.14-.243-1.847-.705-2.319-.477-.488-1.319-.862-2.824-1.025-1.487-.161-2.192.138-2.533.529-.269.307-.437.808-.438 1.578v.021c0 .265.021.562.063.893Zm-1.626 0c.042-.331.063-.628.063-.894v-.02c-.001-.77-.169-1.271-.438-1.578-.341-.391-1.046-.69-2.533-.529-1.505.163-2.347.537-2.824 1.025-.462.472-.705 1.179-.705 2.319 0 1.211.175 1.926.558 2.361.365.414 1.084.751 2.657.751 1.21 0 1.902-.394 2.344-.938.475-.584.742-1.44.878-2.497Z"></path><path d="M14.5 14.25a1 1 0 0 1 1 1v2a1 1 0 0 1-2 0v-2a1 1 0 0 1 1-1Zm-5 0a1 1 0 0 1 1 1v2a1 1 0 0 1-2 0v-2a1 1 0 0 1 1-1Z"></path>
</svg>
      <div>
        <div class="color-fg-default h4">GitHub Copilot</div>
        Write better code with AI
      </div>

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary d-flex flex-items-center Link--has-description pb-lg-3" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;security&quot;,&quot;context&quot;:&quot;product&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;security_link_product_navbar&quot;}" href="https://github.com/features/security">
      <svg aria-hidden="true" height="24" viewBox="0 0 24 24" version="1.1" width="24" data-view-component="true" class="octicon octicon-shield-check color-fg-subtle mr-3">
    <path d="M16.53 9.78a.75.75 0 0 0-1.06-1.06L11 13.19l-1.97-1.97a.75.75 0 0 0-1.06 1.06l2.5 2.5a.75.75 0 0 0 1.06 0l5-5Z"></path><path d="m12.54.637 8.25 2.675A1.75 1.75 0 0 1 22 4.976V10c0 6.19-3.771 10.704-9.401 12.83a1.704 1.704 0 0 1-1.198 0C5.77 20.705 2 16.19 2 10V4.976c0-.758.489-1.43 1.21-1.664L11.46.637a1.748 1.748 0 0 1 1.08 0Zm-.617 1.426-8.25 2.676a.249.249 0 0 0-.173.237V10c0 5.46 3.28 9.483 8.43 11.426a.199.199 0 0 0 .14 0C17.22 19.483 20.5 15.461 20.5 10V4.976a.25.25 0 0 0-.173-.237l-8.25-2.676a.253.253 0 0 0-.154 0Z"></path>
</svg>
      <div>
        <div class="color-fg-default h4">Security</div>
        Find and fix vulnerabilities
      </div>

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary d-flex flex-items-center Link--has-description pb-lg-3" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;actions&quot;,&quot;context&quot;:&quot;product&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;actions_link_product_navbar&quot;}" href="https://github.com/features/actions">
      <svg aria-hidden="true" height="24" viewBox="0 0 24 24" version="1.1" width="24" data-view-component="true" class="octicon octicon-workflow color-fg-subtle mr-3">
    <path d="M1 3a2 2 0 0 1 2-2h6.5a2 2 0 0 1 2 2v6.5a2 2 0 0 1-2 2H7v4.063C7 16.355 7.644 17 8.438 17H12.5v-2.5a2 2 0 0 1 2-2H21a2 2 0 0 1 2 2V21a2 2 0 0 1-2 2h-6.5a2 2 0 0 1-2-2v-2.5H8.437A2.939 2.939 0 0 1 5.5 15.562V11.5H3a2 2 0 0 1-2-2Zm2-.5a.5.5 0 0 0-.5.5v6.5a.5.5 0 0 0 .5.5h6.5a.5.5 0 0 0 .5-.5V3a.5.5 0 0 0-.5-.5ZM14.5 14a.5.5 0 0 0-.5.5V21a.5.5 0 0 0 .5.5H21a.5.5 0 0 0 .5-.5v-6.5a.5.5 0 0 0-.5-.5Z"></path>
</svg>
      <div>
        <div class="color-fg-default h4">Actions</div>
        Automate any workflow
      </div>

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary d-flex flex-items-center Link--has-description pb-lg-3" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;codespaces&quot;,&quot;context&quot;:&quot;product&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;codespaces_link_product_navbar&quot;}" href="https://github.com/features/codespaces">
      <svg aria-hidden="true" height="24" viewBox="0 0 24 24" version="1.1" width="24" data-view-component="true" class="octicon octicon-codespaces color-fg-subtle mr-3">
    <path d="M3.5 3.75C3.5 2.784 4.284 2 5.25 2h13.5c.966 0 1.75.784 1.75 1.75v7.5A1.75 1.75 0 0 1 18.75 13H5.25a1.75 1.75 0 0 1-1.75-1.75Zm-2 12c0-.966.784-1.75 1.75-1.75h17.5c.966 0 1.75.784 1.75 1.75v4a1.75 1.75 0 0 1-1.75 1.75H3.25a1.75 1.75 0 0 1-1.75-1.75ZM5.25 3.5a.25.25 0 0 0-.25.25v7.5c0 .138.112.25.25.25h13.5a.25.25 0 0 0 .25-.25v-7.5a.25.25 0 0 0-.25-.25Zm-2 12a.25.25 0 0 0-.25.25v4c0 .138.112.25.25.25h17.5a.25.25 0 0 0 .25-.25v-4a.25.25 0 0 0-.25-.25Z"></path><path d="M10 17.75a.75.75 0 0 1 .75-.75h6.5a.75.75 0 0 1 0 1.5h-6.5a.75.75 0 0 1-.75-.75Zm-4 0a.75.75 0 0 1 .75-.75h.5a.75.75 0 0 1 0 1.5h-.5a.75.75 0 0 1-.75-.75Z"></path>
</svg>
      <div>
        <div class="color-fg-default h4">Codespaces</div>
        Instant dev environments
      </div>

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary d-flex flex-items-center Link--has-description pb-lg-3" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;issues&quot;,&quot;context&quot;:&quot;product&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;issues_link_product_navbar&quot;}" href="https://github.com/features/issues">
      <svg aria-hidden="true" height="24" viewBox="0 0 24 24" version="1.1" width="24" data-view-component="true" class="octicon octicon-issue-opened color-fg-subtle mr-3">
    <path d="M12 1c6.075 0 11 4.925 11 11s-4.925 11-11 11S1 18.075 1 12 5.925 1 12 1ZM2.5 12a9.5 9.5 0 0 0 9.5 9.5 9.5 9.5 0 0 0 9.5-9.5A9.5 9.5 0 0 0 12 2.5 9.5 9.5 0 0 0 2.5 12Zm9.5 2a2 2 0 1 1-.001-3.999A2 2 0 0 1 12 14Z"></path>
</svg>
      <div>
        <div class="color-fg-default h4">Issues</div>
        Plan and track work
      </div>

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary d-flex flex-items-center Link--has-description pb-lg-3" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;code_review&quot;,&quot;context&quot;:&quot;product&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;code_review_link_product_navbar&quot;}" href="https://github.com/features/code-review">
      <svg aria-hidden="true" height="24" viewBox="0 0 24 24" version="1.1" width="24" data-view-component="true" class="octicon octicon-code-review color-fg-subtle mr-3">
    <path d="M10.3 6.74a.75.75 0 0 1-.04 1.06l-2.908 2.7 2.908 2.7a.75.75 0 1 1-1.02 1.1l-3.5-3.25a.75.75 0 0 1 0-1.1l3.5-3.25a.75.75 0 0 1 1.06.04Zm3.44 1.06a.75.75 0 1 1 1.02-1.1l3.5 3.25a.75.75 0 0 1 0 1.1l-3.5 3.25a.75.75 0 1 1-1.02-1.1l2.908-2.7-2.908-2.7Z"></path><path d="M1.5 4.25c0-.966.784-1.75 1.75-1.75h17.5c.966 0 1.75.784 1.75 1.75v12.5a1.75 1.75 0 0 1-1.75 1.75h-9.69l-3.573 3.573A1.458 1.458 0 0 1 5 21.043V18.5H3.25a1.75 1.75 0 0 1-1.75-1.75ZM3.25 4a.25.25 0 0 0-.25.25v12.5c0 .138.112.25.25.25h2.5a.75.75 0 0 1 .75.75v3.19l3.72-3.72a.749.749 0 0 1 .53-.22h10a.25.25 0 0 0 .25-.25V4.25a.25.25 0 0 0-.25-.25Z"></path>
</svg>
      <div>
        <div class="color-fg-default h4">Code Review</div>
        Manage code changes
      </div>

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary d-flex flex-items-center Link--has-description pb-lg-3" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;discussions&quot;,&quot;context&quot;:&quot;product&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;discussions_link_product_navbar&quot;}" href="https://github.com/features/discussions">
      <svg aria-hidden="true" height="24" viewBox="0 0 24 24" version="1.1" width="24" data-view-component="true" class="octicon octicon-comment-discussion color-fg-subtle mr-3">
    <path d="M1.75 1h12.5c.966 0 1.75.784 1.75 1.75v9.5A1.75 1.75 0 0 1 14.25 14H8.061l-2.574 2.573A1.458 1.458 0 0 1 3 15.543V14H1.75A1.75 1.75 0 0 1 0 12.25v-9.5C0 1.784.784 1 1.75 1ZM1.5 2.75v9.5c0 .138.112.25.25.25h2a.75.75 0 0 1 .75.75v2.19l2.72-2.72a.749.749 0 0 1 .53-.22h6.5a.25.25 0 0 0 .25-.25v-9.5a.25.25 0 0 0-.25-.25H1.75a.25.25 0 0 0-.25.25Z"></path><path d="M22.5 8.75a.25.25 0 0 0-.25-.25h-3.5a.75.75 0 0 1 0-1.5h3.5c.966 0 1.75.784 1.75 1.75v9.5A1.75 1.75 0 0 1 22.25 20H21v1.543a1.457 1.457 0 0 1-2.487 1.03L15.939 20H10.75A1.75 1.75 0 0 1 9 18.25v-1.465a.75.75 0 0 1 1.5 0v1.465c0 .138.112.25.25.25h5.5a.75.75 0 0 1 .53.22l2.72 2.72v-2.19a.75.75 0 0 1 .75-.75h2a.25.25 0 0 0 .25-.25v-9.5Z"></path>
</svg>
      <div>
        <div class="color-fg-default h4">Discussions</div>
        Collaborate outside of code
      </div>

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary d-flex flex-items-center Link--has-description" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;code_search&quot;,&quot;context&quot;:&quot;product&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;code_search_link_product_navbar&quot;}" href="https://github.com/features/code-search">
      <svg aria-hidden="true" height="24" viewBox="0 0 24 24" version="1.1" width="24" data-view-component="true" class="octicon octicon-code-square color-fg-subtle mr-3">
    <path d="M10.3 8.24a.75.75 0 0 1-.04 1.06L7.352 12l2.908 2.7a.75.75 0 1 1-1.02 1.1l-3.5-3.25a.75.75 0 0 1 0-1.1l3.5-3.25a.75.75 0 0 1 1.06.04Zm3.44 1.06a.75.75 0 1 1 1.02-1.1l3.5 3.25a.75.75 0 0 1 0 1.1l-3.5 3.25a.75.75 0 1 1-1.02-1.1l2.908-2.7-2.908-2.7Z"></path><path d="M2 3.75C2 2.784 2.784 2 3.75 2h16.5c.966 0 1.75.784 1.75 1.75v16.5A1.75 1.75 0 0 1 20.25 22H3.75A1.75 1.75 0 0 1 2 20.25Zm1.75-.25a.25.25 0 0 0-.25.25v16.5c0 .138.112.25.25.25h16.5a.25.25 0 0 0 .25-.25V3.75a.25.25 0 0 0-.25-.25Z"></path>
</svg>
      <div>
        <div class="color-fg-default h4">Code Search</div>
        Find more, search less
      </div>

</a></li>

                </ul>
              </div>
          </div>
          <div class="HeaderMenu-column px-lg-4">
              <div class="border-bottom pb-3 pb-lg-0 border-lg-bottom-0 border-bottom-0">
                    <span class="d-block h4 color-fg-default my-1" id="product-explore-heading">Explore</span>
                <ul class="list-style-none f5" aria-labelledby="product-explore-heading">
                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;all_features&quot;,&quot;context&quot;:&quot;product&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;all_features_link_product_navbar&quot;}" href="https://github.com/features">
      All features

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary Link--external" target="_blank" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;documentation&quot;,&quot;context&quot;:&quot;product&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;documentation_link_product_navbar&quot;}" href="https://docs.github.com">
      Documentation

    <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-link-external HeaderMenu-external-icon color-fg-subtle">
    <path d="M3.75 2h3.5a.75.75 0 0 1 0 1.5h-3.5a.25.25 0 0 0-.25.25v8.5c0 .138.112.25.25.25h8.5a.25.25 0 0 0 .25-.25v-3.5a.75.75 0 0 1 1.5 0v3.5A1.75 1.75 0 0 1 12.25 14h-8.5A1.75 1.75 0 0 1 2 12.25v-8.5C2 2.784 2.784 2 3.75 2Zm6.854-1h4.146a.25.25 0 0 1 .25.25v4.146a.25.25 0 0 1-.427.177L13.03 4.03 9.28 7.78a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042l3.75-3.75-1.543-1.543A.25.25 0 0 1 10.604 1Z"></path>
</svg>
</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary Link--external" target="_blank" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;github_skills&quot;,&quot;context&quot;:&quot;product&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;github_skills_link_product_navbar&quot;}" href="https://skills.github.com">
      GitHub Skills

    <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-link-external HeaderMenu-external-icon color-fg-subtle">
    <path d="M3.75 2h3.5a.75.75 0 0 1 0 1.5h-3.5a.25.25 0 0 0-.25.25v8.5c0 .138.112.25.25.25h8.5a.25.25 0 0 0 .25-.25v-3.5a.75.75 0 0 1 1.5 0v3.5A1.75 1.75 0 0 1 12.25 14h-8.5A1.75 1.75 0 0 1 2 12.25v-8.5C2 2.784 2.784 2 3.75 2Zm6.854-1h4.146a.25.25 0 0 1 .25.25v4.146a.25.25 0 0 1-.427.177L13.03 4.03 9.28 7.78a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042l3.75-3.75-1.543-1.543A.25.25 0 0 1 10.604 1Z"></path>
</svg>
</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary Link--external" target="_blank" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;blog&quot;,&quot;context&quot;:&quot;product&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;blog_link_product_navbar&quot;}" href="https://github.blog">
      Blog

    <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-link-external HeaderMenu-external-icon color-fg-subtle">
    <path d="M3.75 2h3.5a.75.75 0 0 1 0 1.5h-3.5a.25.25 0 0 0-.25.25v8.5c0 .138.112.25.25.25h8.5a.25.25 0 0 0 .25-.25v-3.5a.75.75 0 0 1 1.5 0v3.5A1.75 1.75 0 0 1 12.25 14h-8.5A1.75 1.75 0 0 1 2 12.25v-8.5C2 2.784 2.784 2 3.75 2Zm6.854-1h4.146a.25.25 0 0 1 .25.25v4.146a.25.25 0 0 1-.427.177L13.03 4.03 9.28 7.78a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042l3.75-3.75-1.543-1.543A.25.25 0 0 1 10.604 1Z"></path>
</svg>
</a></li>

                </ul>
              </div>
          </div>
      </div>
</li>

                  <li class="HeaderMenu-item position-relative flex-wrap flex-justify-between flex-items-center d-block d-lg-flex flex-lg-nowrap flex-lg-items-center js-details-container js-header-menu-item">
      <button type="button" class="HeaderMenu-link border-0 width-full width-lg-auto px-0 px-lg-2 py-lg-2 no-wrap d-flex flex-items-center flex-justify-between js-details-target" aria-expanded="false">
        Solutions
        <svg opacity="0.5" aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-chevron-down HeaderMenu-icon ml-1">
    <path d="M12.78 5.22a.749.749 0 0 1 0 1.06l-4.25 4.25a.749.749 0 0 1-1.06 0L3.22 6.28a.749.749 0 1 1 1.06-1.06L8 8.939l3.72-3.719a.749.749 0 0 1 1.06 0Z"></path>
</svg>
      </button>
      <div class="HeaderMenu-dropdown dropdown-menu rounded m-0 p-0 pt-2 pt-lg-4 position-relative position-lg-absolute left-0 left-lg-n3 d-lg-flex flex-wrap dropdown-menu-wide">
          <div class="HeaderMenu-column px-lg-4 border-lg-right mb-4 mb-lg-0 pr-lg-7">
              <div class="border-bottom pb-3 pb-lg-0 border-lg-bottom-0 pb-lg-3 mb-3 mb-lg-0">
                    <span class="d-block h4 color-fg-default my-1" id="solutions-by-company-size-heading">By company size</span>
                <ul class="list-style-none f5" aria-labelledby="solutions-by-company-size-heading">
                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;enterprises&quot;,&quot;context&quot;:&quot;solutions&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;enterprises_link_solutions_navbar&quot;}" href="https://github.com/enterprise">
      Enterprises

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;small_and_medium_teams&quot;,&quot;context&quot;:&quot;solutions&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;small_and_medium_teams_link_solutions_navbar&quot;}" href="https://github.com/team">
      Small and medium teams

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;startups&quot;,&quot;context&quot;:&quot;solutions&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;startups_link_solutions_navbar&quot;}" href="https://github.com/enterprise/startups">
      Startups

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;nonprofits&quot;,&quot;context&quot;:&quot;solutions&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;nonprofits_link_solutions_navbar&quot;}" href="/solutions/industry/nonprofits">
      Nonprofits

</a></li>

                </ul>
              </div>
              <div class="border-bottom pb-3 pb-lg-0 border-lg-bottom-0">
                    <span class="d-block h4 color-fg-default my-1" id="solutions-by-use-case-heading">By use case</span>
                <ul class="list-style-none f5" aria-labelledby="solutions-by-use-case-heading">
                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;devsecops&quot;,&quot;context&quot;:&quot;solutions&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;devsecops_link_solutions_navbar&quot;}" href="/solutions/use-case/devsecops">
      DevSecOps

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;devops&quot;,&quot;context&quot;:&quot;solutions&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;devops_link_solutions_navbar&quot;}" href="/solutions/use-case/devops">
      DevOps

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;ci_cd&quot;,&quot;context&quot;:&quot;solutions&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;ci_cd_link_solutions_navbar&quot;}" href="/solutions/use-case/ci-cd">
      CI/CD

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;view_all_use_cases&quot;,&quot;context&quot;:&quot;solutions&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;view_all_use_cases_link_solutions_navbar&quot;}" href="/solutions/use-case">
      View all use cases

</a></li>

                </ul>
              </div>
          </div>
          <div class="HeaderMenu-column px-lg-4">
              <div class="border-bottom pb-3 pb-lg-0 border-lg-bottom-0">
                    <span class="d-block h4 color-fg-default my-1" id="solutions-by-industry-heading">By industry</span>
                <ul class="list-style-none f5" aria-labelledby="solutions-by-industry-heading">
                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;healthcare&quot;,&quot;context&quot;:&quot;solutions&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;healthcare_link_solutions_navbar&quot;}" href="/solutions/industry/healthcare">
      Healthcare

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;financial_services&quot;,&quot;context&quot;:&quot;solutions&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;financial_services_link_solutions_navbar&quot;}" href="/solutions/industry/financial-services">
      Financial services

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;manufacturing&quot;,&quot;context&quot;:&quot;solutions&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;manufacturing_link_solutions_navbar&quot;}" href="/solutions/industry/manufacturing">
      Manufacturing

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;government&quot;,&quot;context&quot;:&quot;solutions&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;government_link_solutions_navbar&quot;}" href="/solutions/industry/government">
      Government

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;view_all_industries&quot;,&quot;context&quot;:&quot;solutions&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;view_all_industries_link_solutions_navbar&quot;}" href="/solutions/industry">
      View all industries

</a></li>

                </ul>
              </div>
          </div>
         <div class="HeaderMenu-trailing-link rounded-bottom-2 flex-shrink-0 mt-lg-4 px-lg-4 py-4 py-lg-3 f5 text-semibold">
            <a href="/solutions">
              View all solutions
              <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-chevron-right HeaderMenu-trailing-link-icon">
    <path d="M6.22 3.22a.75.75 0 0 1 1.06 0l4.25 4.25a.75.75 0 0 1 0 1.06l-4.25 4.25a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042L9.94 8 6.22 4.28a.75.75 0 0 1 0-1.06Z"></path>
</svg>
</a>         </div>
      </div>
</li>

                    <li class="HeaderMenu-item position-relative flex-wrap flex-justify-between flex-items-center d-block d-lg-flex flex-lg-nowrap flex-lg-items-center js-details-container js-header-menu-item">
      <button type="button" class="HeaderMenu-link border-0 width-full width-lg-auto px-0 px-lg-2 py-lg-2 no-wrap d-flex flex-items-center flex-justify-between js-details-target" aria-expanded="false">
        Resources
        <svg opacity="0.5" aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-chevron-down HeaderMenu-icon ml-1">
    <path d="M12.78 5.22a.749.749 0 0 1 0 1.06l-4.25 4.25a.749.749 0 0 1-1.06 0L3.22 6.28a.749.749 0 1 1 1.06-1.06L8 8.939l3.72-3.719a.749.749 0 0 1 1.06 0Z"></path>
</svg>
      </button>
      <div class="HeaderMenu-dropdown dropdown-menu rounded m-0 p-0 pt-2 pt-lg-4 position-relative position-lg-absolute left-0 left-lg-n3 pb-2 pb-lg-4 d-lg-flex flex-wrap dropdown-menu-wide">
          <div class="HeaderMenu-column px-lg-4 border-lg-right mb-4 mb-lg-0 pr-lg-7">
              <div class="border-bottom pb-3 pb-lg-0 border-lg-bottom-0">
                    <span class="d-block h4 color-fg-default my-1" id="resources-topics-heading">Topics</span>
                <ul class="list-style-none f5" aria-labelledby="resources-topics-heading">
                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;ai&quot;,&quot;context&quot;:&quot;resources&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;ai_link_resources_navbar&quot;}" href="/resources/articles/ai">
      AI

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;devops&quot;,&quot;context&quot;:&quot;resources&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;devops_link_resources_navbar&quot;}" href="/resources/articles/devops">
      DevOps

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;security&quot;,&quot;context&quot;:&quot;resources&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;security_link_resources_navbar&quot;}" href="/resources/articles/security">
      Security

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;software_development&quot;,&quot;context&quot;:&quot;resources&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;software_development_link_resources_navbar&quot;}" href="/resources/articles/software-development">
      Software Development

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;view_all&quot;,&quot;context&quot;:&quot;resources&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;view_all_link_resources_navbar&quot;}" href="/resources/articles">
      View all

</a></li>

                </ul>
              </div>
          </div>
          <div class="HeaderMenu-column px-lg-4">
              <div class="border-bottom pb-3 pb-lg-0 border-lg-bottom-0 border-bottom-0">
                    <span class="d-block h4 color-fg-default my-1" id="resources-explore-heading">Explore</span>
                <ul class="list-style-none f5" aria-labelledby="resources-explore-heading">
                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary Link--external" target="_blank" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;learning_pathways&quot;,&quot;context&quot;:&quot;resources&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;learning_pathways_link_resources_navbar&quot;}" href="https://resources.github.com/learn/pathways">
      Learning Pathways

    <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-link-external HeaderMenu-external-icon color-fg-subtle">
    <path d="M3.75 2h3.5a.75.75 0 0 1 0 1.5h-3.5a.25.25 0 0 0-.25.25v8.5c0 .138.112.25.25.25h8.5a.25.25 0 0 0 .25-.25v-3.5a.75.75 0 0 1 1.5 0v3.5A1.75 1.75 0 0 1 12.25 14h-8.5A1.75 1.75 0 0 1 2 12.25v-8.5C2 2.784 2.784 2 3.75 2Zm6.854-1h4.146a.25.25 0 0 1 .25.25v4.146a.25.25 0 0 1-.427.177L13.03 4.03 9.28 7.78a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042l3.75-3.75-1.543-1.543A.25.25 0 0 1 10.604 1Z"></path>
</svg>
</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary Link--external" target="_blank" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;white_papers_ebooks_webinars&quot;,&quot;context&quot;:&quot;resources&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;white_papers_ebooks_webinars_link_resources_navbar&quot;}" href="https://resources.github.com">
      White papers, Ebooks, Webinars

    <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-link-external HeaderMenu-external-icon color-fg-subtle">
    <path d="M3.75 2h3.5a.75.75 0 0 1 0 1.5h-3.5a.25.25 0 0 0-.25.25v8.5c0 .138.112.25.25.25h8.5a.25.25 0 0 0 .25-.25v-3.5a.75.75 0 0 1 1.5 0v3.5A1.75 1.75 0 0 1 12.25 14h-8.5A1.75 1.75 0 0 1 2 12.25v-8.5C2 2.784 2.784 2 3.75 2Zm6.854-1h4.146a.25.25 0 0 1 .25.25v4.146a.25.25 0 0 1-.427.177L13.03 4.03 9.28 7.78a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042l3.75-3.75-1.543-1.543A.25.25 0 0 1 10.604 1Z"></path>
</svg>
</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;customer_stories&quot;,&quot;context&quot;:&quot;resources&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;customer_stories_link_resources_navbar&quot;}" href="https://github.com/customer-stories">
      Customer Stories

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary Link--external" target="_blank" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;partners&quot;,&quot;context&quot;:&quot;resources&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;partners_link_resources_navbar&quot;}" href="https://partner.github.com">
      Partners

    <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-link-external HeaderMenu-external-icon color-fg-subtle">
    <path d="M3.75 2h3.5a.75.75 0 0 1 0 1.5h-3.5a.25.25 0 0 0-.25.25v8.5c0 .138.112.25.25.25h8.5a.25.25 0 0 0 .25-.25v-3.5a.75.75 0 0 1 1.5 0v3.5A1.75 1.75 0 0 1 12.25 14h-8.5A1.75 1.75 0 0 1 2 12.25v-8.5C2 2.784 2.784 2 3.75 2Zm6.854-1h4.146a.25.25 0 0 1 .25.25v4.146a.25.25 0 0 1-.427.177L13.03 4.03 9.28 7.78a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042l3.75-3.75-1.543-1.543A.25.25 0 0 1 10.604 1Z"></path>
</svg>
</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;executive_insights&quot;,&quot;context&quot;:&quot;resources&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;executive_insights_link_resources_navbar&quot;}" href="https://github.com/solutions/executive-insights">
      Executive Insights

</a></li>

                </ul>
              </div>
          </div>
      </div>
</li>

                <li class="HeaderMenu-item position-relative flex-wrap flex-justify-between flex-items-center d-block d-lg-flex flex-lg-nowrap flex-lg-items-center js-details-container js-header-menu-item">
      <button type="button" class="HeaderMenu-link border-0 width-full width-lg-auto px-0 px-lg-2 py-lg-2 no-wrap d-flex flex-items-center flex-justify-between js-details-target" aria-expanded="false">
        Open Source
        <svg opacity="0.5" aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-chevron-down HeaderMenu-icon ml-1">
    <path d="M12.78 5.22a.749.749 0 0 1 0 1.06l-4.25 4.25a.749.749 0 0 1-1.06 0L3.22 6.28a.749.749 0 1 1 1.06-1.06L8 8.939l3.72-3.719a.749.749 0 0 1 1.06 0Z"></path>
</svg>
      </button>
      <div class="HeaderMenu-dropdown dropdown-menu rounded m-0 p-0 pt-2 pt-lg-4 position-relative position-lg-absolute left-0 left-lg-n3 pb-2 pb-lg-4 px-lg-4">
          <div class="HeaderMenu-column">
              <div class="border-bottom pb-3 pb-lg-0 pb-lg-3 mb-3 mb-lg-0 mb-lg-3">
                <ul class="list-style-none f5" >
                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary d-flex flex-items-center Link--has-description" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;github_sponsors&quot;,&quot;context&quot;:&quot;open_source&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;github_sponsors_link_open_source_navbar&quot;}" href="/sponsors">

      <div>
        <div class="color-fg-default h4">GitHub Sponsors</div>
        Fund open source developers
      </div>

</a></li>

                </ul>
              </div>
              <div class="border-bottom pb-3 pb-lg-0 pb-lg-3 mb-3 mb-lg-0 mb-lg-3">
                <ul class="list-style-none f5" >
                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary d-flex flex-items-center Link--has-description" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;the_readme_project&quot;,&quot;context&quot;:&quot;open_source&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;the_readme_project_link_open_source_navbar&quot;}" href="https://github.com/readme">

      <div>
        <div class="color-fg-default h4">The ReadME Project</div>
        GitHub community articles
      </div>

</a></li>

                </ul>
              </div>
              <div class="border-bottom pb-3 pb-lg-0 border-bottom-0">
                    <span class="d-block h4 color-fg-default my-1" id="open-source-repositories-heading">Repositories</span>
                <ul class="list-style-none f5" aria-labelledby="open-source-repositories-heading">
                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;topics&quot;,&quot;context&quot;:&quot;open_source&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;topics_link_open_source_navbar&quot;}" href="https://github.com/topics">
      Topics

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;trending&quot;,&quot;context&quot;:&quot;open_source&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;trending_link_open_source_navbar&quot;}" href="https://github.com/trending">
      Trending

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;collections&quot;,&quot;context&quot;:&quot;open_source&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;collections_link_open_source_navbar&quot;}" href="https://github.com/collections">
      Collections

</a></li>

                </ul>
              </div>
          </div>
      </div>
</li>

                <li class="HeaderMenu-item position-relative flex-wrap flex-justify-between flex-items-center d-block d-lg-flex flex-lg-nowrap flex-lg-items-center js-details-container js-header-menu-item">
      <button type="button" class="HeaderMenu-link border-0 width-full width-lg-auto px-0 px-lg-2 py-lg-2 no-wrap d-flex flex-items-center flex-justify-between js-details-target" aria-expanded="false">
        Enterprise
        <svg opacity="0.5" aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-chevron-down HeaderMenu-icon ml-1">
    <path d="M12.78 5.22a.749.749 0 0 1 0 1.06l-4.25 4.25a.749.749 0 0 1-1.06 0L3.22 6.28a.749.749 0 1 1 1.06-1.06L8 8.939l3.72-3.719a.749.749 0 0 1 1.06 0Z"></path>
</svg>
      </button>
      <div class="HeaderMenu-dropdown dropdown-menu rounded m-0 p-0 pt-2 pt-lg-4 position-relative position-lg-absolute left-0 left-lg-n3 pb-2 pb-lg-4 px-lg-4">
          <div class="HeaderMenu-column">
              <div class="border-bottom pb-3 pb-lg-0 pb-lg-3 mb-3 mb-lg-0 mb-lg-3">
                <ul class="list-style-none f5" >
                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary d-flex flex-items-center Link--has-description" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;enterprise_platform&quot;,&quot;context&quot;:&quot;enterprise&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;enterprise_platform_link_enterprise_navbar&quot;}" href="/enterprise">
      <svg aria-hidden="true" height="24" viewBox="0 0 24 24" version="1.1" width="24" data-view-component="true" class="octicon octicon-stack color-fg-subtle mr-3">
    <path d="M11.063 1.456a1.749 1.749 0 0 1 1.874 0l8.383 5.316a1.751 1.751 0 0 1 0 2.956l-8.383 5.316a1.749 1.749 0 0 1-1.874 0L2.68 9.728a1.751 1.751 0 0 1 0-2.956Zm1.071 1.267a.25.25 0 0 0-.268 0L3.483 8.039a.25.25 0 0 0 0 .422l8.383 5.316a.25.25 0 0 0 .268 0l8.383-5.316a.25.25 0 0 0 0-.422Z"></path><path d="M1.867 12.324a.75.75 0 0 1 1.035-.232l8.964 5.685a.25.25 0 0 0 .268 0l8.964-5.685a.75.75 0 0 1 .804 1.267l-8.965 5.685a1.749 1.749 0 0 1-1.874 0l-8.965-5.685a.75.75 0 0 1-.231-1.035Z"></path><path d="M1.867 16.324a.75.75 0 0 1 1.035-.232l8.964 5.685a.25.25 0 0 0 .268 0l8.964-5.685a.75.75 0 0 1 .804 1.267l-8.965 5.685a1.749 1.749 0 0 1-1.874 0l-8.965-5.685a.75.75 0 0 1-.231-1.035Z"></path>
</svg>
      <div>
        <div class="color-fg-default h4">Enterprise platform</div>
        AI-powered developer platform
      </div>

</a></li>

                </ul>
              </div>
              <div class="border-bottom pb-3 pb-lg-0 border-bottom-0">
                    <span class="d-block h4 color-fg-default my-1" id="enterprise-available-add-ons-heading">Available add-ons</span>
                <ul class="list-style-none f5" aria-labelledby="enterprise-available-add-ons-heading">
                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary d-flex flex-items-center Link--has-description pb-lg-3" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;advanced_security&quot;,&quot;context&quot;:&quot;enterprise&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;advanced_security_link_enterprise_navbar&quot;}" href="https://github.com/enterprise/advanced-security">
      <svg aria-hidden="true" height="24" viewBox="0 0 24 24" version="1.1" width="24" data-view-component="true" class="octicon octicon-shield-check color-fg-subtle mr-3">
    <path d="M16.53 9.78a.75.75 0 0 0-1.06-1.06L11 13.19l-1.97-1.97a.75.75 0 0 0-1.06 1.06l2.5 2.5a.75.75 0 0 0 1.06 0l5-5Z"></path><path d="m12.54.637 8.25 2.675A1.75 1.75 0 0 1 22 4.976V10c0 6.19-3.771 10.704-9.401 12.83a1.704 1.704 0 0 1-1.198 0C5.77 20.705 2 16.19 2 10V4.976c0-.758.489-1.43 1.21-1.664L11.46.637a1.748 1.748 0 0 1 1.08 0Zm-.617 1.426-8.25 2.676a.249.249 0 0 0-.173.237V10c0 5.46 3.28 9.483 8.43 11.426a.199.199 0 0 0 .14 0C17.22 19.483 20.5 15.461 20.5 10V4.976a.25.25 0 0 0-.173-.237l-8.25-2.676a.253.253 0 0 0-.154 0Z"></path>
</svg>
      <div>
        <div class="color-fg-default h4">Advanced Security</div>
        Enterprise-grade security features
      </div>

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary d-flex flex-items-center Link--has-description pb-lg-3" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;github_copilot&quot;,&quot;context&quot;:&quot;enterprise&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;github_copilot_link_enterprise_navbar&quot;}" href="/features/copilot#enterprise">
      <svg aria-hidden="true" height="24" viewBox="0 0 24 24" version="1.1" width="24" data-view-component="true" class="octicon octicon-copilot color-fg-subtle mr-3">
    <path d="M23.922 16.992c-.861 1.495-5.859 5.023-11.922 5.023-6.063 0-11.061-3.528-11.922-5.023A.641.641 0 0 1 0 16.736v-2.869a.841.841 0 0 1 .053-.22c.372-.935 1.347-2.292 2.605-2.656.167-.429.414-1.055.644-1.517a10.195 10.195 0 0 1-.052-1.086c0-1.331.282-2.499 1.132-3.368.397-.406.89-.717 1.474-.952 1.399-1.136 3.392-2.093 6.122-2.093 2.731 0 4.767.957 6.166 2.093.584.235 1.077.546 1.474.952.85.869 1.132 2.037 1.132 3.368 0 .368-.014.733-.052 1.086.23.462.477 1.088.644 1.517 1.258.364 2.233 1.721 2.605 2.656a.832.832 0 0 1 .053.22v2.869a.641.641 0 0 1-.078.256ZM12.172 11h-.344a4.323 4.323 0 0 1-.355.508C10.703 12.455 9.555 13 7.965 13c-1.725 0-2.989-.359-3.782-1.259a2.005 2.005 0 0 1-.085-.104L4 11.741v6.585c1.435.779 4.514 2.179 8 2.179 3.486 0 6.565-1.4 8-2.179v-6.585l-.098-.104s-.033.045-.085.104c-.793.9-2.057 1.259-3.782 1.259-1.59 0-2.738-.545-3.508-1.492a4.323 4.323 0 0 1-.355-.508h-.016.016Zm.641-2.935c.136 1.057.403 1.913.878 2.497.442.544 1.134.938 2.344.938 1.573 0 2.292-.337 2.657-.751.384-.435.558-1.15.558-2.361 0-1.14-.243-1.847-.705-2.319-.477-.488-1.319-.862-2.824-1.025-1.487-.161-2.192.138-2.533.529-.269.307-.437.808-.438 1.578v.021c0 .265.021.562.063.893Zm-1.626 0c.042-.331.063-.628.063-.894v-.02c-.001-.77-.169-1.271-.438-1.578-.341-.391-1.046-.69-2.533-.529-1.505.163-2.347.537-2.824 1.025-.462.472-.705 1.179-.705 2.319 0 1.211.175 1.926.558 2.361.365.414 1.084.751 2.657.751 1.21 0 1.902-.394 2.344-.938.475-.584.742-1.44.878-2.497Z"></path><path d="M14.5 14.25a1 1 0 0 1 1 1v2a1 1 0 0 1-2 0v-2a1 1 0 0 1 1-1Zm-5 0a1 1 0 0 1 1 1v2a1 1 0 0 1-2 0v-2a1 1 0 0 1 1-1Z"></path>
</svg>
      <div>
        <div class="color-fg-default h4">GitHub Copilot</div>
        Enterprise-grade AI features
      </div>

</a></li>

                    <li>
  <a class="HeaderMenu-dropdown-link d-block no-underline position-relative py-2 Link--secondary d-flex flex-items-center Link--has-description" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;premium_support&quot;,&quot;context&quot;:&quot;enterprise&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;premium_support_link_enterprise_navbar&quot;}" href="/premium-support">
      <svg aria-hidden="true" height="24" viewBox="0 0 24 24" version="1.1" width="24" data-view-component="true" class="octicon octicon-comment-discussion color-fg-subtle mr-3">
    <path d="M1.75 1h12.5c.966 0 1.75.784 1.75 1.75v9.5A1.75 1.75 0 0 1 14.25 14H8.061l-2.574 2.573A1.458 1.458 0 0 1 3 15.543V14H1.75A1.75 1.75 0 0 1 0 12.25v-9.5C0 1.784.784 1 1.75 1ZM1.5 2.75v9.5c0 .138.112.25.25.25h2a.75.75 0 0 1 .75.75v2.19l2.72-2.72a.749.749 0 0 1 .53-.22h6.5a.25.25 0 0 0 .25-.25v-9.5a.25.25 0 0 0-.25-.25H1.75a.25.25 0 0 0-.25.25Z"></path><path d="M22.5 8.75a.25.25 0 0 0-.25-.25h-3.5a.75.75 0 0 1 0-1.5h3.5c.966 0 1.75.784 1.75 1.75v9.5A1.75 1.75 0 0 1 22.25 20H21v1.543a1.457 1.457 0 0 1-2.487 1.03L15.939 20H10.75A1.75 1.75 0 0 1 9 18.25v-1.465a.75.75 0 0 1 1.5 0v1.465c0 .138.112.25.25.25h5.5a.75.75 0 0 1 .53.22l2.72 2.72v-2.19a.75.75 0 0 1 .75-.75h2a.25.25 0 0 0 .25-.25v-9.5Z"></path>
</svg>
      <div>
        <div class="color-fg-default h4">Premium Support</div>
        Enterprise-grade 24/7 support
      </div>

</a></li>

                </ul>
              </div>
          </div>
      </div>
</li>

                <li class="HeaderMenu-item position-relative flex-wrap flex-justify-between flex-items-center d-block d-lg-flex flex-lg-nowrap flex-lg-items-center js-details-container js-header-menu-item">
    <a class="HeaderMenu-link no-underline px-0 px-lg-2 py-3 py-lg-2 d-block d-lg-inline-block" data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;pricing&quot;,&quot;context&quot;:&quot;global&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;pricing_link_global_navbar&quot;}" href="https://github.com/pricing">Pricing</a>
</li>

            </ul>
          </nav>

        <div class="d-flex flex-column flex-lg-row width-full flex-justify-end flex-lg-items-center text-center mt-3 mt-lg-0 text-lg-left ml-lg-3">

<qbsearch-input class="search-input" data-scope="repo:johncsantiago/WhartonLab" data-custom-scopes-path="/search/custom_scopes" data-delete-custom-scopes-csrf="9svgEqlT_VxbXDi4phRAbDTfc1d0dzxofR1CiTgdkN_mPydL08cLslyRH9eeilBLKshSImMozRVZC2plBer34Q" data-max-custom-scopes="10" data-header-redesign-enabled="false" data-initial-value="" data-blackbird-suggestions-path="/search/suggestions" data-jump-to-suggestions-path="/_graphql/GetSuggestedNavigationDestinations" data-current-repository="johncsantiago/WhartonLab" data-current-org="" data-current-owner="johncsantiago" data-logged-in="false" data-copilot-chat-enabled="false" data-nl-search-enabled="false" data-retain-scroll-position="true">
  <div
    class="search-input-container search-with-dialog position-relative d-flex flex-row flex-items-center mr-4 rounded"
    data-action="click:qbsearch-input#searchInputContainerClicked"
  >
      <button
        type="button"
        class="header-search-button placeholder  input-button form-control d-flex flex-1 flex-self-stretch flex-items-center no-wrap width-full py-0 pl-2 pr-0 text-left border-0 box-shadow-none"
        data-target="qbsearch-input.inputButton"
        aria-label="Search or jump to…"
        aria-haspopup="dialog"
        placeholder="Search or jump to..."
        data-hotkey=s,/
        autocapitalize="off"
        data-analytics-event="{&quot;location&quot;:&quot;navbar&quot;,&quot;action&quot;:&quot;searchbar&quot;,&quot;context&quot;:&quot;global&quot;,&quot;tag&quot;:&quot;input&quot;,&quot;label&quot;:&quot;searchbar_input_global_navbar&quot;}"
        data-action="click:qbsearch-input#handleExpand"
      >
        <div class="mr-2 color-fg-muted">
          <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-search">
    <path d="M10.68 11.74a6 6 0 0 1-7.922-8.982 6 6 0 0 1 8.982 7.922l3.04 3.04a.749.749 0 0 1-.326 1.275.749.749 0 0 1-.734-.215ZM11.5 7a4.499 4.499 0 1 0-8.997 0A4.499 4.499 0 0 0 11.5 7Z"></path>
</svg>
        </div>
        <span class="flex-1" data-target="qbsearch-input.inputButtonText">Search or jump to...</span>
          <div class="d-flex" data-target="qbsearch-input.hotkeyIndicator">
            <svg xmlns="http://www.w3.org/2000/svg" width="22" height="20" aria-hidden="true" class="mr-1"><path fill="none" stroke="#979A9C" opacity=".4" d="M3.5.5h12c1.7 0 3 1.3 3 3v13c0 1.7-1.3 3-3 3h-12c-1.7 0-3-1.3-3-3v-13c0-1.7 1.3-3 3-3z"></path><path fill="#979A9C" d="M11.8 6L8 15.1h-.9L10.8 6h1z"></path></svg>

          </div>
      </button>

    <input type="hidden" name="type" class="js-site-search-type-field">

<div class="Overlay--hidden " data-modal-dialog-overlay>
  <modal-dialog data-action="close:qbsearch-input#handleClose cancel:qbsearch-input#handleClose" data-target="qbsearch-input.searchSuggestionsDialog" role="dialog" id="search-suggestions-dialog" aria-modal="true" aria-labelledby="search-suggestions-dialog-header" data-view-component="true" class="Overlay Overlay--width-large Overlay--height-auto">
      <h1 id="search-suggestions-dialog-header" class="sr-only">Search code, repositories, users, issues, pull requests...</h1>
    <div class="Overlay-body Overlay-body--paddingNone">

          <div data-view-component="true">        <div class="search-suggestions position-fixed width-full color-shadow-large border color-fg-default color-bg-default overflow-hidden d-flex flex-column query-builder-container"
          style="border-radius: 12px;"
          data-target="qbsearch-input.queryBuilderContainer"
          hidden
        >
          <!-- '"` --><!-- </textarea></xmp> --></option></form><form id="query-builder-test-form" action="" accept-charset="UTF-8" method="get">
  <query-builder data-target="qbsearch-input.queryBuilder" id="query-builder-query-builder-test" data-filter-key=":" data-view-component="true" class="QueryBuilder search-query-builder">
    <div class="FormControl FormControl--fullWidth">
      <label id="query-builder-test-label" for="query-builder-test" class="FormControl-label sr-only">
        Search
      </label>
      <div
        class="QueryBuilder-StyledInput width-fit "
        data-target="query-builder.styledInput"
      >
          <span id="query-builder-test-leadingvisual-wrap" class="FormControl-input-leadingVisualWrap QueryBuilder-leadingVisualWrap">
            <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-search FormControl-input-leadingVisual">
    <path d="M10.68 11.74a6 6 0 0 1-7.922-8.982 6 6 0 0 1 8.982 7.922l3.04 3.04a.749.749 0 0 1-.326 1.275.749.749 0 0 1-.734-.215ZM11.5 7a4.499 4.499 0 1 0-8.997 0A4.499 4.499 0 0 0 11.5 7Z"></path>
</svg>
          </span>
        <div data-target="query-builder.styledInputContainer" class="QueryBuilder-StyledInputContainer">
          <div
            aria-hidden="true"
            class="QueryBuilder-StyledInputContent"
            data-target="query-builder.styledInputContent"
          ></div>
          <div class="QueryBuilder-InputWrapper">
            <div aria-hidden="true" class="QueryBuilder-Sizer" data-target="query-builder.sizer"></div>
            <input id="query-builder-test" name="query-builder-test" value="" autocomplete="off" type="text" role="combobox" spellcheck="false" aria-expanded="false" aria-describedby="validation-1ce1de90-321f-40a1-becf-61982f7b63ae" data-target="query-builder.input" data-action="
          input:query-builder#inputChange
          blur:query-builder#inputBlur
          keydown:query-builder#inputKeydown
          focus:query-builder#inputFocus
        " data-view-component="true" class="FormControl-input QueryBuilder-Input FormControl-medium" />
          </div>
        </div>
          <span class="sr-only" id="query-builder-test-clear">Clear</span>
          <button role="button" id="query-builder-test-clear-button" aria-labelledby="query-builder-test-clear query-builder-test-label" data-target="query-builder.clearButton" data-action="
                click:query-builder#clear
                focus:query-builder#clearButtonFocus
                blur:query-builder#clearButtonBlur
              " variant="small" hidden="hidden" type="button" data-view-component="true" class="Button Button--iconOnly Button--invisible Button--medium mr-1 px-2 py-0 d-flex flex-items-center rounded-1 color-fg-muted">  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-x-circle-fill Button-visual">
    <path d="M2.343 13.657A8 8 0 1 1 13.658 2.343 8 8 0 0 1 2.343 13.657ZM6.03 4.97a.751.751 0 0 0-1.042.018.751.751 0 0 0-.018 1.042L6.94 8 4.97 9.97a.749.749 0 0 0 .326 1.275.749.749 0 0 0 .734-.215L8 9.06l1.97 1.97a.749.749 0 0 0 1.275-.326.749.749 0 0 0-.215-.734L9.06 8l1.97-1.97a.749.749 0 0 0-.326-1.275.749.749 0 0 0-.734.215L8 6.94Z"></path>
</svg>
</button>

      </div>
      <template id="search-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-search">
    <path d="M10.68 11.74a6 6 0 0 1-7.922-8.982 6 6 0 0 1 8.982 7.922l3.04 3.04a.749.749 0 0 1-.326 1.275.749.749 0 0 1-.734-.215ZM11.5 7a4.499 4.499 0 1 0-8.997 0A4.499 4.499 0 0 0 11.5 7Z"></path>
</svg>
</template>

<template id="code-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-code">
    <path d="m11.28 3.22 4.25 4.25a.75.75 0 0 1 0 1.06l-4.25 4.25a.749.749 0 0 1-1.275-.326.749.749 0 0 1 .215-.734L13.94 8l-3.72-3.72a.749.749 0 0 1 .326-1.275.749.749 0 0 1 .734.215Zm-6.56 0a.751.751 0 0 1 1.042.018.751.751 0 0 1 .018 1.042L2.06 8l3.72 3.72a.749.749 0 0 1-.326 1.275.749.749 0 0 1-.734-.215L.47 8.53a.75.75 0 0 1 0-1.06Z"></path>
</svg>
</template>

<template id="file-code-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-file-code">
    <path d="M4 1.75C4 .784 4.784 0 5.75 0h5.586c.464 0 .909.184 1.237.513l2.914 2.914c.329.328.513.773.513 1.237v8.586A1.75 1.75 0 0 1 14.25 15h-9a.75.75 0 0 1 0-1.5h9a.25.25 0 0 0 .25-.25V6h-2.75A1.75 1.75 0 0 1 10 4.25V1.5H5.75a.25.25 0 0 0-.25.25v2.5a.75.75 0 0 1-1.5 0Zm1.72 4.97a.75.75 0 0 1 1.06 0l2 2a.75.75 0 0 1 0 1.06l-2 2a.749.749 0 0 1-1.275-.326.749.749 0 0 1 .215-.734l1.47-1.47-1.47-1.47a.75.75 0 0 1 0-1.06ZM3.28 7.78 1.81 9.25l1.47 1.47a.751.751 0 0 1-.018 1.042.751.751 0 0 1-1.042.018l-2-2a.75.75 0 0 1 0-1.06l2-2a.751.751 0 0 1 1.042.018.751.751 0 0 1 .018 1.042Zm8.22-6.218V4.25c0 .138.112.25.25.25h2.688l-.011-.013-2.914-2.914-.013-.011Z"></path>
</svg>
</template>

<template id="history-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-history">
    <path d="m.427 1.927 1.215 1.215a8.002 8.002 0 1 1-1.6 5.685.75.75 0 1 1 1.493-.154 6.5 6.5 0 1 0 1.18-4.458l1.358 1.358A.25.25 0 0 1 3.896 6H.25A.25.25 0 0 1 0 5.75V2.104a.25.25 0 0 1 .427-.177ZM7.75 4a.75.75 0 0 1 .75.75v2.992l2.028.812a.75.75 0 0 1-.557 1.392l-2.5-1A.751.751 0 0 1 7 8.25v-3.5A.75.75 0 0 1 7.75 4Z"></path>
</svg>
</template>

<template id="repo-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-repo">
    <path d="M2 2.5A2.5 2.5 0 0 1 4.5 0h8.75a.75.75 0 0 1 .75.75v12.5a.75.75 0 0 1-.75.75h-2.5a.75.75 0 0 1 0-1.5h1.75v-2h-8a1 1 0 0 0-.714 1.7.75.75 0 1 1-1.072 1.05A2.495 2.495 0 0 1 2 11.5Zm10.5-1h-8a1 1 0 0 0-1 1v6.708A2.486 2.486 0 0 1 4.5 9h8ZM5 12.25a.25.25 0 0 1 .25-.25h3.5a.25.25 0 0 1 .25.25v3.25a.25.25 0 0 1-.4.2l-1.45-1.087a.249.249 0 0 0-.3 0L5.4 15.7a.25.25 0 0 1-.4-.2Z"></path>
</svg>
</template>

<template id="bookmark-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-bookmark">
    <path d="M3 2.75C3 1.784 3.784 1 4.75 1h6.5c.966 0 1.75.784 1.75 1.75v11.5a.75.75 0 0 1-1.227.579L8 11.722l-3.773 3.107A.751.751 0 0 1 3 14.25Zm1.75-.25a.25.25 0 0 0-.25.25v9.91l3.023-2.489a.75.75 0 0 1 .954 0l3.023 2.49V2.75a.25.25 0 0 0-.25-.25Z"></path>
</svg>
</template>

<template id="plus-circle-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-plus-circle">
    <path d="M8 0a8 8 0 1 1 0 16A8 8 0 0 1 8 0ZM1.5 8a6.5 6.5 0 1 0 13 0 6.5 6.5 0 0 0-13 0Zm7.25-3.25v2.5h2.5a.75.75 0 0 1 0 1.5h-2.5v2.5a.75.75 0 0 1-1.5 0v-2.5h-2.5a.75.75 0 0 1 0-1.5h2.5v-2.5a.75.75 0 0 1 1.5 0Z"></path>
</svg>
</template>

<template id="circle-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-dot-fill">
    <path d="M8 4a4 4 0 1 1 0 8 4 4 0 0 1 0-8Z"></path>
</svg>
</template>

<template id="trash-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-trash">
    <path d="M11 1.75V3h2.25a.75.75 0 0 1 0 1.5H2.75a.75.75 0 0 1 0-1.5H5V1.75C5 .784 5.784 0 6.75 0h2.5C10.216 0 11 .784 11 1.75ZM4.496 6.675l.66 6.6a.25.25 0 0 0 .249.225h5.19a.25.25 0 0 0 .249-.225l.66-6.6a.75.75 0 0 1 1.492.149l-.66 6.6A1.748 1.748 0 0 1 10.595 15h-5.19a1.75 1.75 0 0 1-1.741-1.575l-.66-6.6a.75.75 0 1 1 1.492-.15ZM6.5 1.75V3h3V1.75a.25.25 0 0 0-.25-.25h-2.5a.25.25 0 0 0-.25.25Z"></path>
</svg>
</template>

<template id="team-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-people">
    <path d="M2 5.5a3.5 3.5 0 1 1 5.898 2.549 5.508 5.508 0 0 1 3.034 4.084.75.75 0 1 1-1.482.235 4 4 0 0 0-7.9 0 .75.75 0 0 1-1.482-.236A5.507 5.507 0 0 1 3.102 8.05 3.493 3.493 0 0 1 2 5.5ZM11 4a3.001 3.001 0 0 1 2.22 5.018 5.01 5.01 0 0 1 2.56 3.012.749.749 0 0 1-.885.954.752.752 0 0 1-.549-.514 3.507 3.507 0 0 0-2.522-2.372.75.75 0 0 1-.574-.73v-.352a.75.75 0 0 1 .416-.672A1.5 1.5 0 0 0 11 5.5.75.75 0 0 1 11 4Zm-5.5-.5a2 2 0 1 0-.001 3.999A2 2 0 0 0 5.5 3.5Z"></path>
</svg>
</template>

<template id="project-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-project">
    <path d="M1.75 0h12.5C15.216 0 16 .784 16 1.75v12.5A1.75 1.75 0 0 1 14.25 16H1.75A1.75 1.75 0 0 1 0 14.25V1.75C0 .784.784 0 1.75 0ZM1.5 1.75v12.5c0 .138.112.25.25.25h12.5a.25.25 0 0 0 .25-.25V1.75a.25.25 0 0 0-.25-.25H1.75a.25.25 0 0 0-.25.25ZM11.75 3a.75.75 0 0 1 .75.75v7.5a.75.75 0 0 1-1.5 0v-7.5a.75.75 0 0 1 .75-.75Zm-8.25.75a.75.75 0 0 1 1.5 0v5.5a.75.75 0 0 1-1.5 0ZM8 3a.75.75 0 0 1 .75.75v3.5a.75.75 0 0 1-1.5 0v-3.5A.75.75 0 0 1 8 3Z"></path>
</svg>
</template>

<template id="pencil-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-pencil">
    <path d="M11.013 1.427a1.75 1.75 0 0 1 2.474 0l1.086 1.086a1.75 1.75 0 0 1 0 2.474l-8.61 8.61c-.21.21-.47.364-.756.445l-3.251.93a.75.75 0 0 1-.927-.928l.929-3.25c.081-.286.235-.547.445-.758l8.61-8.61Zm.176 4.823L9.75 4.81l-6.286 6.287a.253.253 0 0 0-.064.108l-.558 1.953 1.953-.558a.253.253 0 0 0 .108-.064Zm1.238-3.763a.25.25 0 0 0-.354 0L10.811 3.75l1.439 1.44 1.263-1.263a.25.25 0 0 0 0-.354Z"></path>
</svg>
</template>

<template id="copilot-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-copilot">
    <path d="M7.998 15.035c-4.562 0-7.873-2.914-7.998-3.749V9.338c.085-.628.677-1.686 1.588-2.065.013-.07.024-.143.036-.218.029-.183.06-.384.126-.612-.201-.508-.254-1.084-.254-1.656 0-.87.128-1.769.693-2.484.579-.733 1.494-1.124 2.724-1.261 1.206-.134 2.262.034 2.944.765.05.053.096.108.139.165.044-.057.094-.112.143-.165.682-.731 1.738-.899 2.944-.765 1.23.137 2.145.528 2.724 1.261.566.715.693 1.614.693 2.484 0 .572-.053 1.148-.254 1.656.066.228.098.429.126.612.012.076.024.148.037.218.924.385 1.522 1.471 1.591 2.095v1.872c0 .766-3.351 3.795-8.002 3.795Zm0-1.485c2.28 0 4.584-1.11 5.002-1.433V7.862l-.023-.116c-.49.21-1.075.291-1.727.291-1.146 0-2.059-.327-2.71-.991A3.222 3.222 0 0 1 8 6.303a3.24 3.24 0 0 1-.544.743c-.65.664-1.563.991-2.71.991-.652 0-1.236-.081-1.727-.291l-.023.116v4.255c.419.323 2.722 1.433 5.002 1.433ZM6.762 2.83c-.193-.206-.637-.413-1.682-.297-1.019.113-1.479.404-1.713.7-.247.312-.369.789-.369 1.554 0 .793.129 1.171.308 1.371.162.181.519.379 1.442.379.853 0 1.339-.235 1.638-.54.315-.322.527-.827.617-1.553.117-.935-.037-1.395-.241-1.614Zm4.155-.297c-1.044-.116-1.488.091-1.681.297-.204.219-.359.679-.242 1.614.091.726.303 1.231.618 1.553.299.305.784.54 1.638.54.922 0 1.28-.198 1.442-.379.179-.2.308-.578.308-1.371 0-.765-.123-1.242-.37-1.554-.233-.296-.693-.587-1.713-.7Z"></path><path d="M6.25 9.037a.75.75 0 0 1 .75.75v1.501a.75.75 0 0 1-1.5 0V9.787a.75.75 0 0 1 .75-.75Zm4.25.75v1.501a.75.75 0 0 1-1.5 0V9.787a.75.75 0 0 1 1.5 0Z"></path>
</svg>
</template>

<template id="copilot-error-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-copilot-error">
    <path d="M16 11.24c0 .112-.072.274-.21.467L13 9.688V7.862l-.023-.116c-.49.21-1.075.291-1.727.291-.198 0-.388-.009-.571-.029L6.833 5.226a4.01 4.01 0 0 0 .17-.782c.117-.935-.037-1.395-.241-1.614-.193-.206-.637-.413-1.682-.297-.683.076-1.115.231-1.395.415l-1.257-.91c.579-.564 1.413-.877 2.485-.996 1.206-.134 2.262.034 2.944.765.05.053.096.108.139.165.044-.057.094-.112.143-.165.682-.731 1.738-.899 2.944-.765 1.23.137 2.145.528 2.724 1.261.566.715.693 1.614.693 2.484 0 .572-.053 1.148-.254 1.656.066.228.098.429.126.612.012.076.024.148.037.218.924.385 1.522 1.471 1.591 2.095Zm-5.083-8.707c-1.044-.116-1.488.091-1.681.297-.204.219-.359.679-.242 1.614.091.726.303 1.231.618 1.553.299.305.784.54 1.638.54.922 0 1.28-.198 1.442-.379.179-.2.308-.578.308-1.371 0-.765-.123-1.242-.37-1.554-.233-.296-.693-.587-1.713-.7Zm2.511 11.074c-1.393.776-3.272 1.428-5.43 1.428-4.562 0-7.873-2.914-7.998-3.749V9.338c.085-.628.677-1.686 1.588-2.065.013-.07.024-.143.036-.218.029-.183.06-.384.126-.612-.18-.455-.241-.963-.252-1.475L.31 4.107A.747.747 0 0 1 0 3.509V3.49a.748.748 0 0 1 .625-.73c.156-.026.306.047.435.139l14.667 10.578a.592.592 0 0 1 .227.264.752.752 0 0 1 .046.249v.022a.75.75 0 0 1-1.19.596Zm-1.367-.991L5.635 7.964a5.128 5.128 0 0 1-.889.073c-.652 0-1.236-.081-1.727-.291l-.023.116v4.255c.419.323 2.722 1.433 5.002 1.433 1.539 0 3.089-.505 4.063-.934Z"></path>
</svg>
</template>

<template id="workflow-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-workflow">
    <path d="M0 1.75C0 .784.784 0 1.75 0h3.5C6.216 0 7 .784 7 1.75v3.5A1.75 1.75 0 0 1 5.25 7H4v4a1 1 0 0 0 1 1h4v-1.25C9 9.784 9.784 9 10.75 9h3.5c.966 0 1.75.784 1.75 1.75v3.5A1.75 1.75 0 0 1 14.25 16h-3.5A1.75 1.75 0 0 1 9 14.25v-.75H5A2.5 2.5 0 0 1 2.5 11V7h-.75A1.75 1.75 0 0 1 0 5.25Zm1.75-.25a.25.25 0 0 0-.25.25v3.5c0 .138.112.25.25.25h3.5a.25.25 0 0 0 .25-.25v-3.5a.25.25 0 0 0-.25-.25Zm9 9a.25.25 0 0 0-.25.25v3.5c0 .138.112.25.25.25h3.5a.25.25 0 0 0 .25-.25v-3.5a.25.25 0 0 0-.25-.25Z"></path>
</svg>
</template>

<template id="book-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-book">
    <path d="M0 1.75A.75.75 0 0 1 .75 1h4.253c1.227 0 2.317.59 3 1.501A3.743 3.743 0 0 1 11.006 1h4.245a.75.75 0 0 1 .75.75v10.5a.75.75 0 0 1-.75.75h-4.507a2.25 2.25 0 0 0-1.591.659l-.622.621a.75.75 0 0 1-1.06 0l-.622-.621A2.25 2.25 0 0 0 5.258 13H.75a.75.75 0 0 1-.75-.75Zm7.251 10.324.004-5.073-.002-2.253A2.25 2.25 0 0 0 5.003 2.5H1.5v9h3.757a3.75 3.75 0 0 1 1.994.574ZM8.755 4.75l-.004 7.322a3.752 3.752 0 0 1 1.992-.572H14.5v-9h-3.495a2.25 2.25 0 0 0-2.25 2.25Z"></path>
</svg>
</template>

<template id="code-review-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-code-review">
    <path d="M1.75 1h12.5c.966 0 1.75.784 1.75 1.75v8.5A1.75 1.75 0 0 1 14.25 13H8.061l-2.574 2.573A1.458 1.458 0 0 1 3 14.543V13H1.75A1.75 1.75 0 0 1 0 11.25v-8.5C0 1.784.784 1 1.75 1ZM1.5 2.75v8.5c0 .138.112.25.25.25h2a.75.75 0 0 1 .75.75v2.19l2.72-2.72a.749.749 0 0 1 .53-.22h6.5a.25.25 0 0 0 .25-.25v-8.5a.25.25 0 0 0-.25-.25H1.75a.25.25 0 0 0-.25.25Zm5.28 1.72a.75.75 0 0 1 0 1.06L5.31 7l1.47 1.47a.751.751 0 0 1-.018 1.042.751.751 0 0 1-1.042.018l-2-2a.75.75 0 0 1 0-1.06l2-2a.75.75 0 0 1 1.06 0Zm2.44 0a.75.75 0 0 1 1.06 0l2 2a.75.75 0 0 1 0 1.06l-2 2a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042L10.69 7 9.22 5.53a.75.75 0 0 1 0-1.06Z"></path>
</svg>
</template>

<template id="codespaces-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-codespaces">
    <path d="M0 11.25c0-.966.784-1.75 1.75-1.75h12.5c.966 0 1.75.784 1.75 1.75v3A1.75 1.75 0 0 1 14.25 16H1.75A1.75 1.75 0 0 1 0 14.25Zm2-9.5C2 .784 2.784 0 3.75 0h8.5C13.216 0 14 .784 14 1.75v5a1.75 1.75 0 0 1-1.75 1.75h-8.5A1.75 1.75 0 0 1 2 6.75Zm1.75-.25a.25.25 0 0 0-.25.25v5c0 .138.112.25.25.25h8.5a.25.25 0 0 0 .25-.25v-5a.25.25 0 0 0-.25-.25Zm-2 9.5a.25.25 0 0 0-.25.25v3c0 .138.112.25.25.25h12.5a.25.25 0 0 0 .25-.25v-3a.25.25 0 0 0-.25-.25Z"></path><path d="M7 12.75a.75.75 0 0 1 .75-.75h4.5a.75.75 0 0 1 0 1.5h-4.5a.75.75 0 0 1-.75-.75Zm-4 0a.75.75 0 0 1 .75-.75h.5a.75.75 0 0 1 0 1.5h-.5a.75.75 0 0 1-.75-.75Z"></path>
</svg>
</template>

<template id="comment-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-comment">
    <path d="M1 2.75C1 1.784 1.784 1 2.75 1h10.5c.966 0 1.75.784 1.75 1.75v7.5A1.75 1.75 0 0 1 13.25 12H9.06l-2.573 2.573A1.458 1.458 0 0 1 4 13.543V12H2.75A1.75 1.75 0 0 1 1 10.25Zm1.75-.25a.25.25 0 0 0-.25.25v7.5c0 .138.112.25.25.25h2a.75.75 0 0 1 .75.75v2.19l2.72-2.72a.749.749 0 0 1 .53-.22h4.5a.25.25 0 0 0 .25-.25v-7.5a.25.25 0 0 0-.25-.25Z"></path>
</svg>
</template>

<template id="comment-discussion-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-comment-discussion">
    <path d="M1.75 1h8.5c.966 0 1.75.784 1.75 1.75v5.5A1.75 1.75 0 0 1 10.25 10H7.061l-2.574 2.573A1.458 1.458 0 0 1 2 11.543V10h-.25A1.75 1.75 0 0 1 0 8.25v-5.5C0 1.784.784 1 1.75 1ZM1.5 2.75v5.5c0 .138.112.25.25.25h1a.75.75 0 0 1 .75.75v2.19l2.72-2.72a.749.749 0 0 1 .53-.22h3.5a.25.25 0 0 0 .25-.25v-5.5a.25.25 0 0 0-.25-.25h-8.5a.25.25 0 0 0-.25.25Zm13 2a.25.25 0 0 0-.25-.25h-.5a.75.75 0 0 1 0-1.5h.5c.966 0 1.75.784 1.75 1.75v5.5A1.75 1.75 0 0 1 14.25 12H14v1.543a1.458 1.458 0 0 1-2.487 1.03L9.22 12.28a.749.749 0 0 1 .326-1.275.749.749 0 0 1 .734.215l2.22 2.22v-2.19a.75.75 0 0 1 .75-.75h1a.25.25 0 0 0 .25-.25Z"></path>
</svg>
</template>

<template id="organization-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-organization">
    <path d="M1.75 16A1.75 1.75 0 0 1 0 14.25V1.75C0 .784.784 0 1.75 0h8.5C11.216 0 12 .784 12 1.75v12.5c0 .085-.006.168-.018.25h2.268a.25.25 0 0 0 .25-.25V8.285a.25.25 0 0 0-.111-.208l-1.055-.703a.749.749 0 1 1 .832-1.248l1.055.703c.487.325.779.871.779 1.456v5.965A1.75 1.75 0 0 1 14.25 16h-3.5a.766.766 0 0 1-.197-.026c-.099.017-.2.026-.303.026h-3a.75.75 0 0 1-.75-.75V14h-1v1.25a.75.75 0 0 1-.75.75Zm-.25-1.75c0 .138.112.25.25.25H4v-1.25a.75.75 0 0 1 .75-.75h2.5a.75.75 0 0 1 .75.75v1.25h2.25a.25.25 0 0 0 .25-.25V1.75a.25.25 0 0 0-.25-.25h-8.5a.25.25 0 0 0-.25.25ZM3.75 6h.5a.75.75 0 0 1 0 1.5h-.5a.75.75 0 0 1 0-1.5ZM3 3.75A.75.75 0 0 1 3.75 3h.5a.75.75 0 0 1 0 1.5h-.5A.75.75 0 0 1 3 3.75Zm4 3A.75.75 0 0 1 7.75 6h.5a.75.75 0 0 1 0 1.5h-.5A.75.75 0 0 1 7 6.75ZM7.75 3h.5a.75.75 0 0 1 0 1.5h-.5a.75.75 0 0 1 0-1.5ZM3 9.75A.75.75 0 0 1 3.75 9h.5a.75.75 0 0 1 0 1.5h-.5A.75.75 0 0 1 3 9.75ZM7.75 9h.5a.75.75 0 0 1 0 1.5h-.5a.75.75 0 0 1 0-1.5Z"></path>
</svg>
</template>

<template id="rocket-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-rocket">
    <path d="M14.064 0h.186C15.216 0 16 .784 16 1.75v.186a8.752 8.752 0 0 1-2.564 6.186l-.458.459c-.314.314-.641.616-.979.904v3.207c0 .608-.315 1.172-.833 1.49l-2.774 1.707a.749.749 0 0 1-1.11-.418l-.954-3.102a1.214 1.214 0 0 1-.145-.125L3.754 9.816a1.218 1.218 0 0 1-.124-.145L.528 8.717a.749.749 0 0 1-.418-1.11l1.71-2.774A1.748 1.748 0 0 1 3.31 4h3.204c.288-.338.59-.665.904-.979l.459-.458A8.749 8.749 0 0 1 14.064 0ZM8.938 3.623h-.002l-.458.458c-.76.76-1.437 1.598-2.02 2.5l-1.5 2.317 2.143 2.143 2.317-1.5c.902-.583 1.74-1.26 2.499-2.02l.459-.458a7.25 7.25 0 0 0 2.123-5.127V1.75a.25.25 0 0 0-.25-.25h-.186a7.249 7.249 0 0 0-5.125 2.123ZM3.56 14.56c-.732.732-2.334 1.045-3.005 1.148a.234.234 0 0 1-.201-.064.234.234 0 0 1-.064-.201c.103-.671.416-2.273 1.15-3.003a1.502 1.502 0 1 1 2.12 2.12Zm6.94-3.935c-.088.06-.177.118-.266.175l-2.35 1.521.548 1.783 1.949-1.2a.25.25 0 0 0 .119-.213ZM3.678 8.116 5.2 5.766c.058-.09.117-.178.176-.266H3.309a.25.25 0 0 0-.213.119l-1.2 1.95ZM12 5a1 1 0 1 1-2 0 1 1 0 0 1 2 0Z"></path>
</svg>
</template>

<template id="shield-check-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-shield-check">
    <path d="m8.533.133 5.25 1.68A1.75 1.75 0 0 1 15 3.48V7c0 1.566-.32 3.182-1.303 4.682-.983 1.498-2.585 2.813-5.032 3.855a1.697 1.697 0 0 1-1.33 0c-2.447-1.042-4.049-2.357-5.032-3.855C1.32 10.182 1 8.566 1 7V3.48a1.75 1.75 0 0 1 1.217-1.667l5.25-1.68a1.748 1.748 0 0 1 1.066 0Zm-.61 1.429.001.001-5.25 1.68a.251.251 0 0 0-.174.237V7c0 1.36.275 2.666 1.057 3.859.784 1.194 2.121 2.342 4.366 3.298a.196.196 0 0 0 .154 0c2.245-.957 3.582-2.103 4.366-3.297C13.225 9.666 13.5 8.358 13.5 7V3.48a.25.25 0 0 0-.174-.238l-5.25-1.68a.25.25 0 0 0-.153 0ZM11.28 6.28l-3.5 3.5a.75.75 0 0 1-1.06 0l-1.5-1.5a.749.749 0 0 1 .326-1.275.749.749 0 0 1 .734.215l.97.97 2.97-2.97a.751.751 0 0 1 1.042.018.751.751 0 0 1 .018 1.042Z"></path>
</svg>
</template>

<template id="heart-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-heart">
    <path d="m8 14.25.345.666a.75.75 0 0 1-.69 0l-.008-.004-.018-.01a7.152 7.152 0 0 1-.31-.17 22.055 22.055 0 0 1-3.434-2.414C2.045 10.731 0 8.35 0 5.5 0 2.836 2.086 1 4.25 1 5.797 1 7.153 1.802 8 3.02 8.847 1.802 10.203 1 11.75 1 13.914 1 16 2.836 16 5.5c0 2.85-2.045 5.231-3.885 6.818a22.066 22.066 0 0 1-3.744 2.584l-.018.01-.006.003h-.002ZM4.25 2.5c-1.336 0-2.75 1.164-2.75 3 0 2.15 1.58 4.144 3.365 5.682A20.58 20.58 0 0 0 8 13.393a20.58 20.58 0 0 0 3.135-2.211C12.92 9.644 14.5 7.65 14.5 5.5c0-1.836-1.414-3-2.75-3-1.373 0-2.609.986-3.029 2.456a.749.749 0 0 1-1.442 0C6.859 3.486 5.623 2.5 4.25 2.5Z"></path>
</svg>
</template>

<template id="server-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-server">
    <path d="M1.75 1h12.5c.966 0 1.75.784 1.75 1.75v4c0 .372-.116.717-.314 1 .198.283.314.628.314 1v4a1.75 1.75 0 0 1-1.75 1.75H1.75A1.75 1.75 0 0 1 0 12.75v-4c0-.358.109-.707.314-1a1.739 1.739 0 0 1-.314-1v-4C0 1.784.784 1 1.75 1ZM1.5 2.75v4c0 .138.112.25.25.25h12.5a.25.25 0 0 0 .25-.25v-4a.25.25 0 0 0-.25-.25H1.75a.25.25 0 0 0-.25.25Zm.25 5.75a.25.25 0 0 0-.25.25v4c0 .138.112.25.25.25h12.5a.25.25 0 0 0 .25-.25v-4a.25.25 0 0 0-.25-.25ZM7 4.75A.75.75 0 0 1 7.75 4h4.5a.75.75 0 0 1 0 1.5h-4.5A.75.75 0 0 1 7 4.75ZM7.75 10h4.5a.75.75 0 0 1 0 1.5h-4.5a.75.75 0 0 1 0-1.5ZM3 4.75A.75.75 0 0 1 3.75 4h.5a.75.75 0 0 1 0 1.5h-.5A.75.75 0 0 1 3 4.75ZM3.75 10h.5a.75.75 0 0 1 0 1.5h-.5a.75.75 0 0 1 0-1.5Z"></path>
</svg>
</template>

<template id="globe-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-globe">
    <path d="M8 0a8 8 0 1 1 0 16A8 8 0 0 1 8 0ZM5.78 8.75a9.64 9.64 0 0 0 1.363 4.177c.255.426.542.832.857 1.215.245-.296.551-.705.857-1.215A9.64 9.64 0 0 0 10.22 8.75Zm4.44-1.5a9.64 9.64 0 0 0-1.363-4.177c-.307-.51-.612-.919-.857-1.215a9.927 9.927 0 0 0-.857 1.215A9.64 9.64 0 0 0 5.78 7.25Zm-5.944 1.5H1.543a6.507 6.507 0 0 0 4.666 5.5c-.123-.181-.24-.365-.352-.552-.715-1.192-1.437-2.874-1.581-4.948Zm-2.733-1.5h2.733c.144-2.074.866-3.756 1.58-4.948.12-.197.237-.381.353-.552a6.507 6.507 0 0 0-4.666 5.5Zm10.181 1.5c-.144 2.074-.866 3.756-1.58 4.948-.12.197-.237.381-.353.552a6.507 6.507 0 0 0 4.666-5.5Zm2.733-1.5a6.507 6.507 0 0 0-4.666-5.5c.123.181.24.365.353.552.714 1.192 1.436 2.874 1.58 4.948Z"></path>
</svg>
</template>

<template id="issue-opened-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-issue-opened">
    <path d="M8 9.5a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3Z"></path><path d="M8 0a8 8 0 1 1 0 16A8 8 0 0 1 8 0ZM1.5 8a6.5 6.5 0 1 0 13 0 6.5 6.5 0 0 0-13 0Z"></path>
</svg>
</template>

<template id="device-mobile-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-device-mobile">
    <path d="M3.75 0h8.5C13.216 0 14 .784 14 1.75v12.5A1.75 1.75 0 0 1 12.25 16h-8.5A1.75 1.75 0 0 1 2 14.25V1.75C2 .784 2.784 0 3.75 0ZM3.5 1.75v12.5c0 .138.112.25.25.25h8.5a.25.25 0 0 0 .25-.25V1.75a.25.25 0 0 0-.25-.25h-8.5a.25.25 0 0 0-.25.25ZM8 13a1 1 0 1 1 0-2 1 1 0 0 1 0 2Z"></path>
</svg>
</template>

<template id="package-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-package">
    <path d="m8.878.392 5.25 3.045c.54.314.872.89.872 1.514v6.098a1.75 1.75 0 0 1-.872 1.514l-5.25 3.045a1.75 1.75 0 0 1-1.756 0l-5.25-3.045A1.75 1.75 0 0 1 1 11.049V4.951c0-.624.332-1.201.872-1.514L7.122.392a1.75 1.75 0 0 1 1.756 0ZM7.875 1.69l-4.63 2.685L8 7.133l4.755-2.758-4.63-2.685a.248.248 0 0 0-.25 0ZM2.5 5.677v5.372c0 .09.047.171.125.216l4.625 2.683V8.432Zm6.25 8.271 4.625-2.683a.25.25 0 0 0 .125-.216V5.677L8.75 8.432Z"></path>
</svg>
</template>

<template id="credit-card-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-credit-card">
    <path d="M10.75 9a.75.75 0 0 0 0 1.5h1.5a.75.75 0 0 0 0-1.5h-1.5Z"></path><path d="M0 3.75C0 2.784.784 2 1.75 2h12.5c.966 0 1.75.784 1.75 1.75v8.5A1.75 1.75 0 0 1 14.25 14H1.75A1.75 1.75 0 0 1 0 12.25ZM14.5 6.5h-13v5.75c0 .138.112.25.25.25h12.5a.25.25 0 0 0 .25-.25Zm0-2.75a.25.25 0 0 0-.25-.25H1.75a.25.25 0 0 0-.25.25V5h13Z"></path>
</svg>
</template>

<template id="play-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-play">
    <path d="M8 0a8 8 0 1 1 0 16A8 8 0 0 1 8 0ZM1.5 8a6.5 6.5 0 1 0 13 0 6.5 6.5 0 0 0-13 0Zm4.879-2.773 4.264 2.559a.25.25 0 0 1 0 .428l-4.264 2.559A.25.25 0 0 1 6 10.559V5.442a.25.25 0 0 1 .379-.215Z"></path>
</svg>
</template>

<template id="gift-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-gift">
    <path d="M2 2.75A2.75 2.75 0 0 1 4.75 0c.983 0 1.873.42 2.57 1.232.268.318.497.668.68 1.042.183-.375.411-.725.68-1.044C9.376.42 10.266 0 11.25 0a2.75 2.75 0 0 1 2.45 4h.55c.966 0 1.75.784 1.75 1.75v2c0 .698-.409 1.301-1 1.582v4.918A1.75 1.75 0 0 1 13.25 16H2.75A1.75 1.75 0 0 1 1 14.25V9.332C.409 9.05 0 8.448 0 7.75v-2C0 4.784.784 4 1.75 4h.55c-.192-.375-.3-.8-.3-1.25ZM7.25 9.5H2.5v4.75c0 .138.112.25.25.25h4.5Zm1.5 0v5h4.5a.25.25 0 0 0 .25-.25V9.5Zm0-4V8h5.5a.25.25 0 0 0 .25-.25v-2a.25.25 0 0 0-.25-.25Zm-7 0a.25.25 0 0 0-.25.25v2c0 .138.112.25.25.25h5.5V5.5h-5.5Zm3-4a1.25 1.25 0 0 0 0 2.5h2.309c-.233-.818-.542-1.401-.878-1.793-.43-.502-.915-.707-1.431-.707ZM8.941 4h2.309a1.25 1.25 0 0 0 0-2.5c-.516 0-1 .205-1.43.707-.337.392-.646.975-.879 1.793Z"></path>
</svg>
</template>

<template id="code-square-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-code-square">
    <path d="M0 1.75C0 .784.784 0 1.75 0h12.5C15.216 0 16 .784 16 1.75v12.5A1.75 1.75 0 0 1 14.25 16H1.75A1.75 1.75 0 0 1 0 14.25Zm1.75-.25a.25.25 0 0 0-.25.25v12.5c0 .138.112.25.25.25h12.5a.25.25 0 0 0 .25-.25V1.75a.25.25 0 0 0-.25-.25Zm7.47 3.97a.75.75 0 0 1 1.06 0l2 2a.75.75 0 0 1 0 1.06l-2 2a.749.749 0 0 1-1.275-.326.749.749 0 0 1 .215-.734L10.69 8 9.22 6.53a.75.75 0 0 1 0-1.06ZM6.78 6.53 5.31 8l1.47 1.47a.749.749 0 0 1-.326 1.275.749.749 0 0 1-.734-.215l-2-2a.75.75 0 0 1 0-1.06l2-2a.751.751 0 0 1 1.042.018.751.751 0 0 1 .018 1.042Z"></path>
</svg>
</template>

<template id="device-desktop-icon">
  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-device-desktop">
    <path d="M14.25 1c.966 0 1.75.784 1.75 1.75v7.5A1.75 1.75 0 0 1 14.25 12h-3.727c.099 1.041.52 1.872 1.292 2.757A.752.752 0 0 1 11.25 16h-6.5a.75.75 0 0 1-.565-1.243c.772-.885 1.192-1.716 1.292-2.757H1.75A1.75 1.75 0 0 1 0 10.25v-7.5C0 1.784.784 1 1.75 1ZM1.75 2.5a.25.25 0 0 0-.25.25v7.5c0 .138.112.25.25.25h12.5a.25.25 0 0 0 .25-.25v-7.5a.25.25 0 0 0-.25-.25ZM9.018 12H6.982a5.72 5.72 0 0 1-.765 2.5h3.566a5.72 5.72 0 0 1-.765-2.5Z"></path>
</svg>
</template>

        <div class="position-relative">
                <ul
                  role="listbox"
                  class="ActionListWrap QueryBuilder-ListWrap"
                  aria-label="Suggestions"
                  data-action="
                    combobox-commit:query-builder#comboboxCommit
                    mousedown:query-builder#resultsMousedown
                  "
                  data-target="query-builder.resultsList"
                  data-persist-list=false
                  id="query-builder-test-results"
                ></ul>
        </div>
      <div class="FormControl-inlineValidation" id="validation-1ce1de90-321f-40a1-becf-61982f7b63ae" hidden="hidden">
        <span class="FormControl-inlineValidation--visual">
          <svg aria-hidden="true" height="12" viewBox="0 0 12 12" version="1.1" width="12" data-view-component="true" class="octicon octicon-alert-fill">
    <path d="M4.855.708c.5-.896 1.79-.896 2.29 0l4.675 8.351a1.312 1.312 0 0 1-1.146 1.954H1.33A1.313 1.313 0 0 1 .183 9.058ZM7 7V3H5v4Zm-1 3a1 1 0 1 0 0-2 1 1 0 0 0 0 2Z"></path>
</svg>
        </span>
        <span></span>
</div>    </div>
    <div data-target="query-builder.screenReaderFeedback" aria-live="polite" aria-atomic="true" class="sr-only"></div>
</query-builder></form>
          <div class="d-flex flex-row color-fg-muted px-3 text-small color-bg-default search-feedback-prompt">
            <a target="_blank" href="https://docs.github.com/search-github/github-code-search/understanding-github-code-search-syntax" data-view-component="true" class="Link color-fg-accent text-normal ml-2">Search syntax tips</a>            <div class="d-flex flex-1"></div>
          </div>
        </div>
</div>

    </div>
</modal-dialog></div>
  </div>
  <div data-action="click:qbsearch-input#retract" class="dark-backdrop position-fixed" hidden data-target="qbsearch-input.darkBackdrop"></div>
  <div class="color-fg-default">

<dialog-helper>
  <dialog data-target="qbsearch-input.feedbackDialog" data-action="close:qbsearch-input#handleDialogClose cancel:qbsearch-input#handleDialogClose" id="feedback-dialog" aria-modal="true" aria-labelledby="feedback-dialog-title" aria-describedby="feedback-dialog-description" data-view-component="true" class="Overlay Overlay-whenNarrow Overlay--size-medium Overlay--motion-scaleFade Overlay--disableScroll">
    <div data-view-component="true" class="Overlay-header">
  <div class="Overlay-headerContentWrap">
    <div class="Overlay-titleWrap">
      <h1 class="Overlay-title " id="feedback-dialog-title">
        Provide feedback
      </h1>

    </div>
    <div class="Overlay-actionWrap">
      <button data-close-dialog-id="feedback-dialog" aria-label="Close" type="button" data-view-component="true" class="close-button Overlay-closeButton"><svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-x">
    <path d="M3.72 3.72a.75.75 0 0 1 1.06 0L8 6.94l3.22-3.22a.749.749 0 0 1 1.275.326.749.749 0 0 1-.215.734L9.06 8l3.22 3.22a.749.749 0 0 1-.326 1.275.749.749 0 0 1-.734-.215L8 9.06l-3.22 3.22a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042L6.94 8 3.72 4.78a.75.75 0 0 1 0-1.06Z"></path>
</svg></button>
    </div>
  </div>

</div>
      <scrollable-region data-labelled-by="feedback-dialog-title">
        <div data-view-component="true" class="Overlay-body">        <!-- '"` --><!-- </textarea></xmp> --></option></form><form id="code-search-feedback-form" data-turbo="false" action="/search/feedback" accept-charset="UTF-8" method="post"><input type="hidden" data-csrf="true" name="authenticity_token" value="Gnv9J0nc1T6qJOWPayE/tvSbvOMqarncba7luzEjyE4B6HfBqQo1X5oRuQGHd3z96LHyWXPn1FVZuRVJpHJmVA==" />
          <p>We read every piece of feedback, and take your input very seriously.</p>
          <textarea name="feedback" class="form-control width-full mb-2" style="height: 120px" id="feedback"></textarea>
          <input name="include_email" id="include_email" aria-label="Include my email address so I can be contacted" class="form-control mr-2" type="checkbox">
          <label for="include_email" style="font-weight: normal">Include my email address so I can be contacted</label>
</form></div>
      </scrollable-region>
      <div data-view-component="true" class="Overlay-footer Overlay-footer--alignEnd">          <button data-close-dialog-id="feedback-dialog" type="button" data-view-component="true" class="btn">    Cancel
</button>
          <button form="code-search-feedback-form" data-action="click:qbsearch-input#submitFeedback" type="submit" data-view-component="true" class="btn-primary btn">    Submit feedback
</button>
</div>
</dialog></dialog-helper>

    <custom-scopes data-target="qbsearch-input.customScopesManager">

<dialog-helper>
  <dialog data-target="custom-scopes.customScopesModalDialog" data-action="close:qbsearch-input#handleDialogClose cancel:qbsearch-input#handleDialogClose" id="custom-scopes-dialog" aria-modal="true" aria-labelledby="custom-scopes-dialog-title" aria-describedby="custom-scopes-dialog-description" data-view-component="true" class="Overlay Overlay-whenNarrow Overlay--size-medium Overlay--motion-scaleFade Overlay--disableScroll">
    <div data-view-component="true" class="Overlay-header Overlay-header--divided">
  <div class="Overlay-headerContentWrap">
    <div class="Overlay-titleWrap">
      <h1 class="Overlay-title " id="custom-scopes-dialog-title">
        Saved searches
      </h1>
        <h2 id="custom-scopes-dialog-description" class="Overlay-description">Use saved searches to filter your results more quickly</h2>
    </div>
    <div class="Overlay-actionWrap">
      <button data-close-dialog-id="custom-scopes-dialog" aria-label="Close" type="button" data-view-component="true" class="close-button Overlay-closeButton"><svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-x">
    <path d="M3.72 3.72a.75.75 0 0 1 1.06 0L8 6.94l3.22-3.22a.749.749 0 0 1 1.275.326.749.749 0 0 1-.215.734L9.06 8l3.22 3.22a.749.749 0 0 1-.326 1.275.749.749 0 0 1-.734-.215L8 9.06l-3.22 3.22a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042L6.94 8 3.72 4.78a.75.75 0 0 1 0-1.06Z"></path>
</svg></button>
    </div>
  </div>

</div>
      <scrollable-region data-labelled-by="custom-scopes-dialog-title">
        <div data-view-component="true" class="Overlay-body">        <div data-target="custom-scopes.customScopesModalDialogFlash"></div>

        <div hidden class="create-custom-scope-form" data-target="custom-scopes.createCustomScopeForm">
        <!-- '"` --><!-- </textarea></xmp> --></option></form><form id="custom-scopes-dialog-form" data-turbo="false" action="/search/custom_scopes" accept-charset="UTF-8" method="post"><input type="hidden" data-csrf="true" name="authenticity_token" value="gDdkiAcoikPH5ymnEah8LHWkTGbg3eZZhSZi7OSbtsRr0vxheOYGr/0tb0ZDZRe8FbscLH86ws6ez4Re60q2Jg==" />
          <div data-target="custom-scopes.customScopesModalDialogFlash"></div>

          <input type="hidden" id="custom_scope_id" name="custom_scope_id" data-target="custom-scopes.customScopesIdField">

          <div class="form-group">
            <label for="custom_scope_name">Name</label>
            <auto-check src="/search/custom_scopes/check_name" required only-validate-on-blur="false">
              <input
                type="text"
                name="custom_scope_name"
                id="custom_scope_name"
                data-target="custom-scopes.customScopesNameField"
                class="form-control"
                autocomplete="off"
                placeholder="github-ruby"
                required
                maxlength="50">
              <input type="hidden" data-csrf="true" value="ER0njA4SbDKCdcdr29Vt496k8plpZGDOumpXEAg4U+SVZjpVkphBKqvgnzHlQlWDsrN3UfjwSfJw6gQ2hBPQsA==" />
            </auto-check>
          </div>

          <div class="form-group">
            <label for="custom_scope_query">Query</label>
            <input
              type="text"
              name="custom_scope_query"
              id="custom_scope_query"
              data-target="custom-scopes.customScopesQueryField"
              class="form-control"
              autocomplete="off"
              placeholder="(repo:mona/a OR repo:mona/b) AND lang:python"
              required
              maxlength="500">
          </div>

          <p class="text-small color-fg-muted">
            To see all available qualifiers, see our <a class="Link--inTextBlock" href="https://docs.github.com/search-github/github-code-search/understanding-github-code-search-syntax">documentation</a>.
          </p>
</form>        </div>

        <div data-target="custom-scopes.manageCustomScopesForm">
          <div data-target="custom-scopes.list"></div>
        </div>

</div>
      </scrollable-region>
      <div data-view-component="true" class="Overlay-footer Overlay-footer--alignEnd Overlay-footer--divided">          <button data-action="click:custom-scopes#customScopesCancel" type="button" data-view-component="true" class="btn">    Cancel
</button>
          <button form="custom-scopes-dialog-form" data-action="click:custom-scopes#customScopesSubmit" data-target="custom-scopes.customScopesSubmitButton" type="submit" data-view-component="true" class="btn-primary btn">    Create saved search
</button>
</div>
</dialog></dialog-helper>
    </custom-scopes>
  </div>
</qbsearch-input>

            <div class="position-relative HeaderMenu-link-wrap d-lg-inline-block">
              <a
                href="/login?return_to=https%3A%2F%2Fgithub.com%2Fjohncsantiago%2FWhartonLab%2Fblob%2Fmaster%2FG85R%2520GOandKEGGTermEnrichment.Rmd"
                class="HeaderMenu-link HeaderMenu-link--sign-in HeaderMenu-button flex-shrink-0 no-underline d-none d-lg-inline-flex border border-lg-0 rounded rounded-lg-0 px-2 py-1"
                style="margin-left: 12px;"
                data-hydro-click="{&quot;event_type&quot;:&quot;authentication.click&quot;,&quot;payload&quot;:{&quot;location_in_page&quot;:&quot;site header menu&quot;,&quot;repository_id&quot;:null,&quot;auth_type&quot;:&quot;SIGN_UP&quot;,&quot;originating_url&quot;:&quot;https://github.com/johncsantiago/WhartonLab/blob/master/G85R%20GOandKEGGTermEnrichment.Rmd&quot;,&quot;user_id&quot;:null}}" data-hydro-click-hmac="8157e09adb62d5d215219a237bffb5c7f645d21a9e7ca38ff2b80dd99879c182"
                data-analytics-event="{&quot;category&quot;:&quot;Marketing nav&quot;,&quot;action&quot;:&quot;click to go to homepage&quot;,&quot;label&quot;:&quot;ref_page:Marketing;ref_cta:Sign in;ref_loc:Header&quot;}"
              >
                Sign in
              </a>
            </div>

              <a href="/signup?ref_cta=Sign+up&amp;ref_loc=header+logged+out&amp;ref_page=%2F%3Cuser-name%3E%2F%3Crepo-name%3E%2Fblob%2Fshow&amp;source=header-repo&amp;source_repo=johncsantiago%2FWhartonLab"
                class="HeaderMenu-link HeaderMenu-link--sign-up HeaderMenu-button flex-shrink-0 d-flex d-lg-inline-flex no-underline border color-border-default rounded px-2 py-1"
                data-hydro-click="{&quot;event_type&quot;:&quot;authentication.click&quot;,&quot;payload&quot;:{&quot;location_in_page&quot;:&quot;site header menu&quot;,&quot;repository_id&quot;:null,&quot;auth_type&quot;:&quot;SIGN_UP&quot;,&quot;originating_url&quot;:&quot;https://github.com/johncsantiago/WhartonLab/blob/master/G85R%20GOandKEGGTermEnrichment.Rmd&quot;,&quot;user_id&quot;:null}}" data-hydro-click-hmac="8157e09adb62d5d215219a237bffb5c7f645d21a9e7ca38ff2b80dd99879c182"
                data-analytics-event="{&quot;category&quot;:&quot;Sign up&quot;,&quot;action&quot;:&quot;click to sign up for account&quot;,&quot;label&quot;:&quot;ref_page:/&lt;user-name&gt;/&lt;repo-name&gt;/blob/show;ref_cta:Sign up;ref_loc:header logged out&quot;}"
              >
                Sign up
              </a>
          <button type="button" class="sr-only js-header-menu-focus-trap d-block d-lg-none">Reseting focus</button>
        </div>
      </div>
    </div>
  </div>
</header>

      <div hidden="hidden" data-view-component="true" class="js-stale-session-flash stale-session-flash flash flash-warn flash-full">

        <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-alert">
    <path d="M6.457 1.047c.659-1.234 2.427-1.234 3.086 0l6.082 11.378A1.75 1.75 0 0 1 14.082 15H1.918a1.75 1.75 0 0 1-1.543-2.575Zm1.763.707a.25.25 0 0 0-.44 0L1.698 13.132a.25.25 0 0 0 .22.368h12.164a.25.25 0 0 0 .22-.368Zm.53 3.996v2.5a.75.75 0 0 1-1.5 0v-2.5a.75.75 0 0 1 1.5 0ZM9 11a1 1 0 1 1-2 0 1 1 0 0 1 2 0Z"></path>
</svg>
        <span class="js-stale-session-flash-signed-in" hidden>You signed in with another tab or window. <a class="Link--inTextBlock" href="">Reload</a> to refresh your session.</span>
        <span class="js-stale-session-flash-signed-out" hidden>You signed out in another tab or window. <a class="Link--inTextBlock" href="">Reload</a> to refresh your session.</span>
        <span class="js-stale-session-flash-switched" hidden>You switched accounts on another tab or window. <a class="Link--inTextBlock" href="">Reload</a> to refresh your session.</span>

    <button id="icon-button-8ac41a50-4fb0-4254-a46c-f80af46be707" aria-labelledby="tooltip-9d09fa68-e1a2-483f-a20c-013febb2eba7" type="button" data-view-component="true" class="Button Button--iconOnly Button--invisible Button--medium flash-close js-flash-close">  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-x Button-visual">
    <path d="M3.72 3.72a.75.75 0 0 1 1.06 0L8 6.94l3.22-3.22a.749.749 0 0 1 1.275.326.749.749 0 0 1-.215.734L9.06 8l3.22 3.22a.749.749 0 0 1-.326 1.275.749.749 0 0 1-.734-.215L8 9.06l-3.22 3.22a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042L6.94 8 3.72 4.78a.75.75 0 0 1 0-1.06Z"></path>
</svg>
</button><tool-tip id="tooltip-9d09fa68-e1a2-483f-a20c-013febb2eba7" for="icon-button-8ac41a50-4fb0-4254-a46c-f80af46be707" popover="manual" data-direction="s" data-type="label" data-view-component="true" class="sr-only position-absolute">Dismiss alert</tool-tip>

</div>
    </div>

  <div id="start-of-content" class="show-on-focus"></div>

    <div id="js-flash-container" class="flash-container" data-turbo-replace>

  <template class="js-flash-template">

<div class="flash flash-full   {{ className }}">
  <div >
    <button autofocus class="flash-close js-flash-close" type="button" aria-label="Dismiss this message">
      <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-x">
    <path d="M3.72 3.72a.75.75 0 0 1 1.06 0L8 6.94l3.22-3.22a.749.749 0 0 1 1.275.326.749.749 0 0 1-.215.734L9.06 8l3.22 3.22a.749.749 0 0 1-.326 1.275.749.749 0 0 1-.734-.215L8 9.06l-3.22 3.22a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042L6.94 8 3.72 4.78a.75.75 0 0 1 0-1.06Z"></path>
</svg>
    </button>
    <div aria-atomic="true" role="alert" class="js-flash-alert">

      <div>{{ message }}</div>

    </div>
  </div>
</div>
  </template>
</div>

  <div
    class="application-main "
    data-commit-hovercards-enabled
    data-discussion-hovercards-enabled
    data-issue-and-pr-hovercards-enabled
    data-project-hovercards-enabled
  >
        <div itemscope itemtype="http://schema.org/SoftwareSourceCode" class="">
    <main id="js-repo-pjax-container" >

  <div id="repository-container-header"  class="pt-3 hide-full-screen" style="background-color: var(--page-header-bgColor, var(--color-page-header-bg));" data-turbo-replace>

      <div class="d-flex flex-nowrap flex-justify-end mb-3  px-3 px-lg-5" style="gap: 1rem;">

        <div class="flex-auto min-width-0 width-fit">

  <div class=" d-flex flex-wrap flex-items-center wb-break-word f3 text-normal">
      <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-repo color-fg-muted mr-2">
    <path d="M2 2.5A2.5 2.5 0 0 1 4.5 0h8.75a.75.75 0 0 1 .75.75v12.5a.75.75 0 0 1-.75.75h-2.5a.75.75 0 0 1 0-1.5h1.75v-2h-8a1 1 0 0 0-.714 1.7.75.75 0 1 1-1.072 1.05A2.495 2.495 0 0 1 2 11.5Zm10.5-1h-8a1 1 0 0 0-1 1v6.708A2.486 2.486 0 0 1 4.5 9h8ZM5 12.25a.25.25 0 0 1 .25-.25h3.5a.25.25 0 0 1 .25.25v3.25a.25.25 0 0 1-.4.2l-1.45-1.087a.249.249 0 0 0-.3 0L5.4 15.7a.25.25 0 0 1-.4-.2Z"></path>
</svg>

    <span class="author flex-self-stretch" itemprop="author">
      <a class="url fn" rel="author" data-hovercard-type="user" data-hovercard-url="/users/johncsantiago/hovercard" data-octo-click="hovercard-link-click" data-octo-dimensions="link_type:self" href="/johncsantiago">
        johncsantiago
</a>    </span>
    <span class="mx-1 flex-self-stretch color-fg-muted">/</span>
    <strong itemprop="name" class="mr-2 flex-self-stretch">
      <a data-pjax="#repo-content-pjax-container" data-turbo-frame="repo-content-turbo-frame" href="/johncsantiago/WhartonLab">WhartonLab</a>
    </strong>

    <span></span><span class="Label Label--secondary v-align-middle mr-1">Public</span>
  </div>

        </div>

        <div id="repository-details-container" class="flex-shrink-0" data-turbo-replace style="max-width: 70%;">
            <ul class="pagehead-actions flex-shrink-0 d-none d-md-inline" style="padding: 2px 0;">

  <li>
            <a href="/login?return_to=%2Fjohncsantiago%2FWhartonLab" rel="nofollow" id="repository-details-watch-button" data-hydro-click="{&quot;event_type&quot;:&quot;authentication.click&quot;,&quot;payload&quot;:{&quot;location_in_page&quot;:&quot;notification subscription menu watch&quot;,&quot;repository_id&quot;:null,&quot;auth_type&quot;:&quot;LOG_IN&quot;,&quot;originating_url&quot;:&quot;https://github.com/johncsantiago/WhartonLab/blob/master/G85R%20GOandKEGGTermEnrichment.Rmd&quot;,&quot;user_id&quot;:null}}" data-hydro-click-hmac="8244cbc4c1a4b029c45cd0615ed29bc227ec668b7844c1ca75aab6d397c24db1" aria-label="You must be signed in to change notification settings" data-view-component="true" class="btn-sm btn">    <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-bell mr-2">
    <path d="M8 16a2 2 0 0 0 1.985-1.75c.017-.137-.097-.25-.235-.25h-3.5c-.138 0-.252.113-.235.25A2 2 0 0 0 8 16ZM3 5a5 5 0 0 1 10 0v2.947c0 .05.015.098.042.139l1.703 2.555A1.519 1.519 0 0 1 13.482 13H2.518a1.516 1.516 0 0 1-1.263-2.36l1.703-2.554A.255.255 0 0 0 3 7.947Zm5-3.5A3.5 3.5 0 0 0 4.5 5v2.947c0 .346-.102.683-.294.97l-1.703 2.556a.017.017 0 0 0-.003.01l.001.006c0 .002.002.004.004.006l.006.004.007.001h10.964l.007-.001.006-.004.004-.006.001-.007a.017.017 0 0 0-.003-.01l-1.703-2.554a1.745 1.745 0 0 1-.294-.97V5A3.5 3.5 0 0 0 8 1.5Z"></path>
</svg>Notifications
</a>    <tool-tip id="tooltip-b4447439-038d-46a9-bf1c-ee353a2b7ccc" for="repository-details-watch-button" popover="manual" data-direction="s" data-type="description" data-view-component="true" class="sr-only position-absolute">You must be signed in to change notification settings</tool-tip>

  </li>

  <li>
          <a icon="repo-forked" id="fork-button" href="/login?return_to=%2Fjohncsantiago%2FWhartonLab" rel="nofollow" data-hydro-click="{&quot;event_type&quot;:&quot;authentication.click&quot;,&quot;payload&quot;:{&quot;location_in_page&quot;:&quot;repo details fork button&quot;,&quot;repository_id&quot;:652065041,&quot;auth_type&quot;:&quot;LOG_IN&quot;,&quot;originating_url&quot;:&quot;https://github.com/johncsantiago/WhartonLab/blob/master/G85R%20GOandKEGGTermEnrichment.Rmd&quot;,&quot;user_id&quot;:null}}" data-hydro-click-hmac="335d006439cb21fbc111dc1e7b4b9a54ac69fd9352fc1100ce32016a28ccd086" data-view-component="true" class="btn-sm btn">    <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-repo-forked mr-2">
    <path d="M5 5.372v.878c0 .414.336.75.75.75h4.5a.75.75 0 0 0 .75-.75v-.878a2.25 2.25 0 1 1 1.5 0v.878a2.25 2.25 0 0 1-2.25 2.25h-1.5v2.128a2.251 2.251 0 1 1-1.5 0V8.5h-1.5A2.25 2.25 0 0 1 3.5 6.25v-.878a2.25 2.25 0 1 1 1.5 0ZM5 3.25a.75.75 0 1 0-1.5 0 .75.75 0 0 0 1.5 0Zm6.75.75a.75.75 0 1 0 0-1.5.75.75 0 0 0 0 1.5Zm-3 8.75a.75.75 0 1 0-1.5 0 .75.75 0 0 0 1.5 0Z"></path>
</svg>Fork
    <span id="repo-network-counter" data-pjax-replace="true" data-turbo-replace="true" title="0" data-view-component="true" class="Counter">0</span>
</a>
  </li>

  <li>
        <div data-view-component="true" class="BtnGroup d-flex">
        <a href="/login?return_to=%2Fjohncsantiago%2FWhartonLab" rel="nofollow" data-hydro-click="{&quot;event_type&quot;:&quot;authentication.click&quot;,&quot;payload&quot;:{&quot;location_in_page&quot;:&quot;star button&quot;,&quot;repository_id&quot;:652065041,&quot;auth_type&quot;:&quot;LOG_IN&quot;,&quot;originating_url&quot;:&quot;https://github.com/johncsantiago/WhartonLab/blob/master/G85R%20GOandKEGGTermEnrichment.Rmd&quot;,&quot;user_id&quot;:null}}" data-hydro-click-hmac="dd6f8ce3b9985cac24c44f08eb511ee6753cb65ce5b3eebae54c3f0b04ec78c4" aria-label="You must be signed in to star a repository" data-view-component="true" class="tooltipped tooltipped-sw btn-sm btn">    <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-star v-align-text-bottom d-inline-block mr-2">
    <path d="M8 .25a.75.75 0 0 1 .673.418l1.882 3.815 4.21.612a.75.75 0 0 1 .416 1.279l-3.046 2.97.719 4.192a.751.751 0 0 1-1.088.791L8 12.347l-3.766 1.98a.75.75 0 0 1-1.088-.79l.72-4.194L.818 6.374a.75.75 0 0 1 .416-1.28l4.21-.611L7.327.668A.75.75 0 0 1 8 .25Zm0 2.445L6.615 5.5a.75.75 0 0 1-.564.41l-3.097.45 2.24 2.184a.75.75 0 0 1 .216.664l-.528 3.084 2.769-1.456a.75.75 0 0 1 .698 0l2.77 1.456-.53-3.084a.75.75 0 0 1 .216-.664l2.24-2.183-3.096-.45a.75.75 0 0 1-.564-.41L8 2.694Z"></path>
</svg><span data-view-component="true" class="d-inline">
          Star
</span>          <span id="repo-stars-counter-star" aria-label="0 users starred this repository" data-singular-suffix="user starred this repository" data-plural-suffix="users starred this repository" data-turbo-replace="true" title="0" data-view-component="true" class="Counter js-social-count">0</span>
</a></div>
  </li>

</ul>

        </div>
      </div>

        <div id="responsive-meta-container" data-turbo-replace>
</div>

          <nav data-pjax="#js-repo-pjax-container" aria-label="Repository" data-view-component="true" class="js-repo-nav js-sidenav-container-pjax js-responsive-underlinenav overflow-hidden UnderlineNav px-3 px-md-4 px-lg-5">

  <ul data-view-component="true" class="UnderlineNav-body list-style-none">
      <li data-view-component="true" class="d-inline-flex">
  <a id="code-tab" href="/johncsantiago/WhartonLab" data-tab-item="i0code-tab" data-selected-links="repo_source repo_downloads repo_commits repo_releases repo_tags repo_branches repo_packages repo_deployments repo_attestations /johncsantiago/WhartonLab" data-pjax="#repo-content-pjax-container" data-turbo-frame="repo-content-turbo-frame" data-hotkey="g c" data-analytics-event="{&quot;category&quot;:&quot;Underline navbar&quot;,&quot;action&quot;:&quot;Click tab&quot;,&quot;label&quot;:&quot;Code&quot;,&quot;target&quot;:&quot;UNDERLINE_NAV.TAB&quot;}" aria-current="page" data-view-component="true" class="UnderlineNav-item no-wrap js-responsive-underlinenav-item js-selected-navigation-item selected">

              <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-code UnderlineNav-octicon d-none d-sm-inline">
    <path d="m11.28 3.22 4.25 4.25a.75.75 0 0 1 0 1.06l-4.25 4.25a.749.749 0 0 1-1.275-.326.749.749 0 0 1 .215-.734L13.94 8l-3.72-3.72a.749.749 0 0 1 .326-1.275.749.749 0 0 1 .734.215Zm-6.56 0a.751.751 0 0 1 1.042.018.751.751 0 0 1 .018 1.042L2.06 8l3.72 3.72a.749.749 0 0 1-.326 1.275.749.749 0 0 1-.734-.215L.47 8.53a.75.75 0 0 1 0-1.06Z"></path>
</svg>
        <span data-content="Code">Code</span>
          <span id="code-repo-tab-count" data-pjax-replace="" data-turbo-replace="" title="Not available" data-view-component="true" class="Counter"></span>

</a></li>
      <li data-view-component="true" class="d-inline-flex">
  <a id="issues-tab" href="/johncsantiago/WhartonLab/issues" data-tab-item="i1issues-tab" data-selected-links="repo_issues repo_labels repo_milestones /johncsantiago/WhartonLab/issues" data-pjax="#repo-content-pjax-container" data-turbo-frame="repo-content-turbo-frame" data-hotkey="g i" data-analytics-event="{&quot;category&quot;:&quot;Underline navbar&quot;,&quot;action&quot;:&quot;Click tab&quot;,&quot;label&quot;:&quot;Issues&quot;,&quot;target&quot;:&quot;UNDERLINE_NAV.TAB&quot;}" data-view-component="true" class="UnderlineNav-item no-wrap js-responsive-underlinenav-item js-selected-navigation-item">

              <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-issue-opened UnderlineNav-octicon d-none d-sm-inline">
    <path d="M8 9.5a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3Z"></path><path d="M8 0a8 8 0 1 1 0 16A8 8 0 0 1 8 0ZM1.5 8a6.5 6.5 0 1 0 13 0 6.5 6.5 0 0 0-13 0Z"></path>
</svg>
        <span data-content="Issues">Issues</span>
          <span id="issues-repo-tab-count" data-pjax-replace="" data-turbo-replace="" title="0" hidden="hidden" data-view-component="true" class="Counter">0</span>

</a></li>
      <li data-view-component="true" class="d-inline-flex">
  <a id="pull-requests-tab" href="/johncsantiago/WhartonLab/pulls" data-tab-item="i2pull-requests-tab" data-selected-links="repo_pulls checks /johncsantiago/WhartonLab/pulls" data-pjax="#repo-content-pjax-container" data-turbo-frame="repo-content-turbo-frame" data-hotkey="g p" data-analytics-event="{&quot;category&quot;:&quot;Underline navbar&quot;,&quot;action&quot;:&quot;Click tab&quot;,&quot;label&quot;:&quot;Pull requests&quot;,&quot;target&quot;:&quot;UNDERLINE_NAV.TAB&quot;}" data-view-component="true" class="UnderlineNav-item no-wrap js-responsive-underlinenav-item js-selected-navigation-item">

              <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-git-pull-request UnderlineNav-octicon d-none d-sm-inline">
    <path d="M1.5 3.25a2.25 2.25 0 1 1 3 2.122v5.256a2.251 2.251 0 1 1-1.5 0V5.372A2.25 2.25 0 0 1 1.5 3.25Zm5.677-.177L9.573.677A.25.25 0 0 1 10 .854V2.5h1A2.5 2.5 0 0 1 13.5 5v5.628a2.251 2.251 0 1 1-1.5 0V5a1 1 0 0 0-1-1h-1v1.646a.25.25 0 0 1-.427.177L7.177 3.427a.25.25 0 0 1 0-.354ZM3.75 2.5a.75.75 0 1 0 0 1.5.75.75 0 0 0 0-1.5Zm0 9.5a.75.75 0 1 0 0 1.5.75.75 0 0 0 0-1.5Zm8.25.75a.75.75 0 1 0 1.5 0 .75.75 0 0 0-1.5 0Z"></path>
</svg>
        <span data-content="Pull requests">Pull requests</span>
          <span id="pull-requests-repo-tab-count" data-pjax-replace="" data-turbo-replace="" title="0" hidden="hidden" data-view-component="true" class="Counter">0</span>

</a></li>
      <li data-view-component="true" class="d-inline-flex">
  <a id="actions-tab" href="/johncsantiago/WhartonLab/actions" data-tab-item="i3actions-tab" data-selected-links="repo_actions /johncsantiago/WhartonLab/actions" data-pjax="#repo-content-pjax-container" data-turbo-frame="repo-content-turbo-frame" data-hotkey="g a" data-analytics-event="{&quot;category&quot;:&quot;Underline navbar&quot;,&quot;action&quot;:&quot;Click tab&quot;,&quot;label&quot;:&quot;Actions&quot;,&quot;target&quot;:&quot;UNDERLINE_NAV.TAB&quot;}" data-view-component="true" class="UnderlineNav-item no-wrap js-responsive-underlinenav-item js-selected-navigation-item">

              <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-play UnderlineNav-octicon d-none d-sm-inline">
    <path d="M8 0a8 8 0 1 1 0 16A8 8 0 0 1 8 0ZM1.5 8a6.5 6.5 0 1 0 13 0 6.5 6.5 0 0 0-13 0Zm4.879-2.773 4.264 2.559a.25.25 0 0 1 0 .428l-4.264 2.559A.25.25 0 0 1 6 10.559V5.442a.25.25 0 0 1 .379-.215Z"></path>
</svg>
        <span data-content="Actions">Actions</span>
          <span id="actions-repo-tab-count" data-pjax-replace="" data-turbo-replace="" title="Not available" data-view-component="true" class="Counter"></span>

</a></li>
      <li data-view-component="true" class="d-inline-flex">
  <a id="projects-tab" href="/johncsantiago/WhartonLab/projects" data-tab-item="i4projects-tab" data-selected-links="repo_projects new_repo_project repo_project /johncsantiago/WhartonLab/projects" data-pjax="#repo-content-pjax-container" data-turbo-frame="repo-content-turbo-frame" data-hotkey="g b" data-analytics-event="{&quot;category&quot;:&quot;Underline navbar&quot;,&quot;action&quot;:&quot;Click tab&quot;,&quot;label&quot;:&quot;Projects&quot;,&quot;target&quot;:&quot;UNDERLINE_NAV.TAB&quot;}" data-view-component="true" class="UnderlineNav-item no-wrap js-responsive-underlinenav-item js-selected-navigation-item">

              <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-table UnderlineNav-octicon d-none d-sm-inline">
    <path d="M0 1.75C0 .784.784 0 1.75 0h12.5C15.216 0 16 .784 16 1.75v12.5A1.75 1.75 0 0 1 14.25 16H1.75A1.75 1.75 0 0 1 0 14.25ZM6.5 6.5v8h7.75a.25.25 0 0 0 .25-.25V6.5Zm8-1.5V1.75a.25.25 0 0 0-.25-.25H6.5V5Zm-13 1.5v7.75c0 .138.112.25.25.25H5v-8ZM5 5V1.5H1.75a.25.25 0 0 0-.25.25V5Z"></path>
</svg>
        <span data-content="Projects">Projects</span>
          <span id="projects-repo-tab-count" data-pjax-replace="" data-turbo-replace="" title="0" hidden="hidden" data-view-component="true" class="Counter">0</span>

</a></li>
      <li data-view-component="true" class="d-inline-flex">
  <a id="security-tab" href="/johncsantiago/WhartonLab/security" data-tab-item="i5security-tab" data-selected-links="security overview alerts policy token_scanning code_scanning /johncsantiago/WhartonLab/security" data-pjax="#repo-content-pjax-container" data-turbo-frame="repo-content-turbo-frame" data-hotkey="g s" data-analytics-event="{&quot;category&quot;:&quot;Underline navbar&quot;,&quot;action&quot;:&quot;Click tab&quot;,&quot;label&quot;:&quot;Security&quot;,&quot;target&quot;:&quot;UNDERLINE_NAV.TAB&quot;}" data-view-component="true" class="UnderlineNav-item no-wrap js-responsive-underlinenav-item js-selected-navigation-item">

              <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-shield UnderlineNav-octicon d-none d-sm-inline">
    <path d="M7.467.133a1.748 1.748 0 0 1 1.066 0l5.25 1.68A1.75 1.75 0 0 1 15 3.48V7c0 1.566-.32 3.182-1.303 4.682-.983 1.498-2.585 2.813-5.032 3.855a1.697 1.697 0 0 1-1.33 0c-2.447-1.042-4.049-2.357-5.032-3.855C1.32 10.182 1 8.566 1 7V3.48a1.75 1.75 0 0 1 1.217-1.667Zm.61 1.429a.25.25 0 0 0-.153 0l-5.25 1.68a.25.25 0 0 0-.174.238V7c0 1.358.275 2.666 1.057 3.86.784 1.194 2.121 2.34 4.366 3.297a.196.196 0 0 0 .154 0c2.245-.956 3.582-2.104 4.366-3.298C13.225 9.666 13.5 8.36 13.5 7V3.48a.251.251 0 0 0-.174-.237l-5.25-1.68ZM8.75 4.75v3a.75.75 0 0 1-1.5 0v-3a.75.75 0 0 1 1.5 0ZM9 10.5a1 1 0 1 1-2 0 1 1 0 0 1 2 0Z"></path>
</svg>
        <span data-content="Security">Security</span>
          <include-fragment src="/johncsantiago/WhartonLab/security/overall-count" accept="text/fragment+html"></include-fragment>

</a></li>
      <li data-view-component="true" class="d-inline-flex">
  <a id="insights-tab" href="/johncsantiago/WhartonLab/pulse" data-tab-item="i6insights-tab" data-selected-links="repo_graphs repo_contributors dependency_graph dependabot_updates pulse people community /johncsantiago/WhartonLab/pulse" data-pjax="#repo-content-pjax-container" data-turbo-frame="repo-content-turbo-frame" data-analytics-event="{&quot;category&quot;:&quot;Underline navbar&quot;,&quot;action&quot;:&quot;Click tab&quot;,&quot;label&quot;:&quot;Insights&quot;,&quot;target&quot;:&quot;UNDERLINE_NAV.TAB&quot;}" data-view-component="true" class="UnderlineNav-item no-wrap js-responsive-underlinenav-item js-selected-navigation-item">

              <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-graph UnderlineNav-octicon d-none d-sm-inline">
    <path d="M1.5 1.75V13.5h13.75a.75.75 0 0 1 0 1.5H.75a.75.75 0 0 1-.75-.75V1.75a.75.75 0 0 1 1.5 0Zm14.28 2.53-5.25 5.25a.75.75 0 0 1-1.06 0L7 7.06 4.28 9.78a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042l3.25-3.25a.75.75 0 0 1 1.06 0L10 7.94l4.72-4.72a.751.751 0 0 1 1.042.018.751.751 0 0 1 .018 1.042Z"></path>
</svg>
        <span data-content="Insights">Insights</span>
          <span id="insights-repo-tab-count" data-pjax-replace="" data-turbo-replace="" title="Not available" data-view-component="true" class="Counter"></span>

</a></li>
</ul>
    <div style="visibility:hidden;" data-view-component="true" class="UnderlineNav-actions js-responsive-underlinenav-overflow position-absolute pr-3 pr-md-4 pr-lg-5 right-0">      <action-menu data-select-variant="none" data-view-component="true">
  <focus-group direction="vertical" mnemonics retain>
    <button id="action-menu-b06f3d73-2e5b-4c9a-8154-7b17f5f33f9a-button" popovertarget="action-menu-b06f3d73-2e5b-4c9a-8154-7b17f5f33f9a-overlay" aria-controls="action-menu-b06f3d73-2e5b-4c9a-8154-7b17f5f33f9a-list" aria-haspopup="true" aria-labelledby="tooltip-2edaa656-737c-4dfb-b4c5-e986240a7302" type="button" data-view-component="true" class="Button Button--iconOnly Button--secondary Button--medium UnderlineNav-item">  <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-kebab-horizontal Button-visual">
    <path d="M8 9a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3ZM1.5 9a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3Zm13 0a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3Z"></path>
</svg>
</button><tool-tip id="tooltip-2edaa656-737c-4dfb-b4c5-e986240a7302" for="action-menu-b06f3d73-2e5b-4c9a-8154-7b17f5f33f9a-button" popover="manual" data-direction="s" data-type="label" data-view-component="true" class="sr-only position-absolute">Additional navigation options</tool-tip>

<anchored-position data-target="action-menu.overlay" id="action-menu-b06f3d73-2e5b-4c9a-8154-7b17f5f33f9a-overlay" anchor="action-menu-b06f3d73-2e5b-4c9a-8154-7b17f5f33f9a-button" align="start" side="outside-bottom" anchor-offset="normal" popover="auto" data-view-component="true">
  <div data-view-component="true" class="Overlay Overlay--size-auto">

      <div data-view-component="true" class="Overlay-body Overlay-body--paddingNone">          <action-list>
  <div data-view-component="true">
    <ul aria-labelledby="action-menu-b06f3d73-2e5b-4c9a-8154-7b17f5f33f9a-button" id="action-menu-b06f3d73-2e5b-4c9a-8154-7b17f5f33f9a-list" role="menu" data-view-component="true" class="ActionListWrap--inset ActionListWrap">
        <li hidden="hidden" data-menu-item="i0code-tab" data-targets="action-list.items" role="none" data-view-component="true" class="ActionListItem">

    <a tabindex="-1" id="item-ffbd1f9f-20d9-4440-b6ff-4b9088c05edf" href="/johncsantiago/WhartonLab" role="menuitem" data-view-component="true" class="ActionListContent ActionListContent--visual16">
        <span class="ActionListItem-visual ActionListItem-visual--leading">
          <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-code">
    <path d="m11.28 3.22 4.25 4.25a.75.75 0 0 1 0 1.06l-4.25 4.25a.749.749 0 0 1-1.275-.326.749.749 0 0 1 .215-.734L13.94 8l-3.72-3.72a.749.749 0 0 1 .326-1.275.749.749 0 0 1 .734.215Zm-6.56 0a.751.751 0 0 1 1.042.018.751.751 0 0 1 .018 1.042L2.06 8l3.72 3.72a.749.749 0 0 1-.326 1.275.749.749 0 0 1-.734-.215L.47 8.53a.75.75 0 0 1 0-1.06Z"></path>
</svg>
        </span>

        <span data-view-component="true" class="ActionListItem-label">
          Code
</span>      
</a>

</li>
        <li hidden="hidden" data-menu-item="i1issues-tab" data-targets="action-list.items" role="none" data-view-component="true" class="ActionListItem">

    <a tabindex="-1" id="item-54d1bbaa-754e-49ba-b0b3-4747d0f94401" href="/johncsantiago/WhartonLab/issues" role="menuitem" data-view-component="true" class="ActionListContent ActionListContent--visual16">
        <span class="ActionListItem-visual ActionListItem-visual--leading">
          <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-issue-opened">
    <path d="M8 9.5a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3Z"></path><path d="M8 0a8 8 0 1 1 0 16A8 8 0 0 1 8 0ZM1.5 8a6.5 6.5 0 1 0 13 0 6.5 6.5 0 0 0-13 0Z"></path>
</svg>
        </span>

        <span data-view-component="true" class="ActionListItem-label">
          Issues
</span>      
</a>

</li>
        <li hidden="hidden" data-menu-item="i2pull-requests-tab" data-targets="action-list.items" role="none" data-view-component="true" class="ActionListItem">

    <a tabindex="-1" id="item-a199e744-b7f6-4d15-8672-a7c0cad70009" href="/johncsantiago/WhartonLab/pulls" role="menuitem" data-view-component="true" class="ActionListContent ActionListContent--visual16">
        <span class="ActionListItem-visual ActionListItem-visual--leading">
          <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-git-pull-request">
    <path d="M1.5 3.25a2.25 2.25 0 1 1 3 2.122v5.256a2.251 2.251 0 1 1-1.5 0V5.372A2.25 2.25 0 0 1 1.5 3.25Zm5.677-.177L9.573.677A.25.25 0 0 1 10 .854V2.5h1A2.5 2.5 0 0 1 13.5 5v5.628a2.251 2.251 0 1 1-1.5 0V5a1 1 0 0 0-1-1h-1v1.646a.25.25 0 0 1-.427.177L7.177 3.427a.25.25 0 0 1 0-.354ZM3.75 2.5a.75.75 0 1 0 0 1.5.75.75 0 0 0 0-1.5Zm0 9.5a.75.75 0 1 0 0 1.5.75.75 0 0 0 0-1.5Zm8.25.75a.75.75 0 1 0 1.5 0 .75.75 0 0 0-1.5 0Z"></path>
</svg>
        </span>

        <span data-view-component="true" class="ActionListItem-label">
          Pull requests
</span>      
</a>

</li>
        <li hidden="hidden" data-menu-item="i3actions-tab" data-targets="action-list.items" role="none" data-view-component="true" class="ActionListItem">

    <a tabindex="-1" id="item-f343815f-09aa-4fef-b51d-fea509f84738" href="/johncsantiago/WhartonLab/actions" role="menuitem" data-view-component="true" class="ActionListContent ActionListContent--visual16">
        <span class="ActionListItem-visual ActionListItem-visual--leading">
          <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-play">
    <path d="M8 0a8 8 0 1 1 0 16A8 8 0 0 1 8 0ZM1.5 8a6.5 6.5 0 1 0 13 0 6.5 6.5 0 0 0-13 0Zm4.879-2.773 4.264 2.559a.25.25 0 0 1 0 .428l-4.264 2.559A.25.25 0 0 1 6 10.559V5.442a.25.25 0 0 1 .379-.215Z"></path>
</svg>
        </span>

        <span data-view-component="true" class="ActionListItem-label">
          Actions
</span>      
</a>

</li>
        <li hidden="hidden" data-menu-item="i4projects-tab" data-targets="action-list.items" role="none" data-view-component="true" class="ActionListItem">

    <a tabindex="-1" id="item-65a98546-9e8a-467e-bf20-732f4f55dc02" href="/johncsantiago/WhartonLab/projects" role="menuitem" data-view-component="true" class="ActionListContent ActionListContent--visual16">
        <span class="ActionListItem-visual ActionListItem-visual--leading">
          <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-table">
    <path d="M0 1.75C0 .784.784 0 1.75 0h12.5C15.216 0 16 .784 16 1.75v12.5A1.75 1.75 0 0 1 14.25 16H1.75A1.75 1.75 0 0 1 0 14.25ZM6.5 6.5v8h7.75a.25.25 0 0 0 .25-.25V6.5Zm8-1.5V1.75a.25.25 0 0 0-.25-.25H6.5V5Zm-13 1.5v7.75c0 .138.112.25.25.25H5v-8ZM5 5V1.5H1.75a.25.25 0 0 0-.25.25V5Z"></path>
</svg>
        </span>

        <span data-view-component="true" class="ActionListItem-label">
          Projects
</span>      
</a>

</li>
        <li hidden="hidden" data-menu-item="i5security-tab" data-targets="action-list.items" role="none" data-view-component="true" class="ActionListItem">

    <a tabindex="-1" id="item-06043754-6f11-4072-b197-78f6e45bebdc" href="/johncsantiago/WhartonLab/security" role="menuitem" data-view-component="true" class="ActionListContent ActionListContent--visual16">
        <span class="ActionListItem-visual ActionListItem-visual--leading">
          <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-shield">
    <path d="M7.467.133a1.748 1.748 0 0 1 1.066 0l5.25 1.68A1.75 1.75 0 0 1 15 3.48V7c0 1.566-.32 3.182-1.303 4.682-.983 1.498-2.585 2.813-5.032 3.855a1.697 1.697 0 0 1-1.33 0c-2.447-1.042-4.049-2.357-5.032-3.855C1.32 10.182 1 8.566 1 7V3.48a1.75 1.75 0 0 1 1.217-1.667Zm.61 1.429a.25.25 0 0 0-.153 0l-5.25 1.68a.25.25 0 0 0-.174.238V7c0 1.358.275 2.666 1.057 3.86.784 1.194 2.121 2.34 4.366 3.297a.196.196 0 0 0 .154 0c2.245-.956 3.582-2.104 4.366-3.298C13.225 9.666 13.5 8.36 13.5 7V3.48a.251.251 0 0 0-.174-.237l-5.25-1.68ZM8.75 4.75v3a.75.75 0 0 1-1.5 0v-3a.75.75 0 0 1 1.5 0ZM9 10.5a1 1 0 1 1-2 0 1 1 0 0 1 2 0Z"></path>
</svg>
        </span>

        <span data-view-component="true" class="ActionListItem-label">
          Security
</span>      
</a>

</li>
        <li hidden="hidden" data-menu-item="i6insights-tab" data-targets="action-list.items" role="none" data-view-component="true" class="ActionListItem">

    <a tabindex="-1" id="item-b05a8f3e-1d44-4f52-9a0f-23cd410a2461" href="/johncsantiago/WhartonLab/pulse" role="menuitem" data-view-component="true" class="ActionListContent ActionListContent--visual16">
        <span class="ActionListItem-visual ActionListItem-visual--leading">
          <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-graph">
    <path d="M1.5 1.75V13.5h13.75a.75.75 0 0 1 0 1.5H.75a.75.75 0 0 1-.75-.75V1.75a.75.75 0 0 1 1.5 0Zm14.28 2.53-5.25 5.25a.75.75 0 0 1-1.06 0L7 7.06 4.28 9.78a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042l3.25-3.25a.75.75 0 0 1 1.06 0L10 7.94l4.72-4.72a.751.751 0 0 1 1.042.018.751.751 0 0 1 .018 1.042Z"></path>
</svg>
        </span>

        <span data-view-component="true" class="ActionListItem-label">
          Insights
</span>      
</a>

</li>
</ul>    
</div></action-list>

</div>

</div></anchored-position>  </focus-group>
</action-menu></div>
</nav>

  </div>

<turbo-frame id="repo-content-turbo-frame" target="_top" data-turbo-action="advance" class="">
    <div id="repo-content-pjax-container" class="repository-content " >

<react-app
  app-name="react-code-view"
  initial-path="/johncsantiago/WhartonLab/blob/master/G85R%20GOandKEGGTermEnrichment.Rmd"
    style="display: block; min-height: calc(100vh - 64px);"
  data-attempted-ssr="true"
  data-ssr="true"
  data-lazy="false"
  data-alternate="false"
  data-data-router-enabled="false"
>

  <script type="application/json" data-target="react-app.embeddedData">{"payload":{"allShortcutsEnabled":false,"fileTree":{"":{"items":[{"name":"A4V_BodySections_RNAseq","path":"A4V_BodySections_RNAseq","contentType":"directory"},{"name":"Anna qPCR data","path":"Anna qPCR data","contentType":"directory"},{"name":"Barron","path":"Barron","contentType":"directory"},{"name":"ConspiracyNetwork","path":"ConspiracyNetwork","contentType":"directory"},{"name":"DeficiencyScreen","path":"DeficiencyScreen","contentType":"directory"},{"name":"DinaFolder","path":"DinaFolder","contentType":"directory"},{"name":"DrosophilaMotifsSingletonData","path":"DrosophilaMotifsSingletonData","contentType":"directory"},{"name":"G85R OXPHOS KEGG Pathways","path":"G85R OXPHOS KEGG Pathways","contentType":"directory"},{"name":"GeneralDataFiles","path":"GeneralDataFiles","contentType":"directory"},{"name":"Grants","path":"Grants","contentType":"directory"},{"name":"KEGG_Image_Files","path":"KEGG_Image_Files","contentType":"directory"},{"name":"Metabolomics","path":"Metabolomics","contentType":"directory"},{"name":"Papers","path":"Papers","contentType":"directory"},{"name":"RussoPaper","path":"RussoPaper","contentType":"directory"},{"name":"TKT_RNAseq","path":"TKT_RNAseq","contentType":"directory"},{"name":"alexa","path":"alexa","contentType":"directory"},{"name":"gbb_RNAseq","path":"gbb_RNAseq","contentType":"directory"},{"name":"gbbmeetingfigures_files","path":"gbbmeetingfigures_files","contentType":"directory"},{"name":"mtDNA-nucDNA analysis","path":"mtDNA-nucDNA analysis","contentType":"directory"},{"name":"qPCR","path":"qPCR","contentType":"directory"},{"name":".DS_Store","path":".DS_Store","contentType":"file"},{"name":"20240507_pharate_axonal_branching.Rmd","path":"20240507_pharate_axonal_branching.Rmd","contentType":"file"},{"name":"20240507_pharate_axonal_branching.html","path":"20240507_pharate_axonal_branching.html","contentType":"file"},{"name":"20240507_pharate_axonal_branching_Raw_Data.csv","path":"20240507_pharate_axonal_branching_Raw_Data.csv","contentType":"file"},{"name":"20240513_pharate_axonal_morphology_Raw_Data.csv","path":"20240513_pharate_axonal_morphology_Raw_Data.csv","contentType":"file"},{"name":"20240514_AxonalMorphologyResults.Rmd","path":"20240514_AxonalMorphologyResults.Rmd","contentType":"file"},{"name":"20240514_AxonalMorphologyResults.html","path":"20240514_AxonalMorphologyResults.html","contentType":"file"},{"name":"20240514_AxonalMorphologyResults.pdf","path":"20240514_AxonalMorphologyResults.pdf","contentType":"file"},{"name":"20240925_Santiago_LabMeeting.pptx","path":"20240925_Santiago_LabMeeting.pptx","contentType":"file"},{"name":"20241011_arc1presentation.pptx","path":"20241011_arc1presentation.pptx","contentType":"file"},{"name":"A4V BMP Signaling Genes Analysis.Rmd","path":"A4V BMP Signaling Genes Analysis.Rmd","contentType":"file"},{"name":"A4V-BMP-Signaling-Genes-Analysis.html","path":"A4V-BMP-Signaling-Genes-Analysis.html","contentType":"file"},{"name":"A4V_mt_vs_nuc_genesPCA.png","path":"A4V_mt_vs_nuc_genesPCA.png","contentType":"file"},{"name":"Aleah_Thesis_Figures.Rmd","path":"Aleah_Thesis_Figures.Rmd","contentType":"file"},{"name":"Arc1CcsNetwork.Rmd","path":"Arc1CcsNetwork.Rmd","contentType":"file"},{"name":"Arc1CcsNetwork.html","path":"Arc1CcsNetwork.html","contentType":"file"},{"name":"Arc1MeetingFigures.Rmd","path":"Arc1MeetingFigures.Rmd","contentType":"file"},{"name":"Arc1Sod1Network.csv","path":"Arc1Sod1Network.csv","contentType":"file"},{"name":"BMP Signaling Genes annotated.csv","path":"BMP Signaling Genes annotated.csv","contentType":"file"},{"name":"BMP Signaling Genes.csv","path":"BMP Signaling Genes.csv","contentType":"file"},{"name":"Cami qPCR Thesis Figures.Rmd","path":"Cami qPCR Thesis Figures.Rmd","contentType":"file"},{"name":"Cami-qPCR-Thesis-Figures.html","path":"Cami-qPCR-Thesis-Figures.html","contentType":"file"},{"name":"CamiCode.Rmd","path":"CamiCode.Rmd","contentType":"file"},{"name":"CamiCode.html","path":"CamiCode.html","contentType":"file"},{"name":"CamiData.Rmd","path":"CamiData.Rmd","contentType":"file"},{"name":"Cami_qPCR_dCt_data.csv","path":"Cami_qPCR_dCt_data.csv","contentType":"file"},{"name":"Cami_qPCR_dCt_data.xlsx","path":"Cami_qPCR_dCt_data.xlsx","contentType":"file"},{"name":"Cami_qpcr 2_28_24 - Sheet1.Rmd","path":"Cami_qpcr 2_28_24 - Sheet1.Rmd","contentType":"file"},{"name":"Cami_qpcr 2_28_24 - Sheet1.csv","path":"Cami_qpcr 2_28_24 - Sheet1.csv","contentType":"file"},{"name":"Dphalerata_male.tif.png","path":"Dphalerata_male.tif.png","contentType":"file"},{"name":"EmilyNetwork.Rmd","path":"EmilyNetwork.Rmd","contentType":"file"},{"name":"EmilyNetwork.csv","path":"EmilyNetwork.csv","contentType":"file"},{"name":"EmilyNetwork.html","path":"EmilyNetwork.html","contentType":"file"},{"name":"EmilyTBPH_qPCR.csv","path":"EmilyTBPH_qPCR.csv","contentType":"file"},{"name":"EmilyTBPH_qPCR.xlsx","path":"EmilyTBPH_qPCR.xlsx","contentType":"file"},{"name":"EmilyTBPHanalysis.Rmd","path":"EmilyTBPHanalysis.Rmd","contentType":"file"},{"name":"Emily_tbph_eclosion.csv","path":"Emily_tbph_eclosion.csv","contentType":"file"},{"name":"FlashTalkModifierData.csv","path":"FlashTalkModifierData.csv","contentType":"file"},{"name":"FlashTalkModifierFigures.Rmd","path":"FlashTalkModifierFigures.Rmd","contentType":"file"},{"name":"FlashTalkModifierFigures.html","path":"FlashTalkModifierFigures.html","contentType":"file"},{"name":"Functions_OmicsQuickPlotFunctions.R","path":"Functions_OmicsQuickPlotFunctions.R","contentType":"file"},{"name":"G85R GOandKEGGTermEnrichment.Rmd","path":"G85R GOandKEGGTermEnrichment.Rmd","contentType":"file"},{"name":"G85R-GOandKEGGTermEnrichment.html","path":"G85R-GOandKEGGTermEnrichment.html","contentType":"file"},{"name":"G85R.GOandKEGGPlottingFunctions.R","path":"G85R.GOandKEGGPlottingFunctions.R","contentType":"file"},{"name":"G85R_OXPHOS_KEGG_Pathways.Rmd","path":"G85R_OXPHOS_KEGG_Pathways.Rmd","contentType":"file"},{"name":"GOandKEGGPlottingFunctions.R","path":"GOandKEGGPlottingFunctions.R","contentType":"file"},{"name":"GOandKEGGTermEnrichment.Rmd","path":"GOandKEGGTermEnrichment.Rmd","contentType":"file"},{"name":"GOandKEGGTermEnrichment.html","path":"GOandKEGGTermEnrichment.html","contentType":"file"},{"name":"GeneExpressionBoxPlots.Rmd","path":"GeneExpressionBoxPlots.Rmd","contentType":"file"},{"name":"GeneExpressionBoxPlots.html","path":"GeneExpressionBoxPlots.html","contentType":"file"},{"name":"JohnSantiago_mtDNAxnucDNAqPCR_Notes.docx","path":"JohnSantiago_mtDNAxnucDNAqPCR_Notes.docx","contentType":"file"},{"name":"KristiPresentationData.Rmd","path":"KristiPresentationData.Rmd","contentType":"file"},{"name":"KristiPresentationData.html","path":"KristiPresentationData.html","contentType":"file"},{"name":"NAD.edges.csv","path":"NAD.edges.csv","contentType":"file"},{"name":"NewEmilyNetwork.csv","path":"NewEmilyNetwork.csv","contentType":"file"},{"name":"NewEmilyNetwork.xlsx","path":"NewEmilyNetwork.xlsx","contentType":"file"},{"name":"Nguyen Serpe 2024 scRNA-seq VNC.csv","path":"Nguyen Serpe 2024 scRNA-seq VNC.csv","contentType":"file"},{"name":"OmicsQuickPlotFunctions.Rmd","path":"OmicsQuickPlotFunctions.Rmd","contentType":"file"},{"name":"OmicsQuickPlotFunctions.html","path":"OmicsQuickPlotFunctions.html","contentType":"file"},{"name":"Paranhos 2021 SOD1 Review.pdf","path":"Paranhos 2021 SOD1 Review.pdf","contentType":"file"},{"name":"PunnettSquareApp.Rmd","path":"PunnettSquareApp.Rmd","contentType":"file"},{"name":"PunnettSquareApp.html","path":"PunnettSquareApp.html","contentType":"file"},{"name":"QiagenADArrayGenes.csv","path":"QiagenADArrayGenes.csv","contentType":"file"},{"name":"Russo Fig 1 eclosion.csv","path":"Russo Fig 1 eclosion.csv","contentType":"file"},{"name":"Russo Fig 1 eclosion_TukeyDATA.csv","path":"Russo Fig 1 eclosion_TukeyDATA.csv","contentType":"file"},{"name":"SG Kankel.csv","path":"SG Kankel.csv","contentType":"file"},{"name":"SG Kankel.xlsx","path":"SG Kankel.xlsx","contentType":"file"},{"name":"Serpe 2024 GliaSubtypes.csv","path":"Serpe 2024 GliaSubtypes.csv","contentType":"file"},{"name":"Serpe 2024 InterneuronClusters.csv","path":"Serpe 2024 InterneuronClusters.csv","contentType":"file"},{"name":"Serpe 2024 MNClusters.csv","path":"Serpe 2024 MNClusters.csv","contentType":"file"},{"name":"SerpeSCdata.Rmd","path":"SerpeSCdata.Rmd","contentType":"file"},{"name":"SerpeSCdata.html","path":"SerpeSCdata.html","contentType":"file"},{"name":"SerpeSCdata.pdf","path":"SerpeSCdata.pdf","contentType":"file"},{"name":"Shortcuts_OmicsQuickPlotFunctions.Rmd","path":"Shortcuts_OmicsQuickPlotFunctions.Rmd","contentType":"file"},{"name":"Shortcuts_OmicsQuickPlotFunctions.html","path":"Shortcuts_OmicsQuickPlotFunctions.html","contentType":"file"},{"name":"UB7 - G85R Pharate Leg Axonal Branching \u0026 Mitochondria  - CSV_Branching.csv","path":"UB7 - G85R Pharate Leg Axonal Branching \u0026 Mitochondria  - CSV_Branching.csv","contentType":"file"},{"name":"arc1presentation.pptx","path":"arc1presentation.pptx","contentType":"file"},{"name":"dme00010.GRFCvWTFC  and  GRMCvWTMC.multi.png","path":"dme00010.GRFCvWTFC  and  GRMCvWTMC.multi.png","contentType":"file"},{"name":"dme00010.GRxWT.FC_GRxWT.MC.multi.png","path":"dme00010.GRxWT.FC_GRxWT.MC.multi.png","contentType":"file"},{"name":"dme00010.png","path":"dme00010.png","contentType":"file"},{"name":"dme00010.xml","path":"dme00010.xml","contentType":"file"},{"name":"dme00030.G85RvsWT_FemaleandMale.multi.png","path":"dme00030.G85RvsWT_FemaleandMale.multi.png","contentType":"file"},{"name":"dme00030.png","path":"dme00030.png","contentType":"file"},{"name":"dme00030.xml","path":"dme00030.xml","contentType":"file"},{"name":"dme00190.A4V_FandM_9T.png","path":"dme00190.A4V_FandM_9T.png","contentType":"file"},{"name":"dme00190.G85RandWT_CxOE_Genes with GR.CxOE_Metab.multi.png","path":"dme00190.G85RandWT_CxOE_Genes with GR.CxOE_Metab.multi.png","contentType":"file"},{"name":"dme00190.gbbKOxgbbWT.multi.png","path":"dme00190.gbbKOxgbbWT.multi.png","contentType":"file"},{"name":"dme00190.png","path":"dme00190.png","contentType":"file"},{"name":"dme00190.xml","path":"dme00190.xml","contentType":"file"},{"name":"dme00563.A4V_FandM_9T.multi.png","path":"dme00563.A4V_FandM_9T.multi.png","contentType":"file"},{"name":"dme00563.A4V_FandM_9T.png","path":"dme00563.A4V_FandM_9T.png","contentType":"file"},{"name":"dme00563.png","path":"dme00563.png","contentType":"file"},{"name":"dme00563.xml","path":"dme00563.xml","contentType":"file"},{"name":"dme00760.A9MTvS9MT  and  A40FTvS40FT.multi.png","path":"dme00760.A9MTvS9MT  and  A40FTvS40FT.multi.png","contentType":"file"},{"name":"dme00760.GRFCvWTFC  and  GRMCvWTMC.multi.png","path":"dme00760.GRFCvWTFC  and  GRMCvWTMC.multi.png","contentType":"file"},{"name":"dme00760.gbbKOxgbbWT.multi.png","path":"dme00760.gbbKOxgbbWT.multi.png","contentType":"file"},{"name":"dme00760.png","path":"dme00760.png","contentType":"file"},{"name":"dme00760.xml","path":"dme00760.xml","contentType":"file"},{"name":"gbbmeetingfigures.Rmd","path":"gbbmeetingfigures.Rmd","contentType":"file"},{"name":"mt.nuc PCA for grant.Rmd","path":"mt.nuc PCA for grant.Rmd","contentType":"file"},{"name":"mt.nuc-PCA-for-grant.html","path":"mt.nuc-PCA-for-grant.html","contentType":"file"},{"name":"mt.nuc.OXPHOS_PCAdata.csv","path":"mt.nuc.OXPHOS_PCAdata.csv","contentType":"file"},{"name":"~$arc1presentation.pptx","path":"~$arc1presentation.pptx","contentType":"file"}],"totalCount":123}},"fileTreeProcessingTime":17.475392,"foldersToFetch":[],"repo":{"id":652065041,"defaultBranch":"master","name":"WhartonLab","ownerLogin":"johncsantiago","currentUserCanPush":false,"isFork":false,"isEmpty":false,"createdAt":"2023-06-11T01:12:53.000Z","ownerAvatar":"https://avatars.githubusercontent.com/u/61519162?v=4","public":true,"private":false,"isOrgOwned":false},"codeLineWrapEnabled":false,"symbolsExpanded":false,"treeExpanded":true,"refInfo":{"name":"master","listCacheKey":"v0:1686446179.162525","canEdit":false,"refType":"branch","currentOid":"19a37d6cf6739514e404cc231008021c1d4ca7e8"},"path":"G85R GOandKEGGTermEnrichment.Rmd","currentUser":null,"blob":{"rawLines":["---","title: \"GO and KEGG Term Enrichment\"","author: \"John Santiago\"","date: \"2024-05-08\"","output: html_document","---","","```{r setup, include=FALSE}","","knitr::opts_chunk$set(echo = TRUE)","","```","","","","```{r echo = F, message = F}","","# Conditions","#Geno is WT or GR","#TKT is either \"C\", \"OE\", or \"Df\"","#Sex is \"F\" or \"M\"","","Geno1 = \"GR\"","TKT1 = \"C\"","Sex1 = \"F\"","","Geno2 = \"WT\"","TKT2 = \"C\"","Sex2 = \"F\"","","##Conditions for second comparison","Geno3 = \"GR\"","TKT3 = \"C\"","Sex3 = \"M\"","","Geno4 = \"WT\"","TKT4 = \"C\"","Sex4 = \"M\"","","","","","","```","","```{r echo = F, message = F}","","source('/Users/johncsantiago/Documents/GitHub/WhartonLab/G85R.GOandKEGGPlottingFunctions.R')","","plot.data = G85R.data(Geno1, Geno2, Geno3, Geno4,","                      Sex1, Sex2, Sex3, Sex4,","                      TKT1, TKT2, TKT3, TKT4)","","```","","","```{r echo = F}","","##Uses data from primary comparison","##Plots genes with an FDR \u003c= 0.05","##point size is relative to maximum mean expression in the two conditions","","G85R.volcano(plot.data)","","```","","","```{r echo = F}","","##Uses data from primary comparison","##point size is relative to max mean expression in any condition","##grey points indicate less than 1 fold difference between the comparisons","G85R.compare.genes(plot.data)","","```","","","```{r echo = F}","","##Uses data from primary comparison","##point size is relative to mean expression in variable condition","G85R.DEG.venn(plot.data)","","```","","```{r echo = F, include = F, warning = F}","","####Uses sigDEGs from primary comparison","cat.data = G85R.Enrichment(plot.data)","","```","","","```{r echo = F}","","","##which ranks of categories to use","##a vector of numbers specifying which ranks to use","##eg. c(1, 5:6) will use the first, fifth and sixth most sig. cats.","##can also just type 'sigs' ","usecats=c(1:20)","#usecats = 'sigs'  ","","G85R.KEGG20(cat.data, plot.data, usecats)","","```","","","```{r echo = F, include = T}","","##can be the exact name or 5 digit code in quotes","specific.kegg = \"00980\"","","G85R.plotkegg.genes(specific.kegg, plot.data)","","```","","","```{r echo = F, message=F, fig.width=7}","","##5 digits from KEGG ID","PathwayID = \"00010\"","","#options include 'primary' or 'both'","use = 'both'","","","G85R.KEGG.diagram(PathwayID, plot.data, use, sigs.only = F)","","","```","","","```{r echo = F, include = T, warnings = F}","","##can be the exact name or 5 digit code in quotes","specific.kegg = \"00010\"","","G85R.comparekegg.genes(specific.kegg, plot.data)","","```","","","```{r echo = F}","","## BP, CC, or MF","GO.ontology = 'BP'","","##can select which categories ot include or say 'sigs.only'","usecats=c(1:20)","#usecats = 'sigs.only'","","G85R.GO20(cat.data, GO.ontology, plot.data, usecats)","","```","","","```{r echo = F}","","##the 7 digits of a GOterm id or the exact term name","specific.go = '0005811'","","G85R.plotgo.genes(specific.go, plot.data)","","```","","","```{r echo = F, warning = F}","","##the 7 digits of a GOterm id or the exact term name","specific.go = '0005811'","","G85R.comparego.genes(specific.go, plot.data)","","```","","","```{r echo = F, warning = F, message = F}","","","## BP, CC, or MF","GO.ontology = 'BP'","","rt = tree(cat.data, GO.ontology)","  ","treemapPlot(rt)","","```","","","```{r echo = F, warning = F}","","##a parent category generated in the treemapPlot function","parent.category = 'neurogenesis'","","G85R.plotparentcat.genes(parent.category, plot.data)","","```","","","```{r echo = F}","","##a parent category generated in the treemapPlot function","##parent.category = 'small molecule biosynthetic process'","","G85R.compareparent.genes(parent.category, plot.data)","##dot size is relative to the max FDR of either condition","","","```  "],"stylingDirectives":[[[0,3,"pl-s"]],[[0,5,"pl-ent"],[7,36,"pl-s"],[7,8,"pl-pds"],[35,36,"pl-pds"]],[[0,6,"pl-ent"],[8,23,"pl-s"],[8,9,"pl-pds"],[22,23,"pl-pds"]],[[0,4,"pl-ent"],[6,18,"pl-s"],[6,7,"pl-pds"],[17,18,"pl-pds"]],[[0,6,"pl-ent"],[8,21,"pl-s"]],[[0,3,"pl-s"]],[],[[0,3,"pl-s"],[3,5,"pl-en"],[27,27,"pl-c1"]],[[0,0,"pl-c1"]],[[0,34,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[],[],[[0,3,"pl-s"],[3,5,"pl-en"],[28,28,"pl-c1"]],[[0,0,"pl-c1"]],[[0,12,"pl-c1"]],[[0,17,"pl-c1"]],[[0,33,"pl-c1"]],[[0,18,"pl-c1"]],[[0,0,"pl-c1"]],[[0,12,"pl-c1"]],[[0,10,"pl-c1"]],[[0,10,"pl-c1"]],[[0,0,"pl-c1"]],[[0,12,"pl-c1"]],[[0,10,"pl-c1"]],[[0,10,"pl-c1"]],[[0,0,"pl-c1"]],[[0,34,"pl-c1"]],[[0,12,"pl-c1"]],[[0,10,"pl-c1"]],[[0,10,"pl-c1"]],[[0,0,"pl-c1"]],[[0,12,"pl-c1"]],[[0,10,"pl-c1"]],[[0,10,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[[0,3,"pl-s"],[3,5,"pl-en"],[28,28,"pl-c1"]],[[0,0,"pl-c1"]],[[0,92,"pl-c1"]],[[0,0,"pl-c1"]],[[0,49,"pl-c1"]],[[0,45,"pl-c1"]],[[0,45,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[],[[0,3,"pl-s"],[3,5,"pl-en"],[15,15,"pl-c1"]],[[0,0,"pl-c1"]],[[0,35,"pl-c1"]],[[0,33,"pl-c1"]],[[0,73,"pl-c1"]],[[0,0,"pl-c1"]],[[0,23,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[],[[0,3,"pl-s"],[3,5,"pl-en"],[15,15,"pl-c1"]],[[0,0,"pl-c1"]],[[0,35,"pl-c1"]],[[0,64,"pl-c1"]],[[0,74,"pl-c1"]],[[0,29,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[],[[0,3,"pl-s"],[3,5,"pl-en"],[15,15,"pl-c1"]],[[0,0,"pl-c1"]],[[0,35,"pl-c1"]],[[0,65,"pl-c1"]],[[0,24,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[[0,3,"pl-s"],[3,5,"pl-en"],[41,41,"pl-c1"]],[[0,0,"pl-c1"]],[[0,40,"pl-c1"]],[[0,37,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[],[[0,3,"pl-s"],[3,5,"pl-en"],[15,15,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"]],[[0,34,"pl-c1"]],[[0,51,"pl-c1"]],[[0,67,"pl-c1"]],[[0,28,"pl-c1"]],[[0,15,"pl-c1"]],[[0,19,"pl-c1"]],[[0,0,"pl-c1"]],[[0,41,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[],[[0,3,"pl-s"],[3,5,"pl-en"],[28,28,"pl-c1"]],[[0,0,"pl-c1"]],[[0,49,"pl-c1"]],[[0,23,"pl-c1"]],[[0,0,"pl-c1"]],[[0,45,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[],[[0,3,"pl-s"],[3,5,"pl-en"],[39,39,"pl-c1"]],[[0,0,"pl-c1"]],[[0,23,"pl-c1"]],[[0,19,"pl-c1"]],[[0,0,"pl-c1"]],[[0,36,"pl-c1"]],[[0,12,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"]],[[0,59,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[],[[0,3,"pl-s"],[3,5,"pl-en"],[42,42,"pl-c1"]],[[0,0,"pl-c1"]],[[0,49,"pl-c1"]],[[0,23,"pl-c1"]],[[0,0,"pl-c1"]],[[0,48,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[],[[0,3,"pl-s"],[3,5,"pl-en"],[15,15,"pl-c1"]],[[0,0,"pl-c1"]],[[0,16,"pl-c1"]],[[0,18,"pl-c1"]],[[0,0,"pl-c1"]],[[0,59,"pl-c1"]],[[0,15,"pl-c1"]],[[0,22,"pl-c1"]],[[0,0,"pl-c1"]],[[0,52,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[],[[0,3,"pl-s"],[3,5,"pl-en"],[15,15,"pl-c1"]],[[0,0,"pl-c1"]],[[0,52,"pl-c1"]],[[0,23,"pl-c1"]],[[0,0,"pl-c1"]],[[0,41,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[],[[0,3,"pl-s"],[3,5,"pl-en"],[28,28,"pl-c1"]],[[0,0,"pl-c1"]],[[0,52,"pl-c1"]],[[0,23,"pl-c1"]],[[0,0,"pl-c1"]],[[0,44,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[],[[0,3,"pl-s"],[3,5,"pl-en"],[41,41,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"]],[[0,16,"pl-c1"]],[[0,18,"pl-c1"]],[[0,0,"pl-c1"]],[[0,32,"pl-c1"]],[[0,2,"pl-c1"]],[[0,15,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[],[[0,3,"pl-s"],[3,5,"pl-en"],[28,28,"pl-c1"]],[[0,0,"pl-c1"]],[[0,57,"pl-c1"]],[[0,32,"pl-c1"]],[[0,0,"pl-c1"]],[[0,52,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[],[[0,3,"pl-s"],[3,5,"pl-en"],[15,15,"pl-c1"]],[[0,0,"pl-c1"]],[[0,57,"pl-c1"]],[[0,57,"pl-c1"]],[[0,0,"pl-c1"]],[[0,52,"pl-c1"]],[[0,57,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"]],[[0,0,"pl-c1"],[0,3,"pl-s"]],[],[],[],[],[],[],[],[],[]],"colorizedLines":null,"csv":null,"csvError":null,"dependabotInfo":{"showConfigurationBanner":false,"configFilePath":null,"networkDependabotPath":"/johncsantiago/WhartonLab/network/updates","dismissConfigurationNoticePath":"/settings/dismiss-notice/dependabot_configuration_notice","configurationNoticeDismissed":null},"displayName":"G85R GOandKEGGTermEnrichment.Rmd","displayUrl":"https://github.com/johncsantiago/WhartonLab/blob/master/G85R%20GOandKEGGTermEnrichment.Rmd?raw=true","headerInfo":{"blobSize":"3.31 KB","deleteTooltip":"You must be signed in to make or propose changes","editTooltip":"You must be signed in to make or propose changes","ghDesktopPath":"https://desktop.github.com","isGitLfs":false,"onBranch":true,"shortPath":"8d30178","siteNavLoginPath":"/login?return_to=https%3A%2F%2Fgithub.com%2Fjohncsantiago%2FWhartonLab%2Fblob%2Fmaster%2FG85R%2520GOandKEGGTermEnrichment.Rmd","isCSV":false,"isRichtext":false,"toc":null,"lineInfo":{"truncatedLoc":"219","truncatedSloc":"116"},"mode":"file"},"image":false,"isCodeownersFile":null,"isPlain":false,"isValidLegacyIssueTemplate":false,"issueTemplate":null,"discussionTemplate":null,"language":"RMarkdown","languageID":313,"large":false,"planSupportInfo":{"repoIsFork":null,"repoOwnedByCurrentUser":null,"requestFullPath":"/johncsantiago/WhartonLab/blob/master/G85R%20GOandKEGGTermEnrichment.Rmd","showFreeOrgGatedFeatureMessage":null,"showPlanSupportBanner":null,"upgradeDataAttributes":null,"upgradePath":null},"publishBannersInfo":{"dismissActionNoticePath":"/settings/dismiss-notice/publish_action_from_dockerfile","releasePath":"/johncsantiago/WhartonLab/releases/new?marketplace=true","showPublishActionBanner":false},"rawBlobUrl":"https://github.com/johncsantiago/WhartonLab/raw/refs/heads/master/G85R%20GOandKEGGTermEnrichment.Rmd","renderImageOrRaw":false,"richText":null,"renderedFileInfo":null,"shortPath":null,"symbolsEnabled":true,"tabSize":8,"topBannersInfo":{"overridingGlobalFundingFile":false,"globalPreferredFundingPath":null,"showInvalidCitationWarning":false,"citationHelpUrl":"https://docs.github.com/github/creating-cloning-and-archiving-repositories/creating-a-repository-on-github/about-citation-files","actionsOnboardingTip":null},"truncated":false,"viewable":true,"workflowRedirectUrl":null,"symbols":{"timed_out":false,"not_analyzed":true,"symbols":[]}},"copilotInfo":null,"copilotAccessAllowed":false,"modelsAccessAllowed":false,"csrf_tokens":{"/johncsantiago/WhartonLab/branches":{"post":"dIL6kyg-OOr0ZHS6M7LC7kiRDT7np3TS2GwIBfSUt2B5vsrsbfKfr5woixUGZCJbRVCfaPccXby3DapTEFGqnw"},"/repos/preferences":{"post":"EIjR0pOceLytEYQ0FwJp8y50walEA1hiYHc7wmg6bch6XdzoxWJ_nO0sL6fo0cEgYGh9dh2HjCSh2AgxNut1XQ"}}},"title":"WhartonLab/G85R GOandKEGGTermEnrichment.Rmd at master · johncsantiago/WhartonLab","appPayload":{"helpUrl":"https://docs.github.com","findFileWorkerPath":"/assets-cdn/worker/find-file-worker-2e45fb5767a8.js","findInFileWorkerPath":"/assets-cdn/worker/find-in-file-worker-86832803a5fc.js","githubDevUrl":null,"enabled_features":{"code_nav_ui_events":false,"overview_shared_code_dropdown_button":false,"react_blob_overlay":false,"copilot_conversational_ux_embedding_update":false,"copilot_smell_icebreaker_ux":true,"accessible_code_button":true}}}</script>
  <div data-target="react-app.reactRoot"><style data-styled="true" data-styled-version="5.3.11">.hOfjFo{padding:0;}/*!sc*/
.oDGAe{max-width:100%;margin-left:auto;margin-right:auto;display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-flex-wrap:wrap;-ms-flex-wrap:wrap;flex-wrap:wrap;}/*!sc*/
.kowOcT{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-flex:1 1 100%;-ms-flex:1 1 100%;flex:1 1 100%;-webkit-flex-wrap:wrap;-ms-flex-wrap:wrap;flex-wrap:wrap;max-width:100%;}/*!sc*/
.gISSDQ{width:100%;}/*!sc*/
@media screen and (min-width:544px){.gISSDQ{width:100%;}}/*!sc*/
@media screen and (min-width:768px){.gISSDQ{width:auto;}}/*!sc*/
.cEmWSE{display:none;-webkit-order:1;-ms-flex-order:1;order:1;width:100%;margin-left:0;margin-right:0;-webkit-flex-direction:column-reverse;-ms-flex-direction:column-reverse;flex-direction:column-reverse;margin-bottom:0;min-width:0;}/*!sc*/
@media screen and (min-width:768px){.cEmWSE{width:auto;margin-top:0 !important;margin-bottom:0 !important;position:-webkit-sticky;position:sticky;top:0px;max-height:100vh !important;-webkit-flex-direction:row;-ms-flex-direction:row;flex-direction:row;margin-right:0;height:100vh;}}/*!sc*/
@media print,screen and (max-width:1011px) and (min-width:768px){.cEmWSE{display:none;}}/*!sc*/
.hPvFuC{margin-left:0;margin-right:0;display:none;margin-top:0;}/*!sc*/
@media screen and (min-width:768px){.hPvFuC{margin-left:0 !important;margin-right:0 !important;}}/*!sc*/
.fFSoPl{--pane-min-width:256px;--pane-max-width-diff:511px;--pane-max-width:calc(100vw - var(--pane-max-width-diff));width:100%;padding:0;}/*!sc*/
@media screen and (min-width:544px){}/*!sc*/
@media screen and (min-width:768px){.fFSoPl{width:clamp(var(--pane-min-width),var(--pane-width),var(--pane-max-width));overflow:auto;}}/*!sc*/
@media screen and (min-width:1280px){.fFSoPl{--pane-max-width-diff:959px;}}/*!sc*/
.bTBnTW{height:100%;position:relative;display:none;margin-left:0;}/*!sc*/
.bHLmSv{position:absolute;inset:0 -2px;cursor:col-resize;background-color:transparent;-webkit-transition-delay:0.1s;transition-delay:0.1s;}/*!sc*/
.bHLmSv:hover{background-color:var(--bgColor-neutral-muted,var(--color-neutral-muted,rgba(175,184,193,0.2)));}/*!sc*/
.iKqMNA{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-flex-direction:column;-ms-flex-direction:column;flex-direction:column;-webkit-order:2;-ms-flex-order:2;order:2;-webkit-flex-basis:0;-ms-flex-preferred-size:0;flex-basis:0;-webkit-box-flex:1;-webkit-flex-grow:1;-ms-flex-positive:1;flex-grow:1;-webkit-flex-shrink:1;-ms-flex-negative:1;flex-shrink:1;min-width:1px;margin-right:auto;}/*!sc*/
@media print{.iKqMNA{display:-webkit-box !important;display:-webkit-flex !important;display:-ms-flexbox !important;display:flex !important;}}/*!sc*/
.FxAyp{width:100%;max-width:100%;margin-left:auto;margin-right:auto;-webkit-box-flex:1;-webkit-flex-grow:1;-ms-flex-positive:1;flex-grow:1;padding:0;}/*!sc*/
.leYMvG{margin-left:auto;margin-right:auto;-webkit-flex-direction:column;-ms-flex-direction:column;flex-direction:column;padding-bottom:40px;max-width:100%;margin-top:0;}/*!sc*/
.KMPzq{display:inherit;}/*!sc*/
.hfKjHv{width:100%;}/*!sc*/
.gZWyZE{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;gap:8px;-webkit-flex-direction:column;-ms-flex-direction:column;flex-direction:column;width:100%;}/*!sc*/
.dwYKDk{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-align-items:start;-webkit-box-align:start;-ms-flex-align:start;align-items:start;-webkit-box-pack:justify;-webkit-justify-content:space-between;-ms-flex-pack:justify;justify-content:space-between;gap:8px;}/*!sc*/
.ibcGmb{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-align-items:start;-webkit-box-align:start;-ms-flex-align:start;align-items:start;min-width:0;}/*!sc*/
.hKaEJF{display:block;margin-right:8px;}/*!sc*/
@media screen and (min-width:1360px){.hKaEJF{display:block;}}/*!sc*/
.XosP{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;font-size:14px;}/*!sc*/
.bCKfWo[data-size="medium"]{color:var(--fgColor-muted,var(--color-fg-muted,#656d76));padding-left:8px;padding-right:8px;display:none;}/*!sc*/
@media screen and (max-width:768px){.bCKfWo[data-size="medium"]{display:block;}}/*!sc*/
.gUkoLg{-webkit-box-pack:center;-webkit-justify-content:center;-ms-flex-pack:center;justify-content:center;}/*!sc*/
.dmxRgG[data-size="medium"]{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;}/*!sc*/
.dmxRgG[data-size="medium"] svg{color:var(--fgColor-muted,var(--color-fg-muted,#656d76));}/*!sc*/
.dmxRgG[data-size="medium"] > span{width:inherit;}/*!sc*/
.bZBlpz{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;width:100%;}/*!sc*/
.lhTYNA{margin-right:4px;color:var(--fgColor-muted,var(--color-fg-muted,#656d76));}/*!sc*/
.dbrgmi{font-size:14px;min-width:0;max-width:125px;overflow:hidden;text-overflow:ellipsis;white-space:nowrap;}/*!sc*/
.dHJiml{-webkit-align-self:center;-ms-flex-item-align:center;align-self:center;display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;padding-left:8px;padding-right:8px;min-width:0;}/*!sc*/
.cEytCf{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-flex-direction:row;-ms-flex-direction:row;flex-direction:row;font-size:16px;min-width:0;-webkit-flex-shrink:1;-ms-flex-negative:1;flex-shrink:1;-webkit-flex-wrap:wrap;-ms-flex-wrap:wrap;flex-wrap:wrap;max-width:100%;-webkit-align-items:center;-webkit-box-align:center;-ms-flex-align:center;align-items:center;}/*!sc*/
.fzFXnm{max-width:100%;}/*!sc*/
.iMnkmv{max-width:100%;list-style:none;display:inline-block;}/*!sc*/
.ghzDag{display:inline-block;max-width:100%;}/*!sc*/
.kHuKdh{font-weight:600;}/*!sc*/
.jGhzSQ{font-weight:600;display:inline-block;max-width:100%;font-size:16px;}/*!sc*/
.faNtbn{min-height:32px;display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-align-items:start;-webkit-box-align:start;-ms-flex-align:start;align-items:start;}/*!sc*/
.fmQaBv{margin-left:4px;margin-right:4px;}/*!sc*/
.dwNhzn[data-size="medium"][data-no-visuals]{border-top-left-radius:0;border-bottom-left-radius:0;display:none;}/*!sc*/
.fGwBZA[data-size="medium"][data-no-visuals]{color:var(--fgColor-muted,var(--color-fg-muted,#656d76));}/*!sc*/
.dJxjrT{margin-left:16px;margin-right:16px;}/*!sc*/
.eFxKDQ{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-flex-direction:column;-ms-flex-direction:column;flex-direction:column;}/*!sc*/
.dzCJzi{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-flex-direction:row;-ms-flex-direction:row;flex-direction:row;-webkit-flex-wrap:wrap;-ms-flex-wrap:wrap;flex-wrap:wrap;-webkit-box-pack:justify;-webkit-justify-content:space-between;-ms-flex-pack:justify;justify-content:space-between;-webkit-align-items:center;-webkit-box-align:center;-ms-flex-align:center;align-items:center;gap:8px;min-width:273px;padding:8px;}/*!sc*/
@media screen and (min-width:544px){.dzCJzi{-webkit-flex-wrap:nowrap;-ms-flex-wrap:nowrap;flex-wrap:nowrap;}}/*!sc*/
.ldRxiI{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-flex-direction:row;-ms-flex-direction:row;flex-direction:row;}/*!sc*/
.fVkfyA{width:100%;height:-webkit-fit-content;height:-moz-fit-content;height:fit-content;min-width:0;margin-right:0;}/*!sc*/
.gNAmSV{height:40px;padding-left:4px;padding-bottom:16px;}/*!sc*/
.jNEwzY{-webkit-align-items:center;-webkit-box-align:center;-ms-flex-align:center;align-items:center;}/*!sc*/
.bsDwxw{font-size:12px;-webkit-flex:auto;-ms-flex:auto;flex:auto;padding-right:16px;color:var(--fgColor-muted,var(--color-fg-muted,#656d76));min-width:0;}/*!sc*/
.jdLMhu{top:0px;z-index:4;background:var(--bgColor-default,var(--color-canvas-default));position:-webkit-sticky;position:sticky;}/*!sc*/
.tOISc{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-flex-direction:column;-ms-flex-direction:column;flex-direction:column;width:100%;position:absolute;}/*!sc*/
.hqwSEx{display:none;min-width:0;padding-top:8px;padding-bottom:8px;}/*!sc*/
.bDVoEr{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-align-items:center;-webkit-box-align:center;-ms-flex-align:center;align-items:center;overflow:hidden;margin-left:8px;margin-right:8px;-webkit-flex-direction:row;-ms-flex-direction:row;flex-direction:row;-webkit-box-pack:justify;-webkit-justify-content:space-between;-ms-flex-pack:justify;justify-content:space-between;width:100%;}/*!sc*/
.kYLlPM{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-align-items:center;-webkit-box-align:center;-ms-flex-align:center;align-items:center;}/*!sc*/
.gYjEmn{margin-left:4px;margin-right:8px;}/*!sc*/
.kGqOLL{text-overflow:ellipsis;overflow:hidden;display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;}/*!sc*/
.fHind{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-flex-direction:row;-ms-flex-direction:row;flex-direction:row;font-size:14px;min-width:0;-webkit-flex-shrink:1;-ms-flex-negative:1;flex-shrink:1;-webkit-flex-wrap:wrap;-ms-flex-wrap:wrap;flex-wrap:wrap;max-width:100%;-webkit-align-items:center;-webkit-box-align:center;-ms-flex-align:center;align-items:center;}/*!sc*/
.dnZoUW{font-weight:600;display:inline-block;max-width:100%;font-size:14px;}/*!sc*/
.jRZWlf[data-size="small"]{color:var(--fgColor-default,var(--color-fg-default,#1F2328));margin-left:8px;}/*!sc*/
.kTvpNk{padding-left:8px;padding-top:8px;padding-bottom:8px;display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-flex:1;-ms-flex:1;flex:1;-webkit-align-items:center;-webkit-box-align:center;-ms-flex-align:center;align-items:center;-webkit-box-pack:justify;-webkit-justify-content:space-between;-ms-flex-pack:justify;justify-content:space-between;background-color:var(--bgColor-muted,var(--color-canvas-subtle,#f6f8fa));border:1px solid var(--borderColor-default,var(--color-border-default));border-radius:6px 6px 0px 0px;}/*!sc*/
.iNMjfP{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-align-items:center;-webkit-box-align:center;-ms-flex-align:center;align-items:center;gap:8px;min-width:0;}/*!sc*/
.fefCSX{display:block;position:relative;-webkit-box-flex:1;-webkit-flex-grow:1;-ms-flex-positive:1;flex-grow:1;margin-top:-1px;margin-bottom:-1px;--separator-color:transparent;}/*!sc*/
.fefCSX:not(:last-child){margin-right:1px;}/*!sc*/
.fefCSX:not(:last-child):after{background-color:var(--separator-color);content:"";position:absolute;right:-2px;top:8px;bottom:8px;width:1px;}/*!sc*/
.fefCSX:focus-within:has(:focus-visible){--separator-color:transparent;}/*!sc*/
.fefCSX:first-child{margin-left:-1px;}/*!sc*/
.fefCSX:last-child{margin-right:-1px;}/*!sc*/
.idgUkN{display:block;position:relative;-webkit-box-flex:1;-webkit-flex-grow:1;-ms-flex-positive:1;flex-grow:1;margin-top:-1px;margin-bottom:-1px;--separator-color:var(--borderColor-default,var(--color-border-default,#d0d7de));}/*!sc*/
.idgUkN:not(:last-child){margin-right:1px;}/*!sc*/
.idgUkN:not(:last-child):after{background-color:var(--separator-color);content:"";position:absolute;right:-2px;top:8px;bottom:8px;width:1px;}/*!sc*/
.idgUkN:focus-within:has(:focus-visible){--separator-color:transparent;}/*!sc*/
.idgUkN:first-child{margin-left:-1px;}/*!sc*/
.idgUkN:last-child{margin-right:-1px;}/*!sc*/
.kcLCKF{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-align-items:center;-webkit-box-align:center;-ms-flex-align:center;align-items:center;gap:8px;margin-right:8px;}/*!sc*/
.kVWtTz{gap:8px;}/*!sc*/
.gWqxTd{padding-left:8px;padding-right:8px;}/*!sc*/
.gWqxTd linkButtonSx:hover:not([disabled]){-webkit-text-decoration:none;text-decoration:none;}/*!sc*/
.gWqxTd linkButtonSx:focus:not([disabled]){-webkit-text-decoration:none;text-decoration:none;}/*!sc*/
.gWqxTd linkButtonSx:active:not([disabled]){-webkit-text-decoration:none;text-decoration:none;}/*!sc*/
.ivobqY[data-size="small"][data-no-visuals]{border-top-left-radius:0;border-bottom-left-radius:0;}/*!sc*/
.heuRGy[data-size="small"][data-no-visuals]{color:var(--fgColor-muted,var(--color-fg-muted,#656d76));position:relative;}/*!sc*/
.ffkqe[data-size="small"][data-no-visuals]{color:var(--fgColor-muted,var(--color-fg-muted,#656d76));}/*!sc*/
.hGyMdv{border:1px solid;border-top:none;border-color:var(--borderColor-default,var(--color-border-default,#d0d7de));border-radius:0px 0px 6px 6px;min-width:273px;}/*!sc*/
.dceWRL{background-color:var(--bgColor-default,var(--color-canvas-default));border:0px;border-width:0;border-radius:0px 0px 6px 6px;padding:0;min-width:0;margin-top:46px;}/*!sc*/
.dGXHv{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-flex:1;-ms-flex:1;flex:1;padding-top:8px;padding-bottom:8px;-webkit-flex-direction:column;-ms-flex-direction:column;flex-direction:column;-webkit-box-pack:justify;-webkit-justify-content:space-between;-ms-flex-pack:justify;justify-content:space-between;min-width:0;position:relative;}/*!sc*/
.bpDFns{position:relative;}/*!sc*/
.iJOeCH{-webkit-flex:1;-ms-flex:1;flex:1;position:relative;min-width:0;}/*!sc*/
.jewUnv{tab-size:8;isolation:isolate;position:relative;overflow:auto;max-width:unset;}/*!sc*/
.cCoXib{position:fixed;top:0;right:0;height:100%;width:15px;-webkit-transition:-webkit-transform 0.3s;-webkit-transition:transform 0.3s;transition:transform 0.3s;z-index:1;}/*!sc*/
.cCoXib:hover{-webkit-transform:scaleX(1.5);-ms-transform:scaleX(1.5);transform:scaleX(1.5);}/*!sc*/
data-styled.g1[id="Box-sc-g0xbh4-0"]{content:"hOfjFo,oDGAe,kowOcT,gISSDQ,cEmWSE,hPvFuC,fFSoPl,bTBnTW,bHLmSv,iKqMNA,FxAyp,leYMvG,KMPzq,hfKjHv,gZWyZE,dwYKDk,ibcGmb,hKaEJF,XosP,bCKfWo,gUkoLg,dmxRgG,bZBlpz,lhTYNA,dbrgmi,dHJiml,cEytCf,fzFXnm,iMnkmv,ghzDag,kHuKdh,jGhzSQ,faNtbn,fmQaBv,dwNhzn,fGwBZA,dJxjrT,eFxKDQ,dzCJzi,ldRxiI,fVkfyA,gNAmSV,jNEwzY,bsDwxw,jdLMhu,tOISc,hqwSEx,bDVoEr,kYLlPM,gYjEmn,kGqOLL,fHind,dnZoUW,jRZWlf,kTvpNk,iNMjfP,fefCSX,idgUkN,kcLCKF,kVWtTz,gWqxTd,ivobqY,heuRGy,ffkqe,hGyMdv,dceWRL,dGXHv,bpDFns,iJOeCH,jewUnv,cCoXib,"}/*!sc*/
.eMMFM{min-width:0;}/*!sc*/
.eMMFM:where([data-size='small']){font-size:var(--text-body-size-small,0.75rem);line-height:var(--text-body-lineHeight-small,1.6666);}/*!sc*/
.eMMFM:where([data-size='medium']){font-size:var(--text-body-size-medium,0.875rem);line-height:var(--text-body-lineHeight-medium,1.4285);}/*!sc*/
.eMMFM:where([data-size='large']){font-size:var(--text-body-size-large,1rem);line-height:var(--text-body-lineHeight-large,1.5);}/*!sc*/
.eMMFM:where([data-weight='light']){font-weight:var(--base-text-weight-light,300);}/*!sc*/
.eMMFM:where([data-weight='normal']){font-weight:var(--base-text-weight-normal,400);}/*!sc*/
.eMMFM:where([data-weight='medium']){font-weight:var(--base-text-weight-medium,500);}/*!sc*/
.eMMFM:where([data-weight='semibold']){font-weight:var(--base-text-weight-semibold,600);}/*!sc*/
.HlHVj{padding-left:4px;padding-right:4px;font-weight:400;color:var(--fgColor-muted,var(--color-fg-muted,#656d76));font-size:16px;}/*!sc*/
.HlHVj:where([data-size='small']){font-size:var(--text-body-size-small,0.75rem);line-height:var(--text-body-lineHeight-small,1.6666);}/*!sc*/
.HlHVj:where([data-size='medium']){font-size:var(--text-body-size-medium,0.875rem);line-height:var(--text-body-lineHeight-medium,1.4285);}/*!sc*/
.HlHVj:where([data-size='large']){font-size:var(--text-body-size-large,1rem);line-height:var(--text-body-lineHeight-large,1.5);}/*!sc*/
.HlHVj:where([data-weight='light']){font-weight:var(--base-text-weight-light,300);}/*!sc*/
.HlHVj:where([data-weight='normal']){font-weight:var(--base-text-weight-normal,400);}/*!sc*/
.HlHVj:where([data-weight='medium']){font-weight:var(--base-text-weight-medium,500);}/*!sc*/
.HlHVj:where([data-weight='semibold']){font-weight:var(--base-text-weight-semibold,600);}/*!sc*/
.lauzFl{padding-left:4px;padding-right:4px;font-weight:400;color:var(--fgColor-muted,var(--color-fg-muted,#656d76));font-size:14px;}/*!sc*/
.lauzFl:where([data-size='small']){font-size:var(--text-body-size-small,0.75rem);line-height:var(--text-body-lineHeight-small,1.6666);}/*!sc*/
.lauzFl:where([data-size='medium']){font-size:var(--text-body-size-medium,0.875rem);line-height:var(--text-body-lineHeight-medium,1.4285);}/*!sc*/
.lauzFl:where([data-size='large']){font-size:var(--text-body-size-large,1rem);line-height:var(--text-body-lineHeight-large,1.5);}/*!sc*/
.lauzFl:where([data-weight='light']){font-weight:var(--base-text-weight-light,300);}/*!sc*/
.lauzFl:where([data-weight='normal']){font-weight:var(--base-text-weight-normal,400);}/*!sc*/
.lauzFl:where([data-weight='medium']){font-weight:var(--base-text-weight-medium,500);}/*!sc*/
.lauzFl:where([data-weight='semibold']){font-weight:var(--base-text-weight-semibold,600);}/*!sc*/
data-styled.g3[id="Text__StyledText-sc-17v1xeu-0"]{content:"eMMFM,HlHVj,lauzFl,"}/*!sc*/
.jkNcAv{border:0;font-size:inherit;font-family:inherit;background-color:transparent;-webkit-appearance:none;color:inherit;width:100%;}/*!sc*/
.jkNcAv:focus{outline:0;}/*!sc*/
data-styled.g14[id="UnstyledTextInput__ToggledUnstyledTextInput-sc-14ypya-0"]{content:"jkNcAv,"}/*!sc*/
.hLzFvi{font-size:14px;line-height:var(--base-size-20);color:var(--fgColor-default,var(--color-fg-default,#1F2328));vertical-align:middle;background-color:var(--bgColor-default,var(--color-canvas-default,#ffffff));border:1px solid var(--control-borderColor-rest,var(--borderColor-default,var(--color-border-default,#d0d7de)));border-radius:6px;outline:none;box-shadow:var(--shadow-inset,var(--color-primer-shadow-inset,inset 0 1px 0 rgba(208,215,222,0.2)));display:-webkit-inline-box;display:-webkit-inline-flex;display:-ms-inline-flexbox;display:inline-flex;-webkit-align-items:stretch;-webkit-box-align:stretch;-ms-flex-align:stretch;align-items:stretch;min-height:var(--base-size-32);overflow:hidden;--inner-action-size:var(--base-size-24);}/*!sc*/
.hLzFvi input,.hLzFvi textarea{cursor:text;}/*!sc*/
.hLzFvi select{cursor:pointer;}/*!sc*/
.hLzFvi input::-webkit-input-placeholder,.hLzFvi textarea::-webkit-input-placeholder,.hLzFvi select::-webkit-input-placeholder{color:var(---control-fgColor-placeholder,var(--fgColor-muted,var(--color-fg-muted,#656d76)));}/*!sc*/
.hLzFvi input::-moz-placeholder,.hLzFvi textarea::-moz-placeholder,.hLzFvi select::-moz-placeholder{color:var(---control-fgColor-placeholder,var(--fgColor-muted,var(--color-fg-muted,#656d76)));}/*!sc*/
.hLzFvi input:-ms-input-placeholder,.hLzFvi textarea:-ms-input-placeholder,.hLzFvi select:-ms-input-placeholder{color:var(---control-fgColor-placeholder,var(--fgColor-muted,var(--color-fg-muted,#656d76)));}/*!sc*/
.hLzFvi input::placeholder,.hLzFvi textarea::placeholder,.hLzFvi select::placeholder{color:var(---control-fgColor-placeholder,var(--fgColor-muted,var(--color-fg-muted,#656d76)));}/*!sc*/
.hLzFvi:where([data-trailing-action][data-focused]),.hLzFvi:where(:not([data-trailing-action]):focus-within){border-color:var(--fgColor-accent,var(--color-accent-fg,#0969da));outline:2px solid var(--fgColor-accent,var(--color-accent-fg,#0969da));outline-offset:-1px;}/*!sc*/
.hLzFvi > textarea{padding:var(--base-size-12);}/*!sc*/
.hLzFvi:where([data-contrast]){background-color:var(--bgColor-inset,var(--color-canvas-inset,#f6f8fa));}/*!sc*/
.hLzFvi:where([data-disabled]){color:var(--fgColor-disabled,var(--color-primer-fg-disabled,#8c959f));background-color:var(--control-bgColor-disabled,var(--color-input-disabled-bg,rgba(175,184,193,0.2)));box-shadow:none;border-color:var(--control-borderColor-disabled,var(--borderColor-default,var(--color-border-default,#d0d7de)));}/*!sc*/
.hLzFvi:where([data-disabled]) input,.hLzFvi:where([data-disabled]) textarea,.hLzFvi:where([data-disabled]) select{cursor:not-allowed;}/*!sc*/
.hLzFvi:where([data-monospace]){font-family:var(--fontStack-monospace,SFMono-Regular,Consolas,"Liberation Mono",Menlo,Courier,monospace);}/*!sc*/
.hLzFvi:where([data-validation='error']){border-color:var(--borderColor-danger-emphasis,var(--color-danger-emphasis,#cf222e));}/*!sc*/
.hLzFvi:where([data-validation='error']):where([data-trailing-action][data-focused]),.hLzFvi:where([data-validation='error']):where(:not([data-trailing-action])):focus-within{border-color:var(--fgColor-accent,var(--color-accent-fg,#0969da));outline:2px solid var(--fgColor-accent,var(--color-accent-fg,#0969da));outline-offset:-1px;}/*!sc*/
.hLzFvi:where([data-validation='success']){border-color:var(--bgColor-success-emphasis,var(--color-success-emphasis,#1f883d));}/*!sc*/
.hLzFvi:where([data-block]){width:100%;display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;-webkit-align-self:stretch;-ms-flex-item-align:stretch;align-self:stretch;}/*!sc*/
@media (min-width:768px){.hLzFvi{font-size:var(--text-body-size-medium);}}/*!sc*/
.hLzFvi:where([data-size='small']){--inner-action-size:var(--base-size-20);min-height:var(--base-size-28);padding-top:3px;padding-right:var(--base-size-8);padding-bottom:3px;padding-left:var(--base-size-8);font-size:var(--text-body-size-small);line-height:var(--base-size-20);}/*!sc*/
.hLzFvi:where([data-size='large']){--inner-action-size:var(--base-size-28);height:var(--base-size-40);padding-top:10px;padding-right:var(--base-size-8);padding-bottom:10px;padding-left:var(--base-size-8);}/*!sc*/
.hLzFvi:where([data-variant='small']){min-height:28px;padding-top:3px;padding-right:var(--base-size-8);padding-bottom:3px;padding-left:var(--base-size-8);font-size:(--text-body-size-small);line-height:var(--base-size-20);}/*!sc*/
.hLzFvi:where([data-variant='large']){padding-top:10px;padding-right:var(--base-size-8);padding-bottom:10px;padding-left:var(--base-size-8);font-size:var(--text-title-size-medium);}/*!sc*/
.hLzFvi{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;min-width:160px;}/*!sc*/
data-styled.g15[id="TextInputWrapper__StyledTextInputBaseWrapper-sc-1mqhpbi-0"]{content:"hLzFvi,"}/*!sc*/
.iHYdQq{background-repeat:no-repeat;background-position:right 8px center;padding-right:0;padding-left:0;}/*!sc*/
.iHYdQq > :not(:last-child){margin-right:8px;}/*!sc*/
.iHYdQq .TextInput-icon,.iHYdQq .TextInput-action{-webkit-align-self:center;-ms-flex-item-align:center;align-self:center;color:var(--fgColor-muted,var(--color-fg-muted,#656d76));-webkit-flex-shrink:0;-ms-flex-negative:0;flex-shrink:0;}/*!sc*/
.iHYdQq > input,.iHYdQq > select{padding-right:0;padding-left:0;}/*!sc*/
.iHYdQq:where([data-leading-visual]){padding-left:var(--base-size-12);}/*!sc*/
.iHYdQq:where([data-trailing-visual]:not([data-trailing-action])){padding-right:var(--base-size-12);}/*!sc*/
.iHYdQq:where(:not([data-leading-visual])) > input,.iHYdQq:where(:not([data-leading-visual])) > select{padding-left:var(--base-size-12);}/*!sc*/
.iHYdQq:where(:not([data-trailing-visual]):not([data-trailing-action])) > input,.iHYdQq:where(:not([data-trailing-visual]):not([data-trailing-action])) > select{padding-right:var(--base-size-12);}/*!sc*/
.iHYdQq{display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;min-width:160px;}/*!sc*/
data-styled.g16[id="TextInputWrapper__StyledTextInputWrapper-sc-1mqhpbi-1"]{content:"iHYdQq,"}/*!sc*/
.jOyaRH{display:none;}/*!sc*/
.jOyaRH[popover]{position:absolute;padding:0.5em 0.75em;width:-webkit-max-content;width:-moz-max-content;width:max-content;margin:auto;-webkit-clip:auto;clip:auto;white-space:normal;font:normal normal 11px/1.5 -apple-system,BlinkMacSystemFont,"Segoe UI","Noto Sans",Helvetica,Arial,sans-serif,"Apple Color Emoji","Segoe UI Emoji";-webkit-font-smoothing:subpixel-antialiased;color:var(--tooltip-fgColor,var(--fgColor-onEmphasis,var(--color-fg-on-emphasis,#ffffff)));text-align:center;word-wrap:break-word;background:var(--tooltip-bgColor,var(--bgColor-emphasis,var(--color-neutral-emphasis-plus,#24292f)));border-radius:6px;border:0;opacity:0;max-width:250px;inset:auto;overflow:visible;}/*!sc*/
.jOyaRH[popover]:popover-open{display:block;}/*!sc*/
.jOyaRH[popover].\:popover-open{display:block;}/*!sc*/
@media (forced-colors:active){.jOyaRH{outline:1px solid transparent;}}/*!sc*/
.jOyaRH::after{position:absolute;display:block;right:0;left:0;height:var(--overlay-offset,0.25rem);content:'';}/*!sc*/
.jOyaRH[data-direction='n']::after,.jOyaRH[data-direction='ne']::after,.jOyaRH[data-direction='nw']::after{top:100%;}/*!sc*/
.jOyaRH[data-direction='s']::after,.jOyaRH[data-direction='se']::after,.jOyaRH[data-direction='sw']::after{bottom:100%;}/*!sc*/
.jOyaRH[data-direction='w']::after{position:absolute;display:block;height:100%;width:8px;content:'';bottom:0;left:100%;}/*!sc*/
.jOyaRH[data-direction='e']::after{position:absolute;display:block;height:100%;width:8px;content:'';bottom:0;right:100%;margin-left:-8px;}/*!sc*/
@-webkit-keyframes tooltip-appear{from{opacity:0;}to{opacity:1;}}/*!sc*/
@keyframes tooltip-appear{from{opacity:0;}to{opacity:1;}}/*!sc*/
.jOyaRH:popover-open,.jOyaRH:popover-open::before{-webkit-animation-name:tooltip-appear;animation-name:tooltip-appear;-webkit-animation-duration:0.1s;animation-duration:0.1s;-webkit-animation-fill-mode:forwards;animation-fill-mode:forwards;-webkit-animation-timing-function:ease-in;animation-timing-function:ease-in;-webkit-animation-delay:0s;animation-delay:0s;}/*!sc*/
.jOyaRH.\:popover-open,.jOyaRH.\:popover-open::before{-webkit-animation-name:tooltip-appear;animation-name:tooltip-appear;-webkit-animation-duration:0.1s;animation-duration:0.1s;-webkit-animation-fill-mode:forwards;animation-fill-mode:forwards;-webkit-animation-timing-function:ease-in;animation-timing-function:ease-in;-webkit-animation-delay:0s;animation-delay:0s;}/*!sc*/
data-styled.g17[id="Tooltip__StyledTooltip-sc-e45c7z-0"]{content:"jOyaRH,"}/*!sc*/
.hWlpPn{position:relative;display:inline-block;}/*!sc*/
.hWlpPn::after{position:absolute;z-index:1000000;display:none;padding:0.5em 0.75em;font:normal normal 11px/1.5 -apple-system,BlinkMacSystemFont,"Segoe UI","Noto Sans",Helvetica,Arial,sans-serif,"Apple Color Emoji","Segoe UI Emoji";-webkit-font-smoothing:subpixel-antialiased;color:var(--tooltip-fgColor,var(--fgColor-onEmphasis,var(--color-fg-on-emphasis,#ffffff)));text-align:center;-webkit-text-decoration:none;text-decoration:none;text-shadow:none;text-transform:none;-webkit-letter-spacing:normal;-moz-letter-spacing:normal;-ms-letter-spacing:normal;letter-spacing:normal;word-wrap:break-word;white-space:pre;pointer-events:none;content:attr(aria-label);background:var(--tooltip-bgColor,var(--bgColor-emphasis,var(--color-neutral-emphasis-plus,#24292f)));border-radius:6px;opacity:0;}/*!sc*/
@-webkit-keyframes tooltip-appear{from{opacity:0;}to{opacity:1;}}/*!sc*/
@keyframes tooltip-appear{from{opacity:0;}to{opacity:1;}}/*!sc*/
.hWlpPn:hover::after,.hWlpPn:active::after,.hWlpPn:focus::after,.hWlpPn:focus-within::after{display:inline-block;-webkit-text-decoration:none;text-decoration:none;-webkit-animation-name:tooltip-appear;animation-name:tooltip-appear;-webkit-animation-duration:0.1s;animation-duration:0.1s;-webkit-animation-fill-mode:forwards;animation-fill-mode:forwards;-webkit-animation-timing-function:ease-in;animation-timing-function:ease-in;-webkit-animation-delay:0s;animation-delay:0s;}/*!sc*/
.hWlpPn.tooltipped-no-delay:hover::after,.hWlpPn.tooltipped-no-delay:active::after,.hWlpPn.tooltipped-no-delay:focus::after,.hWlpPn.tooltipped-no-delay:focus-within::after{-webkit-animation-delay:0s;animation-delay:0s;}/*!sc*/
.hWlpPn.tooltipped-multiline:hover::after,.hWlpPn.tooltipped-multiline:active::after,.hWlpPn.tooltipped-multiline:focus::after,.hWlpPn.tooltipped-multiline:focus-within::after{display:table-cell;}/*!sc*/
.hWlpPn.tooltipped-s::after,.hWlpPn.tooltipped-se::after,.hWlpPn.tooltipped-sw::after{top:100%;right:50%;margin-top:6px;}/*!sc*/
.hWlpPn.tooltipped-se::after{right:auto;left:50%;margin-left:-16px;}/*!sc*/
.hWlpPn.tooltipped-sw::after{margin-right:-16px;}/*!sc*/
.hWlpPn.tooltipped-n::after,.hWlpPn.tooltipped-ne::after,.hWlpPn.tooltipped-nw::after{right:50%;bottom:100%;margin-bottom:6px;}/*!sc*/
.hWlpPn.tooltipped-ne::after{right:auto;left:50%;margin-left:-16px;}/*!sc*/
.hWlpPn.tooltipped-nw::after{margin-right:-16px;}/*!sc*/
.hWlpPn.tooltipped-s::after,.hWlpPn.tooltipped-n::after{-webkit-transform:translateX(50%);-ms-transform:translateX(50%);transform:translateX(50%);}/*!sc*/
.hWlpPn.tooltipped-w::after{right:100%;bottom:50%;margin-right:6px;-webkit-transform:translateY(50%);-ms-transform:translateY(50%);transform:translateY(50%);}/*!sc*/
.hWlpPn.tooltipped-e::after{bottom:50%;left:100%;margin-left:6px;-webkit-transform:translateY(50%);-ms-transform:translateY(50%);transform:translateY(50%);}/*!sc*/
.hWlpPn.tooltipped-multiline::after{width:-webkit-max-content;width:-moz-max-content;width:max-content;max-width:250px;word-wrap:break-word;white-space:pre-line;border-collapse:separate;}/*!sc*/
.hWlpPn.tooltipped-multiline.tooltipped-s::after,.hWlpPn.tooltipped-multiline.tooltipped-n::after{right:auto;left:50%;-webkit-transform:translateX(-50%);-ms-transform:translateX(-50%);transform:translateX(-50%);}/*!sc*/
.hWlpPn.tooltipped-multiline.tooltipped-w::after,.hWlpPn.tooltipped-multiline.tooltipped-e::after{right:100%;}/*!sc*/
.hWlpPn.tooltipped-align-right-2::after{right:0;margin-right:0;}/*!sc*/
.hWlpPn.tooltipped-align-left-2::after{left:0;margin-left:0;}/*!sc*/
data-styled.g18[id="Tooltip__TooltipBase-sc-17tf59c-0"]{content:"hWlpPn,"}/*!sc*/
.eAtkQz{display:inline-block;overflow:hidden;text-overflow:ellipsis;vertical-align:top;white-space:nowrap;max-width:125px;max-width:100%;}/*!sc*/
data-styled.g20[id="Truncate__StyledTruncate-sc-23o1d2-0"]{content:"eAtkQz,"}/*!sc*/
.dbGjOi{--segmented-control-button-inner-padding:12px;--segmented-control-button-bg-inset:4px;--segmented-control-outer-radius:6px;background-color:transparent;border-color:transparent;border-radius:var(--segmented-control-outer-radius);border-width:0;color:currentColor;cursor:pointer;font-family:inherit;font-size:inherit;font-weight:600;padding:0;height:100%;width:100%;}/*!sc*/
.dbGjOi:focus:not(:disabled){box-shadow:none;outline:2px solid var(--fgColor-accent,var(--color-accent-fg,#0969da));outline-offset:-1px;}/*!sc*/
.dbGjOi:focus:not(:disabled):not(:focus-visible){outline:solid 1px transparent;}/*!sc*/
.dbGjOi:focus-visible:not(:disabled){box-shadow:none;outline:2px solid var(--fgColor-accent,var(--color-accent-fg,#0969da));outline-offset:-1px;}/*!sc*/
.dbGjOi .segmentedControl-content{-webkit-align-items:center;-webkit-box-align:center;-ms-flex-align:center;align-items:center;background-color:var(--controlKnob-bgColor-rest,var(--color-segmented-control-button-bg,#ffffff));border-color:var(--controlKnob-borderColor-rest,var(--color-segmented-control-button-hover-active-selected-border,#8c959f));border-style:solid;border-width:1px;border-radius:var(--segmented-control-outer-radius);display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;height:100%;-webkit-box-pack:center;-webkit-justify-content:center;-ms-flex-pack:center;justify-content:center;padding-left:var(--segmented-control-button-inner-padding);padding-right:var(--segmented-control-button-inner-padding);}/*!sc*/
.dbGjOi svg{fill:var(--fgColor-muted,var(--color-fg-muted,#656d76));}/*!sc*/
.dbGjOi:focus:focus-visible:not(:last-child):after{width:0;}/*!sc*/
.dbGjOi .segmentedControl-text:after{content:"Code";display:block;font-weight:600;height:0;overflow:hidden;pointer-events:none;-webkit-user-select:none;-moz-user-select:none;-ms-user-select:none;user-select:none;visibility:hidden;}/*!sc*/
@media (pointer:coarse){.dbGjOi:before{content:"";position:absolute;left:0;right:0;-webkit-transform:translateY(-50%);-ms-transform:translateY(-50%);transform:translateY(-50%);top:50%;min-height:44px;}}/*!sc*/
.bHmvop{--segmented-control-button-inner-padding:12px;--segmented-control-button-bg-inset:4px;--segmented-control-outer-radius:6px;background-color:transparent;border-color:transparent;border-radius:var(--segmented-control-outer-radius);border-width:0;color:currentColor;cursor:pointer;font-family:inherit;font-size:inherit;font-weight:400;padding:var(--segmented-control-button-bg-inset);height:100%;width:100%;}/*!sc*/
.bHmvop:focus:not(:disabled){box-shadow:none;outline:2px solid var(--fgColor-accent,var(--color-accent-fg,#0969da));outline-offset:-1px;}/*!sc*/
.bHmvop:focus:not(:disabled):not(:focus-visible){outline:solid 1px transparent;}/*!sc*/
.bHmvop:focus-visible:not(:disabled){box-shadow:none;outline:2px solid var(--fgColor-accent,var(--color-accent-fg,#0969da));outline-offset:-1px;}/*!sc*/
.bHmvop .segmentedControl-content{-webkit-align-items:center;-webkit-box-align:center;-ms-flex-align:center;align-items:center;background-color:transparent;border-color:transparent;border-style:solid;border-width:1px;border-radius:calc(var(--segmented-control-outer-radius) - var(--segmented-control-button-bg-inset) / 2);display:-webkit-box;display:-webkit-flex;display:-ms-flexbox;display:flex;height:100%;-webkit-box-pack:center;-webkit-justify-content:center;-ms-flex-pack:center;justify-content:center;padding-left:calc(var(--segmented-control-button-inner-padding) - var(--segmented-control-button-bg-inset));padding-right:calc(var(--segmented-control-button-inner-padding) - var(--segmented-control-button-bg-inset));}/*!sc*/
.bHmvop svg{fill:var(--fgColor-muted,var(--color-fg-muted,#656d76));}/*!sc*/
.bHmvop:hover .segmentedControl-content{background-color:var(--controlTrack-bgColor-hover,var(--color-segmented-control-button-hover-bg,rgba(175,184,193,0.2)));}/*!sc*/
.bHmvop:active .segmentedControl-content{background-color:var(--controlTrack-bgColor-active,var(--color-segmented-control-button-hover-active-bg,rgba(175,184,193,0.4)));}/*!sc*/
.bHmvop:focus:focus-visible:not(:last-child):after{width:0;}/*!sc*/
.bHmvop .segmentedControl-text:after{content:"Blame";display:block;font-weight:600;height:0;overflow:hidden;pointer-events:none;-webkit-user-select:none;-moz-user-select:none;-ms-user-select:none;user-select:none;visibility:hidden;}/*!sc*/
@media (pointer:coarse){.bHmvop:before{content:"";position:absolute;left:0;right:0;-webkit-transform:translateY(-50%);-ms-transform:translateY(-50%);transform:translateY(-50%);top:50%;min-height:44px;}}/*!sc*/
data-styled.g106[id="SegmentedControlButton__SegmentedControlButtonStyled-sc-8lkgxl-0"]{content:"dbGjOi,bHmvop,"}/*!sc*/
.lawgDG{background-color:var(--controlTrack-bgColor-rest,var(--color-segmented-control-bg,#eaeef2));border-radius:6px;border:1px solid;border-color:var(--controlTrack-borderColor-rest,transparent);display:-webkit-inline-box;display:-webkit-inline-flex;display:-ms-inline-flexbox;display:inline-flex;font-size:14px;height:28px;margin:0;padding:0;}/*!sc*/
data-styled.g108[id="SegmentedControl__SegmentedControlList-sc-1rzig82-0"]{content:"lawgDG,"}/*!sc*/
body[data-page-layout-dragging="true"]{cursor:col-resize;}/*!sc*/
body[data-page-layout-dragging="true"] *{-webkit-user-select:none;-moz-user-select:none;-ms-user-select:none;user-select:none;}/*!sc*/
data-styled.g109[id="sc-global-gbKrvU1"]{content:"sc-global-gbKrvU1,"}/*!sc*/
</style><meta data-hydrostats="publish"/> <!-- --> <!-- --> <button hidden="" data-testid="header-permalink-button" data-hotkey-scope="read-only-cursor-text-area"></button><button hidden=""></button><div><div style="--sticky-pane-height:100vh;--spacing:var(--spacing-none)" class="Box-sc-g0xbh4-0 hOfjFo"><div class="Box-sc-g0xbh4-0 oDGAe"><div class="Box-sc-g0xbh4-0 kowOcT"><div tabindex="0" class="Box-sc-g0xbh4-0 gISSDQ"><div class="Box-sc-g0xbh4-0 cEmWSE"><div class="Box-sc-g0xbh4-0 hPvFuC"></div><div style="--pane-width:320px" class="Box-sc-g0xbh4-0 fFSoPl"></div><div class="Box-sc-g0xbh4-0 bTBnTW"><div role="slider" aria-label="Draggable pane splitter" aria-valuemin="0" aria-valuemax="0" aria-valuenow="0" aria-valuetext="Pane width 0 pixels" tabindex="0" class="Box-sc-g0xbh4-0 bHLmSv"></div></div></div></div><div class="Box-sc-g0xbh4-0 iKqMNA"><div class="Box-sc-g0xbh4-0"></div><div class="Box-sc-g0xbh4-0 FxAyp"><div data-selector="repos-split-pane-content" tabindex="0" class="Box-sc-g0xbh4-0 leYMvG"><div class="Box-sc-g0xbh4-0 KMPzq"><div class="Box-sc-g0xbh4-0 hfKjHv container"><div class="px-3 pt-3 pb-0" id="StickyHeader"><div class="Box-sc-g0xbh4-0 gZWyZE"><div class="Box-sc-g0xbh4-0 dwYKDk"><div class="Box-sc-g0xbh4-0 ibcGmb react-code-view-header-wrap--narrow"><div class="Box-sc-g0xbh4-0 hKaEJF"><h2 class="Box-sc-g0xbh4-0 XosP prc-Heading-Heading-6CmGO"><button style="--button-color:fg.muted" type="button" aria-label="Expand file tree" data-testid="expand-file-tree-button-mobile" class="Box-sc-g0xbh4-0 bCKfWo prc-Button-ButtonBase-c50BI" data-loading="false" data-size="medium" data-variant="invisible" aria-describedby=":Rld9lab:-loading-announcement"><span data-component="buttonContent" class="Box-sc-g0xbh4-0 gUkoLg prc-Button-ButtonContent-HKbr-"><span data-component="leadingVisual" class="prc-Button-Visual-2epfX prc-Button-VisualWrap-Db-eB"><svg aria-hidden="true" focusable="false" class="octicon octicon-arrow-left" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="M7.78 12.53a.75.75 0 0 1-1.06 0L2.47 8.28a.75.75 0 0 1 0-1.06l4.25-4.25a.751.751 0 0 1 1.042.018.751.751 0 0 1 .018 1.042L4.81 7h7.44a.75.75 0 0 1 0 1.5H4.81l2.97 2.97a.75.75 0 0 1 0 1.06Z"></path></svg></span><span data-component="text" class="prc-Button-Label-pTQ3x">Files</span></span></button><span role="tooltip" aria-label="Expand file tree" id="expand-button-file-tree-button" class="Tooltip__TooltipBase-sc-17tf59c-0 hWlpPn tooltipped-se"><button data-component="IconButton" type="button" data-testid="expand-file-tree-button" aria-controls="repos-file-tree" class="prc-Button-ButtonBase-c50BI position-relative ExpandFileTreeButton-module__expandButton--gL4is ExpandFileTreeButton-module__filesButtonBreakpoint--WfX9t fgColor-muted prc-Button-IconButton-szpyj" data-loading="false" data-no-visuals="true" data-size="medium" data-variant="invisible" aria-describedby=":R35d9lab:-loading-announcement" aria-labelledby="expand-button-file-tree-button"><svg aria-hidden="true" focusable="false" class="octicon octicon-sidebar-collapse" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="M6.823 7.823a.25.25 0 0 1 0 .354l-2.396 2.396A.25.25 0 0 1 4 10.396V5.604a.25.25 0 0 1 .427-.177Z"></path><path d="M1.75 0h12.5C15.216 0 16 .784 16 1.75v12.5A1.75 1.75 0 0 1 14.25 16H1.75A1.75 1.75 0 0 1 0 14.25V1.75C0 .784.784 0 1.75 0ZM1.5 1.75v12.5c0 .138.112.25.25.25H9.5v-13H1.75a.25.25 0 0 0-.25.25ZM11 14.5h3.25a.25.25 0 0 0 .25-.25V1.75a.25.25 0 0 0-.25-.25H11Z"></path></svg></button></span><button hidden="" data-testid="" data-hotkey-scope="read-only-cursor-text-area"></button></h2></div><div class="react-code-view-header-mb--narrow mr-2"><button type="button" aria-haspopup="true" aria-expanded="false" tabindex="0" aria-label="master branch" data-testid="anchor-button" class="Box-sc-g0xbh4-0 dmxRgG prc-Button-ButtonBase-c50BI ref-selector-class" data-loading="false" data-size="medium" data-variant="default" aria-describedby="branch-picker-repos-header-ref-selector-wide-loading-announcement" id="branch-picker-repos-header-ref-selector-wide"><span data-component="buttonContent" class="Box-sc-g0xbh4-0 gUkoLg prc-Button-ButtonContent-HKbr-"><span data-component="text" class="prc-Button-Label-pTQ3x"><div class="Box-sc-g0xbh4-0 bZBlpz"><div class="Box-sc-g0xbh4-0 lhTYNA"><svg aria-hidden="true" focusable="false" class="octicon octicon-git-branch" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="M9.5 3.25a2.25 2.25 0 1 1 3 2.122V6A2.5 2.5 0 0 1 10 8.5H6a1 1 0 0 0-1 1v1.128a2.251 2.251 0 1 1-1.5 0V5.372a2.25 2.25 0 1 1 1.5 0v1.836A2.493 2.493 0 0 1 6 7h4a1 1 0 0 0 1-1v-.628A2.25 2.25 0 0 1 9.5 3.25Zm-6 0a.75.75 0 1 0 1.5 0 .75.75 0 0 0-1.5 0Zm8.25-.75a.75.75 0 1 0 0 1.5.75.75 0 0 0 0-1.5ZM4.25 12a.75.75 0 1 0 0 1.5.75.75 0 0 0 0-1.5Z"></path></svg></div><div class="Box-sc-g0xbh4-0 dbrgmi ref-selector-button-text-container"><span class="Text__StyledText-sc-17v1xeu-0 eMMFM"> <!-- -->master</span></div></div></span><span data-component="trailingVisual" class="prc-Button-Visual-2epfX prc-Button-VisualWrap-Db-eB"><svg aria-hidden="true" focusable="false" class="octicon octicon-triangle-down" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="m4.427 7.427 3.396 3.396a.25.25 0 0 0 .354 0l3.396-3.396A.25.25 0 0 0 11.396 7H4.604a.25.25 0 0 0-.177.427Z"></path></svg></span></span></button><button hidden="" data-hotkey-scope="read-only-cursor-text-area"></button></div><div class="Box-sc-g0xbh4-0 dHJiml react-code-view-header-mb--narrow"><div class="Box-sc-g0xbh4-0 cEytCf"><nav data-testid="breadcrumbs" aria-labelledby="repos-header-breadcrumb-heading" id="repos-header-breadcrumb" class="Box-sc-g0xbh4-0 fzFXnm"><h2 class="sr-only ScreenReaderHeading-module__userSelectNone--vW4Cq prc-Heading-Heading-6CmGO" data-testid="screen-reader-heading" id="repos-header-breadcrumb-heading">Breadcrumbs</h2><ol class="Box-sc-g0xbh4-0 iMnkmv"><li class="Box-sc-g0xbh4-0 ghzDag"><a class="Box-sc-g0xbh4-0 kHuKdh prc-Link-Link-85e08" sx="[object Object]" data-testid="breadcrumbs-repo-link" href="/johncsantiago/WhartonLab/tree/master">WhartonLab</a></li></ol></nav><div data-testid="breadcrumbs-filename" class="Box-sc-g0xbh4-0 ghzDag"><span class="Text__StyledText-sc-17v1xeu-0 HlHVj" aria-hidden="true">/</span><h1 class="Box-sc-g0xbh4-0 jGhzSQ prc-Heading-Heading-6CmGO" tabindex="-1" id="file-name-id">G85R GOandKEGGTermEnrichment.Rmd</h1></div><button data-component="IconButton" type="button" class="prc-Button-ButtonBase-c50BI ml-2 prc-Button-IconButton-szpyj" data-loading="false" data-no-visuals="true" data-size="small" data-variant="invisible" aria-describedby=":Rftd9lab:-loading-announcement" aria-labelledby=":R1td9lab:"><svg aria-hidden="true" focusable="false" class="octicon octicon-copy" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="M0 6.75C0 5.784.784 5 1.75 5h1.5a.75.75 0 0 1 0 1.5h-1.5a.25.25 0 0 0-.25.25v7.5c0 .138.112.25.25.25h7.5a.25.25 0 0 0 .25-.25v-1.5a.75.75 0 0 1 1.5 0v1.5A1.75 1.75 0 0 1 9.25 16h-7.5A1.75 1.75 0 0 1 0 14.25Z"></path><path d="M5 1.75C5 .784 5.784 0 6.75 0h7.5C15.216 0 16 .784 16 1.75v7.5A1.75 1.75 0 0 1 14.25 11h-7.5A1.75 1.75 0 0 1 5 9.25Zm1.75-.25a.25.25 0 0 0-.25.25v7.5c0 .138.112.25.25.25h7.5a.25.25 0 0 0 .25-.25v-7.5a.25.25 0 0 0-.25-.25Z"></path></svg></button><span class="Tooltip__StyledTooltip-sc-e45c7z-0 jOyaRH CopyToClipboardButton-module__tooltip--Dq1IB" data-direction="nw" aria-label="Copy path" aria-hidden="true" id=":R1td9lab:">Copy path</span></div></div></div><div class="react-code-view-header-element--wide"><div class="Box-sc-g0xbh4-0 faNtbn"><div class="d-flex gap-2"> <div><div class="Box-sc-g0xbh4-0 fmQaBv"><span class="TextInputWrapper__StyledTextInputBaseWrapper-sc-1mqhpbi-0 hLzFvi TextInputWrapper__StyledTextInputWrapper-sc-1mqhpbi-1 iHYdQq TextInput-wrapper" data-leading-visual="true" data-trailing-visual="true" aria-busy="false"><span class="TextInput-icon" id=":R5j6d9lab:" aria-hidden="true"><svg aria-hidden="true" focusable="false" class="octicon octicon-search" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="M10.68 11.74a6 6 0 0 1-7.922-8.982 6 6 0 0 1 8.982 7.922l3.04 3.04a.749.749 0 0 1-.326 1.275.749.749 0 0 1-.734-.215ZM11.5 7a4.499 4.499 0 1 0-8.997 0A4.499 4.499 0 0 0 11.5 7Z"></path></svg></span><input type="text" aria-label="Go to file" role="combobox" aria-controls="file-results-list" aria-expanded="false" aria-haspopup="dialog" autoCorrect="off" spellcheck="false" placeholder="Go to file" aria-describedby=":R5j6d9lab: :R5j6d9labH1:" data-component="input" class="UnstyledTextInput__ToggledUnstyledTextInput-sc-14ypya-0 jkNcAv" value=""/><span class="TextInput-icon" id=":R5j6d9labH1:" aria-hidden="true"></span></span></div><button hidden="" data-testid="" data-hotkey-scope="read-only-cursor-text-area"></button><button hidden=""></button></div><button hidden="" data-testid="" data-hotkey-scope="read-only-cursor-text-area"></button><button hidden=""></button><button hidden="" data-testid="" data-hotkey-scope="read-only-cursor-text-area"></button><button hidden=""></button><button type="button" class="Box-sc-g0xbh4-0 dwNhzn prc-Button-ButtonBase-c50BI" data-loading="false" data-no-visuals="true" data-size="medium" data-variant="default" aria-describedby=":R2l6d9lab:-loading-announcement"><span data-component="buttonContent" class="Box-sc-g0xbh4-0 gUkoLg prc-Button-ButtonContent-HKbr-"><span data-component="text" class="prc-Button-Label-pTQ3x">Blame</span></span></button><button hidden="" data-testid="" data-hotkey-scope="read-only-cursor-text-area"></button><button data-component="IconButton" type="button" aria-label="More file actions" title="More file actions" data-testid="more-file-actions-button-nav-menu-wide" aria-haspopup="true" aria-expanded="false" tabindex="0" class="Box-sc-g0xbh4-0 fGwBZA prc-Button-ButtonBase-c50BI js-blob-dropdown-click prc-Button-IconButton-szpyj" data-loading="false" data-no-visuals="true" data-size="medium" data-variant="default" aria-describedby=":R156d9lab:-loading-announcement" id=":R156d9lab:"><svg aria-hidden="true" focusable="false" class="octicon octicon-kebab-horizontal" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="M8 9a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3ZM1.5 9a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3Zm13 0a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3Z"></path></svg></button> </div></div></div><div class="react-code-view-header-element--narrow"><div class="Box-sc-g0xbh4-0 faNtbn"><div class="d-flex gap-2"> <button hidden="" data-testid="" data-hotkey-scope="read-only-cursor-text-area"></button><button hidden=""></button><button hidden="" data-testid="" data-hotkey-scope="read-only-cursor-text-area"></button><button hidden=""></button><button type="button" class="Box-sc-g0xbh4-0 dwNhzn prc-Button-ButtonBase-c50BI" data-loading="false" data-no-visuals="true" data-size="medium" data-variant="default" aria-describedby=":R2l7d9lab:-loading-announcement"><span data-component="buttonContent" class="Box-sc-g0xbh4-0 gUkoLg prc-Button-ButtonContent-HKbr-"><span data-component="text" class="prc-Button-Label-pTQ3x">Blame</span></span></button><button hidden="" data-testid="" data-hotkey-scope="read-only-cursor-text-area"></button><button data-component="IconButton" type="button" aria-label="More file actions" title="More file actions" data-testid="more-file-actions-button-nav-menu-narrow" aria-haspopup="true" aria-expanded="false" tabindex="0" class="Box-sc-g0xbh4-0 fGwBZA prc-Button-ButtonBase-c50BI js-blob-dropdown-click prc-Button-IconButton-szpyj" data-loading="false" data-no-visuals="true" data-size="medium" data-variant="default" aria-describedby=":R157d9lab:-loading-announcement" id=":R157d9lab:"><svg aria-hidden="true" focusable="false" class="octicon octicon-kebab-horizontal" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="M8 9a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3ZM1.5 9a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3Zm13 0a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3Z"></path></svg></button> </div></div></div></div></div></div></div></div><div class="Box-sc-g0xbh4-0 dJxjrT react-code-view-bottom-padding"> <div class="Box-sc-g0xbh4-0 eFxKDQ"></div> <!-- --> <!-- --> </div><div class="Box-sc-g0xbh4-0 dJxjrT"> <!-- --> <!-- --> <div class="d-flex flex-column border rounded-2 mb-3 pl-1"><div class="Box-sc-g0xbh4-0 dzCJzi"><h2 class="sr-only ScreenReaderHeading-module__userSelectNone--vW4Cq prc-Heading-Heading-6CmGO" data-testid="screen-reader-heading">Latest commit</h2><div style="width:120px" class="Skeleton Skeleton--text" data-testid="loading"> </div><div class="d-flex flex-shrink-0 gap-2"><div data-testid="latest-commit-details" class="d-none d-sm-flex flex-items-center"></div><div class="d-flex gap-2"><h2 class="sr-only ScreenReaderHeading-module__userSelectNone--vW4Cq prc-Heading-Heading-6CmGO" data-testid="screen-reader-heading">History</h2><a href="/johncsantiago/WhartonLab/commits/master/G85R%20GOandKEGGTermEnrichment.Rmd" class="prc-Button-ButtonBase-c50BI d-none d-lg-flex LinkButton-module__code-view-link-button--xvCGA flex-items-center fgColor-default" data-loading="false" data-size="small" data-variant="invisible" aria-describedby=":R5dlal9lab:-loading-announcement"><span data-component="buttonContent" data-align="center" class="prc-Button-ButtonContent-HKbr-"><span data-component="leadingVisual" class="prc-Button-Visual-2epfX prc-Button-VisualWrap-Db-eB"><svg aria-hidden="true" focusable="false" class="octicon octicon-history" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="m.427 1.927 1.215 1.215a8.002 8.002 0 1 1-1.6 5.685.75.75 0 1 1 1.493-.154 6.5 6.5 0 1 0 1.18-4.458l1.358 1.358A.25.25 0 0 1 3.896 6H.25A.25.25 0 0 1 0 5.75V2.104a.25.25 0 0 1 .427-.177ZM7.75 4a.75.75 0 0 1 .75.75v2.992l2.028.812a.75.75 0 0 1-.557 1.392l-2.5-1A.751.751 0 0 1 7 8.25v-3.5A.75.75 0 0 1 7.75 4Z"></path></svg></span><span data-component="text" class="prc-Button-Label-pTQ3x"><span class="fgColor-default">History</span></span></span></a><div class="d-sm-none"></div><div class="d-flex d-lg-none"><span role="tooltip" aria-label="History" id="history-icon-button-tooltip" class="Tooltip__TooltipBase-sc-17tf59c-0 hWlpPn tooltipped-n"><a href="/johncsantiago/WhartonLab/commits/master/G85R%20GOandKEGGTermEnrichment.Rmd" class="prc-Button-ButtonBase-c50BI LinkButton-module__code-view-link-button--xvCGA flex-items-center fgColor-default" data-loading="false" data-size="small" data-variant="invisible" aria-describedby=":Rpdlal9lab:-loading-announcement history-icon-button-tooltip"><span data-component="buttonContent" data-align="center" class="prc-Button-ButtonContent-HKbr-"><span data-component="leadingVisual" class="prc-Button-Visual-2epfX prc-Button-VisualWrap-Db-eB"><svg aria-hidden="true" focusable="false" class="octicon octicon-history" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="m.427 1.927 1.215 1.215a8.002 8.002 0 1 1-1.6 5.685.75.75 0 1 1 1.493-.154 6.5 6.5 0 1 0 1.18-4.458l1.358 1.358A.25.25 0 0 1 3.896 6H.25A.25.25 0 0 1 0 5.75V2.104a.25.25 0 0 1 .427-.177ZM7.75 4a.75.75 0 0 1 .75.75v2.992l2.028.812a.75.75 0 0 1-.557 1.392l-2.5-1A.751.751 0 0 1 7 8.25v-3.5A.75.75 0 0 1 7.75 4Z"></path></svg></span></span></a></span></div></div></div></div></div><div class="Box-sc-g0xbh4-0 ldRxiI"><div class="Box-sc-g0xbh4-0 fVkfyA container"><div class="Box-sc-g0xbh4-0 gNAmSV react-code-size-details-banner"><div class="Box-sc-g0xbh4-0 jNEwzY react-code-size-details-banner"><div class="Box-sc-g0xbh4-0 bsDwxw text-mono"><div title="3.31 KB" data-testid="blob-size" class="Truncate__StyledTruncate-sc-23o1d2-0 eAtkQz"><span>219 lines (116 loc) · 3.31 KB</span></div></div></div></div><div class="Box-sc-g0xbh4-0 jdLMhu react-blob-view-header-sticky" id="repos-sticky-header"><div class="Box-sc-g0xbh4-0 tOISc"><div class="react-blob-sticky-header"><div class="Box-sc-g0xbh4-0 hqwSEx"><div class="Box-sc-g0xbh4-0 bDVoEr"><div class="Box-sc-g0xbh4-0 kYLlPM"><div class="Box-sc-g0xbh4-0 gYjEmn"><button type="button" aria-haspopup="true" aria-expanded="false" tabindex="0" aria-label="master branch" data-testid="anchor-button" class="Box-sc-g0xbh4-0 dmxRgG prc-Button-ButtonBase-c50BI ref-selector-class" data-loading="false" data-size="medium" data-variant="default" aria-describedby="branch-picker-repos-header-ref-selector-loading-announcement" id="branch-picker-repos-header-ref-selector"><span data-component="buttonContent" class="Box-sc-g0xbh4-0 gUkoLg prc-Button-ButtonContent-HKbr-"><span data-component="text" class="prc-Button-Label-pTQ3x"><div class="Box-sc-g0xbh4-0 bZBlpz"><div class="Box-sc-g0xbh4-0 lhTYNA"><svg aria-hidden="true" focusable="false" class="octicon octicon-git-branch" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="M9.5 3.25a2.25 2.25 0 1 1 3 2.122V6A2.5 2.5 0 0 1 10 8.5H6a1 1 0 0 0-1 1v1.128a2.251 2.251 0 1 1-1.5 0V5.372a2.25 2.25 0 1 1 1.5 0v1.836A2.493 2.493 0 0 1 6 7h4a1 1 0 0 0 1-1v-.628A2.25 2.25 0 0 1 9.5 3.25Zm-6 0a.75.75 0 1 0 1.5 0 .75.75 0 0 0-1.5 0Zm8.25-.75a.75.75 0 1 0 0 1.5.75.75 0 0 0 0-1.5ZM4.25 12a.75.75 0 1 0 0 1.5.75.75 0 0 0 0-1.5Z"></path></svg></div><div class="Box-sc-g0xbh4-0 dbrgmi ref-selector-button-text-container"><span class="Text__StyledText-sc-17v1xeu-0 eMMFM"> <!-- -->master</span></div></div></span><span data-component="trailingVisual" class="prc-Button-Visual-2epfX prc-Button-VisualWrap-Db-eB"><svg aria-hidden="true" focusable="false" class="octicon octicon-triangle-down" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="m4.427 7.427 3.396 3.396a.25.25 0 0 0 .354 0l3.396-3.396A.25.25 0 0 0 11.396 7H4.604a.25.25 0 0 0-.177.427Z"></path></svg></span></span></button><button hidden="" data-hotkey-scope="read-only-cursor-text-area"></button></div><div class="Box-sc-g0xbh4-0 kGqOLL"><div class="Box-sc-g0xbh4-0 fHind"><nav data-testid="breadcrumbs" aria-labelledby="sticky-breadcrumb-heading" id="sticky-breadcrumb" class="Box-sc-g0xbh4-0 fzFXnm"><h2 class="sr-only ScreenReaderHeading-module__userSelectNone--vW4Cq prc-Heading-Heading-6CmGO" data-testid="screen-reader-heading" id="sticky-breadcrumb-heading">Breadcrumbs</h2><ol class="Box-sc-g0xbh4-0 iMnkmv"><li class="Box-sc-g0xbh4-0 ghzDag"><a class="Box-sc-g0xbh4-0 kHuKdh prc-Link-Link-85e08" sx="[object Object]" data-testid="breadcrumbs-repo-link" href="/johncsantiago/WhartonLab/tree/master">WhartonLab</a></li></ol></nav><div data-testid="breadcrumbs-filename" class="Box-sc-g0xbh4-0 ghzDag"><span class="Text__StyledText-sc-17v1xeu-0 lauzFl" aria-hidden="true">/</span><h1 class="Box-sc-g0xbh4-0 dnZoUW prc-Heading-Heading-6CmGO" tabindex="-1" id="sticky-file-name-id">G85R GOandKEGGTermEnrichment.Rmd</h1></div></div></div></div><button style="--button-color:fg.default" type="button" class="Box-sc-g0xbh4-0 jRZWlf prc-Button-ButtonBase-c50BI" data-loading="false" data-size="small" data-variant="invisible" aria-describedby=":Riptal9lab:-loading-announcement"><span data-component="buttonContent" class="Box-sc-g0xbh4-0 gUkoLg prc-Button-ButtonContent-HKbr-"><span data-component="leadingVisual" class="prc-Button-Visual-2epfX prc-Button-VisualWrap-Db-eB"><svg aria-hidden="true" focusable="false" class="octicon octicon-arrow-up" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="M3.47 7.78a.75.75 0 0 1 0-1.06l4.25-4.25a.75.75 0 0 1 1.06 0l4.25 4.25a.751.751 0 0 1-.018 1.042.751.751 0 0 1-1.042.018L9 4.81v7.44a.75.75 0 0 1-1.5 0V4.81L4.53 7.78a.75.75 0 0 1-1.06 0Z"></path></svg></span><span data-component="text" class="prc-Button-Label-pTQ3x">Top</span></span></button></div></div></div><div class="Box-sc-g0xbh4-0 kTvpNk"><h2 class="sr-only ScreenReaderHeading-module__userSelectNone--vW4Cq prc-Heading-Heading-6CmGO" data-testid="screen-reader-heading">File metadata and controls</h2><div class="Box-sc-g0xbh4-0 iNMjfP"><ul aria-label="File view" class="SegmentedControl__SegmentedControlList-sc-1rzig82-0 lawgDG" data-size="small"><li class="Box-sc-g0xbh4-0 fefCSX" data-selected="true"><button aria-current="true" class="SegmentedControlButton__SegmentedControlButtonStyled-sc-8lkgxl-0 dbGjOi" type="button"><span class="segmentedControl-content"><div class="Box-sc-g0xbh4-0 segmentedControl-text" data-text="Code">Code</div></span></button></li><li class="Box-sc-g0xbh4-0 idgUkN"><button aria-current="false" class="SegmentedControlButton__SegmentedControlButtonStyled-sc-8lkgxl-0 bHmvop" type="button"><span class="segmentedControl-content"><div class="Box-sc-g0xbh4-0 segmentedControl-text" data-text="Blame">Blame</div></span></button></li></ul><button hidden="" data-testid="" data-hotkey-scope="read-only-cursor-text-area"></button><button hidden="" data-testid="" data-hotkey-scope="read-only-cursor-text-area"></button><div class="Box-sc-g0xbh4-0 jNEwzY react-code-size-details-in-header"><div class="Box-sc-g0xbh4-0 bsDwxw text-mono"><div title="3.31 KB" data-testid="blob-size" class="Truncate__StyledTruncate-sc-23o1d2-0 eAtkQz"><span>219 lines (116 loc) · 3.31 KB</span></div></div></div></div><div class="Box-sc-g0xbh4-0 kcLCKF"><div class="Box-sc-g0xbh4-0 kVWtTz react-blob-header-edit-and-raw-actions"><div class="Box-sc-g0xbh4-0 prc-ButtonGroup-ButtonGroup-vcMeG"><div><a href="https://github.com/johncsantiago/WhartonLab/raw/refs/heads/master/G85R%20GOandKEGGTermEnrichment.Rmd" data-testid="raw-button" class="Box-sc-g0xbh4-0 gWqxTd prc-Button-ButtonBase-c50BI" data-loading="false" data-no-visuals="true" data-size="small" data-variant="default" aria-describedby=":R5csptal9lab:-loading-announcement"><span data-component="buttonContent" class="Box-sc-g0xbh4-0 gUkoLg prc-Button-ButtonContent-HKbr-"><span data-component="text" class="prc-Button-Label-pTQ3x">Raw</span></span></a></div><div><button data-component="IconButton" type="button" aria-label="Copy raw content" data-testid="copy-raw-button" class="prc-Button-ButtonBase-c50BI prc-Button-IconButton-szpyj" data-loading="false" data-no-visuals="true" data-size="small" data-variant="default" aria-describedby=":Rpcsptal9lab:-loading-announcement"><svg aria-hidden="true" focusable="false" class="octicon octicon-copy" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="M0 6.75C0 5.784.784 5 1.75 5h1.5a.75.75 0 0 1 0 1.5h-1.5a.25.25 0 0 0-.25.25v7.5c0 .138.112.25.25.25h7.5a.25.25 0 0 0 .25-.25v-1.5a.75.75 0 0 1 1.5 0v1.5A1.75 1.75 0 0 1 9.25 16h-7.5A1.75 1.75 0 0 1 0 14.25Z"></path><path d="M5 1.75C5 .784 5.784 0 6.75 0h7.5C15.216 0 16 .784 16 1.75v7.5A1.75 1.75 0 0 1 14.25 11h-7.5A1.75 1.75 0 0 1 5 9.25Zm1.75-.25a.25.25 0 0 0-.25.25v7.5c0 .138.112.25.25.25h7.5a.25.25 0 0 0 .25-.25v-7.5a.25.25 0 0 0-.25-.25Z"></path></svg></button></div><div><span role="tooltip" aria-label="Download raw file" id=":Rdcsptal9lab:" class="Tooltip__TooltipBase-sc-17tf59c-0 hWlpPn tooltipped-n"><button data-component="IconButton" type="button" aria-label="Download raw content" data-testid="download-raw-button" class="Box-sc-g0xbh4-0 ivobqY prc-Button-ButtonBase-c50BI prc-Button-IconButton-szpyj" data-loading="false" data-no-visuals="true" data-size="small" data-variant="default" aria-describedby=":Rtcsptal9lab:-loading-announcement"><svg aria-hidden="true" focusable="false" class="octicon octicon-download" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="M2.75 14A1.75 1.75 0 0 1 1 12.25v-2.5a.75.75 0 0 1 1.5 0v2.5c0 .138.112.25.25.25h10.5a.25.25 0 0 0 .25-.25v-2.5a.75.75 0 0 1 1.5 0v2.5A1.75 1.75 0 0 1 13.25 14Z"></path><path d="M7.25 7.689V2a.75.75 0 0 1 1.5 0v5.689l1.97-1.969a.749.749 0 1 1 1.06 1.06l-3.25 3.25a.749.749 0 0 1-1.06 0L4.22 6.78a.749.749 0 1 1 1.06-1.06l1.97 1.969Z"></path></svg></button></span></div></div><button hidden="" data-testid="raw-button-shortcut" data-hotkey-scope="read-only-cursor-text-area"></button><button hidden="" data-testid="copy-raw-button-shortcut" data-hotkey-scope="read-only-cursor-text-area"></button><button hidden="" data-testid="download-raw-button-shortcut" data-hotkey-scope="read-only-cursor-text-area"></button></div><span role="tooltip" aria-label="Open symbols panel" id=":R5sptal9lab:" class="Tooltip__TooltipBase-sc-17tf59c-0 hWlpPn tooltipped-nw"><button data-component="IconButton" type="button" aria-label="Symbols" aria-pressed="false" aria-expanded="false" aria-controls="symbols-pane" data-testid="symbols-button" class="Box-sc-g0xbh4-0 heuRGy prc-Button-ButtonBase-c50BI prc-Button-IconButton-szpyj" data-loading="false" data-no-visuals="true" data-size="small" data-variant="invisible" aria-describedby="symbols-button-loading-announcement" id="symbols-button"><svg aria-hidden="true" focusable="false" class="octicon octicon-code-square" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="M0 1.75C0 .784.784 0 1.75 0h12.5C15.216 0 16 .784 16 1.75v12.5A1.75 1.75 0 0 1 14.25 16H1.75A1.75 1.75 0 0 1 0 14.25Zm1.75-.25a.25.25 0 0 0-.25.25v12.5c0 .138.112.25.25.25h12.5a.25.25 0 0 0 .25-.25V1.75a.25.25 0 0 0-.25-.25Zm7.47 3.97a.75.75 0 0 1 1.06 0l2 2a.75.75 0 0 1 0 1.06l-2 2a.749.749 0 0 1-1.275-.326.749.749 0 0 1 .215-.734L10.69 8 9.22 6.53a.75.75 0 0 1 0-1.06ZM6.78 6.53 5.31 8l1.47 1.47a.749.749 0 0 1-.326 1.275.749.749 0 0 1-.734-.215l-2-2a.75.75 0 0 1 0-1.06l2-2a.751.751 0 0 1 1.042.018.751.751 0 0 1 .018 1.042Z"></path></svg></button></span><div class="react-blob-header-edit-and-raw-actions-combined"><button data-component="IconButton" type="button" aria-label="Edit and raw actions" title="More file actions" data-testid="more-file-actions-button" aria-haspopup="true" aria-expanded="false" tabindex="0" class="Box-sc-g0xbh4-0 ffkqe prc-Button-ButtonBase-c50BI js-blob-dropdown-click prc-Button-IconButton-szpyj" data-loading="false" data-no-visuals="true" data-size="small" data-variant="invisible" aria-describedby=":Rnsptal9lab:-loading-announcement" id=":Rnsptal9lab:"><svg aria-hidden="true" focusable="false" class="octicon octicon-kebab-horizontal" viewBox="0 0 16 16" width="16" height="16" fill="currentColor" style="display:inline-block;user-select:none;vertical-align:text-bottom;overflow:visible"><path d="M8 9a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3ZM1.5 9a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3Zm13 0a1.5 1.5 0 1 0 0-3 1.5 1.5 0 0 0 0 3Z"></path></svg></button></div></div></div></div><div></div></div><div class="Box-sc-g0xbh4-0 hGyMdv"><section aria-labelledby="file-name-id-wide file-name-id-mobile" class="Box-sc-g0xbh4-0 dceWRL"><div class="Box-sc-g0xbh4-0 dGXHv"><div id="highlighted-line-menu-positioner" class="position-relative"><div id="copilot-button-positioner" class="Box-sc-g0xbh4-0 bpDFns"><div class="Box-sc-g0xbh4-0 iJOeCH"><div class="Box-sc-g0xbh4-0 jewUnv react-code-file-contents" role="presentation" aria-hidden="true" data-tab-size="8" data-paste-markdown-skip="true" data-hpc="true"><div class="react-line-numbers" style="pointer-events:auto"><div data-line-number="1" class="react-line-number react-code-text" style="padding-right:16px">1</div><div data-line-number="2" class="react-line-number react-code-text" style="padding-right:16px">2</div><div data-line-number="3" class="react-line-number react-code-text" style="padding-right:16px">3</div><div data-line-number="4" class="react-line-number react-code-text" style="padding-right:16px">4</div><div data-line-number="5" class="react-line-number react-code-text" style="padding-right:16px">5</div><div data-line-number="6" class="react-line-number react-code-text" style="padding-right:16px">6</div><div data-line-number="7" class="react-line-number react-code-text" style="padding-right:16px">7</div><div data-line-number="8" class="react-line-number react-code-text" style="padding-right:16px">8</div><div data-line-number="9" class="react-line-number react-code-text" style="padding-right:16px">9</div><div data-line-number="10" class="react-line-number react-code-text" style="padding-right:16px">10</div><div data-line-number="11" class="react-line-number react-code-text" style="padding-right:16px">11</div><div data-line-number="12" class="react-line-number react-code-text" style="padding-right:16px">12</div><div data-line-number="13" class="react-line-number react-code-text" style="padding-right:16px">13</div><div data-line-number="14" class="react-line-number react-code-text" style="padding-right:16px">14</div><div data-line-number="15" class="react-line-number react-code-text" style="padding-right:16px">15</div><div data-line-number="16" class="react-line-number react-code-text" style="padding-right:16px">16</div><div data-line-number="17" class="react-line-number react-code-text" style="padding-right:16px">17</div><div data-line-number="18" class="react-line-number react-code-text" style="padding-right:16px">18</div><div data-line-number="19" class="react-line-number react-code-text" style="padding-right:16px">19</div><div data-line-number="20" class="react-line-number react-code-text" style="padding-right:16px">20</div><div data-line-number="21" class="react-line-number react-code-text" style="padding-right:16px">21</div><div data-line-number="22" class="react-line-number react-code-text" style="padding-right:16px">22</div><div data-line-number="23" class="react-line-number react-code-text" style="padding-right:16px">23</div><div data-line-number="24" class="react-line-number react-code-text" style="padding-right:16px">24</div><div data-line-number="25" class="react-line-number react-code-text" style="padding-right:16px">25</div><div data-line-number="26" class="react-line-number react-code-text" style="padding-right:16px">26</div><div data-line-number="27" class="react-line-number react-code-text" style="padding-right:16px">27</div><div data-line-number="28" class="react-line-number react-code-text" style="padding-right:16px">28</div><div data-line-number="29" class="react-line-number react-code-text" style="padding-right:16px">29</div><div data-line-number="30" class="react-line-number react-code-text" style="padding-right:16px">30</div><div data-line-number="31" class="react-line-number react-code-text" style="padding-right:16px">31</div><div data-line-number="32" class="react-line-number react-code-text" style="padding-right:16px">32</div><div data-line-number="33" class="react-line-number react-code-text" style="padding-right:16px">33</div><div data-line-number="34" class="react-line-number react-code-text" style="padding-right:16px">34</div><div data-line-number="35" class="react-line-number react-code-text" style="padding-right:16px">35</div><div data-line-number="36" class="react-line-number react-code-text" style="padding-right:16px">36</div><div data-line-number="37" class="react-line-number react-code-text" style="padding-right:16px">37</div><div data-line-number="38" class="react-line-number react-code-text" style="padding-right:16px">38</div><div data-line-number="39" class="react-line-number react-code-text" style="padding-right:16px">39</div><div data-line-number="40" class="react-line-number react-code-text" style="padding-right:16px">40</div><div data-line-number="41" class="react-line-number react-code-text" style="padding-right:16px">41</div><div data-line-number="42" class="react-line-number react-code-text" style="padding-right:16px">42</div><div data-line-number="43" class="react-line-number react-code-text" style="padding-right:16px">43</div><div data-line-number="44" class="react-line-number react-code-text" style="padding-right:16px">44</div><div data-line-number="45" class="react-line-number react-code-text" style="padding-right:16px">45</div><div data-line-number="46" class="react-line-number react-code-text" style="padding-right:16px">46</div><div data-line-number="47" class="react-line-number react-code-text" style="padding-right:16px">47</div><div data-line-number="48" class="react-line-number react-code-text" style="padding-right:16px">48</div><div data-line-number="49" class="react-line-number react-code-text" style="padding-right:16px">49</div><div data-line-number="50" class="react-line-number react-code-text" style="padding-right:16px">50</div><div data-line-number="51" class="react-line-number react-code-text" style="padding-right:16px">51</div><div data-line-number="52" class="react-line-number react-code-text" style="padding-right:16px">52</div><div data-line-number="53" class="react-line-number react-code-text" style="padding-right:16px">53</div><div data-line-number="54" class="react-line-number react-code-text" style="padding-right:16px">54</div><div data-line-number="55" class="react-line-number react-code-text" style="padding-right:16px">55</div><div data-line-number="56" class="react-line-number react-code-text" style="padding-right:16px">56</div><div data-line-number="57" class="react-line-number react-code-text" style="padding-right:16px">57</div><div data-line-number="58" class="react-line-number react-code-text" style="padding-right:16px">58</div><div data-line-number="59" class="react-line-number react-code-text" style="padding-right:16px">59</div><div data-line-number="60" class="react-line-number react-code-text" style="padding-right:16px">60</div><div data-line-number="61" class="react-line-number react-code-text" style="padding-right:16px">61</div><div data-line-number="62" class="react-line-number react-code-text" style="padding-right:16px">62</div><div data-line-number="63" class="react-line-number react-code-text" style="padding-right:16px">63</div><div data-line-number="64" class="react-line-number react-code-text" style="padding-right:16px">64</div><div data-line-number="65" class="react-line-number react-code-text" style="padding-right:16px">65</div><div data-line-number="66" class="react-line-number react-code-text" style="padding-right:16px">66</div><div data-line-number="67" class="react-line-number react-code-text" style="padding-right:16px">67</div><div data-line-number="68" class="react-line-number react-code-text" style="padding-right:16px">68</div><div data-line-number="69" class="react-line-number react-code-text" style="padding-right:16px">69</div><div data-line-number="70" class="react-line-number react-code-text" style="padding-right:16px">70</div><div data-line-number="71" class="react-line-number react-code-text" style="padding-right:16px">71</div><div data-line-number="72" class="react-line-number react-code-text" style="padding-right:16px">72</div><div data-line-number="73" class="react-line-number react-code-text" style="padding-right:16px">73</div><div data-line-number="74" class="react-line-number react-code-text" style="padding-right:16px">74</div><div data-line-number="75" class="react-line-number react-code-text" style="padding-right:16px">75</div><div data-line-number="76" class="react-line-number react-code-text" style="padding-right:16px">76</div><div data-line-number="77" class="react-line-number react-code-text" style="padding-right:16px">77</div><div data-line-number="78" class="react-line-number react-code-text" style="padding-right:16px">78</div><div data-line-number="79" class="react-line-number react-code-text" style="padding-right:16px">79</div><div data-line-number="80" class="react-line-number react-code-text" style="padding-right:16px">80</div><div data-line-number="81" class="react-line-number react-code-text" style="padding-right:16px">81</div><div data-line-number="82" class="react-line-number react-code-text" style="padding-right:16px">82</div><div data-line-number="83" class="react-line-number react-code-text" style="padding-right:16px">83</div><div data-line-number="84" class="react-line-number react-code-text" style="padding-right:16px">84</div><div data-line-number="85" class="react-line-number react-code-text" style="padding-right:16px">85</div><div data-line-number="86" class="react-line-number react-code-text" style="padding-right:16px">86</div><div data-line-number="87" class="react-line-number react-code-text" style="padding-right:16px">87</div><div data-line-number="88" class="react-line-number react-code-text" style="padding-right:16px">88</div><div data-line-number="89" class="react-line-number react-code-text" style="padding-right:16px">89</div><div data-line-number="90" class="react-line-number react-code-text" style="padding-right:16px">90</div><div data-line-number="91" class="react-line-number react-code-text" style="padding-right:16px">91</div><div data-line-number="92" class="react-line-number react-code-text" style="padding-right:16px">92</div><div data-line-number="93" class="react-line-number react-code-text" style="padding-right:16px">93</div><div data-line-number="94" class="react-line-number react-code-text" style="padding-right:16px">94</div><div data-line-number="95" class="react-line-number react-code-text" style="padding-right:16px">95</div><div data-line-number="96" class="react-line-number react-code-text" style="padding-right:16px">96</div><div data-line-number="97" class="react-line-number react-code-text" style="padding-right:16px">97</div><div data-line-number="98" class="react-line-number react-code-text" style="padding-right:16px">98</div><div data-line-number="99" class="react-line-number react-code-text" style="padding-right:16px">99</div><div data-line-number="100" class="react-line-number react-code-text" style="padding-right:16px">100</div><div data-line-number="101" class="react-line-number react-code-text" style="padding-right:16px">101</div><div data-line-number="102" class="react-line-number react-code-text" style="padding-right:16px">102</div><div data-line-number="103" class="react-line-number react-code-text" style="padding-right:16px">103</div><div data-line-number="104" class="react-line-number react-code-text" style="padding-right:16px">104</div><div data-line-number="105" class="react-line-number react-code-text" style="padding-right:16px">105</div><div data-line-number="106" class="react-line-number react-code-text" style="padding-right:16px">106</div><div data-line-number="107" class="react-line-number react-code-text" style="padding-right:16px">107</div><div data-line-number="108" class="react-line-number react-code-text" style="padding-right:16px">108</div><div data-line-number="109" class="react-line-number react-code-text" style="padding-right:16px">109</div><div data-line-number="110" class="react-line-number react-code-text" style="padding-right:16px">110</div><div data-line-number="111" class="react-line-number react-code-text" style="padding-right:16px">111</div><div data-line-number="112" class="react-line-number react-code-text" style="padding-right:16px">112</div><div data-line-number="113" class="react-line-number react-code-text" style="padding-right:16px">113</div><div data-line-number="114" class="react-line-number react-code-text" style="padding-right:16px">114</div><div data-line-number="115" class="react-line-number react-code-text" style="padding-right:16px">115</div><div data-line-number="116" class="react-line-number react-code-text" style="padding-right:16px">116</div><div data-line-number="117" class="react-line-number react-code-text" style="padding-right:16px">117</div><div data-line-number="118" class="react-line-number react-code-text" style="padding-right:16px">118</div><div data-line-number="119" class="react-line-number react-code-text" style="padding-right:16px">119</div><div data-line-number="120" class="react-line-number react-code-text" style="padding-right:16px">120</div><div data-line-number="121" class="react-line-number react-code-text" style="padding-right:16px">121</div><div data-line-number="122" class="react-line-number react-code-text" style="padding-right:16px">122</div><div data-line-number="123" class="react-line-number react-code-text" style="padding-right:16px">123</div><div data-line-number="124" class="react-line-number react-code-text" style="padding-right:16px">124</div><div data-line-number="125" class="react-line-number react-code-text" style="padding-right:16px">125</div><div data-line-number="126" class="react-line-number react-code-text" style="padding-right:16px">126</div><div data-line-number="127" class="react-line-number react-code-text" style="padding-right:16px">127</div><div data-line-number="128" class="react-line-number react-code-text" style="padding-right:16px">128</div><div data-line-number="129" class="react-line-number react-code-text" style="padding-right:16px">129</div><div data-line-number="130" class="react-line-number react-code-text" style="padding-right:16px">130</div><div data-line-number="131" class="react-line-number react-code-text" style="padding-right:16px">131</div><div data-line-number="132" class="react-line-number react-code-text" style="padding-right:16px">132</div><div data-line-number="133" class="react-line-number react-code-text" style="padding-right:16px">133</div><div data-line-number="134" class="react-line-number react-code-text" style="padding-right:16px">134</div><div data-line-number="135" class="react-line-number react-code-text" style="padding-right:16px">135</div><div data-line-number="136" class="react-line-number react-code-text" style="padding-right:16px">136</div><div data-line-number="137" class="react-line-number react-code-text" style="padding-right:16px">137</div><div data-line-number="138" class="react-line-number react-code-text" style="padding-right:16px">138</div><div data-line-number="139" class="react-line-number react-code-text" style="padding-right:16px">139</div><div data-line-number="140" class="react-line-number react-code-text" style="padding-right:16px">140</div><div data-line-number="141" class="react-line-number react-code-text" style="padding-right:16px">141</div><div data-line-number="142" class="react-line-number react-code-text" style="padding-right:16px">142</div><div data-line-number="143" class="react-line-number react-code-text" style="padding-right:16px">143</div><div data-line-number="144" class="react-line-number react-code-text" style="padding-right:16px">144</div><div data-line-number="145" class="react-line-number react-code-text" style="padding-right:16px">145</div><div data-line-number="146" class="react-line-number react-code-text" style="padding-right:16px">146</div><div data-line-number="147" class="react-line-number react-code-text" style="padding-right:16px">147</div><div data-line-number="148" class="react-line-number react-code-text" style="padding-right:16px">148</div><div data-line-number="149" class="react-line-number react-code-text" style="padding-right:16px">149</div><div data-line-number="150" class="react-line-number react-code-text" style="padding-right:16px">150</div><div data-line-number="151" class="react-line-number react-code-text" style="padding-right:16px">151</div><div data-line-number="152" class="react-line-number react-code-text" style="padding-right:16px">152</div><div data-line-number="153" class="react-line-number react-code-text" style="padding-right:16px">153</div><div data-line-number="154" class="react-line-number react-code-text" style="padding-right:16px">154</div><div data-line-number="155" class="react-line-number react-code-text" style="padding-right:16px">155</div><div data-line-number="156" class="react-line-number react-code-text" style="padding-right:16px">156</div><div data-line-number="157" class="react-line-number react-code-text" style="padding-right:16px">157</div><div data-line-number="158" class="react-line-number react-code-text" style="padding-right:16px">158</div><div data-line-number="159" class="react-line-number react-code-text" style="padding-right:16px">159</div><div data-line-number="160" class="react-line-number react-code-text" style="padding-right:16px">160</div><div data-line-number="161" class="react-line-number react-code-text" style="padding-right:16px">161</div><div data-line-number="162" class="react-line-number react-code-text" style="padding-right:16px">162</div><div data-line-number="163" class="react-line-number react-code-text" style="padding-right:16px">163</div><div data-line-number="164" class="react-line-number react-code-text" style="padding-right:16px">164</div><div data-line-number="165" class="react-line-number react-code-text" style="padding-right:16px">165</div><div data-line-number="166" class="react-line-number react-code-text" style="padding-right:16px">166</div><div data-line-number="167" class="react-line-number react-code-text" style="padding-right:16px">167</div><div data-line-number="168" class="react-line-number react-code-text" style="padding-right:16px">168</div><div data-line-number="169" class="react-line-number react-code-text" style="padding-right:16px">169</div><div data-line-number="170" class="react-line-number react-code-text" style="padding-right:16px">170</div><div data-line-number="171" class="react-line-number react-code-text" style="padding-right:16px">171</div><div data-line-number="172" class="react-line-number react-code-text" style="padding-right:16px">172</div><div data-line-number="173" class="react-line-number react-code-text" style="padding-right:16px">173</div><div data-line-number="174" class="react-line-number react-code-text" style="padding-right:16px">174</div><div data-line-number="175" class="react-line-number react-code-text" style="padding-right:16px">175</div><div data-line-number="176" class="react-line-number react-code-text" style="padding-right:16px">176</div><div data-line-number="177" class="react-line-number react-code-text" style="padding-right:16px">177</div><div data-line-number="178" class="react-line-number react-code-text" style="padding-right:16px">178</div><div data-line-number="179" class="react-line-number react-code-text" style="padding-right:16px">179</div><div data-line-number="180" class="react-line-number react-code-text" style="padding-right:16px">180</div><div data-line-number="181" class="react-line-number react-code-text" style="padding-right:16px">181</div><div data-line-number="182" class="react-line-number react-code-text" style="padding-right:16px">182</div><div data-line-number="183" class="react-line-number react-code-text" style="padding-right:16px">183</div><div data-line-number="184" class="react-line-number react-code-text" style="padding-right:16px">184</div><div data-line-number="185" class="react-line-number react-code-text" style="padding-right:16px">185</div><div data-line-number="186" class="react-line-number react-code-text" style="padding-right:16px">186</div><div data-line-number="187" class="react-line-number react-code-text" style="padding-right:16px">187</div><div data-line-number="188" class="react-line-number react-code-text" style="padding-right:16px">188</div><div data-line-number="189" class="react-line-number react-code-text" style="padding-right:16px">189</div><div data-line-number="190" class="react-line-number react-code-text" style="padding-right:16px">190</div><div data-line-number="191" class="react-line-number react-code-text" style="padding-right:16px">191</div><div data-line-number="192" class="react-line-number react-code-text" style="padding-right:16px">192</div><div data-line-number="193" class="react-line-number react-code-text" style="padding-right:16px">193</div><div data-line-number="194" class="react-line-number react-code-text" style="padding-right:16px">194</div><div data-line-number="195" class="react-line-number react-code-text" style="padding-right:16px">195</div><div data-line-number="196" class="react-line-number react-code-text" style="padding-right:16px">196</div><div data-line-number="197" class="react-line-number react-code-text" style="padding-right:16px">197</div><div data-line-number="198" class="react-line-number react-code-text" style="padding-right:16px">198</div><div data-line-number="199" class="react-line-number react-code-text" style="padding-right:16px">199</div><div data-line-number="200" class="react-line-number react-code-text" style="padding-right:16px">200</div><div data-line-number="201" class="react-line-number react-code-text" style="padding-right:16px">201</div><div data-line-number="202" class="react-line-number react-code-text" style="padding-right:16px">202</div><div data-line-number="203" class="react-line-number react-code-text" style="padding-right:16px">203</div><div data-line-number="204" class="react-line-number react-code-text" style="padding-right:16px">204</div><div data-line-number="205" class="react-line-number react-code-text" style="padding-right:16px">205</div><div data-line-number="206" class="react-line-number react-code-text" style="padding-right:16px">206</div><div data-line-number="207" class="react-line-number react-code-text" style="padding-right:16px">207</div><div data-line-number="208" class="react-line-number react-code-text" style="padding-right:16px">208</div><div data-line-number="209" class="react-line-number react-code-text" style="padding-right:16px">209</div><div data-line-number="210" class="react-line-number react-code-text" style="padding-right:16px">210</div></div><div class="react-code-lines"><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC1" class="react-file-line html-div" data-testid="code-cell" data-line-number="1" style="position:relative"><span class="pl-s">---</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC2" class="react-file-line html-div" data-testid="code-cell" data-line-number="2" style="position:relative"><span class="pl-ent">title</span>: <span class="pl-s"><span class="pl-pds">&quot;</span>GO and KEGG Term Enrichment<span class="pl-pds">&quot;</span></span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC3" class="react-file-line html-div" data-testid="code-cell" data-line-number="3" style="position:relative"><span class="pl-ent">author</span>: <span class="pl-s"><span class="pl-pds">&quot;</span>John Santiago<span class="pl-pds">&quot;</span></span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC4" class="react-file-line html-div" data-testid="code-cell" data-line-number="4" style="position:relative"><span class="pl-ent">date</span>: <span class="pl-s"><span class="pl-pds">&quot;</span>2024-05-08<span class="pl-pds">&quot;</span></span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC5" class="react-file-line html-div" data-testid="code-cell" data-line-number="5" style="position:relative"><span class="pl-ent">output</span>: <span class="pl-s">html_document</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC6" class="react-file-line html-div" data-testid="code-cell" data-line-number="6" style="position:relative"><span class="pl-s">---</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC7" class="react-file-line html-div" data-testid="code-cell" data-line-number="7" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC8" class="react-file-line html-div" data-testid="code-cell" data-line-number="8" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> setup, include=FALSE}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC9" class="react-file-line html-div" data-testid="code-cell" data-line-number="9" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC10" class="react-file-line html-div" data-testid="code-cell" data-line-number="10" style="position:relative"><span class="pl-c1">knitr::opts_chunk$set(echo = TRUE)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC11" class="react-file-line html-div" data-testid="code-cell" data-line-number="11" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC12" class="react-file-line html-div" data-testid="code-cell" data-line-number="12" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC13" class="react-file-line html-div" data-testid="code-cell" data-line-number="13" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC14" class="react-file-line html-div" data-testid="code-cell" data-line-number="14" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC15" class="react-file-line html-div" data-testid="code-cell" data-line-number="15" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC16" class="react-file-line html-div" data-testid="code-cell" data-line-number="16" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F, message = F}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC17" class="react-file-line html-div" data-testid="code-cell" data-line-number="17" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC18" class="react-file-line html-div" data-testid="code-cell" data-line-number="18" style="position:relative"><span class="pl-c1"># Conditions</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC19" class="react-file-line html-div" data-testid="code-cell" data-line-number="19" style="position:relative"><span class="pl-c1">#Geno is WT or GR</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC20" class="react-file-line html-div" data-testid="code-cell" data-line-number="20" style="position:relative"><span class="pl-c1">#TKT is either &quot;C&quot;, &quot;OE&quot;, or &quot;Df&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC21" class="react-file-line html-div" data-testid="code-cell" data-line-number="21" style="position:relative"><span class="pl-c1">#Sex is &quot;F&quot; or &quot;M&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC22" class="react-file-line html-div" data-testid="code-cell" data-line-number="22" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC23" class="react-file-line html-div" data-testid="code-cell" data-line-number="23" style="position:relative"><span class="pl-c1">Geno1 = &quot;GR&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC24" class="react-file-line html-div" data-testid="code-cell" data-line-number="24" style="position:relative"><span class="pl-c1">TKT1 = &quot;C&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC25" class="react-file-line html-div" data-testid="code-cell" data-line-number="25" style="position:relative"><span class="pl-c1">Sex1 = &quot;F&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC26" class="react-file-line html-div" data-testid="code-cell" data-line-number="26" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC27" class="react-file-line html-div" data-testid="code-cell" data-line-number="27" style="position:relative"><span class="pl-c1">Geno2 = &quot;WT&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC28" class="react-file-line html-div" data-testid="code-cell" data-line-number="28" style="position:relative"><span class="pl-c1">TKT2 = &quot;C&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC29" class="react-file-line html-div" data-testid="code-cell" data-line-number="29" style="position:relative"><span class="pl-c1">Sex2 = &quot;F&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC30" class="react-file-line html-div" data-testid="code-cell" data-line-number="30" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC31" class="react-file-line html-div" data-testid="code-cell" data-line-number="31" style="position:relative"><span class="pl-c1">##Conditions for second comparison</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC32" class="react-file-line html-div" data-testid="code-cell" data-line-number="32" style="position:relative"><span class="pl-c1">Geno3 = &quot;GR&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC33" class="react-file-line html-div" data-testid="code-cell" data-line-number="33" style="position:relative"><span class="pl-c1">TKT3 = &quot;C&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC34" class="react-file-line html-div" data-testid="code-cell" data-line-number="34" style="position:relative"><span class="pl-c1">Sex3 = &quot;M&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC35" class="react-file-line html-div" data-testid="code-cell" data-line-number="35" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC36" class="react-file-line html-div" data-testid="code-cell" data-line-number="36" style="position:relative"><span class="pl-c1">Geno4 = &quot;WT&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC37" class="react-file-line html-div" data-testid="code-cell" data-line-number="37" style="position:relative"><span class="pl-c1">TKT4 = &quot;C&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC38" class="react-file-line html-div" data-testid="code-cell" data-line-number="38" style="position:relative"><span class="pl-c1">Sex4 = &quot;M&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC39" class="react-file-line html-div" data-testid="code-cell" data-line-number="39" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC40" class="react-file-line html-div" data-testid="code-cell" data-line-number="40" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC41" class="react-file-line html-div" data-testid="code-cell" data-line-number="41" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC42" class="react-file-line html-div" data-testid="code-cell" data-line-number="42" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC43" class="react-file-line html-div" data-testid="code-cell" data-line-number="43" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC44" class="react-file-line html-div" data-testid="code-cell" data-line-number="44" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC45" class="react-file-line html-div" data-testid="code-cell" data-line-number="45" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC46" class="react-file-line html-div" data-testid="code-cell" data-line-number="46" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F, message = F}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC47" class="react-file-line html-div" data-testid="code-cell" data-line-number="47" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC48" class="react-file-line html-div" data-testid="code-cell" data-line-number="48" style="position:relative"><span class="pl-c1">source(&#039;/Users/johncsantiago/Documents/GitHub/WhartonLab/G85R.GOandKEGGPlottingFunctions.R&#039;)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC49" class="react-file-line html-div" data-testid="code-cell" data-line-number="49" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC50" class="react-file-line html-div" data-testid="code-cell" data-line-number="50" style="position:relative"><span class="pl-c1">plot.data = G85R.data(Geno1, Geno2, Geno3, Geno4,</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC51" class="react-file-line html-div" data-testid="code-cell" data-line-number="51" style="position:relative"><span class="pl-c1">                      Sex1, Sex2, Sex3, Sex4,</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC52" class="react-file-line html-div" data-testid="code-cell" data-line-number="52" style="position:relative"><span class="pl-c1">                      TKT1, TKT2, TKT3, TKT4)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC53" class="react-file-line html-div" data-testid="code-cell" data-line-number="53" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC54" class="react-file-line html-div" data-testid="code-cell" data-line-number="54" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC55" class="react-file-line html-div" data-testid="code-cell" data-line-number="55" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC56" class="react-file-line html-div" data-testid="code-cell" data-line-number="56" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC57" class="react-file-line html-div" data-testid="code-cell" data-line-number="57" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC58" class="react-file-line html-div" data-testid="code-cell" data-line-number="58" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC59" class="react-file-line html-div" data-testid="code-cell" data-line-number="59" style="position:relative"><span class="pl-c1">##Uses data from primary comparison</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC60" class="react-file-line html-div" data-testid="code-cell" data-line-number="60" style="position:relative"><span class="pl-c1">##Plots genes with an FDR &lt;= 0.05</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC61" class="react-file-line html-div" data-testid="code-cell" data-line-number="61" style="position:relative"><span class="pl-c1">##point size is relative to maximum mean expression in the two conditions</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC62" class="react-file-line html-div" data-testid="code-cell" data-line-number="62" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC63" class="react-file-line html-div" data-testid="code-cell" data-line-number="63" style="position:relative"><span class="pl-c1">G85R.volcano(plot.data)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC64" class="react-file-line html-div" data-testid="code-cell" data-line-number="64" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC65" class="react-file-line html-div" data-testid="code-cell" data-line-number="65" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC66" class="react-file-line html-div" data-testid="code-cell" data-line-number="66" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC67" class="react-file-line html-div" data-testid="code-cell" data-line-number="67" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC68" class="react-file-line html-div" data-testid="code-cell" data-line-number="68" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC69" class="react-file-line html-div" data-testid="code-cell" data-line-number="69" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC70" class="react-file-line html-div" data-testid="code-cell" data-line-number="70" style="position:relative"><span class="pl-c1">##Uses data from primary comparison</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC71" class="react-file-line html-div" data-testid="code-cell" data-line-number="71" style="position:relative"><span class="pl-c1">##point size is relative to max mean expression in any condition</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC72" class="react-file-line html-div" data-testid="code-cell" data-line-number="72" style="position:relative"><span class="pl-c1">##grey points indicate less than 1 fold difference between the comparisons</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC73" class="react-file-line html-div" data-testid="code-cell" data-line-number="73" style="position:relative"><span class="pl-c1">G85R.compare.genes(plot.data)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC74" class="react-file-line html-div" data-testid="code-cell" data-line-number="74" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC75" class="react-file-line html-div" data-testid="code-cell" data-line-number="75" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC76" class="react-file-line html-div" data-testid="code-cell" data-line-number="76" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC77" class="react-file-line html-div" data-testid="code-cell" data-line-number="77" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC78" class="react-file-line html-div" data-testid="code-cell" data-line-number="78" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC79" class="react-file-line html-div" data-testid="code-cell" data-line-number="79" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC80" class="react-file-line html-div" data-testid="code-cell" data-line-number="80" style="position:relative"><span class="pl-c1">##Uses data from primary comparison</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC81" class="react-file-line html-div" data-testid="code-cell" data-line-number="81" style="position:relative"><span class="pl-c1">##point size is relative to mean expression in variable condition</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC82" class="react-file-line html-div" data-testid="code-cell" data-line-number="82" style="position:relative"><span class="pl-c1">G85R.DEG.venn(plot.data)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC83" class="react-file-line html-div" data-testid="code-cell" data-line-number="83" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC84" class="react-file-line html-div" data-testid="code-cell" data-line-number="84" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC85" class="react-file-line html-div" data-testid="code-cell" data-line-number="85" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC86" class="react-file-line html-div" data-testid="code-cell" data-line-number="86" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F, include = F, warning = F}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC87" class="react-file-line html-div" data-testid="code-cell" data-line-number="87" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC88" class="react-file-line html-div" data-testid="code-cell" data-line-number="88" style="position:relative"><span class="pl-c1">####Uses sigDEGs from primary comparison</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC89" class="react-file-line html-div" data-testid="code-cell" data-line-number="89" style="position:relative"><span class="pl-c1">cat.data = G85R.Enrichment(plot.data)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC90" class="react-file-line html-div" data-testid="code-cell" data-line-number="90" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC91" class="react-file-line html-div" data-testid="code-cell" data-line-number="91" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC92" class="react-file-line html-div" data-testid="code-cell" data-line-number="92" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC93" class="react-file-line html-div" data-testid="code-cell" data-line-number="93" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC94" class="react-file-line html-div" data-testid="code-cell" data-line-number="94" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC95" class="react-file-line html-div" data-testid="code-cell" data-line-number="95" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC96" class="react-file-line html-div" data-testid="code-cell" data-line-number="96" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC97" class="react-file-line html-div" data-testid="code-cell" data-line-number="97" style="position:relative"><span class="pl-c1">##which ranks of categories to use</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC98" class="react-file-line html-div" data-testid="code-cell" data-line-number="98" style="position:relative"><span class="pl-c1">##a vector of numbers specifying which ranks to use</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC99" class="react-file-line html-div" data-testid="code-cell" data-line-number="99" style="position:relative"><span class="pl-c1">##eg. c(1, 5:6) will use the first, fifth and sixth most sig. cats.</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC100" class="react-file-line html-div" data-testid="code-cell" data-line-number="100" style="position:relative"><span class="pl-c1">##can also just type &#039;sigs&#039; </span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC101" class="react-file-line html-div" data-testid="code-cell" data-line-number="101" style="position:relative"><span class="pl-c1">usecats=c(1:20)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC102" class="react-file-line html-div" data-testid="code-cell" data-line-number="102" style="position:relative"><span class="pl-c1">#usecats = &#039;sigs&#039;  </span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC103" class="react-file-line html-div" data-testid="code-cell" data-line-number="103" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC104" class="react-file-line html-div" data-testid="code-cell" data-line-number="104" style="position:relative"><span class="pl-c1">G85R.KEGG20(cat.data, plot.data, usecats)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC105" class="react-file-line html-div" data-testid="code-cell" data-line-number="105" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC106" class="react-file-line html-div" data-testid="code-cell" data-line-number="106" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC107" class="react-file-line html-div" data-testid="code-cell" data-line-number="107" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC108" class="react-file-line html-div" data-testid="code-cell" data-line-number="108" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC109" class="react-file-line html-div" data-testid="code-cell" data-line-number="109" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F, include = T}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC110" class="react-file-line html-div" data-testid="code-cell" data-line-number="110" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC111" class="react-file-line html-div" data-testid="code-cell" data-line-number="111" style="position:relative"><span class="pl-c1">##can be the exact name or 5 digit code in quotes</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC112" class="react-file-line html-div" data-testid="code-cell" data-line-number="112" style="position:relative"><span class="pl-c1">specific.kegg = &quot;00980&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC113" class="react-file-line html-div" data-testid="code-cell" data-line-number="113" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC114" class="react-file-line html-div" data-testid="code-cell" data-line-number="114" style="position:relative"><span class="pl-c1">G85R.plotkegg.genes(specific.kegg, plot.data)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC115" class="react-file-line html-div" data-testid="code-cell" data-line-number="115" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC116" class="react-file-line html-div" data-testid="code-cell" data-line-number="116" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC117" class="react-file-line html-div" data-testid="code-cell" data-line-number="117" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC118" class="react-file-line html-div" data-testid="code-cell" data-line-number="118" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC119" class="react-file-line html-div" data-testid="code-cell" data-line-number="119" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F, message=F, fig.width=7}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC120" class="react-file-line html-div" data-testid="code-cell" data-line-number="120" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC121" class="react-file-line html-div" data-testid="code-cell" data-line-number="121" style="position:relative"><span class="pl-c1">##5 digits from KEGG ID</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC122" class="react-file-line html-div" data-testid="code-cell" data-line-number="122" style="position:relative"><span class="pl-c1">PathwayID = &quot;00010&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC123" class="react-file-line html-div" data-testid="code-cell" data-line-number="123" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC124" class="react-file-line html-div" data-testid="code-cell" data-line-number="124" style="position:relative"><span class="pl-c1">#options include &#039;primary&#039; or &#039;both&#039;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC125" class="react-file-line html-div" data-testid="code-cell" data-line-number="125" style="position:relative"><span class="pl-c1">use = &#039;both&#039;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC126" class="react-file-line html-div" data-testid="code-cell" data-line-number="126" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC127" class="react-file-line html-div" data-testid="code-cell" data-line-number="127" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC128" class="react-file-line html-div" data-testid="code-cell" data-line-number="128" style="position:relative"><span class="pl-c1">G85R.KEGG.diagram(PathwayID, plot.data, use, sigs.only = F)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC129" class="react-file-line html-div" data-testid="code-cell" data-line-number="129" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC130" class="react-file-line html-div" data-testid="code-cell" data-line-number="130" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC131" class="react-file-line html-div" data-testid="code-cell" data-line-number="131" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC132" class="react-file-line html-div" data-testid="code-cell" data-line-number="132" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC133" class="react-file-line html-div" data-testid="code-cell" data-line-number="133" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC134" class="react-file-line html-div" data-testid="code-cell" data-line-number="134" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F, include = T, warnings = F}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC135" class="react-file-line html-div" data-testid="code-cell" data-line-number="135" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC136" class="react-file-line html-div" data-testid="code-cell" data-line-number="136" style="position:relative"><span class="pl-c1">##can be the exact name or 5 digit code in quotes</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC137" class="react-file-line html-div" data-testid="code-cell" data-line-number="137" style="position:relative"><span class="pl-c1">specific.kegg = &quot;00010&quot;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC138" class="react-file-line html-div" data-testid="code-cell" data-line-number="138" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC139" class="react-file-line html-div" data-testid="code-cell" data-line-number="139" style="position:relative"><span class="pl-c1">G85R.comparekegg.genes(specific.kegg, plot.data)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC140" class="react-file-line html-div" data-testid="code-cell" data-line-number="140" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC141" class="react-file-line html-div" data-testid="code-cell" data-line-number="141" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC142" class="react-file-line html-div" data-testid="code-cell" data-line-number="142" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC143" class="react-file-line html-div" data-testid="code-cell" data-line-number="143" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC144" class="react-file-line html-div" data-testid="code-cell" data-line-number="144" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC145" class="react-file-line html-div" data-testid="code-cell" data-line-number="145" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC146" class="react-file-line html-div" data-testid="code-cell" data-line-number="146" style="position:relative"><span class="pl-c1">## BP, CC, or MF</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC147" class="react-file-line html-div" data-testid="code-cell" data-line-number="147" style="position:relative"><span class="pl-c1">GO.ontology = &#039;BP&#039;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC148" class="react-file-line html-div" data-testid="code-cell" data-line-number="148" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC149" class="react-file-line html-div" data-testid="code-cell" data-line-number="149" style="position:relative"><span class="pl-c1">##can select which categories ot include or say &#039;sigs.only&#039;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC150" class="react-file-line html-div" data-testid="code-cell" data-line-number="150" style="position:relative"><span class="pl-c1">usecats=c(1:20)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC151" class="react-file-line html-div" data-testid="code-cell" data-line-number="151" style="position:relative"><span class="pl-c1">#usecats = &#039;sigs.only&#039;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC152" class="react-file-line html-div" data-testid="code-cell" data-line-number="152" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC153" class="react-file-line html-div" data-testid="code-cell" data-line-number="153" style="position:relative"><span class="pl-c1">G85R.GO20(cat.data, GO.ontology, plot.data, usecats)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC154" class="react-file-line html-div" data-testid="code-cell" data-line-number="154" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC155" class="react-file-line html-div" data-testid="code-cell" data-line-number="155" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC156" class="react-file-line html-div" data-testid="code-cell" data-line-number="156" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC157" class="react-file-line html-div" data-testid="code-cell" data-line-number="157" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC158" class="react-file-line html-div" data-testid="code-cell" data-line-number="158" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC159" class="react-file-line html-div" data-testid="code-cell" data-line-number="159" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC160" class="react-file-line html-div" data-testid="code-cell" data-line-number="160" style="position:relative"><span class="pl-c1">##the 7 digits of a GOterm id or the exact term name</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC161" class="react-file-line html-div" data-testid="code-cell" data-line-number="161" style="position:relative"><span class="pl-c1">specific.go = &#039;0005811&#039;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC162" class="react-file-line html-div" data-testid="code-cell" data-line-number="162" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC163" class="react-file-line html-div" data-testid="code-cell" data-line-number="163" style="position:relative"><span class="pl-c1">G85R.plotgo.genes(specific.go, plot.data)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC164" class="react-file-line html-div" data-testid="code-cell" data-line-number="164" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC165" class="react-file-line html-div" data-testid="code-cell" data-line-number="165" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC166" class="react-file-line html-div" data-testid="code-cell" data-line-number="166" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC167" class="react-file-line html-div" data-testid="code-cell" data-line-number="167" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC168" class="react-file-line html-div" data-testid="code-cell" data-line-number="168" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F, warning = F}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC169" class="react-file-line html-div" data-testid="code-cell" data-line-number="169" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC170" class="react-file-line html-div" data-testid="code-cell" data-line-number="170" style="position:relative"><span class="pl-c1">##the 7 digits of a GOterm id or the exact term name</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC171" class="react-file-line html-div" data-testid="code-cell" data-line-number="171" style="position:relative"><span class="pl-c1">specific.go = &#039;0005811&#039;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC172" class="react-file-line html-div" data-testid="code-cell" data-line-number="172" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC173" class="react-file-line html-div" data-testid="code-cell" data-line-number="173" style="position:relative"><span class="pl-c1">G85R.comparego.genes(specific.go, plot.data)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC174" class="react-file-line html-div" data-testid="code-cell" data-line-number="174" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC175" class="react-file-line html-div" data-testid="code-cell" data-line-number="175" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC176" class="react-file-line html-div" data-testid="code-cell" data-line-number="176" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC177" class="react-file-line html-div" data-testid="code-cell" data-line-number="177" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC178" class="react-file-line html-div" data-testid="code-cell" data-line-number="178" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F, warning = F, message = F}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC179" class="react-file-line html-div" data-testid="code-cell" data-line-number="179" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC180" class="react-file-line html-div" data-testid="code-cell" data-line-number="180" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC181" class="react-file-line html-div" data-testid="code-cell" data-line-number="181" style="position:relative"><span class="pl-c1">## BP, CC, or MF</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC182" class="react-file-line html-div" data-testid="code-cell" data-line-number="182" style="position:relative"><span class="pl-c1">GO.ontology = &#039;BP&#039;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC183" class="react-file-line html-div" data-testid="code-cell" data-line-number="183" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC184" class="react-file-line html-div" data-testid="code-cell" data-line-number="184" style="position:relative"><span class="pl-c1">rt = tree(cat.data, GO.ontology)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC185" class="react-file-line html-div" data-testid="code-cell" data-line-number="185" style="position:relative"><span class="pl-c1">  </span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC186" class="react-file-line html-div" data-testid="code-cell" data-line-number="186" style="position:relative"><span class="pl-c1">treemapPlot(rt)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC187" class="react-file-line html-div" data-testid="code-cell" data-line-number="187" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC188" class="react-file-line html-div" data-testid="code-cell" data-line-number="188" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC189" class="react-file-line html-div" data-testid="code-cell" data-line-number="189" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC190" class="react-file-line html-div" data-testid="code-cell" data-line-number="190" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC191" class="react-file-line html-div" data-testid="code-cell" data-line-number="191" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F, warning = F}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC192" class="react-file-line html-div" data-testid="code-cell" data-line-number="192" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC193" class="react-file-line html-div" data-testid="code-cell" data-line-number="193" style="position:relative"><span class="pl-c1">##a parent category generated in the treemapPlot function</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC194" class="react-file-line html-div" data-testid="code-cell" data-line-number="194" style="position:relative"><span class="pl-c1">parent.category = &#039;neurogenesis&#039;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC195" class="react-file-line html-div" data-testid="code-cell" data-line-number="195" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC196" class="react-file-line html-div" data-testid="code-cell" data-line-number="196" style="position:relative"><span class="pl-c1">G85R.plotparentcat.genes(parent.category, plot.data)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC197" class="react-file-line html-div" data-testid="code-cell" data-line-number="197" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC198" class="react-file-line html-div" data-testid="code-cell" data-line-number="198" style="position:relative"><span class="pl-s">```</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC199" class="react-file-line html-div" data-testid="code-cell" data-line-number="199" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC200" class="react-file-line html-div" data-testid="code-cell" data-line-number="200" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC201" class="react-file-line html-div" data-testid="code-cell" data-line-number="201" style="position:relative"><span class="pl-s">```</span><span class="pl-en">{r</span> echo = F}</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC202" class="react-file-line html-div" data-testid="code-cell" data-line-number="202" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC203" class="react-file-line html-div" data-testid="code-cell" data-line-number="203" style="position:relative"><span class="pl-c1">##a parent category generated in the treemapPlot function</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC204" class="react-file-line html-div" data-testid="code-cell" data-line-number="204" style="position:relative"><span class="pl-c1">##parent.category = &#039;small molecule biosynthetic process&#039;</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC205" class="react-file-line html-div" data-testid="code-cell" data-line-number="205" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC206" class="react-file-line html-div" data-testid="code-cell" data-line-number="206" style="position:relative"><span class="pl-c1">G85R.compareparent.genes(parent.category, plot.data)</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC207" class="react-file-line html-div" data-testid="code-cell" data-line-number="207" style="position:relative"><span class="pl-c1">##dot size is relative to the max FDR of either condition</span></div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC208" class="react-file-line html-div" data-testid="code-cell" data-line-number="208" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC209" class="react-file-line html-div" data-testid="code-cell" data-line-number="209" style="position:relative">
</div></div></div><div class="react-code-text react-code-line-contents" style="min-height:auto"><div><div id="LC210" class="react-file-line html-div" data-testid="code-cell" data-line-number="210" style="position:relative"><span class="pl-s">```</span>  </div></div></div></div></div></div><div id="copilot-button-container"></div></div><div id="highlighted-line-menu-container"></div></div></div><button hidden="" data-testid="hotkey-button" data-hotkey-scope="read-only-cursor-text-area"></button><button hidden=""></button></section></div></div></div> <!-- --> <!-- --> </div></div></div><div class="Box-sc-g0xbh4-0"></div></div></div></div></div><div id="find-result-marks-container" class="Box-sc-g0xbh4-0 cCoXib"></div><button hidden="" data-testid="" data-hotkey-scope="read-only-cursor-text-area"></button><button hidden=""></button></div> <!-- --> <!-- --> <script type="application/json" id="__PRIMER_DATA_:R0:__">{"resolvedServerColorMode":"day"}</script></div>
</react-app>
</turbo-frame>

  </div>

</turbo-frame>

    </main>
  </div>

  </div>

          <footer class="footer pt-8 pb-6 f6 color-fg-muted p-responsive" role="contentinfo" >
  <h2 class='sr-only'>Footer</h2>

  <div class="d-flex flex-justify-center flex-items-center flex-column-reverse flex-lg-row flex-wrap flex-lg-nowrap">
    <div class="d-flex flex-items-center flex-shrink-0 mx-2">
      <a aria-label="Homepage" title="GitHub" class="footer-octicon mr-2" href="https://github.com">
        <svg aria-hidden="true" height="24" viewBox="0 0 24 24" version="1.1" width="24" data-view-component="true" class="octicon octicon-mark-github">
    <path d="M12.5.75C6.146.75 1 5.896 1 12.25c0 5.089 3.292 9.387 7.863 10.91.575.101.79-.244.79-.546 0-.273-.014-1.178-.014-2.142-2.889.532-3.636-.704-3.866-1.35-.13-.331-.69-1.352-1.18-1.625-.402-.216-.977-.748-.014-.762.906-.014 1.553.834 1.769 1.179 1.035 1.74 2.688 1.25 3.349.948.1-.747.402-1.25.733-1.538-2.559-.287-5.232-1.279-5.232-5.678 0-1.25.445-2.285 1.178-3.09-.115-.288-.517-1.467.115-3.048 0 0 .963-.302 3.163 1.179.92-.259 1.897-.388 2.875-.388.977 0 1.955.13 2.875.388 2.2-1.495 3.162-1.179 3.162-1.179.633 1.581.23 2.76.115 3.048.733.805 1.179 1.825 1.179 3.09 0 4.413-2.688 5.39-5.247 5.678.417.36.776 1.05.776 2.128 0 1.538-.014 2.774-.014 3.162 0 .302.216.662.79.547C20.709 21.637 24 17.324 24 12.25 24 5.896 18.854.75 12.5.75Z"></path>
</svg>
</a>
      <span>
        &copy; 2025 GitHub,&nbsp;Inc.
      </span>
    </div>

    <nav aria-label="Footer">
      <h3 class="sr-only" id="sr-footer-heading">Footer navigation</h3>

      <ul class="list-style-none d-flex flex-justify-center flex-wrap mb-2 mb-lg-0" aria-labelledby="sr-footer-heading">

          <li class="mx-2">
            <a data-analytics-event="{&quot;category&quot;:&quot;Footer&quot;,&quot;action&quot;:&quot;go to Terms&quot;,&quot;label&quot;:&quot;text:terms&quot;}" href="https://docs.github.com/site-policy/github-terms/github-terms-of-service" data-view-component="true" class="Link--secondary Link">Terms</a>
          </li>

          <li class="mx-2">
            <a data-analytics-event="{&quot;category&quot;:&quot;Footer&quot;,&quot;action&quot;:&quot;go to privacy&quot;,&quot;label&quot;:&quot;text:privacy&quot;}" href="https://docs.github.com/site-policy/privacy-policies/github-privacy-statement" data-view-component="true" class="Link--secondary Link">Privacy</a>
          </li>

          <li class="mx-2">
            <a data-analytics-event="{&quot;category&quot;:&quot;Footer&quot;,&quot;action&quot;:&quot;go to security&quot;,&quot;label&quot;:&quot;text:security&quot;}" href="https://github.com/security" data-view-component="true" class="Link--secondary Link">Security</a>
          </li>

          <li class="mx-2">
            <a data-analytics-event="{&quot;category&quot;:&quot;Footer&quot;,&quot;action&quot;:&quot;go to status&quot;,&quot;label&quot;:&quot;text:status&quot;}" href="https://www.githubstatus.com/" data-view-component="true" class="Link--secondary Link">Status</a>
          </li>

          <li class="mx-2">
            <a data-analytics-event="{&quot;category&quot;:&quot;Footer&quot;,&quot;action&quot;:&quot;go to docs&quot;,&quot;label&quot;:&quot;text:docs&quot;}" href="https://docs.github.com/" data-view-component="true" class="Link--secondary Link">Docs</a>
          </li>

          <li class="mx-2">
            <a data-analytics-event="{&quot;category&quot;:&quot;Footer&quot;,&quot;action&quot;:&quot;go to contact&quot;,&quot;label&quot;:&quot;text:contact&quot;}" href="https://support.github.com?tags=dotcom-footer" data-view-component="true" class="Link--secondary Link">Contact</a>
          </li>

          <li class="mx-2" >
  <cookie-consent-link>
    <button
      type="button"
      class="Link--secondary underline-on-hover border-0 p-0 color-bg-transparent"
      data-action="click:cookie-consent-link#showConsentManagement"
      data-analytics-event="{&quot;location&quot;:&quot;footer&quot;,&quot;action&quot;:&quot;cookies&quot;,&quot;context&quot;:&quot;subfooter&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;cookies_link_subfooter_footer&quot;}"
    >
      Manage cookies
    </button>
  </cookie-consent-link>
</li>

<li class="mx-2">
  <cookie-consent-link>
    <button
      type="button"
      class="Link--secondary underline-on-hover border-0 p-0 color-bg-transparent"
      data-action="click:cookie-consent-link#showConsentManagement"
      data-analytics-event="{&quot;location&quot;:&quot;footer&quot;,&quot;action&quot;:&quot;dont_share_info&quot;,&quot;context&quot;:&quot;subfooter&quot;,&quot;tag&quot;:&quot;link&quot;,&quot;label&quot;:&quot;dont_share_info_link_subfooter_footer&quot;}"
    >
      Do not share my personal information
    </button>
  </cookie-consent-link>
</li>

      </ul>
    </nav>
  </div>
</footer>

    <ghcc-consent id="ghcc" class="position-fixed bottom-0 left-0" style="z-index: 999999" data-initial-cookie-consent-allowed="" data-cookie-consent-required="false"></ghcc-consent>

  <div id="ajax-error-message" class="ajax-error-message flash flash-error" hidden>
    <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-alert">
    <path d="M6.457 1.047c.659-1.234 2.427-1.234 3.086 0l6.082 11.378A1.75 1.75 0 0 1 14.082 15H1.918a1.75 1.75 0 0 1-1.543-2.575Zm1.763.707a.25.25 0 0 0-.44 0L1.698 13.132a.25.25 0 0 0 .22.368h12.164a.25.25 0 0 0 .22-.368Zm.53 3.996v2.5a.75.75 0 0 1-1.5 0v-2.5a.75.75 0 0 1 1.5 0ZM9 11a1 1 0 1 1-2 0 1 1 0 0 1 2 0Z"></path>
</svg>
    <button type="button" class="flash-close js-ajax-error-dismiss" aria-label="Dismiss error">
      <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-x">
    <path d="M3.72 3.72a.75.75 0 0 1 1.06 0L8 6.94l3.22-3.22a.749.749 0 0 1 1.275.326.749.749 0 0 1-.215.734L9.06 8l3.22 3.22a.749.749 0 0 1-.326 1.275.749.749 0 0 1-.734-.215L8 9.06l-3.22 3.22a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042L6.94 8 3.72 4.78a.75.75 0 0 1 0-1.06Z"></path>
</svg>
    </button>
    You can’t perform that action at this time.
  </div>

    <template id="site-details-dialog">
  <details class="details-reset details-overlay details-overlay-dark lh-default color-fg-default hx_rsm" open>
    <summary role="button" aria-label="Close dialog"></summary>
    <details-dialog class="Box Box--overlay d-flex flex-column anim-fade-in fast hx_rsm-dialog hx_rsm-modal">
      <button class="Box-btn-octicon m-0 btn-octicon position-absolute right-0 top-0" type="button" aria-label="Close dialog" data-close-dialog>
        <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-x">
    <path d="M3.72 3.72a.75.75 0 0 1 1.06 0L8 6.94l3.22-3.22a.749.749 0 0 1 1.275.326.749.749 0 0 1-.215.734L9.06 8l3.22 3.22a.749.749 0 0 1-.326 1.275.749.749 0 0 1-.734-.215L8 9.06l-3.22 3.22a.751.751 0 0 1-1.042-.018.751.751 0 0 1-.018-1.042L6.94 8 3.72 4.78a.75.75 0 0 1 0-1.06Z"></path>
</svg>
      </button>
      <div class="octocat-spinner my-6 js-details-dialog-spinner"></div>
    </details-dialog>
  </details>
</template>

    <div class="Popover js-hovercard-content position-absolute" style="display: none; outline: none;">
  <div class="Popover-message Popover-message--bottom-left Popover-message--large Box color-shadow-large" style="width:360px;">
  </div>
</div>

    <template id="snippet-clipboard-copy-button">
  <div class="zeroclipboard-container position-absolute right-0 top-0">
    <clipboard-copy aria-label="Copy" class="ClipboardButton btn js-clipboard-copy m-2 p-0" data-copy-feedback="Copied!" data-tooltip-direction="w">
      <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-copy js-clipboard-copy-icon m-2">
    <path d="M0 6.75C0 5.784.784 5 1.75 5h1.5a.75.75 0 0 1 0 1.5h-1.5a.25.25 0 0 0-.25.25v7.5c0 .138.112.25.25.25h7.5a.25.25 0 0 0 .25-.25v-1.5a.75.75 0 0 1 1.5 0v1.5A1.75 1.75 0 0 1 9.25 16h-7.5A1.75 1.75 0 0 1 0 14.25Z"></path><path d="M5 1.75C5 .784 5.784 0 6.75 0h7.5C15.216 0 16 .784 16 1.75v7.5A1.75 1.75 0 0 1 14.25 11h-7.5A1.75 1.75 0 0 1 5 9.25Zm1.75-.25a.25.25 0 0 0-.25.25v7.5c0 .138.112.25.25.25h7.5a.25.25 0 0 0 .25-.25v-7.5a.25.25 0 0 0-.25-.25Z"></path>
</svg>
      <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-check js-clipboard-check-icon color-fg-success d-none m-2">
    <path d="M13.78 4.22a.75.75 0 0 1 0 1.06l-7.25 7.25a.75.75 0 0 1-1.06 0L2.22 9.28a.751.751 0 0 1 .018-1.042.751.751 0 0 1 1.042-.018L6 10.94l6.72-6.72a.75.75 0 0 1 1.06 0Z"></path>
</svg>
    </clipboard-copy>
  </div>
</template>
<template id="snippet-clipboard-copy-button-unpositioned">
  <div class="zeroclipboard-container">
    <clipboard-copy aria-label="Copy" class="ClipboardButton btn btn-invisible js-clipboard-copy m-2 p-0 d-flex flex-justify-center flex-items-center" data-copy-feedback="Copied!" data-tooltip-direction="w">
      <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-copy js-clipboard-copy-icon">
    <path d="M0 6.75C0 5.784.784 5 1.75 5h1.5a.75.75 0 0 1 0 1.5h-1.5a.25.25 0 0 0-.25.25v7.5c0 .138.112.25.25.25h7.5a.25.25 0 0 0 .25-.25v-1.5a.75.75 0 0 1 1.5 0v1.5A1.75 1.75 0 0 1 9.25 16h-7.5A1.75 1.75 0 0 1 0 14.25Z"></path><path d="M5 1.75C5 .784 5.784 0 6.75 0h7.5C15.216 0 16 .784 16 1.75v7.5A1.75 1.75 0 0 1 14.25 11h-7.5A1.75 1.75 0 0 1 5 9.25Zm1.75-.25a.25.25 0 0 0-.25.25v7.5c0 .138.112.25.25.25h7.5a.25.25 0 0 0 .25-.25v-7.5a.25.25 0 0 0-.25-.25Z"></path>
</svg>
      <svg aria-hidden="true" height="16" viewBox="0 0 16 16" version="1.1" width="16" data-view-component="true" class="octicon octicon-check js-clipboard-check-icon color-fg-success d-none">
    <path d="M13.78 4.22a.75.75 0 0 1 0 1.06l-7.25 7.25a.75.75 0 0 1-1.06 0L2.22 9.28a.751.751 0 0 1 .018-1.042.751.751 0 0 1 1.042-.018L6 10.94l6.72-6.72a.75.75 0 0 1 1.06 0Z"></path>
</svg>
    </clipboard-copy>
  </div>
</template>

    </div>

    <div id="js-global-screen-reader-notice" class="sr-only mt-n1" aria-live="polite" aria-atomic="true" ></div>
    <div id="js-global-screen-reader-notice-assertive" class="sr-only mt-n1" aria-live="assertive" aria-atomic="true"></div>
  </body>
</html>