### 2025-03-17 RussoPaper/20250314_RussoManuscript_Fig1Revisions.Rmd

---
title: "Russo Fig1 Replots"
author: "John Santiago"
date: "2025-03-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F}

B = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Russo Fig 1 eclosion.csv")

B = B[c("w1118", "WTloxP", "G85Rhr", 
        "WTcr", "G85Rcr", "A4V.A4V",
        "null26.null26", "WTcr.null26",
        "G85Rcr.null26", "A4V.X16",
        "A4V.G85Rcr")]

mean.B = setNames(rep(0, ncol(B)), colnames(B))
i=1
while(i<=ncol(B)){
  mean.B[i] = mean(na.omit(B[,i]))
  i=i+1
}

B.labels = c("w1118",
             "WTLoxP/WTLoxP", "G85R/G85R(HR)",
             "WT/WT(CR)", "G85R/G85R(CR)",
             "A4V/A4V", "Sod1null/Sod1null", 
             "WT/Sod1null", "G85R(CR)/Sod1null",
             "A4V/X16", "A4V/G85R(CR)")

```

```{r echo = F, fig.width=10, fig.height=7}

par(mar = c(9,4,4,2))

barplot(height = mean.B,
        ylim = c(0, 100),
        col = "black",
        las = 2,
        ylab = "Eclosion rate (%)",
        main = "Eclosion from the pupal case",
        xaxt = "n")
axis(side = 1,
     at = c(.7 + (c(0:10)*1.2)),
     lwd = 3,
     line = 0,
     labels = B.labels,
     xpd = T,
     las = 2)
axis(side = 2,
     at = c(0,20,40,60,80,100),
     lwd = 3,
     line = 0,
     labels = F,
     xpd = T)
lines(x = c(-0.5,13.5),
     y = c(0,0),
     lwd = 3,
     xpd = T)

#abline(h = 100, lty = 2, lwd = 3)

```

```{r echo = F}

C = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 1C pharate_bouton_counts.xlsx - counts.csv")

W1118 = data.frame(Genotype = "W1118",
                   Boutons = na.omit(C$W1118),
                   x = 1)
W1118 = W1118[order(W1118$Boutons),]
W1118$x[c(3)] = .85
W1118$x[c(4)] = 1.15

SILENT = data.frame(Genotype = "SILENT",
                   Boutons = na.omit(C$SILENT),
                   x = 2)
SILENT = SILENT[order(SILENT$Boutons),]
SILENT$x[c(3)] = 1.85
SILENT$x[c(4)] = 2.15

G85Rcr = data.frame(Genotype = "G85Rcr",
                   Boutons = na.omit(C$G85Rcr),
                   x = 3)
G85Rcr = G85Rcr[order(G85Rcr$Boutons),]
G85Rcr$x[c(4,10)] = 2.875
G85Rcr$x[c(5,9)] = 3.125
G85Rcr$x[c(1)] = 2.925
G85Rcr$x[c(2)] = 3.075
G85Rcr$x[c(6)] = 2.75
G85Rcr$x[c(7)] = 3.25

A4V = data.frame(Genotype = "A4V",
                   Boutons = na.omit(C$A4V),
                   x = 4)
A4V = A4V[order(A4V$Boutons),]
A4V$x[c(4, 8)] = 3.875
A4V$x[c(5, 9)] = 4.125
A4V$x[c(6)] = 3.75
A4V$x[c(7)] = 4.25
A4V$x[c(1)] = 3.925
A4V$x[c(2)] = 4.075

C = rbind(W1118, rbind(SILENT,
                       rbind(G85Rcr, A4V)))

C$Genotype = factor(C$Genotype, levels = c("W1118", "SILENT", "G85Rcr", "A4V"))

C.labels = c("w1118", "WT(CR)", "G85R", "A4V")

hex = function(color, alpha){
  rgbvals = col2rgb(color)
  rgb2hex = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
  return(rgb2hex)
}

```

```{r echo = F, fig.width=10, fig.height=7}

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0, 150),
     xlim = c(.5, 4.5),
     xaxt = "n",
     ylab="Bouton Number",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = "Pharate muscle VM NMJ bouton counts")

boxplot(C$Boutons~C$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0, 150),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = c(hex("navy", .25),
                hex("skyblue", .25),
                hex('orange', .25),
                hex("deeppink", .25)),
        add = T, 
        xaxt = "n",
        frame = F)

points(x=as.numeric(C$x),
       y=C$Boutons,
       cex=2.25,
       pch=21,
       bg=c('navy',"skyblue", 'orange', "deeppink")[as.numeric(C$Genotype)])

axis(side = 1,
     at = c(1:4),
     lwd = 3,
     line = -.9,
     labels = C.labels,
     xpd = T)

axis(side = 2,
     at = c(0, 50, 100, 150),
     lwd = 3,
     line = 0,
     labels = F,
     xpd = T)

lines(x = c(0.35,4.5),
     y = c(0,0),
     lwd = 3,
     xpd = T)

lines(x = c(1, 4),
     y = c(140,140),
     lwd = 3)

lines(x = c(3, 3),
     y = c(60,140),
     lwd = 3)

lines(x = c(1, 1),
     y = c(120,140),
     lwd = 3)

lines(x = c(2, 2),
     y = c(120,140),
     lwd = 3)

lines(x = c(4, 4),
     y = c(120,140),
     lwd = 3)

text(x = 3,
     y = 145,
    labels = "*",
    cex = 2)

```

```{r echo = F}

D = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 1D NMJ_quantification_larva.csv")

w1118 = data.frame(Genotype = "w1118",
                   Boutons = na.omit(D$w1118),
                   x = 1)
w1118 = w1118[order(w1118$Boutons),]
w1118$x[c(3, 8, 15)] = .85
w1118$x[c(4, 9, 16)] = 1.15
w1118$x[c(5, 10, 13)] = .7
w1118$x[c(6, 11, 14)] = 1.3

LoxP = data.frame(Genotype = "LoxP",
                   Boutons = na.omit(D$LoxP),
                   x = 2)
LoxP = LoxP[order(LoxP$Boutons),]
LoxP$x[c(3)] = 1.85
LoxP$x[c(4)] = 2.15
LoxP$x[c(5)] = 1.7
LoxP$x[c(6)] = 2.3

G85R_HR = data.frame(Genotype = "G85R_HR",
                   Boutons = na.omit(D$G85R_HR),
                   x = 3)
G85R_HR = G85R_HR[order(G85R_HR$Boutons),]
G85R_HR$x[c(2, 5, 12)] = 2.85
G85R_HR$x[c(3, 6, 13)] = 3.15
G85R_HR$x[c(7, 10)] = 2.7
G85R_HR$x[c(8, 11)] = 3.3

silent = data.frame(Genotype = "silent",
                   Boutons = na.omit(D$silent),
                   x = 4)
silent = silent[order(silent$Boutons),]
silent$x[c(2, 5, 12, 14, 16)] = 3.85
silent$x[c(3, 6, 11, 15, 17)] = 4.15
silent$x[c(7, 10)] = 3.7
silent$x[c(8, 9)] = 4.3

G85R_WG = data.frame(Genotype = "G85R_WG",
                   Boutons = na.omit(D$G85R_WG),
                   x = 5)
G85R_WG = G85R_WG[order(G85R_WG$Boutons),]
G85R_WG$x[c(2)] = 4.85
G85R_WG$x[c(3)] = 5.15
G85R_WG$x[c(4)] = 4.7
G85R_WG$x[c(5)] = 5.3

A4V = data.frame(Genotype = "A4V",
                   Boutons = na.omit(D$A4V),
                   x = 6)
A4V = A4V[order(A4V$Boutons),]
A4V$x[c(1, 3)] = 5.85
A4V$x[c(2, 4)] = 6.15

D = rbind(w1118, rbind(LoxP,
    rbind(G85R_HR, 
    rbind(silent,
    rbind(G85R_WG, A4V)))))

D$Genotype = factor(D$Genotype, levels = c("w1118", "LoxP", "G85R_HR", "silent", "G85R_WG", "A4V"))

D.labels = c("w1118", "WTLoxP", "G85R(HR)", "WT(CR)", "G85R", "A4V")

```

```{r echo = F, fig.width=10, fig.height=7}

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0, 150),
     xlim = c(.5, 6.5),
     xaxt = "n",
     ylab="Bouton Number",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = "Larval muscle 6/7 NMJ bouton counts")

boxplot(D$Boutons~D$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0, 150),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = c(hex("navy", .25),
                hex("gold2", .25),
                hex("turquoise4", .25),
                hex("skyblue", .25),
                hex('orange', .25),
                hex("deeppink", .25)),
        add = T, 
        xaxt = "n",
        frame = F)

points(x=as.numeric(D$x),
       y=D$Boutons,
       cex=2.25,
       pch=21,
       bg=c('navy', "gold2", "turquoise4", 
            "skyblue", 'orange', "deeppink")[as.numeric(D$Genotype)])

axis(side = 1,
     at = c(1:6),
     lwd = 3,
     line = -.9,
     labels = D.labels,
     xpd = T)

axis(side = 2,
     at = c(0, 50, 100, 150),
     lwd = 3,
     line = 0,
     labels = F,
     xpd = T)

lines(x = c(0.25,6.5),
     y = c(0,0),
     lwd = 3,
     xpd = T)

```

```{r echo = F, fig.width=10, fig.height=7}

E = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/Figure 1E LarvalLocomotion.csv")

E = E[c("w1118", "WTLoxP.WTLoxP", "G85R.G85R.HR.", "WT.WT",
        "G85R.G85R.CR.", "A4V.A4V",
        "A4V.X16", "A4V.G85R.CR.")]

E.labels = c("w1118", "WTLoxP", "G85R(HR)", 
             "WT(CR)", "G85R", "A4V",
             "A4V/X16", "A4V/G85R")

n.E = setNames(rep(0, ncol(E)), colnames(E))
mean.E = n.E
sd.E = n.E

i=1
while(i<= ncol(E)){
  mean.E[i] = mean(na.omit(E[,i]))
  sd.E[i] = sd(na.omit(E[,i]))
  n.E[i] = (length(na.omit(E[,i])))
  i=i+1
}

stderr.E = sd.E/(sqrt(n.E))

bar.colors = c(hex("navy", .75),
                hex("gold2", .75),
                hex("turquoise4", .75),
                hex("skyblue", .75),
                hex('orange', .75),
                hex("deeppink", .75),
                hex('grey', .75),
                hex("darkgreen", .75))

border.colors = c('navy', "gold2", "turquoise4",
                  "skyblue", 'orange', "deeppink", 
                  "grey", "green4")

tick.x = c(.7 + (c(0:7)*1.2))

par(mar = c(8,4,4,2), lwd = 3)

barplot(height = mean.E,
        ylim = c(0, 80),
        col = border.colors,
        las = 2,
        ylab = "Distance traveled (mm)",
        main = "Larval locomotion",
        border = border.colors,
        xaxt = "n")

axis(side = 1,
     at = tick.x,
     lwd = 3,
     line = 0,
     labels = E.labels,
     xpd = T)

axis(side = 2,
     at = c(0,20,40,60,80),
     lwd = 3,
     line = 0,
     labels = F,
     xpd = T)
lines(x = c(-0.2,13.5),
     y = c(0,0),
     lwd = 3,
     xpd = T)

i = 1
while(i <= length(stderr.E)){
  lines(x = tick.x[c(i,i)],
        y = c(mean.E[i], (mean.E[i] + stderr.E[i])),
        lwd = 3,
        col = border.colors[i])

  lines(x = c((tick.x[i] - .25), (tick.x[i] + .25)),
        y = c((mean.E[i] + stderr.E[i]), (mean.E[i] + stderr.E[i])),
        lwd = 3,
        col = border.colors[i])

  i = i +1
}

text(x = tick.x[3],
     y = (mean.E[3] + stderr.E[3]) + 5,
    labels = "*",
    cex = 2)

```