### 2025-03-24 GeneralDataFiles/CompleteMetabolicNetwork.Rmd
fixed edge

---
title: "Full Metabolic Network"
author: "John Santiago"
date: "2025-03-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F}

library(visNetwork)

git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/refs/heads/master/GeneralDataFiles/"

G85R.geneFC = read.csv(paste0(git.dir, "TKT.EdgeR.FCTable.csv"), row.names = 1)
G85R.metabFC = read.csv(paste0(git.dir, "MetaboliteFCs.csv"), row.names = 1)

raw.nodes = read.csv(paste0(git.dir, "combined.nodes.csv"), row.names = 1)
raw.edges = read.csv(paste0(git.dir, "combined.edges.csv"))

temp = colorRampPalette(c('royalblue',"dodgerblue",'red', "firebrick"))(401)

```

```{r echo = F}

nodes = raw.nodes
edges = raw.edges

nodes[nodes$shape == "circle", "size"] = 20
nodes$shape = 'dot'

node.colors = G85R.metabFC[nodes$data.id, "GR.CxOE"]
node.colors = temp[(200*node.colors/max(abs(na.exclude(node.colors)))) + 201]
node.colors[na.action(na.exclude(node.colors))] = "white"

nodes$color.background = node.colors

```

```{r echo = F}

visNetwork(nodes, edges)%>%
  visNodes(physics = F)%>%
  visLayout(hierarchical=F, improvedLayout=T)%>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = TRUE)%>%
  visEdges(arrows = visedges$arrows)

```