### 2025-04-11 GeneralDataFiles/CompleteMetabolicNetwork.Rmd
Updated Russo Data

---
title: "Full Metabolic Network"
author: "John Santiago"
date: "2025-03-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F}

library(visNetwork)

git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/refs/heads/master/GeneralDataFiles/"

GeneIDKey = read.csv(paste0(git.dir, "GeneIDKey.csv"), row.names = 1)
G85R.geneFC = read.csv(paste0(git.dir, "TKT.EdgeR.FCTable.csv"), row.names = 1)
G85R.metabFC = read.csv(paste0(git.dir, "MetaboliteFCs.csv"), row.names = 1)
G85R.metabFC[,2:ncol(G85R.metabFC)] = log2(G85R.metabFC[,2:ncol(G85R.metabFC)])

raw.nodes = read.csv(paste0(git.dir, "combined.nodes.csv"), row.names = 1)
raw.edges = read.csv(paste0(git.dir, "combined.edges.csv"))

temp = colorRampPalette(c("dodgerblue", "grey",'red'))(401)

```

```{r echo = F}
##"GRxWT.C"  "GRxWT.Df" "GRxWT.OE" "GR.CxDf" "WT.CxDf"  "GR.CxOE"  "WT.CxOE"
metab = "GR.CxDf"

enzyme = "GRF.CxDf"

nodes = raw.nodes
edges = raw.edges

nodes[nodes$shape == "dot", "label"] = NA

nodes[nodes$shape == "circle", "size"] = 20
nodes$shape = 'dot'

node.colors = (G85R.metabFC[nodes$data.id, metab])
node.colors = temp[(200*node.colors/max(abs(na.exclude(node.colors)))) + 201]
node.colors[na.action(na.exclude(node.colors))] = "white"

nodes$color.background = node.colors

nodes$title = paste0(nodes$id,
                     "<br>FC: ", 
                     signif((G85R.metabFC[nodes$data.id, metab]), 3))

edges$title = paste0(GeneIDKey[edges$FBgn, "Symbol"],
                     "<br>FC: ", 
                     signif(G85R.geneFC[edges$FBgn,
                                 enzyme],3))

edges.colors = G85R.geneFC[edges$FBgn, enzyme]
edges$width = abs(edges.colors) * 10

edges.colors = temp[(200*edges.colors/max(abs(na.exclude(edges.colors)))) + 202]
edges.colors[na.action(na.exclude(edges.colors))] = "white"

edges$color = edges.colors

#edges = edges[c(224,225, 226,227),]
#edges$title = c("A","B","C", "D")

```

```{r echo = F}

visNetwork(nodes, edges)%>%
  visNodes(physics = F)%>%
  #visLayout(hierarchical=F, improvedLayout=T)%>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = T)%>%
  visEdges(arrows = visedges$arrows)

```

```{r echo = F}

nodes = raw.nodes
edges = raw.edges

nodes[nodes$shape == "circle", "size"] = 20
nodes$shape = "dot"

visNetwork(nodes, edges)%>%
  visNodes(physics = F)

```