### 2025-05-14 RussoPaper/Prelim_UBLegMorphologyDataAnalysis.Rmd

---
title: "Jayden Stats Homework"
author: "John Santiago"
date: "2025-04-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F, warning=FALSE, message=FALSE}

library(kableExtra)
library(gtools)
library(gplots)

temp = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/UBLegMorphologyData.csv")

temp$Genotype[temp$Genotype == "WT"] = "WT(cr)"
temp$Genotype[temp$Genotype == "G85R"] = "G85R(cr)"

temp$Genotype = factor(temp$Genotype, levels = unique(temp$Genotype))

temp.titles = c("Is the leg short?",
                "Is there scrunching or curving?",
                "Is the axonal bundle straight?",
                "Does the axonal bundle split or are there chunks missing?",
                "Does the branching look organized?",
                "Is the branching following a pattern tiling?",
                "Is there branching in all aspects of the femur?",
                "Is the mitochonria in the entire axonal bundle?",
                "Do all branches have mitochondria?")

std.colors = setNames(c('navy', "gold2", "turquoise4", "royalblue",
                        'orange', "deeppink", "grey", "green4"),
                      c("w1118", "WT(LoxP)", "G85R(hr)", "WT(cr)",
                        "G85R(cr)", "A4V", "A4V/X16", "A4V/G85R(cr)"))

hex = function(color, alpha){
  rgb2hex = color
  i = 1
  while(i <= length(color)){
    rgbvals = col2rgb(color[i])
    rgb2hex[i] = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
    i = i+1
  }
  return(rgb2hex)
}

x.stagger.odd = c(0, -.1, .1, -.2, .2, -.3, .3, -.4, .4)
x.stagger.even = c(-.05, .05, -.15, .15, -.25, .25, -.35, .35)

```

```{r echo = F}
UB.barplot = function(i){
  ##bar temp data
  btd = temp[,i]

  WT = factor(temp[temp$Genotype == "WT(cr)", i], levels = c(0, 1))
  GR = factor(temp[temp$Genotype == "G85R(cr)", i], levels = c(0, 1))
  fish.data = table(WT,
                    GR)
  fish.data

  pval = as.numeric(fisher.test(fish.data)[1][1])

  bar.temp = data.frame(WT = c(length(btd[temp$Genotype == "WT(cr)" &
                                            btd == 0]),
                               length(btd[temp$Genotype == "WT(cr)" &
                                            btd == 1])),
                        GR = c(length(btd[temp$Genotype == "G85R(cr)" &
                                            btd == 0]),
                               length(btd[temp$Genotype == "G85R(cr)" &
                                            btd == 1])))

  bar.temp = as.matrix(bar.temp)
  btd.names = unique(temp$Genotype)

  opar <- par(lwd = 3)

  par(mar = c(3,4,4,5.5))

  plot(x = NA,
       y = NA,
       type = 'n',
       ylim = c(0, 10),
       xlim = c(1, 6),
       xaxt = "n",
       ylab= colnames(temp)[i],
       yaxt = "n",
       xlab = NA,
       frame = F,
       main = temp.titles[i-5])

  grid(nx=NA, ny=NULL, col = "grey50", lwd = .75)

  barplot(height = bar.temp,
          ylim = c(0, 10),
          beside = T,
          col = "white",
          las = 2,
          ylab= colnames(temp)[i],
          main = colnames(temp)[i],
          xaxt = "n",
          yaxt = "n",
          add = T)

  barplot(height = bar.temp,
          ylim = c(0, 10),
          beside = T,
          col = hex(c("firebrick", "royalblue"), .5),
          las = 2,
          ylab= colnames(temp)[i],
          main = colnames(temp)[i],
          xaxt = "n",
          yaxt = "n",
          add = T)

  axis(side = 1,
       at = c(2,5),
       lwd = 0,
       line = -0.5,
       labels = btd.names,
       xpd = T)

  lines(x = c(.75, 6.25),
       y = c(0,0),
       lwd = 3,
       xpd = T)

  axis(side = 2,
       at = c(0:10),
       lwd = 2,
       line = 0,
       labels = F,
       xpd = T,
       las = 2)

  axis(side = 2,
       at = c(0, 5, 10),
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)

  legend('topright',
         inset=c(-.15,0),
         legend = c("Yes", "No"), 
         fill = hex(c("firebrick", "royalblue"), .5),
         cex = 1,
         bty = 'n',
         pt.cex = 1,
         xpd = T)

  text(x = 6.35,
       y = 8.0,
      label = paste0("pval = ", signif(pval, 3)),
      xpd = T,
      adj = 0)
}

```

```{r echo = F}
i=5

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(400, 650),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= "Femur Length",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = "Femur Length")

axis(side = 1,
     at = c(1,2),
     lwd = 3,
     line = -.45,
     labels = unique(temp$Genotype),
     xpd = T)

lines(x = c(0.4, 2.5),
     y = c(400,400),
     lwd = 3,
     xpd = T)

axis(side = 2,
     at = c(400, 450, 500, 550, 600, 650),
     lwd = 3,
     line = 0,
     labels = T,
     las = 2,
     xpd = T)

boxplot(temp[,i]~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(400, 650),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        yaxt = "n",
        frame = F)
points(x = as.numeric(temp$Genotype),
       y = temp[,i],
       cex = 2.25,
       pch = 21,
       bg= std.colors[temp$Genotype])

```

```{r echo = F, fig.width=10, fig.height=7, include = F}
i=6

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0, 1.25),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= colnames(temp)[i],
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = colnames(temp)[i])

axis(side = 1,
     at = c(1,2),
     lwd = 3,
     line = -.9,
     labels = unique(temp$Genotype),
     xpd = T)

lines(x = c(0.5, 2.5),
     y = c(0,0),
     lwd = 3,
     xpd = T)

axis(side = 2,
     at = c(0, .5, 1),
     lwd = 3,
     line = 0,
     labels = F,
     xpd = T)

boxplot(temp[,i]~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0, 1.25),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        yaxt = "n",
        frame = F)

```

```{r echo = F, warning=F}

i = 6

UB.barplot(i)

```

```{r echo = F, fig.width=10, fig.height=7, include = F}
i=7

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(-0.25, 1.25),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= colnames(temp)[i],
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = colnames(temp)[i])

boxplot(temp[,i]~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(-0.25, 1.25),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        frame = F)

xpos = as.numeric(temp$Genotype)
ones = temp[,i] == 1
zeros = temp[,i] == 0

if(odd(length(xpos[ones & xpos == 1]))){
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.odd[1:length(xpos[ones & xpos == 1])]
}else{
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.even[1:length(xpos[ones & xpos == 1])]  
}

if(odd(length(xpos[ones & xpos == 2]))){
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.odd[1:length(xpos[ones & xpos == 2])]
}else{
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.even[1:length(xpos[ones & xpos == 2])] 
}

if(odd(length(xpos[zeros & xpos == 1]))){
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.odd[1:length(xpos[zeros & xpos == 1])]
}else{
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.even[1:length(xpos[zeros & xpos == 1])]  
}

if(odd(length(xpos[zeros & xpos == 2]))){
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.odd[1:length(xpos[zeros & xpos == 2])]
}else{
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.even[1:length(xpos[zeros & xpos == 2])] 
}

points(x = xpos,
       y = temp[,i],
       cex = 2.25,
       pch = 21,
       bg= std.colors[temp$Genotype])

```

```{r echo = F, warning=F}

i = 7

UB.barplot(i)

```

```{r echo = F}

i=15

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0.5, 2.5),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= colnames(temp)[i],
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = "Leg Length (Total)")

boxplot(temp[,i]~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0.5, 2.5),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        yaxt = "n",
        frame = F)

xpos = as.numeric(temp$Genotype)
ones = temp[,i] == 1
zeros = temp[,i] == 2

pval = t.test(x = temp[1:9, i],
              y = temp[10:18, i],
              paired = F)[[3]]

if(odd(length(xpos[ones & xpos == 1]))){
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.odd[1:length(xpos[ones & xpos == 1])]
}else{
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.even[1:length(xpos[ones & xpos == 1])]  
}

if(odd(length(xpos[ones & xpos == 2]))){
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.odd[1:length(xpos[ones & xpos == 2])]
}else{
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.even[1:length(xpos[ones & xpos == 2])] 
}

if(odd(length(xpos[zeros & xpos == 1]))){
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.odd[1:length(xpos[zeros & xpos == 1])]
}else{
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.even[1:length(xpos[zeros & xpos == 1])]  
}

if(odd(length(xpos[zeros & xpos == 2]))){
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.odd[1:length(xpos[zeros & xpos == 2])]
}else{
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.even[1:length(xpos[zeros & xpos == 2])] 
}

points(x = xpos,
       y = temp[,i],
       cex = 2.25,
       pch = 21,
       bg= std.colors[temp$Genotype])

  axis(side = 1,
       at = c(1,2),
       lwd = 3,
       line = -.45,
       labels = unique(temp$Genotype),
       xpd = T)

    lines(x = c(0.4, 2.5),
       y = c(0.5,0.5),
       lwd = 3,
       xpd = T)

      axis(side = 2,
       at = c(0.5, 1, 1.5, 2, 2.5),
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)

  text(x = 2.25,
       y = 2.5,
       label = paste0("pval = ", signif(pval, 3)))

```

```{r echo = F, fig.width=10, fig.height=7, include = F}
i=9

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(-0.25, 1.25),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= colnames(temp)[i],
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = colnames(temp)[i])

boxplot(temp[,i]~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(-0.25, 1.25),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        frame = F)

xpos = as.numeric(temp$Genotype)
ones = temp[,i] == 1
zeros = temp[,i] == 0

if(odd(length(xpos[ones & xpos == 1]))){
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.odd[1:length(xpos[ones & xpos == 1])]
}else{
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.even[1:length(xpos[ones & xpos == 1])]  
}

if(odd(length(xpos[ones & xpos == 2]))){
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.odd[1:length(xpos[ones & xpos == 2])]
}else{
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.even[1:length(xpos[ones & xpos == 2])] 
}

if(odd(length(xpos[zeros & xpos == 1]))){
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.odd[1:length(xpos[zeros & xpos == 1])]
}else{
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.even[1:length(xpos[zeros & xpos == 1])]  
}

if(odd(length(xpos[zeros & xpos == 2]))){
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.odd[1:length(xpos[zeros & xpos == 2])]
}else{
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.even[1:length(xpos[zeros & xpos == 2])] 
}

points(x = xpos,
       y = temp[,i],
       cex = 2.25,
       pch = 21,
       bg= std.colors[temp$Genotype])

```

```{r echo = F, warning=F}

i = 8

UB.barplot(i)

```

```{r echo = F, fig.width=10, fig.height=7, include = F}
i=10

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(-0.25, 1.25),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= colnames(temp)[i],
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = colnames(temp)[i])

boxplot(temp[,i]~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(-0.25, 1.25),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        frame = F)

xpos = as.numeric(temp$Genotype)
ones = temp[,i] == 1
zeros = temp[,i] == 0

if(odd(length(xpos[ones & xpos == 1]))){
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.odd[1:length(xpos[ones & xpos == 1])]
}else{
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.even[1:length(xpos[ones & xpos == 1])]  
}

if(odd(length(xpos[ones & xpos == 2]))){
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.odd[1:length(xpos[ones & xpos == 2])]
}else{
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.even[1:length(xpos[ones & xpos == 2])] 
}

if(odd(length(xpos[zeros & xpos == 1]))){
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.odd[1:length(xpos[zeros & xpos == 1])]
}else{
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.even[1:length(xpos[zeros & xpos == 1])]  
}

if(odd(length(xpos[zeros & xpos == 2]))){
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.odd[1:length(xpos[zeros & xpos == 2])]
}else{
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.even[1:length(xpos[zeros & xpos == 2])] 
}

points(x = xpos,
       y = temp[,i],
       cex = 2.25,
       pch = 21,
       bg= std.colors[temp$Genotype])

```

```{r echo = F, warning=F}

i = 9

UB.barplot(i)

```

```{r echo = F}

i=16

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(-0.5, 2.5),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= colnames(temp)[i],
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = "Axonal Bundle (Total)")

boxplot(temp[,i]~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(-0.5, 2.5),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        yaxt = "n",
        frame = F)

xpos = as.numeric(temp$Genotype)
ones = temp[,i] == 1
zeros = temp[,i] == 0
twos = temp[,i] == 2

pval = t.test(x = temp[1:9, i],
              y = temp[10:18, i],
              paired = F)[[3]]

if(odd(length(xpos[twos & xpos == 1]))){
  xpos[twos & xpos == 1] = xpos[twos & xpos == 1] + x.stagger.odd[1:length(xpos[twos & xpos == 1])]
}else{
  xpos[twos & xpos == 1] = xpos[twos & xpos == 1] + x.stagger.even[1:length(xpos[twos & xpos == 1])]  
}

if(odd(length(xpos[twos & xpos == 2]))){
  xpos[twos & xpos == 2] = xpos[twos & xpos == 2] + x.stagger.odd[1:length(xpos[twos & xpos == 2])]
}else{
  xpos[twos & xpos == 2] = xpos[twos & xpos == 2] + x.stagger.even[1:length(xpos[twos & xpos == 2])] 
}

if(odd(length(xpos[ones & xpos == 1]))){
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.odd[1:length(xpos[ones & xpos == 1])]
}else{
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.even[1:length(xpos[ones & xpos == 1])]  
}

if(odd(length(xpos[ones & xpos == 2]))){
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.odd[1:length(xpos[ones & xpos == 2])]
}else{
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.even[1:length(xpos[ones & xpos == 2])] 
}

if(odd(length(xpos[zeros & xpos == 1]))){
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.odd[1:length(xpos[zeros & xpos == 1])]
}else{
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.even[1:length(xpos[zeros & xpos == 1])]  
}

if(odd(length(xpos[zeros & xpos == 2]))){
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.odd[1:length(xpos[zeros & xpos == 2])]
}else{
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.even[1:length(xpos[zeros & xpos == 2])] 
}

points(x = xpos,
       y = temp[,i],
       cex = 2.25,
       pch = 21,
       bg= std.colors[temp$Genotype])

  axis(side = 1,
       at = c(1,2),
       lwd = 3,
       line = -.45,
       labels = unique(temp$Genotype),
       xpd = T)

    lines(x = c(0.4, 2.5),
       y = c(-0.5,-0.5),
       lwd = 3,
       xpd = T)

      axis(side = 2,
       at = c(-0.5, 0, 0.5, 1, 1.5, 2, 2.5),
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)

  text(x = 2.25,
       y = 2.5,
       label = paste0("pval = ", signif(pval, 3)))

```

```{r echo = F, warning=F}

i = 10

UB.barplot(i)

```

```{r echo = F, fig.width=10, fig.height=7, include = F}
i=11

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(-0.25, 1.25),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= colnames(temp)[i],
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = colnames(temp)[i])

boxplot(temp[,i]~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(-0.25, 1.25),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        frame = F)

xpos = as.numeric(temp$Genotype)
ones = temp[,i] == 1
zeros = temp[,i] == 0

if(odd(length(xpos[ones & xpos == 1]))){
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.odd[1:length(xpos[ones & xpos == 1])]
}else{
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.even[1:length(xpos[ones & xpos == 1])]  
}

if(odd(length(xpos[ones & xpos == 2]))){
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.odd[1:length(xpos[ones & xpos == 2])]
}else{
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.even[1:length(xpos[ones & xpos == 2])] 
}

if(odd(length(xpos[zeros & xpos == 1]))){
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.odd[1:length(xpos[zeros & xpos == 1])]
}else{
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.even[1:length(xpos[zeros & xpos == 1])]  
}

if(odd(length(xpos[zeros & xpos == 2]))){
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.odd[1:length(xpos[zeros & xpos == 2])]
}else{
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.even[1:length(xpos[zeros & xpos == 2])] 
}

points(x = xpos,
       y = temp[,i],
       cex = 2.25,
       pch = 21,
       bg= std.colors[temp$Genotype])

```

```{r echo = F, warning=F}

i = 11

UB.barplot(i)

```

```{r echo = F, fig.width=10, fig.height=7, include = F}
i=12

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(-0.25, 1.25),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= colnames(temp)[i],
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = colnames(temp)[i])

boxplot(temp[,i]~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(-0.25, 1.25),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        frame = F)

xpos = as.numeric(temp$Genotype)
ones = temp[,i] == 1
zeros = temp[,i] == 0

if(odd(length(xpos[ones & xpos == 1]))){
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.odd[1:length(xpos[ones & xpos == 1])]
}else{
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.even[1:length(xpos[ones & xpos == 1])]  
}

if(odd(length(xpos[ones & xpos == 2]))){
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.odd[1:length(xpos[ones & xpos == 2])]
}else{
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.even[1:length(xpos[ones & xpos == 2])] 
}

if(odd(length(xpos[zeros & xpos == 1]))){
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.odd[1:length(xpos[zeros & xpos == 1])]
}else{
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.even[1:length(xpos[zeros & xpos == 1])]  
}

if(odd(length(xpos[zeros & xpos == 2]))){
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.odd[1:length(xpos[zeros & xpos == 2])]
}else{
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.even[1:length(xpos[zeros & xpos == 2])] 
}

points(x = xpos,
       y = temp[,i],
       cex = 2.25,
       pch = 21,
       bg= std.colors[temp$Genotype])

```

```{r echo = F, warning=F}

i = 12

UB.barplot(i)

```

```{r echo = F}

i=17

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0.5, 3.5),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= colnames(temp)[i],
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = "Axonal Branching (Total)")

boxplot(temp[,i]~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0.5, 3.5),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        yaxt = "n",
        frame = F)

xpos = as.numeric(temp$Genotype)
ones = temp[,i] == 1
zeros = temp[,i] == 3
twos = temp[,i] == 2

pval = t.test(x = temp[1:9, i],
              y = temp[10:18, i],
              paired = F)[[3]]

if(odd(length(xpos[twos & xpos == 1]))){
  xpos[twos & xpos == 1] = xpos[twos & xpos == 1] + x.stagger.odd[1:length(xpos[twos & xpos == 1])]
}else{
  xpos[twos & xpos == 1] = xpos[twos & xpos == 1] + x.stagger.even[1:length(xpos[twos & xpos == 1])]  
}

if(odd(length(xpos[twos & xpos == 2]))){
  xpos[twos & xpos == 2] = xpos[twos & xpos == 2] + x.stagger.odd[1:length(xpos[twos & xpos == 2])]
}else{
  xpos[twos & xpos == 2] = xpos[twos & xpos == 2] + x.stagger.even[1:length(xpos[twos & xpos == 2])] 
}

if(odd(length(xpos[ones & xpos == 1]))){
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.odd[1:length(xpos[ones & xpos == 1])]
}else{
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.even[1:length(xpos[ones & xpos == 1])]  
}

if(odd(length(xpos[ones & xpos == 2]))){
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.odd[1:length(xpos[ones & xpos == 2])]
}else{
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.even[1:length(xpos[ones & xpos == 2])] 
}

if(odd(length(xpos[zeros & xpos == 1]))){
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.odd[1:length(xpos[zeros & xpos == 1])]
}else{
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.even[1:length(xpos[zeros & xpos == 1])]  
}

if(odd(length(xpos[zeros & xpos == 2]))){
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.odd[1:length(xpos[zeros & xpos == 2])]
}else{
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.even[1:length(xpos[zeros & xpos == 2])] 
}

points(x = xpos,
       y = temp[,i],
       cex = 2.25,
       pch = 21,
       bg= std.colors[temp$Genotype])

  axis(side = 1,
       at = c(1,2),
       lwd = 3,
       line = -.45,
       labels = unique(temp$Genotype),
       xpd = T)

    lines(x = c(0.4, 2.5),
       y = c(0.5,0.5),
       lwd = 3,
       xpd = T)

      axis(side = 2,
       at = c(0.5, 1, 1.5, 2, 2.5, 3, 3.5),
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)

  text(x = 2.25,
       y = 3.5,
       label = paste0("pval = ", signif(pval, 3)))

```

```{r echo = F, fig.width=10, fig.height=7, include = F}
i=13

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(-0.25, 1.25),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= colnames(temp)[i],
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = colnames(temp)[i])

boxplot(temp[,i]~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(-0.25, 1.25),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        frame = F)

xpos = as.numeric(temp$Genotype)
ones = temp[,i] == 1
zeros = temp[,i] == 0

if(odd(length(xpos[ones & xpos == 1]))){
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.odd[1:length(xpos[ones & xpos == 1])]
}else{
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.even[1:length(xpos[ones & xpos == 1])]  
}

if(odd(length(xpos[ones & xpos == 2]))){
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.odd[1:length(xpos[ones & xpos == 2])]
}else{
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.even[1:length(xpos[ones & xpos == 2])] 
}

if(odd(length(xpos[zeros & xpos == 1]))){
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.odd[1:length(xpos[zeros & xpos == 1])]
}else{
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.even[1:length(xpos[zeros & xpos == 1])]  
}

if(odd(length(xpos[zeros & xpos == 2]))){
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.odd[1:length(xpos[zeros & xpos == 2])]
}else{
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.even[1:length(xpos[zeros & xpos == 2])] 
}

points(x = xpos,
       y = temp[,i],
       cex = 2.25,
       pch = 21,
       bg= std.colors[temp$Genotype])

```

```{r echo = F, warning=F}

i = 13

UB.barplot(i)

```

```{r echo = F, fig.width=10, fig.height=7, include = F}
i=14

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(-0.25, 1.25),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= colnames(temp)[i],
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = colnames(temp)[i])

boxplot(temp[,i]~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(-0.25, 1.25),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        frame = F)

xpos = as.numeric(temp$Genotype)
ones = temp[,i] == 1
zeros = temp[,i] == 0

if(odd(length(xpos[ones & xpos == 1]))){
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.odd[1:length(xpos[ones & xpos == 1])]
}else{
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.even[1:length(xpos[ones & xpos == 1])]  
}

if(odd(length(xpos[ones & xpos == 2]))){
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.odd[1:length(xpos[ones & xpos == 2])]
}else{
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.even[1:length(xpos[ones & xpos == 2])] 
}

if(odd(length(xpos[zeros & xpos == 1]))){
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.odd[1:length(xpos[zeros & xpos == 1])]
}else{
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.even[1:length(xpos[zeros & xpos == 1])]  
}

if(odd(length(xpos[zeros & xpos == 2]))){
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.odd[1:length(xpos[zeros & xpos == 2])]
}else{
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.even[1:length(xpos[zeros & xpos == 2])] 
}

points(x = xpos,
       y = temp[,i],
       cex = 2.25,
       pch = 21,
       bg= std.colors[temp$Genotype])

```

```{r echo = F, warning=F}

i = 14

UB.barplot(i)

```

```{r echo = F}

i=18

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(-0.5, 2.5),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= colnames(temp)[i],
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = "Mitochondria (Total)")

boxplot(temp[,i]~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(-0.5, 2.5),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        yaxt = "n",
        frame = F)

xpos = as.numeric(temp$Genotype)
ones = temp[,i] == 1
zeros = temp[,i] == 0
twos = temp[,i] == 2

pval = t.test(x = temp[1:9, i],
              y = temp[10:18, i],
              paired = F)[[3]]

if(odd(length(xpos[twos & xpos == 1]))){
  xpos[twos & xpos == 1] = xpos[twos & xpos == 1] + x.stagger.odd[1:length(xpos[twos & xpos == 1])]
}else{
  xpos[twos & xpos == 1] = xpos[twos & xpos == 1] + x.stagger.even[1:length(xpos[twos & xpos == 1])]  
}

if(odd(length(xpos[twos & xpos == 2]))){
  xpos[twos & xpos == 2] = xpos[twos & xpos == 2] + x.stagger.odd[1:length(xpos[twos & xpos == 2])]
}else{
  xpos[twos & xpos == 2] = xpos[twos & xpos == 2] + x.stagger.even[1:length(xpos[twos & xpos == 2])] 
}

if(odd(length(xpos[ones & xpos == 1]))){
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.odd[1:length(xpos[ones & xpos == 1])]
}else{
  xpos[ones & xpos == 1] = xpos[ones & xpos == 1] + x.stagger.even[1:length(xpos[ones & xpos == 1])]  
}

if(odd(length(xpos[ones & xpos == 2]))){
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.odd[1:length(xpos[ones & xpos == 2])]
}else{
  xpos[ones & xpos == 2] = xpos[ones & xpos == 2] + x.stagger.even[1:length(xpos[ones & xpos == 2])] 
}

if(odd(length(xpos[zeros & xpos == 1]))){
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.odd[1:length(xpos[zeros & xpos == 1])]
}else{
  xpos[zeros & xpos == 1] = xpos[zeros & xpos == 1] + x.stagger.even[1:length(xpos[zeros & xpos == 1])]  
}

if(odd(length(xpos[zeros & xpos == 2]))){
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.odd[1:length(xpos[zeros & xpos == 2])]
}else{
  xpos[zeros & xpos == 2] = xpos[zeros & xpos == 2] + x.stagger.even[1:length(xpos[zeros & xpos == 2])] 
}

points(x = xpos,
       y = temp[,i],
       cex = 2.25,
       pch = 21,
       bg= std.colors[temp$Genotype])

  axis(side = 1,
       at = c(1,2),
       lwd = 3,
       line = -.45,
       labels = unique(temp$Genotype),
       xpd = T)

    lines(x = c(0.4, 2.5),
       y = c(-0.5,-0.5),
       lwd = 3,
       xpd = T)

      axis(side = 2,
       at = c(-0.5, 0, 0.5, 1, 1.5, 2, 2.5),
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)

  text(x = 2.25,
       y = 2.5,
       label = paste0("pval = ", signif(pval, 3)))

```

```{r echo = F}

i=19

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0, 10),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= colnames(temp)[i],
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = "Overall Morphology Score")

grid(nx=NA, ny=NULL, col = "grey50", lwd = .75)

boxplot(temp[,i]~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0, 10),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        yaxt = "n",
        frame = F)

xpos = as.numeric(temp$Genotype)
ones = temp[,i] == 1
zeros = temp[,i] == 0
twos = temp[,i] == 2

pval = t.test(x = temp[1:9, i],
              y = temp[10:18, i],
              paired = F)[[3]]

xpos = c(-.05, -.05, -.05, .05, .05, -.15, .15, 0, .05,
         0, -.05, -.05, .05, 0, .05, -.05, .05, 0) + as.numeric(temp$Genotype)

points(x = xpos,
       y = temp[,i],
       cex = 2.25,
       pch = 21,
       bg= std.colors[temp$Genotype])

  axis(side = 1,
       at = c(1,2),
       lwd = 3,
       line = -.45,
       labels = unique(temp$Genotype),
       xpd = T)

    lines(x = c(0.4, 2.5),
       y = c(0,0),
       lwd = 3,
       xpd = T)

      axis(side = 2,
       at = c(0, 5, 10),
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)

  text(x = 2.25,
       y = 9.5,
       label = paste0("pval = ", signif(pval, 3)))

```

```{r echo = F}

i=20

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0, 4),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= colnames(temp)[i],
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = "Supervised Morphology Score\n(Axonal Bundle and Mitochondria)")

grid(nx=NA, ny=NULL, col = "grey50", lwd = .75)

boxplot(temp[,i]~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0, 4),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        yaxt = "n",
        frame = F)

  axis(side = 1,
       at = c(1,2),
       lwd = 3,
       line = -.45,
       labels = unique(temp$Genotype),
       xpd = T)

    lines(x = c(0.4, 2.5),
       y = c(0,0),
       lwd = 3,
       xpd = T)

pval = t.test(x = temp[1:9, i],
              y = temp[10:18, i],
              paired = F)[[3]]

xpos = c(0, -.1, 0, .1, -.1, .1, -.2, 0, .2,
         -.05, .05, -.15, -.05, -.05, .15, 0, .05, .05) + as.numeric(temp$Genotype)

points(x = xpos,
       y = temp[,i],
       cex = 2.25,
       pch = 21,
       bg= std.colors[temp$Genotype])

      axis(side = 2,
       at = c(0:4),
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)

  text(x = 2.25,
       y = 3.75,
       label = paste0("pval = ", signif(pval, 3)))

```

```{r echo = F, include = F}

heat.data = temp[,6:14]
row.names(heat.data) = paste0(temp$Genotype,temp$Animal..)
heat.data = as.matrix(t(heat.data))

heat.colors = colorRampPalette(c("firebrick", "deepskyblue"))
sample.colors = c(rep("darkgreen", 9), rep("gold2", 9))
category.colors = c(rep("darkgreen", 2), 
                    rep("navy", 2),
                    rep("deeppink", 3), 
                    rep("gold2", 2))

heatmap.2(heat.data,
          Colv = T,
          Rowv = T,
          margins = c(7, 15),
          trace = "none",
          ColSideColors = sample.colors,
          RowSideColors = category.colors,
          col = heat.colors)
```

```{r echo = F}

# Assume:
# - data_matrix: 18 x 9 matrix
# - group_labels: factor of length 18 with two levels (e.g., "WT", "KO")

find_best_category_combos <- function(data_matrix, group_labels, max_k = 9) {
  n_categories <- ncol(data_matrix)
  category_names <- colnames(data_matrix)

  results <- list()

  for (k in 1:max_k) {
    combs <- combn(n_categories, k, simplify = FALSE)
    for (combo in combs) {
      summed <- rowSums(data_matrix[, combo, drop = FALSE])
      test <- t.test(summed ~ group_labels)
      results[[paste(combo, collapse = ",")]] <- list(
        categories = category_names[combo],
        mean_diff = abs(diff(tapply(summed, group_labels, mean))),
        p_value = test$p.value
      )
    }
  }

  # Convert to data frame
  results_df <- do.call(rbind, lapply(results, as.data.frame))
  results_df$combo <- rownames(results_df)
  results_df <- results_df[order(results_df$p_value), ]
  return(results_df)
}

```

```{r echo = F, warning=F}

data_matrix = t(heat.data)
group_labels = temp$Genotype

temp.combos = find_best_category_combos(data_matrix, group_labels, max_k = 9)

row.names(temp.combos) = c(1:nrow(temp.combos))

best.combo = data_matrix[,c(2,3,4,7,8)]

sum.best.combo = apply(best.combo, MARGIN = 1, sum)

best.pval = t.test(sum.best.combo[1:9], sum.best.combo[10:18])[[3]]

supervised.combo = temp.combos[temp.combos$combo == "3,4,8,9.1",]

```

```{r echo = F}

par(mar = c(5,7,5,2))

plot(x = NA,
     y = NA,
     type = 'n',
     ylim = c(0, 5),
     xlim = c(.5, 2.5),
     xaxt = "n",
     ylab= "Morphology Score",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = "Most Significantly Different Combination of Traits")

grid(nx=NA, ny=NULL, col = "grey50", lwd = .75)

boxplot(sum.best.combo~temp$Genotype,
        xlab="",
        las = 2,
        cex.axis = 1,
        ylim = c(0, 5),
        boxwex = .75,
        boxlwd = 2,
        lwd = 2,
        cex.ticks = 2,
        col = hex(std.colors[unique(temp$Genotype)], .25),
        add = T, 
        xaxt = "n",
        yaxt = "n",
        frame = F)

  axis(side = 1,
       at = c(1,2),
       lwd = 3,
       line = -.45,
       labels = unique(temp$Genotype),
       xpd = T)

    lines(x = c(0.4, 2.5),
       y = c(0,0),
       lwd = 3,
       xpd = T)

pval = t.test(x = sum.best.combo[1:9],
              y = sum.best.combo[10:18],
              paired = F)[[3]]

xpos = c(-.05, 0, -.1, -.1, .05, -.1, .1, .1, .1,
         -.05, 0, -.1, .1, 0, .05, -.2, .2, 0) + as.numeric(temp$Genotype)

points(x = xpos,
       y = sum.best.combo,
       cex = 2.25,
       pch = 21,
       bg= std.colors[temp$Genotype])

      axis(side = 2,
       at = c(0:5),
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)

  text(x = 2.25,
       y = 4.75,
       label = paste0("pval = ", signif(pval, 3)))

```