### 2025-06-13 DeficiencyScreen/Prelim_Analysis.Rmd
updated df data

---
title: "Defficiency Screen Functional Analysis"
author: "John Santiago"
date: "11/28/2020"
output: html_document
---

```{r setup, include=FALSE}
library(rgl)
knitr::opts_chunk$set(echo = FALSE, fig.align="center")
knitr::knit_hooks$set(webgl = hook_webgl)
```

```{r echo = F, include=F}
##Load Libraries

library(org.Dm.eg.db)
##library(goseq)
library(rrvgo)
library(plotly)
```

```{r echo = F, include =F}
##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"

DfKey = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/DfKey.csv"), row.names = 1)
old = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/DeficiencyModifierGenes.csv"),row.names = 1)

GeneIDKey = read.csv(paste0(git.dir,"GeneralDataFiles/GeneIDKey.csv"),row.names = 1)

dfdata = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/20250602_UpdatedModiferData.csv"))
row.names(dfdata) = dfdata$Fbab
dfgenes = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/GenesInDeficiency.csv"))

fulldf = cbind(dfgenes, dfdata[dfgenes$FBab,])
row.names(fulldf) = 1:nrow(fulldf)
fulldf$Modifier[fulldf$Modifier ==  "Supressor"] = "Suppressor"
fulldf$Modifier[fulldf$Modifier ==  "No effect"] = "No Effect"
new=fulldf

olddf = DfKey$Deficiency
newdf = unique(new$Deficiency)
oldfb = old$FBgn
newfb = unique(new$FBgn)
oldsym = old$Symbol
newsym = unique(new$Symbol)

#old = cbind(old, GeneIDKey[row.names(old), c("Start", "End", "CHR")])
old = GeneIDKey[, c("FBgn", "Start", "End", "CHR")]

old.left = apply(old[,c("Start", "End")], MARGIN = 1, min)
old.right = apply(old[,c("Start", "End")], MARGIN = 1, max)
Df.left = apply(DfKey[,c("Start", "End")], MARGIN = 1, min)
Df.right = apply(DfKey[,c("Start", "End")], MARGIN = 1, max)

genesinolddf = list()
i=1
while(i <= nrow(DfKey)){
  temp = na.omit(old[((old.left <= Df.left[i] & old.right >= Df.right[i]) |
           (old.left >= Df.left[i] & old.left <= Df.right[i]) |
           (old.right >= Df.left[i] & old.right <= Df.right[i])) &
            old$CHR == DfKey$Chr[i],])

  genesinolddf[[DfKey$Deficiency[i]]] = temp$FBgn
  i=i+1
}

#new.loc = GeneIDKey[new$FBgn,c("Start","End","CHR")]
new.loc = GeneIDKey[,c("FBgn", "Start","End","CHR")]
new.left = apply(new.loc[,c("Start", "End")], MARGIN = 1, min)
new.right = apply(new.loc[,c("Start", "End")], MARGIN = 1, max)
newDf.left = apply(new[,c("Start", "End")], MARGIN = 1, min)
newDf.right = apply(new[,c("Start", "End")], MARGIN = 1, max)

new.keep = !duplicated(paste0(new$Deficiency, newDf.left, newDf.right))
newDf.left = newDf.left[new.keep]
newDf.right = newDf.right[new.keep]
newDf = new$Deficiency[new.keep]
newChr = new$Chr[new.keep]

genesinnewdf = list()
i=1
while(i <= length(newDf)){
      temp = na.omit(new.loc[((new.left <= newDf.left[i] & new.right >= newDf.right[i]) |
           (new.left >= newDf.left[i] & new.left <= newDf.right[i]) |
           (new.right >= newDf.left[i] & new.right <= newDf.right[i])) &
            new.loc$CHR == newChr[i],])

    genesinnewdf[[newDf[i]]] = temp$FBgn
  i=i+1
}

```

```{r echo = F}

newdfonly = setdiff(newdf, olddf)
olddfonly = setdiff(olddf, newdf)

newfbonly = setdiff(newfb, oldfb)
oldfbonly = setdiff(oldfb, newfb)

newsymonly = setdiff(newsym, oldsym)
oldsymonly = setdiff(oldsym, newsym)

fbinnewdfonly = unique(fulldf[fulldf$Deficiency %in% newdfonly, "FBgn"])
fbinolddfonly = unique(unlist(genesinolddf[olddfonly]))

```

```{r echo = F}

  ##Sample titles as strings. Only fill in up to your number of selected categories
  set1 = "all old FBgn"
  set2 = "all new FBgn"
  set3 = "FBgn in Old Unique Df"
  set4 = "FBgn in New Unique Df"

  ##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
  s1 = oldfb
  s2 = newfb
  s3 = fbinolddfonly
  s4 = fbinnewdfonly

  main = ""

  grid.newpage()
  tempvenn=draw.quad.venn(area1=length(s1),area2=length(s2),area3=length(s3),area4=length(s4), n12=length(intersect(s1,s2)),n13 = length(intersect(s1,s3)),n14 = length(intersect(s1,s4)),n23 = length(intersect(s2,s3)),n24 = length(intersect(s2,s4)),n34 = length(intersect(s3,s4)),n123 =  length(intersect(s1,intersect(s2,s3))),n124 =  length(intersect(s1,intersect(s2,s4))), n134 = length(intersect(s1,intersect(s3,s4))),n234 = length(intersect(s2,intersect(s3,s4))),n1234 = length(intersect(s1,intersect(s2,intersect(s3,s4)))),category =c(set1,set2,set3,set4),fill = c("blue", "red","green","yellow"),cex=2,cat.cex = 1, title = "main")

```

```{r echo = F}
##kegg = readRDS(gzcon(url(paste0(git.dir,"DeficiencyScreen/DataFiles/kegg.symbol2path.RData"))))
##kegg.names = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/KEGG.names.csv"), row.names = 1)

Moderate.GO = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/ModerateGOAnalysis.csv"), row.names = 1)
Strong.GO = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/StrongGOAnalysis.csv"), row.names = 1)
Repress.GO = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/RepressorGOAnalysis.csv"), row.names = 1)
Lethal.GO = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/LethalGOAnalysis.csv"), row.names = 1)
Enhancer.GO = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/EnhancerGOAnalysis.csv"), row.names = 1)

genesingo=as.list(org.Dm.egGO2ALLEGS)

```

```{r echo = F}

##dfbar = data.frame(Enhancer = c(DfKey[DfKey$Modifier == "Enhancer",], dfdata[dfdata$Modifier == "Enhancer",]),
##                   NoEffect = c(DfKey[DfKey$Modifier == "No Effect",], dfdata[dfdata$Modifier == "No Effect",]),
##                   Lethal = c(DfKey[DfKey$Modifier == "Lethal",], dfdata[dfdata$Modifier == "Lethal",]),
##                   c(DfKey[DfKey$Modifier == "Enhancer",], dfdata[dfdata$Modifier == "Enhancer",]),
##                   Moderate = c(DfKey[DfKey$Modifier == "Enhancer",], dfdata[dfdata$Modifier == "Enhancer",]),)

#old = hist(as.numeric(DfKey$MLE))
#new = hist(as.numeric(dfdata$MLE))

hex = function(color, alpha){
  rgb2hex = color
  i = 1
  while(i <= length(color)){
    rgbvals = col2rgb(color[i])
    rgb2hex[i] = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
    i = i+1
  }
  return(rgb2hex)
}

```

```{r echo = F, warning=F}

hist(na.omit(as.numeric(dfdata$MLE)), col = hex('yellow', .25),
     breaks = seq(-4, 7, by = 1),
     xlab = "MLE",
     main = "Old Df MLE data vs New")
hist(na.omit(as.numeric(DfKey$MLE)), add = T, col = hex('blue', .2),
     breaks = seq(-4, 7, by = 1))
legend("topright", legend = c("old data", "new data"), fill = c("blue", "yellow"))
```

```{r include=F, echo=F}

sigs = unique(fulldf[fulldf$Modifier == "Enhancer", "Symbol"])
genes = setNames(rep(0, length(unique(fulldf$Symbol))), unique(fulldf$Symbol))
genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol", method = "Hypergeometric")

GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category
Enhancer.GO = GO.wall
##write.csv(Suppressor.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/SupressorGOAnalysis.csv")

  ##KEGG=goseq(pwf,gene2cat=kegg)
##KEGG=goseq(pwf,gene2cat=kegg, method = "Hypergeometric")

##KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
##row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
##KEGG$Name=kegg.names[row.names(KEGG),1]
##Repress.KEGG = KEGG
##write.csv(Repress.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/RepressorKEGGAnalysis.csv")

##sigs = Df.data[Df.data$Moderate == 1 , "Symbol"]
##genes = setNames(rep(0, nrow(Df.data)), Df.data$Symbol)
##genes[sigs] = 1
##pwf=nullp(genes,"dm3","geneSymbol")
  ##GO.wall=goseq(pwf,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol", method = "Hypergeometric")

##GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
##row.names(GO.wall) = GO.wall$category
##Moderate.GO = GO.wall
##write.csv(Moderate.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/ModerateGOAnalysis.csv")

  ##KEGG=goseq(pwf,gene2cat=kegg)
##KEGG=goseq(pwf,gene2cat=kegg, method = "Hypergeometric")

##KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
##row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
##KEGG$Name=kegg.names[row.names(KEGG),1]
##Moderate.KEGG = KEGG
##write.csv(Moderate.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/ModerateKEGGAnalysis.csv")

##sigs = Df.data[Df.data$Strong == 1 , "Symbol"]
##genes = setNames(rep(0, nrow(Df.data)), Df.data$Symbol)
##genes[sigs] = 1
##pwf=nullp(genes,"dm3","geneSymbol")
  ##GO.wall=goseq(pwf,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol", method = "Hypergeometric")

##GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
##row.names(GO.wall) = GO.wall$category
##Strong.GO = GO.wall
##write.csv(Strong.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/StrongGOAnalysis.csv")

  ##KEGG=goseq(pwf,gene2cat=kegg)
##KEGG=goseq(pwf,gene2cat=kegg, method = "Hypergeometric")

##KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
##row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
##KEGG$Name=kegg.names[row.names(KEGG),1]
##Strong.KEGG = KEGG
##write.csv(Strong.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/StrongKEGGAnalysis.csv")

##sigs = Df.data[Df.data$Lethal == 1, "Symbol"]
##genes = setNames(rep(0, nrow(Df.data)), Df.data$Symbol)
##genes[sigs] = 1
##pwf=nullp(genes,"dm3","geneSymbol")
  ##GO.wall=goseq(pwf,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol", method = "Hypergeometric")

##GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
##row.names(GO.wall) = GO.wall$category
##Lethal.GO = GO.wall
##write.csv(Lethal.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/LethalGOAnalysis.csv")

  ##KEGG=goseq(pwf,gene2cat=kegg)
##KEGG=goseq(pwf,gene2cat=kegg, method = "Hypergeometric")

##KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
##row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
##KEGG$Name=kegg.names[row.names(KEGG),1]
##Lethal.KEGG = KEGG
##write.csv(Lethal.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/LethalKEGGAnalysis.csv")

##sigs = Df.data[Df.data$Enhancer == 1, "Symbol"]
##genes = setNames(rep(0, nrow(Df.data)), Df.data$Symbol)
##genes[sigs] = 1
##pwf=nullp(genes,"dm3","geneSymbol")
  ##GO.wall=goseq(pwf,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol", method = "Hypergeometric")

##GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
##row.names(GO.wall) = GO.wall$category
##Enhance.GO = GO.wall
##write.csv(Enhance.GO, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/EnhancerGOAnalysis.csv")

  ##KEGG=goseq(pwf,gene2cat=kegg)
##KEGG=goseq(pwf,gene2cat=kegg, method = "Hypergeometric")

##KEGG$adjp=p.adjust(KEGG$over_represented_pvalue,method="BH")
##row.names(KEGG)=substr(as.character(KEGG[,1]),6,13)
##KEGG$Name=kegg.names[row.names(KEGG),1]
##Enhance.KEGG = KEGG
##write.csv(Enhance.KEGG, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/EnhancerKEGGAnalysis.csv")

```

```{r echo = F, include =F}

GO20 = function(GO.data, GO.ontology){
GO.data = GO.data[GO.data$ontology == GO.ontology,
                  c("category", "term","adjp", "numDEInCat", "numInCat")]

data = GO.data[20:1,c("term","adjp","numDEInCat", "numInCat", "category")]
data$Score = -log10(data$adjp)
data$term = factor(data$term, levels = c(unique(data$term)))

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

fig <- plot_ly(data, 
               x = ~numDEInCat, y = ~term, 
               type = 'bar',
               name = "term",
               marker = list(colorscale = list(c(0,.5,1), c("brown","gold", "steelblue")),
    reversescale =T,
                             colorbar = list(title = "-log10(FDR)",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                             color = ~Score,
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$term,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = "Top 20 Most Enriched GO Terms",
           margin = m)

  return(fig)
}

```

```{r echo = F, include =F}

GO.data = Enhancer.GO
GO.ontology = "BP"
GOBP.fig = GO20(GO.data, GO.ontology)

GO.ontology = "CC"
GOCC.fig = GO20(GO.data, GO.ontology)

GO.ontology = "MF"
GOMF.fig = GO20(GO.data, GO.ontology)

```

```{r, fig.width= 10}

GOBP.fig

```

```{r, fig.width= 10}

GOCC.fig

```

```{r, fig.width= 10}

GOMF.fig

```

```{r include=F, echo=F}

GO.data = Enhancer.GO
GO.ontology = "BP"

go_analysis = GO.data$category[GO.data$adjp <= .05 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$adjp), go_analysis)

simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

reducedTerms <- reduceSimMatrix(simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r fig.height=7, fig.width=10}

heatmapPlot(simMatrix,
            reducedTerms,
            annotateParent=TRUE,
            annotationLabel="parentTerm",
            fontsize=6,
            main = "Repressor Gene Enriched Biological Process GO terms")

```

```{r}

scatterPlot(simMatrix, reducedTerms)

```

```{r, fig.width=10}
 use.data   = simMatrix
    pca <- prcomp(t(use.data), scale.=TRUE) 
    pc1=scale(pca$x[,1])
    pc2=scale(pca$x[,2])
    eigs <- pca$sdev^2
    ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(PC1=pc1,
                        PC2=pc2,
                        sdev=pca$sdev,
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group = reducedTerms[row.names(use.data), "parentTerm"],
                        Score = reducedTerms[row.names(use.data), "score"],
                        Term = reducedTerms[row.names(use.data), "term"],
                        TotalDE = GO.data[row.names(use.data), "numDEInCat"],
                        Total = GO.data[row.names(use.data), "numInCat"],
                        SigVal = signif(GO.data[row.names(use.data), "adjp"],3))

    pca.data$Group = factor(pca.data$Group, levels = unique(pca.data$Group))
    pca.data$Score = 20*(1+pca.data$Score/mean(pca.data$Score))

  fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 ##symbol = ~Group,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c("firebrick","gold","deepskyblue"),
                 marker = list(size = pca.data$Score,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Term, 
                                   "\nTotal DE: ", pca.data$TotalDE, 
                                   "\nTotal in Cat.: ", pca.data$Total, 
                                   "\nFDR: ", pca.data$SigVal)) %>%
    layout(xaxis = list(title = paste0("PC1", "(", ve[1], "%)")),
           yaxis = list(title = paste0("PC2", "(", ve[2], "%)")),
           title = "Enriched GO term similarity")

  fig

```

```{r fig.align='left'}

treemapPlot(reducedTerms)

```

```{r include = F, echo = F}

GOID = "0006334"
GenesInGO = genesingo[paste0("GO:", GOID)]

GenesInGO = Df.data[Df.data$ensembl %in% GenesInGO[[1]],]

```

```{r}

go2gene = setNames(list(as.character(genesingo[[1]])), names(genesingo)[1])
i=2
while(i<=length(genesingo)){
  temp = setNames(list(as.character(genesingo[[i]])), names(genesingo)[i])
  go2gene = c(go2gene, temp)
  i=i+1
}

```