### 2025-06-18 GeneralDataFiles/CompleteMetabolicNetwork.Rmd

---
title: "Full Metabolic Network"
author: "John Santiago"
date: "2025-03-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo = F}

library(visNetwork)

git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/refs/heads/master/GeneralDataFiles/"

GeneIDKey = read.csv(paste0(git.dir, "GeneIDKey.csv"), row.names = 1)
G85R.geneFDR = read.csv(paste0(git.dir, "TKT.EdgeR.FDRTable.csv"), row.names = 1)
G85R.geneFC = read.csv(paste0(git.dir, "TKT.EdgeR.FCTable.csv"), row.names = 1)
G85R.metabFC = read.csv(paste0(git.dir, "MetaboliteFCs.csv"), row.names = 1)
G85R.metabFC[,2:ncol(G85R.metabFC)] = log2(G85R.metabFC[,2:ncol(G85R.metabFC)])
G85R.metabFDR = read.csv(paste0(git.dir, "MetaboliteFDRs.csv"), row.names = 1)

raw.nodes = read.csv(paste0(git.dir, "combined.nodes.csv"), row.names = 1)
raw.nodes$x = raw.nodes$x*1
raw.nodes$y = raw.nodes$y*1

raw.edges = read.csv(paste0(git.dir, "combined.edges.csv"))

temp = colorRampPalette(c("navy", "royalblue","white", "deeppink", "firebrick"))(401)
temp2 = colorRampPalette(c("turquoise", "grey", "deeppink"))(401)

```

```{r echo = F}

##metabolomics Choices
##"GRxWT.C"  "GRxWT.Df" "GRxWT.OE" "GR.CxDf" "WT.CxDf"  "GR.CxOE"  "WT.CxOE"
metab = "GRxWT.C"

##Transcriptome choices
##"GRF.CxDf" "GRF.CxOE"  "WTF.CxDf" "WTF.CxOE"  "GRxWT.FC" "GRxWT.FDf"
##"GRxWT.FOE" "GRM.CxDf"  "GRM.CxOE" "WTM.CxDf"  "WTM.CxOE"  "GRxWT.MC"
##"GRxWT.MDf" "GRxWT.MOE" "GR.FxM" "WT.FxM" 
enzyme = "GRxWT.FDf"

nodes = raw.nodes
edges = raw.edges

nodes[nodes$shape == "dot", "label"] = NA

nodes.color = G85R.metabFC[nodes[nodes$shape == "circle", "data.id"], metab]
nodes.color = temp[((nodes.color/max(abs(na.omit(nodes.color))))*200)+201]
nodes[nodes$shape == "circle", "color.background"] = nodes.color
nodes$color.background[na.action(na.exclude(nodes$color.background))] = "white"

node.size = G85R.metabFDR[nodes[nodes$shape == "circle", "data.id"], metab]
node.size[na.action(na.exclude(node.size))] = 1
node.size = -log10(node.size)+1
node.size = node.size*10
node.size[node.size>80] = 50
node.size[node.size < (-log10(.2)+1)*10] =10
nodes[nodes$shape == "circle", "size"] = node.size

nodes$shape = 'dot'

nodes$title = paste0(nodes$id,
                     "<br>-log2(FC): ", 
                     signif((G85R.metabFC[nodes$data.id, metab]), 3),
                     "<br>FDR: ", 
                     signif((G85R.metabFDR[nodes$data.id, metab]), 3))

edges$title = paste0(GeneIDKey[edges$FBgn, "Symbol"],
                     "<br>-log2(FC): ", 
                     signif(G85R.geneFC[edges$FBgn,
                                 enzyme],3),
                     "<br>FDR: ", 
                     signif(G85R.geneFDR[edges$FBgn,
                                 enzyme],3))

edges.width = G85R.geneFDR[edges$FBgn, enzyme]
edges.width[na.action(na.exclude(edges.width))] = 1
edges.width = (-log10(edges.width))+1
edges.width = edges.width * 2

edges.width[edges.width>6] = 6
edges.width[edges.width < (-log10(.2)+1)*2] =1

edges$width = edges.width

edges.colors = G85R.geneFC[edges$FBgn, enzyme]
edges.colors = temp[((edges.colors/max(abs(na.omit(edges.colors))))*200)+201]
edges.colors[na.action(na.exclude(edges.colors))] = "grey"

edges$color = edges.colors

visNetwork(nodes, edges, background = "black")%>%
  visNodes(physics = F)%>%
  #visLayout(hierarchical=F, improvedLayout=T)%>%
  visOptions(highlightNearest = TRUE, nodesIdSelection = list(enabled = T, values = nodes[nodes$size > 0, "label"]))%>%
  visNodes(color = list(border = "white"), font = list(color = "gold", size = 14, background = "black"))%>%
  visEdges(arrows = visedges$arrows, physics = T)

```

```{r echo = F}

nodes = raw.nodes
edges = raw.edges

nodes[nodes$shape == "dot", "label"] = NA
nodes[nodes$shape == "circle", "size"] = 20
nodes.color = G85R.metabFC[nodes[nodes$shape == "circle", "data.id"], metab]
nodes.color = temp[((nodes.color/max(abs(na.omit(nodes.color))))*200)+201]
nodes[nodes$shape == "circle", "color.background"] = nodes.color
nodes$color.background[na.action(na.exclude(nodes$color.background))] = "ghostwhite"
nodes[nodes$shape == "circle", "size"] = 20
nodes$shape = "dot"

edges$color = "black"

PPP.genes = c("FBgn0023477",
              "FBgn0050499",
              "FBgn0050410",
              "FBgn0004654",
              "FBgn0030239",
              "FBgn0037147",
              "FBgn0004057",
              "FBgn0037607",
              "FBgn0036784")

edges[edges$FBgn %in% c(PPP.genes), "color"] = "black"
edges[edges$FBgn %in% c(PPP.genes), "width"] = 2.5
#edges[ grep("FBgn0036784", edges$FBgn), "color"] = "darkgreen"
#edges[ grep("FBgn0037607", edges$FBgn), "color"] = "darkorange"

visNetwork(nodes, edges)%>%
  visNodes(physics = F, font = list(size = 18, background = hex("white", .6)))

#edges[grep('red', edges$color),]

```