### 2025-08-10 DeficiencyScreen/Prelim_Analysis.Rmd
ana update

---
title: "Defficiency Screen Functional Analysis"
author: "John Santiago"
date: "11/28/2020"
output: html_document
---

```{r setup, include=FALSE}
library(rgl)
knitr::opts_chunk$set(echo = FALSE, fig.align="center")
knitr::knit_hooks$set(webgl = hook_webgl)
```

```{r echo = F, include=F}
##Load Libraries

library(org.Dm.eg.db)
library(goseq)
library(rrvgo)
library(plotly)
library(VennDiagram)
genesingo=as.list(org.Dm.egGO2ALLEGS)

hex = function(color, alpha){
  rgb2hex = color
  i = 1
  while(i <= length(color)){
    rgbvals = col2rgb(color[i])
    rgb2hex[i] = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
    i = i+1
  }
  return(rgb2hex)
}

```

```{r echo = F, include =F}
##load all files from github
git.dir = "https://raw.githubusercontent.com/johncsantiago/WhartonLab/master/"

DfKey = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/DfKey.csv"))
##colnames(DfKey)[c(1,3)] = c("BDSC", "FBab")
##DfKey$FollowUp = "FALSE"
##DfKey$Excluded = "FALSE"

old.dfdata = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/DeficiencyModifierGenes.csv"),row.names)

##GeneIDKey = read.csv(paste0(git.dir,"GeneralDataFiles/GeneIDKey.csv"),row.names = 1)

##dfdata = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/20250602_UpdatedModiferData.csv"))
##colnames(dfdata)[3] = "FBab"
##dfdata$Excluded = "FALSE"

##DfKey[DfKey$FBab %in% setdiff(DfKey$FBab, dfdata$FBab), "Excluded"] = "TRUE"

##fullDfKey = rbind(dfdata, DfKey[DfKey$Excluded == TRUE, colnames(dfdata)])

##fullDfKey$Modifier[fullDfKey$Modifier == "No effect"] = "No Effect"

##fullDfKey$Modifier[fullDfKey$Modifier == "Supressor"] = "Suppressor"

##fullDfKey$Modifier[fullDfKey$Modifier == "Strong Suppressor"] = "Suppressor"

##changed start location for FBab0002465
###was listed as 1 but on Flybase it listed a different start site

##write.csv(fullDfKey, "/Users/johncsantiago/Documents/GitHub/WhartonLab/DeficiencyScreen/DataFiles/20250613_FullDfData.csv")

##dfgenes = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/GenesInDeficiency.csv"))

##fulldf = cbind(dfgenes, dfdata[dfgenes$FBab,])
##row.names(fulldf) = 1:nrow(fulldf)
##fulldf$Modifier[fulldf$Modifier ==  "Supressor"] = "Suppressor"
##fulldf$Modifier[fulldf$Modifier ==  "No effect"] = "No Effect"
##new=fulldf

##olddf = DfKey$Deficiency
##newdf = unique(new$Deficiency)
##oldfb = old$FBgn
##newfb = unique(new$FBgn)
##oldsym = old$Symbol
##newsym = unique(new$Symbol)

fullDfKey = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/20250618_FullDfData.csv"),row.names = 1)

row.names(fullDfKey) = fullDfKey$FBab

#Build a fresh gene id data table
gene.chr = as.data.frame(org.Dm.egCHR)
row.names(gene.chr) = gene.chr$gene_id
gene.start = as.list(org.Dm.egCHRLOC)
gene.start = lapply(gene.start, min)
gene.start = unlist(gene.start)
gene.end = as.list(org.Dm.egCHRLOCEND)
gene.end = lapply(gene.end, max)
gene.end = unlist(gene.end)
FBid = as.data.frame(org.Dm.egFLYBASE)
row.names(FBid) = FBid$gene_id
Symbols = as.data.frame(org.Dm.egSYMBOL)
row.names(Symbols) = Symbols$gene_id

gene.location = data.frame(FBgn = FBid[names(gene.start), "flybase_id"],
                           Symbol = Symbols[names(gene.start), "symbol"],
                           ensembl = names(gene.start),
                           CHR = gene.chr[names(gene.start), "chromosome"],
                           Start = abs(gene.start),
                           End = abs(gene.end[names(gene.start)]),
                           Strand = gene.start/abs(gene.start))

gl = apply(gene.location[,c("Start", "End")], MARGIN = 1, min)
gr = apply(gene.location[,c("Start", "End")], MARGIN = 1, max)
Dfl = apply(fullDfKey[,c("Start", "End")], MARGIN = 1, min)
Dfr = apply(fullDfKey[,c("Start", "End")], MARGIN = 1, max)

genesindf = data.frame(Df = character(),
                       FBgn = character(),
                       Symbol = character())
i=1
while(i <= nrow(fullDfKey)){
  temp = na.omit(gene.location[((gl <= Dfl[i] & gr >= Dfr[i]) |
                               (gl >= Dfl[i] & gl <= Dfr[i]) |
                               (gr >= Dfl[i] & gr <= Dfr[i])) &
                               gene.location$CHR == fullDfKey$Chr[i],])

  if(nrow(temp) > 0){
    temp$Df = fullDfKey$Deficiency[i]
    genesindf = rbind(genesindf, temp[, c("Df", "FBgn", "Symbol")])
  }
  #genesindf[[fullDfKey$Deficiency[i]]] = temp$FBgn

  i=i+1
}

python.genesindf = read.csv("/Users/johncsantiago/Documents/GitHub/Python/DeficiencyData/Output.csv", row.names = 1)

py.gid = python.genesindf[,c(3,1,2)]
colnames(py.gid) = c("Df", "FBgn", "Symbol")
py.gid$Df = fullDfKey[py.gid$Df, "Deficiency"]

##Dealing with nomenclature issues
##identifying genes with an FBgn found using python method that were not identified using org.Dm.eg.db approach

py.only = py.gid[py.gid$FBgn %in% setdiff(py.gid$FBgn, gene.location$FBgn),]

##of these 23 genes, how many have a symbol only found with python

py.only = py.only[py.only$Symbol %in% setdiff(py.only$Symbol, gene.location$Symbol),]

##only 8 genes that were not found using the org.Dm.eg.db approach by symbol or FBgn

##unique(py.only$Symbol)
#"Est17" found under CG10175 in R data
#"Dld" found under CG7430 in R data

##NOT in the R data
#"lncRNA:CR46532" 
#"asRNA:PS4"     
#"lncRNA:CR46530" 
#"lncRNA:CR46535"
#"lncRNA:CR46529"
#"lncRNA:CR46534"

py.only.FBgn = c("FBgn0289441", "FBgn0288728", "FBgn0289439", "FBgn0289444", "FBgn0289438", "FBgn0289443")

py.only.sym = c("lncRNA:CR46532", "asRNA:PS4", "lncRNA:CR46530", "lncRNA:CR46535", "lncRNA:CR46529", "lncRNA:CR46534")

py.only.ens = c("86703978", "86703974", "86703976", "86703981", "86703975", "86703980")

py.only.CHR = c("3R", "3R", "3R", "2L", "2L", "2R")

py.only.Start = c(7341783, 16720127, 13056818, 7595504, 9747046, 20149996)

py.only.End = c(7344884, 16729959, 13057676, 7597500, 9747785, 20150499)

py.only.Strand = c(-1, 1, -1, -1, -1, -1)

add.gene.locations = data.frame(FBgn = py.only.FBgn,
                                Symbol = py.only.sym,
                                ensembl = py.only.ens,
                                CHR = py.only.CHR,
                                Start = py.only.Start,
                                End = py.only.End,
                                Strand = py.only.Strand)
gene.location = rbind(gene.location, add.gene.locations)

## redo genes in each Df
gl = apply(gene.location[,c("Start", "End")], MARGIN = 1, min)
gr = apply(gene.location[,c("Start", "End")], MARGIN = 1, max)
Dfl = apply(fullDfKey[,c("Start", "End")], MARGIN = 1, min)
Dfr = apply(fullDfKey[,c("Start", "End")], MARGIN = 1, max)

genesindf = data.frame(Df = character(),
                       FBgn = character(),
                       Symbol = character())
i=1
while(i <= nrow(fullDfKey)){
  temp = na.omit(gene.location[((gl <= Dfl[i] & gr >= Dfr[i]) |
                               (gl >= Dfl[i] & gl <= Dfr[i]) |
                               (gr >= Dfl[i] & gr <= Dfr[i])) &
                               gene.location$CHR == fullDfKey$Chr[i],])

  if(nrow(temp) > 0){
    temp$Df = fullDfKey$Deficiency[i]
    genesindf = rbind(genesindf, temp[, c("Df", "FBgn", "Symbol")])
  }
  i=i+1
}

```

```{r echo = F}
#all the current Df data after Jonah omissions
#fullDfKey

##genes in the current Df data
##genesindf

##the old Df data before Jonah omissions
DfKey = read.csv(paste0(git.dir, "DeficiencyScreen/DataFiles/DfKey.csv"))

##Genes that were originally associated with the old Df data
old.dfdata = read.csv(paste0(git.dir,"DeficiencyScreen/DataFiles/DeficiencyModifierGenes.csv"),row.names = 1)

```

```{r echo = F}

  ##Sample titles as strings. Only fill in up to your number of selected categories
  set1 = "All old Df"
  set2 = "All new Df"
  set3 = "Old Suppressors/Strong Suppressors"
  set4 = "New Suppressors"

  ##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
  s1 = DfKey$FBab..
  s2 = fullDfKey$FBab[fullDfKey$Omit == F]
  s3 = DfKey$FBab..[grep("Sup",DfKey$Modifier)]
  s4 = fullDfKey[fullDfKey$Omit == F,]
  s4 = s4$FBab[grep("Sup", s4$Modifier)]
  main = ""

  grid.newpage()
  tempvenn=draw.quad.venn(area1=length(s1),area2=length(s2),area3=length(s3),area4=length(s4), n12=length(intersect(s1,s2)),n13 = length(intersect(s1,s3)),n14 = length(intersect(s1,s4)),n23 = length(intersect(s2,s3)),n24 = length(intersect(s2,s4)),n34 = length(intersect(s3,s4)),n123 =  length(intersect(s1,intersect(s2,s3))),n124 =  length(intersect(s1,intersect(s2,s4))), n134 = length(intersect(s1,intersect(s3,s4))),n234 = length(intersect(s2,intersect(s3,s4))),n1234 = length(intersect(s1,intersect(s2,intersect(s3,s4)))),category =c(set1,set2,set3,set4),fill = c("blue", "red","green","yellow"),cex=2,cat.cex = 1, title = "main")

```

```{r echo = F}

trimmed.Df = fullDfKey[fullDfKey$Omit == "FALSE",]
trimmed.genes = genesindf[genesindf$Df %in% trimmed.Df$Deficiency,]
#trimmed.sup = fullDfKey[fullDfKey$Omit == "FALSE" & fullDfKey$MLE >= 2,] 
trimmed.sup = fullDfKey[(fullDfKey$Omit == "FALSE" & fullDfKey$MLE == 2) &
                         fullDfKey$Modifier ==  "No Effect",] 
#trimmed.sup = fullDfKey[fullDfKey$Omit == "FALSE" & fullDfKey$Modifier ==  "Suppressor",] 
trimmed.supgenes = genesindf[genesindf$Df %in% trimmed.sup$Deficiency,]

sigs = unique(trimmed.supgenes$Symbol)
genes = setNames(rep(0, length(unique(trimmed.genes$Symbol))), unique(trimmed.genes$Symbol))

#randoms = sample(1:length(genes), 6061)

genes[sigs] = 1
#genes[randoms] = 1
pwf=nullp(genes,"dm3","geneSymbol")
##GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol", method = "Hypergeometric")

GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category

#suppressor.GO = GO.wall
#twodayplus.GO = GO.wall
#twotothree.GO = GO.wall

```

```{r echo = F, warning=F, message=F}
##GO analysis of all Df with MLE 3 or greater, omitting Dfs with escapers
sigs.Df = fullDfKey[fullDfKey$Omit == "FALSE" & fullDfKey$MLE > 3,]
sigs = genesindf[genesindf$Df %in% sigs.Df$Deficiency, "Symbol"]

genes.Df = fullDfKey[fullDfKey$Omit == "FALSE",]
genes = genesindf[genesindf$Df %in% genes.Df$Deficiency, "Symbol"]
sigs = unique(trimmed.supgenes$Symbol)
genes = setNames(rep(0, length(unique(genes))), unique(genes))

genes[sigs] = 1
pwf=nullp(genes,"dm3","geneSymbol")
GO.wall=goseq(pwf,"dm3","geneSymbol")
GO.wall$adjp=p.adjust(GO.wall$over_represented_pvalue,method="BH")
row.names(GO.wall) = GO.wall$category

sup.GO = GO.wall

```

```{r echo = F}

df.genesingo = gene.location[gene.location$ensembl %in% intersect(gene.location$ensembl, as.character(genesingo[["GO:0030527"]])),]

summary(factor(genesindf[genesindf$FBgn %in% df.genesingo$FBgn, "Df"], levels = unique(genesindf[genesindf$FBgn %in% df.genesingo$FBgn, "Df"])))

fullDfKey[fullDfKey$Deficiency == "Df(2L)ED1473",]

```

```{r echo = F, warning=F, include = F}

hist(na.omit(as.numeric(dfdata$MLE)), col = hex('yellow', .25),
     breaks = seq(-4, 7, by = 1),
     xlab = "MLE",
     main = "Old Df MLE data vs New")
hist(na.omit(as.numeric(DfKey$MLE)), add = T, col = hex('blue', .2),
     breaks = seq(-4, 7, by = 1))
legend("topright", legend = c("old data", "new data"), fill = c("blue", "yellow"))
```

```{r echo = F}

  ##Sample titles as strings. Only fill in up to your number of selected categories
  set1 = "all old FBgn"
  set2 = "all new FBgn"
  set3 = "FBgn in Old Unique Df"
  set4 = "FBgn in New Unique Df"

  ##items to be compared (ex: gene names) as a 1 dimensional vectors. Only give inputs up to your number of selected categories
  s1 = oldfb
  s2 = newfb
  s3 = fbinolddfonly
  s4 = fbinnewdfonly

  main = ""

  grid.newpage()
  tempvenn=draw.quad.venn(area1=length(s1),area2=length(s2),area3=length(s3),area4=length(s4), n12=length(intersect(s1,s2)),n13 = length(intersect(s1,s3)),n14 = length(intersect(s1,s4)),n23 = length(intersect(s2,s3)),n24 = length(intersect(s2,s4)),n34 = length(intersect(s3,s4)),n123 =  length(intersect(s1,intersect(s2,s3))),n124 =  length(intersect(s1,intersect(s2,s4))), n134 = length(intersect(s1,intersect(s3,s4))),n234 = length(intersect(s2,intersect(s3,s4))),n1234 = length(intersect(s1,intersect(s2,intersect(s3,s4)))),category =c(set1,set2,set3,set4),fill = c("blue", "red","green","yellow"),cex=2,cat.cex = 1, title = "main")

```

```{r echo = F, include =F}

GO20 = function(GO.data, GO.ontology){
GO.data = GO.data[GO.data$ontology == GO.ontology,
                  c("category", "term","adjp", "numDEInCat", "numInCat")]

data = GO.data[20:1,c("term","adjp","numDEInCat", "numInCat", "category")]
data$Score = -log10(data$adjp)
data$term = factor(data$term, levels = c(unique(data$term)))

m <- list(
  l = 50,
  r = 50,
  b = 100,
  t = 100,
  pad = 4
)

fig <- plot_ly(data, 
               x = ~numDEInCat, y = ~term, 
               type = 'bar',
               name = "term",
               marker = list(colorscale = list(c(0,.5,1), c("brown","gold", "steelblue")),
    reversescale =T,
                             colorbar = list(title = "-log10(FDR)",
                                      len = .8, outlinewidth = 0,
                                      tickformat = ".2f",
                                      tick0 = 0),##, dtick = .01),
                             color = ~Score,
                             line = list(color = "black", width = 1.5)),
    hoverinfo = "text",
    hovertext = paste("Sample:", data$term,
                      "\nGO ID: ", data$category,
                      "\nTotal DE: ", data$numDEInCat,
                      "\nTotal in Cat.: ", data$numInCat,
                      "\nFDR: ", signif(data$adjp, 3))) %>%
    layout(xaxis = list(title = "Number of Genes"),
           yaxis = list(title = ""),
           title = "Top 20 Most Enriched GO Terms",
           margin = m)

  return(fig)
}

```

```{r echo = F, include =F}

GO.data = sup.GO
GO.ontology = "BP"
GOBP.fig = GO20(GO.data, GO.ontology)

GO.ontology = "CC"
GOCC.fig = GO20(GO.data, GO.ontology)

GO.ontology = "MF"
GOMF.fig = GO20(GO.data, GO.ontology)

```

```{r, fig.width= 10}

GOBP.fig

```

```{r, fig.width= 10}

GOCC.fig

```

```{r, fig.width= 10}

GOMF.fig

```

```{r echo = F}

sig.sup.GOcats = sup.GO[sup.GO$adjp <= .05, "category"]

ss.gocat.genes = genesingo[sig.sup.GOcats]

ss.gocat1.genes = (gene.location[gene.location$ensembl %in% as.character(ss.gocat.genes[[1]]),])

ss.gocat1.dfs = genesindf[genesindf$FBgn %in% ss.gocat1.genes$FBgn,]

(unique(ss.gocat1.dfs$Df))

```

```{r include=F, echo=F}

GO.data = Enhancer.GO
GO.ontology = "BP"

go_analysis = GO.data$category[GO.data$adjp <= .05 & GO.data$ontology == GO.ontology]
scores = setNames(-log10(GO.data$adjp), go_analysis)

simMatrix <- calculateSimMatrix(go_analysis,
                                orgdb="org.Dm.eg.db",
                                ont=GO.ontology,
                                method="Rel")

reducedTerms <- reduceSimMatrix(simMatrix,
                                scores,
                                threshold=0.7,
                                orgdb="org.Dm.eg.db")

```

```{r fig.height=7, fig.width=10}

heatmapPlot(simMatrix,
            reducedTerms,
            annotateParent=TRUE,
            annotationLabel="parentTerm",
            fontsize=6,
            main = "Repressor Gene Enriched Biological Process GO terms")

```

```{r}

scatterPlot(simMatrix, reducedTerms)

```

```{r, fig.width=10}
 use.data   = simMatrix
    pca <- prcomp(t(use.data), scale.=TRUE) 
    pc1=scale(pca$x[,1])
    pc2=scale(pca$x[,2])
    eigs <- pca$sdev^2
    ve=signif(((eigs / sum(eigs))*100)[1:3],4)

    pca.data=data.frame(PC1=pc1,
                        PC2=pc2,
                        sdev=pca$sdev,
                        Percent1=ve[1],
                        Percent2=ve[2],
                        Group = reducedTerms[row.names(use.data), "parentTerm"],
                        Score = reducedTerms[row.names(use.data), "score"],
                        Term = reducedTerms[row.names(use.data), "term"],
                        TotalDE = GO.data[row.names(use.data), "numDEInCat"],
                        Total = GO.data[row.names(use.data), "numInCat"],
                        SigVal = signif(GO.data[row.names(use.data), "adjp"],3))

    pca.data$Group = factor(pca.data$Group, levels = unique(pca.data$Group))
    pca.data$Score = 20*(1+pca.data$Score/mean(pca.data$Score))

  fig  = plot_ly(data = pca.data, 
                 x = ~PC1, 
                 y = ~PC2,
                 name = ~Group,
                 type = 'scatter',
                 mode = 'markers',
                 showlegend = T,
                 color = ~Group,
                 ##symbol = ~Group,
                 ##symbols = as.character(seq_along(unique(point.data$Set))),
                 ##colors = c("firebrick","gold","deepskyblue"),
                 marker = list(size = pca.data$Score,
                               line = list(color = "black", width = 1.5)),
                 hoverinfo = "text",
                 hovertext = paste("Sample:", pca.data$Term, 
                                   "\nTotal DE: ", pca.data$TotalDE, 
                                   "\nTotal in Cat.: ", pca.data$Total, 
                                   "\nFDR: ", pca.data$SigVal)) %>%
    layout(xaxis = list(title = paste0("PC1", "(", ve[1], "%)")),
           yaxis = list(title = paste0("PC2", "(", ve[2], "%)")),
           title = "Enriched GO term similarity")

  fig

```

```{r fig.align='left'}

treemapPlot(reducedTerms)

```

```{r include = F, echo = F}

GOID = "0006334"
GenesInGO = genesingo[paste0("GO:", GOID)]

GenesInGO = Df.data[Df.data$ensembl %in% GenesInGO[[1]],]

```

```{r}

go2gene = setNames(list(as.character(genesingo[[1]])), names(genesingo)[1])
i=2
while(i<=length(genesingo)){
  temp = setNames(list(as.character(genesingo[[i]])), names(genesingo)[i])
  go2gene = c(go2gene, temp)
  i=i+1
}

```