### 2025-08-10 RussoPaper/AnaDataAnalysis.Rmd

---
title: "Ana Data"
author: "John Santiago"
date: "2025-06-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=F, message=FALSE, warning=FALSE}

library(kableExtra)
library(gtools)
library(gplots)
library(beeswarm)

temp = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/20250809_AllAnaData.csv", row.names = 1)
temp = temp[temp$Boutons != "x",]
temp$Boutons = as.numeric(temp$Boutons)
temp$Genotype[temp$Genotype == "WT-cr/LoxP"] = "WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "G85R"] = "G85R(cr/hr)"
temp$Genotype[temp$Genotype == "G85Rcr/G85Rhr"] = "G85R(cr/hr)"

temp$Genotype[temp$Genotype == "OK371 > gbb [WT]"] = "OK371-gbb; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371 > gbb [G85R]"] = "OK371-gbb; G85R(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>gbb [WT]"] = "OK371-gbb; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>gbb [G85R]"] = "OK371-gbb; G85R(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>gbb [WTcr/LoxP]"] = "OK371-gbb; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>gbb [WT-cr/LoxP]"] = "OK371-gbb; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>gbb [G85Rcr/G85Rhr]"] = "OK371-gbb; G85R(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>gbb [G85cr/G85Rhr]"] = "OK371-gbb; G85R(cr/LoxP)"

temp$Genotype[temp$Genotype == "OK371 > saxA [WT]"] = "OK371-saxA; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371 > saxA [G85R]"] = "OK371-saxA; G85R(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371> saxA [WT-cr/LoxP]"] = "OK371-saxA; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>saxA G85R"] = "OK371-saxA; G85R(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>saxA [WTcr/LoxP]"] = "OK371-saxA; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>saxA [WT-cr/LoxP]"] = "OK371-saxA; WT(cr/LoxP)"
temp$Genotype[temp$Genotype == "OK371>saxA [G85Rcr/G85Rhr]"] = "OK371-saxA; G85R(cr/LoxP)"

temp$Genotype = factor(temp$Genotype, levels = unique(temp$Genotype))

std.colors = setNames(c('navy', "orange2", 
                        "darkgreen", "turquoise3",
                        "firebrick","deeppink"),
                      c("WT(cr/LoxP)" , "G85R(cr/hr)",
                        "OK371-gbb; WT(cr/LoxP)", "OK371-gbb; G85R(cr/LoxP)",
                        "OK371-saxA; WT(cr/LoxP)", "OK371-saxA; G85R(cr/LoxP)"))

hex = function(color, alpha){
  rgb2hex = color
  i = 1
  while(i <= length(color)){
    rgbvals = col2rgb(color[i])
    rgb2hex[i] = rgb(rgbvals[1]/255, rgbvals[2]/255, rgbvals[3]/255, alpha)
    i = i+1
  }
  return(rgb2hex)
}

```

```{r echo = F}

plot.bouts = function(boutdata){

boutmeans = c(mean(boutdata[boutdata$Genotype == geno[1], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[2], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[3], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[4], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[5], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[6], "Boutons"]))

par(mar=c(6.25, 4.1, 4.1, 5.6))

plot(x = c(.75,1.25),
     y = rep(boutmeans[1], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1],
     ylim = c(0, max(boutdata$Boutons)*1.1),
     xlim = c(.5, length(na.omit(boutmeans))+.5),
     xaxt = "n",
     ylab="Bouton Number",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = paste("Larval muscle Segment", unique(boutdata$Segment), "NMJ", unique(boutdata$NMJ), "bouton counts"))

lines(x = c(.65,1.35),
     y = rep(boutmeans[1], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(1.65,2.35),
     y = rep(boutmeans[2], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

if(length(na.omit(boutmeans)>2)){
lines(x = c(2.65,3.35),
     y = rep(boutmeans[3], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[3])

lines(x = c(3.65,4.35),
     y = rep(boutmeans[4], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[4])

lines(x = c(4.65,5.35),
     y = rep(boutmeans[5], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[5])

lines(x = c(5.65,6.35),
     y = rep(boutmeans[6], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[6])
}

beeswarm(boutdata$Boutons~boutdata$Genotype,
       pch = 21,
       method = "center",
       bg = hex(unique(std.colors[as.numeric(boutdata$Genotype)]),.5),
       add = T)

axis(side = 1,
     at = c(1:length(na.omit(boutmeans))),
     labels = c(labels)[1:length(na.omit(boutmeans))],
     las = 2,
     lwd = 0,
     line = -.75)

    lines(x = c(.225 + .3/length(na.omit(boutmeans)), length(na.omit(boutmeans))+.5),
       y = c(0,0),
       lwd = 3,
       xpd = T)

  y.at = ceiling(max(boutdata$Boutons)*.11)
  y.at = c(0,10*1:y.at)

  axis(side = 2,
       at = y.at,
       lwd = 2,
       line = 0,
       labels = F,
       xpd = T,
       las = 2)

  axis(side = 2,
       at = y.at,
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)
}

```

```{r echo = F, fig.width = 7}

Boutons = temp[temp$Data == "Boutons",]
boutdata = Boutons[Boutons$NMJ == "4" &
                        Boutons$Segment == "3",]
boutdata$Boutons = as.numeric(boutdata$Boutons)

geno = unique(boutdata$Genotype)
labels = c("WT", "G85R",
           "OK-gbb WT", "OK-gbb G85R",
           "OK-SaxA WT", "OK-SaxA G85R")

plot.bouts(boutdata)

```

```{r echo = F}

btd = boutdata

t.data = list("WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "WT(cr/LoxP)"],
                    "G85R(cr/hr)" = btd$Boutons[btd$Genotype == "G85R(cr/hr)"],
                    "OK371-gbb; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-gbb; WT(cr/LoxP)"],
                    "OK371-gbb; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-gbb; G85R(cr/LoxP)"],
                    "OK371-saxA; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-saxA; WT(cr/LoxP)"],
                    "OK371-saxA; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-saxA; G85R(cr/LoxP)"])

t.table = data.frame(Genotype1 = character(),
                     Genotype2 = character(),
                     pval = numeric())

i=1
while(i <= length(t.data)){
  j = i+1
  while(j <= length(t.data)){
    t.temp = data.frame(Genotype1 = geno[i],
                        Genotype2 = geno[j],
                        pval = t.test(x = t.data[[i]],
                                      y = t.data[[j]],
                                      paired = F)[[3]])
    t.table = rbind(t.table, t.temp)
      j = j+1
  }
  i = i+1
}

t.stripe = t.table$pval <= .05
knitr::kable(t.table,
             caption = 'All t-tests for segment 3, NMJ 4 Bouton Counts.')  %>%
kable_styling() %>%
row_spec(c(1:nrow(t.table))[t.table$pval <= .05], bold = T, color = "black", background = "gold")

```

```{r echo = F, fig.width = 7}

Boutons = temp[temp$Data == "Boutons",]
boutdata = Boutons[Boutons$NMJ == "4" &
                        Boutons$Segment == "5",]
boutdata = boutdata[boutdata$Boutons != "x",]
boutdata$Boutons = as.numeric(boutdata$Boutons)

geno = unique(boutdata$Genotype)
labels = c("WT", "G85R",
           "OK-gbb WT", "OK-gbb G85R",
           "OK-SaxA WT", "OK-SaxA G85R")

plot.bouts(boutdata)

```

```{r echo = F}

btd = boutdata

t.data = list("WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "WT(cr/LoxP)"],
                    "G85R(cr/hr)" = btd$Boutons[btd$Genotype == "G85R(cr/hr)"],
                    "OK371-gbb; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-gbb; WT(cr/LoxP)"],
                    "OK371-gbb; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-gbb; G85R(cr/LoxP)"],
                    "OK371-saxA; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-saxA; WT(cr/LoxP)"],
                    "OK371-saxA; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-saxA; G85R(cr/LoxP)"])

t.table = data.frame(Genotype1 = character(),
                     Genotype2 = character(),
                     pval = numeric())

i=1
while(i <= length(t.data)){
  j = i+1
  while(j <= length(t.data)){
    t.temp = data.frame(Genotype1 = geno[i],
                        Genotype2 = geno[j],
                        pval = t.test(x = t.data[[i]],
                                      y = t.data[[j]],
                                      paired = F)[[3]])
    t.table = rbind(t.table, t.temp)
      j = j+1
  }
  i = i+1
}

t.stripe = t.table$pval <= .05
knitr::kable(t.table,
             caption = 'All t-tests for segment 3, NMJ 4 Bouton Counts.')  %>%
kable_styling() %>%
row_spec(c(1:nrow(t.table))[t.table$pval <= .05], bold = T, color = "black", background = "gold")

```

```{r echo = F, fig.width = 7}

Boutons = temp[temp$Data == "Boutons",]
boutdata = Boutons[Boutons$NMJ == "6-7" &
                        Boutons$Segment == "3",]
boutdata$Boutons = as.numeric(boutdata$Boutons)

geno = unique(boutdata$Genotype)
labels = c("WT", "G85R",
           "OK-gbb WT", "OK-gbb G85R",
           "OK-SaxA WT", "OK-SaxA G85R")

plot.bouts(boutdata)

```

```{r echo = F}

btd = boutdata

t.data = list("WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "WT(cr/LoxP)"],
                    "G85R(cr/hr)" = btd$Boutons[btd$Genotype == "G85R(cr/hr)"],
                    "OK371-gbb; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-gbb; WT(cr/LoxP)"],
                    "OK371-gbb; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-gbb; G85R(cr/LoxP)"],
                    "OK371-saxA; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-saxA; WT(cr/LoxP)"],
                    "OK371-saxA; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-saxA; G85R(cr/LoxP)"])

t.table = data.frame(Genotype1 = character(),
                     Genotype2 = character(),
                     pval = numeric())

i=1
while(i <= length(t.data)){
  j = i+1
  while(j <= length(t.data)){
    t.temp = data.frame(Genotype1 = geno[i],
                        Genotype2 = geno[j],
                        pval = t.test(x = t.data[[i]],
                                      y = t.data[[j]],
                                      paired = F)[[3]])
    t.table = rbind(t.table, t.temp)
      j = j+1
  }
  i = i+1
}

t.stripe = t.table$pval <= .05
knitr::kable(t.table,
             caption = 'All t-tests for segment 3, NMJ 6-7 Bouton Counts.')  %>%
kable_styling() %>%
row_spec(c(1:nrow(t.table))[t.table$pval <= .05], bold = T, color = "black", background = "gold")

```

```{r echo = F, fig.width = 7}

Boutons = temp[temp$Data == "Boutons",]
boutdata = Boutons[Boutons$NMJ == "6-7" &
                        Boutons$Segment == "5",]
boutdata = boutdata[boutdata$Boutons != "x",]
boutdata$Boutons = as.numeric(boutdata$Boutons)
geno = unique(boutdata$Genotype)
labels = c("WT", "G85R",
           "OK-gbb WT", "OK-gbb G85R",
           "OK-SaxA WT", "OK-SaxA G85R")

plot.bouts(boutdata)

```

```{r echo = F}

btd = boutdata

t.data = list("WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "WT(cr/LoxP)"],
                    "G85R(cr/hr)" = btd$Boutons[btd$Genotype == "G85R(cr/hr)"],
                    "OK371-gbb; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-gbb; WT(cr/LoxP)"],
                    "OK371-gbb; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-gbb; G85R(cr/LoxP)"],
                    "OK371-saxA; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-saxA; WT(cr/LoxP)"],
                    "OK371-saxA; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "OK371-saxA; G85R(cr/LoxP)"])

t.table = data.frame(Genotype1 = character(),
                     Genotype2 = character(),
                     pval = numeric())

i=1
while(i <= length(t.data)){
  j = i+1
  while(j <= length(t.data)){
    t.temp = data.frame(Genotype1 = geno[i],
                        Genotype2 = geno[j],
                        pval = t.test(x = t.data[[i]],
                                      y = t.data[[j]],
                                      paired = F)[[3]])
    t.table = rbind(t.table, t.temp)
      j = j+1
  }
  i = i+1
}

t.stripe = t.table$pval <= .05
knitr::kable(t.table,
             caption = 'All t-tests for segment 3, NMJ 6-7 Bouton Counts.')  %>%
kable_styling() %>%
row_spec(c(1:nrow(t.table))[t.table$pval <= .05], bold = T, color = "black", background = "gold")

```

```{r echo = F, include = F}

Boutons = temp[temp$Data == "Boutons",]
boutdata = Boutons[Boutons$NMJ == "4" &
                        Boutons$Segment == "4",]

geno = unique(boutdata$Genotype)
labels = c("WT", "G85R",
           "OK-gbb WT", "OK-gbb G85R",
           "OK-SaxA WT", "OK-SaxA G85R")

plot.bouts(boutdata)

```

```{r echo = F, include = F}

Boutons = temp[temp$Data == "Boutons",]
boutdata = Boutons[Boutons$NMJ == "6-7" &
                        Boutons$Segment == "4",]

geno = unique(boutdata$Genotype)
labels = c("WT", "G85R",
           "OK-gbb WT", "OK-gbb G85R",
           "OK-SaxA WT", "OK-SaxA G85R")

plot.bouts(boutdata)

```

```{r echo = F, include = F}

Boutons = temp[temp$Data == "Boutons",]
boutdata = Boutons[Boutons$NMJ == "4" &
                        Boutons$Segment == "5",]

geno = unique(boutdata$Genotype)
labels = c("WT", "G85R",
           "OK-gbb WT", "OK-gbb G85R",
           "OK-SaxA WT", "OK-SaxA G85R")

plot.bouts(boutdata)

```

```{r echo = F, include = F}

Boutons = temp[temp$Data == "Boutons",]
boutdata = Boutons[Boutons$NMJ == "6-7" &
                        Boutons$Segment == "5",]

geno = unique(boutdata$Genotype)
labels = c("WT", "G85R",
           "OK-gbb WT", "OK-gbb G85R",
           "OK-SaxA WT", "OK-SaxA G85R")

plot.bouts(boutdata)

```

```{r echo = F, include = F}

Boutons = temp[temp$Data == "Boutons",]
boutdata = Boutons[Boutons$NMJ == "4" &
                        Boutons$Segment == "6",]

geno = unique(boutdata$Genotype)
labels = c("WT", "G85R",
           "OK-gbb WT", "OK-gbb G85R",
           "OK-SaxA WT", "OK-SaxA G85R")

plot.bouts(boutdata)

```

```{r echo = F, include = F}

Boutons = temp[temp$Data == "Boutons",]
boutdata = Boutons[Boutons$NMJ == "6-7" &
                        Boutons$Segment == "6",]

geno = unique(boutdata$Genotype)
labels = c("WT", "G85R",
           "OK-gbb WT", "OK-gbb G85R",
           "OK-SaxA WT", "OK-SaxA G85R")

plot.bouts(boutdata)

```

```{r echo = F, fig.width = 10}

Boutons = temp[temp$Data == "Boutons",]

#boutdata = Boutons[Boutons$NMJ == "4" &
#                     (Boutons$Genotype == "WT(cr/LoxP)" | #Boutons$Genotype == "G85R(cr/hr)"),]

boutdata = Boutons[Boutons$NMJ == "4",]
boutdata$Genotype = paste(boutdata$Segment, boutdata$Genotype)

geno = unique(boutdata$Genotype)
geno = geno[c(1,5, 9:12, 
              2,6, 
              3,7, 13:16,
              4,8)]

boutdata$Genotype = factor(boutdata$Genotype, levels = geno)

boutmeans = c(mean(boutdata[boutdata$Genotype == geno[1], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[2], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[3], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[4], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[5], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[6], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[7], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[8], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[9], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[10], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[11], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[12], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[13], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[14], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[15], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[16], "Boutons"]))

par(mar=c(4, 4.1, 4.1, 5.6))

plot(x = c(.75,1.25),
     y = rep(boutmeans[1], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1],
     ylim = c(0, max(boutdata$Boutons)*1.1),
     xlim = c(.5, length(na.omit(boutmeans))+.5),
     xaxt = "n",
     ylab="Bouton Number",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = paste("Larval muscle NMJ 4 bouton counts, segments 3-6"))

lines(x = c(.65,1.35),
     y = rep(boutmeans[1], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(1.65,2.35),
     y = rep(boutmeans[2], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

lines(x = c(2.65,3.35),
     y = rep(boutmeans[3], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[3])

lines(x = c(3.65,4.35),
     y = rep(boutmeans[4], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[4])

lines(x = c(4.65,5.35),
     y = rep(boutmeans[5], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[5])

lines(x = c(5.65,6.35),
     y = rep(boutmeans[6], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[6])

lines(x = c(6.65,7.35),
     y = rep(boutmeans[7], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(7.65,8.35),
     y = rep(boutmeans[8], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

lines(x = c(8.65,9.35),
     y = rep(boutmeans[9], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(9.65,10.35),
     y = rep(boutmeans[10], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

lines(x = c(10.65,11.35),
     y = rep(boutmeans[11], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[3])

lines(x = c(11.65,12.35),
     y = rep(boutmeans[12], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[4])

lines(x = c(12.65,13.35),
     y = rep(boutmeans[13], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[5])

lines(x = c(13.65,14.35),
     y = rep(boutmeans[14], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[6])

lines(x = c(14.65,15.35),
     y = rep(boutmeans[15], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(15.65,16.35),
     y = rep(boutmeans[16], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

beeswarm(boutdata$Boutons~boutdata$Genotype,
       pch = 21,
       method = "center",
       bg = hex(std.colors[c(1:6, 1:2, 1:6, 1:2)],.5),
       add = T)

lines(x = c(6.5,6.5),
      y = c(-2.5,70),
      lty = 2,
      xpd = T)

lines(x = c(8.5,8.5),
      y = c(-2.5,70),
      lty = 2,
      xpd = T)

lines(x = c(14.5,14.5),
      y = c(-2.5,70),
      lty = 2,
      xpd = T)

axis(side = 1,
     at = c(3.5, 7.5, 11.5, 15.5),
     labels = c(3,4,5,6),
     las = 1,
     lwd = 0,
     line = -.75)

axis(side = 1,
     at = c(8.5),
     labels = c("Larval Segment"),
     las = 1,
     lwd = 0,
     cex = 2,
     line = 0.5)

    lines(x = c(-.10, length(na.omit(boutmeans))+.5),
       y = c(0,0),
       lwd = 3,
       xpd = T)

  y.at = c(0, 10*1:7)

  axis(side = 2,
       at = y.at,
       lwd = 2,
       line = 0,
       labels = F,
       xpd = T,
       las = 2)

  axis(side = 2,
       at = y.at,
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)

  legend('topright',
       inset=c(-0.14,0),
       legend = c('WT(cr/LoxP)', 'G85R(cr/hr)', 
                  "OK371-gbb; WT(cr/LoxP)", "OK371-gbb; G85R(cr/LoxP)",
                  "OK371-saxA; WT(cr/LoxP)","OK371-saxA; G85R(cr/LoxP)"), 
       fill = std.colors[1:6],
       cex = .5,
       bty = 'n',
       pt.cex = .4,
       xpd = T)

```

```{r echo = F}

btd = boutdata

t.data = list("3 WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "3 WT(cr/LoxP)"],
              "3 G85R(cr/hr)" = btd$Boutons[btd$Genotype == "3 G85R(cr/hr)"],
              "3 OK371-gbb; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "3 OK371-gbb; WT(cr/LoxP)"],
              "3 OK371-gbb; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "3 OK371-gbb; G85R(cr/LoxP)"],
              "3 OK371-saxA; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "3 OK371-saxA; WT(cr/LoxP)"],
              "3 OK371-saxA; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "3 OK371-saxA; G85R(cr/LoxP)"],
              "4 WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "4 WT(cr/LoxP)"],
              "4 G85R(cr/hr)" = btd$Boutons[btd$Genotype == "4 G85R(cr/hr)"],
              "5 WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "5 WT(cr/LoxP)"],
              "5 G85R(cr/hr)" = btd$Boutons[btd$Genotype == "5 G85R(cr/hr)"],
              "5 OK371-gbb; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "5 OK371-gbb; WT(cr/LoxP)"],
              "5 OK371-gbb; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "5 OK371-gbb; G85R(cr/LoxP)"],
              "5 OK371-saxA; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "5 OK371-saxA; WT(cr/LoxP)"],
              "5 OK371-saxA; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "5 OK371-saxA; G85R(cr/LoxP)"],
              "6 WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "6 WT(cr/LoxP)"],
              "6 G85R(cr/hr)" = btd$Boutons[btd$Genotype == "6 G85R(cr/hr)"])

t.table = data.frame(Genotype1 = character(),
                     Genotype2 = character(),
                     pval = numeric())

i=1
while(i <= length(t.data)){
  j = i+1
  while(j <= length(t.data)){
    t.temp = data.frame(Genotype1 = geno[i],
                        Genotype2 = geno[j],
                        pval = t.test(x = t.data[[i]],
                                      y = t.data[[j]],
                                      paired = F)[[3]])
    t.table = rbind(t.table, t.temp)
      j = j+1
  }
  i = i+1
}

t.stripe = t.table$pval <= .05
knitr::kable(t.table,
             caption = 'All t-tests for NMJ 4, all segments Bouton Counts.')  %>%
kable_styling() %>%
row_spec(c(1:nrow(t.table))[t.table$pval <= .05], bold = T, color = "black", background = "gold")

```

```{r echo = F, fig.width = 7, include = F}

Boutons = temp[temp$Data == "Boutons",]
boutdata = Boutons[Boutons$NMJ == "6-7" &
                     (Boutons$Genotype == "WT(cr/LoxP)" | Boutons$Genotype == "G85R(cr/hr)"),]
boutdata$Genotype = paste(boutdata$Segment, boutdata$Genotype)

geno = unique(boutdata$Genotype)
geno = geno[c(1,5,2,6,3,7,4,8)]

boutdata$Genotype = factor(boutdata$Genotype, levels = geno)

boutmeans = c(mean(boutdata[boutdata$Genotype == geno[1], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[2], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[3], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[4], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[5], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[6], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[7], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[8], "Boutons"]))

par(mar=c(4, 4.1, 4.1, 5.6))

plot(x = c(.75,1.25),
     y = rep(boutmeans[1], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1],
     ylim = c(0, max(boutdata$Boutons)*1.1),
     xlim = c(.5, length(na.omit(boutmeans))+.5),
     xaxt = "n",
     ylab="Bouton Number",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = paste("Larval muscle NMJ 6-7 bouton counts, segments 3-6"))

lines(x = c(.65,1.35),
     y = rep(boutmeans[1], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(1.65,2.35),
     y = rep(boutmeans[2], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

lines(x = c(2.65,3.35),
     y = rep(boutmeans[3], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(3.65,4.35),
     y = rep(boutmeans[4], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

lines(x = c(4.65,5.35),
     y = rep(boutmeans[5], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(5.65,6.35),
     y = rep(boutmeans[6], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

lines(x = c(6.65,7.35),
     y = rep(boutmeans[7], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(7.65,8.35),
     y = rep(boutmeans[8], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

beeswarm(boutdata$Boutons~boutdata$Genotype,
       pch = 21,
       method = "center",
       bg = hex(std.colors[1:2],.5),
       add = T)

lines(x = c(2.5,2.5),
      y = c(-5,100),
      lty = 2,
      xpd = T)

lines(x = c(4.5,4.5),
      y = c(-5,100),
      lty = 2,
      xpd = T)

lines(x = c(6.5,6.5),
      y = c(-5,100),
      lty = 2,
      xpd = T)

axis(side = 1,
     at = c(1.5, 3.5, 5.5, 7.5),
     labels = c(3,4,5,6),
     las = 1,
     lwd = 0,
     line = -.75)

axis(side = 1,
     at = c(4.5),
     labels = c("Larval Segment"),
     las = 1,
     lwd = 0,
     cex = 2,
     line = 0.5)

    lines(x = c(.2, length(na.omit(boutmeans))+.5),
       y = c(0,0),
       lwd = 3,
       xpd = T)

  y.at = c(0, 10*1:10)

  axis(side = 2,
       at = y.at,
       lwd = 2,
       line = 0,
       labels = F,
       xpd = T,
       las = 2)

  axis(side = 2,
       at = y.at,
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)

  legend('topright',
       inset=c(-0.185,0),
       legend = c('WT(cr/LoxP)', 'G85R(cr/hr)'), 
       fill = std.colors[1:2],
       cex = 1,
       bty = 'n',
       pt.cex = .5,
       xpd = T)

```

```{r echo = F, include = F}

btd = boutdata

t.data = list("3 WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "3 WT(cr/LoxP)"],
              "3 G85R(cr/hr)" = btd$Boutons[btd$Genotype == "3 G85R(cr/hr)"],
              "4 WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "4 WT(cr/LoxP)"],
              "4 G85R(cr/hr)" = btd$Boutons[btd$Genotype == "4 G85R(cr/hr)"],
              "5 WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "5 WT(cr/LoxP)"],
              "5 G85R(cr/hr)" = btd$Boutons[btd$Genotype == "5 G85R(cr/hr)"],
              "6 WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "6 WT(cr/LoxP)"],
              "6 G85R(cr/hr)" = btd$Boutons[btd$Genotype == "6 G85R(cr/hr)"])

t.table = data.frame(Genotype1 = character(),
                     Genotype2 = character(),
                     pval = numeric())

i=1
while(i <= length(t.data)){
  j = i+1
  while(j <= length(t.data)){
    t.temp = data.frame(Genotype1 = geno[i],
                        Genotype2 = geno[j],
                        pval = t.test(x = t.data[[i]],
                                      y = t.data[[j]],
                                      paired = F)[[3]])
    t.table = rbind(t.table, t.temp)
      j = j+1
  }
  i = i+1
}

t.stripe = t.table$pval <= .05
knitr::kable(t.table,
             caption = 'All t-tests for NMJ 6-7, all segments Bouton Counts.')  %>%
kable_styling() %>%
row_spec(c(1:nrow(t.table))[t.table$pval <= .05], bold = T, color = "black", background = "gold")

```

```{r echo = F, fig.width = 10}

Boutons = temp[temp$Data == "Boutons",]

#boutdata = Boutons[Boutons$NMJ == "6-7" &
#                     (Boutons$Genotype == "WT(cr/LoxP)" | #Boutons$Genotype == "G85R(cr/hr)"),]

boutdata = Boutons[Boutons$NMJ == "6-7",]
boutdata$Genotype = paste(boutdata$Segment, boutdata$Genotype)

geno = unique(boutdata$Genotype)
geno = geno[c(1,5, 9:12, 
              2,6, 
              3,7, 13:16,
              4,8)]

boutdata$Genotype = factor(boutdata$Genotype, levels = geno)

boutmeans = c(mean(boutdata[boutdata$Genotype == geno[1], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[2], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[3], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[4], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[5], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[6], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[7], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[8], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[9], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[10], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[11], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[12], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[13], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[14], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[15], "Boutons"]),
              mean(boutdata[boutdata$Genotype == geno[16], "Boutons"]))

par(mar=c(4, 4.1, 4.1, 5.6))

plot(x = c(.75,1.25),
     y = rep(boutmeans[1], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1],
     ylim = c(0, max(boutdata$Boutons)*1.1),
     xlim = c(.5, length(na.omit(boutmeans))+.5),
     xaxt = "n",
     ylab="Bouton Number",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = paste("Larval muscle NMJ 6-7 bouton counts, segments 3-6"))

lines(x = c(.65,1.35),
     y = rep(boutmeans[1], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(1.65,2.35),
     y = rep(boutmeans[2], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

lines(x = c(2.65,3.35),
     y = rep(boutmeans[3], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[3])

lines(x = c(3.65,4.35),
     y = rep(boutmeans[4], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[4])

lines(x = c(4.65,5.35),
     y = rep(boutmeans[5], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[5])

lines(x = c(5.65,6.35),
     y = rep(boutmeans[6], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[6])

lines(x = c(6.65,7.35),
     y = rep(boutmeans[7], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(7.65,8.35),
     y = rep(boutmeans[8], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

lines(x = c(8.65,9.35),
     y = rep(boutmeans[9], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(9.65,10.35),
     y = rep(boutmeans[10], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

lines(x = c(10.65,11.35),
     y = rep(boutmeans[11], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[3])

lines(x = c(11.65,12.35),
     y = rep(boutmeans[12], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[4])

lines(x = c(12.65,13.35),
     y = rep(boutmeans[13], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[5])

lines(x = c(13.65,14.35),
     y = rep(boutmeans[14], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[6])

lines(x = c(14.65,15.35),
     y = rep(boutmeans[15], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(15.65,16.35),
     y = rep(boutmeans[16], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

beeswarm(boutdata$Boutons~boutdata$Genotype,
       pch = 21,
       method = "center",
       bg = hex(std.colors[c(1:6, 1:2, 1:6, 1:2)],.5),
       add = T)

lines(x = c(6.5,6.5),
      y = c(-2.5,70),
      lty = 2,
      xpd = T)

lines(x = c(8.5,8.5),
      y = c(-2.5,70),
      lty = 2,
      xpd = T)

lines(x = c(14.5,14.5),
      y = c(-2.5,70),
      lty = 2,
      xpd = T)

axis(side = 1,
     at = c(3.5, 7.5, 11.5, 15.5),
     labels = c(3,4,5,6),
     las = 1,
     lwd = 0,
     line = -.75)

axis(side = 1,
     at = c(8.5),
     labels = c("Larval Segment"),
     las = 1,
     lwd = 0,
     cex = 2,
     line = 0.5)

    lines(x = c(-.10, length(na.omit(boutmeans))+.5),
       y = c(0,0),
       lwd = 3,
       xpd = T)

  y.at = c(0, 10*1:7)

  axis(side = 2,
       at = y.at,
       lwd = 2,
       line = 0,
       labels = F,
       xpd = T,
       las = 2)

  axis(side = 2,
       at = y.at,
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)

  legend('topright',
       inset=c(-0.14,0),
       legend = c('WT(cr/LoxP)', 'G85R(cr/hr)', 
                  "OK371-gbb; WT(cr/LoxP)", "OK371-gbb; G85R(cr/LoxP)",
                  "OK371-saxA; WT(cr/LoxP)","OK371-saxA; G85R(cr/LoxP)"), 
       fill = std.colors[1:6],
       cex = .5,
       bty = 'n',
       pt.cex = .4,
       xpd = T)

```

```{r echo = F}

btd = boutdata

t.data = list("3 WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "3 WT(cr/LoxP)"],
              "3 G85R(cr/hr)" = btd$Boutons[btd$Genotype == "3 G85R(cr/hr)"],
              "3 OK371-gbb; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "3 OK371-gbb; WT(cr/LoxP)"],
              "3 OK371-gbb; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "3 OK371-gbb; G85R(cr/LoxP)"],
              "3 OK371-saxA; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "3 OK371-saxA; WT(cr/LoxP)"],
              "3 OK371-saxA; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "3 OK371-saxA; G85R(cr/LoxP)"],
              "4 WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "4 WT(cr/LoxP)"],
              "4 G85R(cr/hr)" = btd$Boutons[btd$Genotype == "4 G85R(cr/hr)"],
              "5 WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "5 WT(cr/LoxP)"],
              "5 G85R(cr/hr)" = btd$Boutons[btd$Genotype == "5 G85R(cr/hr)"],
              "5 OK371-gbb; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "5 OK371-gbb; WT(cr/LoxP)"],
              "5 OK371-gbb; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "5 OK371-gbb; G85R(cr/LoxP)"],
              "5 OK371-saxA; WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "5 OK371-saxA; WT(cr/LoxP)"],
              "5 OK371-saxA; G85R(cr/LoxP)" = btd$Boutons[btd$Genotype == "5 OK371-saxA; G85R(cr/LoxP)"],
              "6 WT(cr/LoxP)" = btd$Boutons[btd$Genotype == "6 WT(cr/LoxP)"],
              "6 G85R(cr/hr)" = btd$Boutons[btd$Genotype == "6 G85R(cr/hr)"])

t.table = data.frame(Genotype1 = character(),
                     Genotype2 = character(),
                     pval = numeric())

i=1
while(i <= length(t.data)){
  j = i+1
  while(j <= length(t.data)){
    t.temp = data.frame(Genotype1 = geno[i],
                        Genotype2 = geno[j],
                        pval = t.test(x = t.data[[i]],
                                      y = t.data[[j]],
                                      paired = F)[[3]])
    t.table = rbind(t.table, t.temp)
      j = j+1
  }
  i = i+1
}

t.stripe = t.table$pval <= .05
knitr::kable(t.table,
             caption = 'All t-tests for NMJ 4, all segments Bouton Counts.')  %>%
kable_styling() %>%
row_spec(c(1:nrow(t.table))[t.table$pval <= .05], bold = T, color = "black", background = "gold")

```

```{r echo = F}

VNC = temp[temp$Data == "VNC",]
VNC$CTCF = VNC$CTCF/1000
geno = unique(VNC$Genotype)
VNC$color = as.numeric(VNC$Genotype) + ((VNC$Experiment -1)*6)
VNC$color = c('navy', "orange2",
               "darkgreen", "turquoise4",
               "firebrick","deeppink3",
               "royalblue", 'gold',
               "green", 'cyan',
               "red", "pink") [VNC$color]

labels = c("WT", "G85R",
           "OK-gbb WT", "OK-gbb G85R",
           "OK-SaxA WT", "OK-SaxA G85R")

vncmeans = c(mean(VNC[VNC$Genotype == geno[1], "CTCF"]),
              mean(VNC[VNC$Genotype == geno[2], "CTCF"]),
              mean(VNC[VNC$Genotype == geno[3], "CTCF"]),
              mean(VNC[VNC$Genotype == geno[4], "CTCF"]),
              mean(VNC[VNC$Genotype == geno[5], "CTCF"]),
              mean(VNC[VNC$Genotype == geno[6], "CTCF"]))

par(mar=c(6.25, 4.1, 4.1, 5.6))

plot(x = c(.75,1.25),
     y = rep(vncmeans[1], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1],
     ylim = c(0, max(VNC$CTCF)*1.1),
     xlim = c(.5, length(na.omit(vncmeans))+.5),
     xaxt = "n",
     ylab="Fluorescence Intensity/1000",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = paste("Larval VNC fluorescence"))

lines(x = c(.65,1.35),
     y = rep(vncmeans[1], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(1.65,2.35),
     y = rep(vncmeans[2], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

if(length(na.omit(vncmeans)>2)){
lines(x = c(2.65,3.35),
     y = rep(vncmeans[3], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[3])

lines(x = c(3.65,4.35),
     y = rep(vncmeans[4], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[4])

lines(x = c(4.65,5.35),
     y = rep(vncmeans[5], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[5])

lines(x = c(5.65,6.35),
     y = rep(vncmeans[6], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[6])
}

beeswarm(VNC$CTCF~VNC$Genotype,
       #pch = 21,
       pwpch = 20+VNC$Experiment,
       method = "center",
       #bg = hex(unique(std.colors[as.numeric(VNC$Genotype)]),.5),
       pwbg = hex(VNC$color, .75),
       add = T)

axis(side = 1,
     at = c(1:length(na.omit(vncmeans))),
     labels = c(labels)[1:length(na.omit(vncmeans))],
     las = 2,
     lwd = 0,
     line = -.75)

    lines(x = c(.225 + .3/length(na.omit(vncmeans)), length(na.omit(vncmeans))+.5),
       y = c(0,0),
       lwd = 3,
       xpd = T)

  y.at = c(0,500, 1000, 1500)

  axis(side = 2,
       at = y.at,
       lwd = 2,
       line = 0,
       labels = F,
       xpd = T,
       las = 2)

  axis(side = 2,
       at = y.at,
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)

```

```{r echo = F}
btd = temp[temp$Data == "VNC",]

t.data = list("WT(cr/LoxP)" = btd$CTCF[btd$Genotype == "WT(cr/LoxP)"],
                    "G85R(cr/hr)" = btd$CTCF[btd$Genotype == "G85R(cr/hr)"],
                    "OK371-gbb; WT(cr/LoxP)" = btd$CTCF[btd$Genotype == "OK371-gbb; WT(cr/LoxP)"],
                    "OK371-gbb; G85R(cr/LoxP)" = btd$CTCF[btd$Genotype == "OK371-gbb; G85R(cr/LoxP)"],
                    "OK371-saxA; WT(cr/LoxP)" = btd$CTCF[btd$Genotype == "OK371-saxA; WT(cr/LoxP)"],
                    "OK371-saxA; G85R(cr/LoxP)" = btd$CTCF[btd$Genotype == "OK371-saxA; G85R(cr/LoxP)"])

t.table = data.frame(Genotype1 = character(),
                     Genotype2 = character(),
                     pval = numeric())

i=1
while(i <= length(t.data)){
  j = i+1
  while(j <= length(t.data)){
    t.temp = data.frame(Genotype1 = geno[i],
                        Genotype2 = geno[j],
                        pval = t.test(x = t.data[[i]],
                                      y = t.data[[j]],
                                      paired = F)[[3]])
    t.table = rbind(t.table, t.temp)
      j = j+1
  }
  i = i+1
}

t.stripe = t.table$pval <= .05
knitr::kable(t.table,
             caption = 'All t-tests for VNC data.')  %>%
kable_styling() %>%
row_spec(c(1:nrow(t.table))[t.table$pval <= .05], bold = T, color = "black", background = "gold")

```

```{r echo = F}

NMJ = temp[temp$Data == "NMJ" & temp$NMJ == "4",]
NMJ$CTCF = NMJ$CTCF/1000
geno = unique(NMJ$Genotype)
NMJ$color = as.numeric(NMJ$Genotype) + ((NMJ$Experiment -1)*6)
NMJ$color = c('navy', "orange2",
               "darkgreen", "turquoise4",
               "firebrick","deeppink3",
               "royalblue", 'gold',
               "green", 'cyan',
               "red", "pink") [NMJ$color]

labels = c("WT", "G85R",
           "OK-gbb WT", "OK-gbb G85R",
           "OK-SaxA WT", "OK-SaxA G85R")

nmjmeans = c(mean(NMJ[NMJ$Genotype == geno[1], "CTCF"]),
              mean(NMJ[NMJ$Genotype == geno[2], "CTCF"]),
              mean(NMJ[NMJ$Genotype == geno[3], "CTCF"]),
              mean(NMJ[NMJ$Genotype == geno[4], "CTCF"]),
              mean(NMJ[NMJ$Genotype == geno[5], "CTCF"]),
              mean(NMJ[NMJ$Genotype == geno[6], "CTCF"]))

par(mar=c(6.25, 4.1, 4.1, 5.6))

plot(x = c(.75,1.25),
     y = rep(nmjmeans[1], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1],
     ylim = c(0, max(NMJ$CTCF)*1.1),
     xlim = c(.5, length(na.omit(nmjmeans))+.5),
     xaxt = "n",
     ylab="Fluorescence Intensity/1000",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = paste("Larval segment 3, NMJ 4 fluorescence"))

lines(x = c(.65,1.35),
     y = rep(nmjmeans[1], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(1.65,2.35),
     y = rep(nmjmeans[2], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

if(length(na.omit(nmjmeans)>2)){
lines(x = c(2.65,3.35),
     y = rep(nmjmeans[3], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[3])

lines(x = c(3.65,4.35),
     y = rep(nmjmeans[4], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[4])

lines(x = c(4.65,5.35),
     y = rep(nmjmeans[5], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[5])

lines(x = c(5.65,6.35),
     y = rep(nmjmeans[6], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[6])
}

beeswarm(NMJ$CTCF~NMJ$Genotype,
       #pch = 21,
       pwpch = 20+NMJ$Experiment,
       method = "center",
       #bg = hex(unique(std.colors[as.numeric(NMJ$Genotype)]),.5),
       pwbg = hex(NMJ$color, .75),
       add = T)

axis(side = 1,
     at = c(1:length(na.omit(nmjmeans))),
     labels = c(labels)[1:length(na.omit(nmjmeans))],
     las = 2,
     lwd = 0,
     line = -.75)

    lines(x = c(.225 + .3/length(na.omit(nmjmeans)), length(na.omit(nmjmeans))+.5),
       y = c(0,0),
       lwd = 3,
       xpd = T)

  y.at = c(0, 500, 1000, 1500, 2000)

  axis(side = 2,
       at = y.at,
       lwd = 2,
       line = 0,
       labels = F,
       xpd = T,
       las = 2)

  axis(side = 2,
       at = y.at,
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)

```

```{r echo = F}
btd = temp[temp$Data == "NMJ" & temp$NMJ == "4",]

t.data = list("WT(cr/LoxP)" = btd$CTCF[btd$Genotype == "WT(cr/LoxP)"],
                    "G85R(cr/hr)" = btd$CTCF[btd$Genotype == "G85R(cr/hr)"],
                    "OK371-gbb; WT(cr/LoxP)" = btd$CTCF[btd$Genotype == "OK371-gbb; WT(cr/LoxP)"],
                    "OK371-gbb; G85R(cr/LoxP)" = btd$CTCF[btd$Genotype == "OK371-gbb; G85R(cr/LoxP)"],
                    "OK371-saxA; WT(cr/LoxP)" = btd$CTCF[btd$Genotype == "OK371-saxA; WT(cr/LoxP)"],
                    "OK371-saxA; G85R(cr/LoxP)" = btd$CTCF[btd$Genotype == "OK371-saxA; G85R(cr/LoxP)"])

t.table = data.frame(Genotype1 = character(),
                     Genotype2 = character(),
                     pval = numeric())

i=1
while(i <= length(t.data)){
  j = i+1
  while(j <= length(t.data)){
    t.temp = data.frame(Genotype1 = geno[i],
                        Genotype2 = geno[j],
                        pval = t.test(x = t.data[[i]],
                                      y = t.data[[j]],
                                      paired = F)[[3]])
    t.table = rbind(t.table, t.temp)
      j = j+1
  }
  i = i+1
}

t.stripe = t.table$pval <= .05
knitr::kable(t.table,
             caption = 'All t-tests for VNC data.')  %>%
kable_styling() %>%
row_spec(c(1:nrow(t.table))[t.table$pval <= .05], bold = T, color = "black", background = "gold")

```

```{r echo = F}

NMJ = temp[temp$Data == "NMJ" & temp$NMJ == "6-7",]
NMJ$CTCF = NMJ$CTCF/1000
geno = unique(NMJ$Genotype)
NMJ$color = as.numeric(NMJ$Genotype) + ((NMJ$Experiment -1)*6)
NMJ$color = c('navy', "orange2",
               "darkgreen", "turquoise4",
               "firebrick","deeppink3",
               "royalblue", 'gold',
               "green", 'cyan',
               "red", "pink") [NMJ$color]

labels = c("WT", "G85R",
           "OK-gbb WT", "OK-gbb G85R",
           "OK-SaxA WT", "OK-SaxA G85R")

nmjmeans = c(mean(NMJ[NMJ$Genotype == geno[1], "CTCF"]),
              mean(NMJ[NMJ$Genotype == geno[2], "CTCF"]),
              mean(NMJ[NMJ$Genotype == geno[3], "CTCF"]),
              mean(NMJ[NMJ$Genotype == geno[4], "CTCF"]),
              mean(NMJ[NMJ$Genotype == geno[5], "CTCF"]),
              mean(NMJ[NMJ$Genotype == geno[6], "CTCF"]))

par(mar=c(6.25, 4.1, 4.1, 5.6))

plot(x = c(.75,1.25),
     y = rep(nmjmeans[1], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1],
     ylim = c(0, max(NMJ$CTCF)*1.1),
     xlim = c(.5, length(na.omit(nmjmeans))+.5),
     xaxt = "n",
     ylab="Fluorescence Intensity/1000",
     yaxt = "n",
     xlab = NA,
     frame = F,
     main = paste("Larval segment 3, NMJ 6-7 fluorescence"))

lines(x = c(.65,1.35),
     y = rep(nmjmeans[1], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[1])

lines(x = c(1.65,2.35),
     y = rep(nmjmeans[2], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[2])

if(length(na.omit(nmjmeans)>2)){
lines(x = c(2.65,3.35),
     y = rep(nmjmeans[3], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[3])

lines(x = c(3.65,4.35),
     y = rep(nmjmeans[4], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[4])

lines(x = c(4.65,5.35),
     y = rep(nmjmeans[5], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[5])

lines(x = c(5.65,6.35),
     y = rep(nmjmeans[6], 2),
     type = 'l',
     lwd = 3,
     col = std.colors[6])
}

beeswarm(NMJ$CTCF~NMJ$Genotype,
       #pch = 21,
       pwpch = 20+NMJ$Experiment,
       method = "center",
       #bg = hex(unique(std.colors[as.numeric(NMJ$Genotype)]),.5),
       pwbg = hex(NMJ$color, .75),
       add = T)

axis(side = 1,
     at = c(1:length(na.omit(nmjmeans))),
     labels = c(labels)[1:length(na.omit(nmjmeans))],
     las = 2,
     lwd = 0,
     line = -.75)

    lines(x = c(.225 + .3/length(na.omit(nmjmeans)), length(na.omit(nmjmeans))+.5),
       y = c(0,0),
       lwd = 3,
       xpd = T)

  y.at = c(0, 1000, 2000, 3000, 4000, 5000)

  axis(side = 2,
       at = y.at,
       lwd = 2,
       line = 0,
       labels = F,
       xpd = T,
       las = 2)

  axis(side = 2,
       at = y.at,
       lwd = 3,
       line = 0,
       labels = T,
       xpd = T,
       las = 2)

```

```{r echo = F}
btd = temp[temp$Data == "NMJ" & temp$NMJ == "6-7",]

t.data = list("WT(cr/LoxP)" = btd$CTCF[btd$Genotype == "WT(cr/LoxP)"],
                    "G85R(cr/hr)" = btd$CTCF[btd$Genotype == "G85R(cr/hr)"],
                    "OK371-gbb; WT(cr/LoxP)" = btd$CTCF[btd$Genotype == "OK371-gbb; WT(cr/LoxP)"],
                    "OK371-gbb; G85R(cr/LoxP)" = btd$CTCF[btd$Genotype == "OK371-gbb; G85R(cr/LoxP)"],
                    "OK371-saxA; WT(cr/LoxP)" = btd$CTCF[btd$Genotype == "OK371-saxA; WT(cr/LoxP)"],
                    "OK371-saxA; G85R(cr/LoxP)" = btd$CTCF[btd$Genotype == "OK371-saxA; G85R(cr/LoxP)"])

t.table = data.frame(Genotype1 = character(),
                     Genotype2 = character(),
                     pval = numeric())

i=1
while(i <= length(t.data)){
  j = i+1
  while(j <= length(t.data)){
    t.temp = data.frame(Genotype1 = geno[i],
                        Genotype2 = geno[j],
                        pval = t.test(x = t.data[[i]],
                                      y = t.data[[j]],
                                      paired = F)[[3]])
    t.table = rbind(t.table, t.temp)
      j = j+1
  }
  i = i+1
}

t.stripe = t.table$pval <= .05
knitr::kable(t.table,
             caption = 'All t-tests for VNC data.')  %>%
kable_styling() %>%
row_spec(c(1:nrow(t.table))[t.table$pval <= .05], bold = T, color = "black", background = "gold")

```

```{r echo = F, include=F}
library(corrplot)
cor.table = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/20250701_AllAnaData_CorrelationTable.csv",header = F)
row.names(cor.table) = paste0(cor.table[,1], ".", cor.table[,2], ".", cor.table[,3])
colnames(cor.table) = paste0(cor.table[1,], ".", cor.table[2,], ".", cor.table[3,], ".", cor.table[4,])

cor.table= cor.table[,-grep("Raw", colnames(cor.table))]

cor.table[1,grep("Norm", colnames(cor.table))] = substr(cor.table[1,grep("Norm", colnames(cor.table))], 8, nchar(cor.table[1,grep("Norm", colnames(cor.table))]))

cor.matrix = cor.table[5:nrow(cor.table), 4:ncol(cor.table)]
cor.matrix = cor.matrix[grep("WT", row.names(cor.matrix)), ]

i = 1
while(i<=ncol(cor.matrix)){
  cor.matrix[,i] = as.numeric(cor.matrix[,i])
  i = i + 1
}

cor.data = as.matrix(cor.matrix)
cor.data[,1:ncol(cor.data)] = 0
cor.data = cor.data[1:ncol(cor.data),]
row.names(cor.data) = colnames(cor.data)

cor.data = data.frame(Factor1 = character(),
                      Factor2 = character(),
                      Corr = numeric())
i=1
while(i<=ncol(cor.matrix)){
  j=1
  while(j<=ncol(cor.matrix)){
    temp = na.omit(cor.matrix[,c(i,j)])
    #cor.data[j,i] = cor(temp[,1], temp[,2])
    temp.cor.data = data.frame(Factor1 = colnames(cor.matrix)[i],
                               Factor2 = colnames(cor.matrix)[j],
                               Corr = cor(temp[,1], temp[,2]))
    cor.data = rbind(cor.data, temp.cor.data)
    j=j+1
  }
  i=i+1
}

cor.data[cor.data$Factor1 == cor.data$Factor2, "Corr"] = NA

cor.data$Factor1 = factor(cor.data$Factor1, levels = unique(cor.data$Factor1))
cor.data$Factor2 = factor(cor.data$Factor2, levels = unique(cor.data$Factor2))

WT.cor.data = cor.data

cor.matrix = cor.table[5:nrow(cor.table), 4:ncol(cor.table)]
cor.matrix = cor.matrix[grep("G85R", row.names(cor.matrix)), ]

i = 1
while(i<=ncol(cor.matrix)){
  cor.matrix[,i] = as.numeric(cor.matrix[,i])
  i = i + 1
}

cor.data = as.matrix(cor.matrix)
cor.data[,1:ncol(cor.data)] = 0
cor.data = cor.data[1:ncol(cor.data),]
row.names(cor.data) = colnames(cor.data)

cor.data = data.frame(Factor1 = character(),
                      Factor2 = character(),
                      Corr = numeric())
i=1
while(i<=ncol(cor.matrix)){
  j=1
  while(j<=ncol(cor.matrix)){
    temp = na.omit(cor.matrix[,c(i,j)])
    #cor.data[j,i] = cor(temp[,1], temp[,2])
    temp.cor.data = data.frame(Factor1 = colnames(cor.matrix)[i],
                               Factor2 = colnames(cor.matrix)[j],
                               Corr = cor(temp[,1], temp[,2]))
    cor.data = rbind(cor.data, temp.cor.data)
    j=j+1
  }
  i=i+1
}

cor.data[cor.data$Factor1 == cor.data$Factor2, "Corr"] = NA

cor.data$Factor1 = factor(cor.data$Factor1, levels = unique(cor.data$Factor1))
cor.data$Factor2 = factor(cor.data$Factor2, levels = unique(cor.data$Factor2))

GR.cor.data = cor.data

WT.cor.data = WT.cor.data[as.numeric(WT.cor.data$Factor1) < as.numeric(WT.cor.data$Factor2),]
WT.cor.data$Geno = "WT"
GR.cor.data = GR.cor.data[as.numeric(GR.cor.data$Factor1) > as.numeric(GR.cor.data$Factor2),]
GR.cor.data$Geno = "G85R"

cor.data = rbind(WT.cor.data, GR.cor.data)

```

```{r echo = F, fig.height=8, fig.width=8, include=F}
bg.colors = colorRampPalette(c("royalblue", "deepskyblue", "yellow", "deeppink", "violetred"))(200)

plot.bg = bg.colors[100*(cor.data$Corr + 1)]

#plot.bg[cor.data$Corr == 2] = "white"

plot.cex = 2*abs(cor.data$Corr)+.5

#plot.cex[cor.data$Corr == 2] = 2.5

par(mar = c(6,6,4,3))

plot(x = as.numeric(cor.data$Factor1),
     y = as.numeric(cor.data$Factor2),
     pch = 21,
     bg = plot.bg,
     cex = plot.cex,
     xaxt = "n",
     yaxt = "n",
     ylab = "",
     xlab = "",
     bty = "n",
     main = "")

lines(x = c(.5,23),
      y = c(.5,23),
      lty = 2,
      lwd = 2)

lines(x = c(.5, .5),
      y = c(.5,21.5),
      lty = 1,
      lwd = 1,
      xpd = T)

lines(x = c(-0.25, -0.25),
      y = c(-.25,21.5),
      lty = 2,
      lwd = .5,
      xpd = T)

lines(x = c(-1.25, -1.25),
      y = c(-1.25,21.5),
      lty = 2,
      lwd = .5,
      xpd = T)

lines(x = c(-2, -2),
      y = c(-2,21.5),
      lty = 2,
      lwd = .5,
      xpd = T)

lines(x = c(.5,21.5),
      y = c(.5, .5),
      lty = 1,
      lwd = 1,
      xpd = T)

lines(y = c(-.25, -.25),
      x = c(-.25,21.5),
      lty = 2,
      lwd = .5,
      xpd = T)

text(x = 22,
     y = 0,
     labels = "Segment",
     xpd = T,
     cex = .5)

text(x = 22,
     y = -.75,
     labels = "NMJ",
     xpd = T,
     cex = .5)

text(x = 22,
     y = -1.6,
     labels = "Side",
     xpd = T,
     cex = .5)

lines(y = c(-1.25, -1.25),
      x = c(-1.25,21.5),
      lty = 2,
      lwd = .5,
      xpd = T)

lines(y = c(-2, -2),
      x = c(-2,21.5),
      lty = 2,
      lwd = .5,
      xpd = T)

axis(side = 1,
     at = c(8.5, 17, 19.5),
     lwd = 0,
     line = 2,
     labels = c("Boutons", "VNC", "NMJ"),
     xpd = T,
     las = 1,
     cex.axis = .75)

axis(side = 1,
     at = c(17, 19.5),
     lwd = 0,
     line = 3,
     labels = c("pMad", "pMad"),
     xpd = T,
     las = 1,
     cex.axis = .75)

axis(side = 1,
     at = c(1:21),
     lwd = 0,
     line = 1,
     labels = cor.table[2, 4:ncol(cor.table)],
     xpd = T,
     las = 1,
     cex.axis = .5)

axis(side = 1,
     at = c(1:21),
     lwd = 0,
     line = -.25,
     labels = cor.table[3, 4:ncol(cor.table)],
     xpd = T,
     las = 1,
     cex.axis = .5)

axis(side = 1,
     at = c(1:21),
     lwd = 0,
     line = -1.5,
     labels = cor.table[4, 4:ncol(cor.table)],
     xpd = T,
     las = 1,
     cex.axis = .5)

axis(side = 2,
     at = c(8.5, 17, 19.5),
     lwd = 0,
     line = 3,
     labels = c("Boutons", "VNC", "NMJ"),
     xpd = T,
     srt = 270,
     cex.axis = .75)

axis(side = 2,
     at = c(17, 19.5),
     lwd = 0,
     line = 2.5,
     labels = c("pMad", "pMad"),
     xpd = T,
     srt = 270,
     cex.axis = .75)

axis(side = 2,
     at = c(1:21),
     lwd = 0,
     line = 1.5,
     labels = cor.table[2, 4:ncol(cor.table)],
     xpd = T,
     las = 1,
     cex.axis = .5)

axis(side = 2,
     at = c(1:21),
     lwd = 0,
     line = 0.25,
     labels = cor.table[3, 4:ncol(cor.table)],
     xpd = T,
     las = 1,
     cex.axis = .5)

axis(side = 2,
     at = c(1:21),
     lwd = 0,
     line = -1,
     labels = cor.table[4, 4:ncol(cor.table)],
     xpd = T,
     las = 1,
     cex.axis = .5)

arrows(x0 = 22,
       y0 = 21,
       x1 = 22,
       y1 = 13,
       lwd = 2,
       xpd = T,
       length = .1)

arrows(x0 = 21,
       y0 = 22,
       x1 = 13,
       y1 = 22,
       lwd = 2,
       xpd = T,
       length = .1)

text(x = 17,
     y = 22.5,
     labels = ("Silent"),
     xpd = T)

text(x = 22.5,
     y = 17,
     labels = ("G85R"),
     srt = 270,
     xpd = T)

axis(side = 3,
     label = "Correlation for WT and G85R, All Lines",
     at = 11,
     line = 1.5,
     cex = 1.5,
     lwd = 0)

```

```{r echo = F, include=F}

cor.table = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/20250701_AllAnaData_CorrelationTable.csv",header = F)

cor.table= cor.table[,-grep("Raw", cor.table[1,])]

cor.table[1,grep("Norm", cor.table[1,])] = substr(cor.table[1,grep("Norm", cor.table[1,])], 8, nchar(cor.table[1,grep("Norm", cor.table[1,])]))

cor.tablea = cor.table[,cor.table[2,] == "a"]
cor.tablea = cbind(cor.tablea, cor.table[,20])
cor.tablea = cbind(cor.tablea, cor.table[,1])

cor.tableb = cor.table[,cor.table[2,] == "b"]
cor.tableb = cbind(cor.tableb, cor.table[,20])
cor.tableb = cbind(cor.tableb, cor.table[,1])

colnames(cor.tablea) = paste0(cor.tablea[1,], ".", cor.tablea[3,], ".", cor.tablea[4,])
colnames(cor.tableb) = paste0(cor.tableb[1,], ".", cor.tableb[3,], ".", cor.tableb[4,])

cor.table = rbind(cor.tablea[c(1,3:nrow(cor.tablea)),],cor.tableb[5:nrow(cor.tableb),])

cor.matrix = cor.table[grep("WT", cor.table[,ncol(cor.table)]), 1:(ncol(cor.table)-1)]

i = 1
while(i<=ncol(cor.matrix)){
  cor.matrix[,i] = as.numeric(cor.matrix[,i])
  i = i + 1
}

cor.data = as.matrix(cor.matrix)
cor.data[,1:ncol(cor.data)] = 0
cor.data = cor.data[1:ncol(cor.data),]
row.names(cor.data) = colnames(cor.data)

cor.data = data.frame(Factor1 = character(),
                      Factor2 = character(),
                      Corr = numeric())
i=1
while(i<=ncol(cor.matrix)){
  j=1
  while(j<=ncol(cor.matrix)){
    temp = na.omit(cor.matrix[,c(i,j)])
    #cor.data[j,i] = cor(temp[,1], temp[,2])
    temp.cor.data = data.frame(Factor1 = colnames(cor.matrix)[i],
                               Factor2 = colnames(cor.matrix)[j],
                               Corr = cor(temp[,1], temp[,2]))
    cor.data = rbind(cor.data, temp.cor.data)
    j=j+1
  }
  i=i+1
}

cor.data[cor.data$Factor1 == cor.data$Factor2, "Corr"] = NA

cor.data$Factor1 = factor(cor.data$Factor1, levels = unique(cor.data$Factor1))
cor.data$Factor2 = factor(cor.data$Factor2, levels = unique(cor.data$Factor2))

WT.cor.data = cor.data

cor.matrix = cor.table[grep("G85R", cor.table[,ncol(cor.table)]), 1:(ncol(cor.table)-1)]

i = 1
while(i<=ncol(cor.matrix)){
  cor.matrix[,i] = as.numeric(cor.matrix[,i])
  i = i + 1
}

cor.data = as.matrix(cor.matrix)
cor.data[,1:ncol(cor.data)] = 0
cor.data = cor.data[1:ncol(cor.data),]
row.names(cor.data) = colnames(cor.data)

cor.data = data.frame(Factor1 = character(),
                      Factor2 = character(),
                      Corr = numeric())
i=1
while(i<=ncol(cor.matrix)){
  j=1
  while(j<=ncol(cor.matrix)){
    temp = na.omit(cor.matrix[,c(i,j)])
    #cor.data[j,i] = cor(temp[,1], temp[,2])
    temp.cor.data = data.frame(Factor1 = colnames(cor.matrix)[i],
                               Factor2 = colnames(cor.matrix)[j],
                               Corr = cor(temp[,1], temp[,2]))
    cor.data = rbind(cor.data, temp.cor.data)
    j=j+1
  }
  i=i+1
}

cor.data[cor.data$Factor1 == cor.data$Factor2, "Corr"] = NA

cor.data$Factor1 = factor(cor.data$Factor1, levels = unique(cor.data$Factor1))
cor.data$Factor2 = factor(cor.data$Factor2, levels = unique(cor.data$Factor2))

GR.cor.data = cor.data

WT.cor.data = WT.cor.data[as.numeric(WT.cor.data$Factor1) < as.numeric(WT.cor.data$Factor2),]
WT.cor.data$Geno = "WT"
GR.cor.data = GR.cor.data[as.numeric(GR.cor.data$Factor1) > as.numeric(GR.cor.data$Factor2),]
GR.cor.data$Geno = "G85R"

cor.data = rbind(WT.cor.data, GR.cor.data)

```

```{r echo = F, fig.height=8, fig.width=8, include=F}
bg.colors = colorRampPalette(c("royalblue", "deepskyblue", "yellow", "deeppink", "violetred"))(200)

plot.bg = bg.colors[100*(cor.data$Corr + 1)]

#plot.bg[cor.data$Corr == 2] = "white"

plot.cex = 2*abs(cor.data$Corr)+.5

#plot.cex[cor.data$Corr == 2] = 2.5

par(mar = c(8,8,4,3))

plot(x = as.numeric(cor.data$Factor1),
     y = as.numeric(cor.data$Factor2),
     pch = 21,
     bg = plot.bg,
     cex = plot.cex,
     xaxt = "n",
     yaxt = "n",
     ylab = "",
     xlab = "",
     bty = "n",
     main = "")

lines(x = c(.5,14),
      y = c(.5,14),
      lty = 2,
      lwd = 2)

lines(x = c(.5, .5),
      y = c(.5,11.5),
      lty = 1,
      lwd = 1,
      xpd = T)

lines(x = c(.5,11.5),
      y = c(.5, .5),
      lty = 1,
      lwd = 1,
      xpd = T)

lines(x = c(-0.25, -0.25),
      y = c(-.25,11.5),
      lty = 2,
      lwd = .5,
      xpd = T)

lines(x = c(-1, -1),
      y = c(-1,11.5),
      lty = 2,
      lwd = .5,
      xpd = T)

lines(y = c(-.25, -.25),
      x = c(-.25,11.5),
      lty = 2,
      lwd = .5,
      xpd = T)

lines(y = c(-1, -1),
      x = c(-1,11.5),
      lty = 2,
      lwd = .5,
      xpd = T)

text(x = 11.8,
     y = 0,
     labels = "NMJ",
     xpd = T,
     cex = .8)

text(x = 11.8,
     y = -.5,
     labels = "Segment",
     xpd = T,
     cex = .8)

arrows(x0 = 11.5,
       y0 = 10.5,
       x1 = 11.5,
       y1 = 7,
       lwd = 2,
       xpd = T,
       length = .1)

arrows(x0 = 10.5,
       y0 = 11.5,
       x1 = 7,
       y1 = 11.5,
       lwd = 2,
       xpd = T,
       length = .1)

text(x = 10,
     y = 11.75,
     labels = ("Silent"),
     xpd = T)

text(x = 11.75,
     y = 10,
     labels = ("G85R"),
     srt = 270,
     xpd = T)

axis(side = 1,
     at = c(4.5, 9.5, 11),
     lwd = 0,
     line = 3.5,
     labels = c("Boutons", "NMJ", "VNC"),
     xpd = T,
     las = 1,
     cex.axis = 1)

axis(side = 1,
     at = c(9.5, 11),
     lwd = 0,
     line = 5,
     labels = c("pMad", "pMad"),
     xpd = T,
     las = 1,
     cex.axis = 1)

axis(side = 2,
     at = c(9.5, 11),
     lwd = 0,
     line = 5,
     labels = c("NMJ", "VNC"),
     xpd = T,
     srt = 270,
     cex.axis = 1)

axis(side = 2,
     at = c(4.5, 9.5, 11),
     lwd = 0,
     line = 3.5,
     labels = c("Boutons", "pMad", "pMad"),
     xpd = T,
     srt = 270,
     cex.axis = 1)

axis(side = 1,
     at = c(1:11),
     lwd = 0,
     line = 0,
     labels = cor.table[2, 1:11],
     xpd = T,
     las = 1,
     cex.axis = 1)

axis(side = 1,
     at = c(1:11),
     lwd = 0,
     line = 2,
     labels = cor.table[3, 1:11],
     xpd = T,
     las = 1,
     cex.axis = 1)

axis(side = 2,
     at = c(1:11),
     lwd = 0,
     line = -0.25,
     labels = cor.table[2, 1:11],
     xpd = T,
     las = 1,
     cex.axis = 1)

axis(side = 2,
     at = c(1:11),
     lwd = 0,
     line = 2,
     labels = cor.table[3, 1:11],
     xpd = T,
     las = 1,
     cex.axis = 1)

axis(side = 3,
     label = "Correlation for WT and G85R, All Lines",
     at = 6,
     line = 1.5,
     cex = 1.5,
     lwd = 0)

```

```{r echo = F, include=F}

cor.table = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/20250701_AllAnaData_CorrelationTable.csv",header = F)

cor.table= cor.table[,-grep("Raw", cor.table[1,])]

cor.table[1,grep("Norm", cor.table[1,])] = substr(cor.table[1,grep("Norm", cor.table[1,])], 8, nchar(cor.table[1,grep("Norm", cor.table[1,])]))

cor.tablea = cor.table[,cor.table[2,] == "a"]
cor.tablea = cbind(cor.tablea, cor.table[,20])
cor.tablea = cbind(cor.tablea, cor.table[,1])

cor.tableb = cor.table[,cor.table[2,] == "b"]
cor.tableb = cbind(cor.tableb, cor.table[,20])
cor.tableb = cbind(cor.tableb, cor.table[,1])

colnames(cor.tablea) = paste0(cor.tablea[1,], ".", cor.tablea[3,], ".", cor.tablea[4,])
colnames(cor.tableb) = paste0(cor.tableb[1,], ".", cor.tableb[3,], ".", cor.tableb[4,])

cor.table = rbind(cor.tablea[c(1,3:nrow(cor.tablea)),],cor.tableb[5:nrow(cor.tableb),])

##cor.matrix = cor.table[grep("WT-cr/LoxP", cor.table[,ncol(cor.table)]), 1:(ncol(cor.table)-1)]

cor.matrix = cor.table[cor.table[,ncol(cor.table)] == "WT-cr/LoxP", 1:(ncol(cor.table)-1)]

i = 1
while(i<=ncol(cor.matrix)){
  cor.matrix[,i] = as.numeric(cor.matrix[,i])
  i = i + 1
}

cor.data = as.matrix(cor.matrix)
cor.data[,1:ncol(cor.data)] = 0
cor.data = cor.data[1:ncol(cor.data),]
row.names(cor.data) = colnames(cor.data)

cor.data = data.frame(Factor1 = character(),
                      Factor2 = character(),
                      Corr = numeric())
i=1
while(i<=ncol(cor.matrix)){
  j=1
  while(j<=ncol(cor.matrix)){
    temp = na.omit(cor.matrix[,c(i,j)])
    #cor.data[j,i] = cor(temp[,1], temp[,2])
    temp.cor.data = data.frame(Factor1 = colnames(cor.matrix)[i],
                               Factor2 = colnames(cor.matrix)[j],
                               Corr = cor(temp[,1], temp[,2]))
    cor.data = rbind(cor.data, temp.cor.data)
    j=j+1
  }
  i=i+1
}

cor.data[cor.data$Factor1 == cor.data$Factor2, "Corr"] = NA

cor.data$Factor1 = factor(cor.data$Factor1, levels = unique(cor.data$Factor1))
cor.data$Factor2 = factor(cor.data$Factor2, levels = unique(cor.data$Factor2))

WT.cor.data = cor.data

#cor.matrix = cor.table[grep("G85Rcr/G85Rhr", cor.table[,ncol(cor.table)]), 1:(ncol(cor.table)-1)]

cor.matrix = cor.table[cor.table[,ncol(cor.table)] == "G85Rcr/G85Rhr", 1:(ncol(cor.table)-1)]

i = 1
while(i<=ncol(cor.matrix)){
  cor.matrix[,i] = as.numeric(cor.matrix[,i])
  i = i + 1
}

cor.data = as.matrix(cor.matrix)
cor.data[,1:ncol(cor.data)] = 0
cor.data = cor.data[1:ncol(cor.data),]
row.names(cor.data) = colnames(cor.data)

cor.data = data.frame(Factor1 = character(),
                      Factor2 = character(),
                      Corr = numeric())
i=1
while(i<=ncol(cor.matrix)){
  j=1
  while(j<=ncol(cor.matrix)){
    temp = na.omit(cor.matrix[,c(i,j)])
    #cor.data[j,i] = cor(temp[,1], temp[,2])
    temp.cor.data = data.frame(Factor1 = colnames(cor.matrix)[i],
                               Factor2 = colnames(cor.matrix)[j],
                               Corr = cor(temp[,1], temp[,2]))
    cor.data = rbind(cor.data, temp.cor.data)
    j=j+1
  }
  i=i+1
}

cor.data[cor.data$Factor1 == cor.data$Factor2, "Corr"] = NA

cor.data$Factor1 = factor(cor.data$Factor1, levels = unique(cor.data$Factor1))
cor.data$Factor2 = factor(cor.data$Factor2, levels = unique(cor.data$Factor2))

GR.cor.data = cor.data

WT.cor.data = WT.cor.data[as.numeric(WT.cor.data$Factor1) < as.numeric(WT.cor.data$Factor2),]
WT.cor.data$Geno = "WT"
GR.cor.data = GR.cor.data[as.numeric(GR.cor.data$Factor1) > as.numeric(GR.cor.data$Factor2),]
GR.cor.data$Geno = "G85R"

cor.data = rbind(WT.cor.data, GR.cor.data)

```

```{r echo = F, fig.height=8, fig.width=8, include=F}
bg.colors = colorRampPalette(c("royalblue", "deepskyblue", "yellow", "deeppink", "violetred"))(200)

plot.bg = bg.colors[100*(cor.data$Corr + 1)]

#plot.bg[cor.data$Corr == 2] = "white"

plot.cex = 2*abs(cor.data$Corr)+.5

#plot.cex[cor.data$Corr == 2] = 2.5

par(mar = c(8,8,4,3))

plot(x = as.numeric(cor.data$Factor1),
     y = as.numeric(cor.data$Factor2),
     pch = 21,
     bg = plot.bg,
     cex = plot.cex,
     xaxt = "n",
     yaxt = "n",
     ylab = "",
     xlab = "",
     bty = "n",
     main = "")

lines(x = c(.5,14),
      y = c(.5,14),
      lty = 2,
      lwd = 2)

lines(x = c(.5, .5),
      y = c(.5,11.5),
      lty = 1,
      lwd = 1,
      xpd = T)

lines(x = c(.5,11.5),
      y = c(.5, .5),
      lty = 1,
      lwd = 1,
      xpd = T)

lines(x = c(-0.25, -0.25),
      y = c(-.25,11.5),
      lty = 2,
      lwd = .5,
      xpd = T)

lines(x = c(-1, -1),
      y = c(-1,11.5),
      lty = 2,
      lwd = .5,
      xpd = T)

lines(y = c(-.25, -.25),
      x = c(-.25,11.5),
      lty = 2,
      lwd = .5,
      xpd = T)

lines(y = c(-1, -1),
      x = c(-1,11.5),
      lty = 2,
      lwd = .5,
      xpd = T)

text(x = 11.8,
     y = 0,
     labels = "NMJ",
     xpd = T,
     cex = .8)

text(x = 11.8,
     y = -.5,
     labels = "Segment",
     xpd = T,
     cex = .8)

arrows(x0 = 11.5,
       y0 = 10.5,
       x1 = 11.5,
       y1 = 7,
       lwd = 2,
       xpd = T,
       length = .1)

arrows(x0 = 10.5,
       y0 = 11.5,
       x1 = 7,
       y1 = 11.5,
       lwd = 2,
       xpd = T,
       length = .1)

text(x = 10,
     y = 11.75,
     labels = ("Silent"),
     xpd = T)

text(x = 11.75,
     y = 10,
     labels = ("G85R"),
     srt = 270,
     xpd = T)

axis(side = 1,
     at = c(4.5, 9.5, 11),
     lwd = 0,
     line = 3.5,
     labels = c("Boutons", "NMJ", "VNC"),
     xpd = T,
     las = 1,
     cex.axis = 1)

axis(side = 1,
     at = c(9.5, 11),
     lwd = 0,
     line = 5,
     labels = c("pMad", "pMad"),
     xpd = T,
     las = 1,
     cex.axis = 1)

axis(side = 2,
     at = c(9.5, 11),
     lwd = 0,
     line = 5,
     labels = c("NMJ", "VNC"),
     xpd = T,
     srt = 270,
     cex.axis = 1)

axis(side = 2,
     at = c(4.5, 9.5, 11),
     lwd = 0,
     line = 3.5,
     labels = c("Boutons", "pMad", "pMad"),
     xpd = T,
     srt = 270,
     cex.axis = 1)

axis(side = 1,
     at = c(1:11),
     lwd = 0,
     line = 0,
     labels = cor.table[2, 1:11],
     xpd = T,
     las = 1,
     cex.axis = 1)

axis(side = 1,
     at = c(1:11),
     lwd = 0,
     line = 2,
     labels = cor.table[3, 1:11],
     xpd = T,
     las = 1,
     cex.axis = 1)

axis(side = 2,
     at = c(1:11),
     lwd = 0,
     line = -0.25,
     labels = cor.table[2, 1:11],
     xpd = T,
     las = 1,
     cex.axis = 1)

axis(side = 2,
     at = c(1:11),
     lwd = 0,
     line = 2,
     labels = cor.table[3, 1:11],
     xpd = T,
     las = 1,
     cex.axis = 1)

axis(side = 3,
     label = "Correlation for WT and G85R, No Rescues",
     at = 6,
     line = 1.5,
     cex = 1.5,
     lwd = 0)

```

```{r echo = F, fig.height= 8, fig.width= 8, include=F}

cor.table = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/20250701_AllAnaData_CorrelationTable.csv",header = F)

cor.table= cor.table[,-grep("Raw", cor.table[1,])]

cor.table[1,grep("Norm", cor.table[1,])] = substr(cor.table[1,grep("Norm", cor.table[1,])], 8, nchar(cor.table[1,grep("Norm", cor.table[1,])]))

cor.tablea = cor.table[,cor.table[2,] == "a"]
cor.tablea = cbind(cor.tablea, cor.table[,20])
cor.tablea = cbind(cor.tablea, cor.table[,1])

cor.tableb = cor.table[,cor.table[2,] == "b"]
cor.tableb = cbind(cor.tableb, cor.table[,20])
cor.tableb = cbind(cor.tableb, cor.table[,1])

colnames(cor.tablea) = paste0(cor.tablea[1,], ".", cor.tablea[3,], ".", cor.tablea[4,])
colnames(cor.tableb) = paste0(cor.tableb[1,], ".", cor.tableb[3,], ".", cor.tableb[4,])

cor.table = rbind(cor.tablea[c(1,3:nrow(cor.tablea)),],cor.tableb[5:nrow(cor.tableb),])

cor.matrix = cor.table[cor.table[,ncol(cor.table)] == "WT-cr/LoxP" |
                         cor.table[,ncol(cor.table)] == "G85Rcr/G85Rhr",
                       1:ncol(cor.table)]

temp = data.frame(NMJ4 = as.numeric(cor.matrix[cor.matrix[,12] == "WT-cr/LoxP",
                  c("Boutons.4.3")]),
                  NMJ67 = as.numeric(cor.matrix[cor.matrix[,12] == "WT-cr/LoxP",
                  c("Boutons.6-7.3")]))

temp.cor1 = cor(na.omit(temp)[,1], na.omit(temp)[,2])

temp.lm = lm(na.omit(temp)[,1] ~ na.omit(temp)[,2])[[1]]

#temp[is.na(temp)] = 0

temp = na.omit(temp)

plot(x = temp$NMJ67,
     y = temp$NMJ4,
     pch = 21,
     bg = "deepskyblue",
     cex = 1,
     #xaxt = "n",
     #yaxt = "n",
     ylab = "",
     xlab = "",
     bty = "n",
     main = "",
     xlim = c(0, 100),
     ylim = c(0,50))

abline(a = temp.lm[1],
       b = temp.lm[2],
       lwd = 2,
       col = "royalblue")

temp = data.frame(NMJ4 = as.numeric(cor.matrix[cor.matrix[,12] == "WT-cr/LoxP",
                  c("Boutons.4.4")]),
                  NMJ67 = as.numeric(cor.matrix[cor.matrix[,12] == "WT-cr/LoxP",
                  c("Boutons.6-7.4")]))

temp.cor2 = cor(na.omit(temp)[,1], na.omit(temp)[,2])

temp.lm = lm(na.omit(temp)[,1] ~ na.omit(temp)[,2])[[1]]

points(x = temp$NMJ67,
       y = temp$NMJ4,
       pch = 21,
       bg = "deeppink",
       cex = 1)

abline(a = temp.lm[1],
       b = temp.lm[2],
       lwd = 2,
       col = "firebrick")

temp = data.frame(NMJ4 = as.numeric(cor.matrix[cor.matrix[,12] == "WT-cr/LoxP",
                  c("Boutons.4.5")]),
                  NMJ67 = as.numeric(cor.matrix[cor.matrix[,12] == "WT-cr/LoxP",
                  c("Boutons.6-7.5")]))

temp.cor3 = cor(na.omit(temp)[,1], na.omit(temp)[,2])

temp.lm = lm(na.omit(temp)[,1] ~ na.omit(temp)[,2])[[1]]

points(x = temp$NMJ67,
       y = temp$NMJ4,
       pch = 21,
       bg = "lightyellow",
       cex = 1)

abline(a = temp.lm[1],
       b = temp.lm[2],
       lwd = 2,
       col = "gold")

temp = data.frame(NMJ4 = as.numeric(cor.matrix[cor.matrix[,12] == "WT-cr/LoxP",
                  c("Boutons.4.6")]),
                  NMJ67 = as.numeric(cor.matrix[cor.matrix[,12] == "WT-cr/LoxP",
                  c("Boutons.6-7.6")]))

temp.cor4 = cor(na.omit(temp)[,1], na.omit(temp)[,2])

temp.lm = lm(na.omit(temp)[,1] ~ na.omit(temp)[,2])[[1]]

points(x = temp$NMJ67,
       y = temp$NMJ4,
       pch = 21,
       bg = "lightgreen",
       cex = 1)

abline(a = temp.lm[1],
       b = temp.lm[2],
       lwd = 2,
       col = "darkgreen")

legend("topleft",
       legend = c(paste0("3 (cor: ", signif(temp.cor1,2),")"), 
                  paste0("4 (cor: ", signif(temp.cor2,2),")"),
                  paste0("5 (cor: ", signif(temp.cor3,2),")"),
                  paste0("6 (cor: ", signif(temp.cor4,2),")")),
       title = "Segment",
       pt.bg = c("royalblue", "firebrick", "gold", "darkgreen"),
       pch = 21)

```

```{r echo = F, fig.height= 8, fig.width= 8, include=F}

cor.table = read.csv("/Users/johncsantiago/Documents/GitHub/WhartonLab/RussoPaper/20250701_AllAnaData_CorrelationTable.csv",header = F)

cor.table= cor.table[,-grep("Raw", cor.table[1,])]

cor.table[1,grep("Norm", cor.table[1,])] = substr(cor.table[1,grep("Norm", cor.table[1,])], 8, nchar(cor.table[1,grep("Norm", cor.table[1,])]))

cor.tablea = cor.table[,cor.table[2,] == "a"]
cor.tablea = cbind(cor.tablea, cor.table[,20])
cor.tablea = cbind(cor.tablea, cor.table[,1])

cor.tableb = cor.table[,cor.table[2,] == "b"]
cor.tableb = cbind(cor.tableb, cor.table[,20])
cor.tableb = cbind(cor.tableb, cor.table[,1])

colnames(cor.tablea) = paste0(cor.tablea[1,], ".", cor.tablea[3,], ".", cor.tablea[4,])
colnames(cor.tableb) = paste0(cor.tableb[1,], ".", cor.tableb[3,], ".", cor.tableb[4,])

cor.table = rbind(cor.tablea[c(1,3:nrow(cor.tablea)),],cor.tableb[5:nrow(cor.tableb),])

cor.matrix = cor.table[cor.table[,ncol(cor.table)] == "WT-cr/LoxP" |
                         cor.table[,ncol(cor.table)] == "G85Rcr/G85Rhr",
                       1:ncol(cor.table)]

temp = data.frame(NMJ4 = as.numeric(cor.matrix[cor.matrix[,12] == "G85Rcr/G85Rhr",
                  c("Boutons.4.3")]),
                  NMJ67 = as.numeric(cor.matrix[cor.matrix[,12] == "G85Rcr/G85Rhr",
                  c("Boutons.6-7.3")]))

temp.cor1 = cor(na.omit(temp)[,1], na.omit(temp)[,2])

temp.lm = lm(na.omit(temp)[,1] ~ na.omit(temp)[,2])[[1]]

#temp[is.na(temp)] = 0

temp = na.omit(temp)

plot(x = temp$NMJ67,
     y = temp$NMJ4,
     pch = 21,
     bg = "deepskyblue",
     cex = 1,
     #xaxt = "n",
     #yaxt = "n",
     ylab = "",
     xlab = "",
     bty = "n",
     main = "",
     xlim = c(0, 100),
     ylim = c(0,50))

abline(a = temp.lm[1],
       b = temp.lm[2],
       lwd = 2,
       col = "royalblue")

temp = data.frame(NMJ4 = as.numeric(cor.matrix[cor.matrix[,12] == "G85Rcr/G85Rhr",
                  c("Boutons.4.4")]),
                  NMJ67 = as.numeric(cor.matrix[cor.matrix[,12] == "G85Rcr/G85Rhr",
                  c("Boutons.6-7.4")]))

temp.cor2 = cor(na.omit(temp)[,1], na.omit(temp)[,2])

temp.lm = lm(na.omit(temp)[,1] ~ na.omit(temp)[,2])[[1]]

points(x = temp$NMJ67,
       y = temp$NMJ4,
       pch = 21,
       bg = "deeppink",
       cex = 1)

abline(a = temp.lm[1],
       b = temp.lm[2],
       lwd = 2,
       col = "firebrick")

temp = data.frame(NMJ4 = as.numeric(cor.matrix[cor.matrix[,12] == "G85Rcr/G85Rhr",
                  c("Boutons.4.5")]),
                  NMJ67 = as.numeric(cor.matrix[cor.matrix[,12] == "G85Rcr/G85Rhr",
                  c("Boutons.6-7.5")]))

temp.cor3 = cor(na.omit(temp)[,1], na.omit(temp)[,2])

temp.lm = lm(na.omit(temp)[,1] ~ na.omit(temp)[,2])[[1]]

points(x = temp$NMJ67,
       y = temp$NMJ4,
       pch = 21,
       bg = "lightyellow",
       cex = 1)

abline(a = temp.lm[1],
       b = temp.lm[2],
       lwd = 2,
       col = "gold")

temp = data.frame(NMJ4 = as.numeric(cor.matrix[cor.matrix[,12] == "G85Rcr/G85Rhr",
                  c("Boutons.4.6")]),
                  NMJ67 = as.numeric(cor.matrix[cor.matrix[,12] == "G85Rcr/G85Rhr",
                  c("Boutons.6-7.6")]))

temp.cor4 = cor(na.omit(temp)[,1], na.omit(temp)[,2])

temp.lm = lm(na.omit(temp)[,1] ~ na.omit(temp)[,2])[[1]]

points(x = temp$NMJ67,
       y = temp$NMJ4,
       pch = 21,
       bg = "lightgreen",
       cex = 1)

abline(a = temp.lm[1],
       b = temp.lm[2],
       lwd = 2,
       col = "darkgreen")

legend("topleft",
       legend = c(paste0("3 (cor: ", signif(temp.cor1,2),")"), 
                  paste0("4 (cor: ", signif(temp.cor2,2),")"),
                  paste0("5 (cor: ", signif(temp.cor3,2),")"),
                  paste0("6 (cor: ", signif(temp.cor4,2),")")),
       title = "Segment",
       pt.bg = c("royalblue", "firebrick", "gold", "darkgreen"),
       pch = 21)

```